### 2019-11-03 analysis/index.Rmd
model and Xin's simulation

---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

## [Model](model.html)

## [Simulation](simulation-singleG.html)

### 2019-11-03 analysis/model.Rmd
model and Xin's simulation

---
title: "Model"
author: "Siming Zhao"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model
 $$Y = \tilde{X}\gamma + G\theta + \epsilon$$
$Y$: quantitative traits, $N$ x 1 vector, $N$ is number of individuals  
$\tilde{X}$: imputed gene expression (standardized). $N$ x $J$ matrix, $J$ is number of genes.  
$G$: genotype (standardized). $N$ x $M$ matrix, $M$ is number of SNPs.  
$\gamma$: effect sizes for $\tilde{X}$,  $\gamma_j$, effect size for gene $j$.

$$ \gamma_j \sim N(0, \sigma_\gamma^2)$$
$\theta$: effect sizes for $G$,  $\theta_m$, effect size for SNP $m$.  
$$ \theta_m \sim N(0, \sigma_\theta^2)$$

## Model inference
The above model is equivalent to this: 
$$Y \sim N(0,\sigma_\gamma^2XX^T + \sigma_\theta^2GG^T + \sigma_e^2I) $$
This is REML with multiple (more than 2) variance components. Use HE regression function from GEMMA software.

## Simulation
Simulate different $PVE_{SNP}$, $PVE_{gene}$.

$$PVE_{SNP} = \frac{M\sigma_\theta^2}{J\sigma_\gamma^2 + M\sigma_\theta^2 + \sigma_e^2}$$
$$PVE_{gene} = \frac{J\sigma_\gamma^2}{J\sigma_\gamma^2 + M\sigma_\theta^2 + \sigma_e^2}$$
Scenario 1: $PVE_{SNP}=0.2$, $PVE_{gene}=0.2$

### 2019-11-03 analysis/simulation-singleG.Rmd
model and Xin's simulation

---
title: "Simulation -- Single Gene"
author: "Siming Zhao"
date: "11/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simulation
  We performed simulations todemonstrate the importance of accounting for pleiotropic effects ofSNPs when performing TWAS. We take real genotype data of aregion near GNL3 (a top gene in a TWAS of SCZ) from WTCCC(n=  3,000).   We  first  simulate  expression  data  of  the  gene  inthe  region,  assuming  two  independent  cis-eQTLs  with  total  cis-heritability of 30%.  We next simulate phenotype data under twosettings:  (1) the gene has no causal effect, but one of the eQTLsacts on the phenotype (through other mechanism);  (2) the genehas a causal effect, and no other SNPs in the region affect the phenotype.  We compare TWAS (using Lassoas expression models) with a simple “causal TWAS” method, which performs multiple regression of phenotypeagainst all SNPs in the region plus the predicted gene expression. We use SuSiE, a Bayesian variable selectionmethod for regression. For TWAS, we use a stringentp-value threshold of 0.001 (note that we are testing a singlegene). Since SuSie does not report p-values, we call the gene effect significant if it is found in the “credible set”.In the first setting (gene has no causal effect),  we found the false positive (FP) rates of TWAS are muchhigher than the causal TWAS method (Figure 3A). This is because under this model, the association of geneexpression with trait is explained away by the causal SNP. In the second setting (gene has causal effect), whileTWAS has high power as expected,  the causal TWAS method also achieves appreciable power (Figure 3B).These simulations thus demonstrate the feasibility of controlling FP rates while maintaining power.  We believethe full EB framework outlined below will help close the gap of power. Additionally, in our simulations, we benefitfrom the fact that the gene has two independent eQTL signals.  In reality, EB approach would be critical for thesituation where a gene has single eQTL signal.

(A) Gene has no causal effect.  The effectsize of the pleiotropic SNP is measured by the Per-cent of Variance Explained (PVE) for the phenotype.(B) Gene has a causal effect, with the effect size (mea-sured by PVE) shown at the x-axis.  Note that we uselarge effect size (PVE) in simulations to reduce samplesizes needed.

```{r simulation}
library(MASS)
library(susieR)
library(glmnet)

# read genotype data
file <- "data/genotype_GNL3.Rdata"
load(file)
G <- scale(mydata)
N <- nrow(G)
M <- ncol(G)

# Simulation function
# G: genotype matrix. h2.eQTL: h2 of eQTL. n.eQTL: number of eQTL SNPs. gamma: effect size of gene. theta: effect size of SNP.
simulate <- function(G, h2.eQTL, n.eQTL, gamma, theta) {
  N <- dim(G)[1]
  M <- dim(G)[2]

  idx.eQTL <- sample(1:M, n.eQTL)
  alpha <- rep(sqrt(h2.eQTL/n.eQTL), n.eQTL)

  # simulate gene expression data
  X <- G[, idx.eQTL] %*% alpha + rnorm(N, sd=1)

  # simulate phenotype data
  idx.SNP <- idx.eQTL[1] # choose a SNP that is one eQTL
  Y <- X * gamma + as.matrix(G[, idx.SNP]) %*% theta + rnorm(N)

  # return results
  return (list(X=X, Y=Y, idx.eQTL=idx.eQTL, idx.SNP))
}

# Evaluation function
evaluate <- function(G, X, Y) {
  # predicted expression
  cvfit <- cv.glmnet(G, X)
  X.tilde <- predict(cvfit, G, s = "lambda.min")

  # TWAS
  fit <- lm(Y ~ X.tilde)
  pvalue <- summary(fit)$coefficients["X.tilde",4]

  # causal-TWAS
  res <- susie(cbind(G, X.tilde), Y, L=1)
  gene.indicator <- (M+1) %in% res$sets$cs$L1

  # return results
  return (list(twas.pvalue=pvalue, causal.twas.gene=gene.indicator))
}

# Compare TWAS vs. causal-TWAS
h2.eQTL <- 0.3
nsims <- 50

# false positive simulation
gamma <- 0.0
theta <- 0.07
n.eQTL <- 2

twas.pvalue <- array(nsims)
causal.twas.gene <- array(nsims)
for (i in 1:nsims) {
  data <- simulate(G, h2.eQTL, n.eQTL, gamma, theta)
  result <- evaluate(G, data$X, data$Y)
  twas.pvalue[i] <- result$twas.pvalue
  causal.twas.gene[i] <- result$causal.twas.gene
}
cat("FP rate of TWAS = ", sum(twas.pvalue < 0.001)/nsims, "\n")
cat("FP rate of causal TWAS = ", sum(causal.twas.gene==TRUE)/nsims, "\n")

FP.low <- c(0.38, 0.04)
FP.high <- c(0.62, 0.14)
FP.rates <- cbind(FP.low, FP.high)
colors <- c("orange","blue")
barplot(FP.rates, col=colors, xlab="PVE of pleiotropic SNP", ylab="False positive rate", ylim=c(0,0.8), names.arg=c("0.5%", "1%"), beside=TRUE, cex.lab=1.3)
legend("topleft",c("TWAS","causal TWAS"),col=colors,pch=15, cex=1.2)

# power simulation
n.eQTL <- 2
gamma <- 0.18
theta <- 0

twas.pvalue <- array(nsims)
causal.twas.gene <- array(nsims)
for (i in 1:nsims) {
  data <- simulate(G, h2.eQTL, n.eQTL, gamma, theta)
  result <- evaluate(G, data$X, data$Y)
  twas.pvalue[i] <- result$twas.pvalue
  causal.twas.gene[i] <- result$causal.twas.gene
}
cat("Power of TWAS = ", sum(twas.pvalue < 0.001)/nsims, "\n")
cat("Power of causal TWAS = ", sum(causal.twas.gene==TRUE)/nsims, "\n")

# plot the results
power.low <- c(0.7, 0.4)
power.high <- c(0.96, 0.68)
TP.rates <- cbind(power.low, power.high)
colors <- c("orange","blue")
barplot(TP.rates, col=colors, xlab="PVE of gene", ylab="Power", ylim=c(0,1), names.arg=c("0.5%", "1%"), beside=TRUE, cex.lab=1.3)
legend("topleft",c("TWAS","causal TWAS"),col=colors,pch=15, cex=1.2)
```

### 2019-11-05 analysis/simulation-WTCCC.Rmd
gemma

---
title: "simulation-WTCCC"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### WTCCC data

### Fusion results

## Simulate quantitative phenotype
Simulate different $PVE_{SNP}$, $PVE_{gene}$.

$$PVE_{SNP} = \frac{M\sigma_\theta^2}{J\sigma_\gamma^2 + M\sigma_\theta^2 + \sigma_e^2}$$
$$PVE_{gene} = \frac{J\sigma_\gamma^2}{J\sigma_\gamma^2 + M\sigma_\theta^2 + \sigma_e^2}$$
Scenario 1: $PVE_{SNP}=0.2$, $PVE_{gene}=0.2$

### 2019-11-07 analysis/simulation-WTCCC.Rmd
simulate description

@@
 ```
-### WTCCC data
+Use WTCCC data obtained from Peter on CRI.
 
-### Fusion results
+### Simulate cis-expression 
+We simulate cis- expression based on the following model:
 
-## Simulate quantitative phenotype
+$$ \tilde{X} = G\alpha + \xi$$
+For each gene, we sample $K$ eQTLs for this gene, for eQTL $k$, we have
+$$ \alpha_k \sim N(0,\sigma_\alpha^2) $$
+$$ \xi \sim N(0,1) $$
+Then we have the heritability of the gene:
+
+$$\begin{aligned}
+h^2_{eQTL} &= \frac{var(G\alpha)}{var(G\alpha)+var(\xi)} \\
+&= \frac{\Sigma_kvar(G_k) \alpha_k^2}{\Sigma_kvar(G_k) \alpha_k^2 + var(\xi)}\\
+&=  \frac{\Sigma_k\alpha_k^2}{\Sigma_k\alpha_k^2 + var(\xi)}\\
+&\approx \frac{\Sigma_k\sigma_\alpha^2}{\Sigma_k\sigma_\alpha^2 + var(\xi)} \\
+&=\frac{K\sigma_\alpha^2}{1+K\sigma_\alpha^2}
+\end{aligned}$$
+
+Here, we use scaled genotype data, so $var(G)=1$.
+We also have $\alpha^2_k \approx E(\alpha_k^2)=var(\alpha_k^2)=\sigma_\alpha^2$.
+Usually, $K$ ranges from 1 to 5, let $\sigma_\alpha = 0.3$, then $h^2_{eQTL}$ ranges from 0.08 to 0.31, this is consistent with gene cis-heritability in reality.
+
+### Simulate quantitative phenotype
 Simulate different $PVE_{SNP}$, $PVE_{gene}$.

### 2019-11-11 analysis/simulation-WTCCC.Rmd
simulate expression

@@
 ---
-title: "simulation-WTCCC"
+title: "Simulation-WTCCC"
 output: html_document
@@
 $$ \tilde{X} = G\alpha + \xi$$
-For each gene, we sample $K$ eQTLs for this gene, for eQTL $k$, we have
+For each gene, we sample $L$ eQTLs for this gene, for eQTL $k$, we have
 $$ \alpha_k \sim N(0,\sigma_\alpha^2) $$
@@
 ### Simulate quantitative phenotype
-Simulate different $PVE_{SNP}$, $PVE_{gene}$.
 
-$$PVE_{SNP} = \frac{M\sigma_\theta^2}{J\sigma_\gamma^2 + M\sigma_\theta^2 + \sigma_e^2}$$
-$$PVE_{gene} = \frac{J\sigma_\gamma^2}{J\sigma_\gamma^2 + M\sigma_\theta^2 + \sigma_e^2}$$
-Scenario 1: $PVE_{SNP}=0.2$, $PVE_{gene}=0.2$
+$$\begin{align}
+PVE_{SNP} = &\frac{var(G\theta)}{var(G\theta) + var(\tilde{X}\gamma) + \sigma_e^2}\\ 
+\approx &\frac{M\sigma_\theta^2}{M\sigma_\theta^2 + \Sigma_jvar(\tilde{X_j})\gamma_j^2+ \sigma_e^2}\\ 
+\approx &\frac{M\sigma_\theta^2}{M\sigma_\theta^2 + Jvar(\tilde{X})\sigma_\gamma^2+ \sigma_e^2}\\   
+\end{align}$$
+
+$$
+PVE_{expr} \approx \frac{Jvar(\tilde{X})\sigma_\gamma^2}{M\sigma_\theta^2 + Jvar(\tilde{X})\sigma_\gamma^2+ \sigma_e^2}$$
+
+Here, $var(\tilde{X})$ is the cis-heratbility of gene expression. $M$ and $J$ are number of causal SNP and gene respectively.
+
+## Simulation 1: $J$ = No. of all SNPs, $M$ = No. all genes
+We simulate quantitative phenotype under several scenarios.
+First we make all genes with imputed gene expression as causal genes and all SNPs as causal SNPs.
+This matches our polygenic version of the model.
+
+### 1.1 $PVE_{SNP}=0.1$, $PVE_{expr}=0.1$
+We will set $\sigma_e=1$.
+
+### 1.2 $PVE_{SNP}=0.1$, $PVE_{expr}=0.4$
+We will set $\sigma_e=1$.

### 2019-11-17 code/stats_func.R
PVE

@@ -1,0 +1,29 @@
+# Short script to center and scale the columns (i.e., SNPs) of a
+# matrix containing genotypes. Credit to Peter
+library(matrixStats)
+library(Rcpp)
+
+scaleRcpp <- function(x) {
+  cat("Scaling genotype data.\n")
+  # Load the genotype data as a matrix of floating-point numbers.
+  storage.mode(x) <- "double"
+  # Remove all SNPs that do not vary.
+  # cat("Filtering SNPs.\n")
+  # s    <- colSds(x)
+  # x <- x[,s > 0]
+  # gc()
+  # center and scale
+  sourceCpp("scale.cpp")
+  timing <- system.time({
+    x.scaled <- x
+    mu <- colMeans(x)
+    s  <- colSds(x)
+    scale_rcpp(x.scaled,mu,s)
+  })
+  print(timing)
+  # verify
+  cat("Get the largest & smallest column mean:\n")
+  print(range(colMeans(x.scaled)))
+  cat("Get the largest & smallest column s.d.:\n")
+  print(range(colSds(x.scaled)))
+}

### 2019-11-19 analysis/simulation-WTCCC.Rmd
PVE
simulate phenotype

@@
 
-### Simulate cis-expression 
+## Simulate cis-expression 
 We simulate cis- expression based on the following model:
@@
 
-### Simulate quantitative phenotype
+## Simulate quantitative phenotype
 
@@
 
-## Simulation 1: $J$ = No. of all SNPs, $M$ = No. all genes
+Thus in order to get desired $PVE_{SNP}$ and $PVE_expr$, we set $\sigma_\theta^2$ and $\sigma_\gamma^2$ based on the following formula:
+
+$$\sigma_\theta^2 = \frac{PVE_{SNP}}{M(1-PVE_{SNP} - PVE_{expr})}$$
+$$\sigma_\gamma^2 = \frac{PVE_{expr}}{Jvar(\tilde{X})(1-PVE_{SNP} - PVE_{expr})}$$
+
+## Simulation 1: $M$ = No. of all SNPs, $J$ = No. all genes
 We simulate quantitative phenotype under several scenarios.
@@
 ### 1.1 $PVE_{SNP}=0.1$, $PVE_{expr}=0.1$
-We will set $\sigma_e=1$.
+We set $\sigma_e=1$.
 
 ### 1.2 $PVE_{SNP}=0.1$, $PVE_{expr}=0.4$
-We will set $\sigma_e=1$.
+We set $\sigma_e=1$.
+
+## Simulation 2: $M$ = No. of all SNPs, $J$ = ~10% all genes (J=2000)
+
+### 2.1 $PVE_{SNP}=0.1$, $PVE_{expr}=0.1$
+We set $\sigma_e=1$.
+
+### 2.2 $PVE_{SNP}=0.1$, $PVE_{expr}=0.4$
+We set $\sigma_e=1$.
+
+## Simulation 3: $M$ = ~ 10% all SNPs (M=50000), $J$ = No. all genes
+
+### 3.1 $PVE_{SNP}=0.1$, $PVE_{expr}=0.1$ 
+We set $\sigma_e=1$.
+
+### 3.2 $PVE_{SNP}=0.1$, $PVE_{expr}=0.4$
+We set $\sigma_e=1$.
+
+## Simulation 4: $M$ = ~ 10% all SNPs (M=50000), $J$ = ~10% all genes (J=2000)
+
+### 4.1 $PVE_{SNP}=0.1$, $PVE_{expr}=0.1$ 
+We set $\sigma_e=1$.
+
+### 4.2 $PVE_{SNP}=0.1$, $PVE_{expr}=0.4$
+We set $\sigma_e=1$.

### 2019-11-22 analysis/simulation-WTCCC.Rmd
simu20191111

@@
 
-Here, $var(\tilde{X})$ is the cis-heratbility of gene expression. $M$ and $J$ are number of causal SNP and gene respectively.
+Here, $var(\tilde{X})$ is the cis-heratbility of gene expression. $M$ and $J$ are numbers of causal SNP and gene respectively.
 
@@
 
+```{r}
+simudir <- "/home/simingz/causalTWAS/simulations/simulation_WTCCC_20191111/"
+show_res <- function(simutag){
+  load(Sys.glob(paste0(simudir,simutag,"*phenotype.Rd")))
+  outdf1 <- data.frame(truth = c(sigma_gamma^2*J.c, sigma_theta^2*M.c , 1))
+  row.names(outdf1) <- c("sigma_gamma^2","sigma_theta^2","sigma_e^2")
+  res <- readLines(paste0(simudir, simutag, "_VC/output/", simutag, ".log.txt"))
+  outdf1$est <- as.numeric(strsplit(res[24], "  ")[[1]][2:4])
+  outdf1$est.se <- as.numeric(strsplit(res[25], "  ")[[1]][2:4])
+  print(outdf1)
+  outdf2 <- data.frame("est"= c(as.numeric(strsplit(res[20], "  ")[[1]][2:3]), as.numeric(strsplit(res[22], "=")[[1]][2])))
+  row.names(outdf2) <- c("PVE.expr","PVE.snp","PVE.total")
+  outdf2$est.se <- c(as.numeric(strsplit(res[21], "  ")[[1]][2:3]), as.numeric(strsplit(res[23], "=")[[1]][2]))
+  print(outdf2)
+}
+```
+
 ## Simulation 1: $M$ = No. of all SNPs, $J$ = No. all genes
@@
 
-### 1.1 $PVE_{SNP}=0.1$, $PVE_{expr}=0.1$
-We set $\sigma_e=1$.
+### 1.1 $PVE_{SNP} \approx 0.1$, $PVE_{expr} \approx 0.1$
+```{r 1.1}
+show_res("S1.1")
+```
 
-### 1.2 $PVE_{SNP}=0.1$, $PVE_{expr}=0.4$
-We set $\sigma_e=1$.
+### 1.2 $PVE_{SNP}\approx 0.1$, $PVE_{expr}\approx 0.4$
+```{r 1.2}
+show_res("S1.2")
+```
 
@@
 
-### 2.1 $PVE_{SNP}=0.1$, $PVE_{expr}=0.1$
-We set $\sigma_e=1$.
+### 2.1 $PVE_{SNP} \approx 0.1$, $PVE_{expr} \approx 0.1$
+```{r 2.1}
+show_res("S2.1")
+```
 
-### 2.2 $PVE_{SNP}=0.1$, $PVE_{expr}=0.4$
-We set $\sigma_e=1$.
-
+### 2.2 $PVE_{SNP} \approx 0.1$, $PVE_{expr} \approx 0.4$
+```{r 2.2}
+show_res("S2.2")
+```
 ## Simulation 3: $M$ = ~ 10% all SNPs (M=50000), $J$ = No. all genes
 
-### 3.1 $PVE_{SNP}=0.1$, $PVE_{expr}=0.1$ 
-We set $\sigma_e=1$.
+### 3.1 $PVE_{SNP} \approx 0.1$, $PVE_{expr} \approx 0.1$ 
+```{r 3.1}
+show_res("S3.1")
+```
 
-### 3.2 $PVE_{SNP}=0.1$, $PVE_{expr}=0.4$
-We set $\sigma_e=1$.
+### 3.2 $PVE_{SNP} \approx 0.1$, $PVE_{expr} \approx 0.4$
+```{r 3.2}
+show_res("S3.2")
+```
 
@@
 
-### 4.1 $PVE_{SNP}=0.1$, $PVE_{expr}=0.1$ 
-We set $\sigma_e=1$.
+### 4.1 $PVE_{SNP} \approx 0.1$, $PVE_{expr} \approx 0.1$ 
+```{r 4.1}
+show_res("S4.1")
+```
 
-### 4.2 $PVE_{SNP}=0.1$, $PVE_{expr}=0.4$
-We set $\sigma_e=1$.
+### 4.2 $PVE_{SNP} \approx 0.1$, $PVE_{expr} \approx 0.4$
+```{r 4.2}
+show_res("S4.2")
+```

### 2019-11-25 analysis/simulation-WTCCC.Rmd
X tilde

@@
 ## Simulate cis-expression 
-We simulate cis- expression based on the following model:
+We simulate cis-expression based on the following model:
 
@@
 
-### 1.2 $PVE_{SNP}\approx 0.1$, $PVE_{expr}\approx 0.4$
+### 1.2 $PVE_{SNP}\approx 0.4$, $PVE_{expr}\approx 0.1$
 ```{r 1.2}
@@
 
-### 2.2 $PVE_{SNP} \approx 0.1$, $PVE_{expr} \approx 0.4$
+### 2.2 $PVE_{SNP} \approx 0.4$, $PVE_{expr} \approx 0.1$
 ```{r 2.2}
@@
 ```
-## Simulation 3: $M$ = ~ 10% all SNPs (M=50000), $J$ = No. all genes
+## Simulation 3: $M$ = ~ 1% all SNPs (M=5000), $J$ = No. all genes
 
@@
 
-### 3.2 $PVE_{SNP} \approx 0.1$, $PVE_{expr} \approx 0.4$
+### 3.2 $PVE_{SNP} \approx 0.4$, $PVE_{expr} \approx 0.1$
 ```{r 3.2}
@@
 
-## Simulation 4: $M$ = ~ 10% all SNPs (M=50000), $J$ = ~10% all genes (J=2000)
+## Simulation 4: $M$ = ~ 1% all SNPs (M=5000), $J$ = ~10% all genes (J=2000)
 
@@
 
-### 4.2 $PVE_{SNP} \approx 0.1$, $PVE_{expr} \approx 0.4$
+### 4.2 $PVE_{SNP} \approx 0.4$, $PVE_{expr} \approx 0.1$
 ```{r 4.2}

### 2020-03-13 analysis/simulation-WTCCC2.Rmd
v1.0 simulate expression using separate data 20191127

--- 2019-11-25 analysis/simulation-WTCCC.Rmd
@@
 ---
-title: "Simulation-WTCCC"
+title: "Simulation-WTCCC-2"
 output: html_document
@@
 Use WTCCC data obtained from Peter on CRI.
+
+The difference between this analysis and [previous run](simulation-WTCCC.html): used 500 samples from `cad.RData` from WTCCC to simulate expression, then train expression model using `cv.glmnet`.
+We impute expression in using `bd.RData` from WTCCC.
+
+Simulation of phenotype: use the same weights as used in simulate expression using the `'cad.RData` dataset.
+
+The genotype data and other settings are the same.
 
@@
 ## Simulate quantitative phenotype
+We simulate phenotype followin
 
@@
 ```{r}
-simudir <- "/home/simingz/causalTWAS/simulations/simulation_WTCCC_20191111/"
+simudir <- "/home/simingz/causalTWAS/simulations/simulation_WTCCC_20191127/"
 show_res <- function(simutag){
@@
 ```
-
-## Simulation 4: $M$ = ~ 1% all SNPs (M=5000), $J$ = ~10% all genes (J=2000)
-
-### 4.1 $PVE_{SNP} \approx 0.1$, $PVE_{expr} \approx 0.1$ 
-```{r 4.1}
-show_res("S4.1")
-```
-
-### 4.2 $PVE_{SNP} \approx 0.4$, $PVE_{expr} \approx 0.1$
-```{r 4.2}
-show_res("S4.2")
-```

### 2020-03-25 analysis/simulation-WTCCC.Rmd
fix results display bug

@@
 
-$$ \tilde{X} = G\alpha + \xi$$
+$$ X' = G\alpha$$
 For each gene, we sample $L$ eQTLs for this gene, for eQTL $k$, we have
@@
 
+We simulate phenotype following
+
+$$Y = G\alpha\gamma + G\theta + \epsilon$$.
+
+For $\alpha$, we used the same as used in simulating expression (Thus, not $\tilde{X}$).
+$\gamma \sim N(0,\sigma_\gamma^2), \theta \sim N(0, \sigma_\theta^2), \epsilon \sim N (0,1)$.
+To choose proper $\sigma_\gamma$ and $\sigma_\theta$.
+We used the following formula:
+
 $$\begin{align}
@@
 
-Thus in order to get desired $PVE_{SNP}$ and $PVE_expr$, we set $\sigma_\theta^2$ and $\sigma_\gamma^2$ based on the following formula:
+Thus in order to get desired $PVE_{SNP}$ and $PVE_{expr}$, we set $\sigma_\theta^2$ and $\sigma_\gamma^2$ based on the following formula:
 
@@
 simudir <- "/home/simingz/causalTWAS/simulations/simulation_WTCCC_20191111/"
+load("/home/simingz/causalTWAS/WTCCC/bd.toy.RData")
+Gvar <- mean(apply(X,2,var))
 show_res <- function(simutag){
   load(Sys.glob(paste0(simudir,simutag,"*phenotype.Rd")))
-  outdf1 <- data.frame(truth = c(sigma_gamma^2*J.c, sigma_theta^2*M.c , 1))
+  outdf1 <- data.frame(truth = c(sigma_gamma^2*J.c, sigma_theta^2*M.c/Gvar , 1))
   row.names(outdf1) <- c("sigma_gamma^2","sigma_theta^2","sigma_e^2")

### 2020-03-25 analysis/simulation-WTCCC2.Rmd
fix results display bug

@@
 
-Simulation of phenotype: use the same weights as used in simulate expression using the `'cad.RData` dataset.
+Simulation of phenotype: use the same weights as used in simulate expression using the `'cad.RData` dataset, simulate phenotype according to:
+
+$$Y = G\alpha\gamma + G\theta + \epsilon$$.
+For $\alpha$, we used the same as used in simulating expression (Thus, not $\tilde{X}$).
+$\gamma \sim N(0,\sigma_\gamma^2), \theta \sim N(0, \sigma_\theta^2), \epsilon \sim N (0,1)$.
 
@@
 ## Simulate cis-expression 
-We simulate cis-expression based on the following model:
+We simulate cis-expression based on the following model, using 500 samples from cad in WTCCC:
 
@@
 ## Simulate quantitative phenotype
-We simulate phenotype followin
+We simulate phenotype using bd data from WTCCC following
+$$Y = G\alpha\gamma + G\theta + \epsilon$$.
+For $\alpha$, we used the same as used in simulating expression (Thus, not $\tilde{X}$).
+$\gamma \sim N(0,\sigma_\gamma^2), \theta \sim N(0, \sigma_\theta^2), \epsilon \sim N (0,1)$.
 
@@
 simudir <- "/home/simingz/causalTWAS/simulations/simulation_WTCCC_20191127/"
+load("/home/simingz/causalTWAS/WTCCC/bd.toy.RData")
+Gvar <- mean(apply(X,2,var))
 show_res <- function(simutag){
   load(Sys.glob(paste0(simudir,simutag,"*phenotype.Rd")))
-  outdf1 <- data.frame(truth = c(sigma_gamma^2*J.c, sigma_theta^2*M.c , 1))
+  outdf1 <- data.frame(truth = c(sigma_gamma^2*J.c, sigma_theta^2*M.c/Gvar , 1))
   row.names(outdf1) <- c("sigma_gamma^2","sigma_theta^2","sigma_e^2")
@@
 ```
+
+## Discussion
+In this analysis, we see the PVE estimation for expression is much under estimated, this is because many of the gene expression training model has low $R^2$, especially for low heritability ones.
+For a polygenic model, because we are estimating the total herability explained by expression, the ones with low $R^2$ will affect this estimation.
+We are moving to model with a sparse prior for expression.

### 2020-03-25 analysis/polygenic_model.Rmd
polygenic done

--- 2019-11-03 analysis/model.Rmd
@@
 
-## Model
+## Polygenic Model
  $$Y = \tilde{X}\gamma + G\theta + \epsilon$$
@@
 
-## Model inference
+## Polygenic Model Inference
 The above model is equivalent to this: 
@@
 This is REML with multiple (more than 2) variance components. Use HE regression function from GEMMA software.
-
-## Simulation
-Simulate different $PVE_{SNP}$, $PVE_{gene}$.
-
-$$PVE_{SNP} = \frac{M\sigma_\theta^2}{J\sigma_\gamma^2 + M\sigma_\theta^2 + \sigma_e^2}$$
-$$PVE_{gene} = \frac{J\sigma_\gamma^2}{J\sigma_\gamma^2 + M\sigma_\theta^2 + \sigma_e^2}$$
-Scenario 1: $PVE_{SNP}=0.2$, $PVE_{gene}=0.2$

### 2020-03-25 analysis/index.Rmd
gemma
v1.0 simulate expression using separate data 20191127
polygenic done
fix results display bug

@@
   workflowr::wflow_html:
-    toc: false
+    toc: true
 editor_options:
@@
 
-## [Model](model.html)
+# Model
+  [polygenic model](polygenic_model.html)
 
-## [Simulation](simulation-singleG.html)
+# Simulations
+
+## For a single gene
+  * [single gene](simulation-singleG.html)
+
+## Under polygenic model
+  * [Simulation: WTCCC](simulation-WTCCC.html)
+  * [Simulation: WTCCC 2](simulation-WTCCC2.html)

### 2020-03-27 analysis/index.Rmd
veb-boost reference

@@
 # Model
-  [polygenic model](polygenic_model.html)
+  * [Polygenic model](polygenic_model.html)
 
@@
 ## For a single gene
-  * [single gene](simulation-singleG.html)
+  * [Single gene](simulation-singleG.html)
 
@@
   * [Simulation: WTCCC 2](simulation-WTCCC2.html)
+
+# References
+  * VEB-boost: Andrew's [software](https://github.com/andrewg3311/VEB.Boost) and [workflowr website](https://andrewg3311.github.io/VEB-Boost-workflowR/VEB_Boost_Derivation.html)

### 2020-03-31 analysis/VEB-boost-fitting.Rmd
sparse model description

---
title: "VEB-boost"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem description

$$Y_{N \times 1}= X_{N \times J}\beta_{J\times1} + G_{N\times M}\theta_{M\times1} + \epsilon_{N\times1}$$
$$ \epsilon \sim MVN(0,I_N\tau^{-1})$$
Spike and slab prior for $\beta_j$:
$$\beta_j|\pi,\tau \sim (1- \pi_\beta)\delta_0 + \pi_\beta N(0, \sigma_\beta^2/\tau)$$
Spike and slab prior for $\theta_m$:
$$\theta_m | \tau \sim (1- \pi_\theta)\delta_0 + \pi_\theta N(0, \sigma_\theta^2/\tau)$$
We want to estimate $\omega=\{\tau,\pi_\beta, \pi_\theta, \sigma_\beta^2,\sigma_\theta^2\}$

### 2020-03-31 analysis/index.Rmd
sparse model description

--- 2020-03-25 analysis/index.Rmd
@@
 # Model
-  [polygenic model](polygenic_model.html)
+  * [Polygenic model](polygenic_model.html)
+  * [Sparse model](sparse_model.html)
 
@@
   * [Simulation: WTCCC 2](simulation-WTCCC2.html)
+
+# References
+  * VEB-boost
+    * [Problem description](VEB-boost-fitting.html)
+    * Andrew's [software](https://github.com/andrewg3311/VEB.Boost) and [workflowr website](https://andrewg3311.github.io/VEB-Boost-workflowR/VEB_Boost_Derivation.html)

### 2020-03-31 analysis/polygenic_model.Rmd
fix results display bug
sparse model description

@@
 ---
-title: "Model"
+title: "Model- polygenic version"
 author: "Siming Zhao"
@@
 $Y$: quantitative traits, $N$ x 1 vector, $N$ is number of individuals  
-$\tilde{X}$: imputed gene expression (standardized). $N$ x $J$ matrix, $J$ is number of genes.  
+$\tilde{X}$: cis-gene expression (the genetic component of expression). $N$ x $J$ matrix, $J$ is number of genes.  
 $G$: genotype (standardized). $N$ x $M$ matrix, $M$ is number of SNPs.  

### 2020-03-31 analysis/sparse_model.Rmd
sparse model description

---
title: "Sparse model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sparse model for expression and genotype
 $$Y = \tilde{X}\beta + G\theta + \epsilon$$
$Y$: quantitative traits, $N$ x 1 vector, $N$ is number of individuals, the columns are centered.
$\tilde{X}$: the genetic component of gene expression, $N$ x $J$ matrix, $J$ is number of genes, the columns are centered.
$G$: genotype (standardized). $N$ x $M$ matrix, $M$ is number of SNPs.

The error term $\epsilon$ has a normal distribution,

$$\epsilon|\tau \sim N(0, \tau^{-1})$$

We use the spike and slab prior for gene expression effect size $\beta_j$ (effect size for gene $j$).
$$\beta_j|\pi,\tau \sim (1- \pi_\beta)\delta_0 + \pi_\beta N(0, \sigma_\beta^2/\tau)$$
Here $\delta_0$ is point mass at 0 ($\delta_0(\beta_j)=1$ if $\beta_j=0$, otherwise $\delta_0(\beta_j)=0$ ).

We use the spike and slab prior for SNP effect size $\theta_m$ (effect size for SNP $m$).
$$ \theta_m | \tau \sim (1- \pi_\theta)\delta_0 + \pi_\theta N(0, \sigma_\theta^2/\tau)$$

We use the following prior distributions for the hyperparameters $\omega = (\pi_\beta,\pi_\theta, \tau, \sigma_\beta^2,\sigma_\theta^2)$: 
$$\tau \sim Gamma(\kappa_1, \kappa_2)$$

$$log(\pi_\theta) \sim U(log(1/M), log(1))$$
$$log(\pi_\beta) \sim U(log(1/J), log(1))$$
We define $h_{SNP}$ and $h_{expr}$ as follows: 
$$
h_{SNP} := \frac{M\sigma_\theta^2}{M\sigma_\theta^2 + Jvar(\tilde{X})\sigma_\beta^2+ 1}$$
$$
h_{expr} := \frac{Jvar(\tilde{X})\sigma_\beta^2}{M\sigma_\theta^2 + Jvar(\tilde{X})\sigma_\beta^2+ 1}$$

Instead of assigning priors for $\sigma_\beta^2$ and $\sigma_\theta^2$, we assign uniform priors for $h_{SNP}$ and $h_{expr}$:

$$ h_{SNP} \sim U(0,1), h_{expr} \sim U(0,1) $$

## Inference
We are interested in the hyperparameters $\omega = (\pi_\beta, \pi_\theta, \tau, \sigma_\beta^2,\sigma_\theta^2)$. We introduce two vectors of binary indicators ($\gamma_\beta$ = $\{\gamma_{\beta1}, \gamma_{\beta2}, ..., \gamma_{\beta J}\} \in \{0,1\}^J$ and $\gamma_\theta$ = $\{\gamma_{\theta1}, \gamma_{\theta2}, ..., \gamma_{\theta M}\} \in \{0,1\}^M$) that indicates whether the corresponding $\beta_j$ or $\theta_m$ is non-zero.

$$\gamma_{\beta j} \sim Bernoulli(\pi_\beta)$$
$$\gamma_{\theta m} \sim Bernoulli(\pi_\theta)$$
$$ \beta_j|\gamma_{\beta j}=1 \sim  N(0, \sigma_\beta^2/\tau), \beta_j|\gamma_{\beta j}=0 \sim  \delta_0$$
$$ \theta_m|\gamma_{\theta m}=1 \sim  N(0, \sigma_\theta^2/\tau), \theta_m|\gamma_{\theta m}=0 \sim  \delta_0$$
The posterior distrition is

$$P(\pi_\beta, \pi_\theta, \tau, h_{expr}, h_{SNP}, \gamma_\beta, \gamma_\theta|y) \propto P(y|\pi_\beta, \pi_\theta, \tau, h_{expr}, h_{SNP}, \gamma_\beta, \gamma_\theta) P(h_{expr})P(h_{SNP})P(\gamma_\beta|\pi_\beta)P( \gamma_\theta|\pi_\theta)P(\pi_\theta)P(\pi_\beta)$$

### 2020-04-01 analysis/VEB-boost-fitting.Rmd
sparse model descriptoin

@@
 Spike and slab prior for $\beta_j$:
-$$\beta_j|\pi,\tau \sim (1- \pi_\beta)\delta_0 + \pi_\beta N(0, \sigma_\beta^2/\tau)$$
+$$\beta_j|\pi_\beta,\tau \sim (1- \pi_\beta)\delta_0 + \pi_\beta N(0, \sigma_\beta^2/\tau)$$
 Spike and slab prior for $\theta_m$:
-$$\theta_m | \tau \sim (1- \pi_\theta)\delta_0 + \pi_\theta N(0, \sigma_\theta^2/\tau)$$
-We want to estimate $\omega=\{\tau,\pi_\beta, \pi_\theta, \sigma_\beta^2,\sigma_\theta^2\}$
+$$\theta_m | \pi_\theta, \tau \sim (1- \pi_\theta)\delta_0 + \pi_\theta N(0, \sigma_\theta^2/\tau)$$
+We want to estimate $\omega=\{\tau,\pi_\beta, \pi_\theta, \sigma_\beta^2,\sigma_\theta^2\}$ and posterior of $\beta$.

### 2020-04-01 analysis/sparse_model.Rmd
sparse model description

@@
 $Y$: quantitative traits, $N$ x 1 vector, $N$ is number of individuals, the columns are centered.
+
 $\tilde{X}$: the genetic component of gene expression, $N$ x $J$ matrix, $J$ is number of genes, the columns are centered.
+
 $G$: genotype (standardized). $N$ x $M$ matrix, $M$ is number of SNPs.
@@
 We use the spike and slab prior for gene expression effect size $\beta_j$ (effect size for gene $j$).
-$$\beta_j|\pi,\tau \sim (1- \pi_\beta)\delta_0 + \pi_\beta N(0, \sigma_\beta^2/\tau)$$
+$$\beta_j|\pi_\beta,\tau \sim (1- \pi_\beta)\delta_0 + \pi_\beta N(0, \sigma_\beta^2/\tau)$$
 Here $\delta_0$ is point mass at 0 ($\delta_0(\beta_j)=1$ if $\beta_j=0$, otherwise $\delta_0(\beta_j)=0$ ).
@@
 We use the spike and slab prior for SNP effect size $\theta_m$ (effect size for SNP $m$).
-$$ \theta_m | \tau \sim (1- \pi_\theta)\delta_0 + \pi_\theta N(0, \sigma_\theta^2/\tau)$$
+$$ \theta_m | \tau, \pi_\theta\sim (1- \pi_\theta)\delta_0 + \pi_\theta N(0, \sigma_\theta^2/\tau)$$
 

### 2020-04-05 analysis/simulation-WTCCC.Rmd
fix description typo

@@
 
-## Simulate cis-expression 
-We simulate cis-expression based on the following model:
+## Simulate expression 
+We simulate expression based on the following model:
 
-$$ X' = G\alpha$$
+$$ X = G\alpha + \xi$$
 For each gene, we sample $L$ eQTLs for this gene, for eQTL $k$, we have
@@
 
-For $\alpha$, we used the same as used in simulating expression (Thus, not $\tilde{X}$).
+For $\alpha$, we used the same as used in simulating expression.
 $\gamma \sim N(0,\sigma_\gamma^2), \theta \sim N(0, \sigma_\theta^2), \epsilon \sim N (0,1)$.

### 2020-04-05 analysis/simulation-WTCCC2.Rmd
fix description typo

@@
 
-## Simulate cis-expression 
-We simulate cis-expression based on the following model, using 500 samples from cad in WTCCC:
+## Simulate expression 
+We simulate expression based on the following model, using 500 samples from cad in WTCCC:
 
-$$ \tilde{X} = G\alpha + \xi$$
+$$ X = G\alpha + \xi$$
 For each gene, we sample $L$ eQTLs for this gene, for eQTL $k$, we have
@@
 $$Y = G\alpha\gamma + G\theta + \epsilon$$.
-For $\alpha$, we used the same as used in simulating expression (Thus, not $\tilde{X}$).
+For $\alpha$, we used the same as used in simulating expression.
 $\gamma \sim N(0,\sigma_\gamma^2), \theta \sim N(0, \sigma_\theta^2), \epsilon \sim N (0,1)$.

### 2020-04-16 analysis/index.Rmd
mr.ash description
veb-boost-repo

@@
     * [Problem description](VEB-boost-fitting.html)
-    * Andrew's [software](https://github.com/andrewg3311/VEB.Boost) and [workflowr website](https://andrewg3311.github.io/VEB-Boost-workflowR/VEB_Boost_Derivation.html)
+    * Andrew's [software](https://github.com/stephenslab/VEB.Boost) and [workflowr website](https://andrewg3311.github.io/VEB-Boost-workflowR/VEB_Boost_Derivation.html)
+
+  * mr.ash
+    * [mr.ash](https://github.com/stephenslab/mr.ash.alpha) package
+    * [mr.ash workflowr website](https://stephenslab.github.io/mr-ash-workflow/)

### 2020-04-16 analysis/mr.ash.Rmd
veb-boost-repo

---
title: "mr.ash"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mr.ash.alpha); # package implementing proposed methods in the paper
library(Matrix); library(mvtnorm) # miscellaneous packages for matrices
library(glmnet); library(ncvreg); library(L0Learn); # penalized regression methods
library(varbvs); library(susieR); library(BGLR); # Bayesian methods
library(ggplot2); library(cowplot); # plotting
```

## Install mr.ash
Following https://github.com/stephenslab/mr.ash.alpha

## mr.ash examples
```{r workflowr}
setwd("/home/simingz/causalTWAS/mr.ash/mr-ash-workflow")
source('code/method_wrapper.R')
source('code/sim_wrapper.R')
data = simulate_data(n = 100, p = 200, s = 5,
                     design = "indepgauss", # "equicorrgauss",
                     signal = "normal", # "lap", "t", etc.
                     pve = 0.5)
str(data)

method_list  = c("mr.ash","varbvs","bayesb","blasso","susie","enet","lasso","ridge","scad2","mcp2","l0learn")
fit.function = get(paste0("fit.", method_list[1], sep = ""))

fit  = fit.function(data$X, data$y, data$X.test, data$y.test)
```

## mr.ash papers
```{r workflowr-paper}
setwd("/home/simingz/causalTWAS/mr.ash/mr-ash-workflow")
filelist = paste("analysis/", list.files("analysis", pattern = "*_Sourcecode.Rmd"), sep = "")
for (i in 1:length(filelist)) {
  workflowr::wflow_publish(filelist[i])
  cat("file",i,": ",'"',filelist[i],'"'," is published\n",sep = "")
}

filelist = paste("analysis/", list.files("analysis", pattern = "*_Plot.Rmd"), sep = "")
for (i in 1:length(filelist)) {
  workflowr::wflow_publish(filelist[i])
  cat("file",i,": ",'"',filelist[i],'"'," is published\n",sep = "")
}
```

### 2020-05-03 analysis/sparse_model.Rmd
test mr.ash

@@
 $$
-h_{SNP} := \frac{M\sigma_\theta^2}{M\sigma_\theta^2 + Jvar(\tilde{X})\sigma_\beta^2+ 1}$$
+h_{SNP} := \frac{\pi_\theta M\sigma_\theta^2}{\pi_\theta M\sigma_\theta^2 + \pi_\beta Jvar(\tilde{X})\sigma_\beta^2+ 1}$$
 $$
-h_{expr} := \frac{Jvar(\tilde{X})\sigma_\beta^2}{M\sigma_\theta^2 + Jvar(\tilde{X})\sigma_\beta^2+ 1}$$
+h_{expr} := \frac{\pi_\beta Jvar(\tilde{X})\sigma_\beta^2}{\pi_\theta M\sigma_\theta^2 + \pi_\beta Jvar(\tilde{X})\sigma_\beta^2+ 1}$$
 

### 2020-05-03 code/run_test_veb-boost.R

@@ -1,0 +1,51 @@
+library(mr.ash.alpha)
+library(VEB.Boost)
+
+args = commandArgs(trailingOnly=TRUE)
+if (length(args) != 3) {
+  stop("3 arguments must be supplied:
+       * genotype file name
+       * param R file
+       * out file name", call.=FALSE)
+}
+
+codedir <- "/project2/mstephens/causalTWAS/causal-TWAS/code/"
+source(paste0(codedir, "stats_func.R"))
+source(paste0(codedir,"input_reformat.R"))
+source(paste0(codedir,"simulate_phenotype.R"))
+
+# genotype plink to R data file
+pfile <- args[1]
+pfileRd <- paste0(pfile, ".Rd")
+
+if (!file.exists(pfileRd)) {
+  print("format converting ...")
+  plink2Rd(pfile, pfileRd); gc()
+} # pfile.traw.gz & .raw.gz should exist
+
+# load genotype data and scale
+load(pfileRd)
+print("scaling ...")
+dat$G <- scaleRcpp(dat$G)
+
+# simulate phenotype
+source(args[2])
+phenolist <- simulate_phenotype(dat, mode = "snp-only",
+                                pve.expr = pve.expr,
+                                pve.snp = pve.snp,
+                                pi_beta =  pi_beta,
+                                pi_theta = pi_theta,
+                                tau = tau)
+
+# write output
+outname <- args[3]
+save(phenolist, file = paste0(outname, "-pheno.Rd"))
+print(warnings())
+
+# run mr.ash
+mr.ash.fit <- mr.ash(dat$G, phenolist$Y)
+mr.ash.fit$data <- NULL
+save(mr.ash.fit, file = paste0(outname,"-mr.ash-res.Rd"))
+
+
+

### 2020-05-03 code/simulate-WTCCC-phenotype.R

@@ -45,7 +45,6 @@ simulate_phenotype<- function(G.scaled, weightslist, M.c, J.c, PVE.snp, PVE.expr
   PVE.snp.truth <- var(s.theta)/var(Y)
   PVE.expr.truth <- var(expr[,idx.cgene] %*% e.gamma)
 
-
   save(Y, sigma_theta, sigma_gamma, s.theta, e.gamma, idx.cSNP, idx.cgene, M.c, J.c, expr.meanvar, file=paste0(paste(outname,M.c, J.c, PVE.snp, PVE.expr,sep="-"), ".simulated_phenotype.Rd"))
   return(list("Y"=Y,"theta"=s.theta, "gamma"=e.gamma))
 }

### 2020-05-20 analysis/VEB-boost-fitting.Rmd
test mr.ash

--- 2020-03-31 analysis/VEB-boost-fitting.Rmd
@@
 
-## Problem description
+## Basic idea of veb_boost
 
-$$Y_{N \times 1}= X_{N \times J}\beta_{J\times1} + G_{N\times M}\theta_{M\times1} + \epsilon_{N\times1}$$
-$$ \epsilon \sim MVN(0,I_N\tau^{-1})$$
-Spike and slab prior for $\beta_j$:
-$$\beta_j|\pi,\tau \sim (1- \pi_\beta)\delta_0 + \pi_\beta N(0, \sigma_\beta^2/\tau)$$
-Spike and slab prior for $\theta_m$:
-$$\theta_m | \tau \sim (1- \pi_\theta)\delta_0 + \pi_\theta N(0, \sigma_\theta^2/\tau)$$
-We want to estimate $\omega=\{\tau,\pi_\beta, \pi_\theta, \sigma_\beta^2,\sigma_\theta^2\}$
+You want to add up 2 linear regressions, which we might call “learners”.
+A learner is, roughly, a fucntion, that learns a relationshipo between Y and some predictors X (here the Xs are different for your 2 different learners) Let’s write this $Y=L1(X1) + L2(X2) + E$.
+
+The idea is that you can fit this model iteratively, by first applying $L1$, and then applying $L2$, each time applying it to the appropriate residuals.So you apply $L1$ to learn a relationship between $R=Y-L2$ (residuals) and $X1$, and then $L2$ to learn a relationship between $R=Y-L1$ and $X2$ and you just iterate….
+
+## Our model
+
+Model:
+
+$$Y = \mu_1 + \mu_2 + \epsilon$$
+$$ \mu_1= X_{N \times J}\beta_{J\times1}$$
+$$ \mu_2 = G_{N\times M}\theta_{M\times1}$$
+$$ \epsilon \sim MVN(0,\sigma^2I_N) $$
+Priors:
+
+ASH prior for scaled $\beta_j$ and $\theta_m$
+$$\beta_j | g_1, \sigma \sim g^1_\sigma(\cdot)$$
+ASH prior for $\theta_m$:
+$$\theta_m | g_2, \sigma \sim g^2_\sigma(\cdot)$$
+
+## Input required by veb_boost
+
+take $\mu1$ as the example, $\mu2$ works similarly.
+
+* first moment of posterior variational approximation ($E_q(\mu_i)\in R^N$)
+
+$$E_q(\mu_1) = X\bar{\beta}$$
+$\bar{\beta}$ is posterior mean of $\beta$ given by `mr.ash` function.
+
+* second moment of posterior variational approximation ($E_q(\mu^2_i)\in R^N_+$)
+
+$$E_q(\mu_{1}^2) = \sum_{j=1}^J{x^2_j E_q(\beta^2)} + [E_q(\mu_1)]^2 =\sum_{j=1}^J\sum_{k=1}^K{x^2_j\phi_k \frac{\sigma^2 d_j^2 \sigma_k^2}{1+ d_j^2 \sigma_k^2}} + [E_q(\mu_1)]^2$$
+where $$d_j=x^T_jx_j= \left\lVert x_j\right\rVert^2$$.
+
+$d_j, \phi, \sigma, \sigma_k$ are estimates given by `mr.ash` function.
+
+* KL-divergence between variational posterior (q) and prior (p).
+
+Shall we just use ELBO from mr.ash?

### 2020-05-20 analysis/index.Rmd
test mr.ash

@@
   * [Sparse model](sparse_model.html)
+  * [mr.ash model](mrash_model.html)
+
+# Data
+
+## Expression data
+  * [Fusion/TWAS gene expression models](fusion_expr.html)
 
@@
 
-## For a single gene
-  * [Single gene](simulation-singleG.html)
+## Proof of concept
+  * [Proof of concept for this project using a single gene's data](simulation-singleG.html)
 
-## Under polygenic model
-  * [Simulation: WTCCC](simulation-WTCCC.html)
-  * [Simulation: WTCCC 2](simulation-WTCCC2.html)
+## Test polygenic model
+Using WTCCC data
 
-# References
+  * [Test polygenic model 1](simulation-WTCCC.html)
+
+  * [Test polygenic model 2: use imputed expression](simulation-WTCCC2.html)
+
+## Test mr.ash
+Using UKbiobank data (chr22)
+
+  * [Test mr.ash performance: only SNP effect](simulation-mr.ash-ukbchr22.html)
+
+## Test veb_boost
+A simplified version: interative update of two mr.ash component
+
+  * [Using some simple data](simulation_simpleveb-boost.html)
+
+  * [Using UKbiobank data (chr22)](simulation_simpleveb-boost_ukbchr22.html)
+
+# Related software
   * VEB-boost

### 2020-05-20 analysis/simulation-mr.ash-ukbchr22.Rmd
test mr.ash

---
title: "Test mr.ash performance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r plot functions}
library(mr.ash.alpha)
source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")

summary_mr.ash <- function(fit){
  cat("pi1 = ", 1-fit$pi[[1]], "\n")
  pve <- get_pve(fit)
  cat("pve : ", pve, "\n")
}

plot_beta <- function(beta,beta.pm, ...){
  plot( beta, pch=19, col ="darkgreen", ...)
  points(beta.pm, pch =19, col = "red")
  legend("topright", legend=c("true beta", "posterior mean"),
       col=c("darkgreen", "red"), pch=19)
}
```

```{r output dir}
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200429/"
```

## Simulation 1
True parameters:
```{r para 1}
readLines(paste0(outputdir, "param-20200429-1.R"))
```

ASH results:

```{r sim 1}
load(paste0(outputdir, "mr.ash_20200429-1-mr.ash-res.Rd"))
summary_mr.ash(mr.ash.fit)
```

## Simulation 2

True parameters:
```{r para 2}
readLines(paste0(outputdir, "param-20200429-2.R"))
```

ASH results:

```{r sim 2}
load(paste0(outputdir, "mr.ash_20200429-2-mr.ash-res.Rd"))
summary_mr.ash(mr.ash.fit)
```

## Simulation 3

True parameters:
```{r para 3}
readLines(paste0(outputdir, "param-20200429-3.R"))
```

ASH results:

```{r sim 3}
load(paste0(outputdir, "mr.ash_20200429-3-mr.ash-res.Rd"))
summary_mr.ash(mr.ash.fit)
```

## Simulation 4

True parameters:
```{r para 4}
readLines(paste0(outputdir, "param-20200429-4.R"))
```

ASH results:

```{r sim 4}
load(paste0(outputdir, "mr.ash_20200429-4-mr.ash-res.Rd"))
summary_mr.ash(mr.ash.fit)
```

### 2020-05-20 analysis/simulation-simpleveb-boost-ukbchr22.Rmd
test mr.ash

---
title: "Test a simpler version of veb-boost performance on ukbiobank data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r plot functions}
library(mr.ash.alpha)
source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")

summary_mr.ash <- function(fit){
  cat("pi1 = ", 1-fit$pi[[1]], "\n")
  pve <- get_pve(fit)
  cat("pve = ", pve, "\n")
}

plot_beta <- function(beta,beta.pm, ...){
  plot( beta, pch=19, col ="darkgreen", ...)
  points(beta.pm, pch =19, col = "red")
  legend("topright", legend=c("true beta", "posterior mean"),
       col=c("darkgreen", "red"), pch=19)
}

summary_mr.ash2 <- function(g.fit, s.fit, phenores){
  e.b <- rep(0, length(g.fit$beta))
  e.b[phenores$param$idx.cgene] <- phenores$param$e.beta
  cat("gene expression effect: \n")
  summary_mr.ash(g.fit)
  plot_beta(e.b, g.fit$beta)

  s.b <- rep(0, length(s.fit$beta))
  s.b[phenores$param$idx.cSNP] <- phenores$param$s.theta
  cat("snp effect: \n")
  summary_mr.ash(s.fit)
  plot_beta(s.b, s.fit$beta)
}

```

```{r output dir}
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200503/"
```

## Simulation 1
True parameters:
```{r para 1}
load(paste0(outputdir, "mr.ash2_20200503-1-pheno.Rd"))
readLines(paste0(outputdir, "param-20200503-1.R"))
```

MR.ASH results (start from gene):
```{r sim 1}
load(paste0(outputdir, "mr.ash2_20200503-1-mr.ash2.expr-res.Rd"))
g.fit <- mr.ash2.fit$fit1
s.fit <-  mr.ash2.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)
```

MR.ASH results (start from SNP):
```{r sim 1}
load(paste0(outputdir, "mr.ash2_20200503-1-mr.ash2.snp-res.Rd"))
g.fit <- mr.ash2.fit$fit2
s.fit <-  mr.ash2.fit$fit1
summary_mr.ash2(g.fit, s.fit, phenores)
```

## Simulation 2

True parameters:
```{r para 2}
readLines(paste0(outputdir, "param-20200429-2.R"))
```

ASH results:

```{r sim 2}
load(paste0(outputdir, "mr.ash_20200429-2-mr.ash-res.Rd"))
summary_mr.ash(mr.ash.fit)
```

## Simulation 3

True parameters:
```{r para 3}
readLines(paste0(outputdir, "param-20200429-3.R"))
```

ASH results:

```{r sim 3}
load(paste0(outputdir, "mr.ash_20200429-3-mr.ash-res.Rd"))
summary_mr.ash(mr.ash.fit)
```

## Simulation 4

True parameters:
```{r para 4}
readLines(paste0(outputdir, "param-20200429-4.R"))
```

ASH results:

```{r sim 4}
load(paste0(outputdir, "mr.ash_20200429-4-mr.ash-res.Rd"))
summary_mr.ash(mr.ash.fit)
```

### 2020-05-20 analysis/simulation_simpleveb-boost.Rmd
test mr.ash

---
title: "Simple simulations for a simplified version of web-boost"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mr.ash.alpha)
source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")
source("~/causalTWAS/causal-TWAS/code/simple_vebboost.R")
```

## Simulation of data

20 blocks:

* Each block has either gene or SNP effect
* Each block has 99 SNPs and 1 gene. Each gene is linear sum of the previous 3 SNPs. 
* Each block, either the gene or the last eQTL has non-zero effect on trait

The first 4 blocks have gene effect.

```{r generate data}
set.seed(1)
N <- 4000
nblocks <- 20
block.size <- 100
p <- nblocks * block.size
n.eQTL <- 3  # number of eQTLs per gene
sigma.eQTL <- 0.5 # eQTL effect size
sigma.SNP <- 0.1 # effect size of causal SNP on trait
sigma.gene <- 0.1 # effect size of causal gene on trait
X <- matrix(rep(0,0), nrow=N, ncol=0)
gamma.gene <- rep(0, nblocks) # indicator of genes
gamma.gene[1:4] <- 1 
beta <- numeric(0)
SNP.idx <- numeric(0)
for (i in 1:nblocks) {
  # sample SNP data
  X.block.SNP <- matrix(rnorm(N*(block.size-1)), nrow=N, ncol=block.size-1)
  SNP.idx <- c(SNP.idx, 1:(block.size-1) + (i-1)*block.size)

  # generate gene data: use the previous few SNPs as eQTL
  effects.eQTL <- rnorm(n.eQTL, 0, sigma.eQTL)
  X.block.gene <- X.block.SNP[, (block.size - n.eQTL):(block.size - 1)] %*% effects.eQTL
  X.block = cbind(X.block.SNP, X.block.gene)
  X <- cbind(X, X.block)

  # sample beta
  if (gamma.gene[i] == 1) { # gene effect in this block
    beta.SNP <- rep(0, block.size - 1)
    beta.gene <- rnorm(1, 0, sigma.gene)
  } else { # SNP effect in this block
    beta.SNP <- c(rep(0, block.size - 2), rnorm(1, 0, sigma.SNP))
    beta.gene <- 0
  }
  beta.block <- c(beta.SNP, beta.gene)
  beta <- c(beta, beta.block)
}
sigma.e <- 1
y <- X %*% beta + rnorm(N, 0, sigma.e)
gene.idx <- (1:nblocks) * block.size
```

## Run mr.ash

```{r plot functions}
summary_mr.ash <- function(fit){
  cat("pi1 = ", 1-fit$pi[[1]], "\n")
  pve <- get_pve(fit)
  cat("pve : ", pve, "\n")
}

plot_beta <- function(beta,beta.pm, ...){
  plot( beta, pch=19, col ="darkgreen", ...)
  points(beta.pm, pch =19, col = "red")
  legend("topright", legend=c("true beta", "posterior mean"),
       col=c("darkgreen", "red"), pch=19)
}

```

```{r run mr.ash}
fit <- mr.ash(X, y, method="caisa")
summary_mr.ash(fit)

plot_beta(beta[gene.idx], fit$beta[gene.idx], main = "beta for gene effect")
plot_beta(beta[SNP.idx], fit$beta[SNP.idx], main = "beta for SNP effect")
```

## Run a simplified version of veb-boost

### start with gene

```{r run mr.ash2, gene}
X.gene <- X[,gene.idx]
X.SNP <- X[, SNP.idx]
fit <- mr.ash2(X.gene, X.SNP, y)

print("for gene effect: ")
summary_mr.ash(fit$fit1)
plot_beta(beta[gene.idx], fit$fit1$beta, main = "beta for gene effect")

print("for SNP effect: ")
summary_mr.ash(fit$fit2)
plot_beta(beta[SNP.idx], fit$fit2$beta, main = "beta for SNP effect")
```

### start with SNP

```{r run mr.ash2, snp}
X.gene <- X[,gene.idx]
X.SNP <- X[, SNP.idx]
fit <- mr.ash2(X.SNP, X.gene, y)

print("for gene effect: ")
summary_mr.ash(fit$fit2)
plot_beta(beta[gene.idx], fit$fit2$beta, main = "beta for gene effect")

print("for SNP effect: ")
summary_mr.ash(fit$fit1)
plot_beta(beta[SNP.idx], fit$fit1$beta, main = "beta for SNP effect")
```

### 2020-05-20 code/run_test_veb-boost.R

@@ -2,9 +2,10 @@ library(mr.ash.alpha)
 library(VEB.Boost)
 
 args = commandArgs(trailingOnly=TRUE)
-if (length(args) != 3) {
-  stop("3 arguments must be supplied:
+if (length(args) != 4) {
+  stop("4 arguments must be supplied:
        * genotype file name
+       * weight dir
        * param R file
        * out file name", call.=FALSE)
 }
@@ -13,6 +14,7 @@ codedir <- "/project2/mstephens/causalTWAS/causal-TWAS/code/"
 source(paste0(codedir, "stats_func.R"))
 source(paste0(codedir,"input_reformat.R"))
 source(paste0(codedir,"simulate_phenotype.R"))
+source(paste0(codedir,"simple_vebboost.R"))
 
 # genotype plink to R data file
 pfile <- args[1]
@@ -29,8 +31,16 @@ print("scaling ...")
 dat$G <- scaleRcpp(dat$G)
 
 # simulate phenotype
-source(args[2])
-phenolist <- simulate_phenotype(dat, mode = "snp-only",
+weight <- args[2]
+exprres <- cis_expr(dat, weight, method = "best", checksnps = F)
+
+outname <- args[4]
+save(exprres, file = paste0(outname, "-cis-expr.Rd"))
+# load(paste0(outname, "-cis-expr.Rd"))
+dat$expr <-  scaleRcpp(exprres$expr)
+
+source(args[3])
+phenores <- simulate_phenotype(dat, mode = "snp-expr",
                                 pve.expr = pve.expr,
                                 pve.snp = pve.snp,
                                 pi_beta =  pi_beta,
@@ -38,14 +48,23 @@ phenolist <- simulate_phenotype(dat, mode = "snp-only",
                                 tau = tau)
 
 # write output
-outname <- args[3]
-save(phenolist, file = paste0(outname, "-pheno.Rd"))
+save(phenores, file = paste0(outname, "-pheno.Rd"))
 print(warnings())
 
-# run mr.ash
-mr.ash.fit <- mr.ash(dat$G, phenolist$Y)
-mr.ash.fit$data <- NULL
-save(mr.ash.fit, file = paste0(outname,"-mr.ash-res.Rd"))
+
+# run mr.ash2 (a simplified version of veb_boost)
+## start with expression
[…]

### 2020-05-20 analysis/simulation-simpleveb-boost-ukbchr22.Rmd
test mr.ash2

@@
        col=c("darkgreen", "red"), pch=19)
+}
+
+plot_pip <- function(indi, pip) {
+  barplot(indi)
 }
@@
 MR.ASH results (start from gene):
-```{r sim 1}
+```{r sim 1 gene}
 load(paste0(outputdir, "mr.ash2_20200503-1-mr.ash2.expr-res.Rd"))
@@
 MR.ASH results (start from SNP):
-```{r sim 1}
+```{r sim 1 snp}
 load(paste0(outputdir, "mr.ash2_20200503-1-mr.ash2.snp-res.Rd"))
@@
 ```{r para 2}
-readLines(paste0(outputdir, "param-20200429-2.R"))
+load(paste0(outputdir, "mr.ash2_20200503-2-pheno.Rd"))
+readLines(paste0(outputdir, "param-20200503-2.R"))
 ```
 
-ASH results:
+MR.ASH results (start from gene):
+```{r sim 2 gene}
+load(paste0(outputdir, "mr.ash2_20200503-2-mr.ash2.expr-res.Rd"))
+g.fit <- mr.ash2.fit$fit1
+s.fit <-  mr.ash2.fit$fit2
+summary_mr.ash2(g.fit, s.fit, phenores)
+```
 
-```{r sim 2}
-load(paste0(outputdir, "mr.ash_20200429-2-mr.ash-res.Rd"))
-summary_mr.ash(mr.ash.fit)
+MR.ASH results (start from SNP):
+```{r sim 2 snp}
+load(paste0(outputdir, "mr.ash2_20200503-2-mr.ash2.snp-res.Rd"))
+g.fit <- mr.ash2.fit$fit2
+s.fit <-  mr.ash2.fit$fit1
+summary_mr.ash2(g.fit, s.fit, phenores)
 ```
@@
 ```{r para 3}
-readLines(paste0(outputdir, "param-20200429-3.R"))
+load(paste0(outputdir, "mr.ash2_20200503-3-pheno.Rd"))
+readLines(paste0(outputdir, "param-20200503-3.R"))
 ```
 
-ASH results:
+MR.ASH results (start from gene):
+```{r sim 3 gene}
+load(paste0(outputdir, "mr.ash2_20200503-3-mr.ash2.expr-res.Rd"))
+g.fit <- mr.ash2.fit$fit1
+s.fit <-  mr.ash2.fit$fit2
+summary_mr.ash2(g.fit, s.fit, phenores)
+```
 
-```{r sim 3}
-load(paste0(outputdir, "mr.ash_20200429-3-mr.ash-res.Rd"))
-summary_mr.ash(mr.ash.fit)
+MR.ASH results (start from SNP):
+```{r sim 3 snp}
+load(paste0(outputdir, "mr.ash2_20200503-3-mr.ash2.snp-res.Rd"))
+g.fit <- mr.ash2.fit$fit2
+s.fit <-  mr.ash2.fit$fit1
+summary_mr.ash2(g.fit, s.fit, phenores)
 ```
@@
 ```{r para 4}
-readLines(paste0(outputdir, "param-20200429-4.R"))
+load(paste0(outputdir, "mr.ash2_20200503-4-pheno.Rd"))
+readLines(paste0(outputdir, "param-20200503-4.R"))
 ```
 
-ASH results:
+MR.ASH results (start from gene):
+```{r sim 4 gene}
+load(paste0(outputdir, "mr.ash2_20200503-4-mr.ash2.expr-res.Rd"))
+g.fit <- mr.ash2.fit$fit1
+s.fit <-  mr.ash2.fit$fit2
+summary_mr.ash2(g.fit, s.fit, phenores)
+```
 
-```{r sim 4}
-load(paste0(outputdir, "mr.ash_20200429-4-mr.ash-res.Rd"))
-summary_mr.ash(mr.ash.fit)
+MR.ASH results (start from SNP):
+```{r sim 4 snp}
+load(paste0(outputdir, "mr.ash2_20200503-4-mr.ash2.snp-res.Rd"))
+g.fit <- mr.ash2.fit$fit2
+s.fit <-  mr.ash2.fit$fit1
+summary_mr.ash2(g.fit, s.fit, phenores)
 ```

### 2020-05-20 code/run_test_mr.ash.R

@@ -51,7 +51,6 @@ mr.ash.fit$data$y <- NULL
 
 mr.ash.fit$t <- t.mr.ash[3]
 
-
 print("mr.ash finished ... ")
 
 save(mr.ash.fit, file = paste0(outname,"-mr.ash-res.Rd"))

### 2020-05-20 analysis/index.Rmd
test mr.ash2

@@
 ## Test veb_boost
-A simplified version: interative update of two mr.ash component
+A simplified version: interative update of two mr.ash components
 
@@
 
-  * [Using UKbiobank data (chr22)](simulation_simpleveb-boost_ukbchr22.html)
+  * [Using UKbiobank data (chr22)](simulation-simpleveb-boost-ukbchr22.html)
 

### 2020-05-20 analysis/simulation-simpleveb-boost-ukbchr22.Rmd
test mr.ash2

@@
 
-plot_pip <- function(indi, pip) {
-  barplot(indi)
+plot_pip <- function(indi, pip, pipcut = 0.1) {
+  x <- barplot(indi, main = "PIP")
+  points(x, pip, pch=19, col="red")
+  pred <- which(pip > pipcut)
+  real <- which(indi==1)
+  tp <- length(intersect(pred, real))/length(pred)
+  fp <- 1 - tp
+  fn <- length(setdiff(real, pred))/length(real)
+  cat("false positive rate: ", fp)
+  cat("false negative rate: ", fn)
 }
 
-summary_mr.ash2 <- function(g.fit, s.fit, phenores){
+summary_mr.ash2 <- function(g.fit, s.fit, phenores, pipcut = 0.1){
   e.b <- rep(0, length(g.fit$beta))
@@
   plot_beta(e.b, g.fit$beta)
+  indi <- rep(0, length(g.fit$beta))
+  indi[phenores$param$idx.cgene] <- 1
+  pip <- get_pip(g.fit)
+  plot_pip(indi, pip, pipcut = pipcut)
 

### 2020-05-20 analysis/simulation-simpleveb-boost-ukbchr22.Rmd
test mr.ash2

@@
   fn <- length(setdiff(real, pred))/length(real)
-  cat("false positive rate: ", fp)
-  cat("false negative rate: ", fn)
+  cat("false positive rate: ", fp, "\n")
+  cat("false negative rate: ", fn, "\n")
 }

### 2020-05-28 analysis/index.Rmd
mr.ash2s

@@
 ## Test veb_boost
-A simplified version: interative update of two mr.ash components
+A simplified version (mr.ash2): interative update of two mr.ash components
 
@@
   * [Using UKbiobank data (chr22)](simulation-simpleveb-boost-ukbchr22.html)
+
+A simplified version (mr.ash2s): interative update of two mr.ash components, now including updates of sigma2 in each iteration and initiate with mr.ash for all.
+
+  * [Using some simple data](simulation_simpleveb-boost2.html)
+
+  * [Using UKbiobank data (chr22)](simulation-simpleveb-boost2-ukbchr22.html)
+
+## Test SUSIE for fine-mapping
+
+  * [Using some simple data](simulation_susie.html)
+
+  * [Using UKbiobank data (chr22)](simulation-susie-ukbchr22.html)
 

### 2020-05-28 analysis/simulation-simpleveb-boost2-ukbchr22.Rmd
mr.ash2s

---
title: "Test a simpler version of veb-boost performance on ukbiobank data-2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r plot functions}
library(mr.ash.alpha)
source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")

summary_mr.ash <- function(fit){
  cat("pi1 = ", 1-fit$pi[[1]], "\n")
  pve <- get_pve(fit)
  cat("pve = ", pve, "\n")
}

plot_beta <- function(beta,beta.pm, ...){
  plot( beta, pch=19, col ="darkgreen", ...)
  points(beta.pm, pch =19, col = "red")
  legend("topright", legend=c("true beta", "posterior mean"),
       col=c("darkgreen", "red"), pch=19)
}

plot_pip <- function(indi, pip, pipcut = 0.1) {
  x <- barplot(indi, main = "PIP")
  points(x, pip, pch=19, col="red")
  pred <- which(pip > pipcut)
  real <- which(indi==1)
  tp <- length(intersect(pred, real))/length(pred)
  fp <- 1 - tp
  fn <- length(setdiff(real, pred))/length(real)
  cat("false positive rate: ", fp, "\n")
  cat("false negative rate: ", fn, "\n")
}

summary_mr.ash2 <- function(g.fit, s.fit, phenores, pipcut = 0.1){
  e.b <- rep(0, length(g.fit$beta))
  e.b[phenores$param$idx.cgene] <- phenores$param$e.beta
  cat("gene expression effect: \n")
  summary_mr.ash(g.fit)
  plot_beta(e.b, g.fit$beta)
  indi <- rep(0, length(g.fit$beta))
  indi[phenores$param$idx.cgene] <- 1
  pip <- get_pip(g.fit)
  plot_pip(indi, pip, pipcut = pipcut)

  s.b <- rep(0, length(s.fit$beta))
  s.b[phenores$param$idx.cSNP] <- phenores$param$s.theta
  cat("snp effect: \n")
  summary_mr.ash(s.fit)
  plot_beta(s.b, s.fit$beta)
}

```

```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200503/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200527/"
```

## Simulation 1
True parameters:
```{r para 1}
load(paste0(simdatadir, "mr.ash2_20200503-1-pheno.Rd"))
readLines(paste0(simdatadir, "param-20200503-1.R"))
```

MR.ASH2s results (start from gene):
```{r sim 1 gene}
load(paste0(outputdir, "20200527-1-mr.ash2s.expr-res.Rd"))
g.fit <- mr.ash2s.fit$fit1
s.fit <-  mr.ash2s.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)
```

MR.ASH results (start from SNP):
```{r sim 1 snp}
load(paste0(outputdir, "20200527-1-mr.ash2s.snp-res.Rd"))
g.fit <- mr.ash2s.fit$fit2
s.fit <-  mr.ash2s.fit$fit1
summary_mr.ash2(g.fit, s.fit, phenores)
```

## Simulation 2

True parameters:
```{r para 2}
load(paste0(simdatadir, "mr.ash2_20200503-2-pheno.Rd"))
readLines(paste0(simdatadir, "param-20200503-2.R"))
```

MR.ASH results (start from gene):
```{r sim 2 gene}
load(paste0(outputdir, "20200527-2-mr.ash2s.expr-res.Rd"))
g.fit <- mr.ash2s.fit$fit1
s.fit <-  mr.ash2s.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)
```

MR.ASH results (start from SNP):
```{r sim 2 snp}
load(paste0(outputdir, "20200527-2-mr.ash2s.snp-res.Rd"))
g.fit <- mr.ash2s.fit$fit2
s.fit <-  mr.ash2s.fit$fit1
summary_mr.ash2(g.fit, s.fit, phenores)
```

## Simulation 3

True parameters:
```{r para 3}
load(paste0(simdatadir, "mr.ash2_20200503-3-pheno.Rd"))
readLines(paste0(simdatadir, "param-20200503-3.R"))
```

MR.ASH results (start from gene):
```{r sim 3 gene}
load(paste0(outputdir, "20200527-3-mr.ash2s.expr-res.Rd"))
g.fit <- mr.ash2s.fit$fit1
s.fit <-  mr.ash2s.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)
```

MR.ASH results (start from SNP):
```{r sim 3 snp}
load(paste0(outputdir, "20200527-3-mr.ash2s.snp-res.Rd"))
g.fit <- mr.ash2s.fit$fit2
s.fit <-  mr.ash2s.fit$fit1
summary_mr.ash2(g.fit, s.fit, phenores)
```

## Simulation 4

True parameters:
```{r para 4}
load(paste0(simdatadir, "mr.ash2_20200503-4-pheno.Rd"))
readLines(paste0(simdatadir, "param-20200503-4.R"))
```

MR.ASH results (start from gene):
```{r sim 4 gene}
load(paste0(outputdir, "20200527-4-mr.ash2s.expr-res.Rd"))
g.fit <- mr.ash2s.fit$fit1
s.fit <-  mr.ash2s.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)
```

MR.ASH results (start from SNP):
```{r sim 4 snp}
load(paste0(outputdir, "20200527-4-mr.ash2s.snp-res.Rd"))
g.fit <- mr.ash2s.fit$fit2
s.fit <-  mr.ash2s.fit$fit1
summary_mr.ash2(g.fit, s.fit, phenores)
```

### 2020-05-28 analysis/simulation_simpleveb-boost2.Rmd
mr.ash2s

--- 2020-05-20 analysis/simulation_simpleveb-boost.Rmd
@@
 ---
-title: "Simple simulations for a simplified version of web-boost"
+title: "Simple simulations for a simplified version of web-boost-2"
 output: html_document
@@
 source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")
-source("~/causalTWAS/causal-TWAS/code/simple_vebboost.R")
+source("~/causalTWAS/causal-TWAS/code/mr.ash2.R")
 ```
@@
 
-## Run a simplified version of veb-boost
+## Run a simplified version of veb-boost (mr.ash2s)
 
@@
 X.SNP <- X[, SNP.idx]
-fit <- mr.ash2(X.gene, X.SNP, y)
+fit <- mr.ash2s(X.gene, X.SNP, y)
 
@@
 X.SNP <- X[, SNP.idx]
-fit <- mr.ash2(X.SNP, X.gene, y)
+fit <- mr.ash2s(X.SNP, X.gene, y)
 

### 2020-05-28 analysis/simulation_susie.Rmd
mr.ash2s

--- 2020-05-28 analysis/simulation_simpleveb-boost2.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE)
-library(mr.ash.alpha)
-source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")
-source("~/causalTWAS/causal-TWAS/code/mr.ash2.R")
+library(susieR)
 ```
@@
 
-## Run mr.ash
-
 ```{r plot functions}
-summary_mr.ash <- function(fit){
-  cat("pi1 = ", 1-fit$pi[[1]], "\n")
-  pve <- get_pve(fit)
-  cat("pve : ", pve, "\n")
-}
-
-plot_beta <- function(beta,beta.pm, ...){
-  plot( beta, pch=19, col ="darkgreen", ...)
-  points(beta.pm, pch =19, col = "red")
-  legend("topright", legend=c("true beta", "posterior mean"),
-       col=c("darkgreen", "red"), pch=19)
+plot_pip <- function(indi, pip, pipcut = 0.1) {
+  x <- barplot(indi, main = "PIP")
+  points(x, pip, pch=19, col="red")
+  pred <- which(pip > pipcut)
+  real <- which(indi==1)
+  tp <- length(intersect(pred, real))/length(pred)
+  fp <- 1 - tp
+  fn <- length(setdiff(real, pred))/length(real)
+  cat("false positive rate: ", fp, "\n")
+  cat("false negative rate: ", fn, "\n")
 }
@@
 
-```{r run mr.ash}
-fit <- mr.ash(X, y, method="caisa")
-summary_mr.ash(fit)
+## Run susie
 
-plot_beta(beta[gene.idx], fit$beta[gene.idx], main = "beta for gene effect")
-plot_beta(beta[SNP.idx], fit$beta[SNP.idx], main = "beta for SNP effect")
-```
+Fix prior as given, using truth.
 
-## Run a simplified version of veb-boost (mr.ash2s)
+```{r run susie, gene}
+p.snp <- length(SNP.idx)
+p.gene <- length(gene.idx)
 
-### start with gene
+prior.SNP <- 16/2000
+prior.gene <- 4/20
+indi.gene <- c(rep(1,4), rep(0,16))
 
-```{r run mr.ash2, gene}
-X.gene <- X[,gene.idx]
-X.SNP <- X[, SNP.idx]
-fit <- mr.ash2s(X.gene, X.SNP, y)
+prior <- rep(0, dim(X)[2])
+prior[gene.idx] <- prior.gene
+prior[SNP.idx] <- prior.SNP
+
+res <- susie(X,y,L=1, prior_weights=prior)
 
 print("for gene effect: ")
-summary_mr.ash(fit$fit1)
-plot_beta(beta[gene.idx], fit$fit1$beta, main = "beta for gene effect")
-
-print("for SNP effect: ")
-summary_mr.ash(fit$fit2)
-plot_beta(beta[SNP.idx], fit$fit2$beta, main = "beta for SNP effect")
+plot_pip(indi.gene, res$pip[gene.idx])
 ```
-
-### start with SNP
-
-```{r run mr.ash2, snp}
-X.gene <- X[,gene.idx]
-X.SNP <- X[, SNP.idx]
-fit <- mr.ash2s(X.SNP, X.gene, y)
-
-print("for gene effect: ")
-summary_mr.ash(fit$fit2)
-plot_beta(beta[gene.idx], fit$fit2$beta, main = "beta for gene effect")
-
-print("for SNP effect: ")
-summary_mr.ash(fit$fit1)
-plot_beta(beta[SNP.idx], fit$fit1$beta, main = "beta for SNP effect")
-```

### 2020-05-28 code/mr.ash2.R

@@ -1,0 +1,100 @@
+#' add up two mr.ash models.
+#' y = X1 + X2 + e
+#' @param iter iterations
+
+mr.ash2 <- function(X1, X2, y, iter = 20){
+  r1 <- y
+  fit1 <- mr.ash(X1, r1, method = "caisa")
+  for (i in 1:iter) {
+    r2 <- y - X1 %*% fit1$beta
+    fit2 <- mr.ash(X2, r2, method = "caisa")
+
+    r1 <- y - X2 %*% fit2$beta
+    fit1 <- mr.ash(X1, r1, method = "caisa")
+    gc()
+  }
+  fit <- list("fit1" = fit1, "fit2" = fit2)
+  return(fit)
+}
+
+#' add up two mr.ash models.
+#' update of sigma, init with mr.ash, one iter each for X1 and X2.
+mr.ash2s <- function(X1, X2, y, iter = 100){
+  n <- dim(X1)[1]
+  p1 <- dim(X1)[2]
+  p2 <- dim(X2)[2]
+
+  # initiate using mr.ash for X1, X2 combined
+  X <- cbind(X1, X2)
+  fit.init <- mr.ash(X, y)
+
+  beta1 <- fit.init$beta[1: p1, , drop = F]
+  beta2 <- fit.init$beta[(p1+1): (p1 + p2), , drop = F]
+
+  pi1 <- fit.init$pi
+  pi2 <- fit.init$pi
+  sa2.1 <- fit.init$sa2
+  sa2.2 <- fit.init$sa2
+
+  r <- y - X1 %*% beta1 - X2 %*% beta2
+  sigma2 <- fit.init$sigma2
+
+  K <- dim(fit.init$pi)[1]
+
+  rm(fit.init)
+  rm(X);gc()
+
+  # start iteration (from X1)
+  piK1_rec <- rep(0, iter)
+  piK2_rec <- rep(0, iter)
+  sigma2_rec <- rep(0, iter)
+
+  for (i in 1:iter) {
+    y1 <- y - X2 %*% beta2
+    fit1 <- mr.ash(X1, y1,
+                   beta.init = beta1,
+                   sigma2 = sigma2,
+                   sa2 = sa2.1,
+                   pi = pi1,
+                   max.iter = 1)
+
+    y2 <- y - X1 %*% beta1
+
+    fit2 <- mr.ash(X2, y2,
+                   beta.init = beta2,
[…]

### 2020-05-28 code/simple_vebboost.R

@@ -1,15 +1,0 @@
-# fit gene first, then with SNP
-mr.ash2 <- function(X1, X2, y, iter = 20){
-  r1 <- y
-  fit1 <- mr.ash(X1, r1, method = "caisa")
-  for (i in 1:iter) {
-    r2 <- y - X1 %*% fit1$beta
-    fit2 <- mr.ash(X2, r2, method = "caisa")
-
-    r1 <- y - X2 %*% fit2$beta
-    fit1 <- mr.ash(X1, r1, method = "caisa")
-    gc()
-  }
-  fit <- list("fit1" = fit1, "fit2" = fit2)
-  return(fit)
-}

### 2020-06-11 analysis/index.Rmd
plotly
Former-commit-id: a3bcd8a2db83a1e3728cc6c1216da4d0a67c0617

@@
 
-  * [Using UKbiobank data (chr22)](simulation-simpleveb-boost-ukbchr22.html)
+  * [Using UKbiobank data (chr22)](https://users.rcc.uchicago.edu/~simingz/simulation-simpleveb-boost-ukbchr22.html)
 
@@
 
-  * [Using UKbiobank data (chr22)](simulation-simpleveb-boost2-ukbchr22.html)
+  * [Using UKbiobank data (chr22)](https://users.rcc.uchicago.edu/~simingz/simulation-simpleveb-boost2-ukbchr22.html)
 

### 2020-06-11 analysis/simulation-simpleveb-boost-ukbchr22.Rmd
plotly
Former-commit-id: a3bcd8a2db83a1e3728cc6c1216da4d0a67c0617

--- 2020-05-20 analysis/simulation-simpleveb-boost-ukbchr22.Rmd
@@
 library(mr.ash.alpha)
+library(data.table)
+suppressMessages({library(plotly)})
+
 source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")
@@
 
-plot_beta <- function(beta,beta.pm, ...){
-  plot( beta, pch=19, col ="darkgreen", ...)
-  points(beta.pm, pch =19, col = "red")
-  legend("topright", legend=c("true beta", "posterior mean"),
-       col=c("darkgreen", "red"), pch=19)
-}
-
-plot_pip <- function(indi, pip) {
-  barplot(indi)
-}
-
-summary_mr.ash2 <- function(g.fit, s.fit, phenores){
+summary_mr.ash2 <- function(g.fit, s.fit, phenores, pipcut = 0.1){
   e.b <- rep(0, length(g.fit$beta))
@@
   summary_mr.ash(g.fit)
-  plot_beta(e.b, g.fit$beta)
 
@@
   summary_mr.ash(s.fit)
-  plot_beta(s.b, s.fit$beta)
+}
+
+plot_pip <- function(dt){
+  pal <- c("salmon", "darkgreen")
+  pal <- setNames(pal, c("Causal", "Non causal"))
+
+  fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))
+
+  fig <- fig %>% layout(yaxis = list(title = "PIP"))
+  return(fig)
+}
+
+plot_beta <- function(dt){
+  fig <- plot_ly(dt, x = ~p0, y = ~ TrueBeta, type = 'scatter', mode = 'markers', name ="Truth", text = ~paste("Name: ", name))
+  fig <- fig %>% add_trace(y = ~ EstBeta, name = 'Estimated',mode = 'markers')
+
+  fig <- fig %>% layout(yaxis = list(title = "beta"))
+  return(fig)
+}
+
+plot_p <- function(dt){
+  pal <- c("salmon", "darkgreen")
+  pal <- setNames(pal, c("Causal", "Non causal"))
+
+  fig <- plot_ly(dt, x = ~p0, y = ~ PVALUE , type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))
+
+  fig <- fig %>% layout(yaxis = list(title = "p value"))
+  return(fig)
+}
+
+plot_all <- function(mrashres.gene, gwasres.gene, mrashres.snp, gwasres.snp){
+  a <- read.table(mrashres.gene , header =T)
+  a[a$ifcausal == 1, "ifcausal"] = "Causal"
+  a[a$ifcausal == 0, "ifcausal"] = "Non causal"
+
+  p <- fread(gwasres.gene, header =T)
+  p[,"PVALUE"] <- -log10(p[,"PVALUE"])
+  colnames(p)[colnames(p) == "MARKER_ID"] <-  "name"
+  dt.gene <- merge(a, p, by = "name", all.x = T)
+
+  a <- read.table(mrashres.snp , header =T)
+  a[a$ifcausal == 1, "ifcausal"] = "Causal"
+  a[a$ifcausal == 0, "ifcausal"] = "Non causal"
+
+  p <- fread(gwasres.snp, header =T)
+  p[,"PVALUE"] <- -log10(p[,"PVALUE"])
+  colnames(p)[colnames(p) == "MARKER_ID"] <-  "name"
+  dt.snp <- merge(a, p, by = "name", all.x = T)
+
+  fig1 <- plot_pip(dt.gene)
+  fig2 <- plot_beta(dt.gene)
+  fig3 <- plot_p(dt.gene)
+  fig4 <- plot_pip(dt.snp)
+  fig5 <- plot_beta(dt.snp)
+  fig6 <- plot_p(dt.snp)
+
+  fig <- subplot(fig1, fig2, fig3, fig4, fig5, fig6, nrows=6, shareX = T, titleY = T)
+
+  fig <- fig %>% layout(autosize = F, width = 600, height = 800, margin = 0.02,
+    xaxis = list(title = "Genomic position"))
+
+  fig <- fig %>% toWebGL()
+  fig
 }
@@
 MR.ASH results (start from gene):
-```{r sim 1 gene}
+```{r sim 1 gene, out.width = "98%", out.height="95%"}
 load(paste0(outputdir, "mr.ash2_20200503-1-mr.ash2.expr-res.Rd"))
@@
 summary_mr.ash2(g.fit, s.fit, phenores)
+
+tag <- "mr.ash2_20200503-1"
+plot_all(paste0(outputdir, tag, "-mr.ash2.expr-res.expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
+         paste0(outputdir, tag, "-mr.ash2.expr-res.snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
 ```
@@
 MR.ASH results (start from SNP):
-```{r sim 1 snp}
+Very similar to start from gene, therefore, not shown.
+```{r sim 1 snp, eval=F}
 load(paste0(outputdir, "mr.ash2_20200503-1-mr.ash2.snp-res.Rd"))
@@
 summary_mr.ash2(g.fit, s.fit, phenores)
+
+tag <- "mr.ash2_20200503-1"
+plot_all(paste0(outputdir, tag, "-mr.ash2.snp-res.expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
+         paste0(outputdir, tag, "-mr.ash2.snp-res.snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
 ```
@@
 True parameters:
-```{r para 2}
+```{r para 2, eval=T}
 load(paste0(outputdir, "mr.ash2_20200503-2-pheno.Rd"))
@@
 MR.ASH results (start from gene):
-```{r sim 2 gene}
+```{r sim 2 gene, eval=T, out.width = "98%", out.height="95%"}
 load(paste0(outputdir, "mr.ash2_20200503-2-mr.ash2.expr-res.Rd"))
@@
 summary_mr.ash2(g.fit, s.fit, phenores)
-```
 
-MR.ASH results (start from SNP):
-```{r sim 2 snp}
-load(paste0(outputdir, "mr.ash2_20200503-2-mr.ash2.snp-res.Rd"))
-g.fit <- mr.ash2.fit$fit2
-s.fit <-  mr.ash2.fit$fit1
-summary_mr.ash2(g.fit, s.fit, phenores)
+tag <- "mr.ash2_20200503-2"
+plot_all(paste0(outputdir, tag, "-mr.ash2.expr-res.expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
+         paste0(outputdir, tag, "-mr.ash2.expr-res.snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
 ```
@@
 True parameters:
-```{r para 3}
+```{r para 3, eval=T}
 load(paste0(outputdir, "mr.ash2_20200503-3-pheno.Rd"))
@@
 MR.ASH results (start from gene):
-```{r sim 3 gene}
+```{r sim 3 gene, eval=T, out.width = "98%", out.height="95%"}
 load(paste0(outputdir, "mr.ash2_20200503-3-mr.ash2.expr-res.Rd"))
@@
 summary_mr.ash2(g.fit, s.fit, phenores)
-```
 
-MR.ASH results (start from SNP):
-```{r sim 3 snp}
-load(paste0(outputdir, "mr.ash2_20200503-3-mr.ash2.snp-res.Rd"))
-g.fit <- mr.ash2.fit$fit2
-s.fit <-  mr.ash2.fit$fit1
-summary_mr.ash2(g.fit, s.fit, phenores)
+tag <- "mr.ash2_20200503-3"
+plot_all(paste0(outputdir, tag, "-mr.ash2.expr-res.expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
+         paste0(outputdir, tag, "-mr.ash2.expr-res.snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
 ```
@@
 True parameters:
-```{r para 4}
+```{r para 4, eval=T}
 load(paste0(outputdir, "mr.ash2_20200503-4-pheno.Rd"))
@@
 MR.ASH results (start from gene):
-```{r sim 4 gene}
+```{r sim 4 gene, eval=T, out.width = "98%", out.height="95%"}
 load(paste0(outputdir, "mr.ash2_20200503-4-mr.ash2.expr-res.Rd"))
@@
 summary_mr.ash2(g.fit, s.fit, phenores)
+
+tag <- "mr.ash2_20200503-4"
+plot_all(paste0(outputdir, tag, "-mr.ash2.expr-res.expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
+         paste0(outputdir, tag, "-mr.ash2.expr-res.snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
 ```
-
-MR.ASH results (start from SNP):
-```{r sim 4 snp}
-load(paste0(outputdir, "mr.ash2_20200503-4-mr.ash2.snp-res.Rd"))
-g.fit <- mr.ash2.fit$fit2
-s.fit <-  mr.ash2.fit$fit1
-summary_mr.ash2(g.fit, s.fit, phenores)
-```

### 2020-06-11 analysis/simulation-simpleveb-boost2-ukbchr22.Rmd
plotly
Former-commit-id: a3bcd8a2db83a1e3728cc6c1216da4d0a67c0617

@@
 
-plot_beta <- function(beta,beta.pm, ...){
-  plot( beta, pch=19, col ="darkgreen", ...)
-  points(beta.pm, pch =19, col = "red")
-  legend("topright", legend=c("true beta", "posterior mean"),
-       col=c("darkgreen", "red"), pch=19)
-}
-
-plot_pip <- function(indi, pip, pipcut = 0.1) {
-  x <- barplot(indi, main = "PIP")
-  points(x, pip, pch=19, col="red")
-  pred <- which(pip > pipcut)
-  real <- which(indi==1)
-  tp <- length(intersect(pred, real))/length(pred)
-  fp <- 1 - tp
-  fn <- length(setdiff(real, pred))/length(real)
-  cat("false positive rate: ", fp, "\n")
-  cat("false negative rate: ", fn, "\n")
-}
-
 summary_mr.ash2 <- function(g.fit, s.fit, phenores, pipcut = 0.1){
@@
   summary_mr.ash(g.fit)
-  plot_beta(e.b, g.fit$beta)
-  indi <- rep(0, length(g.fit$beta))
-  indi[phenores$param$idx.cgene] <- 1
-  pip <- get_pip(g.fit)
-  plot_pip(indi, pip, pipcut = pipcut)
 
@@
 MR.ASH2s results (start from gene):
-```{r sim 1 gene}
+```{r sim 1 gene, out.width = "98%", out.height="95%"}
 load(paste0(outputdir, "20200527-1-mr.ash2s.expr-res.Rd"))
@@
 summary_mr.ash2(g.fit, s.fit, phenores)
+
+tag <- "20200527-1"
+simtag <- "mr.ash2_20200503-1"
+plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
+         paste0(outputdir, tag, "-mr.ash2s.expr-res.snp.txt"), paste0(simdatadir, simtag, ".snpgwas.txt.gz"))
 ```
@@
 MR.ASH results (start from SNP):
-```{r sim 1 snp}
+Very similar to start from gene, therefore, not shown.
+```{r sim 1 snp, eval = F}
 load(paste0(outputdir, "20200527-1-mr.ash2s.snp-res.Rd"))
@@
 summary_mr.ash2(g.fit, s.fit, phenores)
+
+tag <- "20200527-1"
+simtag <- "mr.ash2_20200503-1"
+plot_all(paste0(outputdir, tag, "-mr.ash2s.snp-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
+         paste0(outputdir, tag, "-mr.ash2s.snp-res.snp.txt"), paste0(simdatadir, simtag, ".snpgwas.txt.gz"))
 ```
@@
 MR.ASH results (start from gene):
-```{r sim 2 gene}
+```{r sim 2 gene, out.width = "98%", out.height="95%"}
 load(paste0(outputdir, "20200527-2-mr.ash2s.expr-res.Rd"))
@@
 summary_mr.ash2(g.fit, s.fit, phenores)
-```
 
-MR.ASH results (start from SNP):
-```{r sim 2 snp}
-load(paste0(outputdir, "20200527-2-mr.ash2s.snp-res.Rd"))
-g.fit <- mr.ash2s.fit$fit2
-s.fit <-  mr.ash2s.fit$fit1
-summary_mr.ash2(g.fit, s.fit, phenores)
+tag <- "20200527-2"
+simtag <- "mr.ash2_20200503-2"
+plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
+         paste0(outputdir, tag, "-mr.ash2s.expr-res.snp.txt"), paste0(simdatadir, simtag, ".snpgwas.txt.gz"))
 ```
@@
 MR.ASH results (start from gene):
-```{r sim 3 gene}
+```{r sim 3 gene, out.width = "98%", out.height="95%"}
 load(paste0(outputdir, "20200527-3-mr.ash2s.expr-res.Rd"))
@@
 summary_mr.ash2(g.fit, s.fit, phenores)
-```
 
-MR.ASH results (start from SNP):
-```{r sim 3 snp}
-load(paste0(outputdir, "20200527-3-mr.ash2s.snp-res.Rd"))
-g.fit <- mr.ash2s.fit$fit2
-s.fit <-  mr.ash2s.fit$fit1
-summary_mr.ash2(g.fit, s.fit, phenores)
+tag <- "20200527-3"
+simtag <- "mr.ash2_20200503-3"
+plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
+         paste0(outputdir, tag, "-mr.ash2s.expr-res.snp.txt"), paste0(simdatadir, simtag, ".snpgwas.txt.gz"))
 ```
@@
 MR.ASH results (start from gene):
-```{r sim 4 gene}
+```{r sim 4 gene, out.width = "98%", out.height="95%"}
 load(paste0(outputdir, "20200527-4-mr.ash2s.expr-res.Rd"))
@@
 summary_mr.ash2(g.fit, s.fit, phenores)
+
+tag <- "20200527-4"
+simtag <- "mr.ash2_20200503-4"
+plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
+         paste0(outputdir, tag, "-mr.ash2s.expr-res.snp.txt"), paste0(simdatadir, simtag, ".snpgwas.txt.gz"))
 ```
-
-MR.ASH results (start from SNP):
-```{r sim 4 snp}
-load(paste0(outputdir, "20200527-4-mr.ash2s.snp-res.Rd"))
-g.fit <- mr.ash2s.fit$fit2
-s.fit <-  mr.ash2s.fit$fit1
-summary_mr.ash2(g.fit, s.fit, phenores)
-```

### 2020-06-11 code/gwas.R

@@ -1,0 +1,70 @@
+library(foreach)
+library(doParallel)
+
+
+#' @description this is a function to run GWAS analysis for small genotype matrix (using glm),for faster results use plink (start from plink files). Credit to http://www.stat-gen.org/tut/tut_analysis.html, requires tabix and bgzip if compress is true.
+#'
+#' @param geno genotype matrix, columns: snps, rows: individuals.
+#' @param pheno phenotype, matrix, ncol = 1
+#' @param snpname a vector of N, correpond to geno type matrix, will be written to output.
+#' @param anno a data frame/matrix, anno column with N rows.
+#' @param compress T/F, if T will be compressed and indexed by tabix. (TODO)
+#'
+GWAA <- function(geno, pheno, snpname = NULL, anno = NULL, outname, family = gaussian, ncore = 1, nSplits = 10, compress = F){
+
+  cl <- makeCluster(ncore)
+  show(cl)
+  registerDoParallel(cl)
+
+  colnames(pheno) <- "pheno"
+
+  if (is.null(snpname)){
+    snpname <- 1: ncol(geno)
+  }
+
+  nSNPs <- ncol(geno)
+  genosplit <- ceiling(nSNPs/nSplits) # number of SNPs in each subset
+
+  snp.start <- seq(1, nSNPs, genosplit) # index of first SNP in group
+  snp.stop <- pmin(snp.start+genosplit-1, nSNPs) # index of last SNP in group
+
+  columns<-c("#CHROM",	"BEGIN",	"END", "MARKER_ID", "Estimate", "Std.Error", "t-value", "PVALUE") # header for LOCUSZOOM.
+  write.table(t(columns), outname, row.names=FALSE, col.names=FALSE, quote=FALSE, sep = "\t")
+
+  foreach (part=1:nSplits) %do% {
+
+    geno.i <- geno[ ,snp.start[part]:snp.stop[part]]
+    snpname.i <- snpname[snp.start[part]:snp.stop[part]]
+    if (!is.null(anno)) {
+      anno.i <- anno[snp.start[part]:snp.stop[part], ]
+    } else {
+      anno.i <- NULL
+    }
+
+    res <- foreach(snp.idx=1: length(snpname.i), .combine='rbind') %dopar% {
+      snp <- geno.i[, snp.idx, drop = F]
+      a <- summary(glm(pheno ~ snp,
+                    family=family,
[…]

### 2020-06-11 code/run_test_susie.R

@@ -1,19 +1,19 @@
-library(mr.ash.alpha)
+library(susieR)
 
 args = commandArgs(trailingOnly=TRUE)
-if (length(args) != 4) {
+if (length(args) != 6) {
   stop("4 arguments must be supplied:
        * genotype file name
        * expr Rd
        * pheno Rd
        * prior.R (will be sourced)
+       * region file (optional, chrnumber<tab>start<tab>end)
        * out file name", call.=FALSE)
 }
 
 codedir <- "/project2/mstephens/causalTWAS/causal-TWAS/code/"
 source(paste0(codedir, "stats_func.R"))
 source(paste0(codedir,"input_reformat.R"))
-source(paste0(codedir,"mr.ash2.R"))
 
 # genotype plink to R data file
 pfile <- args[1]
@@ -40,13 +40,22 @@ load(args[3])
 
 source(args[4]) # priors
 
-outname <- args[5]
+outname <- args[6]
 
-prior.SNP <- 1E-3
-prior.gene <- 0.1
-prior <- c(rep(prior.SNP, p-1), prior.gene)
-res <- susie(X,y,L=1, prior_weights=prior)
+regions <- read.table(args[5], stringsAsFactors = F)
 
+for (i in 1:nrow(regions)){
+  chr <- regions[i, 1]
+  p0 <- regions[i, 2]
+  p1 <- regions[i, 3]
+  X.gene <- exprres$expr[ , (exprres$chrom == chr & (exprres$p0 > p0 | exprres$p1 < p1))]
+  X.SNP <-  dat$G[ , (dat$chr == chr & dat$pos > p0 & dat$pos < p1)]
+
+  prior <- c(rep(prior.gene, dim(X.gene)[2]), rep(prior.SNP, dim(X.SNP)[2]))
+
+  res <- susie(cbind(X.gene, X.SNP), phenores$Y, L=1, prior_weights = prior)
+  save(res, file = paste(outname,chr,p0,p1,"susieres.Rd", sep = "-"))
+}
 
 
 

### 2020-06-11 code/temp.R

@@ -1,39 +1,55 @@
-# simulate data with two groups of effects
-# Two groups of variables: k1 and k2 non-zero effects with effect s.d. sigma1/sigma2.
-# Group 1: SNP effect, Group 2: gene effect.
-# Variables in group 2 are in “complete LD” with some of the varaibles in group 1.
-n <- 3000
-p <- 1000
-p2 <- 100
-p1 <- p - p2
-X <- matrix(rnorm(n*p), nrow=n, ncol=p)
-#X2 <- matrix(rnorm(n*p2), nrow=n, ncol=p2)
-#X1 <- cbind(matrix(rnorm(n*(p1-p2)), nrow=n, ncol=p1-p2), X2)
-#X <- cbind(X1, X2)
-k1 <- 20
-sigma1 <- 0.1
-k2 <- 20
-sigma2 <- 0.2
-beta1 <- c(rnorm(k1, 0, sigma1), rep(0, p1-k1))
-beta2 <- c(rnorm(k2, 0, sigma2), rep(0, p2-k2))
-beta <- c(beta1, beta2)
-sigma.e <- 1
-y <- X %*% beta + rnorm(n, 0, sigma.e)
-# fit a single group
-fit <- mr.ash(X, y, method="caisa")
-beta.single <- fit$beta
-which(abs(beta) > 0.05)
-which(abs(beta.single) > 0.05)
-# iteratively fit two groups using mr.ash
-X1 <- X[, 1:p1]
-X2 <- X[, (p1+1):p]
-r1 <- y
-r2 <- y
-fit2 <- mr.ash(X2, r2, method = "caisa")
-beta2.pm <- fit2$beta
-r1 <- y - X2 %*% beta2.pm
-fit1 <- mr.ash(X1, r1, method="caisa")
-beta1.pm <- fit1$beta
-r2 <- y - X1 %*% beta1.pm
-beta.pm <- c(beta1.pm, beta2.pm)
-which(abs(beta.pm) > 0.05)
+weight <- "/home/simingz/causalTWAS/fusion_weights/YFS.BLOOD.RNAARR"
+wgtfiles <- Sys.glob(file.path(weight, "*.wgt.RDat"))
+wgtposfile <- file.path(dirname(weight), paste0(basename(weight), ".pos"))
+wgtpos <- read.table(wgtposfile, header = T, stringsAsFactors = F)
+
[…]

### 2020-06-11 analysis/simulation-susie-ukbchr22.Rmd
SUSIE

---
title: "Test SUSIE for regions near genes (500kb flanking)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r plot functions}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})

plot_pip <- function(dt){
  pal <- c("salmon", "darkgreen")
  pal <- setNames(pal, c("Causal", "Non causal"))

  fig <- plot_ly(dt, x = ~ pos, y = ~ pip, type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))

  fig <- fig %>% layout(yaxis = list(title = "PIP"))
  return(fig)
}

plot_all <- function(susieres, mrashres.gene, mrashres.snp){
  dt <- read.table(susieres , header =T)

  a <- read.table(mrashres.gene , header =T)
  a[a$ifcausal == 1, "ifcausal"] = "Causal"
  a[a$ifcausal == 0, "ifcausal"] = "Non causal"

  dt.gene <- merge(dt, a, by= "name")

  a <- read.table(mrashres.snp , header =T)
  a[a$ifcausal == 1, "ifcausal"] = "Causal"
  a[a$ifcausal == 0, "ifcausal"] = "Non causal"

  dt.snp <- merge(dt, a, by= "name")

  fig1 <- plot_pip(dt.gene)
  fig2 <- plot_pip(dt.snp)

  fig <- subplot(fig1, fig2, nrows= 2, shareX = T, titleY = T)

  fig <- fig %>% layout(autosize = F, width = 600, height = 600, margin = 0.02,
    xaxis = list(title = "Genomic position"))

  fig <- fig %>% toWebGL()
  fig
}

```

```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200503/"
outputdir <- "~/causalTWAS/simulations/simulation_susietest_20200604/"
mr.ash2dir <-  "~/causalTWAS/simulations/simulation_ashtest_20200527/"
```

## Simulation 1
Priors:
```{r para 1}
readLines(paste0(outputdir, "prior-1.R"))
```

SUSIE results:
```{r sim 1 setup}
mrashtag <- "20200527-1"
susietag <- "20200604-1"
mrashres.gene <- paste0(mr.ash2dir, mrashtag, "-mr.ash2s.expr-res.expr.txt")
mrashres.snp <- paste0(mr.ash2dir, mrashtag, "-mr.ash2s.expr-res.snp.txt")
```

### FALSE positives in mr.ash2

```{r sim 1 3, out.width = "98%", out.height="95%"}
name <- "DGCR14"
susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
print(paste0("SUISE results for regions around ", name))
plot_all(susieres, mrashres.gene, mrashres.snp)
```

```{r sim 1 4, out.width = "98%", out.height="95%"}
name <- "VPREB3"
susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
print(paste0("SUISE results for regions around ", name))
plot_all(susieres, mrashres.gene, mrashres.snp)
```

```{r sim 1 5, out.width = "98%", out.height="95%"}
name <- "TCN2"
susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
print(paste0("SUISE results for regions around ", name))
plot_all(susieres, mrashres.gene, mrashres.snp)
```

### FALSE negatives in mr.ash2
```{r sim 1 6, out.width = "98%", out.height="95%"}
name <- "CBX6"
susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
print(paste0("SUISE results for regions around ", name))
plot_all(susieres, mrashres.gene, mrashres.snp)
```

```{r sim 1 7, out.width = "98%", out.height="95%"}
name <- "CHEK2"
susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
print(paste0("SUISE results for regions around ", name))
plot_all(susieres, mrashres.gene, mrashres.snp)
```

### True positives in mr.ash2
```{r sim 1 8, out.width = "98%", out.height="95%"}
name <- "LMF2"
susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
print(paste0("SUISE results for regions around ", name))
plot_all(susieres, mrashres.gene, mrashres.snp)
```

```{r sim 1 9, out.width = "98%", out.height="95%"}
name <- "DDX17"
susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
print(paste0("SUISE results for regions around ", name))
plot_all(susieres, mrashres.gene, mrashres.snp)
```
```{r sim 1 10, out.width = "98%", out.height="95%"}
name <- "MIF"
susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
print(paste0("SUISE results for regions around ", name))
plot_all(susieres, mrashres.gene, mrashres.snp)
```

### 2020-06-11 code/run_test_susie.R

@@ -7,7 +7,7 @@ if (length(args) != 6) {
        * expr Rd
        * pheno Rd
        * prior.R (will be sourced)
-       * region file (optional, chrnumber<tab>start<tab>end)
+       * region file (optional, chrnumber<tab>start<tab>end<name>)
        * out file name", call.=FALSE)
 }
 
@@ -42,21 +42,38 @@ source(args[4]) # priors
 
 outname <- args[6]
 
-regions <- read.table(args[5], stringsAsFactors = F)
+regions <- read.table(args[5], stringsAsFactors = F, header =T)
 
 for (i in 1:nrow(regions)){
   chr <- regions[i, 1]
   p0 <- regions[i, 2]
   p1 <- regions[i, 3]
-  X.gene <- exprres$expr[ , (exprres$chrom == chr & (exprres$p0 > p0 | exprres$p1 < p1))]
-  X.SNP <-  dat$G[ , (dat$chr == chr & dat$pos > p0 & dat$pos < p1)]
+  name <- regions[i, 4]
+  
+  susieres <- list()
+  print(name)
+  
+  idx.gene <- exprres$chrom == chr & exprres$p0 > p0 & exprres$p1 < p1
+  idx.SNP <- dat$chr == chr & dat$pos > p0 & dat$pos < p1
+  
+  X.gene <- exprres$expr[ , idx.gene, drop = F]
+  X.SNP <-  dat$G[ , idx.SNP, drop = F]
 
   prior <- c(rep(prior.gene, dim(X.gene)[2]), rep(prior.SNP, dim(X.SNP)[2]))
 
-  res <- susie(cbind(X.gene, X.SNP), phenores$Y, L=1, prior_weights = prior)
-  save(res, file = paste(outname,chr,p0,p1,"susieres.Rd", sep = "-"))
+  susieres[["susie"]] <- susie(cbind(X.gene, X.SNP), phenores$Y, L=1, prior_weights = prior)
+  
+  anno.gene <- cbind(colnames(X.gene), exprres$chrom[idx.gene],  exprres$p0[idx.gene])
+  anno.SNP <- cbind(dat$snp[idx.SNP,], dat$chr[idx.SNP,], dat$pos[idx.SNP,])
+  susieres[["anno"]] <- rbind(anno.gene, anno.SNP)
+  
+  save(susieres, file = paste(outname, name, "susieres.Rd", sep = "-"))
+  
[…]

### 2020-06-15 analysis/simulation-susie-ukbchr22.Rmd
susie more genes

@@
 ```
+
+## Simulation 3
+Priors:
+```{r para 3}
+readLines(paste0(outputdir, "prior-3.R"))
+```
+
+SUSIE results:
+```{r sim 3 setup}
+mrashtag <- "20200527-3"
+susietag <- "20200604-3"
+mrashres.gene <- paste0(mr.ash2dir, mrashtag, "-mr.ash2s.expr-res.expr.txt")
+mrashres.snp <- paste0(mr.ash2dir, mrashtag, "-mr.ash2s.expr-res.snp.txt")
+```
+```{r sim 3 1, out.width = "98%", out.height="95%"}
+
+name <- "MCM5"
+susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
+print(paste0("SUISE results for regions around ", name))
+plot_all(susieres, mrashres.gene, mrashres.snp)
+```
+
+```{r sim 3 2, out.width = "98%", out.height="95%"}
+
+name <- "CBX6"
+susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
+print(paste0("SUISE results for regions around ", name))
+plot_all(susieres, mrashres.gene, mrashres.snp)
+```

### 2020-06-18 analysis/index.Rmd
SUSIE
logging and more weights for chr22

@@
 
-  * [Using UKbiobank data (chr22)](https://users.rcc.uchicago.edu/~simingz/simulation-simpleveb-boost2-ukbchr22.html)
+  * [Using UKbiobank data (chr22), fusion weights YFS ](https://users.rcc.uchicago.edu/~simingz/simulation-simpleveb-boost2-ukbchr22.html)
+
+  * [Using UKbiobank data (chr22), fusion weights Adipose subcutaneous](https://users.rcc.uchicago.edu/~simingz/simulation-mr.ash2s-ukbchr22-gtex.adipose.html)
+
+  * [Using UKbiobank data (chr22), weights of H3K27 from Xiaotong](https://users.rcc.uchicago.edu/~simingz/simulation-mr.ash2s-ukbchr22-H3K27.html)
 
@@
 
-  * [Using some simple data](simulation_susie.html)
-
-  * [Using UKbiobank data (chr22)](simulation-susie-ukbchr22.html)
+  * [Using UKbiobank data (chr22), fusion weights YFS](simulation-susie-ukbchr22.html)
 

### 2020-06-18 analysis/simulation-mr.ash2s-ukbchr22-H3K27.Rmd
logging and more weights for chr22

--- 2020-06-11 analysis/simulation-simpleveb-boost2-ukbchr22.Rmd
@@
 ---
-title: "Test a simpler version of veb-boost performance on ukbiobank data-2"
+title: "Test a simpler version of veb-boost performance mr.ash2s on ukbiobank data"
 output: html_document
@@
 ```
+
+We test the performance of mr.ash2s on ukbiobank data (chr22, n sample = 20,000). 
+We use weights from Xiaotong, H3K27ac (trained using FUSION). Total features: > 8,000. on chromosome 22: 296.
 
@@
 ```{r output dir}
-simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200503/"
-outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200527/"
+simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200615/"
+outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200615/"
 ```
@@
 ```{r para 1}
-load(paste0(simdatadir, "mr.ash2_20200503-1-pheno.Rd"))
-readLines(paste0(simdatadir, "param-20200503-1.R"))
+load(paste0(simdatadir, "simu_20200615-1-pheno.Rd"))
+readLines(paste0(simdatadir, "param-20200615-1.R"))
 ```
@@
 ```{r sim 1 gene, out.width = "98%", out.height="95%"}
-load(paste0(outputdir, "20200527-1-mr.ash2s.expr-res.Rd"))
+load(paste0(outputdir, "20200615-1-mr.ash2s.expr-res.Rd"))
 g.fit <- mr.ash2s.fit$fit1
@@
 
-tag <- "20200527-1"
-simtag <- "mr.ash2_20200503-1"
+tag <- "20200615-1"
+simtag <- "20200615-1"
 plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
@@
 ```{r sim 1 snp, eval = F}
-load(paste0(outputdir, "20200527-1-mr.ash2s.snp-res.Rd"))
+load(paste0(outputdir, "20200615-1-mr.ash2s.snp-res.Rd"))
 g.fit <- mr.ash2s.fit$fit2
@@
 
-tag <- "20200527-1"
-simtag <- "mr.ash2_20200503-1"
+tag <- "20200615-1"
+simtag <- "20200615-1"
 plot_all(paste0(outputdir, tag, "-mr.ash2s.snp-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
@@
 ```{r para 2}
-load(paste0(simdatadir, "mr.ash2_20200503-2-pheno.Rd"))
-readLines(paste0(simdatadir, "param-20200503-2.R"))
+load(paste0(simdatadir, "simu_20200615-2-pheno.Rd"))
+readLines(paste0(simdatadir, "param-20200615-2.R"))
 ```
@@
 ```{r sim 2 gene, out.width = "98%", out.height="95%"}
-load(paste0(outputdir, "20200527-2-mr.ash2s.expr-res.Rd"))
+load(paste0(outputdir, "20200615-2-mr.ash2s.expr-res.Rd"))
 g.fit <- mr.ash2s.fit$fit1
@@
 
-tag <- "20200527-2"
-simtag <- "mr.ash2_20200503-2"
+tag <- "20200615-2"
+simtag <- "20200615-2"
 plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
@@
 ```{r para 3}
-load(paste0(simdatadir, "mr.ash2_20200503-3-pheno.Rd"))
-readLines(paste0(simdatadir, "param-20200503-3.R"))
+load(paste0(simdatadir, "simu_20200615-3-pheno.Rd"))
+readLines(paste0(simdatadir, "param-20200615-3.R"))
 ```
@@
 ```{r sim 3 gene, out.width = "98%", out.height="95%"}
-load(paste0(outputdir, "20200527-3-mr.ash2s.expr-res.Rd"))
+load(paste0(outputdir, "20200615-3-mr.ash2s.expr-res.Rd"))
 g.fit <- mr.ash2s.fit$fit1
@@
 
-tag <- "20200527-3"
-simtag <- "mr.ash2_20200503-3"
+tag <- "20200615-3"
+simtag <- "20200615-3"
 plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
@@
 ```{r para 4}
-load(paste0(simdatadir, "mr.ash2_20200503-4-pheno.Rd"))
-readLines(paste0(simdatadir, "param-20200503-4.R"))
+load(paste0(simdatadir, "simu_20200615-4-pheno.Rd"))
+readLines(paste0(simdatadir, "param-20200615-4.R"))
 ```
@@
 ```{r sim 4 gene, out.width = "98%", out.height="95%"}
-load(paste0(outputdir, "20200527-4-mr.ash2s.expr-res.Rd"))
+load(paste0(outputdir, "20200615-4-mr.ash2s.expr-res.Rd"))
 g.fit <- mr.ash2s.fit$fit1
@@
 
-tag <- "20200527-4"
-simtag <- "mr.ash2_20200503-4"
+tag <- "20200615-4"
+simtag <- "20200615-4"
 plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 

### 2020-06-18 code/simulate_phenotype.R

@@ -13,16 +13,24 @@ library(data.table)
 #' @return a matrix, N x J. Missing genotypes will not contribute to cis-expr and give warnings.
 cis_expr <- function(dat, weight, method = "bslmm", checksnps = F){
   exprlist <- list()
-  wgtfiles <- Sys.glob(file.path(weight, "*.wgt.RDat"))
-  wgtposfile <- file.path(dirname(weight), paste0(basename(weight), ".pos"))
+  wgtdir <- dirname(weight)
+  wgtposfile <- file.path(wgtdir, paste0(basename(weight), ".pos"))
 
   wgtpos <- read.table(wgtposfile, header = T, stringsAsFactors = F)
+  
+  wgtpos <- transform(wgtpos, 
+            "ID" = ifelse(duplicated(ID) | duplicated(ID, fromLast=TRUE), 
+                                 paste(ID, ave(ID, ID, FUN=seq_along), sep='_ID'), 
+                                 ID))
+  
+  write.table(wgtpos , file= paste0(wgtposfile, ".IDfixed") , row.names=F, col.names=T, sep="\t", quote = F)
+  
+  print(paste0("Number of genes with weights provided: ", dim(wgtpos)[1]))
 
-  print(paste0("Number of genes with weights provided: ", length(wgtfiles)))
-
-  for (wf in wgtfiles){
+  for (i in 1: dim(wgtpos)[1]){
+    wf <- file.path(wgtdir, wgtpos[i, "WGT"])
     load(wf)
-    gname <- strsplit(basename(wf), "." , fixed = T)[[1]][2]
+    gname <- wgtpos[i, "ID"]
     print(gname)
 
     g.method = method
@@ -33,10 +41,10 @@ cis_expr <- function(dat, weight, method = "bslmm", checksnps = F){
     wgt.idx <- match(rownames(wgt.matrix), dat$snp)
 
     if (checksnps){
-      datsnps <- data.frame("V1" = dat$chr,
-                            "V4" = dat$pos,
-                            "V5" = dat$counted,
-                            "V6" = dat$alt)
+      datsnps <- data.frame("V1" = dat$chr[wgt.idx, 1],
+                            "V4" = dat$pos[wgt.idx, 1],
+                            "V5" = dat$counted[wgt.idx, 1],
+                            "V6" = dat$alt[wgt.idx, 1])
 
[…]

### 2020-06-18 analysis/simulation-mr.ash2s-ukbchr22-gtex.adipose.Rmd
logging and more weights for chr22

--- 2020-06-18 analysis/simulation-mr.ash2s-ukbchr22-H3K27.Rmd
@@
 We test the performance of mr.ash2s on ukbiobank data (chr22, n sample = 20,000). 
-We use weights from Xiaotong, H3K27ac (trained using FUSION). Total features: > 8,000. on chromosome 22: 296.
+We use weights from Fusion website, using [Adipose (subcutaneous) p < 0.01](http://gusevlab.org/projects/fusion/weights/GTEx.Adipose_Subcutaneous.P01.tar.bz2).
+Total features > 8,000. On chr22: 240.
 
@@
 ```{r output dir}
-simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200615/"
-outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200615/"
+simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
+outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
 ```
@@
 ```{r para 1}
-load(paste0(simdatadir, "simu_20200615-1-pheno.Rd"))
-readLines(paste0(simdatadir, "param-20200615-1.R"))
+load(paste0(simdatadir, "simu_20200616-1-pheno.Rd"))
+readLines(paste0(simdatadir, "param-20200616-1.R"))
 ```
@@
 ```{r sim 1 gene, out.width = "98%", out.height="95%"}
-load(paste0(outputdir, "20200615-1-mr.ash2s.expr-res.Rd"))
+load(paste0(outputdir, "20200616-1-mr.ash2s.expr-res.Rd"))
 g.fit <- mr.ash2s.fit$fit1
@@
 
-tag <- "20200615-1"
-simtag <- "20200615-1"
+tag <- "20200616-1"
+simtag <- "20200616-1"
 plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
@@
 ```{r sim 1 snp, eval = F}
-load(paste0(outputdir, "20200615-1-mr.ash2s.snp-res.Rd"))
+load(paste0(outputdir, "20200616-1-mr.ash2s.snp-res.Rd"))
 g.fit <- mr.ash2s.fit$fit2
@@
 
-tag <- "20200615-1"
-simtag <- "20200615-1"
+tag <- "20200616-1"
+simtag <- "20200616-1"
 plot_all(paste0(outputdir, tag, "-mr.ash2s.snp-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
@@
 ```{r para 2}
-load(paste0(simdatadir, "simu_20200615-2-pheno.Rd"))
-readLines(paste0(simdatadir, "param-20200615-2.R"))
+load(paste0(simdatadir, "simu_20200616-2-pheno.Rd"))
+readLines(paste0(simdatadir, "param-20200616-2.R"))
 ```
@@
 ```{r sim 2 gene, out.width = "98%", out.height="95%"}
-load(paste0(outputdir, "20200615-2-mr.ash2s.expr-res.Rd"))
+load(paste0(outputdir, "20200616-2-mr.ash2s.expr-res.Rd"))
 g.fit <- mr.ash2s.fit$fit1
@@
 
-tag <- "20200615-2"
-simtag <- "20200615-2"
+tag <- "20200616-2"
+simtag <- "20200616-2"
 plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
@@
 ```{r para 3}
-load(paste0(simdatadir, "simu_20200615-3-pheno.Rd"))
-readLines(paste0(simdatadir, "param-20200615-3.R"))
+load(paste0(simdatadir, "simu_20200616-3-pheno.Rd"))
+readLines(paste0(simdatadir, "param-20200616-3.R"))
 ```
@@
 ```{r sim 3 gene, out.width = "98%", out.height="95%"}
-load(paste0(outputdir, "20200615-3-mr.ash2s.expr-res.Rd"))
+load(paste0(outputdir, "20200616-3-mr.ash2s.expr-res.Rd"))
 g.fit <- mr.ash2s.fit$fit1
@@
 
-tag <- "20200615-3"
-simtag <- "20200615-3"
+tag <- "20200616-3"
+simtag <- "20200616-3"
 plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
@@
 ```{r para 4}
-load(paste0(simdatadir, "simu_20200615-4-pheno.Rd"))
-readLines(paste0(simdatadir, "param-20200615-4.R"))
+load(paste0(simdatadir, "simu_20200616-4-pheno.Rd"))
+readLines(paste0(simdatadir, "param-20200616-4.R"))
 ```
@@
 ```{r sim 4 gene, out.width = "98%", out.height="95%"}
-load(paste0(outputdir, "20200615-4-mr.ash2s.expr-res.Rd"))
+load(paste0(outputdir, "20200616-4-mr.ash2s.expr-res.Rd"))
 g.fit <- mr.ash2s.fit$fit1
@@
 
-tag <- "20200615-4"
-simtag <- "20200615-4"
+tag <- "20200616-4"
+simtag <- "20200616-4"
 plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 

### 2020-06-21 analysis/simulation-susie-ukbchr22-gtex.adipose.Rmd
reduce mr.ash memory/mr.ashs

--- 2020-06-11 analysis/simulation-susie-ukbchr22.Rmd
@@
 ```{r output dir}
-simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200503/"
-outputdir <- "~/causalTWAS/simulations/simulation_susietest_20200604/"
-mr.ash2dir <-  "~/causalTWAS/simulations/simulation_ashtest_20200527/"
+simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
+outputdir <- "~/causalTWAS/simulations/simulation_susietest_20200616/"
+mr.ash2dir <-  "~/causalTWAS/simulations/simulation_ashtest_20200616/"
 ```
@@
 ```{r para 1}
-readLines(paste0(outputdir, "prior-1.R"))
+mrashtag <- "20200616-1"
+susietag <- "20200616-1"
+readLines(paste0(outputdir, "prior-", susietag, ".R"))
 ```
 
-SUSIE results:
 ```{r sim 1 setup}
-mrashtag <- "20200527-1"
-susietag <- "20200604-1"
 mrashres.gene <- paste0(mr.ash2dir, mrashtag, "-mr.ash2s.expr-res.expr.txt")
@@
 
-### FALSE positives in mr.ash2
+### SUSIE results: PIP > 0.5 or causal genes
 
-```{r sim 1 3, out.width = "98%", out.height="95%"}
-name <- "DGCR14"
+```{r sim 1 examples, out.width = "98%", out.height="95%"}
+nameall <- read.table(paste0(outputdir, "gene-flank500kb-", susietag, ".txt"))[ , 4]
+for (name in nameall){
   susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
@@
   plot_all(susieres, mrashres.gene, mrashres.snp)
+}
 ```
-
-```{r sim 1 4, out.width = "98%", out.height="95%"}
-name <- "VPREB3"
-susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
-print(paste0("SUISE results for regions around ", name))
-plot_all(susieres, mrashres.gene, mrashres.snp)
-```
-
-```{r sim 1 5, out.width = "98%", out.height="95%"}
-name <- "TCN2"
-susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
-print(paste0("SUISE results for regions around ", name))
-plot_all(susieres, mrashres.gene, mrashres.snp)
-```
-
-### FALSE negatives in mr.ash2
-```{r sim 1 6, out.width = "98%", out.height="95%"}
-name <- "CBX6"
-susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
-print(paste0("SUISE results for regions around ", name))
-plot_all(susieres, mrashres.gene, mrashres.snp)
-```
-
-```{r sim 1 7, out.width = "98%", out.height="95%"}
-name <- "CHEK2"
-susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
-print(paste0("SUISE results for regions around ", name))
-plot_all(susieres, mrashres.gene, mrashres.snp)
-```
-
-### True positives in mr.ash2
-```{r sim 1 8, out.width = "98%", out.height="95%"}
-name <- "LMF2"
-susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
-print(paste0("SUISE results for regions around ", name))
-plot_all(susieres, mrashres.gene, mrashres.snp)
-```
-
-```{r sim 1 9, out.width = "98%", out.height="95%"}
-name <- "DDX17"
-susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
-print(paste0("SUISE results for regions around ", name))
-plot_all(susieres, mrashres.gene, mrashres.snp)
-```
-```{r sim 1 10, out.width = "98%", out.height="95%"}
-name <- "MIF"
-susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
-print(paste0("SUISE results for regions around ", name))
-plot_all(susieres, mrashres.gene, mrashres.snp)
-```

### 2020-06-21 code/run_simulate_data.R

@@ -37,20 +37,22 @@ if (!file.exists(pfileRd)) {
 load(pfileRd)
 loginfo("genotype scaling ...")
 dat$G <- scaleRcpp(dat$G)
+gc()
 
 # simulate phenotype
+outname <- args[4]
+loginfo("generating imputed expression file: %s",  paste0(outname, "-cis-expr.Rd"))
+
 weight <- args[2]
 exprres <- cis_expr(dat, weight, method = "best", checksnps = F)
-
-outname <- args[4]
 save(exprres, file = paste0(outname, "-cis-expr.Rd"))
-loginfo("writing imputed expression file: %s",  paste0(outname, "-cis-expr.Rd"))
 
 # load(paste0(outname, "-cis-expr.Rd"))
 loginfo("expr scaling ...")
 dat$expr <-  scaleRcpp(exprres$expr)
 
 source(args[3])
+loginfo("generating phenotype file: %s", paste0(outname, "-pheno.Rd"))
 phenores <- simulate_phenotype(dat, mode = "snp-expr",
                                 pve.expr = pve.expr,
                                 pve.snp = pve.snp,
@@ -60,5 +62,6 @@ phenores <- simulate_phenotype(dat, mode = "snp-expr",
 
 # write output
 save(phenores, file = paste0(outname, "-pheno.Rd"))
-loginfo("writing phenotype file: %s", paste0(outname, "-pheno.Rd"))
-logwarn(warnings())
+
+logwarn(str(summary(warnings())))
+loginfo("simulation done.")

### 2020-06-22 analysis/simulation-mr.ash2s-ukbchr22-gtex.adipose.Rmd
fix sa2 bug; locus zoom

@@
 ## Simulation 1
-True parameters:
-```{r para 1}
-load(paste0(simdatadir, "simu_20200616-1-pheno.Rd"))
-readLines(paste0(simdatadir, "param-20200616-1.R"))
-```
-
 MR.ASH2s results (start from gene):
 ```{r sim 1 gene, out.width = "98%", out.height="95%"}
-load(paste0(outputdir, "20200616-1-mr.ash2s.expr-res.Rd"))
-g.fit <- mr.ash2s.fit$fit1
-s.fit <-  mr.ash2s.fit$fit2
-summary_mr.ash2(g.fit, s.fit, phenores)
+tag <- "20200616-1"
+tag2 <- "expr-res"
 
-tag <- "20200616-1"
-simtag <- "20200616-1"
-plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
-         paste0(outputdir, tag, "-mr.ash2s.expr-res.snp.txt"), paste0(simdatadir, simtag, ".snpgwas.txt.gz"))
+param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+print(param)
 ```
@@
 MR.ASH results (start from SNP):
-Very similar to start from gene, therefore, not shown.
 ```{r sim 1 snp, eval = F}
-load(paste0(outputdir, "20200616-1-mr.ash2s.snp-res.Rd"))
-g.fit <- mr.ash2s.fit$fit2
-s.fit <-  mr.ash2s.fit$fit1
-summary_mr.ash2(g.fit, s.fit, phenores)
+tag <- "20200616-1"
+tag2 <- "snp-res"
 
-tag <- "20200616-1"
-simtag <- "20200616-1"
-plot_all(paste0(outputdir, tag, "-mr.ash2s.snp-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
-         paste0(outputdir, tag, "-mr.ash2s.snp-res.snp.txt"), paste0(simdatadir, simtag, ".snpgwas.txt.gz"))
+param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+print(param)
 ```
@@
 
-True parameters:
-```{r para 2}
-load(paste0(simdatadir, "simu_20200616-2-pheno.Rd"))
-readLines(paste0(simdatadir, "param-20200616-2.R"))
-```
-
 MR.ASH results (start from gene):
 ```{r sim 2 gene, out.width = "98%", out.height="95%"}
-load(paste0(outputdir, "20200616-2-mr.ash2s.expr-res.Rd"))
-g.fit <- mr.ash2s.fit$fit1
-s.fit <-  mr.ash2s.fit$fit2
-summary_mr.ash2(g.fit, s.fit, phenores)
+tag <- "20200616-2"
+tag2 <- "expr-res"
 
+param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+print(param)
+```
+MR.ASH results (start from snp):
+```{r sim 2 snp, out.width = "98%", out.height="95%"}
 tag <- "20200616-2"
-simtag <- "20200616-2"
-plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
-         paste0(outputdir, tag, "-mr.ash2s.expr-res.snp.txt"), paste0(simdatadir, simtag, ".snpgwas.txt.gz"))
+tag2 <- "snp-res"
+
+param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+print(param)
 ```
@@
 
-True parameters:
-```{r para 3}
-load(paste0(simdatadir, "simu_20200616-3-pheno.Rd"))
-readLines(paste0(simdatadir, "param-20200616-3.R"))
-```
-
 MR.ASH results (start from gene):
 ```{r sim 3 gene, out.width = "98%", out.height="95%"}
-load(paste0(outputdir, "20200616-3-mr.ash2s.expr-res.Rd"))
-g.fit <- mr.ash2s.fit$fit1
-s.fit <-  mr.ash2s.fit$fit2
-summary_mr.ash2(g.fit, s.fit, phenores)
+tag <- "20200616-3"
+tag2 <- "expr-res"
 
+param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+print(param)
+```
+MR.ASH results (start from snp):
+```{r sim 3 snp, out.width = "98%", out.height="95%"}
 tag <- "20200616-3"
-simtag <- "20200616-3"
-plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
-         paste0(outputdir, tag, "-mr.ash2s.expr-res.snp.txt"), paste0(simdatadir, simtag, ".snpgwas.txt.gz"))
+tag2 <- "snp-res"
+
+param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+print(param)
 ```
@@
 
-True parameters:
-```{r para 4}
-load(paste0(simdatadir, "simu_20200616-4-pheno.Rd"))
-readLines(paste0(simdatadir, "param-20200616-4.R"))
+MR.ASH results (start from gene):
+```{r sim 4 gene, out.width = "98%", out.height="95%"}
+tag <- "20200616-4"
+tag2 <- "expr-res"
+
+param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+print(param)
 ```
 
+MR.ASH results (start from snp):
+```{r sim 4 snp, out.width = "98%", out.height="95%"}
+tag <- "20200616-4"
+tag2 <- "snp-res"
+
+param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+print(param)
+```
+
+## Simulation 5
+
 MR.ASH results (start from gene):
-```{r sim 4 gene, out.width = "98%", out.height="95%"}
-load(paste0(outputdir, "20200616-4-mr.ash2s.expr-res.Rd"))
-g.fit <- mr.ash2s.fit$fit1
-s.fit <-  mr.ash2s.fit$fit2
-summary_mr.ash2(g.fit, s.fit, phenores)
+```{r sim 5 gene, out.width = "98%", out.height="95%"}
+tag <- "20200616-5"
+tag2 <- "expr-res"
 
-tag <- "20200616-4"
-simtag <- "20200616-4"
-plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
-         paste0(outputdir, tag, "-mr.ash2s.expr-res.snp.txt"), paste0(simdatadir, simtag, ".snpgwas.txt.gz"))
+param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+print(param)
+
+plot_all(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
+         paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
 ```
+
+MR.ASH results (start from SNP):
+```{r sim 5 snp, out.width = "98%", out.height="95%"}
+tag <- "20200616-5"
+tag2 <- "snp-res"
+
+param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+print(param)
+
+plot_all(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
+         paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
+```
+
+## Simulation 6
+
+MR.ASH results (start from gene):
+```{r sim 6 gene, out.width = "98%", out.height="95%"}
+tag <- "20200616-6"
+tag2 <- "expr-res"
+
+param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+print(param)
+
+plot_all(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
+         paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
+```
+
+MR.ASH results (start from SNP):
+```{r sim 6 snp, out.width = "98%", out.height="95%"}
+tag <- "20200616-6"
+tag2 <- "snp-res"
+
+param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+print(param)
+
+plot_all(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
+         paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
+```

### 2020-06-22 analysis/simulation-simpleveb-boost2-ukbchr22.Rmd
fix sa2 bug; locus zoom

@@
 ```
+
+## Simulation 5
+
+MR.ASH results (start from gene):
+```{r sim 5 gene, out.width = "98%", out.height="95%"}
+tag <- "20200527-5"
+tag2 <- "expr-res"
+
+param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+print(param)
+
+plot_all(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
+         paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
+```
+
+MR.ASH results (start from SNP):
+```{r sim 5 snp, out.width = "98%", out.height="95%"}
+tag <- "20200527-5"
+tag2 <- "snp-res"
+
+param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+print(param)
+```
+
+## Simulation 6
+
+MR.ASH results (start from gene):
+```{r sim 6 gene, out.width = "98%", out.height="95%"}
+tag <- "20200527-6"
+tag2 <- "expr-res"
+
+param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+print(param)
+
+plot_all(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
+         paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
+```
+
+MR.ASH results (start from SNP):
+```{r sim 6 snp, out.width = "98%", out.height="95%"}
+tag <- "20200527-6"
+tag2 <- "snp-res"
+
+param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+print(param)
+```

### 2020-06-22 analysis/simulation-susie-ukbchr22.Rmd
fix sa2 bug; locus zoom

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(eval=F)
 ```
@@
 ```
+
+Zoomlocus view for CBX7 region:
+```{r zoomlocus, out.width = "85%", eval = T}
+file.copy("/home/simingz/causalTWAS/simulations/simulation_ashtest_20200503/snp_200622_CBX7/chr22_39026778-40048655.pdf", "output")
+knitr::include_graphics("output/chr22_39026778-40048655.pdf")
+```

### 2020-06-22 code/fit_mr.ash.R

@@ -35,7 +35,8 @@ get_phi <- function(fit) {
 }
 
 get_pve <- function(fit) {
-  n <- dim(fit$data$Z)[1]
+  y <- data.frame(y = fit$data$y)
+  n <- dim(y)[1]
   v <- fit$data$sa2 %*% fit$pi
   v1 <- sum(fit$data$w)/n * v
   pve <- v1/(v1 + fit$sigma2)

### 2020-06-22 code/gen_mr.ash2_output.R

@@ -1,5 +1,29 @@
-
 # global variables: phenores, exprres
+
+para_mr.ash <- function(fit){
+  pi1 <-  1-fit$pi[[1]]
+  pve <- get_pve(fit)
+  p <- c(pi1, pve)
+  names(p) <- c("pi1", "pve")
+  return(p)
+}
+
+para_mr.ash2 <- function(g.fit, s.fit){
+  
+  gpara <- para_mr.ash(g.fit)
+  names(gpara) <- paste0("gene.", names(gpara))
+  
+  spara <- para_mr.ash(s.fit)
+  names(spara) <- paste0("snp.", names(spara))
+  
+  outdf <- data.frame("estimated" = c(gpara,spara))
+  outdf$truth <- c(phenores$param$J.c/phenores$param$J,
+                   phenores$param$pve.expr.truth,
+                   phenores$param$M.c/phenores$param$M,
+                   phenores$param$pve.snp.truth)
+  
+  return(outdf)
+}
 
 gen_mr.ash2_output <- function(g.fit, s.fit, outname){
   e.b <- rep(0, length(g.fit$beta))
@@ -48,5 +72,7 @@ gen_mr.ash2_output <- function(g.fit, s.fit, outname){
   c.outdf[, "p0"] <- c.outdf[, "p0"] - 1 # bed format, for locuszoom
 
   write.table(c.outdf, file= paste0(outname, ".causal-snpexpr.bed") , row.names=F, col.names=F, sep="\t", quote = F)
-
+  
+  para <- para_mr.ash2(g.fit, s.fit)
+  write.table(para , file= paste0(outname, ".param.txt") , row.names=T, col.names=T, sep="\t", quote = F)
 }

### 2020-06-23 analysis/simulation-susie-ukbchr22.Rmd
cbx7 locuszoom

--- 2020-06-15 analysis/simulation-susie-ukbchr22.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(eval=T)
 ```
@@
 ```
+
+[Zoomlocus view for CBX7 region](https://users.rcc.uchicago.edu/~simingz/chr22_39026778-40048655.pdf)
+```{r zoomlocus, out.width = "85%", include = F}
+file.copy("/home/simingz/causalTWAS/simulations/simulation_ashtest_20200503/snp_200622_CBX7/chr22_39026778-40048655.pdf", "~/public_html/chr22_39026778-40048655.pdf")
+```

### 2020-06-24 analysis/simulation-susie-ukbchr22-gtex.adipose.Rmd
susie more examples

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, warning = F)
 ```
@@
 
-  fig <- plot_ly(dt, x = ~ pos, y = ~ pip, type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))
+  fig <- plot_ly(dt, x = ~ pos, y = ~ pip, type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name), width = 600, height = 400)
 
-  fig <- fig %>% layout(yaxis = list(title = "PIP"))
+  fig <- fig %>% layout( yaxis = list(title = "PIP"), plot_bgcolor='#ebf6fc')
   return(fig)
@@
 
-plot_all <- function(susieres, mrashres.gene, mrashres.snp){
+plot_all <- function(susieres, mrashres.gene, mrashres.snp, title){
   dt <- read.table(susieres , header =T)
@@
 
-  fig <- fig %>% layout(autosize = F, width = 600, height = 600, margin = 0.02,
+  fig <- fig %>% layout(title = title, autosize = F,  margin = 0.02,
     xaxis = list(title = "Genomic position"))
@@
 
-## Simulation 1
-Priors:
-```{r para 1}
-mrashtag <- "20200616-1"
-susietag <- "20200616-1"
-readLines(paste0(outputdir, "prior-", susietag, ".R"))
+## Simulation 5
+```{r sim 5 setup}
+tag <- "20200616-5"
+tag2 <- "expr-res"
+mrashres.gene <- paste0(mr.ash2dir, tag, "-mr.ash2s.", tag2, ".expr.txt")
+mrashres.snp <- paste0(mr.ash2dir, tag, "-mr.ash2s.", tag2, ".snp.txt")
 ```
 
-```{r sim 1 setup}
-mrashres.gene <- paste0(mr.ash2dir, mrashtag, "-mr.ash2s.expr-res.expr.txt")
-mrashres.snp <- paste0(mr.ash2dir, mrashtag, "-mr.ash2s.expr-res.snp.txt")
+### SUSIE results: PIP > 0.3 or causal genes
+
+```{r sim 5 examples, include= F, echo = F, eval = F, warning= F}
+# displaying not right
+resall <- Sys.glob(paste0(outputdir, tag, ".", tag2, "*-susieres.txt"))
+l <- htmltools::tagList()
+res <- resall[1]
+for (i in 1:8){
+  res <- resall[i]
+  name <- gsub(".*res-(.+)-susie.*", "\\1", res)
+  l[[i]] <- plot_all(res, mrashres.gene, mrashres.snp, name)
+}
+l
 ```
 
-### SUSIE results: PIP > 0.5 or causal genes
-
-```{r sim 1 examples, out.width = "98%", out.height="95%"}
-nameall <- read.table(paste0(outputdir, "gene-flank500kb-", susietag, ".txt"))[ , 4]
-for (name in nameall){
-  susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
-  print(paste0("SUISE results for regions around ", name))
-  plot_all(susieres, mrashres.gene, mrashres.snp)
-}
+```{r sim 5 examples,  out.width = "98%", out.height="95%", warning= F}
+# displaying not right
+res <- paste0(outputdir, tag, ".", tag2, "-LL22NC03-86G7.1-susieres.txt")
+name <- gsub(".*res-(.+)-susie.*", "\\1", res)
+plot_all(res, mrashres.gene, mrashres.snp, name)
 ```

### 2020-06-24 analysis/simulation-susie-ukbchr22.Rmd
susie more examples

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(eval=T)
+knitr::opts_chunk$set(eval=T, warning = F)
 ```
 
-```{r plot functions}
+```{r plot functions, eval=T}
 library(mr.ash.alpha)
@@
 }
-
 ```
 
-```{r output dir}
+```{r output dir, eval=T}
 simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200503/"
@@
 ## Simulation 1
-Priors:
-```{r para 1}
-readLines(paste0(outputdir, "prior-1.R"))
-```
-
-SUSIE results:
-```{r sim 1 setup}
+```{r sim 1 setup, eval = T}
 mrashtag <- "20200527-1"
@@
 
-```{r sim 1 3, out.width = "98%", out.height="95%"}
+```{r sim 1 3, out.width = "98%", out.height="95%", eval =T}
 name <- "DGCR14"
-susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
-print(paste0("SUISE results for regions around ", name))
-plot_all(susieres, mrashres.gene, mrashres.snp)
+susieres <- paste0(outputdir, susietag, "-", name, "-susieres.txt")
+plot_all(susieres, mrashres.gene, mrashres.snp, name)
 ```
@@
 susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
-print(paste0("SUISE results for regions around ", name))
-plot_all(susieres, mrashres.gene, mrashres.snp)
+plot_all(susieres, mrashres.gene, mrashres.snp, name)
 ```
@@
 susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
-print(paste0("SUISE results for regions around ", name))
-plot_all(susieres, mrashres.gene, mrashres.snp)
-```
-
-### True positives in mr.ash2
-```{r sim 1 8, out.width = "98%", out.height="95%"}
-name <- "LMF2"
-susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
-print(paste0("SUISE results for regions around ", name))
-plot_all(susieres, mrashres.gene, mrashres.snp)
-```
-
-```{r sim 1 9, out.width = "98%", out.height="95%"}
-name <- "DDX17"
-susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
-print(paste0("SUISE results for regions around ", name))
-plot_all(susieres, mrashres.gene, mrashres.snp)
-```
-```{r sim 1 10, out.width = "98%", out.height="95%"}
-name <- "MIF"
-susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
-print(paste0("SUISE results for regions around ", name))
-plot_all(susieres, mrashres.gene, mrashres.snp)
+plot_all(susieres, mrashres.gene, mrashres.snp, name)
 ```
@@
 ```{r sim 3 2, out.width = "98%", out.height="95%"}
-
 name <- "CBX6"
 susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
-print(paste0("SUISE results for regions around ", name))
-plot_all(susieres, mrashres.gene, mrashres.snp)
+plot_all(susieres, mrashres.gene, mrashres.snp, name)
 ```
@@
 ```
+
+## Simulation 5
+
+```{r sim 5 setup}
+tag <- "20200527-5"
+tag2 <- "expr-res"
+susietag <- "20200527-5"
+mrashres.gene <- paste0(mr.ash2dir, tag, "-mr.ash2s.", tag2, ".expr.txt")
+mrashres.snp <- paste0(mr.ash2dir, tag, "-mr.ash2s.", tag2, ".snp.txt")
+```
+
+```{r sim 5 1, out.width = "98%", out.height="95%"}
+name <- "CRELD2"
+susieres <- paste0(outputdir, tag, ".", tag2, "-", name, "-susieres.txt")
+plot_all(susieres, mrashres.gene, mrashres.snp, name)
+```

### 2020-06-24 analysis/simulation-susie-ukbchr22-gtex.adipose.Rmd
susie more examples

@@
 
-```{r sim 5 examples, include= F, echo = F, eval = F, warning= F}
+```{r sim 5, include= F, echo = F, eval = F, warning= F}
 # displaying not right

### 2020-06-24 analysis/index.Rmd
susie more examples
susie examples

@@
 
-  * [Using UKbiobank data (chr22), fusion weights YFS](simulation-susie-ukbchr22.html)
+  * [Using UKbiobank data (chr22), fusion weights YFS,  SUSIE `L=1`](simulation-susie-ukbchr22.html)
+  * [Using UKbiobank data (chr22), fusion weights YFS, SUSIE `L=2`](simulation-susie-ukbchr22-L=2.html)
+  * [Using UKbiobank data (chr22), fusion weights YFS,  SUSIE `L=3`](simulation-susie-ukbchr22-L=3.html)
+
+  * [Using UKbiobank data (chr22), fusion weights Adipose, `L=1`](simulation-susie-ukbchr22-gtex.adipose.html)
 

### 2020-06-24 analysis/simulation-susie-ukbchr22-L=2.Rmd
susie examples

--- 2020-06-24 analysis/simulation-susie-ukbchr22.Rmd
@@
 ---
-title: "Test SUSIE for regions near genes (500kb flanking)"
+title: "Test SUSIE (L=2) for regions near genes (500kb flanking)"
 output: html_document
@@
 name <- "DGCR14"
-susieres <- paste0(outputdir, susietag, "-", name, "-susieres.txt")
+susieres <- paste0(outputdir, susietag, ".L2-", name, "-susieres.txt")
 plot_all(susieres, mrashres.gene, mrashres.snp, name)
@@
 name <- "VPREB3"
-susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
+susieres <- paste0(outputdir, susietag,".L2-", name, "-susieres.txt")
 plot_all(susieres, mrashres.gene, mrashres.snp, name)
@@
 name <- "TCN2"
-susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
+susieres <- paste0(outputdir, susietag,".L2-", name, "-susieres.txt")
 plot_all(susieres, mrashres.gene, mrashres.snp, name)
@@
 name <- "CBX6"
-susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
+susieres <- paste0(outputdir, susietag,".L2-", name, "-susieres.txt")
 plot_all(susieres, mrashres.gene, mrashres.snp, name)
@@
 name <- "CHEK2"
-susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
+susieres <- paste0(outputdir, susietag,".L2-", name, "-susieres.txt")
 plot_all(susieres, mrashres.gene, mrashres.snp, name)
@@
 name <- "MCM5"
-susieres <- paste0(outputdir, susietag,"-", name, "-susieres.txt")
+susieres <- paste0(outputdir, susietag,".L2-", name, "-susieres.txt")
 plot_all(susieres, mrashres.gene, mrashres.snp, name)
@@
 
-[Zoomlocus view for CBX7 region](https://users.rcc.uchicago.edu/~simingz/chr22_39026778-40048655.pdf)
+[Zoomlocus view for this region](https://users.rcc.uchicago.edu/~simingz/chr22_39026778-40048655.pdf)
 ```{r zoomlocus, out.width = "85%", include = F}
@@
 name <- "CRELD2"
-susieres <- paste0(outputdir, tag, ".", tag2, "-", name, "-susieres.txt")
+susieres <- paste0(outputdir, tag, ".", tag2, ".L2-", name, "-susieres.txt")
 plot_all(susieres, mrashres.gene, mrashres.snp, name)

### 2020-06-24 analysis/simulation-susie-ukbchr22-L=3.Rmd
susie examples

--- 2020-06-24 analysis/simulation-susie-ukbchr22-L=2.Rmd
@@
 ---
-title: "Test SUSIE (L=2) for regions near genes (500kb flanking)"
+title: "Test SUSIE (L=3) for regions near genes (500kb flanking)"
 output: html_document
@@
 name <- "DGCR14"
-susieres <- paste0(outputdir, susietag, ".L2-", name, "-susieres.txt")
+susieres <- paste0(outputdir, susietag, ".L3-", name, "-susieres.txt")
 plot_all(susieres, mrashres.gene, mrashres.snp, name)
@@
 name <- "VPREB3"
-susieres <- paste0(outputdir, susietag,".L2-", name, "-susieres.txt")
+susieres <- paste0(outputdir, susietag,".L3-", name, "-susieres.txt")
 plot_all(susieres, mrashres.gene, mrashres.snp, name)
@@
 name <- "TCN2"
-susieres <- paste0(outputdir, susietag,".L2-", name, "-susieres.txt")
+susieres <- paste0(outputdir, susietag,".L3-", name, "-susieres.txt")
 plot_all(susieres, mrashres.gene, mrashres.snp, name)
@@
 name <- "CBX6"
-susieres <- paste0(outputdir, susietag,".L2-", name, "-susieres.txt")
+susieres <- paste0(outputdir, susietag,".L3-", name, "-susieres.txt")
 plot_all(susieres, mrashres.gene, mrashres.snp, name)
@@
 name <- "CHEK2"
-susieres <- paste0(outputdir, susietag,".L2-", name, "-susieres.txt")
+susieres <- paste0(outputdir, susietag,".L3-", name, "-susieres.txt")
 plot_all(susieres, mrashres.gene, mrashres.snp, name)
@@
 name <- "MCM5"
-susieres <- paste0(outputdir, susietag,".L2-", name, "-susieres.txt")
+susieres <- paste0(outputdir, susietag,".L3-", name, "-susieres.txt")
 plot_all(susieres, mrashres.gene, mrashres.snp, name)
@@
 name <- "CRELD2"
-susieres <- paste0(outputdir, tag, ".", tag2, ".L2-", name, "-susieres.txt")
+susieres <- paste0(outputdir, tag, ".", tag2, ".L3-", name, "-susieres.txt")
 plot_all(susieres, mrashres.gene, mrashres.snp, name)

### 2020-06-25 analysis/index.Rmd
SUSIE examples

@@
   * [Using UKbiobank data (chr22), fusion weights Adipose, `L=1`](simulation-susie-ukbchr22-gtex.adipose.html)
+  * [Using UKbiobank data (chr22), fusion weights Adipose, `L=2`](simulation-susie-ukbchr22-gtex.adipose-L=2.html)
+   * [Using UKbiobank data (chr22), fusion weights Adipose, `L=3`](simulation-susie-ukbchr22-gtex.adipose-L=3.html)
 

### 2020-06-25 analysis/simulation-susie-ukbchr22-gtex.adipose-L=2.Rmd
SUSIE examples

--- 2020-06-24 analysis/simulation-susie-ukbchr22-gtex.adipose.Rmd
@@
 
-### SUSIE results: PIP > 0.3 or causal genes
+### SUSIE results
 
@@
 # displaying not right
-resall <- Sys.glob(paste0(outputdir, tag, ".", tag2, "*-susieres.txt"))
+resall <- Sys.glob(paste0(outputdir, tag, ".", tag2, ".L2*-susieres.txt"))
 l <- htmltools::tagList()
@@
 # displaying not right
-res <- paste0(outputdir, tag, ".", tag2, "-LL22NC03-86G7.1-susieres.txt")
-name <- gsub(".*res-(.+)-susie.*", "\\1", res)
+res <- paste0(outputdir, tag, ".", tag2, ".L2-LL22NC03-86G7.1-susieres.txt")
+name <- gsub(".*.L2-(.+)-susie.*", "\\1", res)
 plot_all(res, mrashres.gene, mrashres.snp, name)

### 2020-06-25 analysis/simulation-susie-ukbchr22-gtex.adipose-L=3.Rmd
SUSIE examples

--- 2020-06-25 analysis/simulation-susie-ukbchr22-gtex.adipose-L=2.Rmd
@@
 # displaying not right
-resall <- Sys.glob(paste0(outputdir, tag, ".", tag2, ".L2*-susieres.txt"))
+resall <- Sys.glob(paste0(outputdir, tag, ".", tag2, ".L3*-susieres.txt"))
 l <- htmltools::tagList()
@@
 # displaying not right
-res <- paste0(outputdir, tag, ".", tag2, ".L2-LL22NC03-86G7.1-susieres.txt")
-name <- gsub(".*.L2-(.+)-susie.*", "\\1", res)
+res <- paste0(outputdir, tag, ".", tag2, ".L3-LL22NC03-86G7.1-susieres.txt")
+name <- gsub(".*.L3-(.+)-susie.*", "\\1", res)
 plot_all(res, mrashres.gene, mrashres.snp, name)

### 2020-07-03 analysis/simulation-mr.ash2s-ukbchr22-gtex.adipose.Rmd
mr.ash2s lasso beta init

@@
 
-plot_pip <- function(dt){
+plot_pip <- function(dt, ld = F){
+
+  if (isTRUE(ld)) {
+    fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', 
+                   mode = 'markers', color = ~ max_r2, 
+                   text = ~ paste("Name: ", name, 
+                                  "\n", ifcausal,
+                                  "\nmax R^2: ", round(max_r2, digits = 2), 
+                                  "\nMax pair: ", max_pair))
+  } else {
     pal <- c("salmon", "darkgreen")
@@
     fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))
+  }
 
@@
 
-plot_all <- function(mrashres.gene, gwasres.gene, mrashres.snp, gwasres.snp){
+plot_susie_pip <- function(dt) {
+   pal <- c("salmon", "darkgreen")
+   pal <- setNames(pal, c("Causal", "Non causal"))
+
+   fig <- plot_ly(dt, x = ~ p0, y = ~ susiePIP , type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))
+
+   fig <- fig %>% layout(yaxis = list(title = "SUSIE PIP"))
+}
+
+plot_all <- function(mrashres.gene, gwasres.gene, mrashres.snp, gwasres.snp, susieres, ldres){
   a <- read.table(mrashres.gene , header =T)
@@
   a[a$ifcausal == 0, "ifcausal"] = "Non causal"
+
+  ld <- read.table(ldres, header =T)
+  a <- merge(a, ld, by = "name", all.x = T)
+
+  su <- read.table(susieres, header =T)
+  colnames(su)[colnames(su) == "pip"] <-  "susiePIP"
+  a <- merge(a, su, by = "name", all.x = T)
 
@@
 
-  fig1 <- plot_pip(dt.gene)
+  fig1 <- plot_pip(dt.gene, ld = T)  
   fig2 <- plot_beta(dt.gene) 
   fig3 <- plot_p(dt.gene) 
-  fig4 <- plot_pip(dt.snp)
-  fig5 <- plot_beta(dt.snp)
-  fig6 <- plot_p(dt.snp)
+  fig4 <- plot_susie_pip(dt.gene) 
+  fig5 <- plot_pip(dt.snp) 
+  fig6 <- plot_beta(dt.snp) 
+  fig7 <- plot_p(dt.snp) %>% layout( plot_bgcolor='#EFF6FC')
 
-  fig <- subplot(fig1, fig2, fig3, fig4, fig5, fig6, nrows=6, shareX = T, titleY = T)
+  fig <- subplot(fig1, fig2, fig3, fig4, fig5, fig6, fig7, nrows=7, shareX = T, titleY = T)
 
-  fig <- fig %>% layout(autosize = F, width = 600, height = 800, margin = 0.02,
+  fig <- fig %>% layout(autosize = F, width = 600, height = 900, margin = 0.02,
     xaxis = list(title = "Genomic position"))
@@
 outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
+susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200616/"
+```
+
+```{r getfile function}
+get_files <- function(tag, tag2){
+  par <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt")
+  gmrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt")
+  ggwas <- paste0(outputdir, tag, ".exprgwas.txt.gz")
+  smrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt")
+  sgwas <- paste0(outputdir, tag, ".snpgwas.txt.gz")
+  susie <- paste0(susiedir, tag, ".", tag2, ".L3-susieres.txt")
+  rsq <- paste0(outputdir, tag, ".LD.txt")
+  return(tibble::lst(par, gmrash, ggwas, smrash, sgwas, susie, rsq))
+}
 ```
@@
 
-param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+f <- get_files(tag= tag, tag2 = tag2)
+param <- read.table(f$par, header = T, row.names = 1)
 print(param)
-
-plot_all(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
-         paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
+plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
 ```
 
-MR.ASH results (start from SNP):
-```{r sim 5 snp, out.width = "98%", out.height="95%"}
+MR.ASH results (start from SNP): very similar to start from gene.
+```{r sim 5 snp, out.width = "98%", out.height="95%", eval = F}
 tag <- "20200616-5"
@@
 
-param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+f <- get_files(tag= tag, tag2 = tag2)
+param <- read.table(f$par, header = T, row.names = 1)
 print(param)
-
-plot_all(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
-         paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
+plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
 ```
@@
 
-param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+f <- get_files(tag= tag, tag2 = tag2)
+param <- read.table(f$par, header = T, row.names = 1)
 print(param)
-
-plot_all(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
-         paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
+plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
 ```
 
-MR.ASH results (start from SNP):
-```{r sim 6 snp, out.width = "98%", out.height="95%"}
+MR.ASH results (start from SNP): very similar to start from gene
+```{r sim 6 snp, out.width = "98%", out.height="95%", eval = F}
 tag <- "20200616-6"
@@
 
-param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
+f <- get_files(tag= tag, tag2 = tag2)
+param <- read.table(f$par, header = T, row.names = 1)
 print(param)
+plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
+```
 
-plot_all(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
-         paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
+## Simulation 7
+Use lasso trained weight model from FUSION twas. This makes eQTLs (SNP weights != 0) more sparse (<20/gene) compared to BSLMM/BLUP (a few hundred/per gene).
+
+Use beta.init = lasso for mr.ash.
+
+MR.ASH results (use lasso for all):
+```{r sim 7 lasso, out.width = "98%", out.height="95%", eval = T}
+tag <- "20200616-7"
+tag2 <- "lasso"
+
+f <- get_files(tag= tag, tag2 = tag2)
+param <- read.table(f$par, header = T, row.names = 1)
+print(param)
+plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
 ```
+
+MR.ASH results (use lasso for just for SNPs):
+```{r sim 7 lassosnp, out.width = "98%", out.height="95%", eval = T}
+tag <- "20200616-7"
+tag2 <- "lassoSNP"
+
+f <- get_files(tag= tag, tag2 = tag2)
+param <- read.table(f$par, header = T, row.names = 1)
+print(param)
+plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
+```
+
+## Simulation 8
+Use lasso trained weight model from FUSION twas. This makes eQTLs (SNP weights != 0) more sparse (<20/gene) compared to BSLMM/BLUP (a few hundred/per gene).
+
+Use beta.init = lasso for mr.ash.
+
+MR.ASH results (use lasso for all):
+```{r sim 8 lasso, out.width = "98%", out.height="95%", eval = T}
+tag <- "20200616-8"
+tag2 <- "lasso"
+
+f <- get_files(tag= tag, tag2 = tag2)
+param <- read.table(f$par, header = T, row.names = 1)
+print(param)
+plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
+```
+
+MR.ASH results (use lasso for just for SNPs):
+```{r sim 8 lassosnp, out.width = "98%", out.height="95%", eval = T}
+tag <- "20200616-8"
+tag2 <- "lassoSNP"
+
+f <- get_files(tag= tag, tag2 = tag2)
+param <- read.table(f$par, header = T, row.names = 1)
+print(param)
+plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
+```

### 2020-07-03 analysis/simulation_simpleveb-boost2.Rmd
mr.ash2s lasso beta init

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, message = F)
 library(mr.ash.alpha)
@@
 ```
+
+### init with lasso
+
+```{r run mr.ash2, lasso}
+X.gene <- X[,gene.idx]
+X.SNP <- X[, SNP.idx]
+fit <- mr.ash2s(X.gene, X.SNP, y, mr.ash.init = "lasso")
+
+print("for gene effect: ")
+summary_mr.ash(fit$fit1)
+plot_beta(beta[gene.idx], fit$fit1$beta, main = "beta for gene effect")
+
+print("for SNP effect: ")
+summary_mr.ash(fit$fit2)
+plot_beta(beta[SNP.idx], fit$fit2$beta, main = "beta for SNP effect")
+```
+
+### init with lasso SNP
+
+```{r run mr.ash2, lassosnp}
+X.gene <- X[,gene.idx]
+X.SNP <- X[, SNP.idx]
+fit <- mr.ash2s(X.gene, X.SNP, y, mr.ash.init = "lassoX2")
+
+print("for gene effect: ")
+summary_mr.ash(fit$fit1)
+plot_beta(beta[gene.idx], fit$fit1$beta, main = "beta for gene effect")
+
+print("for SNP effect: ")
+summary_mr.ash(fit$fit2)
+plot_beta(beta[SNP.idx], fit$fit2$beta, main = "beta for SNP effect")
+```

### 2020-07-03 code/run_simulate_data.R

@@ -44,7 +44,7 @@ outname <- args[4]
 loginfo("generating imputed expression file: %s",  paste0(outname, "-cis-expr.Rd"))
 
 weight <- args[2]
-exprres <- cis_expr(dat, weight, method = "best", checksnps = F)
+exprres <- cis_expr(dat, weight, method = "lasso", checksnps = F)
 save(exprres, file = paste0(outname, "-cis-expr.Rd"))
 
 # load(paste0(outname, "-cis-expr.Rd"))

### 2020-07-03 code/run_test_susie.R

@@ -50,18 +50,21 @@ regions <- read.table(args[5], stringsAsFactors = F, header =T)
 
 filter <- T
 if (filter) {
-  regions <- regions[regions$ifcausal == 1 | regions$PIP > 0.3, ]
+  #regions <- regions[regions$ifcausal == 1 | regions$PIP > 0.3, ]
+  regions <- regions[regions$PIP > 0.3, ]
   flank = 500000
   regions$p0 <- regions$p0 - flank
   regions$p1 <- regions$p1 + flank
 }
 
 outname <- args[6]
-L = 1
+L = 3
 if (length(args) == 7){
   L = as.numeric(args[7])
   outname <- paste0(outname, ".L", L)
 }
+
+outlist <- list()
 
 for (i in 1:nrow(regions)){
   chr <- regions[i, "chr"]
@@ -92,8 +95,11 @@ for (i in 1:nrow(regions)){
   colnames(outdf) <- c("name", "chr", "pos", "pip")
 
   write.table( outdf , file= paste(outname, name, "susieres.txt", sep = "-")  , row.names=F, col.names= T, sep="\t", quote = F)
+
+  outlist[[name]] <- outdf[outdf[, "name"] == name, ]
 }
 
-
+outgdf <- do.call(rbind, outlist)
+write.table( outgdf , file= paste(outname, "susieres.txt", sep = "-")  , row.names=F, col.names= T, sep="\t", quote = F)
 
 

### 2020-07-03 code/simulate_phenotype.R

@@ -13,18 +13,19 @@ library(data.table)
 #' @return a matrix, N x J. Missing genotypes will not contribute to cis-expr and give warnings.
 cis_expr <- function(dat, weight, method = "bslmm", checksnps = F){
   exprlist <- list()
+  qclist <- list()
   wgtdir <- dirname(weight)
   wgtposfile <- file.path(wgtdir, paste0(basename(weight), ".pos"))
 
   wgtpos <- read.table(wgtposfile, header = T, stringsAsFactors = F)
-  
-  wgtpos <- transform(wgtpos, 
-            "ID" = ifelse(duplicated(ID) | duplicated(ID, fromLast=TRUE), 
-                                 paste(ID, ave(ID, ID, FUN=seq_along), sep='_ID'), 
+
+  wgtpos <- transform(wgtpos,
+            "ID" = ifelse(duplicated(ID) | duplicated(ID, fromLast=TRUE),
+                                 paste(ID, ave(ID, ID, FUN=seq_along), sep='_ID'),
                                  ID))
-  
+
   write.table(wgtpos , file= paste0(wgtposfile, ".IDfixed") , row.names=F, col.names=T, sep="\t", quote = F)
-  
+
   print(paste0("Number of genes with weights provided: ", dim(wgtpos)[1]))
 
   for (i in 1: dim(wgtpos)[1]){
@@ -37,6 +38,7 @@ cis_expr <- function(dat, weight, method = "bslmm", checksnps = F){
     if (g.method == "best"){
       g.method = names(which.max(cv.performance["rsq", ]))
     }
+    if (!(g.method %in% names(cv.performance[1,]))) next
 
     wgt.idx <- match(rownames(wgt.matrix), dat$snp)
 
@@ -51,20 +53,32 @@ cis_expr <- function(dat, weight, method = "bslmm", checksnps = F){
       }
     }
 
-    g <- as.matrix(dat$G[, wgt.idx])
-    g[, is.na(wgt.idx)] <- 0
-    
-    exprlist[[gname]][["expr"]] <- g %*% wgt.matrix[, g.method]
-    exprlist[[gname]][["chr"]] <- wgtpos[wgtpos$ID == gname, "CHR"]
-    exprlist[[gname]][["p0"]] <- wgtpos[wgtpos$ID == gname, "P0"]# position boundary 1
[…]

### 2020-07-21 analysis/simulation_simpleveb-boost2.Rmd
FBM workflow

--- 2020-05-28 analysis/simulation_simpleveb-boost2.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, message = F)
 library(mr.ash.alpha)
+library(biglasso)
+library(bigmemory)
+library(bigstatsr)
 source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")
@@
 ```{r run mr.ash}
-fit <- mr.ash(X, y, method="caisa")
+myorder <- 1:dim(X)[2]
+myorder[gene.idx] <- 1:length(gene.idx)
+myorder[SNP.idx] <- (length(gene.idx) + 1): length(myorder)
+
+X <- cbind(X[, gene.idx], X[, SNP.idx])
+fit <- mr.ash(X, y, method="caisa", update.order = NULL)
 summary_mr.ash(fit)
@@
 plot_beta(beta[SNP.idx], fit$beta[SNP.idx], main = "beta for SNP effect")
+```
+
+```{r run mr.ash}
+myorder <- 1:dim(X)[2]
+myorder[gene.idx] <- (length(SNP.idx) + 1): length(myorder)
+myorder[SNP.idx] <- 1:length(SNP.idx)
+X <- cbind(X[,SNP.idx], X[,gene.idx])
+fit2 <- mr.ash(X, y, method="caisa", update.order = NULL)
+summary_mr.ash(fit2)
+
+plot_beta(beta[gene.idx], fit2$beta[gene.idx], main = "beta for gene effect")
+plot_beta(beta[SNP.idx], fit2$beta[SNP.idx], main = "beta for SNP effect")
 ```
@@
 ```{r run mr.ash2, gene}
-X.gene <- X[,gene.idx]
-X.SNP <- X[, SNP.idx]
-fit <- mr.ash2s(X.gene, X.SNP, y)
+X.gene <- X[, gene.idx]
+X.SNP <- as_FBM(X[, SNP.idx])
+fit <- mr.ash2s(X.gene, X.SNP, y, init.order = "expr-snp")
 
 print("for gene effect: ")
-summary_mr.ash(fit$fit1)
+summary_mr.ash(fit$fit2)
 plot_beta(beta[gene.idx], fit$fit1$beta, main = "beta for gene effect")
@@
 print("for SNP effect: ")
-summary_mr.ash(fit$fit2)
+summary_mr.ash(fit$fit1)
 plot_beta(beta[SNP.idx], fit$fit2$beta, main = "beta for SNP effect")
@@
 ```{r run mr.ash2, snp}
-X.gene <- X[,gene.idx]
-X.SNP <- X[, SNP.idx]
-fit <- mr.ash2s(X.SNP, X.gene, y)
+fit <- mr.ash2s(X.gene, X.SNP, y, init.order = "snp-expr")
 
@@
 summary_mr.ash(fit$fit2)
-plot_beta(beta[gene.idx], fit$fit2$beta, main = "beta for gene effect")
+plot_beta(beta[gene.idx], fit$fit1$beta, main = "beta for gene effect")
 
@@
 summary_mr.ash(fit$fit1)
-plot_beta(beta[SNP.idx], fit$fit1$beta, main = "beta for SNP effect")
+plot_beta(beta[SNP.idx], fit$fit2$beta, main = "beta for SNP effect")
 ```
+
+### init with lasso
+
+```{r run mr.ash2, lasso}
+fit <- mr.ash2s(X.gene, X.SNP, y, mr.ash.init = "lasso")
+
+print("for gene effect: ")
+summary_mr.ash(fit$fit2)
+plot_beta(beta[gene.idx], fit$fit1$beta, main = "beta for gene effect")
+
+print("for SNP effect: ")
+summary_mr.ash(fit$fit1)
+plot_beta(beta[SNP.idx], fit$fit2$beta, main = "beta for SNP effect")
+```
+
+### init with lasso SNP
+
+```{r run mr.ash2, lassosnp}
+fit <- mr.ash2s( X.gene, X.SNP, y, mr.ash.init = "lassoSNP")
+
+print("for gene effect: ")
+summary_mr.ash(fit$fit2)
+plot_beta(beta[gene.idx], fit$fit1$beta, main = "beta for gene effect")
+
+print("for SNP effect: ")
+summary_mr.ash(fit$fit1)
+plot_beta(beta[SNP.idx], fit$fit2$beta, main = "beta for SNP effect")
+```

### 2020-07-21 code/mr.ash2_FBM.R

@@ -265,25 +265,14 @@ mr.ash2s <- function(expr,
                      backingfile = outname,
                      ncores = 1, iter = 30
                      ){
+
+  mr.ash.init <- match.arg(mr.ash.init)
+  init.order <- match.arg(init.order)
+  iter.order <- match.arg(iter.order)
+
   n <- snp$nrow
   p.expr <- dim(expr)[2]
   p.snp <- snp$ncol
-
-  # prepare w
-  print("calculating w")
-  snp.m <- snp[]
-  w.snp <- colSums(snp.m^2)
-  rm(snp.m); gc()
-
-  w.expr <- colSums(expr^2)
-  w = c(w.expr, w.snp)
-
-  # prepare order
-  if (init.order == "es"){
-    o <- 1: (p.expr + p.snp)
-  } else if ( init.order == "se"){
-    o <- c(1: p.expr + p.snp, 1: p.snp)
-  }
 
   # prepare X
   if (file.exists(paste0(backingfile,".bk"))) {
@@ -296,6 +285,15 @@ mr.ash2s <- function(expr,
       X[, ind + p.expr] <<- m[, ind, drop = FALSE]; return(0)
     }, a.combine = 'c', block.size = 2000, ncores = ncores)
   }
+
+  # prepare w
+  print("calculating w")
+  w.snp <- big_apply(snp, a.FUN = function(m, ind) {
+    colSums(m[, ind, drop = FALSE]^2)
+  }, a.combine = 'c', block.size = 2000, ncores = ncores)
+
+  w.expr <- colSums(expr^2)
+  w = c(w.expr, w.snp)
 
   # prepare mr.ash beta.init
   print("prepare mr.ash.init")
@@ -319,6 +317,12 @@ mr.ash2s <- function(expr,
     b <- c(rep(0, p.expr), b.snp[-1])
   }
 
+  # prepare order
+  if (init.order == "es"){
+    o <- 1: (p.expr + p.snp)
+  } else if ( init.order == "se"){
+    o <- c(1: p.expr + p.snp, 1: p.snp)
+  }
 
   # mr.ash for all data as initiation
   print("run mr.ash to initiate")
@@ -336,19 +340,27 @@ mr.ash2s <- function(expr,
   rm(fit.init, X.m); gc()
 
   if (iter.order == "es"){
-    X1 <- expr
-    X2 <- as.matrix(snp$bm())
[…]

### 2020-07-24 analysis/index.Rmd
chr17:22

@@
 
+## Summarize cTWAS results
+We compare cTWAS results with other methods, summarizing results over multiple simulation runs.
+
+  * [Using UKbiobank data (chr22), fusion weights Adipose](simulation-multi-ukbchr22-gtex.adipose.html)
+  * [Using UKbiobank data (chr22), fusion weights Adipose, setting 2](simulation-multi-ukbchr22-gtex.adipose2.html)
+  * [Using UKbiobank data (chr17:22), fusion weights Adipose](simulation-multi-ukbchr17:22-gtex.adipose.html)
+
 # Related software

### 2020-07-24 analysis/simulation-multi-ukbchr17:22-gtex.adipose.Rmd
chr17:22

---
title: "Summarize twas"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  message= F)
```

Run simulation 8 times for ukb chr 17 to chr 22 combined.

```{r plot functions}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})
library(tidyr)
library(plyr)
```
```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
tags <- paste0('20200721-1-', c(2,4:9))
tag2s <- c('zeroes-es', 'zerose-es', 'lassoes-es','lassoes-se')
```

```{r getfile function}
get_files <- function(tag, tag2){
  par <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt")
  rpip <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".rPIP.txt")

  gmrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt")
  smrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt")

  ggwas <- paste0(outputdir, tag, ".exprgwas.txt.gz")
  sgwas <- paste0(outputdir, tag, ".snpgwas.txt.gz")

  gsusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.expr.txt")
  ssusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.snp.txt")

  return(tibble::lst(par, rpip, gmrash, ggwas, smrash, sgwas, gsusie, ssusie))
}
```
# Mr.ash2 parameter estimation
Results for 9 simulations runs, using different initiate and update strategy
```{r param function}
show_param <- function(tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  parf <- lapply(f, '[[', "par")
  param <- do.call(rbind, lapply(parf, function(x) t(read.table(x))[2:1,]))
  knitr::kable(param)
}
```

## NULL; expr-snp; expr-snp

```{r param 1}
show_param(tag2s[1])
```

## NULL; snp-expr; expr-snp

```{r param 2}
show_param(tag2s[2])
```

## lasso; expr-snp; expr-snp

```{r param 3}
show_param(tag2s[3])
```

## lasso; expr-snp; snp-expr

```{r param 4}
show_param(tag2s[4])
```

# Regional mr.ash2s PIP overview
Take simulation 1 (NULL; expr-snp; expr-snp) as examples. We use region size *500kb* and PIP cut off at *0.5* for SUSIE.

```{r rpip 1, fig.height=4, fig.width=10}
f <- get_files(tag= tags[1], tag2 = tag2s[1])
a <- read.table(f[["rpip"]], header = T)
plot(a$p0, a$rPIP, pch =19, col ='salmon', xlab = "position", ylab= "Sum of PIP")
grid()
```

# PIP scatter plot
mr.ash2s PIP vs. susie PIP.
```{r scatterplot}
scatter_plot_PIP<- function(tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  mrashf <- lapply(f, '[[', "gmrash")
  names(mrashf) <- tags

  susief <- lapply(f, '[[', "gsusie")
  names(susief) <- tags

  .tagname <- function(x, flist){
    a <- read.table(flist[[x]], header =T)
    a[, "name"] <- paste0(x, ":", a[, "name"])
    a
  }
  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))

  res <- merge(mrashres, susieres, by = "name", all = T)

  res <- res[complete.cases(res),]
  res <- rename(res, c("PIP" = "mr.ash_PIP", "pip" = "SUSIE_PIP", "pip.null" = "SUSIE_PIP_null") )
  res$ifcausal <- mapvalues(res$ifcausal, 
          from=c(0,1), 
          to=c("Non causal", "Causal"))

  fig1 <- plot_ly(data = res, x = ~ mr.ash_PIP, y = ~ SUSIE_PIP, color = ~ ifcausal, 
                 colors = c( "salmon", "darkgreen"))

  fig2 <- plot_ly(data = res, x = ~ mr.ash_PIP, y = ~ SUSIE_PIP_null, color = ~ ifcausal, 
                 colors = c( "salmon", "darkgreen"))

  fig <- subplot(fig1, fig2, titleX = TRUE, titleY = T, margin = 0.1)
  fig
}
```
## NULL; expr-snp; expr-snp 
```{r sca 1, fig.height=4, fig.width=10}
scatter_plot_PIP(tag2s[1])
```

## NULL; snp-expr; expr-snp

```{r sca 2, fig.height=4, fig.width=10}
scatter_plot_PIP(tag2s[2])
```

## lasso; expr-snp; expr-snp

```{r sca 3, fig.height=4, fig.width=10}
scatter_plot_PIP(tag2s[3])
```

## lasso; expr-snp; snp-expr

```{r sca 4, fig.height=4, fig.width=10}
scatter_plot_PIP(tag2s[4])
```

# ROC curve
```{r roc}
ROC_plot<- function(tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  mrashf <- lapply(f, '[[', "gmrash")
  names(mrashf) <- tags

  susief <- lapply(f, '[[', "gsusie")
  names(susief) <- tags

  gwasf <- lapply(f, '[[', "ggwas")
  names(gwasf) <- tags

  .tagname <- function(x, flist, colnames = NULL){
    a <- read.table(flist[[x]], header =T)
    if (!is.null(colnames)){
      colnames(a) <- colnames
    }
    a[, "name"] <- paste0(x, ":", a[, "name"])
    a
  }
  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))
  gwasres <- do.call(rbind, lapply(tags, .tagname, flist = gwasf, 
                                   colnames =  c("chr",	"p0",	"p1", "name", "Estimate", "Std.Error", "t-value", "PVALUE")))

  res <- merge(mrashres, susieres, by = "name", all = T)
  res <- merge(res, gwasres, by = "name", all = T)

  res <- res[complete.cases(res),]
  res <- rename(res, c("PIP" = "mr.ash", "pip" = "SUSIE", "PVALUE" = "TWAS") )
  res[,"TWAS"] <- -log10(res[, "TWAS"])

  roccolors <-  c("red", "green", "blue")
  methods <- c("mr.ash", "SUSIE", "TWAS")
  plot(0, xlim=c(0,1), ylim=c(0,1), col="white", xlab = "FPR", ylab = "TPR")
  for (i in 1:3){
    method <- methods[i]
    bordered <- res[order(res[,method]),] 
    actuals <- bordered$ifcausal == 1
    sens <- (sum(actuals) - cumsum(actuals))/sum(actuals)
    spec <- cumsum(!actuals)/sum(!actuals)
    lines(1 - spec, sens, type = "l", col = roccolors[i])
    abline(c(0,0),c(1,1))
    auc <- sum(spec*diff(c(0, 1 - sens)))
    cat("AUC for ", method, ": ", auc)
  }
  legend(0.6,0.3, legend= methods, col=roccolors, lty=1, cex=0.8)
  grid()
}
```

## NULL; expr-snp; expr-snp 
```{r roc 1, fig.height=4, fig.width=4}
ROC_plot(tag2s[1])
```

## NULL; snp-expr; expr-snp

```{r roc 2, fig.height=4, fig.width=4}
ROC_plot(tag2s[2])
```

## lasso; expr-snp; expr-snp

```{r roc 3, fig.height=4, fig.width=4}
ROC_plot(tag2s[3])
```

## lasso; expr-snp; snp-expr

```{r roc 4, fig.height=4, fig.width=4}
ROC_plot(tag2s[4])
```

### 2020-07-24 analysis/simulation-multi-ukbchr22-gtex.adipose.Rmd
chr17:22

--- 2020-07-24 analysis/simulation-multi-ukbchr17:22-gtex.adipose.Rmd
@@
 
-Run simulation 8 times for ukb chr 17 to chr 22 combined.
+Run simulation 9 times for ukb chr 22.
 
@@
 ```{r output dir}
-simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
-outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
-susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
-tags <- paste0('20200721-1-', c(2,4:9))
-tag2s <- c('zeroes-es', 'zerose-es', 'lassoes-es','lassoes-se')
+simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
+outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
+susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200616/"
+tags <- paste0('20200616-7-', 1:9)
+tag2s <- c('expr-snp', 'snp-expr', 'lassoexpr-snp','lassoSNPes-es','lassoes-se' )
 ```
@@
 show_param(tag2s[4])
+```
+
+## lassosnp; expr-snp; expr-snp
+
+```{r param 5}
+show_param(tag2s[5])
 ```
@@
 scatter_plot_PIP(tag2s[4])
+```
+
+## lassosnp; expr-snp; expr-snp
+
+```{r sca 5, fig.height=4, fig.width=10}
+scatter_plot_PIP(tag2s[5])
 ```
@@
 ```
+
+## lassosnp; expr-snp; expr-snp
+
+```{r roc 5, fig.height=4, fig.width=4}
+ROC_plot(tag2s[5])
+```

### 2020-07-24 analysis/simulation-multi-ukbchr22-gtex.adipose2.Rmd
chr17:22

--- 2020-07-24 analysis/simulation-multi-ukbchr17:22-gtex.adipose.Rmd
@@
 ```{r output dir}
-simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
-outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
-susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
-tags <- paste0('20200721-1-', c(2,4:9))
+simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
+outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
+susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200616/"
+tags <- paste0('20200616-8-', 1:9)
 tag2s <- c('zeroes-es', 'zerose-es', 'lassoes-es','lassoes-se')

### 2020-07-24 code/input_reformat.R

@@ -166,11 +166,6 @@ pgen2fbm <- function(pfile, select=NULL, scale = F,
   qc <- pgen2mtx(
     pfile, select = select, .transform = fill_FBM, progress = T, part_size = 50 * 1024^2, ... = ...)
 
-  # rm(X);gc()
-  #
-  # X <- readRDS(paste0(backingfile, ".rds"))
-  #
-
   # Verify scaling results
   if (isTRUE(scale)){
     cat("After scaling, mean range:", range(big_scale()(X)[,"center"]), "\n")

### 2020-07-24 code/mr.ash2.R

@@ -273,7 +273,6 @@ mr.ash2s <- function(expr,
   print("calculating w")
   snp.m <- snp[]
   w.snp <- colSums(snp.m^2)
-  rm(snp.m); gc()
 
   w.expr <- colSums(expr^2)
   w = c(w.expr, w.snp)
@@ -286,23 +285,14 @@ mr.ash2s <- function(expr,
   }
 
   # prepare X
-  if (file.exists(paste0(backingfile,".bk"))) {
-    X <- readRDS(paste0(backingfile, ".rds"))
-  } else {
-    X <- as_FBM(expr, backingfile = backingfile)$save()
-    X$add_columns(p.snp)
-
-    a <- big_apply(snp, a.FUN = function(m, ind) {
-      X[, ind + p.expr] <<- m[, ind, drop = FALSE]; return(0)
-    }, a.combine = 'c', block.size = 2000, ncores = ncores)
-  }
+  X <- cbind(expr, snp.m)
 
   # prepare mr.ash beta.init
   print("prepare mr.ash.init")
   if (is.null(mr.ash.init)){
     b <- NULL
   } else if (mr.ash.init == "lasso"){
-    X.bm <- X$bm()
+    X.bm <- as.big.matrix(X)
     time.cvfit <- system.time(
       cvfit <- cv.biglasso(X.bm, y, screen = 'SSR-BEDPP', seed = 1234, ncores = ncores, nfolds = 5)
     )
@@ -310,7 +300,7 @@ mr.ash2s <- function(expr,
     b <- coef(cvfit)
     b <- b[-1] # discard intercept
   } else if (mr.ash.init == "lassoSNP"){
-    snp.bm <- snp$bm()
+    snp.bm <- as.big.matrix(snp.m)
     time.cvfit <- system.time(
       cvfit <- cv.biglasso(snp.bm, y, screen = 'SSR-BEDPP', seed = 1234, ncores = ncores, nfolds = 5)
     )
@@ -322,7 +312,7 @@ mr.ash2s <- function(expr,
 
   # mr.ash for all data as initiation
   print("run mr.ash to initiate")
-  X.m <- as.matrix(X$bm())
+  X.m <- X
   fit.init <- mr.ashs(X.m, y, w = w, beta.init = b, update.order = o)
 
   # start interation
@@ -333,13 +323,13 @@ mr.ash2s <- function(expr,
 
   sigma2 <- fit.init$sigma2
 
[…]

### 2020-07-24 code/mr.ash2_FBM.R

@@ -249,7 +249,7 @@ mr.ash2s_iter <- function(X1, X2,
 #' @param snp FBM
 #' @param expr matrix
 #' @param mr.ash.init beta.init for mr.ash
-#' - "NULL", init mr.ash using default
+#' - "zero", init mr.ash using default
 #' - "lasso", init mr.ash using lasso
 #' - "lassoSNP", init with lasso for SNP and NULL for gene
 #' @param init.order "es", expr-snp; "se", "snp-expr"
@@ -258,7 +258,7 @@ mr.ash2s_iter <- function(X1, X2,
 mr.ash2s <- function(expr,
                      snp,
                      y,
-                     mr.ash.init = c(NULL, "lasso", "lassoSNP"),
+                     mr.ash.init = c("zero", "lasso", "lassoSNP"),
                      init.order = c("es", "se"),
                      iter.order =c("es", "se"),
                      outname = "mr.ash2s",
@@ -297,7 +297,7 @@ mr.ash2s <- function(expr,
 
   # prepare mr.ash beta.init
   print("prepare mr.ash.init")
-  if (is.null(mr.ash.init)){
+  if (mr.ash.init == "zero"){
     b <- NULL
   } else if (mr.ash.init == "lasso"){
     X.bm <- X$bm()

### 2020-07-24 code/run_UKB_process.R

@@ -1,0 +1,57 @@
+codedir <- "/project2/mstephens/causalTWAS/causal-TWAS/code/"
+source(paste0(codedir,"input_reformat.R"))
+
+#----------------------------chr17-22 combined-------------------
+weight <- "/home/simingz/causalTWAS/fusion_weights/Adipose_Subcutaneous"
+method <- "lasso"
+wgtcut <- 1e-8 # abs(wight) > wgtcut
+prune <- c(T, rep(F,9)) # T will be kept
+
+wgtfs <- read.table(paste0(weight,".pos"), header = T)
+wgtdir <-dirname(weight)
+wgtfs <- paste0(wgtdir, "/", wgtfs$WGT)
+snpall <- NULL
+snpannoall <- NULL
+for (wgtf in wgtfs){
+  load(wgtf)
+  snp <- rownames(wgt.matrix[abs(wgt.matrix[, method]) > wgtcut,])
+  snpanno <- snps[snps[,2] %in% snp, ]
+  snpanno <- cbind(snpanno, wgt.matrix[snpanno[,2],])
+  snpall <- c(snpall, snp)
+  snpannoall <- rbind(snpannoall, snpanno)
+}
+
+write.table(snpannoall, file= paste(weight, method, "eqtl.txt", sep= ".") , row.names=F, col.names=T, sep="\t", quote = F)
+
+pfiles <- paste0("/home/simingz/causalTWAS/ukbiobank/ukb_chr", 17:22, "_s20000.pgen")
+
+mtxall <- NULL
+varall <- NULL
+for (pfile in pfiles){
+  pfile <- drop_ext(pfile)
+  pvar <- paste0(pfile, ".pvar")
+  var <- read.table(pvar, header = F)
+  eqtl <- var[,3] %in% snpall
+  prunetag <- rep_len(prune, nrow(var))
+  select <- (1:nrow(var))["|"(eqtl, prunetag)]
+  varall <- rbind(varall, var[select,])
+  mtx <- pgen2mtx(pfile, nb_parts = 1, select=select, progress = T)
+  mtx.scaled <- scaleRcpp(mtx)
+  mtxall <- cbind(mtxall, mtx.scaled)
+}
+
+dat <- list("G"       = mtxall,
+            "chr"     = as.matrix(varall[,1]),
+            "pos"     = as.matrix(varall[,2]),
+            "snp"     = as.matrix(varall[,3]),
+            "counted" = as.matrix(varall[,4]),
[…]

### 2020-07-24 analysis/index.Rmd
chr17:22

--- 2020-06-25 analysis/index.Rmd
@@
 
+## Summarize cTWAS results
+We compare cTWAS results with other methods, summarizing results over multiple simulation runs.
+
+  * [Using UKbiobank data (chr22), fusion weights Adipose](simulation-multi-ukbchr22-gtex.adipose.html)
+  * [Using UKbiobank data (chr22), fusion weights Adipose, setting 2](simulation-multi-ukbchr22-gtex.adipose2.html)
+  * [Using UKbiobank data (chr17:22), fusion weights Adipose](simulation-multi-ukbchr17to22-gtex.adipose.html)
+
 # Related software

### 2020-07-31 analysis/simulation-multi-ukbchr17to22-gtex.adipose.Rmd
workflow and susie

---
title: "Summarize twas"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  message= F)
```

Run simulation 8 times for ukb chr 17 to chr 22 combined. SNPs are downsampled to 1/10, eQTLs defined by FUSION-TWAS (Adipose/GTEx) lasso effect size > 0 were kept, p= 86k, n = 20k.

```{r plot functions}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})
library(tidyr)
library(plyr)
library(kableExtra)
```
```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
tags <- paste0('20200721-1-', c(2,4:9))
tags_ext <- paste0('20200721-1-', 1:50)
tags_ext2 <- paste0('20200721-1-', c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 15, 16, 17, 19, 21, 22, 23, 26, 27, 28, 29, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50))
tag2s <- c('zeroes-es', 'zerose-es', 'lassoes-es','lassoes-se')
```

```{r getfile function}
get_files <- function(tag, tag2){
  par <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt")
  rpip <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".rPIP.txt")

  gmrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt")
  smrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt")

  ggwas <- paste0(outputdir, tag, ".exprgwas.txt.gz")
  sgwas <- paste0(outputdir, tag, ".snpgwas.txt.gz")

  gsusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.expr.txt")
  ssusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.snp.txt")

  return(tibble::lst(par, rpip, gmrash, ggwas, smrash, sgwas, gsusie, ssusie))
}
```
# Mr.ash2 parameter estimation

Results for 7 simulations runs, using different initiate and update strategy

```{r param function}
show_param <- function(tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  parf <- lapply(f, '[[', "par")
  param <- do.call(rbind, lapply(parf, function(x) t(read.table(x))[2:1,]))
  truth <- param[1:(nrow(param)/2)*2-1,]
  est <- param[1:(nrow(param)/2)*2,]
  outdt <- matrix(0, ncol = 2*ncol(param), nrow = nrow(param)/2)
  outdt[,c(1,3,5,7)] <- truth
  outdt[,c(2,4,6,8)] <- est
  outdt <-cbind(1:nrow(outdt),outdt)
  colnames(outdt) <- c("Simulation#", paste0(rep(c("Truth","Est."),4)))
  knitr::kable(outdt) %>%
  kable_styling("striped") %>%
  add_header_above(c(" " = 1, "Gene.pi1" = 2, "Gene.PVE" = 2, "SNP.pi1" = 2, "SNP.PVE" =2))
}
```

## NULL; expr-snp; expr-snp

```{r param 1}
show_param(tag2s[1])
```

## NULL; snp-expr; expr-snp

```{r param 2}
show_param(tag2s[2])
```

## lasso; expr-snp; expr-snp

```{r param 3}
show_param(tag2s[3])
```

## lasso; expr-snp; snp-expr

```{r param 4}
show_param(tag2s[4])
```

# Regional mr.ash2s PIP overview
Take simulation 1 (NULL; expr-snp; expr-snp) as examples. We use region size *500kb* and PIP cut off at *0.2* for SUSIE.

```{r rpip 1, fig.height=4, fig.width=10}
chrom = 18
f <- get_files(tag= tags[1], tag2 = tag2s[1])
allchr <- read.table(f[["rpip"]], header = T)
a <- allchr[allchr["chrom"]==chrom,]
print(paste("plot for chr", chrom))
par(mar=c(5, 4, 4, 6) + 0.1)
with(a, plot(p0, rPIP, col ='salmon', xlab = "position", ylab= "Sum of PIP", type = 'h', lwd = 2))
par(new = T)
with(a, plot(p0, nCausal, pch =19, col = "darkgreen",axes = FALSE, bty = "n", xlab = "", ylab = ""))
axis(side = 4)
mtext(side = 4, line = 3, 'No. causal signals')
legend("topleft",
       legend=c("Mr.ASH PIP", "# Causal"),
       lty=c(1,0), pch=c(NA, 19), col=c("salmon", "darkgreen"))
grid()
```
```{r rpip 2, fig.height=4, fig.width=4}
plot( allchr$rPIP, allchr$nCausal, bg = rgb(red = 1, green = 0, blue = 0, alpha = 0.1), col= "red", pch = 21, cex =2,
     ylab = "# Causal", xlab= "Sum of PIP")
```

# PIP calibration
We run 50 simulations and combine results.
```{r caliplot}
cp_plot <- function(pip, ifcausal, main = "PIP"){
  # ifcausal:0,1
  a_bin <- cut(pip, breaks= seq(0,1, by=0.1))
  PIP_mean = c(by(pip, a_bin, FUN = mean))
  Freq = c(by(ifcausal, a_bin, FUN = mean))
  plot(PIP_mean, Freq, xlim= c(0,1), ylim=c(0,1), pch =19, main = main)
  lines(x = c(0,1), y = c(0,1), col ="red")
}

caliPIP_plot <- function(tags, tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  mrashf <- lapply(f, '[[', "gmrash")
  names(mrashf) <- tags

  susief <- lapply(f, '[[', "gsusie")
  names(susief) <- tags

  .tagname <- function(x, flist){
    a <- read.table(flist[[x]], header =T)
    a[, "name"] <- paste0(x, ":", a[, "name"])
    a
  }
  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))

  res <- merge(mrashres, susieres, by = "name", all = T)

  res <- res[complete.cases(res),]
  res <- rename(res, c("PIP" = "mr.ash_PIP", "pip" = "SUSIE_PIP", "pip.null" = "SUSIE_PIP_null", "pip.w0" = "SUSIE_PIP_w0"))
  par(mfrow=c(1,4), mar=c(5, 6, 4, 1))
  cp_plot(res$mr.ash_PIP, res$ifcausal, main = "Mr.ash PIP")
  cp_plot(res$SUSIE_PIP, res$ifcausal, main = "SUSIE PIP")
  cp_plot(res$SUSIE_PIP_null, res$ifcausal, main = "SUSIE PIP null")
  cp_plot(res$SUSIE_PIP_w0, res$ifcausal, main = "SUSIE PIP weighted null")
}
```

## NULL; expr-snp; expr-snp 
```{r cali 1, fig.height=4, fig.width=14}
res <- caliPIP_plot(tags = tags_ext2, tag2 = "zeroes-es")
```

## Lasso; expr-snp; expr-snp 
```{r cali 2, fig.height=4, fig.width=14}
caliPIP_plot(tags = tags_ext, tag2 = "lassoes-es")
```

# PIP scatter plot
mr.ash2s PIP vs. susie PIP.
```{r scatterplot}
scatter_plot_PIP<- function(tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  mrashf <- lapply(f, '[[', "gmrash")
  names(mrashf) <- tags

  susief <- lapply(f, '[[', "gsusie")
  names(susief) <- tags

  .tagname <- function(x, flist){
    a <- read.table(flist[[x]], header =T)
    a[, "name"] <- paste0(x, ":", a[, "name"])
    a
  }
  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))

  res <- merge(mrashres, susieres, by = "name", all = T)

  res <- res[complete.cases(res),]
  res <- rename(res, c("PIP" = "mr.ash_PIP", "pip" = "SUSIE_PIP", "pip.null" = "SUSIE_PIP_null", "pip.w0" = "SUSIE_PIP_w0") )
  res$ifcausal <- mapvalues(res$ifcausal, 
          from=c(0,1), 
          to=c("Non causal", "Causal"))

  fig1 <- plot_ly(data = res, x = ~ mr.ash_PIP, y = ~ SUSIE_PIP, color = ~ ifcausal, 
                 colors = c( "salmon", "darkgreen"))

  fig2 <- plot_ly(data = res, x = ~ mr.ash_PIP, y = ~ SUSIE_PIP_null, color = ~ ifcausal, 
                 colors = c( "salmon", "darkgreen"))

  fig3 <- plot_ly(data = res, x = ~ mr.ash_PIP, y = ~ SUSIE_PIP_w0, color = ~ ifcausal, 
                 colors = c( "salmon", "darkgreen"))

  fig <- subplot(fig1, fig2, fig3, titleX = TRUE, titleY = T, margin = 0.05)
  fig
}
```
## NULL; expr-snp; expr-snp 
```{r sca 1, fig.height=3, fig.width=10}
scatter_plot_PIP(tag2s[1])
```
## NULL; snp-expr; expr-snp

```{r sca 2, fig.height=3, fig.width=10}
scatter_plot_PIP(tag2s[2])
```
## lasso; expr-snp; expr-snp

```{r sca 3, fig.height=3, fig.width=10}
scatter_plot_PIP(tag2s[3])
```

## lasso; expr-snp; snp-expr

```{r sca 4, fig.height=3, fig.width=10}
scatter_plot_PIP(tag2s[4])
```

# ROC curve
```{r roc}
ROC_plot<- function(tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  mrashf <- lapply(f, '[[', "gmrash")
  names(mrashf) <- tags

  susief <- lapply(f, '[[', "gsusie")
  names(susief) <- tags

  gwasf <- lapply(f, '[[', "ggwas")
  names(gwasf) <- tags

  .tagname <- function(x, flist, colnames = NULL){
    a <- read.table(flist[[x]], header =T)
    if (!is.null(colnames)){
      colnames(a) <- colnames
    }
    a[, "name"] <- paste0(x, ":", a[, "name"])
    a
  }
  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))
  gwasres <- do.call(rbind, lapply(tags, .tagname, flist = gwasf, 
                                   colnames =  c("chr",	"p0",	"p1", "name", "Estimate", "Std.Error", "t-value", "PVALUE")))

  res <- merge(mrashres, susieres, by = "name", all = T)
  res <- merge(res, gwasres, by = "name", all = T)

  res <- res[complete.cases(res),]
  res <- rename(res, c("PIP" = "mr.ash", "pip" = "SUSIE", "pip.null"= "SUSIE.null", "pip.w0" = "SUSIE.w0", "PVALUE" = "TWAS") )
  res[,"TWAS"] <- -log10(res[, "TWAS"])

  roccolors <-  c("red", "green", "orange", "pink", "blue")
  methods <- c("mr.ash", "SUSIE", "SUSIE.null", "SUSIE.w0", "TWAS")
  plot(0, xlim=c(0,1), ylim=c(0,1), col="white", xlab = "FPR", ylab = "TPR")
  for (i in 1:length(methods)){
    method <- methods[i]
    bordered <- res[order(res[,method]),] 
    actuals <- bordered$ifcausal == 1
    sens <- (sum(actuals) - cumsum(actuals))/sum(actuals)
    spec <- cumsum(!actuals)/sum(!actuals)
    lines(1 - spec, sens, type = "l", col = roccolors[i])
    abline(c(0,0),c(1,1))
    auc <- sum(spec*diff(c(0, 1 - sens)))
    cat("AUC for ", method, ": ", auc)
  }
  legend(0.6,0.3, legend= methods, col=roccolors, lty=1, cex=0.5 )
  grid()
}
```

## NULL; expr-snp; expr-snp 
```{r roc 1, fig.height=4, fig.width=4}
ROC_plot(tag2s[1])
```

## NULL; snp-expr; expr-snp

```{r roc 2, fig.height=4, fig.width=4}
ROC_plot(tag2s[2])
```

## lasso; expr-snp; expr-snp

```{r roc 3, fig.height=4, fig.width=4}
ROC_plot(tag2s[3])
```

## lasso; expr-snp; snp-expr

```{r roc 4, fig.height=4, fig.width=4}
ROC_plot(tag2s[4])
```

# Mr.ash2 parameter estimation, n=80k

## lasso; expr-snp; expr-snp

```{r param80k 3}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200730/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200730/"
tags <- paste0('20200730-1-', 1:3)
show_param(tag2s[3])
```

### 2020-07-31 code/gwas.R

@@ -31,7 +31,7 @@ GWAA <- function(geno, pheno, snpname = NULL, anno = NULL, outname, family = gau
   columns<-c("#CHROM",	"BEGIN",	"END", "MARKER_ID", "Estimate", "Std.Error", "t-value", "PVALUE") # header for LOCUSZOOM.
   write.table(t(columns), outname, row.names=FALSE, col.names=FALSE, quote=FALSE, sep = "\t")
 
-  foreach (part=1:nSplits) %do% {
+  for (part in 1:nSplits) {
 
     geno.i <- geno[ ,snp.start[part]:snp.stop[part]]
     snpname.i <- snpname[snp.start[part]:snp.stop[part]]
@@ -55,6 +55,7 @@ GWAA <- function(geno, pheno, snpname = NULL, anno = NULL, outname, family = gau
                 sep = "\t")
 
     cat(sprintf("GWAS SNPs %s%% finished\n", 100 * part/nSplits))
+    rm(geno.i, anno.i, snpname.i);gc()
   }
 
   stopCluster(cl)

### 2020-07-31 code/run_UKB_process.R

@@ -23,7 +23,7 @@ for (wgtf in wgtfs){
 
 write.table(snpannoall, file= paste(weight, method, "eqtl.txt", sep= ".") , row.names=F, col.names=T, sep="\t", quote = F)
 
-pfiles <- paste0("/home/simingz/causalTWAS/ukbiobank/ukb_chr", 17:22, "_s20000.pgen")
+pfiles <- paste0("/home/simingz/causalTWAS/ukbiobank/ukb_chr", 17:22, "_s80000.pgen")
 
 mtxall <- NULL
 varall <- NULL
@@ -46,11 +46,11 @@ dat <- list("G"       = mtxall,
             "snp"     = as.matrix(varall[,3]),
             "counted" = as.matrix(varall[,4]),
             "alt"     = as.matrix(varall[,5]))
-save(dat, file = "/home/simingz/causalTWAS/ukbiobank/ukb_chr17:22_s20000.Rd")
+save(dat, file = "/home/simingz/causalTWAS/ukbiobank/ukb_chr17to22_s80000.Rd")
 
-m <- as_FBM(mtxall, backingfile = "/home/simingz/causalTWAS/ukbiobank/ukb_chr17:22_s20000")$save()
+m <- as_FBM(mtxall, backingfile = "/home/simingz/causalTWAS/ukbiobank/ukb_chr17to22_s80000")$save()
 dat$G <- m
-save(dat, file = "/home/simingz/causalTWAS/ukbiobank/ukb_chr17:22_s20000.FBM.Rd")
+save(dat, file = "/home/simingz/causalTWAS/ukbiobank/ukb_chr17to22_s80000.FBM.Rd")
 
 #----------------------------chr22 only-------------------
 pfile <- "/home/simingz/causalTWAS/ukbiobank/ukb_chr22_s20000"

### 2020-07-31 code/run_gwas_snp.R

@@ -29,7 +29,7 @@ if (!file.exists(pfileRd)) {
 
 load(pfileRd)
 
-dat$G <- dat$G[] # TODO: revise other functions in this script to take FBM as input.
+dat$G <- dat$G[]
 
 snpname = dat$snp[,1]
 anno <- cbind(dat$chr, dat$pos, dat$pos) # EPACT format
@@ -38,7 +38,7 @@ if (length(args) == 3){
 
   geno <- dat$G
 
-  GWAA(geno, pheno, snpname = snpname, anno = anno, outname, family = gaussian, ncore = 3, nSplits = 30, compress = T)
+  GWAA(geno, pheno, snpname = snpname, anno = anno, outname, family = gaussian, ncore = 3, nSplits = 100, compress = T)
 
 } else {
   regions <- read.table(args[4], stringsAsFactors = F)

### 2020-07-31 code/run_test_mr.ash2s.R

@@ -44,7 +44,7 @@ phenores$Y <-  phenores$Y - mean(phenores$Y)
 outname <- args[4]
 
 # run mr.ash2s (a simplified version of veb_boost)
-## run mr.ash2s
+# run mr.ash2s
 loginfo("mr.ash2s initiation (beta: NULL, update order: expr-snp)")
 mr.ash2s.fit <- mr.ash2s(snp = dat$G,
                          expr = dat$expr,

### 2020-08-04 analysis/index.Rmd
change sa2 grid

--- 2020-06-25 analysis/index.Rmd
@@
 
+## Summarize cTWAS results
+We compare cTWAS results with other methods, summarizing results over multiple simulation runs.
+
+  * [Using UKbiobank data (chr22), fusion weights Adipose](simulation-multi-ukbchr22-gtex.adipose.html)
+  * [Using UKbiobank data (chr22), fusion weights Adipose, setting 2](simulation-multi-ukbchr22-gtex.adipose2.html)
+  * [Using UKbiobank data (chr17:22), fusion weights Adipose](simulation-multi-ukbchr17to22-gtex.adipose.html) 
+  * [Using UKbiobank data (chr17:22), fusion weights Adipose, using smaller sa2 grid in mr.ash](simulation-multi-ukbchr17to22-gtex.adipose_sa2_v2.html)
+
 # Related software

### 2020-08-04 analysis/simulation-multi-ukbchr17to22-gtex.adipose_sa2_v2.Rmd
change sa2 grid

---
title: "Summarize twas"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  message= F)
```

Run simulation 8 times for ukb chr 17 to chr 22 combined. SNPs are downsampled to 1/10, eQTLs defined by FUSION-TWAS (Adipose/GTEx) lasso effect size > 0 were kept, p= 86k, n = 20k.

```{r plot functions}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})
library(tidyr)
library(plyr)
library(stringr)
library(kableExtra)
```
```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
tags <- paste0('20200721-1-', 1:7)
tagglob <- '20200721-1-*'
tagextr <- '20200721-1-\\d+'
tag2s <- c('zeroes-es', 'zerose-es', 'lassoes-es','lassoes-se')
```

```{r getfile function}

get_files <- function(tag, tag2){
  par <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt")
  rpip <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".rPIP.txt")

  gmrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt")
  smrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt")

  ggwas <- paste0(outputdir, tag, ".exprgwas.txt.gz")
  sgwas <- paste0(outputdir, tag, ".snpgwas.txt.gz")

  gsusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.expr.txt")
  ssusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.snp.txt")

  return(tibble::lst(par, rpip, gmrash, ggwas, smrash, sgwas, gsusie, ssusie))
}

get_tags <- function(globpattern, extrpattern, tag2){
  lapply(lapply(get_files(globpattern, tag2), Sys.glob), str_extract, pattern = extrpattern)
}
```
# Mr.ash2 parameter estimation

Results for 10 simulations runs, using different initiate and update strategy

```{r param function}
show_param <- function(tags, tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  parf <- lapply(f, '[[', "par")
  param <- do.call(rbind, lapply(parf, function(x) t(read.table(x))[2:1,]))
  truth <- param[1:(nrow(param)/2)*2-1,]
  est <- param[1:(nrow(param)/2)*2,]
  outdt <- matrix(0, ncol = 2*ncol(param), nrow = nrow(param)/2)
  outdt[,c(1,3,5,7)] <- truth
  outdt[,c(2,4,6,8)] <- est
  outdt <-cbind(1:nrow(outdt),outdt)
  colnames(outdt) <- c("Simulation#", paste0(rep(c("Truth","Est."),4)))
  knitr::kable(outdt) %>%
  kable_styling("striped") %>%
  add_header_above(c(" " = 1, "Gene.pi1" = 2, "Gene.PVE" = 2, "SNP.pi1" = 2, "SNP.PVE" =2))
}
```

## NULL; expr-snp; expr-snp

```{r param 1}
show_param(tags, tag2s[1])
```

## NULL; snp-expr; expr-snp

```{r param 2}
show_param(tags, tag2s[2])
```

## lasso; expr-snp; expr-snp

```{r param 3}
show_param(tags, tag2s[3])
```

## lasso; expr-snp; snp-expr

```{r param 4}
show_param(tags, tag2s[4])
```

# Regional mr.ash2s PIP overview
Take simulation 1 (NULL; expr-snp; expr-snp) as examples. We use region size *500kb* and PIP cut off at *0.5* for SUSIE.

```{r rpip 1, fig.height=4, fig.width=10}
chrom = 18
f <- get_files(tag= "20200721-1-2" , tag2 = tag2s[1])
allchr <- read.table(f[["rpip"]], header = T)
a <- allchr[allchr["chrom"]==chrom,]
print(paste("plot for chr", chrom))
par(mar=c(5, 4, 4, 6) + 0.1)
with(a, plot(p0, rPIP, col ='salmon', xlab = "position", ylab= "Sum of PIP", type = 'h', lwd = 2))
par(new = T)
with(a, plot(p0, nCausal, pch =19, col = "darkgreen",axes = FALSE, bty = "n", xlab = "", ylab = ""))
axis(side = 4)
mtext(side = 4, line = 3, 'No. causal signals')
legend("topleft",
       legend=c("Mr.ASH PIP", "# Causal"),
       lty=c(1,0), pch=c(NA, 19), col=c("salmon", "darkgreen"))
grid()
```

# PIP calibration
We run 50 simulations and combine results.
```{r caliplot}
#' s is pip or fdr.
cp_plot <- function(s, ifcausal, mode = c("PIP", "FDR"), main = mode[1]){
  # ifcausal:0,1
  a_bin <- cut(s, breaks= seq(0,1, by=0.1))
  if (mode == "PIP") {
     expected = c(by(s, a_bin, FUN = mean))
     observed = c(by(ifcausal, a_bin, FUN = mean))
  } else if (mode == "FDR"){
     expected = c(by(s, a_bin, FUN = mean))
     observed = 1 - c(by(ifcausal, a_bin, FUN = mean))
  }
  plot(expected, observed, xlim= c(0,1), ylim=c(0,1), pch =19, main = main)
  lines(x = c(0,1), y = c(0,1), col ="red")
}

caliPIP_plot <- function(tags, tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  mrashf <- lapply(f, '[[', "gmrash")
  names(mrashf) <- tags

  susief <- lapply(f, '[[', "gsusie")
  names(susief) <- tags

  .tagname <- function(x, flist){
    a <- read.table(flist[[x]], header =T)
    a[, "name"] <- paste0(x, ":", a[, "name"])
    a
  }
  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))

  res <- merge(mrashres, susieres, by = "name", all = T)

  res <- res[complete.cases(res),]
  res <- rename(res, c("PIP" = "mr.ash_PIP", "pip" = "SUSIE_PIP", "pip.null" = "SUSIE_PIP_null", "pip.w0" = "SUSIE_PIP_w0"))
  par(mfrow=c(1,4), mar=c(5, 6, 4, 1))
  cp_plot(res$mr.ash_PIP, res$ifcausal, main = "Mr.ash PIP")
  cp_plot(res$SUSIE_PIP, res$ifcausal, main = "SUSIE PIP")
  cp_plot(res$SUSIE_PIP_null, res$ifcausal, main = "SUSIE PIP null")
  cp_plot(res$SUSIE_PIP_w0, res$ifcausal, main = "SUSIE PIP weighted null")
}

caliFDR_plot <- function(tags, tag2){

  f <- lapply(tags, get_files, tag2 = tag2)
  gwasf <- lapply(f, '[[', "ggwas")
  names(gwasf) <- tags

  .tagname <- function(x, flist, colnames = NULL){
    a <- read.table(flist[[x]], header =T)
    if (!is.null(colnames)){
      colnames(a) <- colnames
    }
    a[, "name"] <- paste0(x, ":", a[, "name"])
    a
  }

  .addFDR <- function(res){
    res$FDR <- p.adjust(res$PVALUE, method = "fdr")
    res
  }

  gwasres <- do.call(rbind, lapply(lapply(tags, .tagname, flist = gwasf, 
                                   colnames =  c("chr",	"p0",	"p1", "name", 
                                                 "Estimate", "Std.Error", "t-value", "PVALUE")), .addFDR))

  f <- lapply(tags, get_files, tag2 = tag2)
  mrashf <- lapply(f, '[[', "gmrash")
  names(mrashf) <- tags

  susief <- lapply(f, '[[', "gsusie")
  names(susief) <- tags

  .tagname <- function(x, flist){
    a <- read.table(flist[[x]], header =T)
    a[, "name"] <- paste0(x, ":", a[, "name"])
    a
  }
  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))

  res <- merge(mrashres, susieres, by = "name", all = T)

  res <- merge(res, gwasres, by = "name", all = T)

  res <- res[complete.cases(res),]

  cp_plot(res$FDR, res$ifcausal, mode ="FDR", main = "TWAS FDR")
  cat("FDR at bonferroni corrected p = 0.05: ", 1 - mean(res[res$PVALUE < 0.05 /dim(res)[1], "ifcausal"]))
}
```

## NULL; expr-snp; expr-snp 
```{r cali 1, fig.height=4, fig.width=14}
tag2 = "zeroes-es"
tags_ext <- Reduce(intersect, get_tags(tagglob, tagextr, tag2 = tag2)['gsusie'])
res <- caliPIP_plot(tags = tags_ext, tag2 = tag2)
```
```{r cali 1fdr, fig.height=4, fig.width=4}
caliFDR_plot(tags = tags_ext, tag2 = tag2)
```

## Lasso; expr-snp; expr-snp 
```{r cali 2, fig.height=4, fig.width=14}
caliPIP_plot(tags = tags_ext, tag2 = "lassoes-es")
```

```{r cali 2fdr, fig.height=4, fig.width=4}
caliFDR_plot(tags = tags_ext, tag2 = "lassoes-es")
```

# PIP scatter plot
mr.ash2s PIP vs. susie PIP.
```{r scatterplot}
scatter_plot_PIP<- function(tags, tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  mrashf <- lapply(f, '[[', "gmrash")
  names(mrashf) <- tags

  susief <- lapply(f, '[[', "gsusie")
  names(susief) <- tags

  .tagname <- function(x, flist){
    a <- read.table(flist[[x]], header =T)
    a[, "name"] <- paste0(x, ":", a[, "name"])
    a
  }
  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))

  res <- merge(mrashres, susieres, by = "name", all = T)

  res <- res[complete.cases(res),]
  res <- rename(res, c("PIP" = "mr.ash_PIP", "pip" = "SUSIE_PIP", "pip.null" = "SUSIE_PIP_null", "pip.w0" = "SUSIE_PIP_w0") )
  res$ifcausal <- mapvalues(res$ifcausal, 
          from=c(0,1), 
          to=c("Non causal", "Causal"))

  fig1 <- plot_ly(data = res, x = ~ mr.ash_PIP, y = ~ SUSIE_PIP, color = ~ ifcausal, 
                 colors = c( "salmon", "darkgreen"))

  fig2 <- plot_ly(data = res, x = ~ mr.ash_PIP, y = ~ SUSIE_PIP_null, color = ~ ifcausal, 
                 colors = c( "salmon", "darkgreen"))

  fig3 <- plot_ly(data = res, x = ~ mr.ash_PIP, y = ~ SUSIE_PIP_w0, color = ~ ifcausal, 
                 colors = c( "salmon", "darkgreen"))

  fig <- subplot(fig1, fig2, fig3, titleX = TRUE, titleY = T, margin = 0.05)
  fig
}
```
## NULL; expr-snp; expr-snp 
```{r sca 1, fig.height=3, fig.width=10}
scatter_plot_PIP(tags, tag2s[1])
```
## NULL; snp-expr; expr-snp

```{r sca 2, fig.height=3, fig.width=10}
scatter_plot_PIP(tags, tag2s[2])
```
## lasso; expr-snp; expr-snp

```{r sca 3, fig.height=3, fig.width=10}
scatter_plot_PIP(tags, tag2s[3])
```

## lasso; expr-snp; snp-expr

```{r sca 4, fig.height=3, fig.width=10}
scatter_plot_PIP(tags, tag2s[4])
```

# ROC curve
```{r roc}
ROC_plot<- function(tags, tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  mrashf <- lapply(f, '[[', "gmrash")
  names(mrashf) <- tags

  susief <- lapply(f, '[[', "gsusie")
  names(susief) <- tags

  gwasf <- lapply(f, '[[', "ggwas")
  names(gwasf) <- tags

  .tagname <- function(x, flist, colnames = NULL){
    a <- read.table(flist[[x]], header =T)
    if (!is.null(colnames)){
      colnames(a) <- colnames
    }
    a[, "name"] <- paste0(x, ":", a[, "name"])
    a
  }
  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))
  gwasres <- do.call(rbind, lapply(tags, .tagname, flist = gwasf, 
                                   colnames =  c("chr",	"p0",	"p1", "name", "Estimate", "Std.Error", "t-value", "PVALUE")))

  res <- merge(mrashres, susieres, by = "name", all = T)
  res <- merge(res, gwasres, by = "name", all = T)

  res <- res[complete.cases(res),]
  res <- rename(res, c("PIP" = "mr.ash", "pip" = "SUSIE", "pip.null"= "SUSIE.null", "pip.w0" = "SUSIE.w0", "PVALUE" = "TWAS") )
  res[,"TWAS"] <- -log10(res[, "TWAS"])

  roccolors <-  c("red", "green", "orange", "pink", "blue")
  methods <- c("mr.ash", "SUSIE", "SUSIE.null", "SUSIE.w0", "TWAS")
  plot(0, xlim=c(0,1), ylim=c(0,1), col="white", xlab = "FPR", ylab = "TPR")
  for (i in 1:length(methods)){
    method <- methods[i]
    bordered <- res[order(res[,method]),] 
    actuals <- bordered$ifcausal == 1
    sens <- (sum(actuals) - cumsum(actuals))/sum(actuals)
    spec <- cumsum(!actuals)/sum(!actuals)
    lines(1 - spec, sens, type = "l", col = roccolors[i])
    abline(c(0,0),c(1,1))
    auc <- sum(spec*diff(c(0, 1 - sens)))
    cat("AUC for ", method, ": ", auc)
  }
  legend(0.6,0.3, legend= methods, col=roccolors, lty=1, cex=0.5 )
  grid()
}
```

## NULL; expr-snp; expr-snp 
```{r roc 1, fig.height=4, fig.width=4}
tags <- paste0('20200721-1-', c(2,4:9))
ROC_plot(tags, tag2s[2])
```

## NULL; snp-expr; expr-snp

```{r roc 2, fig.height=4, fig.width=4}
ROC_plot(tags, tag2s[2])
```

## lasso; expr-snp; expr-snp

```{r roc 3, fig.height=4, fig.width=4}
ROC_plot(tags, tag2s[3])
```

## lasso; expr-snp; snp-expr

```{r roc 4, fig.height=4, fig.width=4}
ROC_plot(tags, tag2s[4])
```

### 2020-08-04 analysis/simulation-multi-ukbchr22-gtex.adipose2.Rmd
change sa2 grid

@@
 library(plyr)
+library(stringr)
+library(kableExtra)
 ```
@@
 tags <- paste0('20200616-8-', 1:9)
+tagglob <- '20200616-8-*'
+tagextr <- '20200616-8-\\d+'
 tag2s <- c('zeroes-es', 'zerose-es', 'lassoes-es','lassoes-se')
@@
 }
+
+get_tags <- function(globpattern, extrpattern, tag2){
+  lapply(lapply(get_files(globpattern, tag2), Sys.glob), str_extract, pattern = extrpattern)
+}
 ```
@@
 ```{r param function}
-show_param <- function(tag2){
+show_param <- function(tags, tag2){
   f <- lapply(tags, get_files, tag2 = tag2)
@@
   param <- do.call(rbind, lapply(parf, function(x) t(read.table(x))[2:1,]))
-  knitr::kable(param)
+  truth <- param[1:(nrow(param)/2)*2-1,]
+  est <- param[1:(nrow(param)/2)*2,]
+  outdt <- matrix(0, ncol = 2*ncol(param), nrow = nrow(param)/2)
+  outdt[,c(1,3,5,7)] <- truth
+  outdt[,c(2,4,6,8)] <- est
+  outdt <-cbind(1:nrow(outdt),outdt)
+  colnames(outdt) <- c("Simulation#", paste0(rep(c("Truth","Est."),4)))
+  knitr::kable(outdt) %>%
+  kable_styling("striped") %>%
+  add_header_above(c(" " = 1, "Gene.pi1" = 2, "Gene.PVE" = 2, "SNP.pi1" = 2, "SNP.PVE" =2))
 }
@@
 ```{r param 1}
-show_param(tag2s[1])
+show_param(tags = tags, tag2 = tag2s[1])
 ```
@@
 ```{r param 2}
-show_param(tag2s[2])
+show_param(tags = tags, tag2 = tag2s[2])
 ```
@@
 ```{r param 3}
-show_param(tag2s[3])
+show_param(tags = tags, tag2 = tag2s[3])
 ```
@@
 ```{r param 4}
-show_param(tag2s[4])
+show_param(tags = tags, tag2 = tag2s[4])
 ```
@@
 a <- read.table(f[["rpip"]], header = T)
-plot(a$p0, a$rPIP, pch =19, col ='salmon', xlab = "position", ylab= "Sum of PIP")
+par(mar=c(5, 4, 4, 6) + 0.1)
+with(a, plot(p0, rPIP, col ='salmon', xlab = "position", ylab= "Sum of PIP", type = 'h', lwd = 2))
+par(new = T)
+with(a, plot(p0, nCausal, pch =19, col = "darkgreen",axes = FALSE, bty = "n", xlab = "", ylab = ""))
+axis(side = 4)
+mtext(side = 4, line = 3, 'No. causal signals')
+legend("topleft",
+       legend=c("Mr.ASH PIP", "# Causal"),
+       lty=c(1,0), pch=c(NA, 19), col=c("salmon", "darkgreen"))
 grid()
+```
+
+# PIP calibration
+We run 100 simulations and combine results.
+```{r caliplot}
+#' s is pip or fdr.
+cp_plot <- function(s, ifcausal, mode = c("PIP", "FDR"), main = mode[1]){
+  # ifcausal:0,1
+  a_bin <- cut(s, breaks= seq(0,1, by=0.1))
+  if (mode == "PIP") {
+     expected = c(by(s, a_bin, FUN = mean))
+     observed = c(by(ifcausal, a_bin, FUN = mean))
+  } else if (mode == "FDR"){
+     expected = c(by(s, a_bin, FUN = mean))
+     observed = 1 - c(by(ifcausal, a_bin, FUN = mean))
+  }
+  plot(expected, observed, xlim= c(0,1), ylim=c(0,1), pch =19, main = main)
+  lines(x = c(0,1), y = c(0,1), col ="red")
+}
+
+caliPIP_plot <- function(tags, tag2){
+  f <- lapply(tags, get_files, tag2 = tag2)
+  mrashf <- lapply(f, '[[', "gmrash")
+  names(mrashf) <- tags
+
+  susief <- lapply(f, '[[', "gsusie")
+  names(susief) <- tags
+
+  .tagname <- function(x, flist){
+    a <- read.table(flist[[x]], header =T)
+    a[, "name"] <- paste0(x, ":", a[, "name"])
+    a
+  }
+  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
+  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))
+
+  res <- merge(mrashres, susieres, by = "name", all = T)
+
+  res <- res[complete.cases(res),]
+  res <- rename(res, c("PIP" = "mr.ash_PIP", "pip" = "SUSIE_PIP", "pip.null" = "SUSIE_PIP_null", "pip.w0" = "SUSIE_PIP_w0"))
+  par(mfrow=c(1,4), mar=c(5, 6, 4, 1))
+  cp_plot(res$mr.ash_PIP, res$ifcausal, main = "Mr.ash PIP")
+  cp_plot(res$SUSIE_PIP, res$ifcausal, main = "SUSIE PIP")
+  cp_plot(res$SUSIE_PIP_null, res$ifcausal, main = "SUSIE PIP null")
+  cp_plot(res$SUSIE_PIP_w0, res$ifcausal, main = "SUSIE PIP weighted null")
+}
+
+caliFDR_plot <- function(tags, tag2){
+
+  f <- lapply(tags, get_files, tag2 = tag2)
+  gwasf <- lapply(f, '[[', "ggwas")
+  names(gwasf) <- tags
+
+  .tagname <- function(x, flist, colnames = NULL){
+    a <- read.table(flist[[x]], header =T)
+    if (!is.null(colnames)){
+      colnames(a) <- colnames
+    }
+    a[, "name"] <- paste0(x, ":", a[, "name"])
+    a
+  }
+
+  .addFDR <- function(res){
+    res$FDR <- p.adjust(res$PVALUE, method = "fdr")
+    res
+  }
+
+  gwasres <- do.call(rbind, lapply(lapply(tags, .tagname, flist = gwasf, 
+                                   colnames =  c("chr",	"p0",	"p1", "name", 
+                                                 "Estimate", "Std.Error", "t-value", "PVALUE")), .addFDR))
+
+  f <- lapply(tags, get_files, tag2 = tag2)
+  mrashf <- lapply(f, '[[', "gmrash")
+  names(mrashf) <- tags
+
+  susief <- lapply(f, '[[', "gsusie")
+  names(susief) <- tags
+
+  .tagname <- function(x, flist){
+    a <- read.table(flist[[x]], header =T)
+    a[, "name"] <- paste0(x, ":", a[, "name"])
+    a
+  }
+  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
+  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))
+
+  res <- merge(mrashres, susieres, by = "name", all = T)
+
+  res <- merge(res, gwasres, by = "name", all = T)
+
+  res <- res[complete.cases(res),]
+
+  cp_plot(res$FDR, res$ifcausal, mode ="FDR", main = "TWAS FDR")
+  cat("FDR at bonferroni corrected p = 0.05: ", 1 - mean(res[gwasres$PVALUE < 0.05 /dim(res)[1], "ifcausal"]))
+}
+```
+
+## NULL; expr-snp; expr-snp 
+```{r cali 1, fig.height=4, fig.width=14}
+tag2 = "zeroes-es"
+tags_ext <- Reduce(intersect, get_tags(tagglob, tagextr, tag2 = tag2)['gsusie'])
+res <- caliPIP_plot(tags = tags_ext, tag2 = tag2)
+```
+```{r cali 1fdr, fig.height=4, fig.width=4}
+caliFDR_plot(tags = tags_ext, tag2 = tag2)
+```
+
+## Lasso; expr-snp; expr-snp 
+```{r cali 2, fig.height=4, fig.width=14}
+caliPIP_plot(tags = tags_ext, tag2 = "lassoes-es")
+```
+
+```{r cali 2fdr, fig.height=4, fig.width=4}
+caliFDR_plot(tags = tags_ext, tag2 = "lassoes-es")
 ```
@@
 ```{r scatterplot}
-scatter_plot_PIP<- function(tag2){
+scatter_plot_PIP<- function(tags, tag2){
   f <- lapply(tags, get_files, tag2 = tag2)
@@
 ```{r sca 1, fig.height=4, fig.width=10}
-scatter_plot_PIP(tag2s[1])
+scatter_plot_PIP(tags = tags, tag2 = tag2s[1])
 ```
@@
 ```{r sca 2, fig.height=4, fig.width=10}
-scatter_plot_PIP(tag2s[2])
+scatter_plot_PIP(tags = tags, tag2 = tag2s[2])
 ```
@@
 ```{r sca 3, fig.height=4, fig.width=10}
-scatter_plot_PIP(tag2s[3])
+scatter_plot_PIP(tags = tags, tag2 = tag2s[3])
 ```
@@
 ```{r sca 4, fig.height=4, fig.width=10}
-scatter_plot_PIP(tag2s[4])
+scatter_plot_PIP(tags = tags, tag2 = tag2s[4])
 ```
@@
 ```{r roc}
-ROC_plot<- function(tag2){
+ROC_plot<- function(tags, tag2){
   f <- lapply(tags, get_files, tag2 = tag2)
@@
 ```{r roc 1, fig.height=4, fig.width=4}
-ROC_plot(tag2s[1])
+ROC_plot(tags = tags, tag2 = tag2s[1])
 ```
@@
 ```{r roc 2, fig.height=4, fig.width=4}
-ROC_plot(tag2s[2])
+ROC_plot(tags = tags, tag2 = tag2s[2])
 ```
@@
 ```{r roc 3, fig.height=4, fig.width=4}
-ROC_plot(tag2s[3])
+ROC_plot(tags = tags, tag2 = tag2s[3])
 ```
@@
 ```{r roc 4, fig.height=4, fig.width=4}
-ROC_plot(tag2s[4])
+ROC_plot(tags = tags, tag2 = tag2s[4])
 ```

### 2020-08-04 code/run_test_mr.ash2s_temp.R

@@ -45,12 +45,15 @@ outname <- args[4]
 
 # run mr.ash2s (a simplified version of veb_boost)
 # run mr.ash2s
+Sa2 = (2^(c(seq(0,20,2)/200, seq(2,19,2)/20)) - 1)^2
+
+save.image("temp.Rd")
 loginfo("mr.ash2s initiation (beta: NULL, update order: expr-snp)")
 mr.ash2s.fit <- mr.ash2s(snp = dat$G,
                          expr = dat$expr,
                          y= phenores$Y,
                          mr.ash.init = NULL,
-                         init.order = "es", iter = 30, outname = outname, ncores = 5)
+                         init.order = "es", iter = 30, outname = outname, ncores = 5, sa2 = Sa2)
 mr.ash2s_save(mr.ash2s.fit, outname)
 rm(mr.ash2s.fit); gc()
 
@@ -59,7 +62,7 @@ rm(mr.ash2s.fit); gc()
 #                          expr = dat$expr,
 #                          y= phenores$Y,
 #                          mr.ash.init = NULL,
-#                          init.order = "se", iter = 30, outname = outname, ncores = 5)
+#                          init.order = "se", iter = 30, outname = outname, ncores = 5, sa2 = Sa2)
 # mr.ash2s_save(mr.ash2s.fit, outname)
 # rm(mr.ash2s.fit); gc()
 
@@ -69,7 +72,7 @@ mr.ash2s.fit <- mr.ash2s(snp = dat$G,
                          expr = dat$expr,
                          y= phenores$Y,
                          mr.ash.init = "lasso",
-                         init.order = "es", iter = 30, outname = outname, ncores = 5)
+                         init.order = "es", iter = 30, outname = outname, ncores = 5, sa2 = Sa2)
 mr.ash2s_save(mr.ash2s.fit, outname)
 rm(mr.ash2s.fit); gc()
 
@@ -78,7 +81,7 @@ rm(mr.ash2s.fit); gc()
 #                          expr = dat$expr,
 #                          y= phenores$Y,
 #                          mr.ash.init = "lasso",
-#                          init.order = "es", iter.order = "se", iter = 30, outname = outname, ncores = 5)
+#                          init.order = "es", iter.order = "se", iter = 30, outname = outname, ncores = 5, sa2 = Sa2)
 # mr.ash2s_save(mr.ash2s.fit, outname)
 # rm(mr.ash2s.fit); gc()
 #

### 2020-08-04 analysis/simulation-multi-ukbchr17to22-gtex.adipose.Rmd
website hide code

@@
 ---
-title: "Summarize twas"
+title: "Summarize twas: chr17 to 22"
 output: html_document
@@
 library(plyr)
+library(stringr)
 library(kableExtra)
@@
 simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
-outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
-susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
-tags <- paste0('20200721-1-', c(2,4:9))
-tags_ext <- paste0('20200721-1-', 1:50)
-tags_ext2 <- paste0('20200721-1-', c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 15, 16, 17, 19, 21, 22, 23, 26, 27, 28, 29, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50))
+outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/sa2_default/"
+susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/sa2_default/"
+tags <- paste0('20200721-1-', c(2, 4:9))
+tagglob <- '20200721-1-*'
+tagextr <- '20200721-1-\\d+'
 tag2s <- c('zeroes-es', 'zerose-es', 'lassoes-es','lassoes-se')
@@
 ```{r getfile function}
+
 get_files <- function(tag, tag2){
@@
 
-Results for 7 simulations runs, using different initiate and update strategy
+Results for 10 simulations runs, using different initiate and update strategy
 
 ```{r param function}
-show_param <- function(tag2){
+show_param <- function(tags, tag2){
   f <- lapply(tags, get_files, tag2 = tag2)
@@
 ```{r param 1}
-show_param(tag2s[1])
+show_param(tags, tag2s[1])
 ```
@@
 ```{r param 2}
-show_param(tag2s[2])
+show_param(tags, tag2s[2])
 ```
@@
 ```{r param 3}
-show_param(tag2s[3])
+show_param(tags, tag2s[3])
 ```
@@
 ```{r param 4}
-show_param(tag2s[4])
+show_param(tags, tag2s[4])
 ```
@@
 chrom = 18
-f <- get_files(tag= tags[1], tag2 = tag2s[1])
+f <- get_files(tag= "20200721-1-2" , tag2 = tag2s[1])
 allchr <- read.table(f[["rpip"]], header = T)
@@
 ```
-```{r rpip 2, fig.height=4, fig.width=4}
-plot( allchr$rPIP, allchr$nCausal, bg = rgb(red = 1, green = 0, blue = 0, alpha = 0.1), col= "red", pch = 21, cex =2,
-     ylab = "# Causal", xlab= "Sum of PIP")
-```
 
@@
 ```{r caliplot}
-cp_plot <- function(pip, ifcausal, main = "PIP"){
+#' s is pip or fdr.
+cp_plot <- function(s, ifcausal, mode = c("PIP", "FDR"), main = mode[1]){
   # ifcausal:0,1
-  a_bin <- cut(pip, breaks= seq(0,1, by=0.1))
-  PIP_mean = c(by(pip, a_bin, FUN = mean))
-  Freq = c(by(ifcausal, a_bin, FUN = mean))
-  plot(PIP_mean, Freq, xlim= c(0,1), ylim=c(0,1), pch =19, main = main)
+  a_bin <- cut(s, breaks= seq(0,1, by=0.1))
+  if (mode == "PIP") {
+     expected = c(by(s, a_bin, FUN = mean))
+     observed = c(by(ifcausal, a_bin, FUN = mean))
+  } else if (mode == "FDR"){
+     expected = c(by(s, a_bin, FUN = mean))
+     observed = 1 - c(by(ifcausal, a_bin, FUN = mean))
+  }
+  plot(expected, observed, xlim= c(0,1), ylim=c(0,1), pch =19, main = main)
   lines(x = c(0,1), y = c(0,1), col ="red")
@@
 }
+
+caliFDR_plot <- function(tags, tag2){
+
+  f <- lapply(tags, get_files, tag2 = tag2)
+  gwasf <- lapply(f, '[[', "ggwas")
+  names(gwasf) <- tags
+
+  .tagname <- function(x, flist, colnames = NULL){
+    a <- read.table(flist[[x]], header =T)
+    if (!is.null(colnames)){
+      colnames(a) <- colnames
+    }
+    a[, "name"] <- paste0(x, ":", a[, "name"])
+    a
+  }
+
+  .addFDR <- function(res){
+    res$FDR <- p.adjust(res$PVALUE, method = "fdr")
+    res
+  }
+
+  gwasres <- do.call(rbind, lapply(lapply(tags, .tagname, flist = gwasf, 
+                                   colnames =  c("chr",	"p0",	"p1", "name", 
+                                                 "Estimate", "Std.Error", "t-value", "PVALUE")), .addFDR))
+
+  f <- lapply(tags, get_files, tag2 = tag2)
+  mrashf <- lapply(f, '[[', "gmrash")
+  names(mrashf) <- tags
+
+  susief <- lapply(f, '[[', "gsusie")
+  names(susief) <- tags
+
+  .tagname <- function(x, flist){
+    a <- read.table(flist[[x]], header =T)
+    a[, "name"] <- paste0(x, ":", a[, "name"])
+    a
+  }
+  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
+  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))
+
+  res <- merge(mrashres, susieres, by = "name", all = T)
+
+  res <- merge(res, gwasres, by = "name", all = T)
+
+  res <- res[complete.cases(res),]
+
+  cp_plot(res$FDR, res$ifcausal, mode ="FDR", main = "TWAS FDR")
+  cat("FDR at Bonferroni corrected p = 0.05: ", 1 - mean(res[res$PVALUE < 0.05 /dim(res)[1], "ifcausal"]))
+}
 ```
@@
 ```{r cali 1, fig.height=4, fig.width=14}
-res <- caliPIP_plot(tags = tags_ext2, tag2 = "zeroes-es")
+tag2 = "zeroes-es"
+tags_ext <- Reduce(intersect, get_tags(tagglob, tagextr, tag2 = tag2)['gsusie'])
+res <- caliPIP_plot(tags = tags_ext, tag2 = tag2)
+```
+```{r cali 1fdr, fig.height=4, fig.width=4}
+tags_ext2 <- Reduce(intersect, get_tags(tagglob, tagextr, tag2 = tag2)['gsusie'])
+caliFDR_plot(tags = tags_ext2, tag2 = tag2)
 ```
@@
 
+```{r cali 2fdr, fig.height=4, fig.width=4}
+caliFDR_plot(tags = tags_ext2, tag2 = "lassoes-es")
+```
+
 # PIP scatter plot
@@
 ```{r sca 1, fig.height=3, fig.width=10}
-scatter_plot_PIP(tag2s[1])
+scatter_plot_PIP(tags, tag2s[1])
 ```
@@
 ```{r sca 2, fig.height=3, fig.width=10}
-scatter_plot_PIP(tag2s[2])
+scatter_plot_PIP(tags, tag2s[2])
 ```
@@
 ```{r sca 3, fig.height=3, fig.width=10}
-scatter_plot_PIP(tag2s[3])
+scatter_plot_PIP(tags, tag2s[3])
 ```
@@
 ```{r sca 4, fig.height=3, fig.width=10}
-scatter_plot_PIP(tag2s[4])
+scatter_plot_PIP(tags, tag2s[4])
 ```
@@
 ```{r roc 1, fig.height=4, fig.width=4}
-ROC_plot(tag2s[1])
+tags <- paste0('20200721-1-', c(2,4:9))
+ROC_plot(tags, tag2s[2])
 ```
@@
 ```{r roc 2, fig.height=4, fig.width=4}
-ROC_plot(tag2s[2])
+ROC_plot(tags, tag2s[2])
 ```
@@
 ```{r roc 3, fig.height=4, fig.width=4}
-ROC_plot(tag2s[3])
+ROC_plot(tags, tag2s[3])
 ```
@@
 ```{r roc 4, fig.height=4, fig.width=4}
-ROC_plot(tag2s[4])
+ROC_plot(tags, tag2s[4])
 ```
-
-# Mr.ash2 parameter estimation, n=80k
-
-## lasso; expr-snp; expr-snp
-
-```{r param80k 3}
-simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200730/"
-outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200730/"
-tags <- paste0('20200730-1-', 1:3)
-show_param(tag2s[3])
-```

### 2020-08-04 analysis/simulation-multi-ukbchr17to22-gtex.adipose_sa2_v2.Rmd
website hide code

@@
 ---
-title: "Summarize twas"
-output: html_document
+title: "Summarize twas: chr17 to 22"
+date: "`r Sys.Date()`"
+output: workflowr::wflow_html
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE,  message= F)
+knitr::opts_chunk$set(echo = TRUE,  message= F, warning = F)
 ```
+`
 

### 2020-08-04 analysis/simulation-multi-ukbchr22-gtex.adipose2.Rmd
website hide code

@@
 ---
-title: "Summarize twas"
-output: html_document
+title: "Summarize twas: chr22"
+date: "`r Sys.Date()`"
+output: workflowr::wflow_html
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE,  message= F)
+knitr::opts_chunk$set(echo = TRUE,  message= F, warning = F)
 ```

### 2020-08-05 analysis/simulation-multi-ukbchr17to22-gtex.adipose.Rmd
move code-  plot functions

@@
 library(kableExtra)
+source("analysis/summarize_twas_plots.R")
 ```
@@
 
-```{r getfile function}
-
-get_files <- function(tag, tag2){
-  par <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt")
-  rpip <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".rPIP.txt")
-
-  gmrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt")
-  smrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt")
-
-  ggwas <- paste0(outputdir, tag, ".exprgwas.txt.gz")
-  sgwas <- paste0(outputdir, tag, ".snpgwas.txt.gz")
-
-  gsusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.expr.txt")
-  ssusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.snp.txt")
-
-  return(tibble::lst(par, rpip, gmrash, ggwas, smrash, sgwas, gsusie, ssusie))
-}
-
-get_tags <- function(globpattern, extrpattern, tag2){
-  lapply(lapply(get_files(globpattern, tag2), Sys.glob), str_extract, pattern = extrpattern)
-}
-```
 # Mr.ash2 parameter estimation
@@
 Results for 10 simulations runs, using different initiate and update strategy
-
-```{r param function}
-show_param <- function(tags, tag2){
-  f <- lapply(tags, get_files, tag2 = tag2)
-  parf <- lapply(f, '[[', "par")
-  param <- do.call(rbind, lapply(parf, function(x) t(read.table(x))[2:1,]))
-  truth <- param[1:(nrow(param)/2)*2-1,]
-  est <- param[1:(nrow(param)/2)*2,]
-  outdt <- matrix(0, ncol = 2*ncol(param), nrow = nrow(param)/2)
-  outdt[,c(1,3,5,7)] <- truth
-  outdt[,c(2,4,6,8)] <- est
-  outdt <-cbind(1:nrow(outdt),outdt)
-  colnames(outdt) <- c("Simulation#", paste0(rep(c("Truth","Est."),4)))
-  knitr::kable(outdt) %>%
-  kable_styling("striped") %>%
-  add_header_above(c(" " = 1, "Gene.pi1" = 2, "Gene.PVE" = 2, "SNP.pi1" = 2, "SNP.PVE" =2))
-}
-```
 
@@
 We run 50 simulations and combine results.
-```{r caliplot}
-#' s is pip or fdr.
-cp_plot <- function(s, ifcausal, mode = c("PIP", "FDR"), main = mode[1]){
-  # ifcausal:0,1
-  a_bin <- cut(s, breaks= seq(0,1, by=0.1))
-  if (mode == "PIP") {
-     expected = c(by(s, a_bin, FUN = mean))
-     observed = c(by(ifcausal, a_bin, FUN = mean))
-  } else if (mode == "FDR"){
-     expected = c(by(s, a_bin, FUN = mean))
-     observed = 1 - c(by(ifcausal, a_bin, FUN = mean))
-  }
-  plot(expected, observed, xlim= c(0,1), ylim=c(0,1), pch =19, main = main)
-  lines(x = c(0,1), y = c(0,1), col ="red")
-}
-
-caliPIP_plot <- function(tags, tag2){
-  f <- lapply(tags, get_files, tag2 = tag2)
-  mrashf <- lapply(f, '[[', "gmrash")
-  names(mrashf) <- tags
-
-  susief <- lapply(f, '[[', "gsusie")
-  names(susief) <- tags
-
-  .tagname <- function(x, flist){
-    a <- read.table(flist[[x]], header =T)
-    a[, "name"] <- paste0(x, ":", a[, "name"])
-    a
-  }
-  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
-  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))
-
-  res <- merge(mrashres, susieres, by = "name", all = T)
-
-  res <- res[complete.cases(res),]
-  res <- rename(res, c("PIP" = "mr.ash_PIP", "pip" = "SUSIE_PIP", "pip.null" = "SUSIE_PIP_null", "pip.w0" = "SUSIE_PIP_w0"))
-  par(mfrow=c(1,4), mar=c(5, 6, 4, 1))
-  cp_plot(res$mr.ash_PIP, res$ifcausal, main = "Mr.ash PIP")
-  cp_plot(res$SUSIE_PIP, res$ifcausal, main = "SUSIE PIP")
-  cp_plot(res$SUSIE_PIP_null, res$ifcausal, main = "SUSIE PIP null")
-  cp_plot(res$SUSIE_PIP_w0, res$ifcausal, main = "SUSIE PIP weighted null")
-}
-
-caliFDR_plot <- function(tags, tag2){
-
-  f <- lapply(tags, get_files, tag2 = tag2)
-  gwasf <- lapply(f, '[[', "ggwas")
-  names(gwasf) <- tags
-
-  .tagname <- function(x, flist, colnames = NULL){
-    a <- read.table(flist[[x]], header =T)
-    if (!is.null(colnames)){
-      colnames(a) <- colnames
-    }
-    a[, "name"] <- paste0(x, ":", a[, "name"])
-    a
-  }
-
-  .addFDR <- function(res){
-    res$FDR <- p.adjust(res$PVALUE, method = "fdr")
-    res
-  }
-
-  gwasres <- do.call(rbind, lapply(lapply(tags, .tagname, flist = gwasf, 
-                                   colnames =  c("chr",	"p0",	"p1", "name", 
-                                                 "Estimate", "Std.Error", "t-value", "PVALUE")), .addFDR))
-
-  f <- lapply(tags, get_files, tag2 = tag2)
-  mrashf <- lapply(f, '[[', "gmrash")
-  names(mrashf) <- tags
-
-  susief <- lapply(f, '[[', "gsusie")
-  names(susief) <- tags
-
-  .tagname <- function(x, flist){
-    a <- read.table(flist[[x]], header =T)
-    a[, "name"] <- paste0(x, ":", a[, "name"])
-    a
-  }
-  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
-  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))
-
-  res <- merge(mrashres, susieres, by = "name", all = T)
-
-  res <- merge(res, gwasres, by = "name", all = T)
-
-  res <- res[complete.cases(res),]
-
-  cp_plot(res$FDR, res$ifcausal, mode ="FDR", main = "TWAS FDR")
-  cat("FDR at Bonferroni corrected p = 0.05: ", 1 - mean(res[res$PVALUE < 0.05 /dim(res)[1], "ifcausal"]))
-}
-```
 
@@
 mr.ash2s PIP vs. susie PIP.
-```{r scatterplot}
-scatter_plot_PIP<- function(tags, tag2){
-  f <- lapply(tags, get_files, tag2 = tag2)
-  mrashf <- lapply(f, '[[', "gmrash")
-  names(mrashf) <- tags
 
-  susief <- lapply(f, '[[', "gsusie")
-  names(susief) <- tags
-
-  .tagname <- function(x, flist){
-    a <- read.table(flist[[x]], header =T)
-    a[, "name"] <- paste0(x, ":", a[, "name"])
-    a
-  }
-  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
-  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))
-
-  res <- merge(mrashres, susieres, by = "name", all = T)
-
-  res <- res[complete.cases(res),]
-  res <- rename(res, c("PIP" = "mr.ash_PIP", "pip" = "SUSIE_PIP", "pip.null" = "SUSIE_PIP_null", "pip.w0" = "SUSIE_PIP_w0") )
-  res$ifcausal <- mapvalues(res$ifcausal, 
-          from=c(0,1), 
-          to=c("Non causal", "Causal"))
-
-  fig1 <- plot_ly(data = res, x = ~ mr.ash_PIP, y = ~ SUSIE_PIP, color = ~ ifcausal, 
-                 colors = c( "salmon", "darkgreen"))
-
-  fig2 <- plot_ly(data = res, x = ~ mr.ash_PIP, y = ~ SUSIE_PIP_null, color = ~ ifcausal, 
-                 colors = c( "salmon", "darkgreen"))
-
-  fig3 <- plot_ly(data = res, x = ~ mr.ash_PIP, y = ~ SUSIE_PIP_w0, color = ~ ifcausal, 
-                 colors = c( "salmon", "darkgreen"))
-
-  fig <- subplot(fig1, fig2, fig3, titleX = TRUE, titleY = T, margin = 0.05)
-  fig
-}
-```
 ## NULL; expr-snp; expr-snp 
@@
 ```
+
 ## NULL; snp-expr; expr-snp
@@
 ```
+
 ## lasso; expr-snp; expr-snp
@@
 # ROC curve
-```{r roc}
-ROC_plot<- function(tags, tag2){
-  f <- lapply(tags, get_files, tag2 = tag2)
-  mrashf <- lapply(f, '[[', "gmrash")
-  names(mrashf) <- tags
-
-  susief <- lapply(f, '[[', "gsusie")
-  names(susief) <- tags
-
-  gwasf <- lapply(f, '[[', "ggwas")
-  names(gwasf) <- tags
-
-  .tagname <- function(x, flist, colnames = NULL){
-    a <- read.table(flist[[x]], header =T)
-    if (!is.null(colnames)){
-      colnames(a) <- colnames
-    }
-    a[, "name"] <- paste0(x, ":", a[, "name"])
-    a
-  }
-  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
-  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))
-  gwasres <- do.call(rbind, lapply(tags, .tagname, flist = gwasf, 
-                                   colnames =  c("chr",	"p0",	"p1", "name", "Estimate", "Std.Error", "t-value", "PVALUE")))
-
-  res <- merge(mrashres, susieres, by = "name", all = T)
-  res <- merge(res, gwasres, by = "name", all = T)
-
-  res <- res[complete.cases(res),]
-  res <- rename(res, c("PIP" = "mr.ash", "pip" = "SUSIE", "pip.null"= "SUSIE.null", "pip.w0" = "SUSIE.w0", "PVALUE" = "TWAS") )
-  res[,"TWAS"] <- -log10(res[, "TWAS"])
-
-  roccolors <-  c("red", "green", "orange", "pink", "blue")
-  methods <- c("mr.ash", "SUSIE", "SUSIE.null", "SUSIE.w0", "TWAS")
-  plot(0, xlim=c(0,1), ylim=c(0,1), col="white", xlab = "FPR", ylab = "TPR")
-  for (i in 1:length(methods)){
-    method <- methods[i]
-    bordered <- res[order(res[,method]),] 
-    actuals <- bordered$ifcausal == 1
-    sens <- (sum(actuals) - cumsum(actuals))/sum(actuals)
-    spec <- cumsum(!actuals)/sum(!actuals)
-    lines(1 - spec, sens, type = "l", col = roccolors[i])
-    abline(c(0,0),c(1,1))
-    auc <- sum(spec*diff(c(0, 1 - sens)))
-    cat("AUC for ", method, ": ", auc)
-  }
-  legend(0.6,0.3, legend= methods, col=roccolors, lty=1, cex=0.5 )
-  grid()
-}
-```
 

### 2020-08-05 analysis/simulation-multi-ukbchr17to22-gtex.adipose_sa2_v2.Rmd
move code-  plot functions

@@
 We run 50 simulations and combine results.
-```{r caliplot}
-#' s is pip or fdr.
-cp_plot <- function(s, ifcausal, mode = c("PIP", "FDR"), main = mode[1]){
-  # ifcausal:0,1
-  a_bin <- cut(s, breaks= seq(0,1, by=0.1))
-  if (mode == "PIP") {
-     expected = c(by(s, a_bin, FUN = mean))
-     observed = c(by(ifcausal, a_bin, FUN = mean))
-  } else if (mode == "FDR"){
-     expected = c(by(s, a_bin, FUN = mean))
-     observed = 1 - c(by(ifcausal, a_bin, FUN = mean))
-  }
-  plot(expected, observed, xlim= c(0,1), ylim=c(0,1), pch =19, main = main)
-  lines(x = c(0,1), y = c(0,1), col ="red")
-}
-
-caliPIP_plot <- function(tags, tag2){
-  f <- lapply(tags, get_files, tag2 = tag2)
-  mrashf <- lapply(f, '[[', "gmrash")
-  names(mrashf) <- tags
-
-  susief <- lapply(f, '[[', "gsusie")
-  names(susief) <- tags
-
-  .tagname <- function(x, flist){
-    a <- read.table(flist[[x]], header =T)
-    a[, "name"] <- paste0(x, ":", a[, "name"])
-    a
-  }
-  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
-  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))
-
-  res <- merge(mrashres, susieres, by = "name", all = T)
-
-  res <- res[complete.cases(res),]
-  res <- rename(res, c("PIP" = "mr.ash_PIP", "pip" = "SUSIE_PIP", "pip.null" = "SUSIE_PIP_null", "pip.w0" = "SUSIE_PIP_w0"))
-  par(mfrow=c(1,4), mar=c(5, 6, 4, 1))
-  cp_plot(res$mr.ash_PIP, res$ifcausal, main = "Mr.ash PIP")
-  cp_plot(res$SUSIE_PIP, res$ifcausal, main = "SUSIE PIP")
-  cp_plot(res$SUSIE_PIP_null, res$ifcausal, main = "SUSIE PIP null")
-  cp_plot(res$SUSIE_PIP_w0, res$ifcausal, main = "SUSIE PIP weighted null")
-}
-
-caliFDR_plot <- function(tags, tag2){
-
-  f <- lapply(tags, get_files, tag2 = tag2)
-  gwasf <- lapply(f, '[[', "ggwas")
-  names(gwasf) <- tags
-
-  .tagname <- function(x, flist, colnames = NULL){
-    a <- read.table(flist[[x]], header =T)
-    if (!is.null(colnames)){
-      colnames(a) <- colnames
-    }
-    a[, "name"] <- paste0(x, ":", a[, "name"])
-    a
-  }
-
-  .addFDR <- function(res){
-    res$FDR <- p.adjust(res$PVALUE, method = "fdr")
-    res
-  }
-
-  gwasres <- do.call(rbind, lapply(lapply(tags, .tagname, flist = gwasf, 
-                                   colnames =  c("chr",	"p0",	"p1", "name", 
-                                                 "Estimate", "Std.Error", "t-value", "PVALUE")), .addFDR))
-
-  f <- lapply(tags, get_files, tag2 = tag2)
-  mrashf <- lapply(f, '[[', "gmrash")
-  names(mrashf) <- tags
-
-  susief <- lapply(f, '[[', "gsusie")
-  names(susief) <- tags
-
-  .tagname <- function(x, flist){
-    a <- read.table(flist[[x]], header =T)
-    a[, "name"] <- paste0(x, ":", a[, "name"])
-    a
-  }
-  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
-  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))
-
-  res <- merge(mrashres, susieres, by = "name", all = T)
-
-  res <- merge(res, gwasres, by = "name", all = T)
-
-  res <- res[complete.cases(res),]
-
-  cp_plot(res$FDR, res$ifcausal, mode ="FDR", main = "TWAS FDR")
-  cat("FDR at bonferroni corrected p = 0.05: ", 1 - mean(res[res$PVALUE < 0.05 /dim(res)[1], "ifcausal"]))
-}
-```
 

### 2020-08-05 analysis/simulation-multi-ukbchr22-gtex.adipose2.Rmd
move code-  plot functions

@@
 
-```{r getfile function}
-get_files <- function(tag, tag2){
-  par <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt")
-  rpip <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".rPIP.txt")
-
-  gmrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt")
-  smrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt")
-
-  ggwas <- paste0(outputdir, tag, ".exprgwas.txt.gz")
-  sgwas <- paste0(outputdir, tag, ".snpgwas.txt.gz")
-
-  gsusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.expr.txt")
-  ssusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.snp.txt")
-
-  return(tibble::lst(par, rpip, gmrash, ggwas, smrash, sgwas, gsusie, ssusie))
-}
-
-get_tags <- function(globpattern, extrpattern, tag2){
-  lapply(lapply(get_files(globpattern, tag2), Sys.glob), str_extract, pattern = extrpattern)
-}
-```
 # Mr.ash2 parameter estimation
@@
 We run 100 simulations and combine results.
-```{r caliplot}
-#' s is pip or fdr.
-cp_plot <- function(s, ifcausal, mode = c("PIP", "FDR"), main = mode[1]){
-  # ifcausal:0,1
-  a_bin <- cut(s, breaks= seq(0,1, by=0.1))
-  if (mode == "PIP") {
-     expected = c(by(s, a_bin, FUN = mean))
-     observed = c(by(ifcausal, a_bin, FUN = mean))
-  } else if (mode == "FDR"){
-     expected = c(by(s, a_bin, FUN = mean))
-     observed = 1 - c(by(ifcausal, a_bin, FUN = mean))
-  }
-  plot(expected, observed, xlim= c(0,1), ylim=c(0,1), pch =19, main = main)
-  lines(x = c(0,1), y = c(0,1), col ="red")
-}
-
-caliPIP_plot <- function(tags, tag2){
-  f <- lapply(tags, get_files, tag2 = tag2)
-  mrashf <- lapply(f, '[[', "gmrash")
-  names(mrashf) <- tags
-
-  susief <- lapply(f, '[[', "gsusie")
-  names(susief) <- tags
-
-  .tagname <- function(x, flist){
-    a <- read.table(flist[[x]], header =T)
-    a[, "name"] <- paste0(x, ":", a[, "name"])
-    a
-  }
-  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
-  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))
-
-  res <- merge(mrashres, susieres, by = "name", all = T)
-
-  res <- res[complete.cases(res),]
-  res <- rename(res, c("PIP" = "mr.ash_PIP", "pip" = "SUSIE_PIP", "pip.null" = "SUSIE_PIP_null", "pip.w0" = "SUSIE_PIP_w0"))
-  par(mfrow=c(1,4), mar=c(5, 6, 4, 1))
-  cp_plot(res$mr.ash_PIP, res$ifcausal, main = "Mr.ash PIP")
-  cp_plot(res$SUSIE_PIP, res$ifcausal, main = "SUSIE PIP")
-  cp_plot(res$SUSIE_PIP_null, res$ifcausal, main = "SUSIE PIP null")
-  cp_plot(res$SUSIE_PIP_w0, res$ifcausal, main = "SUSIE PIP weighted null")
-}
-
-caliFDR_plot <- function(tags, tag2){
-
-  f <- lapply(tags, get_files, tag2 = tag2)
-  gwasf <- lapply(f, '[[', "ggwas")
-  names(gwasf) <- tags
-
-  .tagname <- function(x, flist, colnames = NULL){
-    a <- read.table(flist[[x]], header =T)
-    if (!is.null(colnames)){
-      colnames(a) <- colnames
-    }
-    a[, "name"] <- paste0(x, ":", a[, "name"])
-    a
-  }
-
-  .addFDR <- function(res){
-    res$FDR <- p.adjust(res$PVALUE, method = "fdr")
-    res
-  }
-
-  gwasres <- do.call(rbind, lapply(lapply(tags, .tagname, flist = gwasf, 
-                                   colnames =  c("chr",	"p0",	"p1", "name", 
-                                                 "Estimate", "Std.Error", "t-value", "PVALUE")), .addFDR))
-
-  f <- lapply(tags, get_files, tag2 = tag2)
-  mrashf <- lapply(f, '[[', "gmrash")
-  names(mrashf) <- tags
-
-  susief <- lapply(f, '[[', "gsusie")
-  names(susief) <- tags
-
-  .tagname <- function(x, flist){
-    a <- read.table(flist[[x]], header =T)
-    a[, "name"] <- paste0(x, ":", a[, "name"])
-    a
-  }
-  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
-  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))
-
-  res <- merge(mrashres, susieres, by = "name", all = T)
-
-  res <- merge(res, gwasres, by = "name", all = T)
-
-  res <- res[complete.cases(res),]
-
-  cp_plot(res$FDR, res$ifcausal, mode ="FDR", main = "TWAS FDR")
-  cat("FDR at bonferroni corrected p = 0.05: ", 1 - mean(res[gwasres$PVALUE < 0.05 /dim(res)[1], "ifcausal"]))
-}
-```
 
@@
 mr.ash2s PIP vs. susie PIP.
-```{r scatterplot}
-scatter_plot_PIP<- function(tags, tag2){
-  f <- lapply(tags, get_files, tag2 = tag2)
-  mrashf <- lapply(f, '[[', "gmrash")
-  names(mrashf) <- tags
 
-  susief <- lapply(f, '[[', "gsusie")
-  names(susief) <- tags
-
-  .tagname <- function(x, flist){
-    a <- read.table(flist[[x]], header =T)
-    a[, "name"] <- paste0(x, ":", a[, "name"])
-    a
-  }
-  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
-  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))
-
-  res <- merge(mrashres, susieres, by = "name", all = T)
-
-  res <- res[complete.cases(res),]
-  res <- rename(res, c("PIP" = "mr.ash_PIP", "pip" = "SUSIE_PIP", "pip.null" = "SUSIE_PIP_null") )
-  res$ifcausal <- mapvalues(res$ifcausal, 
-          from=c(0,1), 
-          to=c("Non causal", "Causal"))
-
-  fig1 <- plot_ly(data = res, x = ~ mr.ash_PIP, y = ~ SUSIE_PIP, color = ~ ifcausal, 
-                 colors = c( "salmon", "darkgreen"))
-
-  fig2 <- plot_ly(data = res, x = ~ mr.ash_PIP, y = ~ SUSIE_PIP_null, color = ~ ifcausal, 
-                 colors = c( "salmon", "darkgreen"))
-
-  fig <- subplot(fig1, fig2, titleX = TRUE, titleY = T, margin = 0.1)
-  fig
-}
-```
 ## NULL; expr-snp; expr-snp 
@@
 # ROC curve
-```{r roc}
-ROC_plot<- function(tags, tag2){
-  f <- lapply(tags, get_files, tag2 = tag2)
-  mrashf <- lapply(f, '[[', "gmrash")
-  names(mrashf) <- tags
-
-  susief <- lapply(f, '[[', "gsusie")
-  names(susief) <- tags
-
-  gwasf <- lapply(f, '[[', "ggwas")
-  names(gwasf) <- tags
-
-  .tagname <- function(x, flist, colnames = NULL){
-    a <- read.table(flist[[x]], header =T)
-    if (!is.null(colnames)){
-      colnames(a) <- colnames
-    }
-    a[, "name"] <- paste0(x, ":", a[, "name"])
-    a
-  }
-  mrashres <- do.call(rbind, lapply(tags, .tagname, flist = mrashf))
-  susieres <- do.call(rbind, lapply(tags, .tagname, flist = susief))
-  gwasres <- do.call(rbind, lapply(tags, .tagname, flist = gwasf, 
-                                   colnames =  c("chr",	"p0",	"p1", "name", "Estimate", "Std.Error", "t-value", "PVALUE")))
-
-  res <- merge(mrashres, susieres, by = "name", all = T)
-  res <- merge(res, gwasres, by = "name", all = T)
-
-  res <- res[complete.cases(res),]
-  res <- rename(res, c("PIP" = "mr.ash", "pip" = "SUSIE", "PVALUE" = "TWAS") )
-  res[,"TWAS"] <- -log10(res[, "TWAS"])
-
-  roccolors <-  c("red", "green", "blue")
-  methods <- c("mr.ash", "SUSIE", "TWAS")
-  plot(0, xlim=c(0,1), ylim=c(0,1), col="white", xlab = "FPR", ylab = "TPR")
-  for (i in 1:3){
-    method <- methods[i]
-    bordered <- res[order(res[,method]),] 
-    actuals <- bordered$ifcausal == 1
-    sens <- (sum(actuals) - cumsum(actuals))/sum(actuals)
-    spec <- cumsum(!actuals)/sum(!actuals)
-    lines(1 - spec, sens, type = "l", col = roccolors[i])
-    abline(c(0,0),c(1,1))
-    auc <- sum(spec*diff(c(0, 1 - sens)))
-    cat("AUC for ", method, ": ", auc)
-  }
-  legend(0.6,0.3, legend= methods, col=roccolors, lty=1, cex=0.8)
-  grid()
-}
-```
 

### 2020-08-05 analysis/index.Rmd
visual chr17to22

@@
 
-## Test veb_boost
-A simplified version (mr.ash2): interative update of two mr.ash components
+## Test mr.ash2s
+A simplified version (mr.ash2) of `Veb_boost`: interative update of two mr.ash components
 
@@
   * [Using UKbiobank data (chr22), fusion weights Adipose subcutaneous](https://users.rcc.uchicago.edu/~simingz/simulation-mr.ash2s-ukbchr22-gtex.adipose.html)
+
+  * [Using UKbiobank data (chr17 to 22), fusion weights Adipose subcutaneous](https://users.rcc.uchicago.edu/~simingz/simulation-mr.ash2s-ukbchr17to22-gtex.adipose.html)
 

### 2020-08-05 analysis/simulation-mr.ash2s-ukbchr17to22-gtex.adipose.Rmd
visual chr17to22

---
title: "Test mr.ash2s on ukbiobank data, chr17-22"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message= F, warning = F)
```

We test the performance of mr.ash2s on ukbiobank data (chr17 to 22, n sample = 20,000). 
We use weights from Fusion website, using [Adipose (subcutaneous) p < 0.01](http://gusevlab.org/projects/fusion/weights/GTEx.Adipose_Subcutaneous.P01.tar.bz2).
Total features > 8,000. On chr17 to 22: ~1500.

```{r plot functions}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})

source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")

plot_pip <- function(dt){

  pal <- c("salmon", "darkgreen")
  pal <- setNames(pal, c("Causal", "Non causal"))

  if ( "max_r2" %in% colnames(dt)) {
    fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', 
                   mode = 'markers', color = ~ ifcausal, 
                   text = ~ paste("Name: ", name, 
                                  "\nmax R^2: ", round(max_r2, digits = 2), 
                                  "\nMax pair: ", max_pair))
  } else {
    fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))
  }

  fig <- fig %>% layout(yaxis = list(title = "PIP"))
  return(fig)
}

plot_beta <- function(dt){
  fig <- plot_ly(dt, x = ~p0, y = ~ TrueBeta, type = 'scatter', mode = 'markers', name ="Truth", text = ~paste("Name: ", name))
  fig <- fig %>% add_trace(y = ~ EstBeta, name = 'Estimated',mode = 'markers')

  fig <- fig %>% layout(yaxis = list(title = "beta"))
  return(fig)
}

plot_p <- function(dt){
  pal <- c("salmon", "darkgreen")
  pal <- setNames(pal, c("Causal", "Non causal"))

  fig <- plot_ly(dt, x = ~p0, y = ~ PVALUE , type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))

  fig <- fig %>% layout(yaxis = list(title = "p value"))
  return(fig)
}

plot_susie_pip <- function(dt) {
   pal <- c("salmon", "darkgreen")
   pal <- setNames(pal, c("Causal", "Non causal"))

   fig <- plot_ly(dt, x = ~ p0, y = ~ susiePIP , type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))

   fig <- fig %>% layout(yaxis = list(title = "SUSIE PIP"))
}

plot_all <- function(mrashres.gene, gwasres.gene, mrashres.snp, gwasres.snp, susieres.gene, susieres.snp, ldres = NULL, chrom=NULL){
  a <- read.table(mrashres.gene , header =T)
  a[a$ifcausal == 1, "ifcausal"] = "Causal"
  a[a$ifcausal == 0, "ifcausal"] = "Non causal"

  if (!is.null(ldres)){
    ld <- read.table(ldres, header =T)
    a <- merge(a, ld, by = "name", all.x = T)}

  su <- read.table(susieres.gene, header =T)
  colnames(su)[colnames(su) == "pip"] <-  "susiePIP"
  a <- merge(a, su, by = "name", all.x = T)

  p <- fread(gwasres.gene, header =T)
  p[,"PVALUE"] <- -log10(p[,"PVALUE"])
  colnames(p)[colnames(p) == "MARKER_ID"] <-  "name"
  dt.gene <- merge(a, p, by = "name", all.x = T)

  a <- read.table(mrashres.snp , header =T)
  a[a$ifcausal == 1, "ifcausal"] = "Causal"
  a[a$ifcausal == 0, "ifcausal"] = "Non causal"

  su <- read.table(susieres.snp, header =T)
  colnames(su)[colnames(su) == "pip"] <-  "susiePIP"
  a <- merge(a, su, by = "name", all.x = T)

  p <- fread(gwasres.snp, header =T)
  p[,"PVALUE"] <- -log10(p[,"PVALUE"])
  colnames(p)[colnames(p) == "MARKER_ID"] <-  "name"
  dt.snp <- merge(a, p, by = "name", all.x = T)

  if (!is.null(chrom)){
    dt.gene <- dt.gene[dt.gene$chr.x == chrom,]
    dt.snp <- dt.snp[dt.snp$chr.x == chrom,]
  }

  fig1 <- plot_pip(dt.gene)  
  fig2 <- plot_beta(dt.gene) 
  fig3 <- plot_p(dt.gene) 
  fig4 <- plot_susie_pip(dt.gene) 
  fig5 <- plot_pip(dt.snp) 
  fig6 <- plot_beta(dt.snp) 
  fig7 <- plot_p(dt.snp)
  fig8 <- plot_susie_pip(dt.snp) %>% layout( plot_bgcolor='#EFF6FC')

  fig <- subplot(fig1, fig2, fig3, fig4, fig5, fig6, fig7, fig8, nrows=8, shareX = T, titleY = T)

  fig <- fig %>% layout(autosize = F, width = 600, height = 1100, margin = 0.02,
    xaxis = list(title = "Genomic position"))

  fig <- fig %>% toWebGL()
  fig
}

```

```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
```

```{r getfile function}
get_files <- function(tag, tag2){
  par <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt")
  gmrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt")
  ggwas <- paste0(outputdir, tag, ".exprgwas.txt.gz")
  smrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt")
  sgwas <- paste0(outputdir, tag, ".snpgwas.txt.gz")
  gsusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.expr.txt")
  ssusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.snp.txt")
  rsq <- paste0(outputdir, tag, ".LD.txt")
  return(tibble::lst(par, gmrash, ggwas, smrash, sgwas, gsusie, ssusie, rsq))
}
```

## Simulation 1
## NULL; expr-snp; expr-snp
```{r sim 1 gene, out.width = "98%", out.height="95%"}
tag <- "20200721-1-2"
tag2s <- c('zeroes-es', 'zerose-es', 'lassoes-es','lassoes-se')
tag2 <-tag2s[1]

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, chrom = "17")
```

```{r sim 2 gene, out.width = "98%", out.height="95%"}
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, chrom = "18")
```

```{r sim 3 gene, out.width = "98%", out.height="95%"}
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, chrom = "19")
```

### 2020-08-05 analysis/simulation-multi-ukbchr17to22-gtex.adipose_sa2_v2.Rmd
visual chr17to22

@@
 susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
-tags <- paste0('20200721-1-', 1:7)
+tags <- paste0('20200721-1-', c(2, 4:9))
 tagglob <- '20200721-1-*'

### 2020-08-06 analysis/simulation-mr.ash2s-ukbchr17to22-gtex.adipose.Rmd
Remove ignored files

@@
     fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', 
-                   mode = 'markers', color = ~ ifcausal, 
+                   mode = 'markers', color = ~ ifcausal, colors = pal,
                    text = ~ paste("Name: ", name, 
-                                  "\nmax R^2: ", round(max_r2, digits = 2), 
-                                  "\nMax pair: ", max_pair))
+                                  "\nmax R^2: ", round(max_r2, digits = 2)))
+                                  #,"\nMax pair: ", max_pair))
   } else {
@@
 ## NULL; expr-snp; expr-snp
-```{r sim 1 gene, out.width = "98%", out.height="95%"}
-tag <- "20200721-1-2"
-tag2s <- c('zeroes-es', 'zerose-es', 'lassoes-es','lassoes-se')
-tag2 <-tag2s[1]
-
+```{r sim 1-4, out.width = "98%", out.height="95%"}
+tag <- "20200721-1-4"
+tag2 <- "zeroes-es"
 f <- get_files(tag= tag, tag2 = tag2)
@@
 print(param)
-plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, chrom = "17")
+```
+```{r 1-4-17, out.width = "98%", out.height="95%"}
+chrom <- "17"
+cat("simulation run: ", tag, "-", tag2, "; chr: ", chrom)
+plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, f$rsq, chrom = chrom)
 ```
 
-```{r sim 2 gene, out.width = "98%", out.height="95%"}
-plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, chrom = "18")
+```{r 1-4-19 gene, out.width = "98%", out.height="95%"}
+chrom <- "19"
+cat("simulation run: ", tag, "-", tag2, "; chr: ", chrom)
+plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, f$rsq, chrom = chrom)
 ```
 
-```{r sim 3 gene, out.width = "98%", out.height="95%"}
-plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, chrom = "19")
+```{r sim 1-8, out.width = "98%", out.height="95%"}
+tag <- "20200721-1-8"
+f <- get_files(tag= tag, tag2 = tag2)
+param <- read.table(f$par, header = T, row.names = 1)
+print(param)
 ```
+```{r 1-8-20, out.width = "98%", out.height="95%"}
+chrom <- "20"
+cat("simulation run: ", tag, "-", tag2, "; chr: ", chrom)
+plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, f$rsq, chrom = chrom)
+```
+
+```{r sim 1-9, out.width = "98%", out.height="95%"}
+tag <- "20200721-1-9"
+f <- get_files(tag= tag, tag2 = tag2)
+param <- read.table(f$par, header = T, row.names = 1)
+print(param)
+```
+```{r 1-9-19, out.width = "98%", out.height="95%"}
+chrom <- "19"
+cat("simulation run: ", tag, "-", tag2, "; chr: ", chrom)
+plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, f$rsq, chrom = chrom)
+```

### 2020-08-06 analysis/simulation-multi-ukbchr17to22-gtex.adipose_sa2_v2.Rmd
Remove ignored files

@@
 chrom = 18
-f <- get_files(tag= "20200721-1-2" , tag2 = tag2s[1])
+f <- get_files(tag= "20200721-1-3" , tag2 = tag2s[1])
 allchr <- read.table(f[["rpip"]], header = T)
@@
 ```
+
+# PIP vs p value
+
+## NULL; expr-snp; expr-snp 
+```{r sca_p 1, fig.height=7, fig.width=7}
+scatter_plot_PIP_p(tags, tag2s[1])
+```

### 2020-08-07 analysis/simulation-multi-ukbchr17to22-gtex.adipose.Rmd
visual chr17to22
Remove ignored files
susie names

@@
 
-Results for 10 simulations runs, using different initiate and update strategy
+Results for 10 simulations runs, using different initialize and update strategy
 

### 2020-08-07 analysis/simulation-multi-ukbchr22-gtex.adipose2.Rmd
Remove ignored files
susie names

@@
 # Mr.ash2 parameter estimation
-Results for 9 simulations runs, using different initiate and update strategy
+Results for 9 simulations runs, using different initialize and update strategy
 ```{r param function}

### 2020-08-12 analysis/simulation-mr.ash2s-ukbchr22-gtex.adipose.Rmd
twas summarize

@@
 MR.ASH results (start from gene):
-```{r sim 5 gene, out.width = "98%", out.height="95%"}
+```{r sim 5 gene, out.width = "98%", out.height="95%", eval = F}
 tag <- "20200616-5"
@@
 MR.ASH results (start from gene):
-```{r sim 6 gene, out.width = "98%", out.height="95%"}
+```{r sim 6 gene, out.width = "98%", out.height="95%", eval = F}
 tag <- "20200616-6"
@@
 MR.ASH results (use lasso for all):
-```{r sim 7 lasso, out.width = "98%", out.height="95%", eval = T}
+```{r sim 7 lasso, out.width = "98%", out.height="95%", eval = F}
 tag <- "20200616-7"
@@
 MR.ASH results (use lasso for just for SNPs):
-```{r sim 7 lassosnp, out.width = "98%", out.height="95%", eval = T}
+```{r sim 7 lassosnp, out.width = "98%", out.height="95%", eval = F}
 tag <- "20200616-7"
@@
 
-MR.ASH results (use lasso for all):
-```{r sim 8 lasso, out.width = "98%", out.height="95%", eval = T}
+### MR.ASH results (use lasso for all):
+```{r sim 8 lasso, out.width = "98%", out.height="95%", eval = F}
 tag <- "20200616-8"
@@
 
-MR.ASH results (use lasso for just for SNPs):
-```{r sim 8 lassosnp, out.width = "98%", out.height="95%", eval = T}
+### MR.ASH results (use lasso for just for SNPs):
+```{r sim 8 lassosnp, out.width = "98%", out.height="95%", eval = F}
 tag <- "20200616-8"
@@
 ```
+
+### More examples from separate runs: 
+```{r sim 20200616-8-7, out.width = "98%", out.height="95%"}
+source("analysis/visual_twas_plots.R")
+tag <- "20200616-8-7"
+tag2 <- "lassoes-es"
+f <- get_files(tag= tag, tag2 = tag2)
+param <- read.table(f$par, header = T, row.names = 1)
+print(param)
+```
+```{r 20200616-8-7, out.width = "98%", out.height="95%"}
+plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, f$rsq)
+```

### 2020-08-12 analysis/visual_twas_plots.R

@@ -1,0 +1,119 @@
+library(mr.ash.alpha)
+library(data.table)
+suppressMessages({library(plotly)})
+
+source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")
+
+get_files <- function(tag, tag2){
+  par <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt")
+  gmrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt")
+  ggwas <- paste0(outputdir, tag, ".exprgwas.txt.gz")
+  smrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt")
+  sgwas <- paste0(outputdir, tag, ".snpgwas.txt.gz")
+  gsusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.expr.txt")
+  ssusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.snp.txt")
+  rsq <- paste0(outputdir, tag, ".LD.txt")
+  return(tibble::lst(par, gmrash, ggwas, smrash, sgwas, gsusie, ssusie, rsq))
+}
+
+plot_pip <- function(dt){
+  
+  pal <- c("salmon", "darkgreen")
+  pal <- setNames(pal, c("Causal", "Non causal"))
+  
+  if ( "max_r2" %in% colnames(dt)) {
+    fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', 
+                   mode = 'markers', color = ~ ifcausal, colors = pal,
+                   text = ~ paste("Name: ", name, 
+                                  "\nmax R^2: ", round(max_r2, digits = 2)))
+    #,"\nMax pair: ", max_pair))
+  } else {
+    fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))
+  }
+  
+  fig <- fig %>% layout(yaxis = list(title = "PIP"))
+  return(fig)
+}
+
+plot_beta <- function(dt){
+  fig <- plot_ly(dt, x = ~p0, y = ~ TrueBeta, type = 'scatter', mode = 'markers', name ="Truth", text = ~paste("Name: ", name))
[…]

### 2020-08-12 code/run_test_mr.ash_gene.R

@@ -1,0 +1,40 @@
+library(logging)
+
+args = commandArgs(trailingOnly=TRUE)
+if (length(args) != 3) {
+  stop("3 arguments must be supplied:
+       * expr Rd
+       * pheno Rd
+       * out file name", call.=FALSE)
+}
+
+addHandler(writeToFile, file="run_mr.ash2s_gene.R.log", level='DEBUG')
+loginfo('script started ... ')
+
+loginfo("input arg 1 (expr): %s ", args[1])
+loginfo("input arg 2 (pheno): %s ", args[2])
+loginfo("input arg 3 (outname): %s ", args[3])
+
+
+codedir <- "/project2/mstephens/causalTWAS/causal-TWAS/code/"
+source(paste0(codedir, "stats_func.R"))
+source(paste0(codedir, "fit_mr.ash.R"))
+
+# load expression Rd file, variable: exprres
+load(args[1])
+loginfo("expr scaling ...")
+dat$expr <-  scaleRcpp(exprres$expr)
+gc()
+
+# load phenotype Rd file, variable: phenores
+load(args[2])
+phenores$Y <-  phenores$Y - mean(phenores$Y)
+
+outname <- args[3]
+
+# run mr.ash2s
+loginfo("run mr.ash for gene expression ...")
+mr.ash.fit <- mr.ash( X = dat$expr,
+                      y = phenores$Y)
+save(mr.ash.fit, file = paste0(outname, ".mr.ashGene.Rd"))
+

### 2020-08-14 analysis/simulation-multi-ukbchr17to22-gtex.adipose.Rmd
susieI

@@
 ```{r rpip 1, fig.height=4, fig.width=10}
-chrom = 18
-f <- get_files(tag= "20200721-1-2" , tag2 = tag2s[1])
+chrom = 17
+f <- get_files(tag= "20200721-1-1" , tag2 = tag2s[1])
 allchr <- read.table(f[["rpip"]], header = T)

### 2020-08-14 analysis/simulation-multi-ukbchr17to22-gtex.adipose_sa2_v2.Rmd
susie names
susieI

@@
 chrom = 18
-f <- get_files(tag= "20200721-1-3" , tag2 = tag2s[1])
-allchr <- read.table(f[["rpip"]], header = T)
+allchr <- read.table(, header = T)
 a <- allchr[allchr["chrom"]==chrom,]

### 2020-08-14 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
susieI

---
title: "Test susieI using chr17 to 22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  message= F, warning = F)
```

Test susieI for ukb chr 17 to chr 22 combined. SNPs are downsampled to 1/10, eQTLs defined by FUSION-TWAS (Adipose/GTEx) lasso effect size > 0 were kept, p= 86k, n = 20k.

susieI is an iterative version of susie, learning parameters (`prior weights`) iteratively using the top region from mr.ash for genes only.

```{r plot functions}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})
library(tidyr)
library(plyr)
library(stringr)
library(kableExtra)
source("analysis/summarize_twas_plots.R")
```
```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
outputdir <- "~/causalTWAS/simulations/simulation_susieI_20200813/"
susiedir <- "~/causalTWAS/simulations/simulation_susieI_20200813/"
tag <- '20200813-1-1'
```

# MR.ASH for genes only

pip distribution

```{r rpip 1, fig.height=4, fig.width=10}
chrom = 17
rpipf <- paste0(outputdir, tag, "-mr.ash.rPIP.txt")
allchr <- read.table(rpipf, header = T)
a <- allchr[allchr["chrom"]==chrom,]
print(paste("plot for chr", chrom))
par(mar=c(5, 4, 4, 6) + 0.1)
with(a, plot(p0, rPIP, col ='salmon', xlab = "position", ylab= "Sum of PIP", type = 'h', lwd = 2))
par(new = T)
with(a, plot(p0, nCausal, pch =19, col = "darkgreen",axes = FALSE, bty = "n", xlab = "", ylab = ""))
axis(side = 4)
mtext(side = 4, line = 3, 'No. causal signals')
legend("topleft",
       legend=c("Mr.ASH PIP", "# Causal"),
       lty=c(1,0), pch=c(NA, 19), col=c("salmon", "darkgreen"))
grid()
```

```{r rpip 1, fig.height=4, fig.width=10}
 hist(a$rPIP, breaks = 100, col = "salmon", xlab = "PIP", main = "PIP distribution")
```

### 2020-08-14 code/ld.R

@@ -1,0 +1,75 @@
+# global variables: phenores, exprres, dat
+library(snpStats)
+
+#' @description calculates measures of linkage disequilibrium
+#' between each column of x and each column of y
+#' need global variables: phenores, exprres, dat
+#' @return the maximum of ld measures of these pairs
+#'
+ld_max <- function(x,  y, stats = "R.squared"){
+  x <- as(round(x),"SnpMatrix")
+  y <- as(round(y),"SnpMatrix")
+  s <- ld(x, y =y, stats = stats)
+  inds <- arrayInd(which.max(s), dim(s))
+  s.max <- max(s)
+  names(s.max) <- paste(rownames(s)[inds[,1]], colnames(s)[inds[,2]], sep = ':')
+  return(s.max)
+}
+
+.eqtl_geno <- function(gname){
+  x <- exprres$exprlist[[gname]]
+  if (is.null(dim(x$wgt))){
+    dim(x$wgt) <- c(1,2)
+    colnames(x$wgt) <- c("idx", "wgt")
+  }
+  eqtlgeno <- dat$G[ , x$wgt[abs(x$wgt[, "wgt"]) > 0, "idx"], drop = F]
+  colnames(eqtlgeno) <- paste(gname, dat$snp[x$wgt[abs(x$wgt[, "wgt"]) > 0, "idx"], 1], sep = ".")
+  return(eqtlgeno)
+}
+
+
+#' @description get ld for genes and all other causal genes or snps
+#' and report max r^2 pairs
+#'
+#'
+get_ld <- function(outname){
+  g.eqtl <- lapply(colnames(exprres$expr), .eqtl_geno)
+
+  causalg.eqtl <-lapply(colnames(exprres$expr)[phenores$param$idx.cgene], .eqtl_geno)
+  cg <- do.call(cbind, causalg.eqtl)
+
+  csnp <- dat$G[ ,phenores$param$idx.cSNP]
+  colnames(csnp) <- dat$snp[phenores$param$idx.cSNP, 1]
+
+  rsq <- lapply(g.eqtl, ld_max, y = cbind(cg,csnp), stats = "R.squared")
+
+  outdf <- cbind(colnames(exprres$expr), rsq, names(rsq))
+  colnames(outdf) <- c("name", "max_r2", "max_pair")
+
[…]

### 2020-08-14 code/run_test_susieI.R

@@ -3,15 +3,13 @@ library(dplyr)
 library(logging)
 
 args = commandArgs(trailingOnly=TRUE)
-if (length(args) < 6) {
-  stop("at least 6 arguments must be supplied:
+if (length(args) < 5) {
+  stop("at least 5 arguments must be supplied:
        * genotype file name
        * expr Rd
        * pheno Rd
-       * param txt
        * mr.ash.res file
        * out file name
-       * L (default 1, L in susie, optional)
        ", call.=FALSE)
 }
 
@@ -23,8 +21,9 @@ source(paste0(codedir,"input_reformat.R"))
 addHandler(writeToFile, file="run_test_susie.R.log", level='DEBUG')
 loginfo('script started ... ')
 
-PIPfilter <- 0.5
+PIPfilter <- 0.02
 chunksize = 500000
+Niter <- 10
 
 loginfo("regional PIP cut: %s ", PIPfilter)
 loginfo("region size: %s", chunksize)
@@ -44,16 +43,9 @@ dat$expr <-  scaleRcpp(exprres$expr)
 load(args[3])
 phenores$Y <-  phenores$Y - mean(phenores$Y)
 
-# run susie using given priors
-param <- read.table(args[4], header = T, row.names = 1)
 
-prior.gene <- param["gene.pi1", "estimated"]
-prior.SNP <- param["snp.pi1", "estimated"]
-
-gres <- paste0(args[5], ".expr.txt")
-sres <- paste0(args[5], ".snp.txt")
-pipres <- rbind(read.table(gres, header =T),
-                 read.table(sres, header =T))
+gres <- paste0(args[4], ".expr.txt")
+pipres <- read.table(gres, header =T)
 
 regions <- NULL
 chroms <- unique(pipres$chr)
@@ -70,69 +62,80 @@ for (chrom in chroms){
   regions <- rbind(regions, itv)
 }
 
-loginfo("No. intervals for chr %s : %s", chrom, nrow(regions))
-write.table(regions , file = paste0(args[5], ".rPIP.txt")  , row.names=F, col.names=T, sep="\t", quote = F)
+loginfo("No. intervals: %s", nrow(regions))
[…]

### 2020-08-19 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
susieI

@@
 tag <- '20200813-1-1'
+Niter <- 10
 ```
 
-# MR.ASH for genes only
+# MR.ASH pip distribution
 
-pip distribution
+* Use mr.ash for gene only:
 
-```{r rpip 1, fig.height=4, fig.width=10}
-chrom = 17
+```{r rpip 1, fig.height=6, fig.width=6}
 rpipf <- paste0(outputdir, tag, "-mr.ash.rPIP.txt")
-allchr <- read.table(rpipf, header = T)
-a <- allchr[allchr["chrom"]==chrom,]
-print(paste("plot for chr", chrom))
-par(mar=c(5, 4, 4, 6) + 0.1)
-with(a, plot(p0, rPIP, col ='salmon', xlab = "position", ylab= "Sum of PIP", type = 'h', lwd = 2))
-par(new = T)
-with(a, plot(p0, nCausal, pch =19, col = "darkgreen",axes = FALSE, bty = "n", xlab = "", ylab = ""))
-axis(side = 4)
-mtext(side = 4, line = 3, 'No. causal signals')
-legend("topleft",
-       legend=c("Mr.ASH PIP", "# Causal"),
-       lty=c(1,0), pch=c(NA, 19), col=c("salmon", "darkgreen"))
-grid()
+
+a <- read.table(rpipf, header = T)
+a.c <- a[a$nCausal > 0, ]
+ax <- pretty(0:max(a$rPIP), n = 30)
+
+par(mfrow=c(2,1))
+h1 <- hist(a$rPIP, breaks = 100, xlab = "PIP", main = "PIP distribution-all", col = "grey"); grid()
+h2 <- hist(a.c$rPIP, breaks = h1$breaks, xlab = "PIP", main = "PIP distribution-causal", col = "salmon");grid()
 ```
 
-```{r rpip 1, fig.height=4, fig.width=10}
- hist(a$rPIP, breaks = 100, col = "salmon", xlab = "PIP", main = "PIP distribution")
+* MR.ash2s results for the same simulated data
+
+```{r rpip 2, fig.height=6, fig.width=6}
+rpipf <- "~/causalTWAS/simulations/simulation_ashtest_20200721/20200721-1-1-mr.ash2s.lassoes-es.rPIP.txt"
+
+a <- read.table(rpipf, header = T)
+a.c <- a[a$nCausal > 0 ,]
+ax <- pretty(0:max(a$rPIP), n = 30) # Make a neat vector for the breakpoints
+
+par(mfrow=c(2,1))
+h1 <- hist(a$rPIP, breaks = 100, xlab = "PIP", main = "PIP distribution-all", col = "grey"); grid()
+h2 <- hist(a.c$rPIP, breaks = h1$breaks, xlab = "PIP", main = "PIP distribution-causal", col = "salmon");grid()
 ```
+
+# susieI genes
+
+Regions: We use regional PIP of both SNP and gene from mr.ash2s > 0.3 , and also require at least one causal gene or SNP is present, also, at least one gene.
+
+Prior change for each iteration:
+```{r iter}
+load(paste0(outputdir, tag, ".susieIres.Rd"))
+df <- cbind(prior.gene_rec, prior.SNP_rec)
+rownames(df) <- paste("Iteration", 1:Niter)
+colnames(df) <- c("Prior.gene", "Prior.SNP")
+df
+```
+
+Results for each region:
+```{r regions}
+regionlist <- readRDS(paste0(outputdir, tag, "_regionlist.rds"))
+ng <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "gene", , drop = F ])[1]))
+ng.c <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "gene" & x$ifcausal == 1, , drop = F])[1]))
+ns.c <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "SNP" & x$ifcausal == 1, , drop = F])[1]))
+ngdf <- data.frame("ngene" = ng, "ncausalgene" = ng.c, "ncausalSNP" = ns.c)
+for (i in Niter){
+  resf <- paste0(outputdir, tag, ".", i, ".susieIres.Rd") 
+  load(resf)
+  ngdf[, paste("Prior.gene", i)] <- unlist(gene.rpiplist)
+}
+ngdf
+```
+
+# susieI shuffle genes
+Shuffle genes to different regions.
+
+Prior change for each iteration:
+```{r iter shuffle}
+tag <- '20200813-1-1-shuffleg'
+load(paste0(outputdir, tag, ".susieIres.Rd"))
+df <- cbind(prior.gene_rec, prior.SNP_rec)
+rownames(df) <- paste("Iteration", 1:Niter)
+colnames(df) <- c("Prior.gene", "Prior.SNP")
+df
+```
+
+Results for each region:
+```{r regions shuffle}
+regionlist <- readRDS(paste0(outputdir, tag, "_regionlist.rds"))
+ng <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "gene", , drop = F ])[1]))
+ng.c <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "gene" & x$ifcausal == 1, , drop = F])[1]))
+ns.c <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "SNP" & x$ifcausal == 1, , drop = F])[1]))
+ngdf <- data.frame("ngene" = ng, "ncausalgene" = ng.c, "ncausalSNP" = ns.c)
+for (i in 1:Niter){
+  resf <- paste0(outputdir, tag, ".", i, ".susieIres.Rd") 
+  load(resf)
+  ngdf[, paste("Prior.gene", i)] <- unlist(gene.rpiplist)
+}
+ngdf
+```

### 2020-08-19 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
susieI

@@
 ngdf <- data.frame("ngene" = ng, "ncausalgene" = ng.c, "ncausalSNP" = ns.c)
-for (i in Niter){
+for (i in 1:Niter){
   resf <- paste0(outputdir, tag, ".", i, ".susieIres.Rd") 

### 2020-08-19 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
susieI bug fix

--- 2020-08-19 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
@@
 ng <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "gene", , drop = F ])[1]))
+ns <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "SNP", , drop = F ])[1]))
 ng.c <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "gene" & x$ifcausal == 1, , drop = F])[1]))
 ns.c <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "SNP" & x$ifcausal == 1, , drop = F])[1]))
-ngdf <- data.frame("ngene" = ng, "ncausalgene" = ng.c, "ncausalSNP" = ns.c)
-for (i in Niter){
+ngdf <- data.frame("ngene" = ng, "nsnp" = ns, "ncausalgene" = ng.c, "ncausalSNP" = ns.c)
+
+cat("causal gene fraction:", sum(ngdf$ncausalgene)/sum(ngdf$ngene) )
+cat("causal snp fraction:", sum(ngdf$ncausalSNP)/sum(ngdf$ns))
+
+for (i in 1:Niter){
   resf <- paste0(outputdir, tag, ".", i, ".susieIres.Rd") 
   load(resf)
-  ngdf[, paste("Prior.gene", i)] <- unlist(gene.rpiplist)
+  ngdf[, paste("Prior.gene", i)] <- unlist(lapply(gene.rpiplist, "mean"))
+  ngdf[, paste("Prior.SNP", i)] <- unlist(lapply(snp.rpiplist, "mean"))
 }
@@
 ng <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "gene", , drop = F ])[1]))
+ns <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "SNP", , drop = F ])[1]))
 ng.c <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "gene" & x$ifcausal == 1, , drop = F])[1]))
 ns.c <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "SNP" & x$ifcausal == 1, , drop = F])[1]))
-ngdf <- data.frame("ngene" = ng, "ncausalgene" = ng.c, "ncausalSNP" = ns.c)
+ngdf <- data.frame("ngene" = ng, "nsnp" = ns, "ncausalgene" = ng.c, "ncausalSNP" = ns.c)
+
+cat("causal gene fraction:", sum(ngdf$ncausalgene)/sum(ngdf$ngene) )
+cat("causal snp fraction:", sum(ngdf$ncausalSNP)/sum(ngdf$ns))
+
 for (i in 1:Niter){
@@
   load(resf)
-  ngdf[, paste("Prior.gene", i)] <- unlist(gene.rpiplist)
+  ngdf[, paste("Prior.gene", i)] <- unlist(lapply(gene.rpiplist, "mean"))
+  ngdf[, paste("Prior.SNP", i)] <- unlist(lapply(snp.rpiplist, "mean"))
 }

### 2020-08-19 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
more iterations

@@
 tag <- '20200813-1-1'
-Niter <- 10
+Niter <- 50
 ```
@@
 
-for (i in 1:Niter){
+for (i in seq(1, Niter, 10)){
   resf <- paste0(outputdir, tag, ".", i, ".susieIres.Rd") 

### 2020-08-21 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
more parameter examples for susieI

@@
 susiedir <- "~/causalTWAS/simulations/simulation_susieI_20200813/"
+mrashdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
 tag <- '20200813-1-1'
-Niter <- 50
+Niter <- 30
 ```
@@
 
+## One example
+
 Prior change for each iteration:
@@
 
+## more examples
+
+Each row represents one simulation run. Runs labeled `1-x` have the same simulation parameters, as you can tell from the true parameters. (and so are runs labled `2-x` and `3-x`)
+
+```{r param}
+show_param <- function(tags){
+  mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt" )
+  param <- do.call(rbind, lapply(mr.ashfs, function(x) t(read.table(x))[2:1,]))
+
+  pars <- cbind(param[seq(1,nrow(param), 2), 1], param[seq(2,nrow(param), 2), 1], param[seq(1,nrow(param), 2), 3], param[seq(2,nrow(param), 2), 3])
+  colnames(pars) <- paste(rep(c("pi1.gene_", "pi1.SNP_"), each = 2), c("truth", "mr.ash2"), sep = "")
+
+  susieIfs <- paste0(outputdir, "20200813-", tags, ".susieIres.Rd" )
+  param.s <- do.call(rbind, lapply(susieIfs, function(x) {load(x); c(tail(prior.gene_rec, 1), tail(prior.SNP_rec,1))}))
+
+  susieIfs <- paste0(outputdir, "20200813-", tags, ".1.txt")
+  param.s.truth <- do.call(rbind, lapply(susieIfs, function(x) {
+    a <- read.table(x, header = T);
+    c(nrow(a[a$ifcausal == 1 & a$type == "gene", , drop = F])/ nrow(a[a$type == "gene", , drop = F]),
+      nrow(a[a$ifcausal == 1 & a$type == "SNP", , drop = F])/ nrow(a[a$type == "SNP", , drop = F]))
+    }))
+
+  pars.s <- cbind(param.s.truth, param.s)[, c(1,3,2,4)]
+  colnames(pars.s) <-  paste(rep(c("pi1.gene_", "pi1.SNP_"), each = 2), c("susietruth", "susieI"), sep = "")
+
+  df <- cbind(pars, pars.s)
+  rownames(df) <- tags
+
+  df %>% 
+  kable(format = "html", col.names = colnames(df)) %>%
+  kable_styling() %>%
+  kableExtra::scroll_box(width = "100%")
+
+}
+
+tags <- paste(rep(1:3, each = 5), 1:5, sep ="-")
+
+show_param(tags)
+```
+
 # susieI shuffle genes
+# one example
 Shuffle genes to different regions.
@@
 df <- cbind(prior.gene_rec, prior.SNP_rec)
-rownames(df) <- paste("Iteration", 1:Niter)
+rownames(df) <- paste("Iteration", 1:nrow(df))
 colnames(df) <- c("Prior.gene", "Prior.SNP")
@@
 
-cat("causal gene fraction:", sum(ngdf$ncausalgene)/sum(ngdf$ngene) )
+cat("causal gene fraction:", sum(ngdf$ncausalgene)/sum(ngdf$ngene))
 cat("causal snp fraction:", sum(ngdf$ncausalSNP)/sum(ngdf$ns))

### 2020-08-28 analysis/index.Rmd
visual chr17to22
susieI
susieI all regions

@@
 
+# Test parameters for susie
+ * [Using UKbiobank data (chr17:22), fusion weights Adipose, test prior's effect for susie](simulation-multi-ukbchr17to22-gtex.adipose_susieprior.html)
+
 ## Summarize cTWAS results
@@
   * [Using UKbiobank data (chr17:22), fusion weights Adipose, using smaller sa2 grid in mr.ash](simulation-multi-ukbchr17to22-gtex.adipose_sa2_v2.html)
+
+## mr.ash + SUSIE iterative
+*  [Using UKbiobank data (chr17:22), fusion weights Adipose](simulation-susieI-ukbchr17to22-gtex.adipose.html)
 

### 2020-08-28 analysis/simulation-multi-ukbchr17to22-gtex.adipose_susieprior.Rmd
susieI all regions

---
title: "Summarize twas: chr17 to 22"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  message= F, warning = F)
```
`

* Run susie with different priors and see how much prior affects results.

* Data: ukb chr 17 to chr 22 combined. SNPs are downsampled to 1/10, eQTLs defined by FUSION-TWAS (Adipose/GTEx) lasso effect size > 0 were kept, p= 86k, n = 20k.

```{r plot functions}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})
library(tidyr)
library(plyr)
library(stringr)
library(kableExtra)
source("analysis/summarize_twas_plots.R")
```

# PIP calibration: filter regions

```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/20200721-1-fixprior_rpip0.5/"
```

We run 100 simulations and run susie using different priors and `L= 1`. We only apply susie for regions with mr.ash2 regional PIP > 0.5.

## Truth

* The true parameters we used to simulate data: 
```{r  true par}
par <- read.table(paste0(susiedir, "20200721-1-fixedprior_1.txt"), header = T)
t(par[,2, drop = F])
```

## Using uniform prior:

```{r cali 0, fig.height=6, fig.width=6}
pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior1.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))

cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
```

## Use overestimated prior

* The prior used is: 
```{r  par1}
par <- read.table(paste0(susiedir, "20200721-1-fixedprior_1.txt"), header = T)
t(par[c(1,3),1, drop = F])
```
* susie results: 
```{r cali 1, fig.height=6, fig.width=14}
par(mfrow=c(1,2))
cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
```

## Use priors close to truth

* The prior used is: 
```{r  par2}
par <- read.table(paste0(susiedir, "20200721-1-fixedprior_2.txt"), header = T)
t(par[c(1,3),1, drop = F])
```

```{r cali 2, fig.height=6, fig.width=14}
pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior2.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))

par(mfrow=c(1,2))
cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
```

## Use under estimated priors

* The prior used is: 
```{r  par3}
par <- read.table(paste0(susiedir, "20200721-1-fixedprior_3.txt"), header = T)
t(par[c(1,3),1, drop = F])
```

* susie results: 
```{r cali 3, fig.height=6, fig.width=14}
pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior3.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))

par(mfrow=c(1,2))
cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
```

# PIP calibration: use all regions

```{r output dir all}
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
```

We run 100 simulations and run susie using different priors and `L= 1`. We only apply susie for all regions.

## Truth

* The true parameters we used to simulate data: 
```{r  true par all}
par <- read.table(paste0(susiedir, "20200721-1-fixedprior_1.txt"), header = T)
t(par[,2, drop = F])
```

## Using uniform prior:

```{r cali 0 all, fig.height=6, fig.width=6}
pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior1.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))

cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
```

## Use overestimated prior

* The prior used is: 
```{r  par1 all}
par <- read.table(paste0(susiedir, "20200721-1-fixedprior_1.txt"), header = T)
t(par[c(1,3),1, drop = F])
```
* susie results: 
```{r cali 1 all, fig.height=6, fig.width=14}
par(mfrow=c(1,2))
cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
```

## Use priors close to truth

* The prior used is: 
```{r  par2 all}
par <- read.table(paste0(susiedir, "20200721-1-fixedprior_2.txt"), header = T)
t(par[c(1,3),1, drop = F])
```

```{r cali 2 all, fig.height=6, fig.width=14}
pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior2.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))

par(mfrow=c(1,2))
cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
```

## Use under estimated priors

* The prior used is: 
```{r  par3 all}
par <- read.table(paste0(susiedir, "20200721-1-fixedprior_3.txt"), header = T)
t(par[c(1,3),1, drop = F])
```

* susie results: 
```{r cali 3 all, fig.height=6, fig.width=14}
pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior3.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))

par(mfrow=c(1,2))
cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
```

### 2020-08-28 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
susieI all regions

@@
 
-# susieI genes
+# susieI genes - setting 1
 
-Regions: We use regional PIP of both SNP and gene from mr.ash2s > 0.3 , and also require at least one causal gene or SNP is present, also, at least one gene.
+* Regions: We use regional PIP of both SNP and gene from mr.ash2s > 0.3, and also require at least one causal gene or SNP is present, also, at least one gene.
+
+* Susie run parameters: `L=1`, `null_weight = 0`, didn't update `null_weight` in iterations.
 
@@
 
-Results for each region:
-```{r regions}
-regionlist <- readRDS(paste0(outputdir, tag, "_regionlist.rds"))
-ng <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "gene", , drop = F ])[1]))
-ns <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "SNP", , drop = F ])[1]))
-ng.c <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "gene" & x$ifcausal == 1, , drop = F])[1]))
-ns.c <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "SNP" & x$ifcausal == 1, , drop = F])[1]))
-ngdf <- data.frame("ngene" = ng, "nsnp" = ns, "ncausalgene" = ng.c, "ncausalSNP" = ns.c)
-
-cat("causal gene fraction:", sum(ngdf$ncausalgene)/sum(ngdf$ngene) )
-cat("causal snp fraction:", sum(ngdf$ncausalSNP)/sum(ngdf$ns))
-
-for (i in seq(1, Niter, 10)){
-  resf <- paste0(outputdir, tag, ".", i, ".susieIres.Rd") 
-  load(resf)
-  ngdf[, paste("Prior.gene", i)] <- unlist(lapply(gene.rpiplist, "mean"))
-  ngdf[, paste("Prior.SNP", i)] <- unlist(lapply(snp.rpiplist, "mean"))
-}
-ngdf
-```
-
 ## more examples
@@
 ```{r param}
-show_param <- function(tags){
-  mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt" )
+show_param <- function(mrashfs, susieIfs, susieIfs2){
   param <- do.call(rbind, lapply(mr.ashfs, function(x) t(read.table(x))[2:1,]))
@@
 
-  susieIfs <- paste0(outputdir, "20200813-", tags, ".susieIres.Rd" )
   param.s <- do.call(rbind, lapply(susieIfs, function(x) {load(x); c(tail(prior.gene_rec, 1), tail(prior.SNP_rec,1))}))
 
-  susieIfs <- paste0(outputdir, "20200813-", tags, ".1.txt")
-  param.s.truth <- do.call(rbind, lapply(susieIfs, function(x) {
+  param.s.truth <- do.call(rbind, lapply(susieIfs2, function(x) {
     a <- read.table(x, header = T);
@@
 
-  df <- cbind(pars, pars.s)
-  rownames(df) <- tags
+  df <- cbind(tags, format(pars, digits = 4), format(pars.s, digits =4))
+  rownames(df) <- NULL
 
   df %>% 
-  kable(format = "html", col.names = colnames(df)) %>%
-  kable_styling() %>%
-  kableExtra::scroll_box(width = "100%")
+  kable("html", escape = F) %>%
+  kable_styling("striped", full_width = F) %>%
+  scroll_box(width = "100%")
 
@@
 
-tags <- paste(rep(1:3, each = 5), 1:5, sep ="-")
+tags <- paste(rep(1:3, each = 5), 1:5, sep = "-")
+mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".1.txt")
 
-show_param(tags)
+show_param(mrashfs, susieIfs, susieIfs2)
+```
+
+# susieI genes - setting 2
+
+* Regions: All regions.
+
+* Susie run parameters: `L=1`. Initiate with `null_weight = 0` and update `null_weight` based on last iteration results.
+
+## One example
+
+Prior change for each iteration:
+```{r iter2}
+load(paste0(outputdir, tag, ".config1.susieIres.Rd"))
+df <- cbind(prior.gene_rec, prior.SNP_rec)
+rownames(df) <- paste("Iteration", 1:Niter)
+colnames(df) <- c("Prior.gene", "Prior.SNP")
+df
+```
+
+## More examples
+
+```{r param2}
+mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".config1.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config1.susieI.txt")
+
+show_param(mrashfs, susieIfs, susieIfs2)
+```
+
+# susieI genes - setting 3
+
+* Regions: All regions.
+
+* Susie run parameters: `L=2`. Initiate with `null_weight = 0` and update `null_weight` based on last iteration results.
+
+## One example
+
+Prior change for each iteration:
+```{r iter3}
+load(paste0(outputdir, tag, ".config2.susieIres.Rd"))
+df <- cbind(prior.gene_rec, prior.SNP_rec)
+rownames(df) <- paste("Iteration", 1:Niter)
+colnames(df) <- c("Prior.gene", "Prior.SNP")
+df
+```
+
+## More examples
+
+```{r param3}
+mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".config2.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config2.susieI.txt")
+
+show_param(mrashfs, susieIfs, susieIfs2)
+```
+
+# susieI genes - setting 4
+
+* Regions: All regions.
+
+* Susie run parameters: `L=1`. Did not update `null_weight` in each iteration.
+
+## One example
+
+Prior change for each iteration:
+```{r iter4}
+load(paste0(outputdir, tag, ".config3.susieIres.Rd"))
+df <- cbind(prior.gene_rec, prior.SNP_rec)
+rownames(df) <- paste("Iteration", 1:Niter)
+colnames(df) <- c("Prior.gene", "Prior.SNP")
+df
+```
+
+## More examples
+
+```{r param4}
+mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".config3.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config3.susieI.txt")
+
+show_param(mrashfs, susieIfs, susieIfs2)
 ```
@@
 # susieI shuffle genes
-# one example
-Shuffle genes to different regions.
+## one example
+* Shuffle genes to different regions. Regions: We use regional PIP of both SNP and gene from mr.ash2s > 0.3 , and also require at least one causal gene or SNP is present, also, at least one gene.
+
+* Susie run parameters: `L=1`, `null_weight = 0`.
 
@@
 ```
-
-Results for each region:
-```{r regions shuffle}
-regionlist <- readRDS(paste0(outputdir, tag, "_regionlist.rds"))
-ng <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "gene", , drop = F ])[1]))
-ns <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "SNP", , drop = F ])[1]))
-ng.c <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "gene" & x$ifcausal == 1, , drop = F])[1]))
-ns.c <- unlist(lapply(lapply(regionlist, '[[', "anno"), function(x) dim(x[x$type == "SNP" & x$ifcausal == 1, , drop = F])[1]))
-ngdf <- data.frame("ngene" = ng, "nsnp" = ns, "ncausalgene" = ng.c, "ncausalSNP" = ns.c)
-
-cat("causal gene fraction:", sum(ngdf$ncausalgene)/sum(ngdf$ngene))
-cat("causal snp fraction:", sum(ngdf$ncausalSNP)/sum(ngdf$ns))
-
-for (i in seq(1, Niter, 10)){
-  resf <- paste0(outputdir, tag, ".", i, ".susieIres.Rd") 
-  load(resf)
-  ngdf[, paste("Prior.gene", i)] <- unlist(lapply(gene.rpiplist, "mean"))
-  ngdf[, paste("Prior.SNP", i)] <- unlist(lapply(snp.rpiplist, "mean"))
-}
-ngdf
-```

### 2020-09-01 analysis/simulation-multi-ukbchr17to22-gtex.adipose_susieprior.Rmd
increase gene pve

@@
 
-# PIP calibration: filter regions
+# Setting 1
 
@@
 outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
-susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/20200721-1-fixprior_rpip0.5/"
+susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
 ```
 
-We run 100 simulations and run susie using different priors and `L= 1`. We only apply susie for regions with mr.ash2 regional PIP > 0.5.
+We run 50 simulations and run susie using different priors and `L= 1`. We apply susie for all regions.
 
@@
 
-# PIP calibration: use all regions
+# Setting 2
 
-```{r output dir all}
-susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
-```
-
-We run 100 simulations and run susie using different priors and `L= 1`. We only apply susie for all regions.
+We run 100 simulations and run susie using different priors and `L= 1`. We apply susie for all regions.
 
@@
 ```{r  true par all}
-par <- read.table(paste0(susiedir, "20200721-1-fixedprior_1.txt"), header = T)
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_1.txt"), header = T)
 t(par[,2, drop = F])
@@
 ```{r cali 0 all, fig.height=6, fig.width=6}
-pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior1.L1.susieres.expr.txt"))
+pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior1.L1.susieres.expr.txt"))
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
@@
 ```{r  par1 all}
-par <- read.table(paste0(susiedir, "20200721-1-fixedprior_1.txt"), header = T)
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_1.txt"), header = T)
 t(par[c(1,3),1, drop = F])
@@
 ```{r  par2 all}
-par <- read.table(paste0(susiedir, "20200721-1-fixedprior_2.txt"), header = T)
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
 t(par[c(1,3),1, drop = F])
@@
 ```{r cali 2 all, fig.height=6, fig.width=14}
-pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior2.L1.susieres.expr.txt"))
+pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.L1.susieres.expr.txt"))
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
@@
 ```{r  par3 all}
-par <- read.table(paste0(susiedir, "20200721-1-fixedprior_3.txt"), header = T)
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_3.txt"), header = T)
 t(par[c(1,3),1, drop = F])
@@
 ```{r cali 3 all, fig.height=6, fig.width=14}
-pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior3.L1.susieres.expr.txt"))
+pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior3.L1.susieres.expr.txt"))
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))

### 2020-09-01 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
increase gene pve

@@
 
-* Susie run parameters: `L=1`, `null_weight = 0`, didn't update `null_weight` in iterations.
+* Susie run parameters: `L=1`, `null_weight = 0`, didn't update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform.
 
@@
 
-  pars <- cbind(param[seq(1,nrow(param), 2), 1], param[seq(2,nrow(param), 2), 1], param[seq(1,nrow(param), 2), 3], param[seq(2,nrow(param), 2), 3])
-  colnames(pars) <- paste(rep(c("pi1.gene_", "pi1.SNP_"), each = 2), c("truth", "mr.ash2"), sep = "")
+  pars <- cbind(param[seq(1,nrow(param), 2), 2], param[seq(1,nrow(param), 2), 4], param[seq(1,nrow(param), 2), 1], param[seq(2,nrow(param), 2), 1], param[seq(1,nrow(param), 2), 3], param[seq(2,nrow(param), 2), 3])
+  colnames(pars) <- c("PVE.gene_truth", "PVE.SNP_truth", paste(rep(c("pi1.gene_", "pi1.SNP_"), each = 2), c("truth", "mr.ash2"), sep = ""))
 
@@
   kable_styling("striped", full_width = F) %>%
-  scroll_box(width = "100%")
-
+   row_spec(c(1:5, 11:15), background = "#FEF3B9") %>%
+  scroll_box(width = "100%", height = "600px", fixed_thead = T)
 }
@@
 
-* Susie run parameters: `L=1`. Initiate with `null_weight = 0` and update `null_weight` based on last iteration results.
+* Susie run parameters: `L=1`. initialize with `null_weight = 0` and update `null_weight` based on last iteration results. We initialize with prior for genes and SNPs as uniform.
 
@@
 ```{r param2}
+tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
 mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
@@
 
-* Susie run parameters: `L=2`. Initiate with `null_weight = 0` and update `null_weight` based on last iteration results.
+* Susie run parameters: `L=2`. initialize with `null_weight = 0` and update `null_weight` based on last iteration results. We initialize with prior for genes and SNPs as uniform.
 
@@
 
-* Susie run parameters: `L=1`. Did not update `null_weight` in each iteration.
+* Susie run parameters: `L=1`. Did not update `null_weight` in each iteration. We initialize with prior for genes and SNPs as uniform.
 
@@
 
+# susieI genes - setting 5
+
+* Regions: All regions.
+
+* Susie run parameters: `L=2`. Update `null_weight` in each iteration. We initialize prior for gene as `0.05` and prior for SNPs as `2.5E-3`, we initialize `null_weight` as 0.
+
+## One example
+
+Prior change for each iteration:
+```{r iter5}
+load(paste0(outputdir, tag, ".config4.susieIres.Rd"))
+df <- cbind(prior.gene_rec, prior.SNP_rec)
+rownames(df) <- paste("Iteration", 1:Niter)
+colnames(df) <- c("Prior.gene", "Prior.SNP")
+df
+```
+
+## More examples
+
+```{r param5}
+mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".config4.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config4.susieI.txt")
+
+show_param(mrashfs, susieIfs, susieIfs2)
+```
+
 # susieI shuffle genes

### 2020-09-03 analysis/index.Rmd
susie prior

--- 2020-08-05 analysis/index.Rmd
@@
 
-  * [Using UKbiobank data (chr17 to 22), fusion weights Adipose subcutaneous](https://users.rcc.uchicago.edu/~simingz/simulation-mr.ash2s-ukbchr17to22-gtex.adipose.html)
-
   * [Using UKbiobank data (chr22), weights of H3K27 from Xiaotong](https://users.rcc.uchicago.edu/~simingz/simulation-mr.ash2s-ukbchr22-H3K27.html)
@@
 
+# Test parameters for susie
+ * [Using UKbiobank data (chr17:22), fusion weights Adipose, test prior's effect for susie](simulation-multi-ukbchr17to22-gtex.adipose_susieprior.html)
+    * [Visualization](https://users.rcc.uchicago.edu/~simingz/simulation-susieP-ukbchr17to22-gtex.adipose.html)
+
 ## Summarize cTWAS results
@@
   * [Using UKbiobank data (chr17:22), fusion weights Adipose, using smaller sa2 grid in mr.ash](simulation-multi-ukbchr17to22-gtex.adipose_sa2_v2.html)
+      *   [Visualization](https://users.rcc.uchicago.edu/~simingz/simulation-mr.ash2s-ukbchr17to22-gtex.adipose.html)
+
+## mr.ash + SUSIE iterative
+*  [Using UKbiobank data (chr17:22), fusion weights Adipose](simulation-susieI-ukbchr17to22-gtex.adipose.html)
 

### 2020-09-03 analysis/simulation-multi-ukbchr17to22-gtex.adipose_susieprior.Rmd
susie prior

@@
 
+## PIP calibration: filter regions
+
 ```{r output dir}
@@
 outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
-susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
+susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/20200721-1-fixprior_rpip0.5/"
 ```
 
-We run 50 simulations and run susie using different priors and `L= 1`. We apply susie for all regions.
+We run 100 simulations and run susie using different priors and `L= 1`. We only apply susie for regions with mr.ash2 regional PIP > 0.5.
 
-## Truth
+### Truth
 
@@
 
-## Using uniform prior:
+### Using uniform prior:
 
@@
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-
 cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
@@
 
-## Use overestimated prior
+### Use overestimated prior
 
@@
 
-## Use priors close to truth
+### Use priors close to truth
 
@@
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-
 par(mfrow=c(1,2))
@@
 
-## Use under estimated priors
+### Use under estimated priors
 
@@
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+par(mfrow=c(1,2))
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
+```
+
+## PIP calibration: all regions
+
+```{r output a dir}
+susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
+```
+
+We run 50 simulations and run susie using different priors and `L= 1`. We apply susie for all regions.
+
+### Truth
+
+* The true parameters we used to simulate data: 
+```{r  true a par}
+par <- read.table(paste0(susiedir, "20200721-1-fixedprior_1.txt"), header = T)
+t(par[,2, drop = F])
+```
+
+### Using uniform prior:
+
+```{r cali 0 a, fig.height=6, fig.width=6}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior1.L1.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+
+cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
+```
+
+### Use overestimated prior
+
+* The prior used is: 
+```{r  par1 a}
+par <- read.table(paste0(susiedir, "20200721-1-fixedprior_1.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+```
+* susie results: 
+```{r cali 1 a, fig.height=6, fig.width=14}
+par(mfrow=c(1,2))
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
+```
+
+### Use priors close to truth
+
+* The prior used is: 
+```{r  par2 a}
+par <- read.table(paste0(susiedir, "20200721-1-fixedprior_2.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+```
+
+```{r cali 2 a, fig.height=6, fig.width=14}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior2.L1.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+
+par(mfrow=c(1,2))
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
+```
+
+### Use under estimated priors
+
+* The prior used is: 
+```{r  par3 a}
+par <- read.table(paste0(susiedir, "20200721-1-fixedprior_3.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+```
+
+* susie results: 
+```{r cali 3 a, fig.height=6, fig.width=14}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior3.L1.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
 
@@
 
-We run 100 simulations and run susie using different priors and `L= 1`. We apply susie for all regions.
+## PIP calibration: filter regions
 
-## Truth
+```{r output dir-2}
+susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/20200721-3-fixprior_causal/"
+```
+
+We run 100 simulations and run susie using different priors and `L= 1`. We only apply susie for regions with at least one causal signal.
+
+### Truth
+
+* The true parameters we used to simulate data: 
+```{r  true par-2}
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_1.txt"), header = T)
+t(par[,2, drop = F])
+```
+
+### Using uniform prior:
+
+```{r cali 0-2, fig.height=6, fig.width=6}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior1.L1.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
+```
+
+### Use overestimated prior
+
+* The prior used is: 
+```{r  par1-2}
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_1.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+```
+* susie results: 
+```{r cali 1-2, fig.height=6, fig.width=14}
+par(mfrow=c(1,2))
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
+```
+
+### Use priors close to truth
+
+* The prior used is: 
+```{r  par2-2}
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+```
+* susie results: 
+```{r cali 2-2, fig.height=6, fig.width=14}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.L1.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+par(mfrow=c(1,2))
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
+```
+
+### Use under estimated priors
+
+* The prior used is: 
+```{r  par3-2}
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_3.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+```
+
+* susie results: 
+```{r cali 3-2, fig.height=6, fig.width=14}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior3.L1.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+par(mfrow=c(1,2))
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
+```
+
+## PIP calibration: all regions
+
+```{r output dir all}
+susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
+```
+
+We run 100 simulations and run susie using different priors and `L= 1`. We only apply susie for all regions.
+
+### Truth
 
@@
 ```
-
+* susie results: 
 ```{r cali 2 all, fig.height=6, fig.width=14}

### 2020-09-03 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
susie prior

@@
 
-* Susie run parameters: `L=2`. Update `null_weight` in each iteration. We initialize prior for gene as `0.05` and prior for SNPs as `2.5E-3`, we initialize `null_weight` as 0.
+* Susie run parameters: `L=5`. Update `null_weight` in each iteration. We initialize prior for gene as `0.05` and prior for SNPs as `2.5E-3`, we initialize `null_weight` as 0.
 
@@
 
+# susieI genes - setting 6
+
+* Regions: All regions.
+
+* Susie run parameters: `L=2`. Update `null_weight` in each iteration. We initialize prior for gene as `0.05` and prior for SNPs as `2.5E-3`, we initialize `null_weight` as 0.
+
+## One example
+
+Prior change for each iteration:
+```{r iter6}
+load(paste0(outputdir, tag, ".config5.susieIres.Rd"))
+df <- cbind(prior.gene_rec, prior.SNP_rec)
+rownames(df) <- paste("Iteration", 1:Niter)
+colnames(df) <- c("Prior.gene", "Prior.SNP")
+df
+```
+
+## More examples
+
+```{r param6}
+mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".config5.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config5.susieI.txt")
+
+show_param(mrashfs, susieIfs, susieIfs2)
+```
+
 # susieI shuffle genes

### 2020-09-03 analysis/simulation-susieP-ukbchr17to22-gtex.adipose.Rmd
susie prior

--- 2020-08-06 analysis/simulation-mr.ash2s-ukbchr17to22-gtex.adipose.Rmd
@@
 
-```{r plot functions}
-library(mr.ash.alpha)
-library(data.table)
-suppressMessages({library(plotly)})
-
-source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")
-
-plot_pip <- function(dt){
-
-  pal <- c("salmon", "darkgreen")
-  pal <- setNames(pal, c("Causal", "Non causal"))
-
-  if ( "max_r2" %in% colnames(dt)) {
-    fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', 
-                   mode = 'markers', color = ~ ifcausal, colors = pal,
-                   text = ~ paste("Name: ", name, 
-                                  "\nmax R^2: ", round(max_r2, digits = 2)))
-                                  #,"\nMax pair: ", max_pair))
-  } else {
-    fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))
-  }
-
-  fig <- fig %>% layout(yaxis = list(title = "PIP"))
-  return(fig)
-}
-
-plot_beta <- function(dt){
-  fig <- plot_ly(dt, x = ~p0, y = ~ TrueBeta, type = 'scatter', mode = 'markers', name ="Truth", text = ~paste("Name: ", name))
-  fig <- fig %>% add_trace(y = ~ EstBeta, name = 'Estimated',mode = 'markers')
-
-  fig <- fig %>% layout(yaxis = list(title = "beta"))
-  return(fig)
-}
-
-plot_p <- function(dt){
-  pal <- c("salmon", "darkgreen")
-  pal <- setNames(pal, c("Causal", "Non causal"))
-
-  fig <- plot_ly(dt, x = ~p0, y = ~ PVALUE , type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))
-
-  fig <- fig %>% layout(yaxis = list(title = "p value"))
-  return(fig)
-}
-
-plot_susie_pip <- function(dt) {
-   pal <- c("salmon", "darkgreen")
-   pal <- setNames(pal, c("Causal", "Non causal"))
-
-   fig <- plot_ly(dt, x = ~ p0, y = ~ susiePIP , type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))
-
-   fig <- fig %>% layout(yaxis = list(title = "SUSIE PIP"))
-}
-
-plot_all <- function(mrashres.gene, gwasres.gene, mrashres.snp, gwasres.snp, susieres.gene, susieres.snp, ldres = NULL, chrom=NULL){
-  a <- read.table(mrashres.gene , header =T)
-  a[a$ifcausal == 1, "ifcausal"] = "Causal"
-  a[a$ifcausal == 0, "ifcausal"] = "Non causal"
-
-  if (!is.null(ldres)){
-    ld <- read.table(ldres, header =T)
-    a <- merge(a, ld, by = "name", all.x = T)}
-
-  su <- read.table(susieres.gene, header =T)
-  colnames(su)[colnames(su) == "pip"] <-  "susiePIP"
-  a <- merge(a, su, by = "name", all.x = T)
-
-  p <- fread(gwasres.gene, header =T)
-  p[,"PVALUE"] <- -log10(p[,"PVALUE"])
-  colnames(p)[colnames(p) == "MARKER_ID"] <-  "name"
-  dt.gene <- merge(a, p, by = "name", all.x = T)
-
-  a <- read.table(mrashres.snp , header =T)
-  a[a$ifcausal == 1, "ifcausal"] = "Causal"
-  a[a$ifcausal == 0, "ifcausal"] = "Non causal"
-
-  su <- read.table(susieres.snp, header =T)
-  colnames(su)[colnames(su) == "pip"] <-  "susiePIP"
-  a <- merge(a, su, by = "name", all.x = T)
-
-  p <- fread(gwasres.snp, header =T)
-  p[,"PVALUE"] <- -log10(p[,"PVALUE"])
-  colnames(p)[colnames(p) == "MARKER_ID"] <-  "name"
-  dt.snp <- merge(a, p, by = "name", all.x = T)
-
-  if (!is.null(chrom)){
-    dt.gene <- dt.gene[dt.gene$chr.x == chrom,]
-    dt.snp <- dt.snp[dt.snp$chr.x == chrom,]
-  }
-
-  fig1 <- plot_pip(dt.gene)  
-  fig2 <- plot_beta(dt.gene) 
-  fig3 <- plot_p(dt.gene) 
-  fig4 <- plot_susie_pip(dt.gene) 
-  fig5 <- plot_pip(dt.snp) 
-  fig6 <- plot_beta(dt.snp) 
-  fig7 <- plot_p(dt.snp)
-  fig8 <- plot_susie_pip(dt.snp) %>% layout( plot_bgcolor='#EFF6FC')
-
-  fig <- subplot(fig1, fig2, fig3, fig4, fig5, fig6, fig7, fig8, nrows=8, shareX = T, titleY = T)
-
-  fig <- fig %>% layout(autosize = F, width = 600, height = 1100, margin = 0.02,
-    xaxis = list(title = "Genomic position"))
-
-  fig <- fig %>% toWebGL()
-  fig
-}
-
-```
-
 ```{r output dir}
@@
 
-```{r getfile function}
-get_files <- function(tag, tag2){
-  par <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt")
-  gmrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt")
-  ggwas <- paste0(outputdir, tag, ".exprgwas.txt.gz")
-  smrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt")
-  sgwas <- paste0(outputdir, tag, ".snpgwas.txt.gz")
-  gsusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.expr.txt")
-  ssusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.snp.txt")
-  rsq <- paste0(outputdir, tag, ".LD.txt")
-  return(tibble::lst(par, gmrash, ggwas, smrash, sgwas, gsusie, ssusie, rsq))
-}
+```{r source file}
+source("analysis/visual_twas_plots.R")
 ```

### 2020-09-03 code/run_test_susie.R

@@ -25,6 +25,8 @@ loginfo('script started ... ')
 
 PIPfilter <- 0
 chunksize = 500000
+ifcausal <- T
+L = 1
 
 loginfo("regional PIP cut: %s ", PIPfilter)
 loginfo("region size: %s", chunksize)
@@ -43,6 +45,9 @@ dat$expr <-  scaleRcpp(exprres$expr)
 # load phenotype Rd file, variable: phenores
 load(args[3])
 phenores$Y <-  phenores$Y - mean(phenores$Y)
+
+# causal signal names
+cau <- c(dat$snp[phenores$param$idx.cSNP,], colnames(exprres$expr)[phenores$param$idx.cgene])
 
 # run susie using given priors
 param <- read.table(args[4], header = T, row.names = 1)
@@ -65,7 +70,7 @@ for (chrom in chroms){
   p1 <- stpos + 1: ceiling((edpos - stpos)/chunksize) * chunksize
   itv <- cbind(p0, p1)
   rPIP <- apply(itv, 1, function(x) sum(pipres.chr[x[1] < pipres.chr$p0 & pipres.chr$p0 < x[2], "PIP"]))
-  nCausal <- apply(itv, 1, function(x) sum(pipres.chr[x[1] < pipres.chr$p0 & pipres.chr$p0 < x[2], "ifcausal"]))
+  nCausal <- apply(itv, 1, function(x) {gnames <- pipres.chr[x[1] < pipres.chr$p0 & pipres.chr$p0 < x[2], "name"]; sum(ifelse(gnames %in% cau, 1, 0))})
   itv <- cbind(chrom, itv, rPIP, nCausal)
   regions <- rbind(regions, itv)
 }
@@ -74,11 +79,16 @@ loginfo("No. intervals for chr %s : %s", chrom, nrow(regions))
 write.table(regions, file = paste0(args[5], ".rPIP.txt"), row.names=F, col.names=T, sep="\t", quote = F)
 
 regions <- regions[regions[, "rPIP"] > PIPfilter,  ]
+
+if (isTRUE(ifcausal)){
+  regions <- regions[regions[, "nCausal"] > 0,  ]
+}
+
 loginfo("No. intervals for chr %s after PIP filter: %s", chrom, nrow(regions))
 
 outname <- args[6]
 
-L = 1
+
 if (length(args) == 7){
   L = as.numeric(args[7])
   outname <- paste0(outname, ".L", L)
[…]

### 2020-09-04 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
susieI more runs

@@
 
-# susieI genes - setting 1
+# susieI (filtered regions) - setting 1
 
@@
 
-# susieI genes - setting 2
+# susieI (filtered regions) - setting 2
+
+* Regions: We use regional PIP of both SNP and gene from mr.ash2s > 0.3.
+
+* Susie run parameters: `L=2`, initialize `null_weight = 0`, update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform.
+
+## One example
+
+Prior change for each iteration:
+```{r iter-6}
+load(paste0(outputdir, tag, ".config6.susieIres.Rd"))
+df <- cbind(prior.gene_rec, prior.SNP_rec)
+rownames(df) <- paste("Iteration", 1:Niter)
+colnames(df) <- c("Prior.gene", "Prior.SNP")
+df
+```
+
+## more examples
+
+Each row represents one simulation run. Runs labeled `1-x` have the same simulation parameters, as you can tell from the true parameters. (and so are runs labled `2-x` and `3-x`)
+
+```{r param-6}
+tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
+mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".config6.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config6.susieI.txt")
+
+show_param(mrashfs, susieIfs, susieIfs2)
+```
+
+# susieI (filtered regions) - setting 3
+
+* Regions: We use regional PIP of both SNP and gene from mr.ash2s > 0.3.
+
+* Susie run parameters: `L=2`, initialize `null_weight = 0`, update `null_weight` in iterations. We initialize with prior for genes and SNPs as 0.05 and 2.5e-3.
+
+## One example
+
+Prior change for each iteration:
+```{r iter-8}
+load(paste0(outputdir, tag, ".config8.susieIres.Rd"))
+df <- cbind(prior.gene_rec, prior.SNP_rec)
+rownames(df) <- paste("Iteration", 1:Niter)
+colnames(df) <- c("Prior.gene", "Prior.SNP")
+df
+```
+
+## more examples
+
+Each row represents one simulation run. Runs labeled `1-x` have the same simulation parameters, as you can tell from the true parameters. (and so are runs labled `2-x` and `3-x`)
+
+```{r param-8}
+tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".config8.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config8.susieI.txt")
+
+show_param(mrashfs, susieIfs, susieIfs2)
+```
+
+# susieI (filtered regions) - setting 4
+
+* Regions: We use regional PIP of both SNP and gene from mr.ash2s > 0.3.
+
+* Susie run parameters: `L=2`, `null_weight = 0`. Didn't update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform.
+
+## One example
+
+Prior change for each iteration:
+```{r iter-7}
+load(paste0(outputdir, tag, ".config7.susieIres.Rd"))
+df <- cbind(prior.gene_rec, prior.SNP_rec)
+rownames(df) <- paste("Iteration", 1:Niter)
+colnames(df) <- c("Prior.gene", "Prior.SNP")
+df
+```
+
+## more examples
+
+Each row represents one simulation run. Runs labeled `1-x` have the same simulation parameters, as you can tell from the true parameters. (and so are runs labled `2-x` and `3-x`)
+
+```{r param-7}
+tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".config7.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config7.susieI.txt")
+
+show_param(mrashfs, susieIfs, susieIfs2)
+```
+
+# susieI (all regions) - setting 1
 
@@
 
-# susieI genes - setting 3
+# susieI (all regions) - setting 2
 
@@
 
-# susieI genes - setting 4
+# susieI (all regions) - setting 3
 
@@
 
-# susieI genes - setting 5
+# susieI (all regions) - setting 4
 
@@
 
-# susieI genes - setting 6
+# susieI (all regions) - setting 5
 

### 2020-09-06 analysis/description.Rmd
pretty plot for Xin's grant

# Analysis description

```{r param}
n <- 20000 # number of samples
p <- 86000 # number of SNPs
```

The genotype data we used is from UKB biobank, randomly selecting n=`r as.character(n)` samples. We use SNP genotype data from chr 17 to chr 22 combined. These genomic regions represents 12.5% of the genome. SNPs are downsampled to 1/10 (randomly), eQTLs (see below for definition of eQTL) were added back. This ends up with p = `p as.charater(p)` SNPs.

Our analysis consists of the following steps:

1. Build expression predictors using another expression-genotype dataset.

The one we used in this analysis is GTEx Adipose tissue v7 dataset. This dataset contains ~ 380 samples. FUSION/TWAS were used train expression model and we used their lasso results. There are around 1700 genes with expression

### 2020-09-06 analysis/simulation-multi-ukbchr17to22-gtex.adipose_sa2_v2.Rmd
pretty plot for Xin's grant

--- 2020-08-06 analysis/simulation-multi-ukbchr17to22-gtex.adipose_sa2_v2.Rmd
@@
 tags_ext <- Reduce(intersect, get_tags(tagglob, tagextr, tag2 = tag2)['gsusie'])
-res <- caliPIP_plot(tags = tags_ext, tag2 = tag2)
+a <- caliPIP_plot(tags = tags_ext, tag2 = tag2)
 ```
 ```{r cali 1fdr, fig.height=4, fig.width=4}
-caliFDR_plot(tags = tags_ext, tag2 = tag2)
+a <- caliFDR_plot(tags = tags_ext, tag2 = tag2)
 ```
@@
 ```{r cali 2, fig.height=4, fig.width=14}
-caliPIP_plot(tags = tags_ext, tag2 = "lassoes-es")
+a <- caliPIP_plot(tags = tags_ext, tag2 = "lassoes-es")
 ```
@@
 ```{r cali 2fdr, fig.height=4, fig.width=4}
-caliFDR_plot(tags = tags_ext, tag2 = "lassoes-es")
+a <- caliFDR_plot(tags = tags_ext, tag2 = "lassoes-es")
 ```
@@
 
-## NULL; expr-snp; expr-snp 
+## Lasso; expr-snp; expr-snp 
 ```{r sca_p 1, fig.height=7, fig.width=7}
-scatter_plot_PIP_p(tags, tag2s[1])
+a <- scatter_plot_PIP_p(tags, "lassoes-es")
 ```
+
+# Pretty plots
+A) PIP calibration plot; B) FDP inflation of TWAS; C) Scatter plot of individual genes: PIPs vs. -log10-p values from TWAS.
+
+We use SNP genotype data from chr 17 to chr 22 combined.
+These genomic regions represents 12.5% of the genome.
+SNPs are downsampled to 1/10 (randomly), eQTLs (see below for definition of eQTL) were added back.
+We used lasso to train expression model using GTEx adipose tissue v7.
+In this simulation, we simulate phenotype with PVE explained by gene as 0.01, by SNP 0.05.
+pi1 for gene 0.05, pi1 for SNP 0.0025.
+
+For our method, we run mr.ash2s for both genes and SNPs (initiated with lasso). Then apply SUSIE for regions with sum of PIP > 0.5. This gives us PIP for individual genes.
+
+For TWAS, we use linear regression for each gene and obtain p values.
+
+```{r pretty data, include = F}
+tag2 <- "lassoes-es"
+
+# A)
+tags_ext <- Reduce(intersect, get_tags(tagglob, tagextr, tag2 = tag2)['gsusie'])
+res1 <- caliPIP_plot(tags = tags_ext, tag2 = tag2)
+
+# B)
+res2 <- caliFDR_plot(tags = tags_ext, tag2 = tag2)
+
+# C)
+tags <- paste0('20200721-1-', c(2,4:9))
+res3 <- scatter_plot_PIP_p(tags, "lassoes-es")
+```
+
+```{r pretty plot, fig.height=3, fig.width=9}
+par(mfrow=c(1,3))
+cp_plot(res1$SUSIE.w0_PIP, res1$ifcausal, main = "SUSIE PIP calibration", col = "firebrick")
+grid()
+cp_plot(res2$FDR, res2$ifcausal, mode ="FDR", main = "TWAS FDP", col= "darkgreen")
+grid()
+plot(res3$TWAS, res3$SUSIE.w0, col = ifelse(res3$ifcausal=="Causal",  "salmon", "cyan3"), xlab = expression('TWAS -log'[10]*'(p)'), ylab = "SUSIE PIP",pch =19)
+grid()
+legend(9.5, 0.6, 
+       legend = c("Causal", "Non causal"), 
+      cex=0.9, pch=19, col=c("salmon", "cyan3"))
+```

### 2020-09-06 analysis/summarize_twas_plots.R

@@ -34,18 +34,18 @@ show_param <- function(tags, tag2){
     add_header_above(c(" " = 1, "Gene.pi1" = 2, "Gene.PVE" = 2, "SNP.pi1" = 2, "SNP.PVE" =2))
 }
 
-cp_plot <- function(s, ifcausal, mode = c("PIP", "FDR"), main = mode[1]){
+cp_plot <- function(s, ifcausal, mode = c("PIP", "FDR"), main = mode[1], ...){
   # ifcausal:0,1
   a_bin <- cut(s, breaks= seq(0,1, by=0.1))
   if (mode == "PIP") {
-    expected = c(by(s, a_bin, FUN = mean))
-    observed = c(by(ifcausal, a_bin, FUN = mean))
+    Expected = c(by(s, a_bin, FUN = mean))
+    Observed = c(by(ifcausal, a_bin, FUN = mean))
   } else if (mode == "FDR"){
-    expected = c(by(s, a_bin, FUN = mean))
-    observed = 1 - c(by(ifcausal, a_bin, FUN = mean))
+    Expected = c(by(s, a_bin, FUN = mean))
+    Observed = 1 - c(by(ifcausal, a_bin, FUN = mean))
   }
-  plot(expected, observed, xlim= c(0,1), ylim=c(0,1), pch =19, main = main)
-  lines(x = c(0,1), y = c(0,1), col ="red")
+  plot(Expected, Observed, xlim= c(0,1), ylim=c(0,1), pch =19, main = main, ...)
+  lines(x = c(0,1), y = c(0,1), col ="grey", lty = 2)
 }
 
 caliPIP_plot <- function(tags, tag2){
@@ -73,6 +73,7 @@ caliPIP_plot <- function(tags, tag2){
   cp_plot(res$SUSIE.w_PIP, res$ifcausal, main = "SUSIE.weighted PIP")
   cp_plot(res$SUSIE.u_PIP, res$ifcausal, main = "SUSIE.uniform PIP")
   cp_plot(res$SUSIE.w0_PIP, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
+  return(res)
 }
 
 
@@ -125,6 +126,7 @@ caliFDR_plot <- function(tags, tag2){
 
   cp_plot(res$FDR, res$ifcausal, mode ="FDR", main = "TWAS FDP")
   cat("FDR at bonferroni corrected p = 0.05: ", 1 - mean(res[res$PVALUE < 0.05 /dim(res)[1], "ifcausal"]))
[…]

### 2020-09-06 code/run_test_susie.R

@@ -95,10 +95,6 @@ loginfo("No. intervals for chr %s after PIP filter: %s", chrom, nrow(regions))
 outname <- args[6]
 
 
-if (length(args) == 7){
-  L = as.numeric(args[7])
-  outname <- paste0(outname, ".L", L)
-}
 loginfo("L = %s for susie run", L)
 
 loginfo("susie started for %s", outname)

### 2020-09-08 analysis/description.Rmd
pretty plot for Xin's grant

@@
 p <- 86000 # number of SNPs
+J <- 1653 # number of genes
 ```
 
-The genotype data we used is from UKB biobank, randomly selecting n=`r as.character(n)` samples.
+The genotype data we used is from UKB biobank, randomly selecting n = `r as.character(n)` samples.
 We use SNP genotype data from chr 17 to chr 22 combined.
@@
 This dataset contains ~ 380 samples.
-FUSION/TWAS were used train expression model and we used their lasso results.
-There are around 1700 genes with expression
+FUSION/TWAS were used to train expression model and we used their lasso results.
+eQTLs are defined as SNPs with abs(effectize) > 1e-8.
+There are `r as.character(J)` genes with expression from chr17 to chr22.
+
+2. Run susie iteratively for regions.
+
+We impute gene expression for our genotype data.
+Next, the analysis is done at the “region” level, which is about 500kb bins along the genome.
+Each bin would contain all the SNPs, as well as all the genes in that bin.
+The “genes” here mean the imputed expression from genotypes using the expression predictors.
+10:18 AM
+So the features of SuSiE are: SNPs and imputed genes (not cis-eQTLs of that gene). (edited) 
+10:22 AM
+Actually I just realize that there may be potential issues here: the cis-eQTLs of the gene may be outside the bins containing the gene.
+In that case, if eQTL is in LD with a causal SNP in another bin, SuSiE may still find that gene as causal, which is false.
+10:22 AM
+This may not be common though, if the bin is relatively large.
+10:23 AM
+Also to give an idea of the size of bin: on average, it contains about 7 genes and 120 SNPs in Siming’s simualation.

### 2020-09-08 analysis/index.Rmd
pretty plot for Xin's grant

@@
  * [Using UKbiobank data (chr17:22), fusion weights Adipose, test prior's effect for susie](simulation-multi-ukbchr17to22-gtex.adipose_susieprior.html)
-    * [Visualization](https://users.rcc.uchicago.edu/~simingz/simulation-susieP-ukbchr17to22-gtex.adipose.html)
+    * [Visualization](https://users.rcc.uchicago.edu/~simingz/simulation-susieP-ukbchr17to22-gtex.adipose.html) (TODO)
 

### 2020-09-08 analysis/simulation-multi-ukbchr17to22-gtex.adipose_sa2_v2.Rmd
pretty plot for Xin's grant

@@
 These genomic regions represents 12.5% of the genome.
-SNPs are downsampled to 1/10 (randomly), eQTLs (see below for definition of eQTL) were added back.
+SNPs are downsampled to 1/10 (randomly), eQTLs used in building the expression model were added back.
 We used lasso to train expression model using GTEx adipose tissue v7.
@@
 
-```{r pretty plot, fig.height=3, fig.width=9}
+```{r pretty plot, fig.height=3.2, fig.width=9}
 par(mfrow=c(1,3))
-cp_plot(res1$SUSIE.w0_PIP, res1$ifcausal, main = "SUSIE PIP calibration", col = "firebrick")
+cp_plot(res1$SUSIE.w0_PIP, res1$ifcausal, main = "SUSIE PIP calibration", col = "firebrick",cex.lab=1.4, cex.axis=1.4, cex.main=1.4, cex.sub=1.4)
 grid()
-cp_plot(res2$FDR, res2$ifcausal, mode ="FDR", main = "TWAS FDP", col= "darkgreen")
+cp_plot(res2$FDR, res2$ifcausal, mode ="FDR", main = "TWAS FDP", col= "darkgreen",cex.lab=1.4, cex.axis=1.4, cex.main=1.4, cex.sub=1.4)
 grid()
-plot(res3$TWAS, res3$SUSIE.w0, col = ifelse(res3$ifcausal=="Causal",  "salmon", "cyan3"), xlab = expression('TWAS -log'[10]*'(p)'), ylab = "SUSIE PIP",pch =19)
+plot(res3$TWAS, res3$SUSIE.w0, col = ifelse(res3$ifcausal=="Causal",  "salmon", "cyan3"), xlab = expression('TWAS -log'[10]*'(p)'), ylab = "SUSIE PIP",pch =19,cex.lab=1.4, cex.axis=1.4, cex.main=1.4, cex.sub=1.4)
 grid()
-legend(9.5, 0.6, 
+legend(8.5, 0.6, 
        legend = c("Causal", "Non causal"), 
-      cex=0.9, pch=19, col=c("salmon", "cyan3"))
+      cex=1.1, pch=19, col=c("salmon", "cyan3"))
 ```

### 2020-09-14 analysis/index.Rmd
susieI plots

--- 2020-09-03 analysis/index.Rmd
@@
 
-# Test parameters for susie
- * [Using UKbiobank data (chr17:22), fusion weights Adipose, test prior's effect for susie](simulation-multi-ukbchr17to22-gtex.adipose_susieprior.html)
-    * [Visualization](https://users.rcc.uchicago.edu/~simingz/simulation-susieP-ukbchr17to22-gtex.adipose.html)
-
 ## Summarize cTWAS results
@@
 *  [Using UKbiobank data (chr17:22), fusion weights Adipose](simulation-susieI-ukbchr17to22-gtex.adipose.html) 
+  *  [Using UKbiobank data (chr17:22), fusion weights Adipose, test prior's effect for susie](simulation-multi-ukbchr17to22-gtex.adipose_susieprior.html)
 

### 2020-09-14 analysis/simulation-multi-ukbchr17to22-gtex.adipose_susieprior.Rmd
susieI plots

--- 2020-09-01 analysis/simulation-multi-ukbchr17to22-gtex.adipose_susieprior.Rmd
@@
 ```
-`
 
@@
 * Data: ukb chr 17 to chr 22 combined. SNPs are downsampled to 1/10, eQTLs defined by FUSION-TWAS (Adipose/GTEx) lasso effect size > 0 were kept, p= 86k, n = 20k.
+
+* For each simulation scenarios(1-x, 3-x), we picked two few parameter sets (priors for gene and snps) estimated by susieI and check if PIPs for genes are calibrated.
 
@@
 
-# Setting 1
+# Simulation 1-x, setting 1
+- filter criteria: rpip 0.3 and causal
 
@@
 outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
-susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
+susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/20200721-1-fixprior_rpip0.3_causal/"
 ```
 
-We run 50 simulations and run susie using different priors and `L= 1`. We apply susie for all regions.
+We run 100 simulations and run susie using different priors and `L= 1`. We only apply susie for regions with mr.ash2 regional PIP > 0.5.
 
@@
 
-## Using uniform prior:
-
-```{r cali 0, fig.height=6, fig.width=6}
-pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior1.L1.susieres.expr.txt"))
-res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-
-cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
-```
-
-## Use overestimated prior
+## Use prior from susie estimates 1.
 
@@
 ```{r cali 1, fig.height=6, fig.width=14}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior1.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
 par(mfrow=c(1,2))
@@
 
-## Use priors close to truth
+## Use prior from susie estimates 2.
 
@@
 ```{r cali 2, fig.height=6, fig.width=14}
-pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior2.L1.susieres.expr.txt"))
+pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior2.susieres.expr.txt"))
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-
 par(mfrow=c(1,2))
@@
 
-## Use under estimated priors
+# simulation 1-x, setting 2
+- filter criteria: rpip 0.3
+
+```{r output a dir}
+susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/20200721-1-fixprior_rpip0.3/"
+```
+
+We run 50 simulations and run susie using different priors and `L= 1`. We apply susie for all regions.
+
+## Truth
+
+* The true parameters we used to simulate data: 
+```{r  true a par}
+par <- read.table(paste0(susiedir, "20200721-1-fixedprior_1.txt"), header = T)
+t(par[,2, drop = F])
+```
+
+## Use prior from susie estimates 1.
 
 * The prior used is: 
-```{r  par3}
-par <- read.table(paste0(susiedir, "20200721-1-fixedprior_3.txt"), header = T)
+```{r  par1 a}
+par <- read.table(paste0(susiedir, "20200721-1-fixedprior_1.txt"), header = T)
 t(par[c(1,3),1, drop = F])
 ```
-
 * susie results: 
-```{r cali 3, fig.height=6, fig.width=14}
-pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior3.L1.susieres.expr.txt"))
+```{r cali 1 a, fig.height=6, fig.width=14}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior1.susieres.expr.txt"))
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-
 par(mfrow=c(1,2))
@@
 
-# Setting 2
+## Use prior from susie estimates 2.
 
-We run 100 simulations and run susie using different priors and `L= 1`. We apply susie for all regions.
-
-## Truth
-
-* The true parameters we used to simulate data: 
-```{r  true par all}
-par <- read.table(paste0(susiedir, "20200721-3-fixedprior_1.txt"), header = T)
-t(par[,2, drop = F])
+* The prior used is: 
+```{r  par2 a}
+par <- read.table(paste0(susiedir, "20200721-1-fixedprior_2.txt"), header = T)
+t(par[c(1,3),1, drop = F])
 ```
 
-## Using uniform prior:
-
-```{r cali 0 all, fig.height=6, fig.width=6}
-pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior1.L1.susieres.expr.txt"))
+```{r cali 2 a, fig.height=6, fig.width=14}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior2.susieres.expr.txt"))
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-
-cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
-```
-
-## Use overestimated prior
-
-* The prior used is: 
-```{r  par1 all}
-par <- read.table(paste0(susiedir, "20200721-3-fixedprior_1.txt"), header = T)
-t(par[c(1,3),1, drop = F])
-```
-* susie results: 
-```{r cali 1 all, fig.height=6, fig.width=14}
 par(mfrow=c(1,2))
@@
 
-## Use priors close to truth
+# Simulation 3-x, setting 1
+- filter criteria: rpip 0.3 and causal
+
+```{r output 3.1 dir}
+susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/20200721-3-fixprior_rpip0.3_causal/"
+```
+
+We run 100 simulations and run susie using different priors and `L= 1`. We only apply susie for regions with mr.ash2 regional PIP > 0.5.
+
+## Truth
+
+* The true parameters we used to simulate data: 
+```{r  true par 3.1}
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_1.txt"), header = T)
+t(par[,2, drop = F])
+```
+
+## Use prior from susie estimates 1.
 
 * The prior used is: 
-```{r  par2 all}
-par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
+```{r  par1-2}
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_1.txt"), header = T)
 t(par[c(1,3),1, drop = F])
 ```
-
-```{r cali 2 all, fig.height=6, fig.width=14}
-pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.L1.susieres.expr.txt"))
+* susie results: 
+```{r cali 1-2, fig.height=6, fig.width=14}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior1.susieres.expr.txt"))
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-
 par(mfrow=c(1,2))
@@
 
-## Use under estimated priors
+## Use prior from susie estimates 2.
 
 * The prior used is: 
-```{r  par3 all}
-par <- read.table(paste0(susiedir, "20200721-3-fixedprior_3.txt"), header = T)
+```{r  par2-2}
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
 t(par[c(1,3),1, drop = F])
 ```
-
 * susie results: 
-```{r cali 3 all, fig.height=6, fig.width=14}
-pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior3.L1.susieres.expr.txt"))
+```{r cali 2-2, fig.height=6, fig.width=14}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.susieres.expr.txt"))
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-
 par(mfrow=c(1,2))
@@
 ```
+
+# Simulation 3-x, setting 2
+- filter criteria: rpip 0.3
+
+```{r output dir 3.2}
+susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/20200721-3-fixprior_rpip0.3/"
+```
+
+We run 100 simulations and run susie using different priors and `L= 1`. We only apply susie for regions with mr.ash2 regional PIP > 0.5.
+
+## Truth
+
+* The true parameters we used to simulate data: 
+```{r  true par3.2}
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_1.txt"), header = T)
+t(par[,2, drop = F])
+```
+
+## Use prior from susie estimates 1.
+
+* The prior used is: 
+```{r  par3.2.1}
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_1.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+```
+* susie results: 
+```{r cali 3.2.1, fig.height=6, fig.width=14}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+par(mfrow=c(1,2))
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
+```
+
+## Use prior from susie estimates 2.
+
+* The prior used is: 
+```{r  par 3.2.2}
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+```
+* susie results: 
+```{r cali 3.2.2, fig.height=6, fig.width=14}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+par(mfrow=c(1,2))
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
+```

### 2020-09-14 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
susieI plots

--- 2020-09-01 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
@@
 ```
-
-Test susieI for ukb chr 17 to chr 22 combined. SNPs are downsampled to 1/10, eQTLs defined by FUSION-TWAS (Adipose/GTEx) lasso effect size > 0 were kept, p= 86k, n = 20k.
-
-susieI is an iterative version of susie, learning parameters (`prior weights`) iteratively using the top region from mr.ash for genes only.
-
-```{r plot functions}
+```{r libs}
 library(mr.ash.alpha)
@@
 ```
-```{r output dir}
+
+```{r des, child = here::here("analysis/description.Rmd")}
+```
+
+```{r dirs}
 simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
@@
 
-# MR.ASH pip distribution
+# Regional minimum p value distribution
+
+* minimum TWAS p value of genes for each region:
+
+```{r rpip twas, fig.height=6, fig.width=6}
+simtag <- "20200721-1-1"
+rpipf <- paste0(simdatadir, simtag, ".pgenerprank.txt")
+
+a <- read.table(rpipf, header = T)
+a.c <- a[a$nCausal > 0, ]
+ax <- pretty(0:max(-log10(a$rpmin)), n = 30)
+
+par(mfrow=c(2,1))
+h1 <- hist(-log10(a$rpmin), breaks = 100, xlab = "PIP", main = "min P value distribution-all", col = "grey"); grid()
+h2 <- hist(-log10(a.c$rpmin), breaks = h1$breaks, xlab = "PIP", main = "min p distribution-causal", col = "salmon");grid()
+```
+
+* minimum TWAS p value of genes and SNPs for each region:
+
+```{r rpip twas2, fig.height=6, fig.width=6}
+rpipf <- paste0(simdatadir, simtag, ".p2rprank.txt")
+
+a <- read.table(rpipf, header = T)
+a.c <- a[a$nCausal > 0, ]
+ax <- pretty(0:max(-log10(a$rpmin)), n = 30)
+
+par(mfrow=c(2,1))
+h1 <- hist(-log10(a$rpmin), breaks = 100, xlab = "PIP", main = "min P value distribution-all", col = "grey"); grid()
+h2 <- hist(-log10(a.c$rpmin), breaks = h1$breaks, xlab = "PIP", main = "min p distribution-causal", col = "salmon");grid()
+```
+
+# MR.ASH regional pip sum distribution
 
@@
 
-```{r rpip 1, fig.height=6, fig.width=6}
-rpipf <- paste0(outputdir, tag, "-mr.ash.rPIP.txt")
+```{r rpip 1, fig.height=6, fig.width=6, eval=F}
+rpipf <- paste0(simdatadir, simtag, "-mr.ash.mrashrPIP.txt")
 
@@
 
-* MR.ash2s results for the same simulated data
+* MR.ash2s for both genes and SNPs using the same data
 
 ```{r rpip 2, fig.height=6, fig.width=6}
-rpipf <- "~/causalTWAS/simulations/simulation_ashtest_20200721/20200721-1-1-mr.ash2s.lassoes-es.rPIP.txt"
+rpipf <- paste0(simdatadir, simtag, "-mr.ash2s.lassoes-es.mrash2srPIP.txt")
 
@@
 
-# susieI genes - setting 1
+# Parameter estimation results
+
+Results: Each row shows parameter estimation results from 5 simulation runs with similar settings (i.e.
+pi1 and PVE for genes and SNPs).
+each row has two plots, one for gene pi1 estimation, one for enrichment (gene pi1/snp pi1).
+Results from each run were represented by one dot, dots with the same color come from the same run.
+horizontal dash lines: simulation truth, `mr.ash2`, mr.ash for gene and SNPs; `susietruth`, the truth in selected regions that were used to run susie iteractively (susieI).
+
+## Regions: causal
 
@@
 
-## One example
-
-Prior change for each iteration:
-```{r iter}
-load(paste0(outputdir, tag, ".susieIres.Rd"))
-df <- cbind(prior.gene_rec, prior.SNP_rec)
-rownames(df) <- paste("Iteration", 1:Niter)
-colnames(df) <- c("Prior.gene", "Prior.SNP")
-df
-```
-
-## more examples
-
-Each row represents one simulation run. Runs labeled `1-x` have the same simulation parameters, as you can tell from the true parameters. (and so are runs labled `2-x` and `3-x`)
-
-```{r param}
+```{r param functions}
 show_param <- function(mrashfs, susieIfs, susieIfs2){
@@
   rownames(df) <- NULL
+  return(df)
 
-  df %>% 
-  kable("html", escape = F) %>%
-  kable_styling("striped", full_width = F) %>%
-   row_spec(c(1:5, 11:15), background = "#FEF3B9") %>%
-  scroll_box(width = "100%", height = "600px", fixed_thead = T)
+  # df %>% 
+  # kable("html", escape = F) %>%
+  # kable_styling("striped", full_width = F) %>%
+  #  row_spec(c(1:5, 11:15), background = "#FEF3B9") %>%
+  # scroll_box(width = "100%", height = "600px", fixed_thead = T)
 }
 
+plot_param <- function(df, ...){
+  df <- apply(df[ , 2:ncol(df)], 2, function(x) as.numeric(x))
+  m <- cbind(df[,"pi1.gene_mr.ash2"], 1:nrow(df), 1 + 1:nrow(df)/nrow(df)/3)
+  st <-  cbind(df[,"pi1.gene_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
+  s <- cbind(df[,"pi1.gene_susieI"], 1:nrow(df), 3 + 1:nrow(df)/nrow(df)/3)
+  t <- df[1,"pi1.gene_truth"]
+  dfp <- rbind(m,st,s)
+  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "gene pi1", xaxt = "n", xlab="", xlim = c(0.8, 3.5), frame.plot=FALSE, ylim = c(0, max(dfp[,1],t) *1.05), ...)
+  axis(side=1, at=1:3, labels = FALSE, tick = F)
+  text(x=1:3, 0, labels = c("mr.ash2", "susieI_truth", "susieI"), xpd = T, pos =1)
+  abline(h=t, lty = 2, col= "salmon", lwd=1.5)
+  grid()
+
+  m <- cbind(df[,"pi1.gene_mr.ash2"]/df[,"pi1.SNP_mr.ash2"], 1:nrow(df), 1 + 1:nrow(df)/nrow(df)/3)
+  st <-  cbind(df[,"pi1.gene_susietruth"]/df[,"pi1.SNP_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
+  s <- cbind(df[,"pi1.gene_susieI"]/df[,"pi1.SNP_susieI"], 1:nrow(df), 3 + 1:nrow(df)/nrow(df)/3)
+  t <- df[1,"pi1.gene_truth"]/df[1,"pi1.SNP_truth"]
+  dfp <- rbind(m,st,s)
+  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "Enrichment (gene/snp)", xaxt = "n", xlab="", xlim = c(0.8, 3.5),frame.plot=FALSE, ylim = c(0, max(dfp[,1],t, 150) *1.05))
+  axis(side=1, at=1:3, labels = FALSE, tick = F)
+  text(x=1:3, 0, labels = c("mr.ash2", "susieI_truth", "susieI"), xpd = T, pos =1)
+  abline(h= t, lty = 2, col= "darkgreen", lwd=1.5)
+  grid()
+}
+```
+
+```{r param, fig.height=8, fig.width=6}
 tags <- paste(rep(1:3, each = 5), 1:5, sep = "-")
@@
 
-show_param(mrashfs, susieIfs, susieIfs2)
+df <- show_param(mrashfs, susieIfs, susieIfs2)
+par(mfrow = c(3,2))
+plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05") 
+plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
 ```
 
-# susieI genes - setting 2
+## Regions: mrash gene+SNP (1)
 
-* Regions: All regions.
+* Regions: We use regional PIP of both SNP and gene from mr.ash2s > 0.3.
 
-* Susie run parameters: `L=1`. initialize with `null_weight = 0` and update `null_weight` based on last iteration results. We initialize with prior for genes and SNPs as uniform.
+* Susie run parameters: `L=2`, initialize `null_weight = 0`, update `null_weight` in iterations.
+We initialize with prior for genes and SNPs as uniform.
+(We also tried initialize with prior for genes and SNPs as 0.05 and 2.5e-3, results are the same.)
 
-## One example
+```{r param-6, fig.height=11, fig.width=6}
+tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
+mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".config6.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config6.susieI.txt")
 
-Prior change for each iteration:
-```{r iter2}
-load(paste0(outputdir, tag, ".config1.susieIres.Rd"))
-df <- cbind(prior.gene_rec, prior.SNP_rec)
-rownames(df) <- paste("Iteration", 1:Niter)
-colnames(df) <- c("Prior.gene", "Prior.SNP")
-df
+df <- show_param(mrashfs, susieIfs, susieIfs2)
+par(mfrow = c(4,2))
+plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
 ```
 
-## More examples
+## Regions: mrash gene+SNP (2)
 
-```{r param2}
+* Regions: We use regional PIP of both SNP and gene from mr.ash2s > 0.3.
+
+* Susie run parameters: `L=2`, `null_weight = 0`. Didn't update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform.
+
+```{r param-7, fig.height=11, fig.width=6}
 tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
-mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
-susieIfs <- paste0(outputdir, "20200813-", tags, ".config1.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config1.susieI.txt")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".config7.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config7.susieI.txt")
 
-show_param(mrashfs, susieIfs, susieIfs2)
+df <- show_param(mrashfs, susieIfs, susieIfs2)
+par(mfrow = c(4,2))
+plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
 ```
 
-# susieI genes - setting 3
+## Regions: mrash gene only
+
+* Regions: We use regional sum of PIP for genes from mr.ash for genes only > 0.1.
+
+* Susie run parameters: `L=2`. Update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform, `null_weight = 0`.
+
+```{r param-9, fig.height=11, fig.width=6}
+tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".config9.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config9.susieI.txt")
+
+df <- show_param(mrashfs, susieIfs, susieIfs2)
+par(mfrow = c(4,2))
+plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
+```
+
+## Regions: TWAS gene only
+
+* Regions: We use regional min p value for genes from twas for genes only, requiring this p value in top 5%.
+
+* Susie run parameters: `L=2`. Update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform, `null_weight = 0`.
+
+```{r param-10, fig.height=11, fig.width=6}
+tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".config10.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config10.susieI.txt")
+
+df <- show_param(mrashfs, susieIfs, susieIfs2)
+par(mfrow = c(4,2))
+plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
+```
+
+## Regions: TWAS gene+SNP
+
+* Regions: We use regional min p value for genes or SNPs from twas for genes and snps, requiring this p value in top 5% for genes or 1% for SNPs.
+
+* Susie run parameters: `L=2`. Update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform, `null_weight = 0`.
+
+```{r param-11, fig.height=11, fig.width=6}
+tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".config11.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config11.susieI.txt")
+
+df <- show_param(mrashfs, susieIfs, susieIfs2)
+par(mfrow = c(4,2))
+plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
+```
+
+## Regions: all (1)
 
@@
 
-## One example
-
-Prior change for each iteration:
-```{r iter3}
-load(paste0(outputdir, tag, ".config2.susieIres.Rd"))
-df <- cbind(prior.gene_rec, prior.SNP_rec)
-rownames(df) <- paste("Iteration", 1:Niter)
-colnames(df) <- c("Prior.gene", "Prior.SNP")
-df
-```
-
-## More examples
-
-```{r param3}
+```{r param3, fig.height=11, fig.width=6}
 mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
@@
 
-show_param(mrashfs, susieIfs, susieIfs2)
+df <- show_param(mrashfs, susieIfs, susieIfs2)
+par(mfrow = c(4,2))
+plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
 ```
 
-# susieI genes - setting 4
+## Regions: all (2)
 
@@
 
-## One example
-
-Prior change for each iteration:
-```{r iter4}
-load(paste0(outputdir, tag, ".config3.susieIres.Rd"))
-df <- cbind(prior.gene_rec, prior.SNP_rec)
-rownames(df) <- paste("Iteration", 1:Niter)
-colnames(df) <- c("Prior.gene", "Prior.SNP")
-df
-```
-
-## More examples
-
-```{r param4}
+```{r param4, fig.height=11, fig.width=6}
 mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
@@
 
-show_param(mrashfs, susieIfs, susieIfs2)
+df <- show_param(mrashfs, susieIfs, susieIfs2)
+par(mfrow = c(4,2))
+plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
 ```
 
-# susieI genes - setting 5
-
-* Regions: All regions.
-
-* Susie run parameters: `L=2`. Update `null_weight` in each iteration. We initialize prior for gene as `0.05` and prior for SNPs as `2.5E-3`, we initialize `null_weight` as 0.
-
-## One example
-
-Prior change for each iteration:
-```{r iter5}
-load(paste0(outputdir, tag, ".config4.susieIres.Rd"))
-df <- cbind(prior.gene_rec, prior.SNP_rec)
-rownames(df) <- paste("Iteration", 1:Niter)
-colnames(df) <- c("Prior.gene", "Prior.SNP")
-df
-```
-
-## More examples
-
-```{r param5}
-mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
-susieIfs <- paste0(outputdir, "20200813-", tags, ".config4.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config4.susieI.txt")
-
-show_param(mrashfs, susieIfs, susieIfs2)
-```
-
-# susieI shuffle genes
-## one example
-* Shuffle genes to different regions. Regions: We use regional PIP of both SNP and gene from mr.ash2s > 0.3 , and also require at least one causal gene or SNP is present, also, at least one gene.
-
-* Susie run parameters: `L=1`, `null_weight = 0`.
-
-Prior change for each iteration:
-```{r iter shuffle}
-tag <- '20200813-1-1-shuffleg'
-load(paste0(outputdir, tag, ".susieIres.Rd"))
-df <- cbind(prior.gene_rec, prior.SNP_rec)
-rownames(df) <- paste("Iteration", 1:nrow(df))
-colnames(df) <- c("Prior.gene", "Prior.SNP")
-df
-```
+## Regions: all (others)
+* Susie run parameters: `L=1`.
+initialize with null_weight = 0 and update null_weight based on last iteration results.
+We initialize with prior for genes and SNPs as uniform.
+The prior for genes is underestimated for low power settings (1-x, 2-x), many converged to close to 0, for high power settings it is better.
+* Susie run parameters: `L=5`.
+initialize with null_weight = 0 and update null_weight based on last iteration results.
+The prior for gene estimation is fine for all runs, however prior for SNPs overestimated about 10 times.
+* We have tried different initiation strategies, but the results remain the same.

### 2020-09-14 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
susieI plots

@@
   dfp <- rbind(m,st,s)
-  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "Enrichment (gene/snp)", xaxt = "n", xlab="", xlim = c(0.8, 3.5),frame.plot=FALSE, ylim = c(0, max(dfp[,1],t, 150) *1.05))
+  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "Enrichment (gene/snp)", xaxt = "n", xlab="", xlim = c(0.8, 3.5),frame.plot=FALSE, ylim = c(0, min(max(dfp[,1],t) *1.05, 150)))
   axis(side=1, at=1:3, labels = FALSE, tick = F)

### 2020-09-14 analysis/simulation-multi-ukbchr17to22-gtex.adipose_susieprior.Rmd
susieI plots

@@
 # Simulation 1-x, setting 1
-- filter criteria: rpip 0.3 and causal
 
@@
 
-We run 100 simulations and run susie using different priors and `L= 1`. We only apply susie for regions with mr.ash2 regional PIP > 0.5.
+We run 100 simulations and run susie using priors obtained from susieI.
+Two examples are shown.
+The setting is the same as in [here](https://stephenslab.github.io/causal-TWAS/simulation-susieI-ukbchr17to22-gtex.adipose.html#regions:_causal).
+Regions are rpip from mr.ash2 > 0.3 and causal.
+We take simulation setting 1-x.
 
@@
 # simulation 1-x, setting 2
-- filter criteria: rpip 0.3
 
@@
 
-We run 50 simulations and run susie using different priors and `L= 1`. We apply susie for all regions.
+We run 100 simulations and run susie using priors obtained from susieI.
+Two examples are shown.
+The setting is the same as in [here](https://stephenslab.github.io/causal-TWAS/simulation-susieI-ukbchr17to22-gtex.adipose.html#regions:_mrash_gene+snp_(1)).
+Regions are rpip from mr.ash2 > 0.3.
+We take simulation setting 1-x.
 
@@
 # Simulation 3-x, setting 1
-- filter criteria: rpip 0.3 and causal
 
@@
 
-We run 100 simulations and run susie using different priors and `L= 1`. We only apply susie for regions with mr.ash2 regional PIP > 0.5.
+We run 100 simulations and run susie using priors obtained from susieI.
+Two examples are shown.
+The setting is the same as in [here](https://stephenslab.github.io/causal-TWAS/simulation-susieI-ukbchr17to22-gtex.adipose.html#regions:_causal).
+Regions are rpip from mr.ash2 > 0.3 and causal.
+We take simulation setting 3-x.
 
@@
 # Simulation 3-x, setting 2
-- filter criteria: rpip 0.3
 
@@
 
-We run 100 simulations and run susie using different priors and `L= 1`. We only apply susie for regions with mr.ash2 regional PIP > 0.5.
+We run 100 simulations and run susie using priors obtained from susieI.
+Two examples are shown.
+The setting is the same as in [here](https://stephenslab.github.io/causal-TWAS/simulation-susieI-ukbchr17to22-gtex.adipose.html#regions:_mrash_gene+snp_(1)).
+Regions are rpip from mr.ash2 > 0.3.
+We take simulation setting 1-x.
 

### 2020-09-18 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
BF

@@
 h2 <- hist(-log10(a.c$rpmin), breaks = h1$breaks, xlab = "PIP", main = "min p distribution-causal", col = "salmon");grid()
+```
+
+# Regional BF distribution
+mean BF of both SNPs and genes in region as the regional BF.
+
+```{r rBF twas, fig.height=6, fig.width=6}
+rpipf <- paste0(simdatadir, simtag, ".rBF.txt")
+
+a <- read.table(rpipf, header = T)
+a.c <- a[a$nCausal > 0, ]
+
+par(mfrow=c(2,1))
+h1 <- hist(a$rBF, breaks = 5000, xlab = "PIP", main = "regional BF distribution-all", col = "grey", xlim = c(0,10)); grid()
+h2 <- hist(a.c$rBF, breaks = h1$breaks, xlab = "PIP", main = "regional BF distribution-causal", col = "salmon",xlim = c(0,10));grid()
 ```
@@
 * We have tried different initiation strategies, but the results remain the same.
+
+## Regions: TWAS BF
+
+* Regions: regional TWAS BF > 2
+
+```{r param 5, fig.height=11, fig.width=6}
+tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
+mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".config12.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config12.susieI.txt")
+
+df <- show_param(mrashfs, susieIfs, susieIfs2)
+par(mfrow = c(4,2))
+plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
+```

### 2020-09-18 code/simulate_phenotype.R

@@ -24,7 +24,7 @@ cis_expr <- function(dat, weight, method = "bslmm", checksnps = F){
                                  paste(ID, ave(ID, ID, FUN=seq_along), sep='_ID'),
                                  ID))
 
-  write.table(wgtpos , file= paste0(wgtposfile, ".IDfixed") , row.names=F, col.names=T, sep="\t", quote = F)
+  write.table(wgtpos, file= paste0(wgtposfile, ".IDfixed") , row.names=F, col.names=T, sep="\t", quote = F)
 
   print(paste0("Number of genes with weights provided: ", dim(wgtpos)[1]))
 
@@ -99,27 +99,32 @@ cis_expr <- function(dat, weight, method = "bslmm", checksnps = F){
   return(list("expr"= expr, "exprlist" = exprlist, "qclist" = qclist, "chrom" = chrom, "p0" = p0, "p1" = p1))
 }
 
-# dat: a list with the following items:
-#   - G: matrix, N x M, genotype, need to be scaled
-#   - snp: matrix, N x 1, SNP name
-#   - ref
-#   - alt
-#   - chr
-#   - pos
-#   - expr: matrix, N x J
-# mode: string, "snp-only", "snp-expr"
-# pve.expr: scalar [0,1]
-# pve.snp: scalar [0,1]
-# pi_beta: scalar [0,1]
-# pi_theta: scalar [0,1]
-# tau: 1 (to add)
+#' @param dat a list with the following items:
+#'   - G: matrix, N x M, genotype, need to be scaled
+#'   - snp: matrix, N x 1, SNP name
+#'   - ref
+#'   - alt
+#'   - chr
+#'   - pos
+#'   - expr: matrix, N x J
+#' @param mode string, "snp-only", "snp-expr"
+#' @param pve.expr scalar [0,1], not used when if `sigma_beta` is given
+#' @param pve.snp scalar [0,1], not used if `sigma_theta` is given
+#' @param pi_beta scalar [0,1]
+#' @param pi_theta scalar [0,1]
+#' @param sigma_beta default NULL, if not given, will calculate from pve.
+#' @param sigma_theta default NULL, if not given, will calculate from pve.
+#' @return a list, Y.g is phenotype determined by genetic, will need to add
+#'  error term ( pve calculation is based on N(0,1)).
[…]

### 2020-09-20 analysis/simulation-multi-ukbchr17to22-gtex.adipose_susieprior.Rmd
susie cali truth

@@
 source("analysis/summarize_twas_plots.R")
+```
+
+# Simulation 1-x, test true prior
+
+```{r output-t dir}
+simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
+outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
+susiedir <- "/home/simingz/causalTWAS/simulations/simulation_susietest_20200721/20200721-1-fixprior_rpip0.5/"
+```
+
+We run 50 simulations and run susie using the true parameter (gene and snp pi1) as prior.
+We run susie for regions with regional sum of pip from mr.ash2s for both genes and snps > 0.5.
+run susie with L=1.
+
+## Truth
+
+* The true parameters we used to simulate data: 
+```{r  true par-t}
+par <- read.table(paste0(susiedir, "20200721-1-fixedprior_1.txt"), header = T)
+t(par[,2, drop = F])
+```
+
+## Use truth as prior
+
+* The prior used is: 
+```{r  par1-t}
+par <- read.table(paste0(susiedir, "20200721-1-fixedprior_2.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+```
+* susie results: 
+```{r cali 1-t, fig.height=6, fig.width=14}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior2.L1.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+par(mfrow=c(1,2))
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
 ```

### 2020-09-21 analysis/simulation-multi-ukbchr17to22-gtex.adipose_susieprior.Rmd
causal susie prior

--- 2020-09-14 analysis/simulation-multi-ukbchr17to22-gtex.adipose_susieprior.Rmd
@@
 source("analysis/summarize_twas_plots.R")
+```
+
+# Test true prior
+
+## Regions rPIP > 0.5, simulation 1-x
+```{r output-t dir}
+simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
+outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
+susiedir <- "/home/simingz/causalTWAS/simulations/simulation_susietest_20200721/20200721-1-fixprior_rpip0.5/"
+```
+
+We run 50 simulations and run susie using the true parameter (gene and snp pi1) as prior.
+We run susie for regions with regional sum of pip from mr.ash2s for both genes and snps > 0.5.
+run susie with L=1.
+
+### Truth
+
+* The true parameters we used to simulate data: 
+```{r  true par-t}
+par <- read.table(paste0(susiedir, "20200721-1-fixedprior_2.txt"), header = T)
+t(par[,2, drop = F])
+```
+
+### Use truth as prior
+
+* The prior used is: 
+```{r  par1-t}
+par <- read.table(paste0(susiedir, "20200721-1-fixedprior_2.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+susiedir <- "/home/simingz/causalTWAS/simulations/simulation_susietest_20200721/20200721-1-fixprior_rpip0.5/"
+```
+* susie results: 
+```{r cali 1-t, fig.height=6, fig.width=14}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior2.L1.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+par(mfrow=c(1,2))
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
+```
+
+## Causal regions, simulation 3-x
+We run susie for regions with causal signals. run susie with L=1.
+
+### Truth
+
+* The true parameters we used to simulate data: 
+```{r  true par-t3}
+susiedir <- "/home/simingz/causalTWAS/simulations/simulation_susietest_20200721/20200721-3-fixprior_causal/"
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
+t(par[,2, drop = F])
+```
+
+### Use truth as prior
+
+* The prior used is: 
+```{r  par1-t3}
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+```
+* susie results: 
+```{r cali 1-t3, fig.height=6, fig.width=14}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.L1.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+par(mfrow=c(1,2))
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
 ```

### 2020-09-23 analysis/simulation-multi-ukbchr17to22-gtex.adipose_susieprior.Rmd
susie prior

--- 2020-09-14 analysis/simulation-multi-ukbchr17to22-gtex.adipose_susieprior.Rmd
@@
 source("analysis/summarize_twas_plots.R")
+```
+
+# Test true prior
+
+## Regions rPIP > 0.5, simulation 1-x
+```{r output-t dir}
+simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
+outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
+susiedir <- "/home/simingz/causalTWAS/simulations/simulation_susietest_20200721/20200721-1-fixprior_rpip0.5/"
+```
+
+We run 50 simulations and run susie using the true parameter (gene and snp pi1) as prior.
+We run susie for regions with regional sum of pip from mr.ash2s for both genes and snps > 0.5.
+run susie with L=1.
+
+### Truth
+
+* The true parameters we used to simulate data: 
+```{r  true par-t}
+par <- read.table(paste0(susiedir, "20200721-1-fixedprior_2.txt"), header = T)
+t(par[,2, drop = F])
+```
+
+### Use truth as prior
+
+* The prior used is: 
+```{r  par1-t}
+par <- read.table(paste0(susiedir, "20200721-1-fixedprior_2.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+susiedir <- "/home/simingz/causalTWAS/simulations/simulation_susietest_20200721/20200721-1-fixprior_rpip0.5/"
+```
+* susie results: 
+```{r cali 1-t, fig.height=4, fig.width=12}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior2.L1.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+par(mfrow=c(1,3)); cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP"); cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP"); cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
+cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
+```
+
+## Causal regions, simulation 3-x
+We run susie for regions with causal signals. run susie with L=1.
+
+### Truth
+
+* The true parameters we used to simulate data: 
+```{r  true par-t3}
+susiedir <- "/home/simingz/causalTWAS/simulations/simulation_susietest_20200721/20200721-3-fixprior_causal/"
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
+t(par[,2, drop = F])
+```
+
+### Use truth as prior
+
+* The prior used is: 
+```{r  par1-t3}
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+```
+* susie results: 
+```{r cali 1-t3, fig.height=4, fig.width=12}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.L1.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+par(mfrow=c(1,3)); 
+cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
+```
+
+# Test over and underestimated gene pi1
+We intentionally use over or underestimated gene pi1 and use the same snp pi1 (close to truth) when running SUSIE. Run susie `L=1`.
+
+## Causal regions, simulation 3-x
+
+* The true parameters we used to simulate data: 
+```{r  par-over-t}
+susiedir <- "/home/simingz/causalTWAS/simulations/simulation_susietest_20200721/20200721-3-fixprior_causal/"
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
+t(par[,2, drop = F])
+```
+
+### Use overestimated gene pi1
+
+* The prior used is: 
+```{r  par1-t3-over}
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_1.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+```
+* susie results: 
+```{r cali 1-t3-over, fig.height=4, fig.width=12}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior1.L1.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+par(mfrow=c(1,3))
+cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
+```
+
+### Use underrestimated gene pi1
+
+* The prior used is: 
+```{r  par1-t3-under}
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_3.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+```
+* susie results: 
+```{r cali 1-t3-under, fig.height=4, fig.width=12}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior3.L1.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+par(mfrow=c(1,3))
+cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
 ```
@@
 * susie results: 
-```{r cali 1, fig.height=6, fig.width=14}
+```{r cali 1, fig.height=4, fig.width=12}
 pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior1.susieres.expr.txt"))
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-par(mfrow=c(1,2))
+par(mfrow=c(1,3))
+cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
 cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
@@
 
-```{r cali 2, fig.height=6, fig.width=14}
+```{r cali 2, fig.height=4, fig.width=12}
 pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior2.susieres.expr.txt"))
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-par(mfrow=c(1,2))
+par(mfrow=c(1,3))
+cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
 cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
@@
 * susie results: 
-```{r cali 1 a, fig.height=6, fig.width=14}
+```{r cali 1 a, fig.height=4, fig.width=12}
 pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior1.susieres.expr.txt"))
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-par(mfrow=c(1,2))
+par(mfrow=c(1,3))
+cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
 cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
@@
 
-```{r cali 2 a, fig.height=6, fig.width=14}
+```{r cali 2 a, fig.height=4, fig.width=12}
 pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior2.susieres.expr.txt"))
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-par(mfrow=c(1,2))
+par(mfrow=c(1,3)) 
+cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
 cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
@@
 * susie results: 
-```{r cali 1-2, fig.height=6, fig.width=14}
+```{r cali 1-2, fig.height=4, fig.width=12}
 pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior1.susieres.expr.txt"))
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-par(mfrow=c(1,2))
+par(mfrow=c(1,3))
+cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
 cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
@@
 * susie results: 
-```{r cali 2-2, fig.height=6, fig.width=14}
+```{r cali 2-2, fig.height=4, fig.width=12}
 pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.susieres.expr.txt"))
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-par(mfrow=c(1,2))
+par(mfrow=c(1,3))
+cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
 cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
@@
 * susie results: 
-```{r cali 3.2.1, fig.height=6, fig.width=14}
+```{r cali 3.2.1, fig.height=4, fig.width=12}
 pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.susieres.expr.txt"))
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-par(mfrow=c(1,2))
+par(mfrow=c(1,3))
+cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
 cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
@@
 * susie results: 
-```{r cali 3.2.2, fig.height=6, fig.width=14}
+```{r cali 3.2.2, fig.height=4, fig.width=12}
 pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.susieres.expr.txt"))
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-par(mfrow=c(1,2))
+par(mfrow=c(1,3)) 
+cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
 cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")

### 2020-09-25 analysis/simulation-multi-ukbchr17to22-gtex.adipose_susieprior.Rmd
susie I multi

@@
 res <- do.call(rbind, lapply(pipfs, read.table, header = T))
-par(mfrow=c(1,3)); cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP"); cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP"); cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
+par(mfrow=c(1,3))
+cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
+```
+
+## Regions rPIP > 0.9, simulation 3-x
+We run 50 simulations and run susie using the true parameter (gene and snp pi1) as prior.
+We run susie for regions with regional sum of pip from mr.ash2s for both genes and snps > 0.9.
+run susie with L= 2.
+
+### Truth
+
+* The true parameters we used to simulate data: 
+```{r  true par-t0.9}
+susiedir <- "/home/simingz/causalTWAS/simulations/simulation_susietest_20200721/20200721-3-fixprior_rpip0.9/"
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
+t(par[,2, drop = F])
+```
+
+### Use truth as prior
+
+* The prior used is: 
+```{r  par1-t0.9}
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+```
+* susie results: 
+```{r cali 1-t0.9, fig.height=4, fig.width=12}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+par(mfrow=c(1,3))
 cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
@@
 ```{r cali 1-t3, fig.height=4, fig.width=12}
+pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.L1.susieres.expr.txt"))
+res <- do.call(rbind, lapply(pipfs, read.table, header = T))
+par(mfrow=c(1,3)); 
+cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
+cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
+cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
+```
+
+## All regions, simulation 3-x
+We run susie for all regions. Run susie with L=1.
+
+### Truth
+
+* The true parameters we used to simulate data: 
+```{r  true par-a3}
+susiedir <- "/project2/mstephens/causalTWAS/simulations/simulation_susietest_20200721/20200721-3-fixprior_allregions/"
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
+t(par[,2, drop = F])
+```
+
+### Use truth as prior
+
+* The prior used is: 
+```{r  par1-a3}
+par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
+t(par[c(1,3),1, drop = F])
+```
+* susie results: 
+```{r cali 1-a3, fig.height=4, fig.width=12}
 pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.L1.susieres.expr.txt"))
@@
 
+## Test susieI results
 # Simulation 1-x, setting 1

### 2020-09-25 code/run_simulate_data.R

@@ -53,6 +53,8 @@ for ( b in 1:length(pfileRds)){
   weight <- args[2]
   exprres <- cis_expr(dat, weight, method = "lasso", checksnps = F)
   save(exprres, file = paste0(outname, "-cis-expr.Rd"))
+  expr <- scaleRcpp(exprres$expr)
+  as_FBM(expr, backingfile = paste0(outname, "-cis-expr.FBM.scaled"), is_read_only = T)$save()
   nlist[[b]] <- c(ncol(dat$G), ncol(exprres$expr), sum(apply(exprres$expr, 2, var))) # genotype is scaled
 }
 rm(dat);gc()

### 2020-10-02 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
batch susieI

@@
 }
+
+gpip_dist <- function(susiefs, ...){
+  dflist <- list()
+  for (f in susiefs){
+    dflist[[f]] <- read.table(f, header =T , stringsAsFactors = F)
+  }
+  df <- do.call(rbind, dflist)
+  hist(df[df$type == "gene", "susie_pip"], xlab = "gene susie PIP", 
+       breaks = 50, ylim = c(0,20), xlim=c(0,1), col = "salmon", ...)
+}
 ```
@@
 plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+```
+
+```{r pipdist,fig.height=3, fig.width=9}
+par(mfrow = c(1,3))
+gpip_dist(susieIfs2[1:5], main = "1-x combined")
+gpip_dist(susieIfs2[6:10], main = "2-x combined")
+gpip_dist(susieIfs2[11:15], main = "3-x combined")
 ```
@@
 ```
+```{r pipdist-6,fig.height=3, fig.width=12}
+par(mfrow = c(1,4))
+gpip_dist(susieIfs2[1:5], main = "1-x combined")
+gpip_dist(susieIfs2[6:10], main = "2-x combined")
+gpip_dist(susieIfs2[11:15], main = "3-x combined")
+gpip_dist(susieIfs2[16:20], main = "4-x combined")
+```
 
@@
 
+```{r pipdist-7,fig.height=3, fig.width=12}
+par(mfrow = c(1,4))
+gpip_dist(susieIfs2[1:5], main = "1-x combined")
+gpip_dist(susieIfs2[6:10], main = "2-x combined")
+gpip_dist(susieIfs2[11:15], main = "3-x combined")
+gpip_dist(susieIfs2[16:20], main = "4-x combined")
+```
+
 ## Regions: mrash gene only
@@
 plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
+```
+```{r pipdist-9,fig.height=3, fig.width=12}
+par(mfrow = c(1,4))
+gpip_dist(susieIfs2[1:5], main = "1-x combined")
+gpip_dist(susieIfs2[6:10], main = "2-x combined")
+gpip_dist(susieIfs2[11:15], main = "3-x combined")
+gpip_dist(susieIfs2[16:20], main = "4-x combined")
 ```
@@
 ```
+```{r pipdist-10,fig.height=3, fig.width=12}
+par(mfrow = c(1,4))
+gpip_dist(susieIfs2[1:5], main = "1-x combined")
+gpip_dist(susieIfs2[6:10], main = "2-x combined")
+gpip_dist(susieIfs2[11:15], main = "3-x combined")
+gpip_dist(susieIfs2[16:20], main = "4-x combined")
+```
 
-## Regions: TWAS gene+SNP
+## Regions: TWAS gene+SNP top
 
@@
 plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
+```
+```{r pipdist-11,fig.height=3, fig.width=12}
+par(mfrow = c(1,4))
+gpip_dist(susieIfs2[1:5], main = "1-x combined")
+gpip_dist(susieIfs2[6:10], main = "2-x combined")
+gpip_dist(susieIfs2[11:15], main = "3-x combined")
+gpip_dist(susieIfs2[16:20], main = "4-x combined")
+```
+
+## Regions: TWAS gene+SNP BF
+
+* Regions: get BF for both genes and SNPs. regional TWAS BF > 2
+
+```{r param 5, fig.height=11, fig.width=6}
+tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
+mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".config12.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config12.susieI.txt")
+
+df <- show_param(mrashfs, susieIfs, susieIfs2)
+par(mfrow = c(4,2))
+plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
+```
+```{r pipdist-5,fig.height=3, fig.width=12}
+par(mfrow = c(1,4))
+gpip_dist(susieIfs2[1:5], main = "1-x combined")
+gpip_dist(susieIfs2[6:10], main = "2-x combined")
+gpip_dist(susieIfs2[11:15], main = "3-x combined")
+gpip_dist(susieIfs2[16:20], main = "4-x combined")
+```
+
+## Regions: TWAS gene+SNP FDR
+
+* Regions: put gene and SNP pvalues together and get FDR. Seelect regions with minimum FDR < 0.05.
+
+```{r param 13, fig.height=11, fig.width=6}
+tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
+mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
+susieIfs <- paste0(outputdir, "20200813-", tags, ".config13.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config13.susieI.txt")
+
+df <- show_param(mrashfs, susieIfs, susieIfs2)
+par(mfrow = c(4,2))
+plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
+plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
+```
+```{r pipdist-13,fig.height=3, fig.width=12}
+par(mfrow = c(1,4))
+gpip_dist(susieIfs2[1:5], main = "1-x combined")
+gpip_dist(susieIfs2[6:10], main = "2-x combined")
+gpip_dist(susieIfs2[11:15], main = "3-x combined")
+gpip_dist(susieIfs2[16:20], main = "4-x combined")
 ```
@@
 ```
+```{r pipdist-3,fig.height=3, fig.width=12}
+par(mfrow = c(1,4))
+gpip_dist(susieIfs2[1:5], main = "1-x combined")
+gpip_dist(susieIfs2[6:10], main = "2-x combined")
+gpip_dist(susieIfs2[11:15], main = "3-x combined")
+gpip_dist(susieIfs2[16:20], main = "4-x combined")
+```
 
@@
 ```
+```{r pipdist-4,fig.height=3, fig.width=12}
+par(mfrow = c(1,4))
+gpip_dist(susieIfs2[1:5], main = "1-x combined")
+gpip_dist(susieIfs2[6:10], main = "2-x combined")
+gpip_dist(susieIfs2[11:15], main = "3-x combined")
+gpip_dist(susieIfs2[16:20], main = "4-x combined")
+```
 
@@
 * We have tried different initiation strategies, but the results remain the same.
-
-## Regions: TWAS BF
-
-* Regions: regional TWAS BF > 2
-
-```{r param 5, fig.height=11, fig.width=6}
-tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
-mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
-susieIfs <- paste0(outputdir, "20200813-", tags, ".config12.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config12.susieI.txt")
-
-df <- show_param(mrashfs, susieIfs, susieIfs2)
-par(mfrow = c(4,2))
-plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
-```

### 2020-10-02 analysis/simulation-susieI-ukbchr17to22B-gtex.adipose.Rmd
batch susieI

--- 2020-09-14 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
@@
 ```{r dirs}
-simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
-outputdir <- "~/causalTWAS/simulations/simulation_susieI_20200813/"
-susiedir <- "~/causalTWAS/simulations/simulation_susieI_20200813/"
-mrashdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
-tag <- '20200813-1-1'
+simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200915/"
+outputdir <- "~/causalTWAS/simulations/simulation_susieI_20200915/"
+susiedir <- "~/causalTWAS/simulations/simulation_susieI_20200915/"
+tag <- '20200915-1-1'
 Niter <- 30
@@
 ```{r rpip twas, fig.height=6, fig.width=6}
-simtag <- "20200721-1-1"
-rpipf <- paste0(simdatadir, simtag, ".pgenerprank.txt")
+simtag <- "20200915-1-1"
+rpipf <- paste0(outputdir, simtag, ".pgenerprank.txt")
 
@@
 ```{r rpip twas2, fig.height=6, fig.width=6}
-rpipf <- paste0(simdatadir, simtag, ".p2rprank.txt")
+rpipf <- paste0(outputdir, simtag, ".p2rprank.txt")
 
@@
 
-# MR.ASH regional pip sum distribution
-
-* Use mr.ash for gene only:
-
-```{r rpip 1, fig.height=6, fig.width=6, eval=F}
-rpipf <- paste0(simdatadir, simtag, "-mr.ash.mrashrPIP.txt")
-
-a <- read.table(rpipf, header = T)
-a.c <- a[a$nCausal > 0, ]
-ax <- pretty(0:max(a$rPIP), n = 30)
-
-par(mfrow=c(2,1))
-h1 <- hist(a$rPIP, breaks = 100, xlab = "PIP", main = "PIP distribution-all", col = "grey"); grid()
-h2 <- hist(a.c$rPIP, breaks = h1$breaks, xlab = "PIP", main = "PIP distribution-causal", col = "salmon");grid()
-```
-
-* MR.ash2s for both genes and SNPs using the same data
-
-```{r rpip 2, fig.height=6, fig.width=6}
-rpipf <- paste0(simdatadir, simtag, "-mr.ash2s.lassoes-es.mrash2srPIP.txt")
-
-a <- read.table(rpipf, header = T)
-a.c <- a[a$nCausal > 0 ,]
-ax <- pretty(0:max(a$rPIP), n = 30) # Make a neat vector for the breakpoints
-
-par(mfrow=c(2,1))
-h1 <- hist(a$rPIP, breaks = 100, xlab = "PIP", main = "PIP distribution-all", col = "grey"); grid()
-h2 <- hist(a.c$rPIP, breaks = h1$breaks, xlab = "PIP", main = "PIP distribution-causal", col = "salmon");grid()
-```
-
 # Parameter estimation results
@@
 
-## Regions: causal
+```{r param functions}
+show_param <- function(phenofs, susieIfs, susieIfs2){
+  pars <- do.call(rbind, lapply(phenofs, function(x) {load(read.table(x, header = F, stringsAsFactors = F)[1,1]); 
+    c(phenores$param$pve.gene.truth,
+      phenores$param$pve.snp.truth,
+      length(phenores$param$idx.cgene)/phenores$param$J,
+      length(phenores$param$idx.cSNP)/phenores$param$M)}))
 
-* Regions: We use regional PIP of both SNP and gene from mr.ash2s > 0.3, and also require at least one causal gene or SNP is present, also, at least one gene.
+  colnames(pars) <- c("PVE.gene_truth", "PVE.SNP_truth", "pi1.gene_truth", "pi1.SNP_truth")
 
-* Susie run parameters: `L=1`, `null_weight = 0`, didn't update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform.
-
-```{r param functions}
-show_param <- function(mrashfs, susieIfs, susieIfs2){
-  param <- do.call(rbind, lapply(mr.ashfs, function(x) t(read.table(x))[2:1,]))
-
-  pars <- cbind(param[seq(1,nrow(param), 2), 2], param[seq(1,nrow(param), 2), 4], param[seq(1,nrow(param), 2), 1], param[seq(2,nrow(param), 2), 1], param[seq(1,nrow(param), 2), 3], param[seq(2,nrow(param), 2), 3])
-  colnames(pars) <- c("PVE.gene_truth", "PVE.SNP_truth", paste(rep(c("pi1.gene_", "pi1.SNP_"), each = 2), c("truth", "mr.ash2"), sep = ""))
-
-  param.s <- do.call(rbind, lapply(susieIfs, function(x) {load(x); c(tail(prior.gene_rec, 1), tail(prior.SNP_rec,1))}))
+  param.s <- do.call(rbind, lapply(susieIfs, function(x) {load(x); c(tail(prior.gene_rec[prior.gene_rec!=0], 1), tail(prior.SNP_rec[prior.SNP_rec!=0],1))}))
 
@@
   df <- apply(df[ , 2:ncol(df)], 2, function(x) as.numeric(x))
-  m <- cbind(df[,"pi1.gene_mr.ash2"], 1:nrow(df), 1 + 1:nrow(df)/nrow(df)/3)
   st <-  cbind(df[,"pi1.gene_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
@@
   t <- df[1,"pi1.gene_truth"]
-  dfp <- rbind(m,st,s)
+  dfp <- rbind(st,s)
   plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "gene pi1", xaxt = "n", xlab="", xlim = c(0.8, 3.5), frame.plot=FALSE, ylim = c(0, max(dfp[,1],t) *1.05), ...)
-  axis(side=1, at=1:3, labels = FALSE, tick = F)
-  text(x=1:3, 0, labels = c("mr.ash2", "susieI_truth", "susieI"), xpd = T, pos =1)
+  axis(side=1, at=1:2, labels = FALSE, tick = F)
+  text(x=1:2, 0, labels = c( "susieI_truth", "susieI"), xpd = T, pos =1)
   abline(h=t, lty = 2, col= "salmon", lwd=1.5)
@@
 
-  m <- cbind(df[,"pi1.gene_mr.ash2"]/df[,"pi1.SNP_mr.ash2"], 1:nrow(df), 1 + 1:nrow(df)/nrow(df)/3)
   st <-  cbind(df[,"pi1.gene_susietruth"]/df[,"pi1.SNP_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
@@
   t <- df[1,"pi1.gene_truth"]/df[1,"pi1.SNP_truth"]
-  dfp <- rbind(m,st,s)
+  dfp <- rbind(st,s)
   plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "Enrichment (gene/snp)", xaxt = "n", xlab="", xlim = c(0.8, 3.5),frame.plot=FALSE, ylim = c(0, min(max(dfp[,1],t) *1.05, 150)))
-  axis(side=1, at=1:3, labels = FALSE, tick = F)
-  text(x=1:3, 0, labels = c("mr.ash2", "susieI_truth", "susieI"), xpd = T, pos =1)
+  axis(side=1, at=1:2, labels = FALSE, tick = F)
+  text(x=1:2, 0, labels = c("susieI_truth", "susieI"), xpd = T, pos =1)
   abline(h= t, lty = 2, col= "darkgreen", lwd=1.5)
@@
 }
+
+gpip_dist <- function(susiefs, ...){
+  dflist <- list()
+  for (f in susiefs){
+    dflist[[f]] <- read.table(f, header =T , stringsAsFactors = F)
+  }
+  df <- do.call(rbind, dflist)
+  hist(df[df$type == "gene", "susie_pip"], xlab = "gene susie PIP", 
+       breaks = 50, ylim = c(0,20), xlim=c(0,1), col = "salmon", ...)
+}
 ```
 
-```{r param, fig.height=8, fig.width=6}
-tags <- paste(rep(1:3, each = 5), 1:5, sep = "-")
-mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
-susieIfs <- paste0(outputdir, "20200813-", tags, ".susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20200813-", tags, ".1.txt")
-
-df <- show_param(mrashfs, susieIfs, susieIfs2)
-par(mfrow = c(3,2))
-plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05") 
-plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-```
-
-## Regions: mrash gene+SNP (1)
-
-* Regions: We use regional PIP of both SNP and gene from mr.ash2s > 0.3.
-
-* Susie run parameters: `L=2`, initialize `null_weight = 0`, update `null_weight` in iterations.
-We initialize with prior for genes and SNPs as uniform.
-(We also tried initialize with prior for genes and SNPs as 0.05 and 2.5e-3, results are the same.)
-
-```{r param-6, fig.height=11, fig.width=6}
-tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
-mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
-susieIfs <- paste0(outputdir, "20200813-", tags, ".config6.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config6.susieI.txt")
-
-df <- show_param(mrashfs, susieIfs, susieIfs2)
-par(mfrow = c(4,2))
-plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
-```
-
-## Regions: mrash gene+SNP (2)
-
-* Regions: We use regional PIP of both SNP and gene from mr.ash2s > 0.3.
-
-* Susie run parameters: `L=2`, `null_weight = 0`. Didn't update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform.
-
-```{r param-7, fig.height=11, fig.width=6}
-tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
-susieIfs <- paste0(outputdir, "20200813-", tags, ".config7.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config7.susieI.txt")
-
-df <- show_param(mrashfs, susieIfs, susieIfs2)
-par(mfrow = c(4,2))
-plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
-```
-
-## Regions: mrash gene only
-
-* Regions: We use regional sum of PIP for genes from mr.ash for genes only > 0.1.
-
-* Susie run parameters: `L=2`. Update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform, `null_weight = 0`.
-
-```{r param-9, fig.height=11, fig.width=6}
-tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
-susieIfs <- paste0(outputdir, "20200813-", tags, ".config9.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config9.susieI.txt")
-
-df <- show_param(mrashfs, susieIfs, susieIfs2)
-par(mfrow = c(4,2))
-plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
-```
-
-## Regions: TWAS gene only
-
-* Regions: We use regional min p value for genes from twas for genes only, requiring this p value in top 5%.
-
-* Susie run parameters: `L=2`. Update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform, `null_weight = 0`.
-
-```{r param-10, fig.height=11, fig.width=6}
-tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
-susieIfs <- paste0(outputdir, "20200813-", tags, ".config10.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config10.susieI.txt")
-
-df <- show_param(mrashfs, susieIfs, susieIfs2)
-par(mfrow = c(4,2))
-plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
-```
-
-## Regions: TWAS gene+SNP
-
-* Regions: We use regional min p value for genes or SNPs from twas for genes and snps, requiring this p value in top 5% for genes or 1% for SNPs.
-
-* Susie run parameters: `L=2`. Update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform, `null_weight = 0`.
-
-```{r param-11, fig.height=11, fig.width=6}
-tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
-susieIfs <- paste0(outputdir, "20200813-", tags, ".config11.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config11.susieI.txt")
-
-df <- show_param(mrashfs, susieIfs, susieIfs2)
-par(mfrow = c(4,2))
-plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
-```
-
-## Regions: all (1)
+## Regions: all
 
@@
 
-* Susie run parameters: `L=2`. initialize with `null_weight = 0` and update `null_weight` based on last iteration results. We initialize with prior for genes and SNPs as uniform.
+* Susie run parameters: `L=2`.
+initialize with `null_weight = 0` and update `null_weight` based on last iteration results.
+We initialize with prior for genes and SNPs as uniform.
+gene.pve ~ 0.01, snp.pve ~ 0.05.
 
-```{r param3, fig.height=11, fig.width=6}
-mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
-susieIfs <- paste0(outputdir, "20200813-", tags, ".config2.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config2.susieI.txt")
+```{r param1, fig.height=6, fig.width=6}
+tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
+phenofs <- paste0(simdatadir, "simu_20200915-", tags, "-pheno.txt")
+susieIfs <- paste0(outputdir, "20200915-", tags, ".config1.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20200915-", tags, ".config1.susieI.txt")
 
-df <- show_param(mrashfs, susieIfs, susieIfs2)
-par(mfrow = c(4,2))
-plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
+df <- show_param(phenofs, susieIfs, susieIfs2)
+par(mfrow = c(2,2))
+plot_param(df[1:5,], main = "low power")
+plot_param(df[6:10,], main = "high power")
 ```
 
-## Regions: all (2)
-
-* Regions: All regions.
-
-* Susie run parameters: `L=1`. Did not update `null_weight` in each iteration. We initialize with prior for genes and SNPs as uniform.
-
-```{r param4, fig.height=11, fig.width=6}
-mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
-susieIfs <- paste0(outputdir, "20200813-", tags, ".config3.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config3.susieI.txt")
-
-df <- show_param(mrashfs, susieIfs, susieIfs2)
-par(mfrow = c(4,2))
-plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
-plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
+```{r pipdist-1,fig.height=3, fig.width=12}
+par(mfrow = c(1,2))
+gpip_dist(susieIfs2[1:5], main = "low power")
+gpip_dist(susieIfs2[6:10], main = "high power")
 ```
-
-## Regions: all (others)
-* Susie run parameters: `L=1`.
-initialize with null_weight = 0 and update null_weight based on last iteration results.
-We initialize with prior for genes and SNPs as uniform.
-The prior for genes is underestimated for low power settings (1-x, 2-x), many converged to close to 0, for high power settings it is better.
-* Susie run parameters: `L=5`.
-initialize with null_weight = 0 and update null_weight based on last iteration results.
-The prior for gene estimation is fine for all runs, however prior for SNPs overestimated about 10 times.
-* We have tried different initiation strategies, but the results remain the same.

### 2020-10-02 code/run_UKB_process.R

@@ -75,8 +75,8 @@ setwd("/home/simingz/causalTWAS/ukbiobank")
 for (i in 1:6) {
   load("/home/simingz/causalTWAS/ukbiobank/ukb_chr17to22_s20000.FBM.Rd")
   tag <- paste0("-B", i)
-  dat$chr <- paste0(dat$chr, tag)
-  dat$snp <- paste0(dat$snp, tag)
+  dat$chr <- as.matrix(paste0(dat$chr, tag), ncol = 1)
+  dat$snp <- as.matrix(paste0(dat$snp, tag), ncol = 1)
   save(dat, file = paste0("ukb_chr17to22_s20000", tag, ".FBM.Rd"))
 }
 
@@ -118,13 +118,12 @@ write.table(outdfall , file= posf, row.names=F, col.names=T, sep="\t", quote = F
 # region files
 setwd("/home/simingz/causalTWAS/simulations/shared_files")
 regfile0 <- "chr17-22-500kb_bins.txt"
-regfile <- "chr17-22-500kb_B_bins.txt"
-reglist <- list()
 for (i in 1:6) {
+  regfile <- paste0("chr17-22-500kb_B", i, "_bins.txt")
   tag <- paste0("-B", i)
   reg <- read.table(regfile0, header =T)
   reg[,1] <- paste0(reg[,1], tag)
-  reglist[[i]] <- reg
+  write.table(reg, file= regfile , row.names=F, col.names=T, sep="\t", quote = F)
 }
-regout <- do.call(rbind, reglist)
-write.table(regout , file= regfile , row.names=F, col.names=T, sep="\t", quote = F)
+
+

### 2020-10-14 analysis/description.Rmd
wg susieI

--- 2020-09-06 analysis/description.Rmd
@@
 
-```{r param}
-n <- 20000 # number of samples
-p <- 86000 # number of SNPs
+```{r parameters}
+n <- 40000 # number of samples
+p <- 656321 # number of SNPs
+J <- 8021 # number of genes
 ```
 
-The genotype data we used is from UKB biobank, randomly selecting n=`r as.character(n)` samples.
-We use SNP genotype data from chr 17 to chr 22 combined.
-These genomic regions represents 12.5% of the genome.
+The genotype data we used is from UKB biobank, randomly selecting n = `r as.character(n)` samples.
+We use SNP genotype data from chr 1 to chr 22 combined from UKB.
 SNPs are downsampled to 1/10 (randomly), eQTLs (see below for definition of eQTL) were added back.
@@
 This dataset contains ~ 380 samples.
-FUSION/TWAS were used train expression model and we used their lasso results.
-There are around 1700 genes with expression
+[FUSION/TWAS](http://gusevlab.org/projects/fusion/) were used to train expression model and we used their lasso results.
+SNPs included in eQTL anlaysis are restricted to cis-locus 500kb on either side of the gene boundary.
+eQTLs are defined as SNPs with abs(effectize) > 1e-8 in lasso results.
+
+2. Impute expression.
+
+We impute gene expression for our genotype data using expression models obtained from step 1.
+There are `r as.character(J)` genes with expression model from chr17 to chr22.
+We imputed expression from genotypes using the expression predictors.
+
+3. Define and select regions
+
+Next, the analysis is done at the “region” level, which is 500kb bins along the genome.
+Each bin would contain all the SNPs, as well as all the genes in that bin.
+We are exploring several ways to select regions that contain true signals, e.g.
+based on regional sum of mr.ash PIP for genes/SNPs, region smallest TWAS p value for gene/SNPs, or regional bayes factors, etc.
+
+4. Run susie iteratively
+We then run susie for each of these regions.
+So the features of SuSiE are: SNPs and "genes" (not cis-eQTLs of that gene).
+We use the same prior for all SNPs and another prior for all "genes" when running SUSIE.
+In some settings, we also run SUSIE with null weight, which is calculated as `1- prior.SNP * n.SNP - prior.gene * n.gene`.
+We obtain the PIP for SNPs and gene in the region.
+After we run susie for all regions (one iteration), we take the average of all SNP PIPs as the prior of SNPs for the next iteration and similarly for the prior for genes.
+
+5. We obtain PIP for genes from the last iteration as results.

### 2020-10-14 analysis/simulation-susieI-ukbWG-gtex.adipose.Rmd
wg susieI

---
title: "Test susieI using whole genome, 4k ukb samples"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  message= F, warning = F)
```
```{r libs}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})
library(tidyr)
library(plyr)
library(stringr)
library(kableExtra)
source("analysis/summarize_twas_plots.R")
```

```{r des, child = here::here("analysis/description.Rmd")}
```

```{r dirs}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
outputdir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
susiedir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
tag <- '20201001-1-1'
Niter <- 20
```

# p value distribution

* TWAS p value of genes:

```{r gene twas, fig.height=6, fig.width=8}
simtag <- "20201001-1-1"
exprgwasf <- paste0(simdatadir, simtag, ".exprgwas.txt.gz")
load(paste0(simdatadir, "simu_", simtag, "-pheno.Rd"))
caulist <- list()
for (chrom in 1:22) {
  load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom ,"_s40000.FBM.Rd"))
  load(paste0(simdatadir, "simu_s40000_GTEXadipose-B", chrom, "-cis-expr.Rd"))
  caulist[[chrom]]<- c(exprres$gnames[phenores$batch[[chrom]]$param$idx.cgene], dat$snp[phenores$batch[[chrom]]$param$idx.cSNP,])
}
cau <- unlist(caulist)

a <- read.table(exprgwasf, header = T)
a$ifcausal <- ifelse(a$MARKER_ID %in% cau, 1, 0)
ax <- pretty(0:max(-log10(a$PVALUE)), n = 30)

par(mfrow=c(3,1))
h1 <- hist(-log10(a$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey"); grid()
h2 <- hist(-log10(a[a$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon");grid()

plot(a[a$X.CHROM ==22, ]$BEGIN, -log10(a[a$X.CHROM ==22, ]$PVALUE), col = a[a$X.CHROM ==22, ]$ifcausal + 1, xlab = paste0("chr", chrom), ylab = "-log10(pvalue)")
points(a[a$X.CHROM ==22 & a$ifcausal ==1, ]$BEGIN, -log10(a[a$X.CHROM ==22 & a$ifcausal ==1, ]$PVALUE), col = "red", pch =19)
```

```{r eval=F}
r <- do.call(cbind, lapply(1:22, function(x) phenores$batch[[x]]$Y.g))
GWAA(exprres$expr, matrix(rowSums(r[, c(1:12,14:21)]) + rnorm(N), ncol=1), snpname = exprres$gnames, anno = anno, "temp", family = gaussian, ncore = 1, nSplits = 1, compress = F)
m <- read.table("temp", header =T, comment.char = "")
m$ifcausal <- ifelse(m$MARKER_ID %in% cau, 1, 0)
plot(m[m$X.CHROM ==22, ]$BEGIN, -log10(m[m$X.CHROM ==22, ]$PVALUE), col = m$ifcausal + 1)

snpname = dat$snp[,1]
anno <- cbind(dat$chr, dat$pos, dat$pos) # EPACT format

GWAA(dat$G, r[,13, drop=F], snpname = snpname, anno = anno, "tempsnp", family = gaussian, ncore = 5, nSplits = 100, compress = F)
```

* TWAS p value of SNPs:

```{r snp twas, fig.height=6, fig.width=8}
simtag <- "20201001-1-1"
snpgwasf <- paste0(simdatadir, simtag, ".snpgwas.txt.gz")

b <- read.table(snpgwasf, header = T)
b$ifcausal <- ifelse(b$MARKER_ID %in% cau, 1, 0)
ax <- pretty(0:max(-log10(b$PVALUE)), n = 30)

par(mfrow=c(3,1))
h1 <- hist(-log10(b$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey"); grid()
h2 <- hist(-log10(b[b$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon");grid()

plot(b[b$X.CHROM ==22, ]$BEGIN, -log10(b[b$X.CHROM ==22, ]$PVALUE), col = b$ifcausal + 1, xlab = paste0("chr", chrom), ylab = "-log10(pvalue)")
points(b[b$X.CHROM ==22 & b$ifcausal ==1, ]$BEGIN, -log10(b[b$X.CHROM ==22 & b$ifcausal ==1, ]$PVALUE), col = "red", pch =19)
```

# Parameter estimation results

Results: Each row shows parameter estimation results from 5 simulation runs with similar settings (i.e. pi1 and PVE for genes and SNPs). each row has two plots, one for gene pi1 estimation, one for enrichment (gene pi1/snp pi1). Results from each run were represented by one dot, dots with the same color come from the same run. horizontal dash lines: simulation truth,  `susietruth`, the truth in selected regions that were used to run susie iteractively (susieI).

```{r param functions}
show_param <- function(phenofs, susieIfs, susieIfs2){
  pars <- do.call(rbind, lapply(phenofs, function(x) {load(x); 
    c(phenores$param$pve.gene.truth,
      phenores$param$pve.snp.truth,
      length(phenores$batch[[1]]$param$idx.cgene)/phenores$batch[[1]]$param$J,
      length(phenores$batch[[1]]$param$idx.cSNP)/phenores$batch[[1]]$param$M)}))

  colnames(pars) <- c("PVE.gene_truth", "PVE.SNP_truth", "pi1.gene_truth", "pi1.SNP_truth")

  param.s <- do.call(rbind, lapply(susieIfs, function(x) {load(x); c(tail(prior.gene_rec[prior.gene_rec!=0], 1), tail(prior.SNP_rec[prior.SNP_rec!=0],1))}))

  param.s.truth <- do.call(rbind, lapply(susieIfs2, function(x) {
    a <- fread(x, header = T);
    c(nrow(a[a$ifcausal == 1 & a$type == "gene" ])/ nrow(a[a$type == "gene"]),
      nrow(a[a$ifcausal == 1 & a$type == "SNP"])/ nrow(a[a$type == "SNP"]))
    }))

  pars.s <- cbind(param.s.truth, param.s)[, c(1,3,2,4)]
  colnames(pars.s) <-  paste(rep(c("pi1.gene_", "pi1.SNP_"), each = 2), c("susietruth", "susieI"), sep = "")

  df <- cbind(tags, format(pars, digits = 4), format(pars.s, digits =4))
  rownames(df) <- NULL
  return(df)

  # df %>% 
  # kable("html", escape = F) %>%
  # kable_styling("striped", full_width = F) %>%
  #  row_spec(c(1:5, 11:15), background = "#FEF3B9") %>%
  # scroll_box(width = "100%", height = "600px", fixed_thead = T)
}

plot_param <- function(df, ...){
  df <- apply(df[ , 2:ncol(df)], 2, function(x) as.numeric(x))
  st <-  cbind(df[,"pi1.gene_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
  s <- cbind(df[,"pi1.gene_susieI"], 1:nrow(df), 3 + 1:nrow(df)/nrow(df)/3)
  t <- df[1,"pi1.gene_truth"]
  dfp <- rbind(st,s)
  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "gene pi1", xaxt = "n", xlab="", xlim = c(0.8, 3.5), frame.plot=FALSE, ylim = c(0, max(dfp[,1],t) *1.05), ...)
  axis(side=1, at=1:2, labels = FALSE, tick = F)
  text(x=2:3, 0, labels = c( "susieI_truth", "susieI"), xpd = T, pos =1)
  abline(h=t, lty = 2, col= "salmon", lwd=1.5)
  grid()

  st <-  cbind(df[,"pi1.gene_susietruth"]/df[,"pi1.SNP_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
  s <- cbind(df[,"pi1.gene_susieI"]/df[,"pi1.SNP_susieI"], 1:nrow(df), 3 + 1:nrow(df)/nrow(df)/3)
  t <- df[1,"pi1.gene_truth"]/df[1,"pi1.SNP_truth"]
  dfp <- rbind(st,s)
  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "Enrichment (gene/snp)", xaxt = "n", xlab="", xlim = c(0.8, 3.5),frame.plot=FALSE, ylim = c(0, min(max(dfp[,1],t) *1.05, 150)))
  axis(side=1, at=1:2, labels = FALSE, tick = F)
  text(x=2:3, 0, labels = c("susieI_truth", "susieI"), xpd = T, pos =1)
  abline(h= t, lty = 2, col= "darkgreen", lwd=1.5)
  grid()
}

gpip_dist <- function(susiefs, ...){
  dflist <- list()
  for (f in susiefs){
    dflist[[f]] <- read.table(f, header =T , stringsAsFactors = F)
  }
  df <- do.call(rbind, dflist)
  hist(df[df$type == "gene", "susie_pip"], xlab = "gene susie PIP", 
       breaks = 50, ylim = c(0,20), xlim=c(0,1), col = "salmon", ...)
}
```

## Regions: all (1)

* Susie run parameters: `L=2`. initialize with `null_weight = 0` and update `null_weight` based on last iteration results. We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.

```{r param1, fig.height=6, fig.width=6}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")

df <- show_param(phenofs, susieIfs, susieIfs2)
par(mfrow = c(2,2))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```

## Regions: causal

* Susie run parameters: `L=2`. initialize with `null_weight = 0` and update `null_weight` based on last iteration results. We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.

```{r param5, fig.height=6, fig.width=6}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config5.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config5.susieI.txt")

df <- show_param(phenofs, susieIfs, susieIfs2)
par(mfrow = c(2,2))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```

### 2020-10-14 analysis/simulation-susieI-ukbchr17to22B-gtex.adipose.Rmd
wg susieI

@@
 ---
-title: "Test susieI using chr17 to 22"
+title: "Test susieI using chr17 to 22, multiple batch"
 output: html_document
@@
 ```
+# Summary
 
-# Regional minimum p value distribution
-
-* minimum TWAS p value of genes for each region:
-
-```{r rpip twas, fig.height=6, fig.width=6}
-simtag <- "20200915-1-1"
-rpipf <- paste0(outputdir, simtag, ".pgenerprank.txt")
-
-a <- read.table(rpipf, header = T)
-a.c <- a[a$nCausal > 0, ]
-ax <- pretty(0:max(-log10(a$rpmin)), n = 30)
-
-par(mfrow=c(2,1))
-h1 <- hist(-log10(a$rpmin), breaks = 100, xlab = "PIP", main = "min P value distribution-all", col = "grey"); grid()
-h2 <- hist(-log10(a.c$rpmin), breaks = h1$breaks, xlab = "PIP", main = "min p distribution-causal", col = "salmon");grid()
-```
-
-* minimum TWAS p value of genes and SNPs for each region:
-
-```{r rpip twas2, fig.height=6, fig.width=6}
-rpipf <- paste0(outputdir, simtag, ".p2rprank.txt")
-
-a <- read.table(rpipf, header = T)
-a.c <- a[a$nCausal > 0, ]
-ax <- pretty(0:max(-log10(a$rpmin)), n = 30)
-
-par(mfrow=c(2,1))
-h1 <- hist(-log10(a$rpmin), breaks = 100, xlab = "PIP", main = "min P value distribution-all", col = "grey"); grid()
-h2 <- hist(-log10(a.c$rpmin), breaks = h1$breaks, xlab = "PIP", main = "min p distribution-causal", col = "salmon");grid()
-```
+This result is problematic. The artificial perfectly correlated LD leads to overestimated pi1.
 

### 2020-10-14 code/run_simulate_pheno.R

@@ -1,0 +1,95 @@
+library(data.table)
+library(logging)
+library(tools)
+
+args = commandArgs(trailingOnly=TRUE)
+if (length(args) != 4) {
+  stop("4 arguments must be supplied:
+       * genotype file name (or multiple file names in a .txt file)
+       * impute expr Rd file (or multiple file names in a .txt file)
+       * param R file
+       * out file name", call.=FALSE)
+}
+
+addHandler(writeToFile, file="run_simulate_phenotype.R.log", level='DEBUG')
+loginfo('script started ... ')
+
+loginfo("input arg 1 (genotype): %s ", args[1])
+loginfo("input arg 2 (imputed expression): %s ", args[2])
+loginfo("input arg 3 (param): %s ", args[3])
+loginfo("input arg 4 (outname): %s ", args[4])
+
+
+codedir <- "/project2/mstephens/causalTWAS/causal-TWAS/code/"
+source(paste0(codedir, "stats_func.R"))
+source(paste0(codedir,"input_reformat.R"))
+source(paste0(codedir,"simulate_phenotype.R"))
+
+
+outname <- args[4]
+
+# prepare input files. Genotype and expression are already scaled.
+pfile <- args[1]
+
+if (file_ext(pfile) == "txt"){
+  pfiles <- read.table(pfile, header = F, stringsAsFactors = F)[,1]
+  efiles <- read.table(args[2], header =F, stringsAsFactors = F)[,1]
+} else {
+  pfiles <- pfile
+  efiles <- args[2]
+}
+pfileRds <- paste0(drop_ext(pfiles), ".FBM.Rd")
+
+
+# set sigma based on pve
+source(args[3])
+set.seed(SED)
+
+nlist <- list()
+for ( b in 1:length(pfileRds)){
+  load(pfileRds[b])
+  load(efiles[b])
+  nlist[[b]] <- c(ncol(dat$G), ncol(exprres$expr))
+}
+ndf <- do.call(rbind, nlist)
+M <- sum(ndf[,1])
+J <- sum(ndf[,2])
+N <- nrow(dat$G)
+expr.meanvar <- 1 # expression has been scaled
+
+J.c <- round(J * pi_beta)
+M.c <- round(M * pi_theta)
+
+sigma_beta <- sqrt(pve.expr / (J.c * expr.meanvar * (1 - pve.snp - pve.expr)))
+sigma_theta <- sqrt(pve.snp / (M.c * (1 - pve.snp - pve.expr)))
+
+# Simulate phenotype
[…]

### 2020-10-14 code/workflow/master_run.sh

@@ -1,20 +1,0 @@
-source activate twas;cd /project2/mstephens/causalTWAS/simulations/simulation_susieI_20200915/;Rscript /project2/mstephens/causalTWAS/causal-TWAS/code//run_test_susieI.R /project2/mstephens/causalTWAS/ukbiobank/ukb_chr17to22_s20000-B_fn.txt /project2/mstephens/causalTWAS/simulations/simulation_ashtest_20200915//simu_20200915-1-1-cis-expr.Rd /project2/mstephens/causalTWAS/simulations/simulation_ashtest_20200915//simu_20200915-1-1-pheno.Rd /home/simingz/causalTWAS/simulations/simulation_ashtest_20200721/20200915-1-1 20200915-1-1.config1 susieI_config1.R
-source activate twas;cd /project2/mstephens/causalTWAS/simulations/simulation_susieI_20200915/;Rscript /project2/mstephens/causalTWAS/causal-TWAS/code//run_test_susieI.R /project2/mstephens/causalTWAS/ukbiobank/ukb_chr17to22_s20000-B_fn.txt /project2/mstephens/causalTWAS/simulations/simulation_ashtest_20200915//simu_20200915-1-2-cis-expr.Rd /project2/mstephens/causalTWAS/simulations/simulation_ashtest_20200915//simu_20200915-1-2-pheno.Rd /home/simingz/causalTWAS/simulations/simulation_ashtest_20200721/20200915-1-2 20200915-1-2.config1 susieI_config1.R
-source activate twas;cd /project2/mstephens/causalTWAS/simulations/simulation_susieI_20200915/;Rscript /project2/mstephens/causalTWAS/causal-TWAS/code//run_test_susieI.R /project2/mstephens/causalTWAS/ukbiobank/ukb_chr17to22_s20000-B_fn.txt /project2/mstephens/causalTWAS/simulations/simulation_ashtest_20200915//simu_20200915-1-3-cis-expr.Rd /project2/mstephens/causalTWAS/simulations/simulation_ashtest_20200915//simu_20200915-1-3-pheno.Rd /home/simingz/causalTWAS/simulations/simulation_ashtest_20200721/20200915-1-3 20200915-1-3.config1 susieI_config1.R
[…]

### 2020-10-14 code/workflow/submit_parallel_master_run.sh.sh

@@ -1,14 +1,0 @@
-#!/bin/sh
-
-#SBATCH --time=1-12:00:00
-#SBATCH --array=1-20
-#SBATCH --nodes=1
-#SBATCH --ntasks-per-node=1
-#SBATCH --mem=40G
-#SBATCH --job-name=master_run.sh
-#SBATCH --output=master_run.sh_%A_%a.out
-#SBATCH --error=master_run.sh_%A_%a.err
-#SBATCH --partition=mstephens
-
-head -n $SLURM_ARRAY_TASK_ID master_run.sh | tail -1 | sh
-mkdir master_run.sh_JOBID_$SLURM_ARRAY_JOB_ID

### 2020-10-20 analysis/description.Rmd
small pvalues

@@
 ```{r parameters}
-n <- 40000 # number of samples
-p <- 656321 # number of SNPs
-J <- 8021 # number of genes
+n <- 20000 # number of samples
+p <- 86000 # number of SNPs
+J <- 1653 # number of genes
 ```
@@
 The genotype data we used is from UKB biobank, randomly selecting n = `r as.character(n)` samples.
-We use SNP genotype data from chr 1 to chr 22 combined from UKB.
+We use SNP genotype data from chr 17 to chr 22 combined.
+These genomic regions represents 12.5% of the genome.
 SNPs are downsampled to 1/10 (randomly), eQTLs (see below for definition of eQTL) were added back.

### 2020-10-20 analysis/simulation-susieI-ukbWG-gtex.adipose.Rmd
small pvalues

@@
 
-```{r des, child = here::here("analysis/description.Rmd")}
+# Analysis description
+
+```{r parameters}
+n <- 40000 # number of samples
+p <- 656321 # number of SNPs
+J <- 8021 # number of genes
 ```
+
+The genotype data we used is from UKB biobank, randomly selecting n = `r as.character(n)` samples.
+We use SNP genotype data from chr 1 to chr 22 combined from UKB.
+SNPs are downsampled to 1/10 (randomly), eQTLs (see below for definition of eQTL) were added back.
+This ends up with p = `p as.charater(p)` SNPs.
+
+Our analysis consists of the following steps:
+
+1. Build expression predictors using another expression-genotype dataset.
+
+The one we used in this analysis is GTEx Adipose tissue v7 dataset.
+This dataset contains ~ 380 samples.
+[FUSION/TWAS](http://gusevlab.org/projects/fusion/) were used to train expression model and we used their lasso results.
+SNPs included in eQTL anlaysis are restricted to cis-locus 500kb on either side of the gene boundary.
+eQTLs are defined as SNPs with abs(effectize) > 1e-8 in lasso results.
+
+2. Impute expression.
+
+We impute gene expression for our genotype data using expression models obtained from step 1.
+There are `r as.character(J)` genes with expression model.
+We imputed expression from genotypes using the expression predictors.
+
+3. Define and select regions
+
+Next, the analysis is done at the “region” level, which is 500kb bins along the genome.
+Each bin would contain all the SNPs, as well as all the genes in that bin.
+We are exploring several ways to select regions that contain true signals, e.g.
+based on regional sum of mr.ash PIP for genes/SNPs, region smallest TWAS p value for gene/SNPs, or regional bayes factors, etc.
+
+4. Run susie iteratively
+We then run susie for each of these regions.
+So the features of SuSiE are: SNPs and "genes" (not cis-eQTLs of that gene).
+We use the same prior for all SNPs and another prior for all "genes" when running SUSIE.
+In some settings, we also run SUSIE with null weight, which is calculated as `1- prior.SNP * n.SNP - prior.gene * n.gene`.
+We obtain the PIP for SNPs and gene in the region.
+After we run susie for all regions (one iteration), we take the average of all SNP PIPs as the prior of SNPs for the next iteration and similarly for the prior for genes.
+
+5. We obtain PIP for genes from the last iteration as results.
 
@@
 susiedir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
-tag <- '20201001-1-1'
-Niter <- 20
+simtag <- '20201001-1-3'
+source('~/causalTWAS/causal-TWAS/code/gwas.R')
+source('~/causalTWAS/causal-TWAS/code/ld.R')
 ```
 
-# p value distribution
-
-* TWAS p value of genes:
-
-```{r gene twas, fig.height=6, fig.width=8}
-simtag <- "20201001-1-1"
+```{r causal}
 exprgwasf <- paste0(simdatadir, simtag, ".exprgwas.txt.gz")
@@
 cau <- unlist(caulist)
+```
 
+# p value distribution
+
+* TWAS p value of genes:
+
+```{r gene twas, fig.height=8, fig.width=8}
+chrom <- 22
 a <- read.table(exprgwasf, header = T)
@@
 
-plot(a[a$X.CHROM ==22, ]$BEGIN, -log10(a[a$X.CHROM ==22, ]$PVALUE), col = a[a$X.CHROM ==22, ]$ifcausal + 1, xlab = paste0("chr", chrom), ylab = "-log10(pvalue)")
-points(a[a$X.CHROM ==22 & a$ifcausal ==1, ]$BEGIN, -log10(a[a$X.CHROM ==22 & a$ifcausal ==1, ]$PVALUE), col = "red", pch =19)
-```
-
-```{r eval=F}
-r <- do.call(cbind, lapply(1:22, function(x) phenores$batch[[x]]$Y.g))
-GWAA(exprres$expr, matrix(rowSums(r[, c(1:12,14:21)]) + rnorm(N), ncol=1), snpname = exprres$gnames, anno = anno, "temp", family = gaussian, ncore = 1, nSplits = 1, compress = F)
-m <- read.table("temp", header =T, comment.char = "")
-m$ifcausal <- ifelse(m$MARKER_ID %in% cau, 1, 0)
-plot(m[m$X.CHROM ==22, ]$BEGIN, -log10(m[m$X.CHROM ==22, ]$PVALUE), col = m$ifcausal + 1)
-
-snpname = dat$snp[,1]
-anno <- cbind(dat$chr, dat$pos, dat$pos) # EPACT format
-
-GWAA(dat$G, r[,13, drop=F], snpname = snpname, anno = anno, "tempsnp", family = gaussian, ncore = 5, nSplits = 100, compress = F)
+plot(a[a$X.CHROM ==chrom, ]$BEGIN, -log10(a[a$X.CHROM ==chrom, ]$PVALUE), col = a[a$X.CHROM ==chrom, ]$ifcausal + 1, xlab = paste0("chr", chrom), ylab = "-log10(pvalue)")
+points(a[a$X.CHROM ==chrom & a$ifcausal ==1, ]$BEGIN, -log10(a[a$X.CHROM ==chrom & a$ifcausal ==1, ]$PVALUE), col = "red", pch =19)
+grid()
 ```
@@
 
-```{r snp twas, fig.height=6, fig.width=8}
-simtag <- "20201001-1-1"
+```{r snp twas, fig.height=8, fig.width=8}
 snpgwasf <- paste0(simdatadir, simtag, ".snpgwas.txt.gz")
 
-b <- read.table(snpgwasf, header = T)
+b <- fread(snpgwasf, header = T)
 b$ifcausal <- ifelse(b$MARKER_ID %in% cau, 1, 0)
@@
 
-plot(b[b$X.CHROM ==22, ]$BEGIN, -log10(b[b$X.CHROM ==22, ]$PVALUE), col = b$ifcausal + 1, xlab = paste0("chr", chrom), ylab = "-log10(pvalue)")
-points(b[b$X.CHROM ==22 & b$ifcausal ==1, ]$BEGIN, -log10(b[b$X.CHROM ==22 & b$ifcausal ==1, ]$PVALUE), col = "red", pch =19)
+plot(b[b$X.CHROM ==chrom, ]$BEGIN, -log10(b[b$X.CHROM ==chrom, ]$PVALUE), col = b[b$X.CHROM ==chrom,]$ifcausal + 1, xlab = paste0("chr", chrom), ylab = "-log10(pvalue)")
+points(b[b$X.CHROM ==chrom & b$ifcausal ==1, ]$BEGIN, -log10(b[b$X.CHROM ==chrom & b$ifcausal ==1, ]$PVALUE), col = "red", pch =19)
+grid()
+```
+
+# Exploration of unusual small p values
+Take chromosom 22 geno type and perform linear regression for against geno type for causal SNPs on other chromosomes.
+We see some very significant p values.
+Take chromosome 2 causal SNPs as an example.
+
+## normal distribution as phenotype
+```{r random, fig.height=4, fig.width=8}
+chrom <- 22
+load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom ,"_s40000.FBM.Rd"))
+load(paste0(simdatadir, "simu_s40000_GTEXadipose-B", chrom, "-cis-expr.Rd"))
+load(paste0(simdatadir, "simu_", simtag, "-pheno.Rd"))
+
+N<- nrow(phenores$Y)
+cat("variance of the phenotype: ", var(phenores$Y), "\n")
+
+snpname = dat$snp[,1]
+anno <- cbind(dat$chr, dat$pos, dat$pos)
+
+outname <- "/home/simingz/causalTWAS/simulations/simulation_susieI_20201001/chr22_plots/chr22snpRandom"
+# GWAA(dat$G, matrix(rnorm(N, sd = sqrt(var(phenores$Y))), ncol =1), snpname = snpname, anno = anno, outname , family = gaussian, ncore = 5, nSplits = 30, compress = F)
+n <- read.table(outname, header =T, comment.char = "")
+plot(n$BEGIN, -log10(n$PVALUE), main = "Random", xlab = "position", ylab = "-log10(p)")
+grid()
+```
+
+## genetic component of chromosome 2 as phenotype
+
+Not all chromosome give small p values. chr2 is the one with the smallest p values.
+```{r chrfixed, fig.height=4, fig.width=8}
+chrom.p <- 2
+r <- do.call(cbind, lapply(1:22, function(x) phenores$batch[[x]]$Y.g))
+cat("variance of the genetic component: ", var(r[,chrom.p]))
+
+outname <- "/home/simingz/causalTWAS/simulations/simulation_susieI_20201001/chr22_plots/chr22snpVschr2fixed"
+# GWAA(dat$G, matrix(r[,chrom.p], ncol =1), snpname = snpname, anno = anno, outname, family = gaussian, ncore = 5, nSplits = 30, compress = F)
+n <- read.table(outname, header =T, comment.char = "", stringsAsFactors = F)
+plot(n$BEGIN, -log10(n$PVALUE), main = "chr2 genetic", xlab = "position", ylab = "-log10(p)")
+grid()
+```
+
+## genotype of causal snps on chromosome 2 as phenotype
+
+genotype of causal snps on chromosome 2 as phenotype.
+There are `r length(phenores$batch[[2]]$param$idx.cSNP)` causal SNPs.
+Not all of them significantly associated with SNP genotype on chr22, but quite a few.
+below are examples (only a subset of SNPs on chr22 are selected to run regression analysis with chr2 causal SNP genotypes.)
+
+```{r geno, fig.height=4, fig.width=8, message=F}
+chrom <- 22
+load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom ,"_s40000.FBM.Rd"))
+geno <- dat$G[, match(n[n$PVALUE < 1e-40, "MARKER_ID"], dat$snp[,1])]
+load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom ,"_s40000.unscaled.FBM.Rd"))
+geno.unscaled <-  dat$G[, match(n[n$PVALUE < 1e-40, "MARKER_ID"], dat$snp[,1])]
+
+snpname = dat$snp[match(n[n$PVALUE < 1e-40, "MARKER_ID"], dat$snp[,1]),1]
+anno <- cbind(dat$chr, dat$pos, dat$pos)[match(n[n$PVALUE < 1e-40, "MARKER_ID"], dat$snp[,1]), ]
+
+load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom.p ,"_s40000.FBM.Rd"))
+geno.c <- dat$G[, phenores$batch[[chrom.p]]$param$idx.cSNP]
+geno.c.names <- dat$snp[phenores$batch[[chrom.p]]$param$idx.cSNP,]
+load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom.p ,"_s40000.unscaled.FBM.Rd"))
+geno.c.unscaled <- dat$G[, phenores$batch[[chrom.p]]$param$idx.cSNP]
+
+ac <- ac <- apply(round(geno.unscaled), 2, table)
+maf <- apply(cbind( (ac[1,] * 2  + ac[2,]) / N/2, 1- (ac[1,] * 2  + ac[2,]) /N/2), 1, min)
+
+geno.c.select <- c(1, 23, 45, 65)
+
+for (i in geno.c.select){
+  outname <- paste0("/home/simingz/causalTWAS/simulations/simulation_susieI_20201001/chr22_plots/chr22snpVschr2causal", i)
+  #GWAA(geno, matrix(geno.c[,i], ncol =1), snpname = snpname, anno = anno, outname, family = gaussian, ncore = 5, nSplits = 10, compress = F)
+  rsq_max <- ld_max(geno.c.unscaled[,i, drop =F], geno.unscaled, stats = "R.squared")
+
+  cat("max LD: ", rsq_max)
+  m <- read.table(outname, header =T, comment.char = "")
+  rbPal <- colorRampPalette(c('red','blue'))
+  plot(anno[,2], -log10(m[m$X.CHROM ==22, ]$PVALUE),  main = paste0("chr", chrom.p, ":", geno.c.names[i]), xlab = "position", ylab = "-log10(p)", col =rbPal(11)[as.numeric(cut(maf,breaks = seq(0,0.5, 0.05)))], pch =20)
+  legend("topleft",title="MAF",legend=seq(0,0.5, 0.05),col =rbPal(11),pch=20, cex = 0.5)
+  grid()
+}
 ```

### 2020-10-20 analysis/simulation-susieI-ukbWG-gtex.adipose_visual.Rmd
small pvalues

---
title: "Test mr.ash2s on ukbiobank data, chr1-22, 40k"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message= F, warning = F)
```
Provide visualization for analysis [here]

```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20201001/"
```

```{r source file}
source("analysis/visual_twas_plots.R")
```

## Simulation- high power

```{r sim 1-4, out.width = "98%", out.height="95%"}
tag <- "20201001-1-4"
f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
```
```{r 1-4-17, out.width = "98%", out.height="95%"}
chrom <- "17"
cat("simulation run: ", tag, "-", tag2, "; chr: ", chrom)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, f$rsq, chrom = chrom)
```

### 2020-10-20 code/ld.R

@@ -17,13 +17,13 @@ ld_max <- function(x,  y, stats = "R.squared"){
 }
 
 .eqtl_geno <- function(gname){
-  x <- exprres$exprlist[[gname]]
-  if (is.null(dim(x$wgt))){
-    dim(x$wgt) <- c(1,2)
-    colnames(x$wgt) <- c("idx", "wgt")
+  x <- exprres$wgtlist[[gname]]
+  if (is.null(dim(x))){
+    dim(x) <- c(1,2)
+    colnames(x) <- c("idx", "wgt")
   }
-  eqtlgeno <- dat$G[ , x$wgt[abs(x$wgt[, "wgt"]) > 0, "idx"], drop = F]
-  colnames(eqtlgeno) <- paste(gname, dat$snp[x$wgt[abs(x$wgt[, "wgt"]) > 0, "idx"], 1], sep = ".")
+  eqtlgeno <- dat$G[ , x[abs(x[, "wgt"]) > 0, "idx"], drop = F]
+  colnames(eqtlgeno) <- paste(gname, dat$snp[x[abs(x[, "wgt"]) > 0, "idx"], 1], sep = ".")
   return(eqtlgeno)
 }
 

### 2020-10-23 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
filtered samples s40.22

--- 2020-10-20 analysis/simulation-susieI-ukbWG-gtex.adipose.Rmd
@@
 ---
-title: "Test susieI using whole genome, 4k ukb samples"
+title: "Test susieI using whole genome, s40.22 samples"
 output: html_document
@@
 ```{r parameters}
-n <- 40000 # number of samples
+n.ori <- 40000 # number of samples
+n <-  22542
 p <- 656321 # number of SNPs
@@
 
-The genotype data we used is from UKB biobank, randomly selecting n = `r as.character(n)` samples.
+The genotype data we used is from UKB biobank, randomly selecting `r as.character(n.ori)` samples.
+We then filtered samples based on relatedness, that ended up with n = `r as.character(n)` samples.
 We use SNP genotype data from chr 1 to chr 22 combined from UKB.
@@
 
-* TWAS p value of SNPs:
+*  p value of SNPs (GWAS):
 
@@
 grid()
-```
-
-# Exploration of unusual small p values
-Take chromosom 22 geno type and perform linear regression for against geno type for causal SNPs on other chromosomes.
-We see some very significant p values.
-Take chromosome 2 causal SNPs as an example.
-
-## normal distribution as phenotype
-```{r random, fig.height=4, fig.width=8}
-chrom <- 22
-load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom ,"_s40000.FBM.Rd"))
-load(paste0(simdatadir, "simu_s40000_GTEXadipose-B", chrom, "-cis-expr.Rd"))
-load(paste0(simdatadir, "simu_", simtag, "-pheno.Rd"))
-
-N<- nrow(phenores$Y)
-cat("variance of the phenotype: ", var(phenores$Y), "\n")
-
-snpname = dat$snp[,1]
-anno <- cbind(dat$chr, dat$pos, dat$pos)
-
-outname <- "/home/simingz/causalTWAS/simulations/simulation_susieI_20201001/chr22_plots/chr22snpRandom"
-# GWAA(dat$G, matrix(rnorm(N, sd = sqrt(var(phenores$Y))), ncol =1), snpname = snpname, anno = anno, outname , family = gaussian, ncore = 5, nSplits = 30, compress = F)
-n <- read.table(outname, header =T, comment.char = "")
-plot(n$BEGIN, -log10(n$PVALUE), main = "Random", xlab = "position", ylab = "-log10(p)")
-grid()
-```
-
-## genetic component of chromosome 2 as phenotype
-
-Not all chromosome give small p values. chr2 is the one with the smallest p values.
-```{r chrfixed, fig.height=4, fig.width=8}
-chrom.p <- 2
-r <- do.call(cbind, lapply(1:22, function(x) phenores$batch[[x]]$Y.g))
-cat("variance of the genetic component: ", var(r[,chrom.p]))
-
-outname <- "/home/simingz/causalTWAS/simulations/simulation_susieI_20201001/chr22_plots/chr22snpVschr2fixed"
-# GWAA(dat$G, matrix(r[,chrom.p], ncol =1), snpname = snpname, anno = anno, outname, family = gaussian, ncore = 5, nSplits = 30, compress = F)
-n <- read.table(outname, header =T, comment.char = "", stringsAsFactors = F)
-plot(n$BEGIN, -log10(n$PVALUE), main = "chr2 genetic", xlab = "position", ylab = "-log10(p)")
-grid()
-```
-
-## genotype of causal snps on chromosome 2 as phenotype
-
-genotype of causal snps on chromosome 2 as phenotype.
-There are `r length(phenores$batch[[2]]$param$idx.cSNP)` causal SNPs.
-Not all of them significantly associated with SNP genotype on chr22, but quite a few.
-below are examples (only a subset of SNPs on chr22 are selected to run regression analysis with chr2 causal SNP genotypes.)
-
-```{r geno, fig.height=4, fig.width=8, message=F}
-chrom <- 22
-load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom ,"_s40000.FBM.Rd"))
-geno <- dat$G[, match(n[n$PVALUE < 1e-40, "MARKER_ID"], dat$snp[,1])]
-load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom ,"_s40000.unscaled.FBM.Rd"))
-geno.unscaled <-  dat$G[, match(n[n$PVALUE < 1e-40, "MARKER_ID"], dat$snp[,1])]
-
-snpname = dat$snp[match(n[n$PVALUE < 1e-40, "MARKER_ID"], dat$snp[,1]),1]
-anno <- cbind(dat$chr, dat$pos, dat$pos)[match(n[n$PVALUE < 1e-40, "MARKER_ID"], dat$snp[,1]), ]
-
-load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom.p ,"_s40000.FBM.Rd"))
-geno.c <- dat$G[, phenores$batch[[chrom.p]]$param$idx.cSNP]
-geno.c.names <- dat$snp[phenores$batch[[chrom.p]]$param$idx.cSNP,]
-load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom.p ,"_s40000.unscaled.FBM.Rd"))
-geno.c.unscaled <- dat$G[, phenores$batch[[chrom.p]]$param$idx.cSNP]
-
-ac <- ac <- apply(round(geno.unscaled), 2, table)
-maf <- apply(cbind( (ac[1,] * 2  + ac[2,]) / N/2, 1- (ac[1,] * 2  + ac[2,]) /N/2), 1, min)
-
-geno.c.select <- c(1, 23, 45, 65)
-
-for (i in geno.c.select){
-  outname <- paste0("/home/simingz/causalTWAS/simulations/simulation_susieI_20201001/chr22_plots/chr22snpVschr2causal", i)
-  #GWAA(geno, matrix(geno.c[,i], ncol =1), snpname = snpname, anno = anno, outname, family = gaussian, ncore = 5, nSplits = 10, compress = F)
-  rsq_max <- ld_max(geno.c.unscaled[,i, drop =F], geno.unscaled, stats = "R.squared")
-
-  cat("max LD: ", rsq_max)
-  m <- read.table(outname, header =T, comment.char = "")
-  rbPal <- colorRampPalette(c('red','blue'))
-  plot(anno[,2], -log10(m[m$X.CHROM ==22, ]$PVALUE),  main = paste0("chr", chrom.p, ":", geno.c.names[i]), xlab = "position", ylab = "-log10(p)", col =rbPal(11)[as.numeric(cut(maf,breaks = seq(0,0.5, 0.05)))], pch =20)
-  legend("topleft",title="MAF",legend=seq(0,0.5, 0.05),col =rbPal(11),pch=20, cex = 0.5)
-  grid()
-}
 ```
@@
 
+## Regions: all (2)
+
+* Susie run parameters: `L=2`.
+initialize with `null_weight = 0` and update `null_weight` based on last iteration results.
+We initialize prior for genes and SNPs using true values.
+gene.pve ~ 0.1, snp.pve ~ 0.5.
+
+```{r param2, fig.height=6, fig.width=6}
+tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
+phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
+susieIfs <- paste0(outputdir, "20201001-", tags, ".config2.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config2.susieI.txt")
+
+df <- show_param(phenofs, susieIfs, susieIfs2)
+par(mfrow = c(2,2))
+plot_param(df[1:5,], main = "low power")
+plot_param(df[6:10,], main = "high power")
+```
+
 ## Regions: causal
@@
 ```
+
+## Regions: filtered (1)
+* We filterd regions with regional minimum FDR of both genes and SNPs < 0.05, round 700 regions left.
+* Susie run parameters: `L=2`.
+initialize with `null_weight = 0` and update `null_weight` based on last iteration results.
+We initialize with prior for genes and SNPs as uniform.
+gene.pve ~ 0.1, snp.pve ~ 0.5.
+
+```{r param3, fig.height=6, fig.width=6}
+tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
+phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
+susieIfs <- paste0(outputdir, "20201001-", tags, ".config3.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config3.susieI.txt")
+
+df <- show_param(phenofs, susieIfs, susieIfs2)
+par(mfrow = c(2,2))
+plot_param(df[1:5,], main = "low power")
+plot_param(df[6:10,], main = "high power")
+```
+
+## Regions: filtered (2)
+* We filterd regions with regional minimum FDR of both genes and SNPs < 0.01, round 400 regions left.
+* Susie run parameters: `L=2`.
+initialize with `null_weight = 0` and update `null_weight` based on last iteration results.
+We initialize with prior for genes and SNPs as uniform.
+gene.pve ~ 0.1, snp.pve ~ 0.5.
+
+```{r param4, fig.height=6, fig.width=6}
+tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
+phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
+susieIfs <- paste0(outputdir, "20201001-", tags, ".config4.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config4.susieI.txt")
+
+df <- show_param(phenofs, susieIfs, susieIfs2)
+par(mfrow = c(2,2))
+plot_param(df[1:5,], main = "low power")
+plot_param(df[6:10,], main = "high power")
+```

### 2020-10-23 code/run_UKB_process.R

@@ -141,6 +141,21 @@ for (chrom in 1:22){
 write.table(regions, file= "/home/simingz/causalTWAS/simulations/shared_files/chr1-22-500kb_bins.txt",
             row.names=F, col.names=T, sep="\t", quote = F)
 
+#-------------------------filter s40000 samples---------------------------------------
+s.s <- read.table("/home/simingz/causalTWAS/ukbiobank/ukbiobank_samples_s40.22.txt", header =F, stringsAsFactors = F)
+s.ori <- read.table("/home/simingz/causalTWAS/ukbiobank/ukbiobank_samples40000.txt", header =F, stringsAsFactors = F)
+
+s.s.idx <- match(s.s[,1], s.ori[,1])
+for (chrom in 1:22){
+  outf <- paste0("/home/simingz/causalTWAS/ukbiobank/ukb_chr", chrom, "_s40000.FBM.Rd")
+  load(outf)
+  G.new <- dat$G[]
+  G.new <- G.new[s.s.idx, ]
+  m <- as_FBM(G.new, backingfile = paste0("/home/simingz/causalTWAS/ukbiobank/ukb_chr", chrom, "_s40.22"), is_read_only = T)$save()
+  dat$G <- m
+  save(dat, file = paste0("/home/simingz/causalTWAS/ukbiobank/ukb_chr", chrom, "_s40.22.FBM.Rd"))
+}
+
 #---------chr17-22 multiple copies with different names---
 setwd("/home/simingz/causalTWAS/ukbiobank")
 for (i in 1:6) {

### 2020-10-23 code/run_simulate_pheno.R

@@ -38,7 +38,7 @@ if (file_ext(pfile) == "txt"){
   pfiles <- pfile
   efiles <- args[2]
 }
-pfileRds <- paste0(drop_ext(pfiles), ".FBM.Rd")
+pfileRds <- paste0(pfiles, ".FBM.Rd")
 
 
 # set sigma based on pve

### 2020-10-23 code/workflow/prep_ukbiobank_n40000_chr12_genotype.sh

@@ -1,0 +1,1 @@
+cd /home/szhao1/uk-biobank/uk-biobank-genotypes; /home/szhao1/uk-biobank/uk-biobank-phenotypes/tools/plink2 --sample /home/szhao1/uk-biobank/uk-biobank-phenotypes/data/genotypes/ukb27386_imp_v3_s487324.sample --keep /home/szhao1/causal-TWAS/data/ukbiobank_samples40000.txt --bgen ukb_imp_chr12_v3.bgen --chr 12 --threads 19 --make-pgen --out ~/causal-TWAS/data/ukb_chr12_s40000 --maf 0.05 --geno 0.05 --memory 30000

### 2020-10-23 code/workflow/prep_ukbiobank_n40000_chr17_genotype.sh

@@ -1,0 +1,1 @@
+cd /home/szhao1/uk-biobank/uk-biobank-genotypes; /home/szhao1/uk-biobank/uk-biobank-phenotypes/tools/plink2 --sample /home/szhao1/uk-biobank/uk-biobank-phenotypes/data/genotypes/ukb27386_imp_v3_s487324.sample --keep /home/szhao1/causal-TWAS/data/ukbiobank_samples40000.txt --bgen ukb_imp_chr17_v3.bgen --chr 17 --threads 19 --make-pgen --out ~/causal-TWAS/data/ukb_chr17_s40000 --maf 0.05 --geno 0.05 --memory 30000

### 2020-10-23 code/workflow/prep_ukbiobank_n40000_chr19_genotype.sh

@@ -1,0 +1,1 @@
+cd /home/szhao1/uk-biobank/uk-biobank-genotypes; /home/szhao1/uk-biobank/uk-biobank-phenotypes/tools/plink2 --sample /home/szhao1/uk-biobank/uk-biobank-phenotypes/data/genotypes/ukb27386_imp_v3_s487324.sample --keep /home/szhao1/causal-TWAS/data/ukbiobank_samples40000.txt --bgen ukb_imp_chr19_v3.bgen --chr 19 --threads 19 --make-pgen --out ~/causal-TWAS/data/ukb_chr19_s40000 --maf 0.05 --geno 0.05 --memory 30000

### 2020-10-23 code/workflow/prep_ukbiobank_n40000_chr3_genotype.sh

@@ -1,0 +1,1 @@
+cd /home/szhao1/uk-biobank/uk-biobank-genotypes; /home/szhao1/uk-biobank/uk-biobank-phenotypes/tools/plink2 --sample /home/szhao1/uk-biobank/uk-biobank-phenotypes/data/genotypes/ukb27386_imp_v3_s487324.sample --keep /home/szhao1/causal-TWAS/data/ukbiobank_samples40000.txt --bgen ukb_imp_chr3_v3.bgen --chr 3 --threads 19 --make-pgen --out ~/causal-TWAS/data/ukb_chr3_s40000 --maf 0.05 --geno 0.05 --memory 30000

### 2020-10-23 code/workflow/prep_ukbiobank_n40000_chr6_genotype.sh

@@ -1,0 +1,1 @@
+cd /home/szhao1/uk-biobank/uk-biobank-genotypes; /home/szhao1/uk-biobank/uk-biobank-phenotypes/tools/plink2 --sample /home/szhao1/uk-biobank/uk-biobank-phenotypes/data/genotypes/ukb27386_imp_v3_s487324.sample --keep /home/szhao1/causal-TWAS/data/ukbiobank_samples40000.txt --bgen ukb_imp_chr6_v3.bgen --chr 6 --threads 19 --make-pgen --out ~/causal-TWAS/data/ukb_chr6_s40000 --maf 0.05 --geno 0.05 --memory 30000

### 2020-10-23 code/workflow/prep_ukbiobank_n40000_chr8_genotype.sh

@@ -1,0 +1,1 @@
+cd /home/szhao1/uk-biobank/uk-biobank-genotypes; /home/szhao1/uk-biobank/uk-biobank-phenotypes/tools/plink2 --sample /home/szhao1/uk-biobank/uk-biobank-phenotypes/data/genotypes/ukb27386_imp_v3_s487324.sample --keep /home/szhao1/causal-TWAS/data/ukbiobank_samples40000.txt --bgen ukb_imp_chr8_v3.bgen --chr 8 --threads 19 --make-pgen --out ~/causal-TWAS/data/ukb_chr8_s40000 --maf 0.05 --geno 0.05 --memory 30000

### 2020-10-23 code/workflow/qsub_prep_ukbiobank_n40000_chr1_genotype.sh

@@ -1,0 +1,10 @@
+
+#!/bin/bash
+
+#PBS -N bgen2plink
+#PBS -S /bin/bash
+#PBS -l mem=64gb
+#PBS -l walltime=48:00:00
+#PBS -l nodes=1:ppn=20
+
+cd /home/szhao1/causal-TWAS/data; bash prep_ukbiobank_n40000_chr1_genotype.sh

### 2020-10-23 code/workflow/qsub_prep_ukbiobank_n40000_chr2_genotype.sh

@@ -1,0 +1,10 @@
+
+#!/bin/bash
+
+#PBS -N bgen2plink
+#PBS -S /bin/bash
+#PBS -l mem=64gb
+#PBS -l walltime=48:00:00
+#PBS -l nodes=1:ppn=20
+
+cd /home/szhao1/causal-TWAS/data; bash prep_ukbiobank_n40000_chr2_genotype.sh

### 2020-10-23 code/workflow/qsub_prep_ukbiobank_n40000_chr5_genotype.sh

@@ -1,0 +1,10 @@
+
+#!/bin/bash
+
+#PBS -N bgen2plink
+#PBS -S /bin/bash
+#PBS -l mem=64gb
+#PBS -l walltime=48:00:00
+#PBS -l nodes=1:ppn=20
+
+cd /home/szhao1/causal-TWAS/data; bash prep_ukbiobank_n40000_chr5_genotype.sh

### 2020-10-23 code/workflow/qsub_prep_ukbiobank_n40000_chr6_genotype.sh

@@ -1,0 +1,10 @@
+
+#!/bin/bash
+
+#PBS -N bgen2plink
+#PBS -S /bin/bash
+#PBS -l mem=64gb
+#PBS -l walltime=48:00:00
+#PBS -l nodes=1:ppn=20
+
+cd /home/szhao1/causal-TWAS/data; bash prep_ukbiobank_n40000_chr6_genotype.sh

### 2020-10-23 code/workflow/qsub_prep_ukbiobank_n40000_chr7_genotype.sh

@@ -1,0 +1,10 @@
+
+#!/bin/bash
+
+#PBS -N bgen2plink
+#PBS -S /bin/bash
+#PBS -l mem=64gb
+#PBS -l walltime=48:00:00
+#PBS -l nodes=1:ppn=20
+
+cd /home/szhao1/causal-TWAS/data; bash prep_ukbiobank_n40000_chr7_genotype.sh

### 2020-10-24 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
qqplot

@@
 The genotype data we used is from UKB biobank, randomly selecting `r as.character(n.ori)` samples.
-We then filtered samples based on relatedness, that ended up with n = `r as.character(n)` samples.
+We then filtered samples based on relatedness, ethics and other qc metrics, that ended up with n = `r as.character(n)` samples.
 We use SNP genotype data from chr 1 to chr 22 combined from UKB.
@@
 source('~/causalTWAS/causal-TWAS/code/ld.R')
+source('~/causalTWAS/causal-TWAS/code/qqplot.R')
 ```
@@
 for (chrom in 1:22) {
-  load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom ,"_s40000.FBM.Rd"))
+  load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom ,"_s40.22.FBM.Rd"))
   load(paste0(simdatadir, "simu_s40000_GTEXadipose-B", chrom, "-cis-expr.Rd"))
@@
 cau <- unlist(caulist)
+```
+# Power estimation
+
+We use gene.pve ~ 0.1, snp.pve ~ 0.5.
+
+* For SNPs, we use $\pi_1 = 2.5e-3$, variance for effect size ~ $0.03^2$, power at 5e-8 p value cutoff:
+```{r pow func} 
+pow <- function(total, n, beta, cutp){
+  rec <- rep(0, total)
+  for (i in 1:total){
+    x <- rnorm(n)
+    y <- x * rnorm(1, sd = beta) + rnorm(n, sd = sqrt(2.5))
+    lm.s <- lm(y~x)
+    pv <- summary(lm.s)$coefficients[2,4]
+    rec[i] <- pv
+  }
+  length(rec[rec < cutp])/length(rec)
+}
+```
+
+```{r snp}
+total <- 1e3
+n <- 22542
+p1 <- pow(total, n, 0.0276, 5e-8)
+print(p1)
+```
+
+* For genes, under low power setting, $\pi_1 = 0.05$, variance for effect size ~ $0.025^2$, power at 1e-5 cutoff:
+
+```{r gene1}
+p2 <- pow(total, n, 0.025, 1e-5)
+print(p2)
+```
+
+For genes, under high power setting, $\pi_1 = 0.02$, variance for effect size ~ $0.045^2$, power at 1e-5 cutoff:
+
+```{r gene2}
+p3 <- pow(total, n, 0.045, 1e-5)
+print(p3)
 ```
@@
 ```{r gene twas, fig.height=8, fig.width=8}
-chrom <- 22
+chrom <- 1
 a <- read.table(exprgwasf, header = T)
@@
 par(mfrow=c(3,1))
-h1 <- hist(-log10(a$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey"); grid()
-h2 <- hist(-log10(a[a$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon");grid()
+h1 <- hist(-log10(a$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey", xlim= c(3,20), ylim =c(0,50)); grid()
+h2 <- hist(-log10(a[a$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon", xlim= c(3,20), ylim =c(0,50));grid()
+
+cat("number of genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5,]))
 
@@
 grid()
+```
+
+* qq plot for genes
+```{r qqgene, fig.height= 5, fig.width=5}
+gg_qqplot(a$PVALUE) +
+  theme_bw(base_size = 24) +
+  theme(
+    axis.ticks = element_line(size = 0.5),
+    panel.grid = element_blank()
+    # panel.grid = element_line(size = 0.5, color = "grey80")
+  )
 ```
@@
 par(mfrow=c(3,1))
-h1 <- hist(-log10(b$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey"); grid()
-h2 <- hist(-log10(b[b$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon");grid()
+h1 <- hist(-log10(b$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey", xlim= c(3,20), ylim =c(0,100)); grid()
+h2 <- hist(-log10(b[b$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon", xlim= c(3,20), ylim =c(0,100));grid()
+
+cat("number of SNPs < 5e-8: ", nrow(b[b$PVALUE < 5e-8,]))
 
@@
 grid()
+```
+
+* qq plot for SNPs
+```{r qqsnp, fig.height= 5, fig.width=5}
+gg_qqplot(b$PVALUE) +
+  theme_bw(base_size = 24) +
+  theme(
+    axis.ticks = element_line(size = 0.5),
+    panel.grid = element_blank()
+    # panel.grid = element_line(size = 0.5, color = "grey80")
+  )
 ```
@@
 ```{r param1, fig.height=6, fig.width=6}
+outputdir <- "~/causalTWAS/simulations/simulation_susieI_20201001/L=2/"
 tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")

### 2020-10-24 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
qqplot

@@
 print(p3)
+# save(p1,p2,p3, file = "data/power_s40.22.Rd")
 ```
@@
 * Susie run parameters: `L=2`.
-initialize with `null_weight = 0` and update `null_weight` based on last iteration results.
 We initialize with prior for genes and SNPs as uniform.
 gene.pve ~ 0.1, snp.pve ~ 0.5.
+Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
 
@@
 * Susie run parameters: `L=2`.
-initialize with `null_weight = 0` and update `null_weight` based on last iteration results.
-We initialize prior for genes and SNPs using true values.
+We initialize with prior for genes and SNPs using their true values.
 gene.pve ~ 0.1, snp.pve ~ 0.5.
+Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
 

### 2020-10-30 analysis/index.Rmd
wg susieI
small pvalues
filtered samples s40.22
log bayes factor

@@
 
-## mr.ash + SUSIE iterative
+## SUSIE iterative
 *  [Using UKbiobank data (chr17:22), fusion weights Adipose](simulation-susieI-ukbchr17to22-gtex.adipose.html) 
   *  [Using UKbiobank data (chr17:22), fusion weights Adipose, test prior's effect for susie](simulation-multi-ukbchr17to22-gtex.adipose_susieprior.html)
+
+* [Using UKbiobank data (whole genome), fusion weights Adipose](simulation-susieI-ukbWG-gtex.adipose_s40.22.html) 
+  * [Visualization](simulation-susieI-ukbWG-gtex.adipose_visual.html)
 

### 2020-10-30 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
qqplot
log bayes factor

@@
 library(stringr)
-library(kableExtra)
 source("analysis/summarize_twas_plots.R")
@@
 cat("number of genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5,]))
+cat("number of causal genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5 & a$ifcausal ==1 ,]))
 
@@
 
-* Susie run parameters: `L=2`.
+* Susie run parameters: `L=1`.
 We initialize with prior for genes and SNPs as uniform.
@@
 ```{r param1, fig.height=6, fig.width=6}
-outputdir <- "~/causalTWAS/simulations/simulation_susieI_20201001/L=2/"
 tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
 phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")
+susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.SERIres.Rd")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.SERI.txt")
 
@@
 
-* Susie run parameters: `L=2`.
+* Susie run parameters: `L=1`.
 We initialize with prior for genes and SNPs using their true values.
@@
 
-* Susie run parameters: `L=2`.
+* Susie run parameters: `L=1`.
 initialize with `null_weight = 0` and update `null_weight` based on last iteration results.
@@
 * We filterd regions with regional minimum FDR of both genes and SNPs < 0.05, round 700 regions left.
-* Susie run parameters: `L=2`.
+* Susie run parameters: `L=1`.
 initialize with `null_weight = 0` and update `null_weight` based on last iteration results.
@@
 * We filterd regions with regional minimum FDR of both genes and SNPs < 0.01, round 400 regions left.
-* Susie run parameters: `L=2`.
+* Susie run parameters: `L=1`.
 initialize with `null_weight = 0` and update `null_weight` based on last iteration results.
@@
 ```
+
+# Plot PIP
+
+Red is causal.
+
+```{r gene twas, fig.height=4, fig.width=10}
+chrom <- 1
+tag <- "20201001-2-1"
+pipfile <-  paste0(outputdir, tag, ".config2.1.SERI.txt")
+a <- fread(pipfile, header = T)
+
+plot(a[a$chr ==chrom, ]$pos, a[a$chr ==chrom, ]$SERpip, col = a[a$chr ==chrom, ]$ifcausal + 1, xlab = paste0("chr", chrom), ylab = "pip", pch = c(1,2)[as.factor(a[a$chr ==chrom, ]$type)])
+points(a[a$chr ==chrom & a$ifcausal ==1, ]$pos, a[a$chr ==chrom & a$ifcausal ==1, ]$SERpip , col = "red", pch = c(19, 17)[as.factor(a[a$chr ==chrom & a$ifcausal ==1 , ]$type)])
+legend("topright",c("gene","SNP"),cex=.8,pch=c(1,2))
+grid()
+```

### 2020-10-30 analysis/simulation-susieI-ukbWG-gtex.adipose_visual.Rmd
log bayes factor

@@
 ---
-title: "Test mr.ash2s on ukbiobank data, chr1-22, 40k"
+title: "Visualize ukb whole genome, s40.22"
 date: "`r Sys.Date()`"
@@
 simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
-outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
-susiedir <- "~/causalTWAS/simulations/simulation_susietest_20201001/"
+outputdir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
+
+source("analysis/visual_twas_plots.R")
 ```
 
-```{r source file}
-source("analysis/visual_twas_plots.R")
+```{r functions file}
+
+plot_all <- function(SERf, phenof, chrom, ldres = NULL){
+  a <- fread(SERf , header =T)
+  load(phenof)
+
+  a <- a[a$chr == chrom,]
+
+  if (!is.null(ldres)){
+    ld <- read.table(ldres, header =T)
+    a <- merge(a, ld, by = "name", all.x = T)}
+
+  a <- a[, ifcausal:=as.character(ifcausal)]
+  a[a$ifcausal == "1", "ifcausal"] = "Causal"
+  a[a$ifcausal == "0", "ifcausal"] = "Non causal"
+
+  colnames(a)[colnames(a) == "SERpip"] <-  "susiePIP"
+  colnames(a)[colnames(a) == "pos"] <-  "p0"
+
+  a[, "PVALUE"] <- -log10(a[, "pvalue"])
+
+  dt.gene <- a[a$type == "gene", ]
+  dt.snp <- a[a$type == "SNP", ]
+
+  fig1 <- plot_susie_pip(dt.gene) 
+  #fig2 <- plot_beta(dt.gene)
+  fig3 <- plot_p(dt.gene)
+
+  fig4 <- plot_susie_pip(dt.snp)
+  #fig5 <- plot_beta(dt.snp) 
+  fig6 <- plot_p(dt.snp) %>% layout( plot_bgcolor='#EFF6FC')
+
+  # fig <- subplot(fig1, fig2, fig3, fig4, fig5, fig6, nrows=6, shareX = T, titleY = T)
+
+  fig <- subplot(fig1, fig3, fig4, fig6, nrows=4, shareX = T, titleY = T)
+
+  fig <- fig %>% layout(autosize = F, width = 600, height = 600, margin = 0.02,
+                        xaxis = list(title = "Genomic position"))
+
+  fig <- fig %>% toWebGL()
+  fig
+}
 ```
@@
 
-```{r sim 1-4, out.width = "98%", out.height="95%"}
-tag <- "20201001-1-4"
-f <- get_files(tag= tag, tag2 = tag2)
-param <- read.table(f$par, header = T, row.names = 1)
-print(param)
+first two rows are for genes, last two rows are for SNPs.
+
+```{r sim 2-1-1, out.width = "98%", out.height="95%"}
+chrom <- 1
+tag <- "20201001-2-1"
+pipfile <-  paste0(outputdir, tag, ".config2.1.SERI.txt")
+phenof <- paste0(simdatadir, "simu_", tag, "-pheno.Rd")
+plot_all(pipfile, phenof, chrom)
 ```
-```{r 1-4-17, out.width = "98%", out.height="95%"}
-chrom <- "17"
-cat("simulation run: ", tag, "-", tag2, "; chr: ", chrom)
-plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, f$rsq, chrom = chrom)
-```

### 2020-10-30 code/SER.R

@@ -1,0 +1,98 @@
+#' @title Bayesian single-effect linear regression of y on X,
+#' adapted from susieR (single_effect_regression.R)
+#'
+#' @param Y An n vector.
+#'
+#' @param X An n by p matrix of covariates. centered and scaled.
+#'
+#' @param V The prior variance.
+#'
+#' @param residual_variance The residual variance.
+#'
+#' @param prior_weights A p vector of prior weights.
+#'
+#' @return a list
+#'
+#' @importFrom Matrix colSums
+#'
+#'
+get_lbf =
+  function (Y, X, V, residual_variance = 1) {
+
+    ytX = crossprod(Y, X)
+    Xty = t(ytX)
+    xTx = colSums(X * X)
+    betahat = (1/xTx) * Xty
+    shat2 = residual_variance/xTx
+
+
+    # log(bf) for each SNP
+    lbf = dnorm(betahat,0,sqrt(V + shat2),log = TRUE) -
+      dnorm(betahat,0,sqrt(shat2),log = TRUE)
+
+    # Deal with special case of infinite shat2 (e.g., happens if X does
+    # not vary).
+    lbf[is.infinite(shat2)] = 0
+
+    z = betahat/sqrt(shat2)
+    z[is.infinite(z)] = 0
+
+    pvalue = 2*pnorm(-abs(z))
+
+    lbfout <- cbind(lbf, betahat, shat2, z, pvalue)
+    colnames(lbfout) <- c("lbf", "betahat", "shat2", "z", "pvalue")
+
+    return(lbfout)
+  }
+
+
+
+#' @title Bayesian single-effect linear regression of y on X,
+#' adapted from susieR (single_effect_regression.R)
+#'
+#' @param lbf a vector
+#'
+#' @return A list with the following elements:
+#'
+#' \item{alpha}{Vector of posterior inclusion probabilities;
+#'   \code{alpha[i]} is posterior probability that that the ith
+#'   coefficient is non-zero.}
+#'
+#' \item{lbf_model}{Log-Bayes factor for the single effect regression.}
+#'
+#' \item{loglik}{The logarithm of the likelihood, \eqn{p(y | X, V)}.}
+#'
+#' @importFrom Matrix colSums
+#'
+#'
+SER =
+  function (lbf, prior_weights = NULL, null_weight = NULL) {
+    if (is.null(prior_weights))
+      prior_weights = c(rep(1/length(lbf)*(1-null_weight), length(lbf)))
+
[…]

### 2020-10-30 code/run_test_susieI.R

@@ -155,6 +155,7 @@ for (b in 1:length(pfiles)){
 
 write.table(regionsall, file= paste0(outname,".", filtertype, ".r.txt" ) , row.names=F, col.names=T, sep="\t", quote = F)
 
+save.image("tempsusie.Rd")
 # start susieI
 loginfo("susie started for %s", outname)
 
@@ -209,7 +210,7 @@ for (iter in 1:Niter){
 
 
       if (isTRUE(ifnullweight)){
-        susieres <- susie( X , phenores$Y, L = L, prior_weights = prior, null_weight = nw)
+        susieres <- susie(X , phenores$Y, L = L, prior_weights = prior/(1-nw), null_weight = nw)
       } else {
         susieres <- susie(X, phenores$Y, L = L, prior_weights = prior)
       }

### 2020-10-31 analysis/simulation-susieI-ukbWG-gtex.adipose_visual.Rmd
logBF

@@
   a[, "PVALUE"] <- -log10(a[, "pvalue"])
+  a[, "lbf"] <- log10(exp(a[, "lbf"]))
 
@@
   fig1 <- plot_susie_pip(dt.gene) 
-  #fig2 <- plot_beta(dt.gene)
+  fig2 <- plot_lbf(dt.gene)
   fig3 <- plot_p(dt.gene)
@@
   fig4 <- plot_susie_pip(dt.snp)
-  #fig5 <- plot_beta(dt.snp) 
+  fig5 <- plot_lbf(dt.snp) 
   fig6 <- plot_p(dt.snp) %>% layout( plot_bgcolor='#EFF6FC')
 
-  # fig <- subplot(fig1, fig2, fig3, fig4, fig5, fig6, nrows=6, shareX = T, titleY = T)
+  fig <- subplot(fig1, fig2, fig3, fig4, fig5, fig6, nrows=6, shareX = T, titleY = T)
 
-  fig <- subplot(fig1, fig3, fig4, fig6, nrows=4, shareX = T, titleY = T)
-
-  fig <- fig %>% layout(autosize = F, width = 600, height = 600, margin = 0.02,
+  fig <- fig %>% layout(autosize = F, width = 600, height = 800, margin = 0.02,
                         xaxis = list(title = "Genomic position"))
@@
 
-first two rows are for genes, last two rows are for SNPs.
+First three rows are for genes, last three rows are for SNPs.
 

### 2020-11-04 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
SERI

@@
 Next, the analysis is done at the “region” level, which is 500kb bins along the genome.
-Each bin would contain all the SNPs, as well as all the genes in that bin.
+We also used LDetect to define regions.
 We are exploring several ways to select regions that contain true signals, e.g.
@@
 
-## Regions: all (1)
-
+## susieI
+* Regions: all regions, 500kb uniform regions.
 * Susie run parameters: `L=1`.
@@
 Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
+Prior variance and residual variance were calculated by SUSIE for each region.
+
+```{r param1susie, fig.height=6, fig.width=6}
+tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
+phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
+susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")
+
+df <- show_param(phenofs, susieIfs, susieIfs2)
+par(mfrow = c(2,2))
+plot_param(df[1:5,], main = "low power")
+plot_param(df[6:10,], main = "high power")
+```
+
+##  Single effect model/EM (1)
+* Regions: all regions, 500kb uniform regions.
+* Run parameters: We initialize with prior for genes and SNPs as uniform.
+gene.pve ~ 0.1, snp.pve ~ 0.5.
+Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
+We use the true prior variance for genes and SNPs.
 
@@
 
-## Regions: all (2)
-
-* Susie run parameters: `L=1`.
-We initialize with prior for genes and SNPs using their true values.
+##  Single effect model/EM (2)
+* Regions: all regions, defined based on LDetect.
+* Run parameters: We initialize with prior for genes and SNPs as uniform.
 gene.pve ~ 0.1, snp.pve ~ 0.5.
 Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
+We use the true prior variance for genes and SNPs.
 
-```{r param2, fig.height=6, fig.width=6}
+```{r param6, fig.height=6, fig.width=6}
 tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
 phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config2.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config2.susieI.txt")
+susieIfs <- paste0(outputdir, "20201001-", tags, ".config6.SERIres.Rd")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config6.SERI.txt")
 
@@
 
-## Regions: causal
+##  Single effect model/EM (3)
+* Regions: regions with at most 1 causal signal, regions are defined based on LDetect.
+* Run parameters: We initialize with prior for genes and SNPs as uniform.
+gene.pve ~ 0.1, snp.pve ~ 0.5.
+Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
+We use the true prior variance for genes and SNPs.
 
-* Susie run parameters: `L=1`.
-initialize with `null_weight = 0` and update `null_weight` based on last iteration results.
-We initialize with prior for genes and SNPs as uniform.
-gene.pve ~ 0.1, snp.pve ~ 0.5.
-
-```{r param5, fig.height=6, fig.width=6}
+```{r param7, fig.height=6, fig.width=6}
 tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
 phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config5.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config5.susieI.txt")
+susieIfs <- paste0(outputdir, "20201001-", tags, ".config7.SERIres.Rd")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config7.SERI.txt")
 
@@
 ```
-
-## Regions: filtered (1)
-* We filterd regions with regional minimum FDR of both genes and SNPs < 0.05, round 700 regions left.
-* Susie run parameters: `L=1`.
-initialize with `null_weight = 0` and update `null_weight` based on last iteration results.
-We initialize with prior for genes and SNPs as uniform.
-gene.pve ~ 0.1, snp.pve ~ 0.5.
-
-```{r param3, fig.height=6, fig.width=6}
-tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
-phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config3.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config3.susieI.txt")
-
-df <- show_param(phenofs, susieIfs, susieIfs2)
-par(mfrow = c(2,2))
-plot_param(df[1:5,], main = "low power")
-plot_param(df[6:10,], main = "high power")
-```
-
-## Regions: filtered (2)
-* We filterd regions with regional minimum FDR of both genes and SNPs < 0.01, round 400 regions left.
-* Susie run parameters: `L=1`.
-initialize with `null_weight = 0` and update `null_weight` based on last iteration results.
-We initialize with prior for genes and SNPs as uniform.
-gene.pve ~ 0.1, snp.pve ~ 0.5.
-
-```{r param4, fig.height=6, fig.width=6}
-tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
-phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config4.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config4.susieI.txt")
-
-df <- show_param(phenofs, susieIfs, susieIfs2)
-par(mfrow = c(2,2))
-plot_param(df[1:5,], main = "low power")
-plot_param(df[6:10,], main = "high power")
-```
-
-# Plot PIP
-
-Red is causal.
-
-```{r gene twas, fig.height=4, fig.width=10}
-chrom <- 1
-tag <- "20201001-2-1"
-pipfile <-  paste0(outputdir, tag, ".config2.1.SERI.txt")
-a <- fread(pipfile, header = T)
-
-plot(a[a$chr ==chrom, ]$pos, a[a$chr ==chrom, ]$SERpip, col = a[a$chr ==chrom, ]$ifcausal + 1, xlab = paste0("chr", chrom), ylab = "pip", pch = c(1,2)[as.factor(a[a$chr ==chrom, ]$type)])
-points(a[a$chr ==chrom & a$ifcausal ==1, ]$pos, a[a$chr ==chrom & a$ifcausal ==1, ]$SERpip , col = "red", pch = c(19, 17)[as.factor(a[a$chr ==chrom & a$ifcausal ==1 , ]$type)])
-legend("topright",c("gene","SNP"),cex=.8,pch=c(1,2))
-grid()
-```

### 2020-11-04 analysis/simulation-susieI-ukbWG-gtex.adipose_visual.Rmd
SERI

@@
 tag <- "20201001-2-1"
-pipfile <-  paste0(outputdir, tag, ".config2.1.SERI.txt")
+pipfile <-  paste0(outputdir, tag, ".config.1.SERI.txt")
 phenof <- paste0(simdatadir, "simu_", tag, "-pheno.Rd")

### 2020-11-04 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
SER

@@
 
+  st <-  cbind(df[,"pi1.SNP_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
+  s <- cbind(df[,"pi1.SNP_susieI"], 1:nrow(df), 3 + 1:nrow(df)/nrow(df)/3)
+  t <- df[1,"pi1.SNP_truth"]
+  dfp <- rbind(st,s)
+  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "SNP pi1", xaxt = "n", xlab="", xlim = c(0.8, 3.5), frame.plot=FALSE, ylim = c(0, max(dfp[,1],t) *1.05), ...)
+  axis(side=1, at=1:2, labels = FALSE, tick = F)
+  text(x=2:3, 0, labels = c( "susieI_truth", "susieI"), xpd = T, pos =1)
+  abline(h=t, lty = 2, col= "salmon", lwd=1.5)
+  grid()
+
   st <-  cbind(df[,"pi1.gene_susietruth"]/df[,"pi1.SNP_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
@@
 
-```{r param1susie, fig.height=6, fig.width=6}
+```{r param1susie, fig.height=6, fig.width=8}
 tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
@@
 df <- show_param(phenofs, susieIfs, susieIfs2)
-par(mfrow = c(2,2))
+par(mfrow = c(2,3))
 plot_param(df[1:5,], main = "low power")
@@
 
-```{r param1, fig.height=6, fig.width=6}
+```{r param1, fig.height=6, fig.width=8}
 tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
@@
 
-```{r param6, fig.height=6, fig.width=6}
+```{r param6, fig.height=6, fig.width=8}
 tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
@@
 
-```{r param7, fig.height=6, fig.width=6}
+```{r param7, fig.height=6, fig.width=8}
 tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")

### 2020-11-13 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
pip cali

@@
 ```{r snp}
+load("data/power_s40.22.Rd")
 total <- 1e3
 n <- 22542
-p1 <- pow(total, n, 0.0276, 5e-8)
+#p1 <- pow(total, n, 0.0276, 5e-8)
 print(p1)
@@
 ```{r gene1}
-p2 <- pow(total, n, 0.025, 1e-5)
+#p2 <- pow(total, n, 0.025, 1e-5)
 print(p2)
@@
 ```{r gene2}
-p3 <- pow(total, n, 0.045, 1e-5)
+#p3 <- pow(total, n, 0.045, 1e-5)
 print(p3)
-# save(p1,p2,p3, file = "data/power_s40.22.Rd")
+#save(p1,p2,p3, file = "data/power_s40.22.Rd")
 ```
@@
 
-cat("number of genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5,]))
-cat("number of causal genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5 & a$ifcausal ==1 ,]))
+cat("number of genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5, ]))
+cat("number of causal genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5 & a$ifcausal ==1, ]))
 
@@
   )
+```
+
+# TWAS FDP
+
+```{r PIPplot functions}
+
+library(ggplot2)
+library(cowplot)
+library(plotrix)
+
+.exob <- function(pips, ifcausal, mode = c("PIP", "FDR")){
+  a_bin <- cut(pips, breaks= seq(0, 1, by=0.1))
+  if (mode == "PIP") {
+    ex = c(by(pips, a_bin, FUN = mean))
+    ob = c(by(ifcausal, a_bin, FUN = mean))
+  } else if (mode == "FDR"){
+    ex = c(by(pips, a_bin, FUN = mean))
+    ob = 1 - c(by(ifcausal, a_bin, FUN = mean))
+  }
+  return(list("expected" = ex, "observed" = ob))
+}
+
+dot_plot = function(df, main) {
+        ggplot(df, aes(x=mean_pip, y=observed_freq)) + 
+          geom_errorbar(aes(ymin=observed_freq-se, ymax=observed_freq+se), colour="black", size = 0.5, width=.01) +
+          geom_point(size=1.5, shape=21, fill="#002b36") + # 21 is filled circle
+          xlab("Expected") +
+          ylab("Observed") +
+          coord_cartesian(ylim=c(0,1), xlim=c(0,1)) +
+          geom_abline(slope=1,intercept=0,colour='red', size=0.2) +
+          ggtitle(main) +
+          expand_limits(y=0) +                        # Expand y range
+          theme_cowplot()}
+
+cp_plot <- function(pips, ifcausal, runtag = NULL, mode = c("PIP", "FDR"), main = mode[1], ...){
+
+  # ifcausal:0,1, runtag: for adding std.
+  if (is.null(runtag)){
+     se = 0 
+  } else{
+    dflist <- list()
+    for (rt in unique(runtag)){
+      pips.rt <- pips[runtag == rt]
+      ifcausal.rt <- ifcausal[runtag == rt]
+      dflist[[rt]] <- .exob(pips.rt, ifcausal.rt, mode = mode)
+    }
+    mean_pip <- colMeans(do.call(rbind, lapply(dflist, '[[', "expected")))
+    observed_freq <- colMeans(do.call(rbind, lapply(dflist, '[[', "observed")))
+    se <- apply(do.call(rbind, lapply(dflist, '[[', "observed")), 2, plotrix::std.error)
+  }
+
+  df <- data.frame("mean_pip" = mean_pip, "observed_freq"= observed_freq, "se" = se)
+  dot_plot(df, main = main)
+
+  #plot(Expected, Observed, xlim= c(0,1), ylim=c(0,1), pch =19, main = main, ...)
+  #lines(x = c(0,1), y = c(0,1), col ="grey", lty = 2)
+}
+
+caliPIP_plot <- function(pipfs, format = "susie", main = "PIP"){
+  df <- NULL
+  for (i in 1:length(pipfs)) {
+    res <- fread(pipfs[i], header = T)
+    res <- data.frame(res[res$type  =="gene", ])
+    res$runtag <- i
+    if (format == "susie"){
+       res <- rename(res, c("susie_pip" = "pip"))
+    } else if (format == "SER"){
+       res <- rename(res, c("SERpip" = "pip"))
+    }
+    res <- res[complete.cases(res),]
+    df <- rbind(df, res)
+  }
+
+  fig <- cp_plot(df$pip, df$ifcausal, df$runtag, mode ="PIP", main = main)
+  return(fig)
+}
+
+# pipfs is just for ifcausal info
+caliFDP_plot <- function(gwasfs, pipfs,  format= "susie",  main = "FDP"){
+  df <- NULL
+  for (i in 1:length(pipfs)) {
+    pipres <- fread(pipfs[i], header = T)
+    pipres <- data.frame(pipres[pipres$type  =="gene", ])
+    pipres$runtag <- i
+    if (format == "susie"){
+       pipres <- rename(pipres, c("susie_pip" = "pip"))
+    } else if (format == "SER"){
+       pipres <- rename(pipres, c("SERpip" = "pip"))
+    }
+    gwasres <- read.table(gwasfs[i], header = T)
+    gwasres <- rename(gwasres, c("MARKER_ID" = "name"))
+    gwasres$FDR <- p.adjust(gwasres$PVALUE, method = "fdr")
+    res <- merge(gwasres, pipres, by = "name", all = T)
+    res <- res[complete.cases(res),]
+    df <- rbind(df, res)
+  }
+  fig <- cp_plot(df$FDR, df$ifcausal, df$runtag, mode ="FDR", main = main)
+  cat("FDP at bonferroni corrected p = 0.05: ", 1 - mean(df[df$PVALUE < 0.05 /dim(df)[1], "ifcausal"]))
+  return(fig)
+}
+
+scatter_plot_PIP_p <- function(pipfs, gwasfs, pipformat = "susie", main ="PIP-p"){
+  df <- NULL
+  for (i in 1:length(pipfs)) {
+    pipres <- fread(pipfs[i], header = T)
+    pipres <- data.frame(pipres[pipres$type  =="gene", ])
+    pipres$runtag <- i
+    if (pipformat == "susie"){
+       pipres <- rename(pipres, c("susie_pip" = "pip"))
+    } else if (pipformat == "SER"){
+       pipres <- rename(pipres, c("SERpip" = "pip"))
+    }
+    gwasres <- read.table(gwasfs[i], header = T)
+    gwasres <- rename(gwasres, c("MARKER_ID" = "name"))
+    res <- merge(gwasres, pipres, by = "name", all = T)
+    res <- res[complete.cases(res),]
+    df <- rbind(df, res)
+  }
+
+  df <- rename(df, c( "PVALUE" = "TWAS.p"))
+  df[,"TWAS.p"] <- -log10(df[, "TWAS.p"])
+
+  # df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1), to=c("Non causal", "Causal"))
+
+  df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1),to=c("darkgreen", "salmon"))
+
+  plot(df$TWAS.p, df$pip, col = df$ifcausal, main = main, xlab = "-log10(TWAS p value)", ylab = "PIP")
+  # fig <- plot_ly(data = df, x = ~ TWAS.p, y = ~ pip, color = ~ ifcausal,
+  #                colors = c( "salmon", "darkgreen"), type ="scatter", text = ~ paste("Name: ", paste0(runtag,":",name),
+  #                                                                  "\nChr: ", chr,  "\nPos:", pos))
+
+}
+
+```
+
+```{r TWAS FDP, fig.height=4, fig.width=8}
+tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
+gwasfs <-  paste0(simdatadir, "20201001-", tags, ".exprgwas.txt.gz")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")
+
+f1 <- caliFDP_plot(gwasfs[1:5], susieIfs2[1:5], format= "susie", main = "TWAS FDP (low power)")
+f2 <- caliFDP_plot(gwasfs[6:10], susieIfs2[6:10], format= "susie", main ="TWAS FDP (high power)")
+
+gridExtra::grid.arrange(f1, f2, ncol =2)
 ```
@@
 
-## susieI
+## susieI (1)
 * Regions: all regions, 500kb uniform regions.
@@
 phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
+gwasfs <-  paste0(simdatadir, "20201001-", tags, ".exprgwas.txt.gz")
 susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.susieIres.Rd")
 susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")
+
+df <- show_param(phenofs, susieIfs, susieIfs2)
+par(mfrow = c(2,3))
+plot_param(df[1:5,], main = "low power")
+plot_param(df[6:10,], main = "high power")
+```
+
+```{r pip1susie, fig.height=4, fig.width=8}
+f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP (low power)" )
+f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
+gridExtra::grid.arrange(f1, f2, ncol =2)
+```
+
+```{r scatter1susie, fig.height=4, fig.width=8}
+par(mfrow=c(1,2))
+f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "susie", main = "low power")
+f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "susie", main = "high power")
+```
+
+## susieI (2)
+* Regions: all regions, LD-defined regions.
+* Susie run parameters: `L=1`.
+We initialize with prior for genes and SNPs as uniform.
+gene.pve ~ 0.1, snp.pve ~ 0.5.
+Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
+Prior variance and residual variance were calculated by SUSIE for each region.
+
+```{r param6susie, fig.height=6, fig.width=8}
+tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
+phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
+susieIfs <- paste0(outputdir, "20201001-", tags, ".config6.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config6.susieI.txt")
+
+df <- show_param(phenofs, susieIfs, susieIfs2)
+par(mfrow = c(2,3))
+plot_param(df[1:5,], main = "low power")
+plot_param(df[6:10,], main = "high power")
+```
+
+## susieI (3)
+* Regions: all regions, LD-defined regions.
+* Susie run parameters: `L=10`.
+We initialize with prior for genes and SNPs as uniform.
+gene.pve ~ 0.1, snp.pve ~ 0.5.
+Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
+Prior variance and residual variance were calculated by SUSIE for each region.
+
+```{r param8susie, fig.height=6, fig.width=8}
+tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
+phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
+susieIfs <- paste0(outputdir, "20201001-", tags, ".config8.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config8.susieI.txt")
 
@@
 
+```{r pip1ser, fig.height=4, fig.width=8}
+f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP (low power)" )
+f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)" )
+gridExtra::grid.arrange(f1, f2, ncol =2)
+```
+
+```{r scatter1ser, fig.height=4, fig.width=10}
+par(mfrow=c(1,2))
+f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "SER", main = "low power")
+f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "SER", main = "high power")
+```
+
 ##  Single effect model/EM (2)
@@
 plot_param(df[6:10,], main = "high power")
+```
+
+```{r pip6ser, fig.height=4, fig.width=8}
+f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP (low power)" )
+f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)" )
+gridExtra::grid.arrange(f1, f2, ncol =2)
+```
+
+```{r scatter6ser, fig.height=4, fig.width=10}
+par(mfrow=c(1,2))
+f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "SER", main = "low power")
+f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "SER", main = "high power")
 ```
@@
 ```
+
+```{r pip7ser, fig.height=4, fig.width=8}
+f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP (low power)" )
+f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)" )
+gridExtra::grid.arrange(f1, f2, ncol =2)
+```
+
+```{r scatter7ser, fig.height=4, fig.width=10}
+par(mfrow=c(1,2))
+f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "SER", main = "low power")
+f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "SER", main = "high power")
+```

### 2020-11-13 analysis/summarize_twas_plots.R

@@ -125,7 +125,7 @@ caliFDR_plot <- function(tags, tag2){
   res <- res[complete.cases(res),]
 
   cp_plot(res$FDR, res$ifcausal, mode ="FDR", main = "TWAS FDP")
-  cat("FDR at bonferroni corrected p = 0.05: ", 1 - mean(res[res$PVALUE < 0.05 /dim(res)[1], "ifcausal"]))
+  cat("FDP at bonferroni corrected p = 0.05: ", 1 - mean(res[res$PVALUE < 0.05 /dim(res)[1], "ifcausal"]))
   return(res)
 }
 

### 2020-11-14 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
more PIP cali

@@
 
+.obn <- function(pips, ifcausal, mode = c("PIP", "FDR")){
+  a_bin <- cut(pips, breaks= seq(0, 1, by=0.1))
+  if (mode == "PIP") {
+    ob = c(by(ifcausal, a_bin, FUN = sum))
+  } else if (mode == "FDR"){
+    ob = c(by((1-ifcausal), a_bin, FUN = sum))
+  }
+  return(ob)
+}
+
+nca_plot <- function(pips, ifcausal, runtag = NULL, mode = c("PIP", "FDR"), main = mode[1], ...){
+
+  # ifcausal:0,1, runtag: for adding std.
+  if (is.null(runtag)){
+     se = 0 
+  } else{
+    dflist <- list()
+    for (rt in unique(runtag)){
+      pips.rt <- pips[runtag == rt]
+      ifcausal.rt <- ifcausal[runtag == rt]
+      dflist[[rt]] <- .obn(pips.rt, ifcausal.rt, mode = mode)
+    }
+    nca_mean <- colMeans(do.call(rbind, dflist))
+    se <- apply(do.call(rbind, dflist), 2, plotrix::std.error)
+  }
+
+  df <- data.frame("ncausal" = nca_mean, "se" = se )
+
+  fig <- ggplot(df) +
+    geom_bar( aes(x=seq(0, 1, by=0.1)[1:10]+0.05, y=ncausal), color ="black", stat="identity", fill="salmon", alpha=0.7, width = 0.1) +
+    geom_errorbar( aes(x= seq(0, 1, by=0.1)[1:10] + 0.05, ymin= ncausal-se, ymax=ncausal+se), width=0.05, colour="black", alpha=0.9, size=0.5) +
+    ggtitle(main) +
+    xlab("PIP") + ylab("No. causal genes")
+    theme_cowplot()
+  return(fig)
+}
+
+ncausal_plot <- function(pipfs, format = "susie", main = "PIP"){
+  df <- NULL
+  for (i in 1:length(pipfs)) {
+    res <- fread(pipfs[i], header = T)
+    res <- data.frame(res[res$type  =="gene", ])
+    res$runtag <- i
+    if (format == "susie"){
+       res <- rename(res, c("susie_pip" = "pip"))
+    } else if (format == "SER"){
+       res <- rename(res, c("SERpip" = "pip"))
+    }
+    res <- res[complete.cases(res),]
+    df <- rbind(df, res)
+  }
+
+  fig <- nca_plot(df$pip, df$ifcausal, df$runtag, mode ="PIP", main = main)
+  return(fig)
+}
+
 .exob <- function(pips, ifcausal, mode = c("PIP", "FDR")){
@@
 }
-
 ```
@@
 
-```{r pip1susie, fig.height=4, fig.width=8}
-f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP (low power)" )
+```{r pip1susie, fig.height=6, fig.width=8}
+f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)" )
 f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
-gridExtra::grid.arrange(f1, f2, ncol =2)
+f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
+f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
+gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
 ```
@@
 
+```{r pip6susie, fig.height=6, fig.width=8}
+f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)" )
+f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
+f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
+f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
+gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
+```
+
+```{r scatter6susie, fig.height=4, fig.width=8}
+par(mfrow=c(1,2))
+f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "susie", main = "low power")
+f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "susie", main = "high power")
+```
+
 ## susieI (3)
@@
 plot_param(df[6:10,], main = "high power")
+```
+
+```{r pip8susie, fig.height=6, fig.width=8}
+f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)" )
+f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
+f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
+f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
+gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
+```
+
+```{r scatter8susie, fig.height=4, fig.width=8}
+par(mfrow=c(1,2))
+f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "susie", main = "low power")
+f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "susie", main = "high power")
 ```
@@
 
-```{r pip1ser, fig.height=4, fig.width=8}
-f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP (low power)" )
-f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)" )
-gridExtra::grid.arrange(f1, f2, ncol =2)
+```{r pip1ser, fig.height=6, fig.width=8}
+f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP(low power)")
+f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)")
+f3 <- ncausal_plot(susieIfs2[1:5], format = "SER", main = "No. Causal Genes (low power)")
+f4 <- ncausal_plot(susieIfs2[6:10], format = "SER", main = "No. Causal Genes (high power)")
+gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
 ```
 
-```{r scatter1ser, fig.height=4, fig.width=10}
+```{r scatter1ser, fig.height=4, fig.width=8}
 par(mfrow=c(1,2))
@@
 
-```{r pip6ser, fig.height=4, fig.width=8}
-f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP (low power)" )
-f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)" )
-gridExtra::grid.arrange(f1, f2, ncol =2)
+```{r pip6ser, fig.height=6, fig.width=8}
+f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP(low power)")
+f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)")
+f3 <- ncausal_plot(susieIfs2[1:5], format = "SER", main = "No. Causal Genes (low power)")
+f4 <- ncausal_plot(susieIfs2[6:10], format = "SER", main = "No. Causal Genes (high power)")
+gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
 ```
 
-```{r scatter6ser, fig.height=4, fig.width=10}
+```{r scatter6ser, fig.height=4, fig.width=8}
 par(mfrow=c(1,2))
@@
 
-```{r pip7ser, fig.height=4, fig.width=8}
-f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP (low power)" )
-f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)" )
-gridExtra::grid.arrange(f1, f2, ncol =2)
+```{r pip7ser, fig.height=6, fig.width=8}
+f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP(low power)")
+f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)")
+f3 <- ncausal_plot(susieIfs2[1:5], format = "SER", main = "No. Causal Genes (low power)")
+f4 <- ncausal_plot(susieIfs2[6:10], format = "SER", main = "No. Causal Genes (high power)")
+gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
 ```
 
-```{r scatter7ser, fig.height=4, fig.width=10}
+```{r scatter7ser, fig.height=4, fig.width=8}
 par(mfrow=c(1,2))

### 2020-11-18 code/run_test_susieI.R
index

@@ -18,6 +18,7 @@ if (length(args) < 5) {
 
 codedir <- "/project2/mstephens/causalTWAS/causal-TWAS/code/"
 source(paste0(codedir, "stats_func.R"))
+source(paste0(codedir, "susie_func.R"))
 source(paste0(codedir,"input_reformat.R"))
 source(paste0(codedir,"susie_filter.R"))
 
@@ -38,8 +39,6 @@ prankfilter.gene <- 1 # for pgene or p2, lowest rank of gene/SNP pvalues > prank
 prankfilter.snp <- 1 # for pgene or p2, lowest rank of gene/SNP pvalues > prankfilter that will be filtered.
 rBFfilter <- 1 # for rBF, regional BF of genes & SNPs < rBFfilter will be filtered
 rfdrfilter <- 1 # for rfdr, if region contains genes or snp with fdr < fdrfilter, it will be kept.
-
-
 ifgene <- F # if limiting to regions with at least one gene.
 ifca <- T # if limiting to regions with at least one causal signal
 regionsfile <- "/home/simingz/causalTWAS/simulations/shared_files/chr1to22-500kb_fn.txt" # need to match input
@@ -49,6 +48,9 @@ ifnullweight <- F # if include/update null weight in iterations.
 prior.gene_init <- NULL
 prior.SNP_init <- NULL
 L <- 1
+plugin_prior_variance <- F
+V.g <- phenores$batch[[1]]$param$sigma_beta ** 2 # prior variance for gene
+V.s <- phenores$batch[[1]]$param$sigma_theta ** 2 # prior variance for SNP
 
 Ncore <- 1
 
@@ -163,6 +165,8 @@ loginfo("susie started for %s", outname)
 cl <- makeCluster(Ncore,outfile="")
 registerDoParallel(cl)
 
+
+varY <- var(phenores$Y)
 prior.SNP_rec <- rep(0, Niter)
 prior.gene_rec <- rep(0, Niter)
 
@@ -204,17 +208,25 @@ for (iter in 1:Niter){
         prior <- c(rep(prior.gene, length(gidx[gidx])), rep(prior.SNP, length(sidx[sidx])))
       }
 
-      nw <- max(0, 1 - sum(prior))
+      if (isTRUE(plugin_prior_variance)){
+        V.scaled <- c(rep(V.g/varY, length(gidx[gidx])), rep(V.s/varY, length(sidx[sidx])))
+        V.scaled <- matrix(rep(V.scaled, each = L), nrow=L)
+      } else{
[…]

### 2020-11-19 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
index
susie different prior variance

@@
 SNPs are downsampled to 1/10 (randomly), eQTLs (see below for definition of eQTL) were added back.
-This ends up with p = `p as.charater(p)` SNPs.
+This ends up with p = `r as.character(p)` SNPs.
 
@@
 
-  df <- data.frame("ncausal" = nca_mean, "se" = se )
+  df <- data.frame("ncausal" = nca_mean, "se" = se)
 
@@
 
+## susieI (4)
+* Regions: all regions, LD-defined regions.
+* Susie run parameters: `L=1`.
+We use true prior variance and plug into SUSIE, so this should be the same as SER/EM (2).
+We initialize with prior for genes and SNPs as uniform.
+gene.pve ~ 0.1, snp.pve ~ 0.5.
+Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
+
+```{r param9susie, fig.height=6, fig.width=8}
+tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
+phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
+susieIfs <- paste0(outputdir, "20201001-", tags, ".config9.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config9.susieI.txt")
+
+df <- show_param(phenofs, susieIfs, susieIfs2)
+par(mfrow = c(2,3))
+plot_param(df[1:5,], main = "low power")
+plot_param(df[6:10,], main = "high power")
+```
+
+```{r pip9susie, fig.height=6, fig.width=8}
+f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)" )
+f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
+f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
+f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
+gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
+```
+
+```{r scatter9susie, fig.height=4, fig.width=8}
+par(mfrow=c(1,2))
+f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "susie", main = "low power")
+f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "susie", main = "high power")
+```
+
+## susieI (5)
+* Regions: all regions, LD-defined regions.
+* Susie run parameters: `L=5`.
+We use true prior variance and plug into SUSIE.
+We initialize with prior for genes and SNPs as uniform.
+gene.pve ~ 0.1, snp.pve ~ 0.5.
+Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
+
+```{r param10susie, fig.height=6, fig.width=8}
+tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
+phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
+susieIfs <- paste0(outputdir, "20201001-", tags, ".config10.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config10.susieI.txt")
+
+df <- show_param(phenofs, susieIfs, susieIfs2)
+par(mfrow = c(2,3))
+plot_param(df[1:5,], main = "low power")
+plot_param(df[6:10,], main = "high power")
+```
+
+```{r pip10susie, fig.height=6, fig.width=8}
+f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)" )
+f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
+f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
+f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high poweunr)")
+gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
+```
+
+```{r scatter10susie, fig.height=4, fig.width=8}
+par(mfrow=c(1,2))
+f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "susie", main = "low power")
+f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "susie", main = "high power")
+```
+
 ##  Single effect model/EM (1)

### 2020-11-20 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
block feature

@@
 ```
+# Block features
+
+```{r block}
+
+a <- read.table(paste0(outputdir, "20201001-1-1.config9.gene.nofilter.r.txt"), header =T)
+b <-  read.table(paste0(outputdir, "20201001-2-1.config9.gene.nofilter.r.txt"), header =T)
+m1 <- read.table(paste0(outputdir, "20201001-1-1.config9.nofilter.r.txt"), header =T)
+m2 <- read.table(paste0(outputdir, "20201001-2-1.config9.nofilter.r.txt"), header =T)
+
+hist(a$p1-a$p0, main = NULL, xlab = "Block size(bp)", col = "salmon")
+
+par(mfrow=c(2,2))
+hist(a$nCausal, xlab = "No.causal genes", col = "salmon", breaks=100, xlim=c(0,15), main ="low power")
+hist(b$nCausal, xlab = "No.causal genes", col = "salmon", breaks=100, xlim=c(0,15), main ="high power")
+hist(m1$nCausal, xlab = "No.causal genes + SNPs", col = "salmon", breaks=100, xlim=c(0,15), main ="low power")
+hist(m2$nCausal, xlab = "No.causal genes + SNPs", col = "salmon", breaks=100, xlim=c(0,15), main ="high power")
+```
 
@@
 
-  # df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1), to=c("Non causal", "Causal"))
-
-  df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1),to=c("darkgreen", "salmon"))
-
-  plot(df$TWAS.p, df$pip, col = df$ifcausal, main = main, xlab = "-log10(TWAS p value)", ylab = "PIP")
-  # fig <- plot_ly(data = df, x = ~ TWAS.p, y = ~ pip, color = ~ ifcausal,
-  #                colors = c( "salmon", "darkgreen"), type ="scatter", text = ~ paste("Name: ", paste0(runtag,":",name),
-  #                                                                  "\nChr: ", chr,  "\nPos:", pos))
+  # df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1),to=c("darkgreen", "salmon"))
+  # plot(df$TWAS.p, df$pip, col = df$ifcausal, main = main, xlab = "-log10(TWAS p value)", ylab = "PIP")
+  # 
+  df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1), to=c("Non causal", "Causal"))
+  fig <- plot_ly(data = df, x = ~ TWAS.p, y = ~ pip, color = ~ ifcausal,
+                 colors = c( "salmon", "darkgreen"), type ="scatter", text = ~ paste("Name: ", paste0(runtag,":",name),
+                                                                   "\nChr: ", chr,  "\nPos:", pos))
 

### 2020-11-20 analysis/simulation-susieI-ukbWG-gtex.adipose_visual.Rmd
pip cali
block feature

@@
 plot_all <- function(SERf, phenof, chrom, ldres = NULL){
-  a <- fread(SERf , header =T)
+  a <- fread(SERf, header =T)
   load(phenof)
@@
 
-```{r sim 2-1-1, out.width = "98%", out.height="95%"}
-chrom <- 1
-tag <- "20201001-2-1"
-pipfile <-  paste0(outputdir, tag, ".config.1.SERI.txt")
+```{r sim 1-3-17, out.width = "98%", out.height="95%"}
+chrom <- 13
+tag <- "20201001-1-2"
+pipfile <-  paste0(outputdir, tag, ".config6.SERI.txt")
 phenof <- paste0(simdatadir, "simu_", tag, "-pheno.Rd")

### 2020-11-24 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
credible set

@@
 
+# L's effect on SuSiE
+
+We run SuSiE on each block with true priors and true prior variances. We tested the difference between using `L=1` and `L=5`. We only did this for the low power setting.
+
+```{r reg functions}
+ncs_ncausal <- function(susief, main = "nCSvs.nCausal") {
+  dt <- fread(susief, header = T)
+  ncausal <- dt[, sum(ifcausal), by=list(b,rn)]
+  nCS <- dt[, max(cs_index), by=list(b,rn)]
+  plot(jitter(ncausal$V1), jitter(nCS$V1), xlab= "No.causal/block", ylab = "No. credible set/block", main= main)
+  return(dt)
+}
+
+sumPIP_ncausal <- function(susief, main = "sumPIPvs.nCausal"){
+  dt <- fread(susief, header = T)
+  ncausal <- dt[, sum(ifcausal), by=list(b,rn)]
+  sumpip <- dt[, sum(susie_pip), by=list(b,rn)]
+  plot(jitter(ncausal$V1), sumpip$V1, xlab= "No.causal/block", ylab = "Regional sum PIP", main = main)
+  return(dt)
+}
+```
+## L=1
+
+Note: the data has been added some noise (both X and Y axes for credible set vs. no.causal, X axis for sum of PIP vs. no.causal) for visualization purposes. 
+```{r L=1, fig.width=8, fig.height=4}
+tags <- paste(rep(1, each = 5), 1:5, sep = "-")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config11.susieI.txt")
+f1 <- caliPIP_plot(susieIfs2, main = "L=1")
+f2 <- ncausal_plot(susieIfs2, format = "susie", main = "No. Causal Genes")
+gridExtra::grid.arrange(f1, f2, ncol =2)
+
+par(mfrow=c(1,2))
+dt <- sumPIP_ncausal(susieIfs2[2], main = "L=1")
+dt <-ncs_ncausal(susieIfs2[2], main = "L=1, CS=0.95")
+```
+
+* True parameters: gene $\pi_1$, `r g <- nrow(dt[type =="gene" & ifcausal !=0])/ nrow(dt[type =="gene"]);g`, SNP $\pi_1$, `r s <-nrow(dt[type =="SNP" & ifcausal !=0])/ nrow(dt[type =="SNP"]);s`.
+[enrichment = `r g/s`].
+* Average PIP: gene `r g <- mean(dt[type =="gene", susie_pip]);g`, SNP `r s <-mean(dt[type =="SNP", susie_pip]);s` . [enrichment = `r g/s`].
+* Proportion in credible set (0.95): gene `r g <-nrow(dt[type =="gene" & cs_index !=0])/ nrow(dt[type =="gene"]);g `, SNP `r s <-nrow(dt[type =="SNP" & cs_index !=0])/ nrow(dt[type =="SNP"]);s`.
+[enrichment = `r g/s`].
+
+## L=5
+
+Note: the data has been added some noise (both X and Y axes for credible set vs. no.causal, X axis for sum of PIP vs. no.causal) for visualization purposes.
+
+```{r L=5, fig.width=8, fig.height=4}
+tags <- paste(rep(1, each = 5), 1:5, sep = "-")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config12.susieI.txt")
+
+f1 <- caliPIP_plot(susieIfs2, main = "L=5")
+f2 <- ncausal_plot(susieIfs2, format = "susie", main = "No. Causal Genes")
+gridExtra::grid.arrange(f1, f2, ncol =2)
+
+par(mfrow=c(1,3))
+dt <- sumPIP_ncausal(susieIfs2[2], main = "L=5")
+dt <- ncs_ncausal(susieIfs2[2], main = "L=5, CS=0.95")
+
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config13.susieI.txt")
+dt2 <- ncs_ncausal(susieIfs2[2], main = "L=5, CS=0.8")
+```
+
+* True parameters: gene $\pi_1$, `r g <- nrow(dt[type =="gene" & ifcausal !=0])/ nrow(dt[type =="gene"]);g`, SNP $\pi_1$, `r s<-nrow(dt[type =="SNP" & ifcausal !=0])/ nrow(dt[type =="SNP"]);s`.
+[enrichment = `r g/s`].
+* Average PIP: gene `r g <- mean(dt[type =="gene", susie_pip]);g`, SNP `r s <-mean(dt[type =="SNP", susie_pip]);s` . [enrichment = `r g/s`].
+* Proportion in credible set (0.95): gene `r g <-nrow(dt[type =="gene" & cs_index !=0])/ nrow(dt[type =="gene"]);g `, SNP `r s <-nrow(dt[type =="SNP" & cs_index !=0])/ nrow(dt[type =="SNP"]);s`.
+[enrichment = `r g/s`].
+* Proportion in credible set (0.8): gene `r g <-nrow(dt2[type =="gene" & cs_index !=0])/ nrow(dt2[type =="gene"]);g `, SNP `r s <-nrow(dt2[type =="SNP" & cs_index !=0])/ nrow(dt2[type =="SNP"]);s`.
+[enrichment = `r g/s`].
+
 # Parameter estimation results

### 2020-11-26 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
filter and rerun

@@
 
+## Select block + susieI (1)
+* Regions: all regions 10 iterations.
+Then filter out regions with probability of having two or more effects > 0.2 to estimate paramters 20 iterations.
+then all regions to get all PIP.
+LD-defined regions.
+* Susie run parameters: `L=1`.
+We use true prior variance and plug into SUSIE.
+We initialize with prior for genes and SNPs as uniform.
+gene.pve ~ 0.1, snp.pve ~ 0.5.
+We use null weight when run susie.
+We run 10 iterations.
+Then filter out regions with probability of having two or more effects.
+Then rerun on selected regions to estimate parameters.
+Lastly, run on all regions with `L=1` to get PIP.
+
+```{r param1susiefl, fig.height=6, fig.width=8}
+tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
+phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
+susieIfs <- paste0(outputdir, "20201001-", tags, ".config14.fl.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config14.fl.susieI.txt")
+
+df <- show_param(phenofs, susieIfs, susieIfs2)
+par(mfrow = c(2,3))
+plot_param(df[1:5,], main = "low power")
+plot_param(df[6:10,], main = "high power")
+
+a <- fread(susieIfs2[1], header =T)
+cat("No.blocks selected for parameter estimation (low power):", nrow(unique(a[, c("b", "rn")])), "\n")
+a <- fread(susieIfs2[6], header =T)
+cat("No.blocks selected for parameter estimation (high power):", nrow(unique(a[, c("b", "rn")])), "\n")
+```
+
+## Select block + susieI (2)
+* Regions: all regions 10 iterations.
+Then filter out regions with probability of having two or more effects > 0.1 to estimate paramters 20 iterations.
+then all regions to get all PIP.
+LD-defined regions.
+* Susie run parameters: `L=1`.
+We use true prior variance and plug into SUSIE.
+We initialize with prior for genes and SNPs as uniform.
+gene.pve ~ 0.1, snp.pve ~ 0.5.
+We use null weight when run susie.
+We run 10 iterations.
+Then filter out regions with probability of having two or more effects.
+Then rerun on selected regions to estimate parameters.
+Lastly, run on all regions with `L=1` to get PIP.
+
+```{r param2susiefl, fig.height=6, fig.width=8}
+tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
+phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
+susieIfs <- paste0(outputdir, "20201001-", tags, ".config15.fl.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config15.fl.susieI.txt")
+
+df <- show_param(phenofs, susieIfs, susieIfs2)
+par(mfrow = c(2,3))
+plot_param(df[1:5,], main = "low power")
+plot_param(df[6:10,], main = "high power")
+
+a <- fread(susieIfs2[1], header =T)
+cat("No.blocks selected for parameter estimation (low power):", nrow(unique(a[, c("b", "rn")])), "\n")
+a <- fread(susieIfs2[6], header =T)
+cat("No.blocks selected for parameter estimation (high power):", nrow(unique(a[, c("b", "rn")])), "\n")
+```
+
 ##  Single effect model/EM (1)

### 2020-11-26 code/run_test_susieI.R

@@ -197,9 +197,10 @@ if (isTRUE(filterandrerun)){
       if (P2 >= P2cut){
         regionlist2[[b]][[rn]] <- NULL
       }
-
     }
   }
+
+  loginfo("Blocks are filtered for parameter estimation: %s blocks left",  sum(unlist(lapply(lapply(regionlist2, names), length))))
 
   pars2 <- susieI(prior.gene_init = prior.gene_init2,
                  prior.SNP_init = prior.SNP_init2,
@@ -208,6 +209,7 @@ if (isTRUE(filterandrerun)){
                  niter = Niter2,
                  Ncore = Ncore)
 
+  loginfo("Parameter estimation done for %s", outname)
 
   susieI(prior.gene_init = pars2[["prior.gene"]],
          prior.SNP_init = pars2[["prior.SNP"]],

### 2020-12-03 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
indididual level pipeline

@@
 
+## Select block + susieI (3)
+* Regions: all regions 5 iterations.
+Then filter out regions with probability of having two or more effects > 0.2 to estimate paramters 20 iterations.
+then all regions to get all PIP.
+LD-defined regions.
+* Susie run parameters: `L=1`.
+We estimate prior variance based on EM and plug into SUSIE.
+We initialize with prior for genes and SNPs as uniform.
+gene.pve ~ 0.1, snp.pve ~ 0.5.
+We use null weight when run susie.
+We run 5 iterations.
+Then filter out regions with probability of having two or more effects.
+Then rerun on selected regions to estimate parameters (20 iterations).
+Lastly, run on all regions with `L=5` to get PIP.
+
+```{r param3susiefl, fig.height=6, fig.width=8}
+tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
+phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
+susieIfs <- paste0(outputdir, "20201001-", tags, ".config16.fl.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config16.fl.susieI.txt")
+
+df <- show_param(phenofs, susieIfs, susieIfs2)
+par(mfrow = c(2,3))
+plot_param(df[1:5,], main = "low power")
+plot_param(df[6:10,], main = "high power")
+
+a <- fread(susieIfs2[1], header =T)
+cat("No.blocks selected for parameter estimation (low power):", nrow(unique(a[, c("b", "rn")])), "\n")
+a <- fread(susieIfs2[6], header =T)
+cat("No.blocks selected for parameter estimation (high power):", nrow(unique(a[, c("b", "rn")])), "\n")
+```
+
+```{r pip3susiefl, fig.height=6, fig.width=8}
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config16.flrerun.susieI.txt")
+f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)")
+f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)")
+f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
+f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
+gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
+```
+
+## Select block + susieI (4)
+* Regions: all regions 5 iterations.
+Then filter out regions with probability of having two or more effects > 0.1 to estimate paramters 20 iterations.
+then all regions to get all PIP.
+LD-defined regions.
+* Susie run parameters: `L=1`.
+We estimate prior variance based on EM and plug into SUSIE.
+We initialize with prior for genes and SNPs as uniform.
+gene.pve ~ 0.1, snp.pve ~ 0.5.
+We use null weight when run susie.
+We run 5 iterations.
+Then filter out regions with probability of having two or more effects.
+Then rerun on selected regions to estimate parameters (20 iterations).
+Lastly, run on all regions with `L=5` to get PIP.
+
+```{r param4susiefl, fig.height=6, fig.width=8}
+tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
+phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
+susieIfs <- paste0(outputdir, "20201001-", tags, ".config17.fl.susieIres.Rd")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config17.fl.susieI.txt")
+
+df <- show_param(phenofs, susieIfs, susieIfs2)
+par(mfrow = c(2,3))
+plot_param(df[1:5,], main = "low power")
+plot_param(df[6:10,], main = "high power")
+
+a <- fread(susieIfs2[1], header =T)
+cat("No.blocks selected for parameter estimation (low power):", nrow(unique(a[, c("b", "rn")])), "\n")
+a <- fread(susieIfs2[6], header =T)
+cat("No.blocks selected for parameter estimation (high power):", nrow(unique(a[, c("b", "rn")])), "\n")
+```
+
+```{r pip4susiefl, fig.height=6, fig.width=8}
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config17.flrerun.susieI.txt")
+f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)")
+f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)")
+f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
+f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
+gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
+```
+
 ##  Single effect model/EM (1)

### 2020-12-11 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
susieI-rss

@@
 ```
+
+##  Summary stats version
+* The run setting is the same as  in "Select block + susieI (3)" except that we used summary statistics and in-sample LD as input.
+
+```{r dirsSS}
+simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
+outputdir <- "~/causalTWAS/simulations/simulation_susieI_rss_20201001/"
+susiedir <- "~/causalTWAS/simulations/simulation_susieI_rss_20201001/"
+```
+```{r param1susieflSS, fig.height=6, fig.width=8}
+tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
+phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
+susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.fl.susieIrssres.Rd")
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.fl.susieIrss.txt")
+
+df <- show_param(phenofs, susieIfs, susieIfs2)
+par(mfrow = c(2,3))
+plot_param(df[1:5,], main = "low power")
+plot_param(df[6:10,], main = "high power")
+
+a <- fread(susieIfs2[1], header =T)
+cat("No.blocks selected for parameter estimation (low power):", nrow(unique(a[, c("b", "rn")])), "\n")
+a <- fread(susieIfs2[6], header =T)
+cat("No.blocks selected for parameter estimation (high power):", nrow(unique(a[, c("b", "rn")])), "\n")
+```
+
+```{r pip1susieflSS, fig.height=6, fig.width=8}
+susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.flrerun.susieIrss.txt")
+f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)")
+f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)")
+f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
+f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
+gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
+```

### 2020-12-11 code/impute_expr_z.R

@@ -1,0 +1,118 @@
+library(data.table)
+# requires tibble
+
+#' @description  generate cis expression of gene given weights
+#' @param dat: a list with the following items:
+#'   - G: matrix, N x M, genotype, need to be scaled
+#'   - snp: matrix, N x 1, SNP name
+#' @param weight a string, pointing to the fusion/twas format of weights
+#' @param method a string,  blup/bslmm/lasso/top1/enet/best
+#'   "best" means the method giving the best cross validation R^2
+#' @param zdf a data frame, with columns "snp", "chr", "pos", "ref", "alt", "z"
+#' @param checksnps T/F, if need to check SNP consistency between weights
+#'    file and genotype. fusion format of weights gives snp information from plink .bim file. checking include chr, pos, A1, A2.
+#' @return a matrix, N x J. Missing genotypes will not contribute to cis-expr and give warnings.
+impute_expr_z <- function(dat, weight, zdf, method = "bslmm", checksnps = F){
+  exprlist <- list()
+  qclist <- list()
+  wgtdir <- dirname(weight)
+  wgtposfile <- file.path(wgtdir, paste0(basename(weight), ".pos"))
+
+  if (!file.exists(paste0(wgtposfile, ".IDfixed"))){
+    wgtpos <- read.table(wgtposfile, header = T, stringsAsFactors = F)
+    wgtpos <- transform(wgtpos,
+                        "ID" = ifelse(duplicated(ID) | duplicated(ID, fromLast=TRUE),
+                                      paste(ID, ave(ID, ID, FUN=seq_along), sep='_ID'),
+                                      ID))
+
+    write.table(wgtpos, file = paste0(wgtposfile, ".IDfixed") , row.names=F, col.names=T, sep="\t", quote = F)
+  }
+  wgtpos <- read.table(paste0(wgtposfile, ".IDfixed"), header = T, stringsAsFactors = F)
+
+  print(paste0("Number of genes with weights provided: ", dim(wgtpos)[1]))
+
+  for (i in 1: dim(wgtpos)[1]){
+    wf <- file.path(wgtdir, wgtpos[i, "WGT"])
+    load(wf)
[…]

### 2020-12-11 code/run_impute_expr_z.R

@@ -1,0 +1,69 @@
+library(data.table)
+library(logging)
+library(tools)
+
+# impute z score using summary statistics
+
+args = commandArgs(trailingOnly=TRUE)
+if (length(args) != 4) {
+  stop("4 arguments must be supplied:
+       * LD reference genotype file name (or multiple file names in a .txt file)
+       * weight dir
+       * sumstats (snp z scores)
+       * out file name", call.=FALSE)
+}
+
+addHandler(writeToFile, file="run_impute_expr_z.R.log", level='DEBUG')
+loginfo('script started ... ')
+
+loginfo("input arg 1 (genotype): %s ", args[1])
+loginfo("input arg 2 (weight): %s ", args[2])
+loginfo("input arg 3 (sumstats): %s ", args[3])
+loginfo("input arg 4 (outname): %s ", args[4])
+
+
+codedir <- "/project2/mstephens/causalTWAS/causal-TWAS/code/"
+source(paste0(codedir, "stats_func.R"))
+source(paste0(codedir,"input_reformat.R"))
+source(paste0(codedir,"impute_expr_z.R"))
+
+weight <- args[2]
+
+zdf <- fread(args[3], header =T)
+setnames(zdf, old = c('t.value','MARKER_ID'), new = c('z','snp'))
+zdf <- data.frame(zdf)
+
+outname <- args[4]
+exprtxt <- file(paste0(outname, "-expr-z.txt"), 'w')
+
+# load genotype data
+pfile <- args[1]
+
+if (file_ext(pfile) == "txt"){
+  pfiles <- read.table(pfile, header = F, stringsAsFactors = F)[,1]
+  outnames <- paste0(outname, "-B", 1:length(pfiles))
+} else {
+  outnames <- outname
+  pfiles <- pfile
+}
+pfileRds <- paste0(pfiles, ".FBM.Rd")
+
+loginfo("start to impute expression ...")
+for ( b in 1:length(pfileRds)){
+  # load genotype
+  pfileRd <- pfileRds[b]
+  outname <- outnames[b]
+  load(pfileRd)
+
+  # impute expression z scores
+  loginfo("generating imputed expression file: %s",  paste0(outname, "-expr-z.Rd"))
+  exprres <- impute_expr_z(dat, weight, zdf, method = "lasso", checksnps = F)
+
+  # covert to FBM
[…]

### 2020-12-11 code/susie_func.R

@@ -1,3 +1,6 @@
+# added susieR::: for all functions from susie
+
+
 # @title update each effect once
 # @param X an n by p matrix of regressor variables
 # @param Y an n vector of response variable
@@ -43,6 +46,35 @@ update_each_effect = function (X, Y, s, estimate_prior_variance = FALSE,
 }
 
 assignInNamespace("update_each_effect", update_each_effect, "susieR")
+
+update_each_effect_rss <- function (R, z, s_init, Sigma, estimate_prior_variance = FALSE,
+          estimate_prior_method = "optim", check_null_threshold = 0)
+{
+  if (!estimate_prior_variance)
+    estimate_prior_method = "none"
+  s = s_init
+  L = nrow(s$alpha)
+  if (L > 0)
+    for (l in 1:L) {
+      s$Rz = s$Rz - R %*% (s$alpha[l, ] * s$mu[l, ])
+      r = z - s$Rz
+      res = single_effect_regression_rss(as.vector(r),
+                                         Sigma, s$V[l], s$pi, estimate_prior_method, check_null_threshold)
+      s$mu[l, ] = res$mu
+      s$alpha[l, ] = res$alpha
+      s$mu2[l, ] = res$mu2
+      s$V[l,] = res$V #different from susieR
+      s$lbf[l] = res$lbf_model
+      s$KL[l] = -res$lbf_model + SER_posterior_e_loglik_rss(R,
+                                                            Sigma, r, res$alpha * res$mu, res$alpha * res$mu2)
+      s$Rz = s$Rz + R %*% (s$alpha[l, ] * s$mu[l, ])
+    }
+  s$Rz = unname(as.matrix(s$Rz))
+  return(s)
+}
+
+assignInNamespace("update_each_effect_rss", update_each_effect, "susieR")
+
 
 # Update a susie fit object in order to initialize susie model.
 init_finalize = function (s, X = NULL, Xr = NULL) {
@@ -90,6 +122,46 @@ init_finalize = function (s, X = NULL, Xr = NULL) {
 }
 
 assignInNamespace("init_finalize", init_finalize ,"susieR")
+
+
+
+init_finalize_rss <- function (s, R = NULL, Rz = NULL){
+  # different from susieR
+  # if (length(s$V) == 1)
+  #   s$V = rep(s$V, nrow(s$alpha))
+  if (!is.numeric(s$sigma2))
+    stop("Input residual variance sigma2 must be numeric")
[…]

### 2021-01-17 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
snakemake workflow using ctwas package

@@
     if (format == "susie"){
-       res <- rename(res, c("susie_pip" = "pip"))
+       res <- rename(res, c("susie_pip" = "pip") )
     } else if (format == "SER"){
-       res <- rename(res, c("SERpip" = "pip"))
+       res <- rename(res, c("SERpip" = "pip"), )
     }
@@
 
-  # df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1),to=c("darkgreen", "salmon"))
-  # plot(df$TWAS.p, df$pip, col = df$ifcausal, main = main, xlab = "-log10(TWAS p value)", ylab = "PIP")
-  # 
-  df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1), to=c("Non causal", "Causal"))
-  fig <- plot_ly(data = df, x = ~ TWAS.p, y = ~ pip, color = ~ ifcausal,
-                 colors = c( "salmon", "darkgreen"), type ="scatter", text = ~ paste("Name: ", paste0(runtag,":",name),
-                                                                   "\nChr: ", chr,  "\nPos:", pos))
+  df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1),to=c("darkgreen", "salmon"))
+  plot(df$TWAS.p, df$pip, col = df$ifcausal, main = main, xlab = "-log10(TWAS p value)", ylab = "PIP")
+
+  # df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1), to=c("Non causal", "Causal"))
+  # fig <- plot_ly(data = df, x = ~ TWAS.p, y = ~ pip, color = ~ ifcausal,
+  #                colors = c( "salmon", "darkgreen"), type ="scatter", text = ~ paste("Name: ", paste0(runtag,":",name),
+  #                                                                  "\nChr: ", chr,  "\nPos:", pos))
 
@@
 
-```{r scatter1susie, fig.height=4, fig.width=8}
-par(mfrow=c(1,2))
-f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "susie", main = "low power")
-f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "susie", main = "high power")
-```
-
 ## susieI (2)
@@
 gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
-```
-
-```{r scatter6susie, fig.height=4, fig.width=8}
-par(mfrow=c(1,2))
-f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "susie", main = "low power")
-f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "susie", main = "high power")
 ```
@@
 
-```{r scatter8susie, fig.height=4, fig.width=8}
-par(mfrow=c(1,2))
-f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "susie", main = "low power")
-f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "susie", main = "high power")
-```
-
 ## susieI (4)
@@
 
-```{r scatter9susie, fig.height=4, fig.width=8}
-par(mfrow=c(1,2))
-f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "susie", main = "low power")
-f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "susie", main = "high power")
-```
-
 ## susieI (5)
@@
 gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
-```
-
-```{r scatter10susie, fig.height=4, fig.width=8}
-par(mfrow=c(1,2))
-f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "susie", main = "low power")
-f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "susie", main = "high power")
 ```
@@
 
+```{r scatter3susiefl, fig.height=4, fig.width=8}
+par(mfrow=c(1,2))
+f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "susie", main = "low power")
+f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "susie", main = "high power")
+```
+
 ## Select block + susieI (4)
@@
 
+```{r scatter4susiefl, fig.height=4, fig.width=8}
+par(mfrow=c(1,2))
+f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "susie", main = "low power")
+f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "susie", main = "high power")
+```
+
 ##  Single effect model/EM (1)
@@
 gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
-```
-
-```{r scatter1ser, fig.height=4, fig.width=8}
-par(mfrow=c(1,2))
-f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "SER", main = "low power")
-f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "SER", main = "high power")
 ```
@@
 
-```{r scatter7ser, fig.height=4, fig.width=8}
-par(mfrow=c(1,2))
-f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "SER", main = "low power")
-f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "SER", main = "high power")
-```
-
 ##  Summary stats version

### 2021-01-17 analysis/simulation-susieI-ukbWG-gtex.adipose_visual.Rmd
snakemake workflow using ctwas package

@@
 
-plot_all <- function(SERf, phenof, chrom, ldres = NULL){
-  a <- fread(SERf, header =T)
+plot_all <- function(susief, pf, phenof, chrom, ldres = NULL){
+  a <- fread(susief, header =T)
+  b1 <- fread(paste0(pf, ".snpgwas.txt.gz"), header =T)
+  b2 <- fread(paste0(pf, ".exprgwas.txt.gz"), header =T)
+  b <- rbind(b1, b2)
+  setnames(b, old = c('X.CHROM', 'PVALUE', 'MARKER_ID'),
+         new = c("chr", 'pvalue', 'name'))
+
   load(phenof)
@@
   a <- a[a$chr == chrom,]
+  b <- b[b$chr == chrom,]
+
+  a <- merge(a, b, by = "name", all.x =T)
 
@@
 
-  colnames(a)[colnames(a) == "SERpip"] <-  "susiePIP"
+  colnames(a)[colnames(a) == "susie_pip"] <-  "susiePIP"
   colnames(a)[colnames(a) == "pos"] <-  "p0"
@@
   a[, "PVALUE"] <- -log10(a[, "pvalue"])
-  a[, "lbf"] <- log10(exp(a[, "lbf"]))
 
@@
 
-  fig1 <- plot_susie_pip(dt.gene) 
-  fig2 <- plot_lbf(dt.gene)
+  fig1 <- plot_susie_pip(dt.gene)  # %>%  add_trace(p, x = list(1), y = list(1), name = " ", showlegend = F)
+  # fig2 <- plot_lbf(dt.gene)
   fig3 <- plot_p(dt.gene)
@@
   fig4 <- plot_susie_pip(dt.snp)
-  fig5 <- plot_lbf(dt.snp) 
+  # fig5 <- plot_lbf(dt.snp) 
   fig6 <- plot_p(dt.snp) %>% layout( plot_bgcolor='#EFF6FC')
 
-  fig <- subplot(fig1, fig2, fig3, fig4, fig5, fig6, nrows=6, shareX = T, titleY = T)
+  fig <- subplot(fig1, fig3, fig4, fig6, nrows=4, shareX = T, titleY = T)
 
-  fig <- fig %>% layout(autosize = F, width = 600, height = 800, margin = 0.02,
+  fig <- fig %>% layout(autosize = F, width = 600, height = 500, margin = 0.02,
                         xaxis = list(title = "Genomic position"))
@@
 ```{r sim 1-3-17, out.width = "98%", out.height="95%"}
-chrom <- 13
-tag <- "20201001-1-2"
-pipfile <-  paste0(outputdir, tag, ".config6.SERI.txt")
+chrom <- 2
+tag <- "20201001-2-5"
+susief <-  paste0(outputdir, tag, ".config16.flrerun.susieI.txt")
+pf <- paste0(simdatadir, tag)
 phenof <- paste0(simdatadir, "simu_", tag, "-pheno.Rd")
-plot_all(pipfile, phenof, chrom)
+plot_all(susief, pf, phenof, chrom)
 ```

### 2021-01-17 code/before_package/SERI_func.R

@@ -1,0 +1,103 @@
+#' @title Bayesian single-effect linear regression of y on X,
+#' adapted from susieR (single_effect_regression.R)
+#'
+#' @param Y An n vector.
+#'
+#' @param X An n by p matrix of covariates. centered and scaled.
+#'
+#' @param V The prior variance.
+#'
+#' @param residual_variance The residual variance.
+#'
+#' @param prior_weights A p vector of prior weights.
+#'
+#' @return a list
+#'
+#' @importFrom Matrix colSums
+#'
+#'
+get_lbf =
+  function (Y, X, V, residual_variance = 1) {
+
+    ytX = crossprod(Y, X)
+    Xty = t(ytX)
+    xTx = colSums(X * X)
+    betahat = (1/xTx) * Xty
+    shat2 = residual_variance/xTx
+
+
+    # log(bf) for each SNP
+    lbf = dnorm(betahat,0,sqrt(V + shat2),log = TRUE) -
+      dnorm(betahat,0,sqrt(shat2),log = TRUE)
+
+    # Deal with special case of infinite shat2 (e.g., happens if X does
+    # not vary).
+    lbf[is.infinite(shat2)] = 0
+
+    z = betahat/sqrt(shat2)
+    z[is.infinite(z)] = 0
+
+    pvalue = 2*pnorm(-abs(z))
+
+    lbfout <- cbind(lbf, betahat, shat2, z, pvalue)
+    colnames(lbfout) <- c("lbf", "betahat", "shat2", "z", "pvalue")
+
+    return(lbfout)
+  }
+
+
+#' @title Bayesian single-effect linear regression of y on X,
+#' adapted from susieR (single_effect_regression.R)
+#'
+#' @param lbf a vector
+#'
+#' @return A list with the following elements:
+#'
+#' \item{alpha}{Vector of posterior inclusion probabilities;
+#'   \code{alpha[i]} is posterior probability that that the ith
+#'   coefficient is non-zero.}
+#'
+#' \item{lbf_model}{Log-Bayes factor for the single effect regression.}
+#'
+#' \item{loglik}{The logarithm of the likelihood, \eqn{p(y | X, V)}.}
+#'
+#' @importFrom Matrix colSums
+#'
+#'
+SER =
+  function (lbf, prior_weights = NULL, null_weight = NULL) {
+    if (is.null(prior_weights) & is.null(null_weight))
+      prior_weights = c(rep(1/length(lbf), length(lbf)))
+
[…]

### 2021-01-17 code/before_package/impute_expr_z.R

@@ -1,0 +1,118 @@
+library(data.table)
+# requires tibble
+
+#' @description  generate cis expression of gene given weights
+#' @param dat: a list with the following items:
+#'   - G: matrix, N x M, genotype, need to be scaled
+#'   - snp: matrix, N x 1, SNP name
+#' @param weight a string, pointing to the fusion/twas format of weights
+#' @param method a string,  blup/bslmm/lasso/top1/enet/best
+#'   "best" means the method giving the best cross validation R^2
+#' @param zdf a data frame, with columns "snp", "chr", "pos", "ref", "alt", "z"
+#' @param checksnps T/F, if need to check SNP consistency between weights
+#'    file and genotype. fusion format of weights gives snp information from plink .bim file. checking include chr, pos, A1, A2.
+#' @return a matrix, N x J. Missing genotypes will not contribute to cis-expr and give warnings.
+impute_expr_z <- function(dat, weight, zdf, method = "bslmm", checksnps = F){
+  exprlist <- list()
+  qclist <- list()
+  wgtdir <- dirname(weight)
+  wgtposfile <- file.path(wgtdir, paste0(basename(weight), ".pos"))
+
+  if (!file.exists(paste0(wgtposfile, ".IDfixed"))){
+    wgtpos <- read.table(wgtposfile, header = T, stringsAsFactors = F)
+    wgtpos <- transform(wgtpos,
+                        "ID" = ifelse(duplicated(ID) | duplicated(ID, fromLast=TRUE),
+                                      paste(ID, ave(ID, ID, FUN=seq_along), sep='_ID'),
+                                      ID))
+
+    write.table(wgtpos, file = paste0(wgtposfile, ".IDfixed") , row.names=F, col.names=T, sep="\t", quote = F)
+  }
+  wgtpos <- read.table(paste0(wgtposfile, ".IDfixed"), header = T, stringsAsFactors = F)
+
+  print(paste0("Number of genes with weights provided: ", dim(wgtpos)[1]))
+
+  for (i in 1: dim(wgtpos)[1]){
+    wf <- file.path(wgtdir, wgtpos[i, "WGT"])
+    load(wf)
[…]

### 2021-01-17 code/before_package/input_reformat.R

@@ -1,0 +1,243 @@
+library(data.table) # requires v.1.12.7 or higher
+library(bigstatsr)
+library(bigreadr)
+library(pgenlibr) #install_github("chrchang/plink-ng", subdir="/2.0/pgenlibr"), Notice that the installation of pgenlibr requires zstd(>=1.4.4). It can be built from source or simply available from conda, pip or brew.
+library(bigmemory)
+
+source(paste0(codedir, "stats_func.R"))
+
+write.gemma.pheno <- function(outfile, y) {
+  if (is.numeric(y))
+    y <- round(y,digits = 6)
+  write.table(y, outfile,quote = FALSE,row.names = FALSE,col.names = FALSE)
+}
+
+write.gemma.geno <- function(outfile, geno, labels, append=F) {
+  geno <- t(geno)
+  geno <- data.table(geno)
+  geno <- round(geno,digits = 3)
+  geno <- cbind(paste0("rs",labels),"A","B",geno) # fake allele for now
+  fwrite(geno, outfile, col.names = FALSE, sep = " ", quote = FALSE, row.names = FALSE, append=append)
+  rm(geno);gc()
+}
+
+write.gemma.geno.chunks <- function(outfile, geno, labels, n_per_chunk){
+
+	l = split(1:ncol(geno), ceiling(seq_along(1:ncol(geno))/n_per_chunk))
+
+	# Write large data frame to csv in chunks
+	count = 1
+	for(i in l){
+	  cat(paste0("writing chunks: ",count ,"/",length(l),"...\n"))
+	  if (count==1){
+		  write.gemma.geno(outfile, geno[,i], labels[i], append=F)
+	  }
+	  else{
+		  write.gemma.geno(outfile, geno[,i], labels[i], append=T)
+	  }
+	  count = count + 1
+	}
+}
+
+gemma.prepare <- function(Rdatafile,
+                       outnamebase, chromosomes = NULL) {
+  # read in Rdata file
+  cat("* loading data.\n")
+  load(Rdatafile)
+  nsample <- dim(X)[1]
+  nSNP <- dim(X)[2]
+  save(nsample,nSNP,file= paste0(outnamebase,".QC.RData"))
+
+  # Write the phenotype data
+  cat("* Writing phenotype to file.\n")
+  phenofile <- paste0(outnamebase,".pheno.txt")
+  write.gemma.pheno(phenofile, y)
+
+  # Write genotype data
[…]

### 2021-01-17 code/before_package/mr.ash2.R

@@ -1,0 +1,359 @@
+library(mr.ash.alpha)
+library(biglasso)
+
+#' this is a simplified version of mr.ash
+#' Difference:
+#' - doesn't have the remove covariates function. no `Z` will be accepted.
+#' - preprocessing (center/scale) for X and y need to be done before running
+#'   this function (no `standardize`, `intercept` arguments).
+#' - precalculation for w can be provided as an argument `w`.
+
+mr.ashs                      = function(X, y, sa2 = NULL,
+                                       K = 20, method = c("caisa","accelerate","block","sigma_scaled","sigma_indep"),
+                                       w = NULL,
+                                       max.iter = 1000, min.iter = 1,
+                                       beta.init = NULL,
+                                       update.pi = TRUE, pi = NULL,
+                                       update.sigma = TRUE, sigma2 = NULL,
+                                       update.order = NULL,
+                                       mixsqpiter = 5, mode = FALSE,
+                                       tol = set_default_tolerance(),
+                                       verbose = TRUE){
+
+  # get sizes
+  n            = dim(X)[1];
+  p            = dim(X)[2];
+
+  # match method
+  method      <- match.arg(method)
+
+  # set default tolerances unless specified
+  tol0         = set_default_tolerance()
+  tol          = modifyList(tol0,tol,keep.null = TRUE)
+
+  data         = list(y = y, alpha = mean(y))
+
+  # initialize beta
+  if ( is.null(beta.init) ){
+    data$beta  = as.vector(double(p));
+  } else {
+    data$beta  = as.vector(beta.init);
+  }
+
+  data$beta[1] = data$beta[1] + 0;      # to make sure beta.init is not modified
+
+  # initialize r
+  r            = data$y - X %*% data$beta;
+
+  # sigma2
+  if ( is.null(sigma2) ) {
+    sigma2 = c(var(r));
+  }
+
+  # set sa2 if missing
+  if ( is.null(sa2) ) {
+    sa2             = (2^((0:19) / 20) - 1)^2
+  }
[…]

### 2021-01-17 code/before_package/mr.ash2_FBM.R

@@ -1,0 +1,385 @@
+library(mr.ash.alpha)
+library(biglasso)
+
+#' this is a simplified version of mr.ash
+#' Difference:
+#' - doesn't have the remove covariates function. no `Z` will be accepted.
+#' - preprocessing (center/scale) for X and y need to be done before running
+#'   this function (no `standardize`, `intercept` arguments).
+#' - precalculation for w can be provided as an argument `w`.
+
+mr.ashs                      = function(X, y, sa2 = NULL,
+                                       K = 20,
+                                       method = c("caisa","accelerate","block","sigma_scaled","sigma_indep"),
+                                       w = NULL,
+                                       max.iter = 1000, min.iter = 1,
+                                       beta.init = NULL,
+                                       update.pi = TRUE, pi = NULL,
+                                       update.sigma = TRUE, sigma2 = NULL,
+                                       update.order = NULL,
+                                       mixsqpiter = 5, mode = FALSE,
+                                       tol = set_default_tolerance(),
+                                       verbose = TRUE){
+
+  # get sizes
+  n            = dim(X)[1];
+  p            = dim(X)[2];
+
+  # match method
+  method      <- match.arg(method)
+
+  # set default tolerances unless specified
+  tol0         = set_default_tolerance()
+  tol          = modifyList(tol0,tol,keep.null = TRUE)
+
+  data         = list(y = y, alpha = mean(y))
+
+  # initialize beta
+  if ( is.null(beta.init) ){
+    data$beta  = as.vector(double(p));
+  } else {
+    data$beta  = as.vector(beta.init);
+  }
+
+  data$beta[1] = data$beta[1] + 0;      # to make sure beta.init is not modified
+
+  # initialize r
+  r            = data$y - X %*% data$beta;
+
+  # sigma2
+  if ( is.null(sigma2) ) {
+    sigma2 = c(var(r));
+  }
+
+  # set sa2 if missing
+  if ( is.null(sa2) ) {
[…]

### 2021-01-17 code/before_package/run_ld.R

@@ -1,0 +1,74 @@
+library(logging)
+library(tools)
+
+args = commandArgs(trailingOnly=TRUE)
+if (length(args) != 4) {
+  stop("3 arguments must be supplied:
+       * genotype file name
+       * expr Rd
+       * pheno Rd
+       * out file name", call.=FALSE)
+}
+
+addHandler(writeToFile, file="run_ld.R.log", level='DEBUG')
+loginfo('script started ... ')
+
+loginfo("input arg 1 (genotype): %s ", args[1])
+loginfo("input arg 2 (expr): %s ", args[2])
+loginfo("input arg 3 (expr): %s ", args[3])
+loginfo("input arg 4 (outname): %s ", args[4])
+
+
+codedir <- "/project2/mstephens/causalTWAS/causal-TWAS/code/"
+source(paste0(codedir,"ld.R"))
+
+pfile <- args[1]
+if (file_ext(pfile) == "txt"){
+  pfiles <- read.table(pfile, header =F, stringsAsFactors = F)[,1]
+  efiles <- read.table(args[2], header =F, stringsAsFactors = F)[,1]
+} else {
+  pfiles <- pfile
+  efiles <- args[2]
+}
+pfileRds <- paste0(pfiles, ".unscaled.FBM.Rd")
+
+
+# load pheno Rd file, variable: pheno
+load(args[3])
+
+outname <- args[4]
+
+load(pfileRds[13])
+load(efiles[13])
+
+geno.t <- list()
+
+geno.t[[1]] <- .eqtl_geno("STAG1")
+
+name <- "rs763581481"
+temp <- dat$G[, dat$snp == name, drop = F]
+colnames(temp) <- name
+geno.t[[name]] <- temp
+
+rsqlist <- list()
+
+for (b in 1:length(pfiles)){
+  # load genotype data
+  load(pfileRds[b])
+
+  # load expression Rd file, variable: exprres
+  load(efiles[b])
+
+  causalg.eqtl <-lapply(exprres$gnames[phenores$batch[[b]]$param$idx.cgene], .eqtl_geno)
+  cg <- do.call(cbind, causalg.eqtl)
+
+  csnp <- dat$G[ ,phenores$batch[[b]]$param$idx.cSNP]
+  colnames(csnp) <- dat$snp[phenores$batch[[b]]$param$idx.cSNP, 1]
+
[…]

### 2021-01-17 code/gen_mr.ash2_output.R

@@ -1,104 +1,0 @@
-
-para_mr.ash <- function(fit){
-  pi1 <-  1-fit$pi[[1]]
-  pve <- get_pve(fit)
-  p <- c(pi1, pve)
-  names(p) <- c("pi1", "pve")
-  return(p)
-}
-
-para_mr.ash2 <- function(g.fit, s.fit){
-
-  gpara <- para_mr.ash(g.fit)
-  names(gpara) <- paste0("gene.", names(gpara))
-
-  spara <- para_mr.ash(s.fit)
-  names(spara) <- paste0("snp.", names(spara))
-
-  outdf <- data.frame("estimated" = c(gpara,spara))
-  outdf$truth <- c(phenores$param$J.c/phenores$param$J,
-                   phenores$param$pve.expr.truth,
-                   phenores$param$M.c/phenores$param$M,
-                   phenores$param$pve.snp.truth)
-
-  return(outdf)
-}
-
-gen_mr.ash_gene_output <- function(g.fit, outname){
-  e.b <- rep(0, length(g.fit$beta))
-  e.b[phenores$param$idx.cgene] <- phenores$param$e.beta
-
-  indi <- rep(0, length(g.fit$beta))
-  indi[phenores$param$idx.cgene] <- 1
-  pip <- get_pip(g.fit)
-
-  g.outdf <- data.frame("chr" = exprres$chrom,
-                        "p0" = exprres$p0,
-                        "p1" = exprres$p1,
-                        "ifcausal"= indi,
-                        "name" = colnames(exprres$expr),
-                        "TrueBeta" = e.b,
-                        "EstBeta" = g.fit$beta,
-                        "PIP" = pip)
-
-  g.outdf.sorted <- g.outdf[with(g.outdf, order(chr, p0)), ]
-  write.table(g.outdf.sorted , file= paste0(outname, ".expr.txt") , row.names=F, col.names=T, sep="\t", quote = F)
-  return(g.outdf.sorted)
-}
-
-gen_mr.ash_snp_output <- function(s.fit, outname){
-  s.b <- rep(0, length(s.fit$beta))
-  s.b[phenores$param$idx.cSNP] <- phenores$param$s.theta
-
-  indi <- rep(0, length(s.fit$beta))
-  indi[phenores$param$idx.cSNP] <- 1
-  pip <- get_pip(s.fit)
-
[…]

### 2021-01-17 code/gwas.R

@@ -22,57 +22,61 @@ GWAA <- function(geno, pheno, snpname = NULL, anno = NULL, outname, family = gau
     snpname <- 1: ncol(geno)
   }
 
-  nSNPs <- ncol(geno)
-  genosplit <- ceiling(nSNPs/nSplits) # number of SNPs in each subset
-
-  snp.start <- seq(1, nSNPs, genosplit) # index of first SNP in group
-  snp.stop <- pmin(snp.start+genosplit-1, nSNPs) # index of last SNP in group
-
-  columns<-c("#CHROM",	"BEGIN",	"END", "MARKER_ID", "Estimate", "Std.Error", "t-value", "PVALUE") # header for LOCUSZOOM.
+  columns<- c(colnames(anno), "id", c("Estimate", "Std.Error", "t-value", "PVALUE"))
   write.table(t(columns), outname, row.names=FALSE, col.names=FALSE, quote=FALSE, sep = "\t")
 
+  nSNPs <- ncol(geno)
+  if (nSNPs > 0){
 
-  if (!is.null(s.idx)){
-    pheno <- pheno[s.idx, ,drop=F]
-  }
+    genosplit <- ceiling(nSNPs/nSplits) # number of SNPs in each subset
 
-  for (part in 1:nSplits) {
-
-    geno.i <- geno[ ,snp.start[part]:snp.stop[part]]
+    snp.start <- seq(1, nSNPs, genosplit) # index of first SNP in group
+    snp.stop <- pmin(snp.start+genosplit-1, nSNPs) # index of last SNP in group
 
     if (!is.null(s.idx)){
-      geno.i <- geno.i[s.idx, ]
+      pheno <- pheno[s.idx, ,drop=F]
     }
 
-    snpname.i <- snpname[snp.start[part]:snp.stop[part]]
-    if (!is.null(anno)) {
-      anno.i <- anno[snp.start[part]:snp.stop[part], ]
-    } else {
-      anno.i <- NULL
+    for (part in 1:length(snp.start)) {
+
+      geno.i <- geno[ , snp.start[part]:snp.stop[part], drop = F]
+
+      if (!is.null(s.idx)){
+        geno.i <- geno.i[s.idx, ]
+      }
+
+      snpname.i <- snpname[snp.start[part]:snp.stop[part]]
+      if (!is.null(anno)) {
+        anno.i <- anno[snp.start[part]:snp.stop[part], ]
+      } else {
+        anno.i <- NULL
[…]

### 2021-01-17 code/run_UKB_phenotype.R

@@ -72,12 +72,12 @@ suppressMessages(library(dplyr))
 
 
 load("/home/szhao1/causal-TWAS/phenotype_data/population_ukb26140_ukb32141.Rd")
-sa <- fread("/home/szhao1/causal-TWAS/genotype_data/ukbiobank_samples70000.txt", header = F,colClasses = "character")[,1]
-output.file <- file.path( "/home/szhao1/causal-TWAS/phenotype_data/population_ukb26140_ukb32141_s70000.csv")
+sa <- data.table::fread("/home/szhao1/causal-TWAS/genotype_data/ukbiobank_samples200k.txt", header = F,colClasses = "character")[,1]
+output.file <- file.path( "/home/szhao1/causal-TWAS/phenotype_data/population_ukb26140_ukb32141_s200k.csv")
 
 # SELECT samples
 colnames(sa) <- "id"
-dat <- inner_join(dat, sa) # N=39993
+dat <- inner_join(dat, sa) # N=39993, 79982, 199968
 
 # Convert all columns except the first one (the first column contains
 # the sample ids) to numeric values, and set all empty strings to NA.
@@ -94,33 +94,34 @@ for (i in 2:n) {
 # When the "genetic ethnic grouping" column is included, this removes
 # any samples that are not marked as being "White British". The
 # "outliers" have value 1 when it is an outlier, NA otherwise.
-# including the genetic ethnic column, filtered ~ 5000 sampels
+# including the genetic ethnic column, filtered ~ 5000, 13000, 33456 samples
 cols <- !(names(dat) == "outliers" | grepl("relatedness_genetic",names(dat)))
 rows <- which(rowSums(is.na(dat[, cols])) == 0)
 dat  <- dat[rows,]
 cat(sprintf("After removing rows with NAs, %d rows remain.\n",nrow(dat)))
 
 # Remove rows with mismatches between self-reported and genetic sex
-# This step should filter out 2 rows.
+# This step should filter out 2,2,11 rows.
 dat <- dat %>% filter(sex == sex_genetic)
 cat(sprintf("After removing sex mismatches, %d rows remain.\n",nrow(dat)))
 
 # Remove "missingness" and "heterozygosity" outliers as defined by UK
-# Biobank. This step should filter out 0 rows. Note that this step
+# Biobank. This step should filter out 0,0,0 rows. Note that this step
 # will remove any samples in which the "missingness" column is greater
 # than 5%.
 dat <- dat %>% filter(is.na(outliers))
[…]

### 2021-01-17 code/run_simulate_pheno.R

@@ -1,95 +1,33 @@
-library(data.table)
+library(ctwas)
 library(logging)
-library(tools)
 
 args = commandArgs(trailingOnly=TRUE)
-if (length(args) != 4) {
-  stop("4 arguments must be supplied:
-       * genotype file name (or multiple file names in a .txt file)
-       * impute expr Rd file (or multiple file names in a .txt file)
+if (length(args) != 5) {
+  stop("5 arguments must be supplied:
+       * genotype file name (multiple file names in a .txt file)
+       * imputed expr file (multiple file names in a .txt file)
        * param R file
-       * out file name", call.=FALSE)
+       * out file name
+       * outputdir", call.=FALSE)
 }
-
-addHandler(writeToFile, file="run_simulate_phenotype.R.log", level='DEBUG')
-loginfo('script started ... ')
-
-loginfo("input arg 1 (genotype): %s ", args[1])
-loginfo("input arg 2 (imputed expression): %s ", args[2])
-loginfo("input arg 3 (param): %s ", args[3])
-loginfo("input arg 4 (outname): %s ", args[4])
-
 
 codedir <- "/project2/mstephens/causalTWAS/causal-TWAS/code/"
-source(paste0(codedir, "stats_func.R"))
-source(paste0(codedir,"input_reformat.R"))
-source(paste0(codedir,"simulate_phenotype.R"))
+source(paste0(codedir, "stats_func.R"))       
+source("~/causalTWAS/causal-TWAS/code/simulate_phenotype.R")
 
-
-outname <- args[4]
-
-# prepare input files. Genotype and expression are already scaled.
-pfile <- args[1]
-
-if (file_ext(pfile) == "txt"){
-  pfiles <- read.table(pfile, header = F, stringsAsFactors = F)[,1]
-  efiles <- read.table(args[2], header =F, stringsAsFactors = F)[,1]
-} else {
-  pfiles <- pfile
-  efiles <- args[2]
-}
-pfileRds <- paste0(pfiles, ".FBM.Rd")
-
-
-# set sigma based on pve
+pgenfs <- read.table(args[1], header = F, stringsAsFactors = F)[,1]
+exprfs <- read.table(args[2], header = F, stringsAsFactors = F)[,1]
 source(args[3])
+outname <- args[4]
+outputdir <- args[5]
+
+loginfo("simulating phenotype for %s", outname)
[…]

### 2021-01-17 code/run_test_mr.ash_gene.R

@@ -1,43 +1,0 @@
-library(logging)
-
-args = commandArgs(trailingOnly=TRUE)
-if (length(args) != 3) {
-  stop("3 arguments must be supplied:
-       * expr Rd
-       * pheno Rd
-       * out file name", call.=FALSE)
-}
-
-addHandler(writeToFile, file="run_mr.ash_gene.R.log", level='DEBUG')
-loginfo('script started ... ')
-
-loginfo("input arg 1 (expr): %s ", args[1])
-loginfo("input arg 2 (pheno): %s ", args[2])
-loginfo("input arg 3 (outname): %s ", args[3])
-
-
-codedir <- "/project2/mstephens/causalTWAS/causal-TWAS/code/"
-source(paste0(codedir, "stats_func.R"))
-source(paste0(codedir, "fit_mr.ash.R"))
-source(paste0(codedir, "gen_mr.ash2_output.R"))
-
-# load expression Rd file, variable: exprres
-load(args[1])
-loginfo("expr scaling ...")
-expr <-  scaleRcpp(exprres$expr)
-gc()
-
-# load phenotype Rd file, variable: phenores
-load(args[2])
-phenores$Y <-  phenores$Y - mean(phenores$Y)
-
-outname <- args[3]
-
-# run mr.ash2s
-loginfo("run mr.ash for gene expression ...")
-mr.ash.fit <- mr.ash( X = expr,
-                      y = phenores$Y)
-save(mr.ash.fit, file = paste0(outname, ".mr.ashG.Rd"))
-
-a <- gen_mr.ash_gene_output(mr.ash.fit, paste0(outname, '-mr.ash'))
-

### 2021-01-17 code/simulate_phenotype.R

@@ -1,14 +1,6 @@
-library(data.table)
-# requires tibble
-#' @param dat a list with the following items:
-#'   - G: matrix, N x M, genotype, need to be scaled
-#'   - snp: matrix, N x 1, SNP name
-#'   - ref
-#'   - alt
-#'   - chr
-#'   - pos
-#'   - expr: matrix, N x J
-#' @param mode string, "snp-only", "snp-expr"
+#' Simulate phenotype
+#' @description simulate Y under the sparse model (spike and slab prior).
+#' will standardize gene expression and SNP genotype when simulate.
 #' @param pve.expr scalar [0,1], not used when if `sigma_beta` is given
 #' @param pve.snp scalar [0,1], not used if `sigma_theta` is given
 #' @param pi_beta scalar [0,1]
@@ -17,67 +9,84 @@ library(data.table)
 #' @param sigma_theta default NULL, if not given, will calculate from pve.
 #' @return a list, Y.g is phenotype determined by genetic, will need to add
 #'  error term ( pve calculation is based on N(0,1)).
-simulate_phenotype<- function(dat,
-                              mode = "snp-expr",
+#' requires tibble, pgenlibr, scaleRcpp
+simulate_phenotype<- function(pgenfs,
+                              exprfs,
                               pve.expr = 0.1,
                               pve.snp = 0.1,
                               pi_beta = 0.01,
                               pi_theta = 0.0001,
                               sigma_beta = NULL,
                               sigma_theta = NULL
-                              ) {
+) {
+  pvarfs <- sapply(pgenfs, prep_pvar, outputdir = outputdir)
+  exprvarfs <- sapply(exprfs, prep_exprvar)
 
-  N <- dim(dat$G)[1]
-  M <- dim(dat$G)[2]
-  J <- dim(dat$expr)[2]
+  pgen1 <- prep_pgen(pgenf = pgenfs[1], pvarfs[1])
+  N <- pgenlibr::GetRawSampleCt(pgen1)
 
-  if (mode == "snp-only"){
-    expr.meanvar <- NULL
-    idx.cgene <- NULL
[…]

### 2021-01-17 code/susieI_parallel_test.R

@@ -1,166 +1,0 @@
-#-------------------------------------------------------------------------------------------
-cl <- makeCluster(Ncore, outfile="")
-registerDoParallel(cl)
-
-
-iter <- 1
-b <- 1
-# test 1 --> not working, low CPU
-a <- dat$G$bm()
-bm.desc <- describe(a)
-
-
-susiefunc <- function(rn, bm.desc){
-  print(c(b, rn))
-  
-  sidx <- regionlist[[b]][[rn]][["sidx"]]
-  
-  a <- bigmemory::attach.big.matrix(bm.desc)
-  sidx.n <-c(1:dim(a)[2])[sidx]
-  x <- a[, sidx.n]
-  1:5
-  
-}
-
-parallel::clusterCall(cl, function() {
-  
-  library(bigmemory)
-  library(susieR)
-})
-
-parallel::clusterExport(cl, c("regionlist", "bm.desc", "b"),
-                        envir=environment())
-
-fold.results <- parallel::parLapply(cl = cl, X = seq_len(length(regionlist[[b]][1:100])), fun = susiefunc, bm.desc = bm.desc)
-
-
-# test 2 --> not working, low CPU
-outdf.b <- foreach(r = 1: length(regionlist[[b]]), .combine = "rbind",
-                   .packages = c("susieR", "bigstatsr")) %dopar% {
-                     
-                     r.s <- rs[r]
-                     print(r.s)
-                     x <- dat$G[, (1+r.s):(50+r.s)]
-                     5
-                   }
-
-# test 3 --> not working, low CPU
-sidx <- regionlist[[b]][["1"]][["sidx"]]
-outdf.b <- foreach(rn = names(regionlist[[b]]), .combine = "rbind",
-                   .packages = c("susieR", "bigstatsr")) %dopar% {
-                     print(c(b, rn))
-                     
-                     X <- dat$G[, sidx]
-                     1: 6
-                   }
-
-#test 4 --> slow, when assigning jobs
-outdf.b.list <- list()
-for (b in 1:22){
-  
-  # load genotype data
-  load(pfileRds[b])
-  
-  # load expression Rd file, variable: exprres
-  load(efiles[b])
-  
-  cau <- c(dat$snp[phenores$batch[[b]]$param$idx.cSNP,],
-           exprres$gnames[phenores$batch[[b]]$param$idx.cgene])
-  
[…]

### 2021-01-17 code/susie_filter.R

@@ -1,117 +1,0 @@
-#' Filter regions
-#' require global variables: `phenores`, `exprres`, `dat`, `cau`
-#' and cutoff for different filter types.
-#' @param reg a matrix/data.frame, with columns chr, p0, p1
-#' @param filtertype "nofilter", "mr.ash", "mr.ash2s", "p2", "pgene", "rBF"
-#' @param filterfile, for "nofilter", filterfile is not used
-filter_region <- function(reg, filtertype, filterfile){
-
-  # prep/filter regions
-  anno <- rbind(cbind(dat$chr, dat$pos, dat$snp, "SNP"),
-                cbind(exprres$chrom, exprres$p0, exprres$gnames, "gene"))
-  colnames(anno) <- c("chrom", "p0", "name", "type")
-
-  regions <- NULL
-  chroms <- unique(reg$chrom)
-
-  if (filtertype == 'nofilter'){
-    for (chrom in chroms){
-      itv <- cbind(reg[reg$chrom == chrom, ]$p0, reg[reg$chrom == chrom, ]$p1)
-      nCausal <- apply(itv, 1, function(x) {gnames <- anno[x[1] < anno[, "p0"] & anno[, "p0"] < x[2], "name"]; sum(ifelse(gnames %in% cau, 1, 0))})
-      itv <- data.frame(chrom, itv, nCausal, stringsAsFactors = F)
-      regions <- rbind(regions, itv)
-    }
-    colnames(regions) <- c("chrom", "p0", "p1", "nCausal")
-    regions.f <- regions
-
-  } else if (filtertype %in% c("mrash2s", "mrash")){
-
-    gres <- paste0(filterfile, ".expr.txt")
-    pipres <- read.table(gres, header =T)
-    if (filtertype == "mrash2s"){
-      sres <- paste0(filterfile, ".snp.txt")
-      pipres <- rbind(read.table(gres, header =T), read.table(sres, header =T))
-    }
-    for (chrom in chroms){
-      pipres.chr <- pipres[pipres$chr == chrom, ]
-      itv <- cbind(reg[reg$chrom == chrom, ]$p0, reg[reg$chrom == chrom, ]$p1)
-      rPIP <- apply(itv, 1, function(x) sum(pipres.chr[x[1] < pipres.chr$p0 & pipres.chr$p0 < x[2], "PIP"]))
[…]

### 2021-01-17 code/workflow/prep_ukbiobank_n40000_chr10_genotype.sh

@@ -1,1 +1,0 @@
-cd /home/szhao1/uk-biobank/uk-biobank-genotypes; /home/szhao1/uk-biobank/uk-biobank-phenotypes/tools/plink2 --sample /home/szhao1/uk-biobank/uk-biobank-phenotypes/data/genotypes/ukb27386_imp_v3_s487324.sample --keep /home/szhao1/causal-TWAS/data/ukbiobank_samples40000.txt --bgen ukb_imp_chr10_v3.bgen --chr 10 --threads 19 --make-pgen --out ~/causal-TWAS/data/ukb_chr10_s40000 --maf 0.05 --geno 0.05 --memory 30000

### 2021-01-17 code/workflow/prep_ukbiobank_n40000_chr15_genotype.sh

@@ -1,1 +1,0 @@
-cd /home/szhao1/uk-biobank/uk-biobank-genotypes; /home/szhao1/uk-biobank/uk-biobank-phenotypes/tools/plink2 --sample /home/szhao1/uk-biobank/uk-biobank-phenotypes/data/genotypes/ukb27386_imp_v3_s487324.sample --keep /home/szhao1/causal-TWAS/data/ukbiobank_samples40000.txt --bgen ukb_imp_chr15_v3.bgen --chr 15 --threads 19 --make-pgen --out ~/causal-TWAS/data/ukb_chr15_s40000 --maf 0.05 --geno 0.05 --memory 30000

### 2021-01-17 code/workflow/prep_ukbiobank_n40000_chr16_genotype.sh

@@ -1,1 +1,0 @@
-cd /home/szhao1/uk-biobank/uk-biobank-genotypes; /home/szhao1/uk-biobank/uk-biobank-phenotypes/tools/plink2 --sample /home/szhao1/uk-biobank/uk-biobank-phenotypes/data/genotypes/ukb27386_imp_v3_s487324.sample --keep /home/szhao1/causal-TWAS/data/ukbiobank_samples40000.txt --bgen ukb_imp_chr16_v3.bgen --chr 16 --threads 19 --make-pgen --out ~/causal-TWAS/data/ukb_chr16_s40000 --maf 0.05 --geno 0.05 --memory 30000

### 2021-01-17 code/workflow/prep_ukbiobank_n40000_chr6_genotype.sh

@@ -1,1 +1,0 @@
-cd /home/szhao1/uk-biobank/uk-biobank-genotypes; /home/szhao1/uk-biobank/uk-biobank-phenotypes/tools/plink2 --sample /home/szhao1/uk-biobank/uk-biobank-phenotypes/data/genotypes/ukb27386_imp_v3_s487324.sample --keep /home/szhao1/causal-TWAS/data/ukbiobank_samples40000.txt --bgen ukb_imp_chr6_v3.bgen --chr 6 --threads 19 --make-pgen --out ~/causal-TWAS/data/ukb_chr6_s40000 --maf 0.05 --geno 0.05 --memory 30000

### 2021-01-17 code/workflow/qsub_prep_ukbiobank_n40000_chr4_genotype.sh

@@ -1,10 +1,0 @@
-
-#!/bin/bash
-
-#PBS -N bgen2plink
-#PBS -S /bin/bash
-#PBS -l mem=64gb
-#PBS -l walltime=48:00:00
-#PBS -l nodes=1:ppn=20
-
-cd /home/szhao1/causal-TWAS/data; bash prep_ukbiobank_n40000_chr4_genotype.sh

### 2021-01-17 code/workflow/qsub_prep_ukbiobank_n40000_chr7_genotype.sh

@@ -1,10 +1,0 @@
-
-#!/bin/bash
-
-#PBS -N bgen2plink
-#PBS -S /bin/bash
-#PBS -l mem=64gb
-#PBS -l walltime=48:00:00
-#PBS -l nodes=1:ppn=20
-
-cd /home/szhao1/causal-TWAS/data; bash prep_ukbiobank_n40000_chr7_genotype.sh

### 2021-01-17 code/workflow/qsub_prep_ukbiobank_n40000_chr8_genotype.sh

@@ -1,10 +1,0 @@
-
-#!/bin/bash
-
-#PBS -N bgen2plink
-#PBS -S /bin/bash
-#PBS -l mem=64gb
-#PBS -l walltime=48:00:00
-#PBS -l nodes=1:ppn=20
-
-cd /home/szhao1/causal-TWAS/data; bash prep_ukbiobank_n40000_chr8_genotype.sh

### 2021-01-17 code/workflow/s40000-chr1to22.sh

@@ -1,22 +1,0 @@
-qsub qsub_prep_ukbiobank_n40000_chr1_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr2_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr3_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr4_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr5_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr6_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr7_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr8_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr9_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr10_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr11_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr12_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr13_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr14_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr15_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr16_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr17_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr18_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr19_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr20_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr21_genotype.sh
-qsub qsub_prep_ukbiobank_n40000_chr22_genotype.sh

### 2021-01-19 analysis/simulation-susieI-ukbWG-gtex.adipose_s80.45.Rmd
faster gwas

--- 2021-01-17 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
@@
 ---
-title: "Test susieI using whole genome, s40.22 samples"
+title: "Test susieI using whole genome, s80.45 samples"
 output: html_document
@@
 ```{r parameters}
-n.ori <- 40000 # number of samples
-n <-  22542
-p <- 656321 # number of SNPs
+n.ori <- 80000 # number of samples
+n <-  45542
+p <- 6656321 # number of SNPs
 J <- 8021 # number of genes
@@
 We use SNP genotype data from chr 1 to chr 22 combined from UKB.
-SNPs are downsampled to 1/10 (randomly), eQTLs (see below for definition of eQTL) were added back.
-This ends up with p = `r as.character(p)` SNPs.
+There total = `r as.character(p)` SNPs.
 
@@
 
-Next, the analysis is done at the “region” level, which is 500kb bins along the genome.
-We also used LDetect to define regions.
-We are exploring several ways to select regions that contain true signals, e.g.
-based on regional sum of mr.ash PIP for genes/SNPs, region smallest TWAS p value for gene/SNPs, or regional bayes factors, etc.
+Next, the analysis is done at the “region” level. We used LDetect to define regions.
 
-4. Run susie iteratively
-We then run susie for each of these regions.
-So the features of SuSiE are: SNPs and "genes" (not cis-eQTLs of that gene).
-We use the same prior for all SNPs and another prior for all "genes" when running SUSIE.
-In some settings, we also run SUSIE with null weight, which is calculated as `1- prior.SNP * n.SNP - prior.gene * n.gene`.
-We obtain the PIP for SNPs and gene in the region.
-After we run susie for all regions (one iteration), we take the average of all SNP PIPs as the prior of SNPs for the next iteration and similarly for the prior for genes.
-
-5. We obtain PIP for genes from the last iteration as results.
+4. Run ctwas_rss
+We downsampled to 1/10 of SNPs to run ctwas_rss
 
 ```{r dirs}
-simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
-outputdir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
-susiedir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
-simtag <- '20201001-1-3'
-source('~/causalTWAS/causal-TWAS/code/gwas.R')
-source('~/causalTWAS/causal-TWAS/code/ld.R')
+library(ctwas)
+library(data.table)
 source('~/causalTWAS/causal-TWAS/code/qqplot.R')
+outputdir <- "~/causalTWAS/simulations/simulation_ctwas_rss_20210117/"
+runtag <- 'ukb-s80.45-adi'
 ```
-
-```{r causal}
-exprgwasf <- paste0(simdatadir, simtag, ".exprgwas.txt.gz")
-load(paste0(simdatadir, "simu_", simtag, "-pheno.Rd"))
-caulist <- list()
-for (chrom in 1:22) {
-  load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom ,"_s40.22.FBM.Rd"))
-  load(paste0(simdatadir, "simu_s40000_GTEXadipose-B", chrom, "-cis-expr.Rd"))
-  caulist[[chrom]]<- c(exprres$gnames[phenores$batch[[chrom]]$param$idx.cgene], dat$snp[phenores$batch[[chrom]]$param$idx.cSNP,])
-}
-cau <- unlist(caulist)
-```
-# Power estimation
-
-We use gene.pve ~ 0.1, snp.pve ~ 0.5.
-
-* For SNPs, we use $\pi_1 = 2.5e-3$, variance for effect size ~ $0.03^2$, power at 5e-8 p value cutoff:
-```{r pow func} 
-pow <- function(total, n, beta, cutp){
-  rec <- rep(0, total)
-  for (i in 1:total){
-    x <- rnorm(n)
-    y <- x * rnorm(1, sd = beta) + rnorm(n, sd = sqrt(2.5))
-    lm.s <- lm(y~x)
-    pv <- summary(lm.s)$coefficients[2,4]
-    rec[i] <- pv
-  }
-  length(rec[rec < cutp])/length(rec)
-}
-```
-
-```{r snp}
-load("data/power_s40.22.Rd")
-total <- 1e3
-n <- 22542
-#p1 <- pow(total, n, 0.0276, 5e-8)
-print(p1)
-```
-
-* For genes, under low power setting, $\pi_1 = 0.05$, variance for effect size ~ $0.025^2$, power at 1e-5 cutoff:
-
-```{r gene1}
-#p2 <- pow(total, n, 0.025, 1e-5)
-print(p2)
-```
-
-For genes, under high power setting, $\pi_1 = 0.02$, variance for effect size ~ $0.045^2$, power at 1e-5 cutoff:
-
-```{r gene2}
-#p3 <- pow(total, n, 0.045, 1e-5)
-print(p3)
-#save(p1,p2,p3, file = "data/power_s40.22.Rd")
-```
-
-# p value distribution
-
-* TWAS p value of genes:
-
-```{r gene twas, fig.height=8, fig.width=8}
-chrom <- 1
-a <- read.table(exprgwasf, header = T)
-a$ifcausal <- ifelse(a$MARKER_ID %in% cau, 1, 0)
-ax <- pretty(0:max(-log10(a$PVALUE)), n = 30)
-
-par(mfrow=c(3,1))
-h1 <- hist(-log10(a$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey", xlim= c(3,20), ylim =c(0,50)); grid()
-h2 <- hist(-log10(a[a$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon", xlim= c(3,20), ylim =c(0,50));grid()
-
-cat("number of genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5, ]))
-cat("number of causal genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5 & a$ifcausal ==1, ]))
-
-plot(a[a$X.CHROM ==chrom, ]$BEGIN, -log10(a[a$X.CHROM ==chrom, ]$PVALUE), col = a[a$X.CHROM ==chrom, ]$ifcausal + 1, xlab = paste0("chr", chrom), ylab = "-log10(pvalue)")
-points(a[a$X.CHROM ==chrom & a$ifcausal ==1, ]$BEGIN, -log10(a[a$X.CHROM ==chrom & a$ifcausal ==1, ]$PVALUE), col = "red", pch =19)
-grid()
-```
-
-* qq plot for genes
-```{r qqgene, fig.height= 5, fig.width=5}
-gg_qqplot(a$PVALUE) +
-  theme_bw(base_size = 24) +
-  theme(
-    axis.ticks = element_line(size = 0.5),
-    panel.grid = element_blank()
-    # panel.grid = element_line(size = 0.5, color = "grey80")
-  )
-```
-
-*  p value of SNPs (GWAS):
-
-```{r snp twas, fig.height=8, fig.width=8}
-snpgwasf <- paste0(simdatadir, simtag, ".snpgwas.txt.gz")
-
-b <- fread(snpgwasf, header = T)
-b$ifcausal <- ifelse(b$MARKER_ID %in% cau, 1, 0)
-ax <- pretty(0:max(-log10(b$PVALUE)), n = 30)
-
-par(mfrow=c(3,1))
-h1 <- hist(-log10(b$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey", xlim= c(3,20), ylim =c(0,100)); grid()
-h2 <- hist(-log10(b[b$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon", xlim= c(3,20), ylim =c(0,100));grid()
-
-cat("number of SNPs < 5e-8: ", nrow(b[b$PVALUE < 5e-8,]))
-
-plot(b[b$X.CHROM ==chrom, ]$BEGIN, -log10(b[b$X.CHROM ==chrom, ]$PVALUE), col = b[b$X.CHROM ==chrom,]$ifcausal + 1, xlab = paste0("chr", chrom), ylab = "-log10(pvalue)")
-points(b[b$X.CHROM ==chrom & b$ifcausal ==1, ]$BEGIN, -log10(b[b$X.CHROM ==chrom & b$ifcausal ==1, ]$PVALUE), col = "red", pch =19)
-grid()
-```
-
-* qq plot for SNPs
-```{r qqsnp, fig.height= 5, fig.width=5}
-gg_qqplot(b$PVALUE) +
-  theme_bw(base_size = 24) +
-  theme(
-    axis.ticks = element_line(size = 0.5),
-    panel.grid = element_blank()
-    # panel.grid = element_line(size = 0.5, color = "grey80")
-  )
-```
-# Block features
-
-```{r block}
-
-a <- read.table(paste0(outputdir, "20201001-1-1.config9.gene.nofilter.r.txt"), header =T)
-b <-  read.table(paste0(outputdir, "20201001-2-1.config9.gene.nofilter.r.txt"), header =T)
-m1 <- read.table(paste0(outputdir, "20201001-1-1.config9.nofilter.r.txt"), header =T)
-m2 <- read.table(paste0(outputdir, "20201001-2-1.config9.nofilter.r.txt"), header =T)
-
-hist(a$p1-a$p0, main = NULL, xlab = "Block size(bp)", col = "salmon")
-
-par(mfrow=c(2,2))
-hist(a$nCausal, xlab = "No.causal genes", col = "salmon", breaks=100, xlim=c(0,15), main ="low power")
-hist(b$nCausal, xlab = "No.causal genes", col = "salmon", breaks=100, xlim=c(0,15), main ="high power")
-hist(m1$nCausal, xlab = "No.causal genes + SNPs", col = "salmon", breaks=100, xlim=c(0,15), main ="low power")
-hist(m2$nCausal, xlab = "No.causal genes + SNPs", col = "salmon", breaks=100, xlim=c(0,15), main ="high power")
-```
-
-# TWAS FDP
-
-```{r PIPplot functions}
-
-library(ggplot2)
-library(cowplot)
-library(plotrix)
-
-.obn <- function(pips, ifcausal, mode = c("PIP", "FDR")){
-  a_bin <- cut(pips, breaks= seq(0, 1, by=0.1))
-  if (mode == "PIP") {
-    ob = c(by(ifcausal, a_bin, FUN = sum))
-  } else if (mode == "FDR"){
-    ob = c(by((1-ifcausal), a_bin, FUN = sum))
-  }
-  return(ob)
-}
-
-nca_plot <- function(pips, ifcausal, runtag = NULL, mode = c("PIP", "FDR"), main = mode[1], ...){
-
-  # ifcausal:0,1, runtag: for adding std.
-  if (is.null(runtag)){
-     se = 0 
-  } else{
-    dflist <- list()
-    for (rt in unique(runtag)){
-      pips.rt <- pips[runtag == rt]
-      ifcausal.rt <- ifcausal[runtag == rt]
-      dflist[[rt]] <- .obn(pips.rt, ifcausal.rt, mode = mode)
-    }
-    nca_mean <- colMeans(do.call(rbind, dflist))
-    se <- apply(do.call(rbind, dflist), 2, plotrix::std.error)
-  }
-
-  df <- data.frame("ncausal" = nca_mean, "se" = se)
-
-  fig <- ggplot(df) +
-    geom_bar( aes(x=seq(0, 1, by=0.1)[1:10]+0.05, y=ncausal), color ="black", stat="identity", fill="salmon", alpha=0.7, width = 0.1) +
-    geom_errorbar( aes(x= seq(0, 1, by=0.1)[1:10] + 0.05, ymin= ncausal-se, ymax=ncausal+se), width=0.05, colour="black", alpha=0.9, size=0.5) +
-    ggtitle(main) +
-    xlab("PIP") + ylab("No. causal genes")
-    theme_cowplot()
-  return(fig)
-}
-
-ncausal_plot <- function(pipfs, format = "susie", main = "PIP"){
-  df <- NULL
-  for (i in 1:length(pipfs)) {
-    res <- fread(pipfs[i], header = T)
-    res <- data.frame(res[res$type  =="gene", ])
-    res$runtag <- i
-    if (format == "susie"){
-       res <- rename(res, c("susie_pip" = "pip") )
-    } else if (format == "SER"){
-       res <- rename(res, c("SERpip" = "pip"), )
-    }
-    res <- res[complete.cases(res),]
-    df <- rbind(df, res)
-  }
-
-  fig <- nca_plot(df$pip, df$ifcausal, df$runtag, mode ="PIP", main = main)
-  return(fig)
-}
-
-.exob <- function(pips, ifcausal, mode = c("PIP", "FDR")){
-  a_bin <- cut(pips, breaks= seq(0, 1, by=0.1))
-  if (mode == "PIP") {
-    ex = c(by(pips, a_bin, FUN = mean))
-    ob = c(by(ifcausal, a_bin, FUN = mean))
-  } else if (mode == "FDR"){
-    ex = c(by(pips, a_bin, FUN = mean))
-    ob = 1 - c(by(ifcausal, a_bin, FUN = mean))
-  }
-  return(list("expected" = ex, "observed" = ob))
-}
-
-dot_plot = function(df, main) {
-        ggplot(df, aes(x=mean_pip, y=observed_freq)) + 
-          geom_errorbar(aes(ymin=observed_freq-se, ymax=observed_freq+se), colour="black", size = 0.5, width=.01) +
-          geom_point(size=1.5, shape=21, fill="#002b36") + # 21 is filled circle
-          xlab("Expected") +
-          ylab("Observed") +
-          coord_cartesian(ylim=c(0,1), xlim=c(0,1)) +
-          geom_abline(slope=1,intercept=0,colour='red', size=0.2) +
-          ggtitle(main) +
-          expand_limits(y=0) +                        # Expand y range
-          theme_cowplot()}
-
-cp_plot <- function(pips, ifcausal, runtag = NULL, mode = c("PIP", "FDR"), main = mode[1], ...){
-
-  # ifcausal:0,1, runtag: for adding std.
-  if (is.null(runtag)){
-     se = 0 
-  } else{
-    dflist <- list()
-    for (rt in unique(runtag)){
-      pips.rt <- pips[runtag == rt]
-      ifcausal.rt <- ifcausal[runtag == rt]
-      dflist[[rt]] <- .exob(pips.rt, ifcausal.rt, mode = mode)
-    }
-    mean_pip <- colMeans(do.call(rbind, lapply(dflist, '[[', "expected")))
-    observed_freq <- colMeans(do.call(rbind, lapply(dflist, '[[', "observed")))
-    se <- apply(do.call(rbind, lapply(dflist, '[[', "observed")), 2, plotrix::std.error)
-  }
-
-  df <- data.frame("mean_pip" = mean_pip, "observed_freq"= observed_freq, "se" = se)
-  dot_plot(df, main = main)
-
-  #plot(Expected, Observed, xlim= c(0,1), ylim=c(0,1), pch =19, main = main, ...)
-  #lines(x = c(0,1), y = c(0,1), col ="grey", lty = 2)
-}
-
-caliPIP_plot <- function(pipfs, format = "susie", main = "PIP"){
-  df <- NULL
-  for (i in 1:length(pipfs)) {
-    res <- fread(pipfs[i], header = T)
-    res <- data.frame(res[res$type  =="gene", ])
-    res$runtag <- i
-    if (format == "susie"){
-       res <- rename(res, c("susie_pip" = "pip"))
-    } else if (format == "SER"){
-       res <- rename(res, c("SERpip" = "pip"))
-    }
-    res <- res[complete.cases(res),]
-    df <- rbind(df, res)
-  }
-
-  fig <- cp_plot(df$pip, df$ifcausal, df$runtag, mode ="PIP", main = main)
-  return(fig)
-}
-
-# pipfs is just for ifcausal info
-caliFDP_plot <- function(gwasfs, pipfs,  format= "susie",  main = "FDP"){
-  df <- NULL
-  for (i in 1:length(pipfs)) {
-    pipres <- fread(pipfs[i], header = T)
-    pipres <- data.frame(pipres[pipres$type  =="gene", ])
-    pipres$runtag <- i
-    if (format == "susie"){
-       pipres <- rename(pipres, c("susie_pip" = "pip"))
-    } else if (format == "SER"){
-       pipres <- rename(pipres, c("SERpip" = "pip"))
-    }
-    gwasres <- read.table(gwasfs[i], header = T)
-    gwasres <- rename(gwasres, c("MARKER_ID" = "name"))
-    gwasres$FDR <- p.adjust(gwasres$PVALUE, method = "fdr")
-    res <- merge(gwasres, pipres, by = "name", all = T)
-    res <- res[complete.cases(res),]
-    df <- rbind(df, res)
-  }
-  fig <- cp_plot(df$FDR, df$ifcausal, df$runtag, mode ="FDR", main = main)
-  cat("FDP at bonferroni corrected p = 0.05: ", 1 - mean(df[df$PVALUE < 0.05 /dim(df)[1], "ifcausal"]))
-  return(fig)
-}
-
-scatter_plot_PIP_p <- function(pipfs, gwasfs, pipformat = "susie", main ="PIP-p"){
-  df <- NULL
-  for (i in 1:length(pipfs)) {
-    pipres <- fread(pipfs[i], header = T)
-    pipres <- data.frame(pipres[pipres$type  =="gene", ])
-    pipres$runtag <- i
-    if (pipformat == "susie"){
-       pipres <- rename(pipres, c("susie_pip" = "pip"))
-    } else if (pipformat == "SER"){
-       pipres <- rename(pipres, c("SERpip" = "pip"))
-    }
-    gwasres <- read.table(gwasfs[i], header = T)
-    gwasres <- rename(gwasres, c("MARKER_ID" = "name"))
-    res <- merge(gwasres, pipres, by = "name", all = T)
-    res <- res[complete.cases(res),]
-    df <- rbind(df, res)
-  }
-
-  df <- rename(df, c( "PVALUE" = "TWAS.p"))
-  df[,"TWAS.p"] <- -log10(df[, "TWAS.p"])
-
-  df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1),to=c("darkgreen", "salmon"))
-  plot(df$TWAS.p, df$pip, col = df$ifcausal, main = main, xlab = "-log10(TWAS p value)", ylab = "PIP")
-
-  # df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1), to=c("Non causal", "Causal"))
-  # fig <- plot_ly(data = df, x = ~ TWAS.p, y = ~ pip, color = ~ ifcausal,
-  #                colors = c( "salmon", "darkgreen"), type ="scatter", text = ~ paste("Name: ", paste0(runtag,":",name),
-  #                                                                  "\nChr: ", chr,  "\nPos:", pos))
-
-}
-```
-
-```{r TWAS FDP, fig.height=4, fig.width=8}
-tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
-gwasfs <-  paste0(simdatadir, "20201001-", tags, ".exprgwas.txt.gz")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")
-
-f1 <- caliFDP_plot(gwasfs[1:5], susieIfs2[1:5], format= "susie", main = "TWAS FDP (low power)")
-f2 <- caliFDP_plot(gwasfs[6:10], susieIfs2[6:10], format= "susie", main ="TWAS FDP (high power)")
-
-gridExtra::grid.arrange(f1, f2, ncol =2)
-```
-
-# L's effect on SuSiE
-
-We run SuSiE on each block with true priors and true prior variances. We tested the difference between using `L=1` and `L=5`. We only did this for the low power setting.
-
-```{r reg functions}
-ncs_ncausal <- function(susief, main = "nCSvs.nCausal") {
-  dt <- fread(susief, header = T)
-  ncausal <- dt[, sum(ifcausal), by=list(b,rn)]
-  nCS <- dt[, max(cs_index), by=list(b,rn)]
-  plot(jitter(ncausal$V1), jitter(nCS$V1), xlab= "No.causal/block", ylab = "No. credible set/block", main= main)
-  return(dt)
-}
-
-sumPIP_ncausal <- function(susief, main = "sumPIPvs.nCausal"){
-  dt <- fread(susief, header = T)
-  ncausal <- dt[, sum(ifcausal), by=list(b,rn)]
-  sumpip <- dt[, sum(susie_pip), by=list(b,rn)]
-  plot(jitter(ncausal$V1), sumpip$V1, xlab= "No.causal/block", ylab = "Regional sum PIP", main = main)
-  return(dt)
-}
-```
-## L=1
-
-Note: the data has been added some noise (both X and Y axes for credible set vs. no.causal, X axis for sum of PIP vs. no.causal) for visualization purposes. 
-```{r L=1, fig.width=8, fig.height=4}
-tags <- paste(rep(1, each = 5), 1:5, sep = "-")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config11.susieI.txt")
-f1 <- caliPIP_plot(susieIfs2, main = "L=1")
-f2 <- ncausal_plot(susieIfs2, format = "susie", main = "No. Causal Genes")
-gridExtra::grid.arrange(f1, f2, ncol =2)
-
-par(mfrow=c(1,2))
-dt <- sumPIP_ncausal(susieIfs2[2], main = "L=1")
-dt <-ncs_ncausal(susieIfs2[2], main = "L=1, CS=0.95")
-```
-
-* True parameters: gene $\pi_1$, `r g <- nrow(dt[type =="gene" & ifcausal !=0])/ nrow(dt[type =="gene"]);g`, SNP $\pi_1$, `r s <-nrow(dt[type =="SNP" & ifcausal !=0])/ nrow(dt[type =="SNP"]);s`.
-[enrichment = `r g/s`].
-* Average PIP: gene `r g <- mean(dt[type =="gene", susie_pip]);g`, SNP `r s <-mean(dt[type =="SNP", susie_pip]);s` . [enrichment = `r g/s`].
-* Proportion in credible set (0.95): gene `r g <-nrow(dt[type =="gene" & cs_index !=0])/ nrow(dt[type =="gene"]);g `, SNP `r s <-nrow(dt[type =="SNP" & cs_index !=0])/ nrow(dt[type =="SNP"]);s`.
-[enrichment = `r g/s`].
-
-## L=5
-
-Note: the data has been added some noise (both X and Y axes for credible set vs. no.causal, X axis for sum of PIP vs. no.causal) for visualization purposes.
-
-```{r L=5, fig.width=8, fig.height=4}
-tags <- paste(rep(1, each = 5), 1:5, sep = "-")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config12.susieI.txt")
-
-f1 <- caliPIP_plot(susieIfs2, main = "L=5")
-f2 <- ncausal_plot(susieIfs2, format = "susie", main = "No. Causal Genes")
-gridExtra::grid.arrange(f1, f2, ncol =2)
-
-par(mfrow=c(1,3))
-dt <- sumPIP_ncausal(susieIfs2[2], main = "L=5")
-dt <- ncs_ncausal(susieIfs2[2], main = "L=5, CS=0.95")
-
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config13.susieI.txt")
-dt2 <- ncs_ncausal(susieIfs2[2], main = "L=5, CS=0.8")
-```
-
-* True parameters: gene $\pi_1$, `r g <- nrow(dt[type =="gene" & ifcausal !=0])/ nrow(dt[type =="gene"]);g`, SNP $\pi_1$, `r s<-nrow(dt[type =="SNP" & ifcausal !=0])/ nrow(dt[type =="SNP"]);s`.
-[enrichment = `r g/s`].
-* Average PIP: gene `r g <- mean(dt[type =="gene", susie_pip]);g`, SNP `r s <-mean(dt[type =="SNP", susie_pip]);s` . [enrichment = `r g/s`].
-* Proportion in credible set (0.95): gene `r g <-nrow(dt[type =="gene" & cs_index !=0])/ nrow(dt[type =="gene"]);g `, SNP `r s <-nrow(dt[type =="SNP" & cs_index !=0])/ nrow(dt[type =="SNP"]);s`.
-[enrichment = `r g/s`].
-* Proportion in credible set (0.8): gene `r g <-nrow(dt2[type =="gene" & cs_index !=0])/ nrow(dt2[type =="gene"]);g `, SNP `r s <-nrow(dt2[type =="SNP" & cs_index !=0])/ nrow(dt2[type =="SNP"]);s`.
-[enrichment = `r g/s`].
 
@@
 ```{r param functions}
-show_param <- function(phenofs, susieIfs, susieIfs2){
+show_param <- function(phenofs, susieIfs, susieIfs2, thin = 1){
   pars <- do.call(rbind, lapply(phenofs, function(x) {load(x); 
@@
       phenores$param$pve.snp.truth,
-      length(phenores$batch[[1]]$param$idx.cgene)/phenores$batch[[1]]$param$J,
-      length(phenores$batch[[1]]$param$idx.cSNP)/phenores$batch[[1]]$param$M)}))
+      length(phenores$batch[[1]]$idx.cgene)/phenores$batch[[1]]$J,
+      length(phenores$batch[[1]]$idx.cSNP)/phenores$batch[[1]]$M)}))
 
@@
 
-  param.s <- do.call(rbind, lapply(susieIfs, function(x) {load(x); c(tail(prior.gene_rec[prior.gene_rec!=0], 1), tail(prior.SNP_rec[prior.SNP_rec!=0],1))}))
+  param.s <- do.call(rbind, lapply(susieIfs, function(x) {load(x); c(tail(group_prior_rec[1,][!is.na(group_prior_rec[1,])], 1), tail(group_prior_rec[2,][!is.na(group_prior_rec[1,])], 1) * thin)}) )
 
@@
 
+  df[,"pi1.SNP_susietruth"] <- 1
   st <-  cbind(df[,"pi1.gene_susietruth"]/df[,"pi1.SNP_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
@@
 
-gpip_dist <- function(susiefs, ...){
-  dflist <- list()
-  for (f in susiefs){
-    dflist[[f]] <- read.table(f, header =T , stringsAsFactors = F)
-  }
-  df <- do.call(rbind, dflist)
-  hist(df[df$type == "gene", "susie_pip"], xlab = "gene susie PIP", 
-       breaks = 50, ylim = c(0,20), xlim=c(0,1), col = "salmon", ...)
-}
 ```
 
-## susieI (1)
-* Regions: all regions, 500kb uniform regions.
-* Susie run parameters: `L=1`.
-We initialize with prior for genes and SNPs as uniform.
-gene.pve ~ 0.1, snp.pve ~ 0.5.
-Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
-Prior variance and residual variance were calculated by SUSIE for each region.
+## ctwas_rss (1)
+* run parameters: downsample 1/10.
 
@@
 tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
-phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-gwasfs <-  paste0(simdatadir, "20201001-", tags, ".exprgwas.txt.gz")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")
+phenofs <- paste0(outputdir, runtag, "_simu", tags, "-pheno.Rd")
+susieIfs <- paste0(outputdir, runtag, "_simu", tags, "_config1.s2.susieIrssres.Rd")
+susieIfs2 <- paste0(outputdir, runtag, "_simu",tags, "_config1.s2.susieIrss.txt")
 
-df <- show_param(phenofs, susieIfs, susieIfs2)
+df <- show_param(phenofs, susieIfs, susieIfs2, thin = 0.1)
 par(mfrow = c(2,3))
@@
 ```
-
-```{r pip1susie, fig.height=6, fig.width=8}
-f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)" )
-f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
-f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
-f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
-gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
-```
-
-## susieI (2)
-* Regions: all regions, LD-defined regions.
-* Susie run parameters: `L=1`.
-We initialize with prior for genes and SNPs as uniform.
-gene.pve ~ 0.1, snp.pve ~ 0.5.
-Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
-Prior variance and residual variance were calculated by SUSIE for each region.
-
-```{r param6susie, fig.height=6, fig.width=8}
-tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
-phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config6.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config6.susieI.txt")
-
-df <- show_param(phenofs, susieIfs, susieIfs2)
-par(mfrow = c(2,3))
-plot_param(df[1:5,], main = "low power")
-plot_param(df[6:10,], main = "high power")
-```
-
-```{r pip6susie, fig.height=6, fig.width=8}
-f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)" )
-f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
-f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
-f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
-gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
-```
-
-## susieI (3)
-* Regions: all regions, LD-defined regions.
-* Susie run parameters: `L=10`.
-We initialize with prior for genes and SNPs as uniform.
-gene.pve ~ 0.1, snp.pve ~ 0.5.
-Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
-Prior variance and residual variance were calculated by SUSIE for each region.
-
-```{r param8susie, fig.height=6, fig.width=8}
-tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
-phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config8.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config8.susieI.txt")
-
-df <- show_param(phenofs, susieIfs, susieIfs2)
-par(mfrow = c(2,3))
-plot_param(df[1:5,], main = "low power")
-plot_param(df[6:10,], main = "high power")
-```
-
-```{r pip8susie, fig.height=6, fig.width=8}
-f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)" )
-f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
-f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
-f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
-gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
-```
-
-## susieI (4)
-* Regions: all regions, LD-defined regions.
-* Susie run parameters: `L=1`.
-We use true prior variance and plug into SUSIE, so this should be the same as SER/EM (2).
-We initialize with prior for genes and SNPs as uniform.
-gene.pve ~ 0.1, snp.pve ~ 0.5.
-Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
-
-```{r param9susie, fig.height=6, fig.width=8}
-tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
-phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config9.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config9.susieI.txt")
-
-df <- show_param(phenofs, susieIfs, susieIfs2)
-par(mfrow = c(2,3))
-plot_param(df[1:5,], main = "low power")
-plot_param(df[6:10,], main = "high power")
-```
-
-```{r pip9susie, fig.height=6, fig.width=8}
-f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)" )
-f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
-f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
-f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
-gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
-```
-
-## susieI (5)
-* Regions: all regions, LD-defined regions.
-* Susie run parameters: `L=5`.
-We use true prior variance and plug into SUSIE.
-We initialize with prior for genes and SNPs as uniform.
-gene.pve ~ 0.1, snp.pve ~ 0.5.
-Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
-
-```{r param10susie, fig.height=6, fig.width=8}
-tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
-phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config10.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config10.susieI.txt")
-
-df <- show_param(phenofs, susieIfs, susieIfs2)
-par(mfrow = c(2,3))
-plot_param(df[1:5,], main = "low power")
-plot_param(df[6:10,], main = "high power")
-```
-
-```{r pip10susie, fig.height=6, fig.width=8}
-f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)" )
-f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
-f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
-f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high poweunr)")
-gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
-```
-
-## Select block + susieI (1)
-* Regions: all regions 10 iterations.
-Then filter out regions with probability of having two or more effects > 0.2 to estimate paramters 20 iterations.
-then all regions to get all PIP.
-LD-defined regions.
-* Susie run parameters: `L=1`.
-We use true prior variance and plug into SUSIE.
-We initialize with prior for genes and SNPs as uniform.
-gene.pve ~ 0.1, snp.pve ~ 0.5.
-We use null weight when run susie.
-We run 10 iterations.
-Then filter out regions with probability of having two or more effects.
-Then rerun on selected regions to estimate parameters.
-Lastly, run on all regions with `L=1` to get PIP.
-
-```{r param1susiefl, fig.height=6, fig.width=8}
-tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
-phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config14.fl.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config14.fl.susieI.txt")
-
-df <- show_param(phenofs, susieIfs, susieIfs2)
-par(mfrow = c(2,3))
-plot_param(df[1:5,], main = "low power")
-plot_param(df[6:10,], main = "high power")
-
-a <- fread(susieIfs2[1], header =T)
-cat("No.blocks selected for parameter estimation (low power):", nrow(unique(a[, c("b", "rn")])), "\n")
-a <- fread(susieIfs2[6], header =T)
-cat("No.blocks selected for parameter estimation (high power):", nrow(unique(a[, c("b", "rn")])), "\n")
-```
-
-## Select block + susieI (2)
-* Regions: all regions 10 iterations.
-Then filter out regions with probability of having two or more effects > 0.1 to estimate paramters 20 iterations.
-then all regions to get all PIP.
-LD-defined regions.
-* Susie run parameters: `L=1`.
-We use true prior variance and plug into SUSIE.
-We initialize with prior for genes and SNPs as uniform.
-gene.pve ~ 0.1, snp.pve ~ 0.5.
-We use null weight when run susie.
-We run 10 iterations.
-Then filter out regions with probability of having two or more effects.
-Then rerun on selected regions to estimate parameters.
-Lastly, run on all regions with `L=1` to get PIP.
-
-```{r param2susiefl, fig.height=6, fig.width=8}
-tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
-phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config15.fl.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config15.fl.susieI.txt")
-
-df <- show_param(phenofs, susieIfs, susieIfs2)
-par(mfrow = c(2,3))
-plot_param(df[1:5,], main = "low power")
-plot_param(df[6:10,], main = "high power")
-
-a <- fread(susieIfs2[1], header =T)
-cat("No.blocks selected for parameter estimation (low power):", nrow(unique(a[, c("b", "rn")])), "\n")
-a <- fread(susieIfs2[6], header =T)
-cat("No.blocks selected for parameter estimation (high power):", nrow(unique(a[, c("b", "rn")])), "\n")
-```
-
-## Select block + susieI (3)
-* Regions: all regions 5 iterations.
-Then filter out regions with probability of having two or more effects > 0.2 to estimate paramters 20 iterations.
-then all regions to get all PIP.
-LD-defined regions.
-* Susie run parameters: `L=1`.
-We estimate prior variance based on EM and plug into SUSIE.
-We initialize with prior for genes and SNPs as uniform.
-gene.pve ~ 0.1, snp.pve ~ 0.5.
-We use null weight when run susie.
-We run 5 iterations.
-Then filter out regions with probability of having two or more effects.
-Then rerun on selected regions to estimate parameters (20 iterations).
-Lastly, run on all regions with `L=5` to get PIP.
-
-```{r param3susiefl, fig.height=6, fig.width=8}
-tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
-phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config16.fl.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config16.fl.susieI.txt")
-
-df <- show_param(phenofs, susieIfs, susieIfs2)
-par(mfrow = c(2,3))
-plot_param(df[1:5,], main = "low power")
-plot_param(df[6:10,], main = "high power")
-
-a <- fread(susieIfs2[1], header =T)
-cat("No.blocks selected for parameter estimation (low power):", nrow(unique(a[, c("b", "rn")])), "\n")
-a <- fread(susieIfs2[6], header =T)
-cat("No.blocks selected for parameter estimation (high power):", nrow(unique(a[, c("b", "rn")])), "\n")
-```
-
-```{r pip3susiefl, fig.height=6, fig.width=8}
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config16.flrerun.susieI.txt")
-f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)")
-f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)")
-f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
-f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
-gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
-```
-
-```{r scatter3susiefl, fig.height=4, fig.width=8}
-par(mfrow=c(1,2))
-f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "susie", main = "low power")
-f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "susie", main = "high power")
-```
-
-## Select block + susieI (4)
-* Regions: all regions 5 iterations.
-Then filter out regions with probability of having two or more effects > 0.1 to estimate paramters 20 iterations.
-then all regions to get all PIP.
-LD-defined regions.
-* Susie run parameters: `L=1`.
-We estimate prior variance based on EM and plug into SUSIE.
-We initialize with prior for genes and SNPs as uniform.
-gene.pve ~ 0.1, snp.pve ~ 0.5.
-We use null weight when run susie.
-We run 5 iterations.
-Then filter out regions with probability of having two or more effects.
-Then rerun on selected regions to estimate parameters (20 iterations).
-Lastly, run on all regions with `L=5` to get PIP.
-
-```{r param4susiefl, fig.height=6, fig.width=8}
-tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
-phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config17.fl.susieIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config17.fl.susieI.txt")
-
-df <- show_param(phenofs, susieIfs, susieIfs2)
-par(mfrow = c(2,3))
-plot_param(df[1:5,], main = "low power")
-plot_param(df[6:10,], main = "high power")
-
-a <- fread(susieIfs2[1], header =T)
-cat("No.blocks selected for parameter estimation (low power):", nrow(unique(a[, c("b", "rn")])), "\n")
-a <- fread(susieIfs2[6], header =T)
-cat("No.blocks selected for parameter estimation (high power):", nrow(unique(a[, c("b", "rn")])), "\n")
-```
-
-```{r pip4susiefl, fig.height=6, fig.width=8}
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config17.flrerun.susieI.txt")
-f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)")
-f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)")
-f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
-f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
-gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
-```
-
-```{r scatter4susiefl, fig.height=4, fig.width=8}
-par(mfrow=c(1,2))
-f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "susie", main = "low power")
-f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "susie", main = "high power")
-```
-
-##  Single effect model/EM (1)
-* Regions: all regions, 500kb uniform regions.
-* Run parameters: We initialize with prior for genes and SNPs as uniform.
-gene.pve ~ 0.1, snp.pve ~ 0.5.
-Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
-We use the true prior variance for genes and SNPs.
-
-```{r param1, fig.height=6, fig.width=8}
-tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
-phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.SERIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.SERI.txt")
-
-df <- show_param(phenofs, susieIfs, susieIfs2)
-par(mfrow = c(2,3))
-plot_param(df[1:5,], main = "low power")
-plot_param(df[6:10,], main = "high power")
-```
-
-```{r pip1ser, fig.height=6, fig.width=8}
-f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP(low power)")
-f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)")
-f3 <- ncausal_plot(susieIfs2[1:5], format = "SER", main = "No. Causal Genes (low power)")
-f4 <- ncausal_plot(susieIfs2[6:10], format = "SER", main = "No. Causal Genes (high power)")
-gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
-```
-
-##  Single effect model/EM (2)
-* Regions: all regions, defined based on LDetect.
-* Run parameters: We initialize with prior for genes and SNPs as uniform.
-gene.pve ~ 0.1, snp.pve ~ 0.5.
-Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
-We use the true prior variance for genes and SNPs.
-
-```{r param6, fig.height=6, fig.width=8}
-tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
-phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config6.SERIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config6.SERI.txt")
-
-df <- show_param(phenofs, susieIfs, susieIfs2)
-par(mfrow = c(2,3))
-plot_param(df[1:5,], main = "low power")
-plot_param(df[6:10,], main = "high power")
-```
-
-```{r pip6ser, fig.height=6, fig.width=8}
-f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP(low power)")
-f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)")
-f3 <- ncausal_plot(susieIfs2[1:5], format = "SER", main = "No. Causal Genes (low power)")
-f4 <- ncausal_plot(susieIfs2[6:10], format = "SER", main = "No. Causal Genes (high power)")
-gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
-```
-
-```{r scatter6ser, fig.height=4, fig.width=8}
-par(mfrow=c(1,2))
-f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "SER", main = "low power")
-f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "SER", main = "high power")
-```
-
-##  Single effect model/EM (3)
-* Regions: regions with at most 1 causal signal, regions are defined based on LDetect.
-* Run parameters: We initialize with prior for genes and SNPs as uniform.
-gene.pve ~ 0.1, snp.pve ~ 0.5.
-Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).
-We use the true prior variance for genes and SNPs.
-
-```{r param7, fig.height=6, fig.width=8}
-tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
-phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config7.SERIres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config7.SERI.txt")
-
-df <- show_param(phenofs, susieIfs, susieIfs2)
-par(mfrow = c(2,3))
-plot_param(df[1:5,], main = "low power")
-plot_param(df[6:10,], main = "high power")
-```
-
-```{r pip7ser, fig.height=6, fig.width=8}
-f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP(low power)")
-f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)")
-f3 <- ncausal_plot(susieIfs2[1:5], format = "SER", main = "No. Causal Genes (low power)")
-f4 <- ncausal_plot(susieIfs2[6:10], format = "SER", main = "No. Causal Genes (high power)")
-gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
-```
-
-##  Summary stats version
-* The run setting is the same as  in "Select block + susieI (3)" except that we used summary statistics and in-sample LD as input.
-
-```{r dirsSS}
-simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
-outputdir <- "~/causalTWAS/simulations/simulation_susieI_rss_20201001/"
-susiedir <- "~/causalTWAS/simulations/simulation_susieI_rss_20201001/"
-```
-```{r param1susieflSS, fig.height=6, fig.width=8}
-tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
-phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
-susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.fl.susieIrssres.Rd")
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.fl.susieIrss.txt")
-
-df <- show_param(phenofs, susieIfs, susieIfs2)
-par(mfrow = c(2,3))
-plot_param(df[1:5,], main = "low power")
-plot_param(df[6:10,], main = "high power")
-
-a <- fread(susieIfs2[1], header =T)
-cat("No.blocks selected for parameter estimation (low power):", nrow(unique(a[, c("b", "rn")])), "\n")
-a <- fread(susieIfs2[6], header =T)
-cat("No.blocks selected for parameter estimation (high power):", nrow(unique(a[, c("b", "rn")])), "\n")
-```
-
-```{r pip1susieflSS, fig.height=6, fig.width=8}
-susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.flrerun.susieIrss.txt")
-f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)")
-f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)")
-f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
-f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
-gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
-```

### 2021-01-19 code/gwas.R

@@ -1,20 +1,6 @@
 library(foreach)
 library(doParallel)
 
-get_geno <- function(genof, mode = mode,
-                   variantidx = NULL, outputdir = NULL){
-  if (mode == "expr"){
-    geno <- read_expr(genof, variantidx = variantidx)
-  } else if (mode == "snp"){
-    pvarf <- prep_pvar(genof, outputdir = outputdir)
-    pgen <- prep_pgen(genof, pvarf)
-    geno <- read_pgen(pgen, variantidx = variantidx)
-  } else {
-    stop("undefined mode")
-  }
-  geno
-}
-
 
 #' @description this is a function to run GWAS analysis for small genotype matrix (using glm),for faster results use plink (start from plink files). Credit to http://www.stat-gen.org/tut/tut_analysis.html, requires tabix and bgzip if compress is true.
 #'
@@ -45,7 +31,23 @@ GWAA <- function(genof, mode = c("snp", "expr"), pheno, snpname, anno = NULL, ou
     snp.start <- seq(1, nSNPs, genosplit) # index of first SNP in group
     snp.stop <- pmin(snp.start+genosplit-1, nSNPs) # index of last SNP in group
 
-    for (part in 1:length(snp.start)) {
+    res <- foreach(part = 1:length(snp.start),.combine='rbind', .packages = "ctwas")  %dopar% {
+
+      cat(sprintf("GWAS %s%% finished\n", part/length(snp.start)))
+
+      get_geno <- function(genof, mode = mode,
+                           variantidx = NULL, outputdir = NULL){
+        if (mode == "expr"){
+          geno <- read_expr(genof, variantidx = variantidx)
+        } else if (mode == "snp"){
+          pvarf <- prep_pvar(genof, outputdir = outputdir)
+          pgen <- prep_pgen(genof, pvarf)
+          geno <- read_pgen(pgen, variantidx = variantidx)
+        } else {
+          stop("undefined mode")
+        }
+        geno
+      }
 
       geno.i <- get_geno(genof, mode = mode,
                          variantidx = snp.start[part]:snp.stop[part],
@@ -58,25 +60,26 @@ GWAA <- function(genof, mode = c("snp", "expr"), pheno, snpname, anno = NULL, ou
         anno.i <- NULL
       }
 
-      res <- foreach(snp.idx=1: length(snpname.i), .combine='rbind') %dopar% {
[…]

### 2021-01-27 analysis/index.Rmd
ctwas para results

@@
 
+## `ctwas` package
+Test the summary statistics version of the `ctwas` algorithm. Now provided as a R package.
+
+* [Using UKbiobank data (22k samples), fusion weights Adipose](simulation-ctwas-ukbWG-gtex.adipose_s40.22.html)
+
+* [Using UKbiobank data (22k samples), fusion weights Adipose, in-sample LD](simulation-ctwas-ukbWG-gtex.adipose_s40.22_2.html)
+
+* [Using UKbiobank data (45k samples), fusion weights Adipose](simulation-ctwas-ukbWG-gtex.adipose_s80.45.html)
+
+* [Using UKbiobank data (110k samples), fusion weights Adipose](simulation-ctwas-ukbWG-gtex.adipose_s200.113.html)
+
+  * [Visualization](simulation-ctwas-ukbWG-gtex.adipose_visual.html)
+
 # Related software

### 2021-01-27 analysis/simulation-ctwas-ukbWG-gtex.adipose_s200.113.Rmd
ctwas para results

---
title: "Simulations to test ctwas summary stats version, 113k samples"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  message= F, warning = F)
```

```{r libs}
library(ctwas)
library(data.table)
suppressMessages({library(plotly)})
library(tidyr)
library(plyr)
library(stringr)
source("~/causalTWAS/causal-TWAS/analysis/summarize_ctwas_plots.R")
source('~/causalTWAS/causal-TWAS/code/qqplot.R')
```

```{r files}
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s200.113/ukb-s200.113_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s200.113/ukb-s200.113.2_pgenfs.txt"
outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210118/" # /
runtag = "ukb-s200.113-adi"
configtags = 1
simutags = paste(rep(1:2, each = length(1:5)), 1:5, sep = "-")

pgenfs <- read.table(pgenfn, header = F, stringsAsFactors = F)[,1]
pvarfs <- sapply(pgenfs, prep_pvar, outputdir = outputdir)

ld_pgenfs <- read.table(ld_pgenfn, header = F, stringsAsFactors = F)[,1]
ld_pvarfs <- sapply(ld_pgenfs, prep_pvar, outputdir = outputdir)

pgens <- lapply(1:length(pgenfs), function(x) prep_pgen(pgenf = pgenfs[x],pvarf = pvarfs[x]))
```

# Analysis description

```{r parameters}
n.ori <- 200000 # number of samples
n <- pgenlibr::GetRawSampleCt(pgens[[1]])
p <- sum(unlist(lapply(pgens, pgenlibr::GetVariantCt))) # number of SNPs
J <- 8021 # number of genes
```

## Data

  * **GWAS summary statistics** we simulated summary statistics data with different causal gene/SNP proportion and PVE. To simulate this data, we need the following:
      * **genotype data** we used is from UKB biobank, randomly selecting `r as.character(n.ori)` samples. We then filtered samples based on relatedness, ethics and other qc metrics, that ended up with **n = `r as.character(n)` samples**.  We use SNP genotype data from chr 1 to chr 22 combined from UKB. There are **total = `r as.character(p)` SNPs**.

      * **Expression models** The one we used in this analysis is GTEx Adipose tissue v7 dataset. This dataset contains ~ 380 samples, **`r as.character(J)` genes** with expression model. [FUSION/TWAS](http://gusevlab.org/projects/fusion/) were used to train expression model and we used their lasso results. SNPs included in eQTL anlaysis are restricted to cis-locus 500kb on either side of the gene boundary. eQTLs are defined as SNPs with abs(effectize) > 1e-8 in lasso results.

      To simulate phenotype data, first we impute gene expression based on expression models, then we set gene/SNP pi1 and PVE, get rough effect size for causal SNPs and genes and simulate phenotype under the sparse model with spike and slab prior.
      Then we performed GWAS for all SNPs and get z scores for each by univariate linear regression.

  * **LD genotype reference**
  We randomly selected 2000 samples to serve as the LD reference.

  * **Expression models**  
  We used GTEx Adipose tissue v7 dataset, the same as used for simulating phenotypes.

## Analysis

1. Get z scores for gene expression. 
We used expression models and LD reference to get z scores for gene expression.

2. Run ctwas_rss
We used LDetect to define regions. `ctwas_rss` algorithm first runs on all regions to get rough estimate for gene and SNP prior. Then run on small regions (having small probablities of having > 1 causal signals based on rough estimates) to get more accurate estimate. To lower computational burden, we downsampled SNPs (0.1), estimate parameters and convert back to orginal scale. Lastly, run susie with given L for all regions and for all genes and SNPs using estimated prior and prior variance.

# Power estimation

```{r power}
simutag <- "1-1"
niter <- 1000
snp.p <- 5e-8
gene.p <- 1e-5
source(paste0(outputdir, "simu", simutag, "_param.R"))
load(paste0(outputdir, runtag, "_simu", simutag, "-pheno.Rd"))
```

We select run `r simutag`  as an example.

* For SNPs. $\pi_1 =$ `r pi_theta` , effect size = `r phenores[["batch"]][[1]][["sigma_theta"]]`, PVE = `r phenores$param$pve.snp.truth`. Power at `r as.character(snp.p)` p value cutoff:

```{r snp}
load("data/power_s200.113.Rd")
# p1 <- pow(niter, n, phenores[["batch"]][[1]][["sigma_theta"]], snp.p)
print(p1)
```

* For genes. $\pi_1 = `r pi_beta`$ , effect size = `r phenores[["batch"]][[1]][["sigma_beta"]]`, PVE = `r phenores$param$pve.gene.truth`. Power at `r as.character(gene.p)` p value cutoff:

```{r gene}
# p2 <- pow(niter, n, phenores[["batch"]][[1]][["sigma_beta"]], gene.p)
print(p2)
# save(p1,p2, file = "data/power_s200.113.Rd")
```

# GWAS/TWAS p value distribution

```{r pvalue}
simutag <- "1-1"
chrom <- 1
source(paste0(outputdir, "simu", simutag, "_param.R"))
load(paste0(outputdir, runtag, "_simu", simutag, "-pheno.Rd"))
```
We select run `r simutag`  as an example.

* For genes. $\pi_1 = `r pi_beta`$ , effect size = `r phenores[["batch"]][[1]][["sigma_beta"]]`, PVE = `r phenores$param$pve.gene.truth`.
TWAS p values and qqplot:

```{r gene-p, message=F, fig.width= 7, fig.height=7}
exprgwasf <- paste0(outputdir, runtag, "_simu", simutag, ".exprgwas.txt.gz")

exprvarf <- paste0(outputdir, runtag, "_chr", chrom, ".exprvar")
exprid <- read_exprvar(exprvarf)[, "id"]
cau <- as.matrix(exprid[phenores[["batch"]][[chrom]][["idx.cgene"]]])
pdist_plot(exprgwasf, chrom, cau)
```
```{r gene-qq, fig.width= 5, fig.height=5}
exprgwas <- fread(exprgwasf, header =T)
gg_qqplot(exprgwas$PVALUE)
```

* For SNPs. $\pi_1 =$ `r pi_theta` , effect size = `r phenores[["batch"]][[1]][["sigma_theta"]]`, PVE = `r phenores$param$pve.snp.truth`.
GWAS p values and qqplot:

```{r snp-p, fig.width= 7, fig.height=7}
snpgwasf <-  paste0(outputdir, runtag, "_simu", simutag, ".snpgwas.txt.gz")

pvarf <- pvarfs[chrom]
snpid <- read_pvar(pvarf)[, "id"]
cau <- as.matrix(snpid[phenores[["batch"]][[chrom]][["idx.cSNP"]]])
pdist_plot(snpgwasf, chrom, cau, thin = 0.1)
```

```{r snp-qq, fig.width= 5, fig.height=5}
snpgwas <- fread(snpgwasf, header =T)
gg_qqplot(snpgwas$PVALUE, thin = 0.1)
```

# `ctwas` results

Results: Each row shows parameter estimation results from 5 simulation runs with similar settings (i.e. pi1 and PVE for genes and SNPs). Results from each run were represented by one dot, dots with the same color come from the same run. `truth`: the true parameters,  `selected_truth`: the truth in selected regions that were used to estimate parameters, `ctwas`: ctwas estimated parameters (using summary statistics as input).

```{r func}
plot_par <- function(configtag, tags){
  source(paste0(outputdir, "config", configtag, ".R"))
  phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
  susieIfs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".s2.susieIrssres.Rd")
  susieIfs2 <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".s2.susieIrss.txt")

  mtx <- show_param(phenofs, susieIfs, susieIfs2, thin = thin)
  par(mfrow=c(1,3))
  cat("simulations ", paste(tags, sep=",") , ": ")
  cat("mean gene PVE:", mean(mtx[, "PVE.gene_truth"]), ",", "mean SNP PVE:", mean(mtx[, "PVE.SNP_truth"]), "\n")
  plot_param(mtx)
}

plot_PIP <- function(configtag, tags){
   phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".s2.susieIrss.txt")

   f1 <- caliPIP_plot(phenofs, susieIfs)
   f2 <- ncausal_plot(phenofs, susieIfs) 
   gridExtra::grid.arrange(f1, f2, ncol =2)
}
```

## ctwas_rss

```{r param1susie, fig.height=3, fig.width=8, message=F}
configtag <- 1

simutags <- paste(1, 1:5, sep = "-")
plot_par(configtag, simutags)
# plot_PIP(configtag, simutags)

simutags <- paste(2, 1:5, sep = "-")
plot_par(configtag, simutags)
# plot_PIP(configtag, simutags)
```

### 2021-01-27 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22.Rmd
ctwas para results

--- 2021-01-27 analysis/simulation-ctwas-ukbWG-gtex.adipose_s200.113.Rmd
@@
 ---
-title: "Simulations to test ctwas summary stats version, 113k samples"
+title: "Simulations to test ctwas summary stats version, 22k samples"
 output: html_document
@@
 ```{r files}
-pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s200.113/ukb-s200.113_pgenfs.txt"
-ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s200.113/ukb-s200.113.2_pgenfs.txt"
-outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210118/" # /
-runtag = "ukb-s200.113-adi"
+pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22_pgenfs.txt"
+ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22.2_pgenfs.txt"
+outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210113/" # /
+runtag = "ukb-s40.22-adi"
 configtags = 1
@@
 ```{r parameters}
-n.ori <- 200000 # number of samples
+n.ori <- 40000 # number of samples
 n <- pgenlibr::GetRawSampleCt(pgens[[1]])
@@
   * **LD genotype reference**
-  We randomly selected 2000 samples to serve as the LD reference.
+  We used 2k samples from UKbiobank with same ethnicity to serve as the LD reference.
 
@@
 ```{r snp}
-load("data/power_s200.113.Rd")
+load("data/power_s40.22.Rd")
 # p1 <- pow(niter, n, phenores[["batch"]][[1]][["sigma_theta"]], snp.p)
@@
 print(p2)
-# save(p1,p2, file = "data/power_s200.113.Rd")
+# save(p1,p2, file = "data/power_s80.45.Rd")
 ```
@@
    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
-   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".s2.susieIrss.txt")
+   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
 
@@
 
-## ctwas_rss
+## ctwas_rss parameter estimation
 
@@
 # plot_PIP(configtag, simutags)
+
+simutags <- paste(3, 1:5, sep = "-")
+plot_par(configtag, simutags)
+
+simutags <- paste(4, 1:5, sep = "-")
+plot_par(configtag, simutags)
+
+simutags <- paste(5, 1:5, sep = "-")
+plot_par(configtag, simutags)
+
+simutags <- paste(6, 1:5, sep = "-")
+plot_par(configtag, simutags)
+
+simutags <- paste(7, 1:5, sep = "-")
+plot_par(configtag, simutags)
+
+simutags <- paste(8, 1:5, sep = "-")
+plot_par(configtag, simutags)
 ```

### 2021-01-27 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22_2.Rmd
ctwas para results

--- 2021-01-27 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22.Rmd
@@
 ---
-title: "Simulations to test ctwas summary stats version, 22k samples"
+title: "Simulations to test ctwas summary stats version, 22k samples, in-sample LD"
 output: html_document
@@
 pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22_pgenfs.txt"
-ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22.2_pgenfs.txt"
-outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210113/" # /
+ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22_pgenfs.txt"
+outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210112/" # /
 runtag = "ukb-s40.22-adi"
@@
   * **LD genotype reference**
-  We used 2k samples from UKbiobank with same ethnicity to serve as the LD reference.
+  We used the same genotype as in simulation to serve as the LD reference.
 
@@
 # plot_PIP(configtag, simutags)
-
-simutags <- paste(3, 1:5, sep = "-")
-plot_par(configtag, simutags)
-
-simutags <- paste(4, 1:5, sep = "-")
-plot_par(configtag, simutags)
-
-simutags <- paste(5, 1:5, sep = "-")
-plot_par(configtag, simutags)
-
-simutags <- paste(6, 1:5, sep = "-")
-plot_par(configtag, simutags)
-
-simutags <- paste(7, 1:5, sep = "-")
-plot_par(configtag, simutags)
-
-simutags <- paste(8, 1:5, sep = "-")
-plot_par(configtag, simutags)
 ```

### 2021-01-27 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45.Rmd
ctwas para results

--- 2021-01-27 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22.Rmd
@@
 ---
-title: "Simulations to test ctwas summary stats version, 22k samples"
+title: "Simulations to test ctwas summary stats version, 45k samples"
 output: html_document
@@
 ```{r files}
-pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22_pgenfs.txt"
-ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22.2_pgenfs.txt"
-outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210113/" # /
-runtag = "ukb-s40.22-adi"
+pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt"
+ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
+outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210117/" # /
+runtag = "ukb-s80.45-adi"
 configtags = 1
@@
 ```{r parameters}
-n.ori <- 40000 # number of samples
+n.ori <- 80000 # number of samples
 n <- pgenlibr::GetRawSampleCt(pgens[[1]])
@@
   * **LD genotype reference**
-  We used 2k samples from UKbiobank with same ethnicity to serve as the LD reference.
+  We randomly selected 2000 samples to serve as the LD reference.
 
@@
 ```{r snp}
-load("data/power_s40.22.Rd")
+load("data/power_s80.45.Rd")
 # p1 <- pow(niter, n, phenores[["batch"]][[1]][["sigma_theta"]], snp.p)
@@
 
-## ctwas_rss parameter estimation
+## ctwas_rss
 

### 2021-01-27 analysis/simulation-ctwas-ukbWG-gtex.adipose_visual.Rmd
ctwas para results

--- 2021-01-17 analysis/simulation-susieI-ukbWG-gtex.adipose_visual.Rmd
@@
 ---
-title: "Visualize ukb whole genome, s40.22"
+title: "Visualize ukb whole genome, using ctwas results"
 date: "`r Sys.Date()`"
@@
 ```{r output dir}
-simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
-outputdir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
+outputdir <- "~/causalTWAS/simulations/simulation_ctwas_rss_20210112/"
 
-source("analysis/visual_twas_plots.R")
+source("~/causalTWAS/causal-TWAS/analysis/summarize_ctwas_plots.R")
+source("~/causalTWAS/causal-TWAS/analysis/visual_twas_plots.R")
 ```
@@
 
-plot_all <- function(susief, pf, phenof, chrom, ldres = NULL){
+plot_all <- function(susief, pf, phenof, chr, ldres = NULL, thin = 1){
   a <- fread(susief, header =T)
+  setnames(a, old = "id", new = "name")
   b1 <- fread(paste0(pf, ".snpgwas.txt.gz"), header =T)
   b2 <- fread(paste0(pf, ".exprgwas.txt.gz"), header =T)
-  b <- rbind(b1, b2)
-  setnames(b, old = c('X.CHROM', 'PVALUE', 'MARKER_ID'),
-         new = c("chr", 'pvalue', 'name'))
+  setnames(b1, old = "pos", new = "p0")
+  b <- rbind(b1, b2, fill = T)
+  setnames(b, old = c('PVALUE', 'id'),
+         new = c('pvalue', 'name'))
 
   load(phenof)
+  cau <- get_causal_id(phenores)
 
-  a <- a[a$chr == chrom,]
-  b <- b[b$chr == chrom,]
+  a <- a[a$chrom == chr,]
+  b <- b[b$chrom == chr,]
 
-  a <- merge(a, b, by = "name", all.x =T)
+  a <- merge(a, b, by = "name")
+  a$ifcausal <- ifelse(a$name %in% cau, "Causal", "Non causal" )
 
@@
     a <- merge(a, ld, by = "name", all.x = T)}
-
-  a <- a[, ifcausal:=as.character(ifcausal)]
-  a[a$ifcausal == "1", "ifcausal"] = "Causal"
-  a[a$ifcausal == "0", "ifcausal"] = "Non causal"
 
@@
 ```{r sim 1-3-17, out.width = "98%", out.height="95%"}
-chrom <- 2
-tag <- "20201001-2-5"
-susief <-  paste0(outputdir, tag, ".config16.flrerun.susieI.txt")
-pf <- paste0(simdatadir, tag)
-phenof <- paste0(simdatadir, "simu_", tag, "-pheno.Rd")
+# TODO, thin option
+chrom <- 1
+runtag <- "ukb-s40.22-adi"
+simutag <- "1-1"
+configtag <- 1
+
+susief <- paste0(outputdir, runtag, "_simu", simutag, "_config", configtag, ".s2.susieIrss.txt")
+pf <- paste0(outputdir, runtag, "_simu",simutag)
+phenof <- paste0(outputdir, runtag, "_simu", simutag, "-pheno.Rd")
 plot_all(susief, pf, phenof, chrom)

### 2021-01-27 analysis/visual_twas_plots.R

@@ -2,7 +2,7 @@ library(mr.ash.alpha)
 library(data.table)
 suppressMessages({library(plotly)})
 
-source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")
+source("~/causalTWAS/causal-TWAS/code/before_package/fit_mr.ash.R")
 
 get_files <- function(tag, tag2){
   par <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt")

### 2021-02-10 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22.Rmd
match ctwas 0.1.10

@@
 print(p2)
-# save(p1,p2, file = "data/power_s80.45.Rd")
+# save(p1,p2, file = "data/power_s40.22.Rd")
 ```
@@
 ```{r func}
-plot_par <- function(configtag, tags){
+plot_par <- function(configtag, runtag, simutags){
   source(paste0(outputdir, "config", configtag, ".R"))
@@
   par(mfrow=c(1,3))
-  cat("simulations ", paste(tags, sep=",") , ": ")
+  cat("simulations ", paste(simutags, sep=",") , ": ")
   cat("mean gene PVE:", mean(mtx[, "PVE.gene_truth"]), ",", "mean SNP PVE:", mean(mtx[, "PVE.SNP_truth"]), "\n")
@@
 
-plot_PIP <- function(configtag, tags){
-   phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+plot_PIP <- function(configtag, runtag,  simutags){
+   phenofs <- paste0(outputdir, "ukb-s40.22-adi", "_simu", simutags, "-pheno.Rd")
    susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
@@
 
-## ctwas_rss parameter estimation
+## Results-last step impact
 
-```{r param1susie, fig.height=3, fig.width=8, message=F}
+After we have the estimated parameters, we obtain PIP for genes in one the following ways:
+(1) run susie for each region using all SNPs
+```{r param1-1, fig.height=3, fig.width=8, message=F, cache=T}
 configtag <- 1
+runtag2 = "config1_ctwas_rss_s3_full/ukb-s40.22-adi"
 
-simutags <- paste(1, 1:5, sep = "-")
-plot_par(configtag, simutags)
-# plot_PIP(configtag, simutags)
+simutags <- paste(1, c(2,4,5), sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag2, simutags)
 
-simutags <- paste(2, 1:5, sep = "-")
-plot_par(configtag, simutags)
-# plot_PIP(configtag, simutags)
+simutags <- paste(2, c(2,4), sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag2, simutags)
+```
+(2) run susie for each region thinned SNPs 
+```{r param1-2, fig.height=3, fig.width=8, message=F}
+configtag <- 1
+runtag2 = "config1_ctwas_rss_s3_thin/ukb-s40.22-adi"
+
+simutags <- paste(1, c(2,4,5), sep = "-")
+plot_PIP(configtag, runtag2, simutags)
+
+simutags <- paste(2, c(2,4), sep = "-")
+plot_PIP(configtag, runtag2, simutags)
+```
+
+(3) run susie for each regions with thinned SNPs, then select regions with strong gene PIP (max PIP > 0.8) and rerun susie with full SNPs.
+```{r param1-3, fig.height=3, fig.width=8, message=F}
+configtag <- 1
+runtag2 = "ukb-s40.22-adi"
+
+simutags <- paste(1, c(2,4,5), sep = "-")
+plot_PIP(configtag, runtag, simutags)
+
+simutags <- paste(2, c(2,4), sep = "-")
+plot_PIP(configtag, runtag2, simutags)
+```
+
+## Results-more simulation settings
+We run the last step of `ctwas_rss` using way (3) as described above.
+
+```{r param2, fig.height=3, fig.width=8, message=F}
+
+configtag <- 1
+runtag = "ukb-s40.22-adi"
 
 simutags <- paste(3, 1:5, sep = "-")
-plot_par(configtag, simutags)
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
 
 simutags <- paste(4, 1:5, sep = "-")
-plot_par(configtag, simutags)
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
 
 simutags <- paste(5, 1:5, sep = "-")
-plot_par(configtag, simutags)
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
 
 simutags <- paste(6, 1:5, sep = "-")
-plot_par(configtag, simutags)
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
 
 simutags <- paste(7, 1:5, sep = "-")
-plot_par(configtag, simutags)
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
 
 simutags <- paste(8, 1:5, sep = "-")
-plot_par(configtag, simutags)
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
+
+simutags <- paste(9, 1:4, sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
+
+simutags <- paste(10, 1:5, sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
 ```

### 2021-02-10 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45.Rmd
match ctwas 0.1.10

@@
 
-plot_PIP <- function(configtag, tags){
-   phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+plot_PIP <- function(configtag, runtag,  simutags){
+   phenofs <- paste0(outputdir, "ukb-s80.45-adi", "_simu", simutags, "-pheno.Rd")
    susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
@@
 
-## ctwas_rss
+## Results-last step impact
 
-```{r param1susie, fig.height=3, fig.width=8, message=F}
+After we have the estimated parameters, we obtain PIP for genes in one the following ways:
+(1) run susie for each region using all SNPs
+```{r param1-1, fig.height=3, fig.width=8, message=F, cache=T}
 configtag <- 1
+runtag2 = "config1_ctwas_rss_s3_full/ukb-s80.45-adi"
 
 simutags <- paste(1, 1:5, sep = "-")
-plot_par(configtag, simutags)
-# plot_PIP(configtag, simutags)
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag2, simutags)
 
 simutags <- paste(2, 1:5, sep = "-")
-plot_par(configtag, simutags)
-# plot_PIP(configtag, simutags)
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
+```
+(2) run susie for each region thinned SNPs 
+```{r param1-2, fig.height=3, fig.width=8, message=F}
+configtag <- 1
+runtag2 = "config1_ctwas_rss_s3_thin/ukb-s80.45-adi"
+
+simutags <- paste(1, 1:5, sep = "-")
+plot_PIP(configtag, runtag2, simutags)
+
+simutags <- paste(2, 1:5, sep = "-")
+plot_PIP(configtag, runtag2, simutags)
+```
+
+(3) run susie for each regions with thinned SNPs, then select regions with strong gene PIP (max PIP > 0.8) and rerun susie with full SNPs.
+```{r param1-3, fig.height=3, fig.width=8, message=F}
+configtag <- 1
+runtag2 = "ukb-s80.45-adi"
+
+simutags <- paste(1, 1:5, sep = "-")
+plot_PIP(configtag, runtag2, simutags)
+
+simutags <- paste(2, 1:5, sep = "-")
+plot_PIP(configtag, runtag2, simutags)
+```
+
+## Results-more simulation settings
+We run the last step of `ctwas_rss` using way (3) as described above.
+
+```{r param2, fig.height=3, fig.width=8, message=F}
+
+configtag <- 1
+runtag = "ukb-s80.45-adi"
 
 simutags <- paste(3, 1:5, sep = "-")
-plot_par(configtag, simutags)
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
 
 simutags <- paste(4, 1:5, sep = "-")
-plot_par(configtag, simutags)
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
 
-simutags <- paste(5, 1:5, sep = "-")
-plot_par(configtag, simutags)
+simutags <- paste(5, 2:3, sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
 
 simutags <- paste(6, 1:5, sep = "-")
-plot_par(configtag, simutags)
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
 
-simutags <- paste(7, 1:5, sep = "-")
-plot_par(configtag, simutags)
+simutags <- paste(7, 2:5, sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
 
 simutags <- paste(8, 1:5, sep = "-")
-plot_par(configtag, simutags)
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
+
+simutags <- paste(9, 1:5, sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
+
+simutags <- paste(10, 1:5, sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
 ```

### 2021-02-10 analysis/simulation-ctwas-ukbWG-gtex.adipose_visual.Rmd
match ctwas 0.1.10

@@
 ```
-Provide visualization for analysis [here]
 
 ```{r output dir}
-outputdir <- "~/causalTWAS/simulations/simulation_ctwas_rss_20210112/"
+outputdir <- "~/causalTWAS/simulations/simulation_ctwas_rss_20210117/"
 
@@
   a$ifcausal <- ifelse(a$name %in% cau, "Causal", "Non causal" )
+
+  if (thin !=1){
+    b <- a[a$type == "gene" | (a$type == "SNP" & a$ifcausal == "Causal") | 
+             (a$type == "SNP" & a$ifcausal == "Non causal" & rbinom(nrow(a), 1, thin) ==1)]
+  }
 
@@
 
-## Simulation- high power
+## Simulation
 
@@
 
-```{r sim 1-3-17, out.width = "98%", out.height="95%"}
-# TODO, thin option
-chrom <- 1
-runtag <- "ukb-s40.22-adi"
-simutag <- "1-1"
+```{r sim 2, out.width = "98%", out.height="95%"}
+chrom <- 7
+runtag <- "ukb-s80.45-adi"
+simutag <- "2-1"
 configtag <- 1
 
-susief <- paste0(outputdir, runtag, "_simu", simutag, "_config", configtag, ".s2.susieIrss.txt")
+susief <- paste0(outputdir, runtag, "_simu", simutag, "_config", configtag, ".susieIrss.txt")
 pf <- paste0(outputdir, runtag, "_simu",simutag)
 phenof <- paste0(outputdir, runtag, "_simu", simutag, "-pheno.Rd")
-plot_all(susief, pf, phenof, chrom)
+plot_all(susief, pf, phenof, chrom, thin = 0.01)
 ```

### 2021-02-19 analysis/index.Rmd
mix normal simulate phenotype

@@
 
+### Using normal distribution as prior distribution for causal effects
+
 * [Using UKbiobank data (22k samples), fusion weights Adipose](simulation-ctwas-ukbWG-gtex.adipose_s40.22.html)
@@
 
+### Using normal mixture distribution as prior distribution for causal effects
+
+* [Using UKbiobank data (22k samples), fusion weights Adipose](simulation-ctwas-ukbWG-gtex.adipose_s40.22_mixnormal.html)
+
+* [Using UKbiobank data (45k samples), fusion weights Adipose](simulation-ctwas-ukbWG-gtex.adipose_s80.45_mixnormal.html)
+
 # Related software

### 2021-02-19 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22_mixnormal.Rmd
mix normal simulate phenotype

--- 2021-02-10 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22.Rmd
@@
 ---
-title: "Simulations to test ctwas summary stats version, 22k samples"
+title: "Simulations (mix normal prior), ctwas_rss, 22k samples"
 output: html_document
@@
 ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22.2_pgenfs.txt"
-outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210113/" # /
+outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210215/" # /
 runtag = "ukb-s40.22-adi"
@@
 
-      To simulate phenotype data, first we impute gene expression based on expression models, then we set gene/SNP pi1 and PVE, get rough effect size for causal SNPs and genes and simulate phenotype under the sparse model with spike and slab prior.
+      To simulate phenotype data, first we impute gene expression based on expression models, then we set gene/SNP pi1 and PVE, get rough effect size for causal SNPs and genes and simulate phenotype under the mixture of normal distributions.
+
+* The prior distribution for causal effect: 
+```{r }
+
+nmix <- 4
+pi_k <- rep(1/nmix, nmix)
+
+ra <- 2**(1: length(pi_k))
+sa2_kr <-  ra / (pi_k * sum(ra))
+
+cat("number of mixtures: ", nmix, "\n")
+cat("variance of each component:", sa2_kr, "\n")
+
+hist(mixtools::rnormmix(2000, lambda = pi_k,
+                        mu = rep(0, length(pi_k)),
+                        sigma = sa2_kr), breaks = 100, main = "prior distribution", xlab = "")
+```
+
 Then we performed GWAS for all SNPs and get z scores for each by univariate linear regression.
@@
 Then run on small regions (having small probablities of having > 1 causal signals based on rough estimates) to get more accurate estimate.
-To lower computational burden, we downsampled SNPs (0.1), estimate parameters and convert back to orginal scale.
-Lastly, run susie with given L for all regions and for all genes and SNPs using estimated prior and prior variance.
+To lower computational burden, we downsampled SNPs (0.1), estimate parameters.
+We run susie for all regions using estimated parameters, select regions with strong gene signals (max gene PIP > 0.8) and rerun using full SNPs for these regions.
 
@@
 
-## Results-last step impact
-
-After we have the estimated parameters, we obtain PIP for genes in one the following ways:
-(1) run susie for each region using all SNPs
-```{r param1-1, fig.height=3, fig.width=8, message=F, cache=T}
-configtag <- 1
-runtag2 = "config1_ctwas_rss_s3_full/ukb-s40.22-adi"
-
-simutags <- paste(1, c(2,4,5), sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag2, simutags)
-
-simutags <- paste(2, c(2,4), sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag2, simutags)
-```
-(2) run susie for each region thinned SNPs 
-```{r param1-2, fig.height=3, fig.width=8, message=F}
-configtag <- 1
-runtag2 = "config1_ctwas_rss_s3_thin/ukb-s40.22-adi"
-
-simutags <- paste(1, c(2,4,5), sep = "-")
-plot_PIP(configtag, runtag2, simutags)
-
-simutags <- paste(2, c(2,4), sep = "-")
-plot_PIP(configtag, runtag2, simutags)
-```
-
-(3) run susie for each regions with thinned SNPs, then select regions with strong gene PIP (max PIP > 0.8) and rerun susie with full SNPs.
-```{r param1-3, fig.height=3, fig.width=8, message=F}
-configtag <- 1
-runtag2 = "ukb-s40.22-adi"
-
-simutags <- paste(1, c(2,4,5), sep = "-")
-plot_PIP(configtag, runtag, simutags)
-
-simutags <- paste(2, c(2,4), sep = "-")
-plot_PIP(configtag, runtag2, simutags)
-```
-
-## Results-more simulation settings
-We run the last step of `ctwas_rss` using way (3) as described above.
-
-```{r param2, fig.height=3, fig.width=8, message=F}
+```{r param, fig.height=3, fig.width=8, message=F}
 
@@
 runtag = "ukb-s40.22-adi"
+
+simutags <- paste(1, 1:5, sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
+
+simutags <- paste(2, 1:5, sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
 
@@
 
-simutags <- paste(9, 1:4, sep = "-")
+simutags <- paste(9, 1:5, sep = "-")
 plot_par(configtag, runtag, simutags)

### 2021-02-19 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_mixnormal.Rmd
mix normal simulate phenotype

--- 2021-02-19 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22_mixnormal.Rmd
@@
 ---
-title: "Simulations (mix normal prior), ctwas_rss, 22k samples"
+title: "Simulations (mix normal prior), ctwas_rss, 45k samples"
 output: html_document
@@
 ```{r files}
-pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22_pgenfs.txt"
-ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22.2_pgenfs.txt"
-outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210215/" # /
-runtag = "ukb-s40.22-adi"
+pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt"
+ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
+outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210216/" # /
+runtag = "ukb-s80.45-adi"
 configtags = 1
@@
 print(p2)
-# save(p1,p2, file = "data/power_s40.22.Rd")
+# save(p1,p2, file = "data/power_s80.45.Rd")
 ```
@@
 plot_PIP <- function(configtag, runtag,  simutags){
-   phenofs <- paste0(outputdir, "ukb-s40.22-adi", "_simu", simutags, "-pheno.Rd")
+   phenofs <- paste0(outputdir, "ukb-s80.45-adi", "_simu", simutags, "-pheno.Rd")
    susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
@@
 
-```{r param, fig.height=3, fig.width=8, message=F}
+```{r param2, fig.height=3, fig.width=8, message=F}
 
 configtag <- 1
-runtag = "ukb-s40.22-adi"
+runtag = "ukb-s80.45-adi"
 

### 2021-03-25 analysis/SNP-harmonization.Rmd
snp harmonization notes

---
title: "SNP-harmonization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SNP nomenclature

I found this paper particularly useful for clarifying SNP nomenclature:

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6099125/

# File formats for SNP information

Different file formats have different names for describing alleles and coding alleles.

* `.bim` format: they use A1 and A2 alleles, they claim A1 is usually the minor allele, see here: https://www.cog-genomics.org/plink/1.9/formats#bim

* `.pvar` format: they use Alt and Ref alleles, see here: https://www.cog-genomics.org/plink/2.0/formats#pvar

* `.traw` format: see here: https://www.cog-genomics.org/plink/2.0/formats#traw
the Ref allele is counted (coded)

* Fusion summary stats files: They use A1, A2. A1 is the effect allele.

* Fusion weights files: They use .bim format for SNP information, so 5th column is Alt, 6th column is Ref.

* SMR files: They use A1, A2. A1 is the effect allele. They also call A1 as Ref allele, A2 as Alt allele.

* ctwas format: `ctwas` package uses R package `pgenlibr` to read data. The Alt allele in .bim or .pvar files are counted, so the effect allele is Alt allele.

# Consistency between GTEx v7 and UKB data

* We check if SNP information provided by FUSION GTEx v7 trained and UKB SNP information are consistent.

UKBiobank SNP information were generated by plink2, we only selected variants with MAF > 0.05 variants and missing genotype rate < 0.05.

```{r consistent}
library(ctwas)
library(plyr)
ld_pgenfn = "~/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22.2_pgenfs.txt"
eqtlfn = "~/causalTWAS/fusion_weights/Adipose_Subcutaneous.lasso.eqtl.txt"

eqtl = read.table(eqtlfn, header = T)
colnames(eqtl)[1:6] <- c("chrom", "id", "cm", "pos", "alt", "ref")
ld_pgenfs <- read.table(ld_pgenfn, header = F, stringsAsFactors = F)[,1]
ld_pvarfs <- sapply(ld_pgenfs, prep_pvar, outputdir = "~/temp/")

i <- 1 # chrom 1 as an example
pvar <- read_pvar(ld_pvarfs[i])

cat("number of eQTLs on chromosome ", i,  "(by lasso): ", nrow(eqtl[eqtl$chrom == i,]))

a <- join(pvar, eqtl, type = "inner", by = c("id", "pos"))
cat("number of eQTLs on chromosome ", i,  "(by lasso, shared with UKB): ", nrow(a))

b <- join(pvar, eqtl, type = "inner")
cat("number of eQTLs on chromosome ", i,  "(by lasso, shared with UKB) that match: ", nrow(b))

cat("number of eQTLs on chromosome ", i,  "(by lasso, shared with UKB) that match after reverse alt and ref: ",
    which(a[!(a$id %in% b$id), 5] != a[!(a$id %in% b$id), 8]))

cat("number of eQTLs on chromosome ", i,  "(by lasso, shared with UKB) that match after flip: ",
sum(a[!(a$id %in% b$id), 5] != a[!(a$id %in% b$id), 8]))
```

# Procedures to harmonize SNPs

See here about some instructions for merging two datasets: 
https://zzz.bwh.harvard.edu/plink/dataman.shtml#flip

It is assumed that reported two alleles are on the same strand, but it can be either positive strand or negative strand.

* FUSION filters out all strand ambiguous SNPs (i.e. AT, TA, CG, GC alleles) and flipped strand unambiguous SNPs if there is a match after flipping. It also switch alt ref SNPs if there is a match after switching.

* SMR switch alt ref SNPs if there is a match after switching. No flipping.

* ctwas currently (v.0.1.11) does not harmonize data and assumes the alleles were harmonized before running ctwas, so they do not need to be flipped or switched.

### 2021-03-25 analysis/SNP-harmonization.Rmd
snp harmonization notes

@@
 
+eQTL information were obtained from FUSION pre-trained models. We used all SNPs with weights != 0 in
+
 ```{r consistent}
@@
 sum(a[!(a$id %in% b$id), 5] != a[!(a$id %in% b$id), 8]))
+
+cat("number of eQTLs on chromosome ", i,  "(by lasso, shared with UKB) that are strand ambiguous: ",
+nrow(a[(a$alt=="A" & a$ref=="T") | (a$alt=="T" & a$ref=="A") | (a$alt=="C" & a$ref=="G") | (a$alt=="G" & a$ref=="C"),]))
 ```

### 2021-04-05 analysis/index.Rmd
snp harmonization notes
add coloc

--- 2021-01-27 analysis/index.Rmd
@@
 # Data
-
-## Expression data
-  * [Fusion/TWAS gene expression models](fusion_expr.html)
+  * [SNP data harmonization notes](SNP-harmonization.html)
 
@@
 
-* [Using UKbiobank data (22k samples), fusion weights Adipose](simulation-ctwas-ukbWG-gtex.adipose_s40.22.html)
+### Using normal distribution as prior distribution for causal effects
+
+* [**Using UKbiobank data (22k samples), fusion weights Adipose**](simulation-ctwas-ukbWG-gtex.adipose_s40.22.html)
 
@@
 
-* [Using UKbiobank data (45k samples), fusion weights Adipose](simulation-ctwas-ukbWG-gtex.adipose_s80.45.html)
+* [**Using UKbiobank data (45k samples), fusion weights Adipose**](simulation-ctwas-ukbWG-gtex.adipose_s80.45.html)
 
@@
 * [Visualization](simulation-ctwas-ukbWG-gtex.adipose_visual.html)
+
+### Using normal mixture distribution as prior distribution for causal effects
+
+* [**Using UKbiobank data (22k samples), fusion weights Adipose**](simulation-ctwas-ukbWG-gtex.adipose_s40.22_mixnormal.html)
+
+* [**Using UKbiobank data (45k samples), fusion weights Adipose**](simulation-ctwas-ukbWG-gtex.adipose_s80.45_mixnormal.html)
 

### 2021-04-05 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22.Rmd
snp harmonization notes
add coloc

@@
 source("~/causalTWAS/causal-TWAS/analysis/summarize_ctwas_plots.R")
+source('~/causalTWAS/causal-TWAS/analysis/summarize_twas-coloc_plots.R')
 source('~/causalTWAS/causal-TWAS/code/qqplot.R')
@@
 outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210113/" # /
+comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210113_compare/"
 runtag = "ukb-s40.22-adi"
@@
 ## Analysis
+
+### ctwas
 
@@
 plot_PIP <- function(configtag, runtag,  simutags){
-   phenofs <- paste0(outputdir, "ukb-s40.22-adi", "_simu", simutags, "-pheno.Rd")
+   phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
    susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
@@
    gridExtra::grid.arrange(f1, f2, ncol =2)
+}
+
+plot_fusion_coloc <- function(configtag, runtag,  simutags){
+    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+    fusioncolocfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.coloc.result")
+
+    f1 <- caliFUSIONp_plot(phenofs, fusioncolocfs)
+    f2 <- ncausalFUSIONp_plot(phenofs, fusioncolocfs)
+    f3 <- caliPP4_plot(phenofs, fusioncolocfs)
+    f4 <- ncausalPP4_plot(phenofs, fusioncolocfs)
+    gridExtra::grid.arrange(f1, f2, ncol=2)
+    gridExtra::grid.arrange(f3, f4, ncol=2)
 }
@@
 
-## Results-more simulation settings
+## Results- More simutations + other software
 We run the last step of `ctwas_rss` using way (3) as described above.
+We run FUSION following default settings and adjust p values by BH method to get expected FDP.
+we ran coloc for all genes with TWAS p < 1e-4. We use PP4 (SNP associate with both traits).
 
@@
 
+simutags <- paste(1, c(2,4,5), sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
+plot_fusion_coloc(configtag, runtag, simutags)
+
+simutags <- paste(2, c(2,4), sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag2, simutags)
+plot_fusion_coloc(configtag, runtag, simutags)
+
 simutags <- paste(3, 1:5, sep = "-")
@@
 plot_PIP(configtag, runtag, simutags)
+plot_fusion_coloc(configtag, runtag, simutags)
 
@@
 plot_PIP(configtag, runtag, simutags)
+plot_fusion_coloc(configtag, runtag, simutags)
 ```

### 2021-04-05 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22_mixnormal.Rmd
add coloc

@@
 outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210215/" # /
+comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210215_compare/"
 runtag = "ukb-s40.22-adi"
@@
 
-# `ctwas` results
+# `ctwas` results + other comparators
 
@@
 `truth`: the true parameters,  `selected_truth`: the truth in selected regions that were used to estimate parameters, `ctwas`: ctwas estimated parameters (using summary statistics as input).
+
+We run FUSION following default settings and adjust p values by BH method to get expected FDP.
+we ran coloc for all genes with TWAS p < 1e-4. We use PP4 (SNP associate with both traits).
 
@@
 }
+
+plot_fusion_coloc <- function(configtag, runtag,  simutags){
+    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+    fusioncolocfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.coloc.result")
+
+    f1 <- caliFUSIONp_plot(phenofs, fusioncolocfs)
+    f2 <- ncausalFUSIONp_plot(phenofs, fusioncolocfs)
+    f3 <- caliPP4_plot(phenofs, fusioncolocfs)
+    f4 <- ncausalPP4_plot(phenofs, fusioncolocfs)
+    gridExtra::grid.arrange(f1, f2, ncol=2)
+    gridExtra::grid.arrange(f3, f4, ncol=2)
+}
 ```

### 2021-04-05 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45.Rmd
add coloc

@@
 outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210117/" # /
+comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210117_compare/"
 runtag = "ukb-s80.45-adi"
@@
 
-## Results-more simulation settings
+## Results-more simulations + compare
 We run the last step of `ctwas_rss` using way (3) as described above.
+We run FUSION following default settings and adjust p values by BH method to get expected FDP.
+we ran coloc for all genes with TWAS p < 1e-4. We use PP4 (SNP associate with both traits).
 
@@
 
+simutags <- paste(1, 1:5, sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag2, simutags)
+plot_fusion_coloc(configtag, runtag, simutags)
+
+simutags <- paste(2, 1:5, sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
+plot_fusion_coloc(configtag, runtag, simutags)
+
 simutags <- paste(3, 1:5, sep = "-")
@@
 plot_PIP(configtag, runtag, simutags)
+plot_fusion_coloc(configtag, runtag, simutags)
 
-simutags <- paste(5, 2:3, sep = "-")
+simutags <- paste(5, 1:5, sep = "-")
 plot_par(configtag, runtag, simutags)
 plot_PIP(configtag, runtag, simutags)
+plot_fusion_coloc(configtag, runtag, simutags)
 
-simutags <- paste(6, 1:5, sep = "-")
+simutags <- paste(6, 2:5, sep = "-")
 plot_par(configtag, runtag, simutags)
 plot_PIP(configtag, runtag, simutags)
+plot_fusion_coloc(configtag, runtag, simutags)
 
-simutags <- paste(7, 2:5, sep = "-")
+simutags <- paste(7, c(1:3,5), sep = "-")
 plot_par(configtag, runtag, simutags)
 plot_PIP(configtag, runtag, simutags)
+plot_fusion_coloc(configtag, runtag, simutags)
 

### 2021-04-05 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_mixnormal.Rmd
add coloc

@@
 outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210216/" # /
+comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210216_compare/"
 runtag = "ukb-s80.45-adi"

### 2021-04-09 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22.Rmd
fusion plot fix bug

@@
     f2 <- ncausalFUSIONp_plot(phenofs, fusioncolocfs)
-    f3 <- caliPP4_plot(phenofs, fusioncolocfs)
-    f4 <- ncausalPP4_plot(phenofs, fusioncolocfs)
+    f3 <- caliPP4_plot(phenofs, fusioncolocfs, twas.p = 0.05/J)
+    f4 <- ncausalPP4_plot(phenofs, fusioncolocfs, twas.p = 0.05/J)
     gridExtra::grid.arrange(f1, f2, ncol=2)

### 2021-04-09 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22_mixnormal.Rmd
fusion plot fix bug

@@
 
-```{r param, fig.height=3, fig.width=8, message=F}
+```{r param, fig.height=3, fig.width=10, message=F}
 

### 2021-04-09 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_mixnormal.Rmd
fusion plot fix bug

@@
 
-```{r param2, fig.height=3, fig.width=8, message=F}
+```{r param2, fig.height=3, fig.width=10, message=F}
 

### 2021-04-22 analysis/SNP-harmonization.Rmd
add coloc
focus results

@@
 * Fusion weights files: They use .bim format for SNP information, so 5th column is Alt, 6th column is Ref.
+Effect size in `wgt.matrix` variable is defined using Alt as effect ellele (5th column in `snps` variable).
 
@@
 * ctwas format: `ctwas` package uses R package `pgenlibr` to read data. The Alt allele in .bim or .pvar files are counted, so the effect allele is Alt allele.
+
+* plink association tests: see [here](http://zzz.bwh.harvard.edu/plink/anal.shtml).
+For the additive effects of SNPs, the direction of the regression coefficient represents the effect of each extra minor allele (i.e.
+a positive regression coefficient means that the minor allele increases risk/phenotype mean).
+If the --beta command is added along with --logistic, then the regression coefficients rather than the odds ratios will be returned.
+NOTE:Elsewhere in plink, the term reference allele is sometimes used to refer to A1, i.e.
+the --reference-allele command can be used to specify which allele is A1.
+Note that in association testing, the odds ratios, etc are typically calculated with A2 as the actual reference allele (i.e.
+a positive OR means A1 increases risk relative to A2).
+
+* GTEx summary statistics: The normalized effect size (NES) of the eQTLs is defined as the slope of the linear regression, and is computed as the effect of the alternative allele (ALT) relative to the reference allele (REF) in the human genome reference GRCh38/hg38 (i.e., the eQTL effect allele is the ALT allele).
+See [here](https://www.gtexportal.org/home/documentationPage).
+
+* GEMMA: "chr    rs  ps  n_miss  allel1  allel0  af  beta    se  l_remle p_wald".
+The 11 columns are: chromosome, SNP ID, basepair position, number of missing values for a given SNP, the effect (coded) allele, the other allele, frequency of the effect allele, effect size, standard error, lambda and p-value ([http://www.xzlab.org/software.html]).
 
@@
 
-eQTL information were obtained from FUSION pre-trained models. We used all SNPs with weights != 0 in
+eQTL information were obtained from FUSION pre-trained models. We used all SNPs with weights != 0 in lasso trained models.
 

### 2021-04-22 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22.Rmd
focus results

--- 2021-04-05 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22.Rmd
@@
 source('~/causalTWAS/causal-TWAS/analysis/summarize_twas-coloc_plots.R')
+source('~/causalTWAS/causal-TWAS/analysis/summarize_focus_plots.R')
 source('~/causalTWAS/causal-TWAS/code/qqplot.R')
@@
     f2 <- ncausalFUSIONp_plot(phenofs, fusioncolocfs)
-    f3 <- caliPP4_plot(phenofs, fusioncolocfs)
-    f4 <- ncausalPP4_plot(phenofs, fusioncolocfs)
+    f3 <- caliPP4_plot(phenofs, fusioncolocfs, twas.p = 0.05/J)
+    f4 <- ncausalPP4_plot(phenofs, fusioncolocfs, twas.p = 0.05/J)
     gridExtra::grid.arrange(f1, f2, ncol=2)
     gridExtra::grid.arrange(f3, f4, ncol=2)
+}
+
+plot_focus <- function(configtag, runtag,  simutags){
+    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+    focusfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous_setalleleUKB.focus.tsv")
+
+    f1 <- califocusPIP_plot(phenofs, focusfs)
+    f2 <- ncausalfocusPIP_plot(phenofs, focusfs)
+    gridExtra::grid.arrange(f1, f2, ncol=2)
 }
@@
 plot_fusion_coloc(configtag, runtag, simutags)
+plot_focus(configtag, runtag, simutags)
 
@@
 plot_fusion_coloc(configtag, runtag, simutags)
+plot_focus(configtag, runtag, simutags)
 ```

### 2021-04-22 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22_041521.Rmd
focus results

--- 2021-04-09 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22_mixnormal.Rmd
@@
 ---
-title: "Simulations (mix normal prior), ctwas_rss, 22k samples"
+title: "Simulations to test ctwas summary stats version, 22k samples"
 output: html_document
@@
 ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22.2_pgenfs.txt"
-outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210215/" # /
-comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210215_compare/"
+outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210415/" # /
+comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210415_compare/"
 runtag = "ukb-s40.22-adi"
@@
 
-      To simulate phenotype data, first we impute gene expression based on expression models, then we set gene/SNP pi1 and PVE, get rough effect size for causal SNPs and genes and simulate phenotype under the mixture of normal distributions.
-
-* The prior distribution for causal effect: 
-```{r }
-
-nmix <- 4
-pi_k <- rep(1/nmix, nmix)
-
-ra <- 2**(1: length(pi_k))
-sa2_kr <-  ra / (pi_k * sum(ra))
-
-cat("number of mixtures: ", nmix, "\n")
-cat("variance of each component:", sa2_kr, "\n")
-
-hist(mixtools::rnormmix(2000, lambda = pi_k,
-                        mu = rep(0, length(pi_k)),
-                        sigma = sa2_kr), breaks = 100, main = "prior distribution", xlab = "")
-```
-
+      To simulate phenotype data, first we impute gene expression based on expression models, then we set gene/SNP pi1 and PVE, get rough effect size for causal SNPs and genes and simulate phenotype under the sparse model with spike and slab prior.
       Then we performed GWAS for all SNPs and get z scores for each by univariate linear regression.
@@
 
+### ctwas
+
 1. Get z scores for gene expression. 
@@
 Then run on small regions (having small probablities of having > 1 causal signals based on rough estimates) to get more accurate estimate.
-To lower computational burden, we downsampled SNPs (0.1), estimate parameters.
-We run susie for all regions using estimated parameters, select regions with strong gene signals (max gene PIP > 0.8) and rerun using full SNPs for these regions.
+To lower computational burden, we downsampled SNPs (0.1), estimate parameters and convert back to orginal scale.
+Lastly, run susie with given L for all regions and for all genes and SNPs using estimated prior and prior variance.
 
@@
 
-# `ctwas` results + other comparators
+# `ctwas` results
 
@@
 `truth`: the true parameters,  `selected_truth`: the truth in selected regions that were used to estimate parameters, `ctwas`: ctwas estimated parameters (using summary statistics as input).
-
-We run FUSION following default settings and adjust p values by BH method to get expected FDP.
-we ran coloc for all genes with TWAS p < 1e-4. We use PP4 (SNP associate with both traits).
 
@@
 
-```{r param, fig.height=3, fig.width=10, message=F}
+```{r param2, fig.height=3, fig.width=8, message=F}
 
@@
 
-simutags <- paste(1, 1:5, sep = "-")
+simutags <- paste(1, c(1:5), sep = "-")
 plot_par(configtag, runtag, simutags)
@@
 plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
+plot_PIP(configtag, runtag2, simutags)
 plot_fusion_coloc(configtag, runtag, simutags)
@@
 
-simutags <- paste(9, 1:5, sep = "-")
+simutags <- paste(9, 1:4, sep = "-")
 plot_par(configtag, runtag, simutags)

### 2021-04-22 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22_mixnormal.Rmd
focus results

@@
 }
+
+plot_focus <- function(configtag, runtag,  simutags){
+    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+    focusfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous_setalleleUKB.focus.tsv")
+
+    f1 <- califocusPIP_plot(phenofs, focusfs)
+    f2 <- ncausalfocusPIP_plot(phenofs, focusfs)
+    gridExtra::grid.arrange(f1, f2, ncol=2)
+}
 ```
@@
 plot_fusion_coloc(configtag, runtag, simutags)
+plot_focus(configtag, runtag, simutags)
 
-simutags <- paste(6, 1:5, sep = "-")
+simutags <- paste(6, 2:5, sep = "-")
 plot_par(configtag, runtag, simutags)
@@
 plot_fusion_coloc(configtag, runtag, simutags)
+plot_focus(configtag, runtag, simutags)
 
-simutags <- paste(7, 1:5, sep = "-")
+simutags <- paste(7, c(1:3,5), sep = "-")
 plot_par(configtag, runtag, simutags)

### 2021-04-22 analysis/summarize_focus_plots.R

@@ -1,0 +1,35 @@
+source("~/causalTWAS/causal-TWAS/analysis/summarize_basic_plots.R")
+
+# coloc PP4 calibration plot
+califocusPIP_plot <- function(phenofs, focusfs, main = "FOCUS PIP Calibration"){
+  cau <- lapply(phenofs, function(x) {load(x);get_causal_id(phenores)})
+  df <- NULL
+  for (i in 1:length(focusfs)) {
+    res <- fread(focusfs[i], header = T)
+    res$runtag <- i
+    res <- res[res$mol_name != "NULL",]
+    res$ifcausal <- ifelse(res$mol_name %in% cau[[i]], 1, 0)
+    df <- rbind(df, res)
+  }
+
+  fig <- cp_plot(df$pip, df$ifcausal, df$runtag, mode ="PIP", main = main)
+  return(fig)
+}
+
+
+ncausalfocusPIP_plot <- function(phenofs, focusfs, main = "FOCUS PIP"){
+
+  cau <- lapply(phenofs, function(x) {load(x);get_causal_id(phenores)})
+
+  df <- NULL
+  for (i in 1:length(focusfs)) {
+    res <- fread(focusfs[i], header = T)
+    res$ifcausal <- ifelse(res$mol_name %in% cau[[i]], 1, 0)
+    res$runtag <- i
+    res <- res[res$mol_name != "NULL",]
+    df <- rbind(df, res)
+  }
+
+  fig <- nca_plot(df$pip, df$ifcausal, df$runtag, mode ="PIP", xmin = 0.5, main = main)
+  return(fig)
+}

### 2021-04-22 code/temp.R

@@ -1,0 +1,14 @@
+setwd("/home/simingz/causalTWAS/GTEx_v7_eQTL")
+lookupfile <- "GTEx_Analysis_2016-01-15_v7_WholeGenomeSeq_635Ind_PASS_AB02_GQ20_HETX_MISS15_PLINKQC.lookup_table.txt.gz"
+egenefile <- "Adipose_Subcutaneous.v7.egenes.txt.gz"
+eqtlfile <- "Adipose_Subcutaneous.v7.signif_variant_gene_pairs.txt.gz"
+
+library(data.table)
+eqtl <- fread(eqtlfile)
+ref <- fread(lookupfile)
+a <- merge(eqtl, ref, all.x = T, by = "variant_id")
+rm(ref, eqtl)
+egene <- fread(egenefile)
+egene <- egene[, 1:11, drop =F]
+out <- merge(a, egene, all.x = T, by = "gene_id")
+write.table(out , file= "Adipose_Subcutaneous.v7.eQTL.txt" , row.names=F, col.names=T, sep="\t", quote = F)

### 2021-05-09 analysis/index.Rmd
added smr-heidi results, scripts matching v0.1.25

@@
 
-* [**Using UKbiobank data (22k samples), fusion weights Adipose**](simulation-ctwas-ukbWG-gtex.adipose_s40.22.html)
+* [**Using UKbiobank data (22k samples), fusion weights Adipose**](simulation-ctwas-ukbWG-gtex.adipose_s40.22_041521.html)
 
@@
 
-* [**Using UKbiobank data (45k samples), fusion weights Adipose**](simulation-ctwas-ukbWG-gtex.adipose_s80.45.html)
+* [**Using UKbiobank data (45k samples), fusion weights Adipose**](simulation-ctwas-ukbWG-gtex.adipose_s80.45_041621.html)
 
@@
 
-* [**Using UKbiobank data (22k samples), fusion weights Adipose**](simulation-ctwas-ukbWG-gtex.adipose_s40.22_mixnormal.html)
+* [**Using UKbiobank data (22k samples), fusion weights Adipose**](simulation-ctwas-ukbWG-gtex.adipose_s40.22_mixnormal_041721.html)
 
-* [**Using UKbiobank data (45k samples), fusion weights Adipose**](simulation-ctwas-ukbWG-gtex.adipose_s80.45_mixnormal.html)
+* [**Using UKbiobank data (45k samples), fusion weights Adipose**](simulation-ctwas-ukbWG-gtex.adipose_s80.45_mixnormal_041821.html)
 

### 2021-05-09 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22_041521.Rmd
added smr-heidi results, scripts matching v0.1.25

@@
 source('~/causalTWAS/causal-TWAS/analysis/summarize_twas-coloc_plots.R')
+source('~/causalTWAS/causal-TWAS/analysis/summarize_focus_plots.R')
+source('~/causalTWAS/causal-TWAS/analysis/summarize_smr_plots.R')
 source('~/causalTWAS/causal-TWAS/code/qqplot.R')
@@
 
+We run FUSION following default settings and adjust p values by BH method to get expected FDP.
+We ran coloc for all genes with TWAS p < 1e-4. We use PP4 (SNP associate with both traits).
+We ran SMR+HEIDI, using eQTL summary statistics GTEx v.7. We filter the results by requiring p_HEIDI > 0.05. The plots 
+are based on SMR p value adjusted by BH method to get expected FDP.
+
 ```{r func}
@@
 }
+
+plot_focus <- function(configtag, runtag,  simutags){
+    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+    focusfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.focus.tsv")
+
+    f1 <- califocusPIP_plot(phenofs, focusfs)
+    f2 <- ncausalfocusPIP_plot(phenofs, focusfs)
+    gridExtra::grid.arrange(f1, f2, ncol=2)
+}
+
+plot_smr <- function(configtag, runtag,  simutags){
+    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+    smrfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.smr")
+
+    f1 <- caliSMRp_plot(phenofs, smrfs)
+    f2 <- ncausalSMRp_plot(phenofs, smrfs)
+    gridExtra::grid.arrange(f1, f2, ncol=2)
+}
 ```
@@
 plot_fusion_coloc(configtag, runtag, simutags)
+plot_focus(configtag, runtag, simutags)
+plot_smr(configtag, runtag, simutags)
 
@@
 plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag2, simutags)
+plot_PIP(configtag, runtag, simutags)
 plot_fusion_coloc(configtag, runtag, simutags)
+plot_focus(configtag, runtag, simutags)
+plot_smr(configtag, runtag, simutags)
 
@@
 plot_fusion_coloc(configtag, runtag, simutags)
+plot_focus(configtag, runtag, simutags)
+plot_smr(configtag, runtag, simutags)
 ```

### 2021-05-09 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22_mixnormal_041721.Rmd
added smr-heidi results, scripts matching v0.1.25

--- 2021-05-09 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22_041521.Rmd
@@
 ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22.2_pgenfs.txt"
-outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210415/" # /
-comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210415_compare/"
+outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210417/" # /
+comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210417_compare/"
 runtag = "ukb-s40.22-adi"
@@
 
-### ctwas
-
 1. Get z scores for gene expression. 
@@
 We run FUSION following default settings and adjust p values by BH method to get expected FDP.
-We ran coloc for all genes with TWAS p < 1e-4. We use PP4 (SNP associate with both traits).
+we ran coloc for all genes with TWAS p < 1e-4. We use PP4 (SNP associate with both traits).
 We ran SMR+HEIDI, using eQTL summary statistics GTEx v.7. We filter the results by requiring p_HEIDI > 0.05. The plots 
@@
 plot_PIP <- function(configtag, runtag,  simutags){
-   phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+   phenofs <- paste0(outputdir, "ukb-s40.22-adi", "_simu", simutags, "-pheno.Rd")
    susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
@@
     gridExtra::grid.arrange(f3, f4, ncol=2)
+
 }
-
 plot_focus <- function(configtag, runtag,  simutags){
@@
 
-simutags <- paste(2, 1:5, sep = "-")
+simutags <- paste(2, c(1,3:5), sep = "-")
 plot_par(configtag, runtag, simutags)

### 2021-05-09 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_041621.Rmd
added smr-heidi results, scripts matching v0.1.25

--- 2021-05-09 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22_041521.Rmd
@@
 ```{r files}
-pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22_pgenfs.txt"
-ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22.2_pgenfs.txt"
-outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210415/" # /
-comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210415_compare/"
-runtag = "ukb-s40.22-adi"
+pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt"
+ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
+outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210418/" # /
+comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210418_compare/"
+runtag = "ukb-s80.45-adi"
 configtags = 1
@@
 ## Analysis
-
-### ctwas
 
@@
 `truth`: the true parameters,  `selected_truth`: the truth in selected regions that were used to estimate parameters, `ctwas`: ctwas estimated parameters (using summary statistics as input).
-
 We run FUSION following default settings and adjust p values by BH method to get expected FDP. 
@@
 plot_PIP <- function(configtag, runtag,  simutags){
-   phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+   phenofs <- paste0(outputdir, "ukb-s80.45-adi", "_simu", simutags, "-pheno.Rd")
    susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
@@
 configtag <- 1
-runtag = "ukb-s40.22-adi"
+runtag = "ukb-s80.45-adi"
 
@@
 
-simutags <- paste(5, 1:5, sep = "-")
+simutags <- paste(5, c(1, 3:5), sep = "-")
 plot_par(configtag, runtag, simutags)
@@
 plot_smr(configtag, runtag, simutags)
+
 ```

### 2021-05-09 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_mixnormal_041821.Rmd
added smr-heidi results, scripts matching v0.1.25

--- 2021-05-09 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_041621.Rmd
@@
 ---
-title: "Simulations to test ctwas summary stats version, 45k samples"
+title: "Simulations (mix normal prior), ctwas_rss, 45k samples"
 output: html_document
@@
 
-      To simulate phenotype data, first we impute gene expression based on expression models, then we set gene/SNP pi1 and PVE, get rough effect size for causal SNPs and genes and simulate phenotype under the sparse model with spike and slab prior.
+      To simulate phenotype data, first we impute gene expression based on expression models, then we set gene/SNP pi1 and PVE, get rough effect size for causal SNPs and genes and simulate phenotype under the mixture of normal distributions.
+
+* The prior distribution for causal effect: 
+```{r }
+
+nmix <- 4
+pi_k <- rep(1/nmix, nmix)
+
+ra <- 2**(1: length(pi_k))
+sa2_kr <-  ra / (pi_k * sum(ra))
+
+cat("number of mixtures: ", nmix, "\n")
+cat("variance of each component:", sa2_kr, "\n")
+
+hist(mixtools::rnormmix(2000, lambda = pi_k,
+                        mu = rep(0, length(pi_k)),
+                        sigma = sa2_kr), breaks = 100, main = "prior distribution", xlab = "")
+```
+
       Then we performed GWAS for all SNPs and get z scores for each by univariate linear regression.
@@
 `truth`: the true parameters,  `selected_truth`: the truth in selected regions that were used to estimate parameters, `ctwas`: ctwas estimated parameters (using summary statistics as input).
+
 We run FUSION following default settings and adjust p values by BH method to get expected FDP.
-We ran coloc for all genes with TWAS p < 1e-4. We use PP4 (SNP associate with both traits).
+we ran coloc for all genes with TWAS p < 1e-4. We use PP4 (SNP associate with both traits).
 We ran SMR+HEIDI, using eQTL summary statistics GTEx v.7. We filter the results by requiring p_HEIDI > 0.05. The plots 
@@
 
-simutags <- paste(2, 1:5, sep = "-")
+simutags <- paste(2, c(1:5), sep = "-")
 plot_par(configtag, runtag, simutags)
@@
 plot_smr(configtag, runtag, simutags)
-
 ```

### 2021-06-08 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_041621.Rmd
matching ctwas v0.1.29

@@
 ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
-outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210418/" # /
-comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210418_compare/"
+outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416/" # /
+comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416_compare/"
 runtag = "ukb-s80.45-adi"
@@
 
-simutags <- paste(1, c(1:5), sep = "-")
+simutags <- paste(1, c(1,2,5), sep = "-")
 plot_par(configtag, runtag, simutags)

### 2021-06-15 code/run_ctwas_rss_LDR2.R
LDR2.R

@@ -1,0 +1,44 @@
+library(ctwas)
+# source('~/causalTWAS/causal-TWAS/code/debug_ctwas_susieI_rss2.R')
+
+args = commandArgs(trailingOnly=TRUE)
+if (length(args) < 7) {
+  stop(" 7 arguments:
+       * zscore file name
+       * ld R directory
+       * weight
+       * config file name (.R)
+       * out expr z file name
+       * out file name
+       * outputdir", call.=FALSE)
+}
+
+z_snp <- data.table::fread(args[1], header = T)
+data.table::setnames(z_snp, old = c("alt", "ref", "t.value"), new = c("A1", "A2", "z"))
+z_snp <- z_snp[, c("id", "A1", "A2", "z")]
+ld_R_dir <- args[2]
+weight <- args[3]
+outname.e <- args[5]
+outname <- args[6]
+outputdir <- args[7]
+
+
+ld_regions_custom <- "/home/simingz/ctwas/inst/extdata/example_regions.bed"
+ncore <- 1
+ncore.rerun <- 1
+prob_single <- 0.8
+thin <- 1
+max_snp_region <- 1e4
+
+source(args[4]) # config
+
+# get gene z score
+ld_exprfs <- paste0(outputdir, "/", outname.e, "_chr", 1:22, ".expr.gz")
+load(file = paste0(outputdir, "/", outname.e, "_z_gene.Rd"))
+
+load( paste0(outputdir, "/", outname, ".s2.susieIrssres.Rd"))
+group_prior <- group_prior_rec[, ncol(group_prior_rec)]
+group_prior_var <- group_prior_var_rec[, ncol(group_prior_var_rec)]
+
+# run ctwas_rss last step
+ctwas_rss(z_gene, z_snp, ld_exprfs, ld_pgenfs = NULL, ld_R_dir = ld_R_dir, ld_regions = "EUR", ld_regions_custom = ld_regions_custom, thin = thin, max_snp_region = max_snp_region, outputdir = outputdir, outname = outname, ncore = ncore, ncore.rerun = ncore.rerun, prob_single = prob_single,  group_prior = group_prior, group_prior_var = group_prior_var, estimate_group_prior = F, estimate_group_prior_var = F)

### 2021-07-22 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22_041521.Rmd
bonferroni Fusion, description

@@
 source('~/causalTWAS/causal-TWAS/analysis/summarize_smr_plots.R')
+source('~/causalTWAS/causal-TWAS/analysis/summarize_mrjti_plots.R')
 source('~/causalTWAS/causal-TWAS/code/qqplot.R')
@@
 ```
-
 # Analysis description
@@
   * **GWAS summary statistics** we simulated summary statistics data with different causal gene/SNP proportion and PVE. To simulate this data, we need the following:
-      * **genotype data** we used is from UKB biobank, randomly selecting `r as.character(n.ori)` samples.
-We then filtered samples based on relatedness, ethics and other qc metrics, that ended up with **n = `r as.character(n)` samples**.
+      * **genotype data** we used is from UKB biobank, randomly selecting `r as.character(n.ori)` samples. We then filtered samples.
+      The columns selected for filtering samples are as follows:
+        * sex (31)
+        * UK Biobank assessment centre (54)
+        * age (21022)
+        * genetic ethnic grouping (22006)
+        * genetic sex (22001)
+        * genotype measurement batch (22000)
+        * missingness (22005)
+        * genetic PCs (22009-0.1 - 22009-0.40)
+        * genetic relatedness pairing (22011)
+        * genetic kinship (22021)
+        * outliers (22027)
+
+        We filtered the samples based on the following criteria:  
+        1.
+Remove all rows in which one or more of the values are missing, aside from the in the "outlier" and "relatedness_genetic" columns.
+This step removes any samples that are not marked as being "White British".
+        2. Remove rows with mismatches between self-reported and genetic sex  
+        3. Remove "missingness" and "heterozygosity" outliers as defined by UK Biobank. 
+        4. Remove any individuals have at least one relative  
+        5. Remove any individuals that have close relatives identified from the "relatendess" calculations that weren't already identified using the kinship calculations.
+
+        This ended up with **n = `r as.character(n)` samples**.
 We use SNP genotype data from chr 1 to chr 22 combined from UKB.
-There are **total = `r as.character(p)` SNPs**.
-
+We select SNPs with minor allele frequency > 0.05 and SNPs with at least 95% genotyping rate (5% missing).
+There are **total = `r as.character(p)` SNPs** passed these filters and go into our analysis.
     * **Expression models** The one we used in this analysis is GTEx Adipose tissue v7 dataset.
@@
   * **LD genotype reference**
-  We used 2k samples from UKbiobank with same ethnicity to serve as the LD reference.
+  We used the genotype of 2k samples from UKbiobank (randomly selected from the samples used in simulations) to serve as the LD reference.
 
@@
 
-We run FUSION following default settings and adjust p values by BH method to get expected FDP.
+We run FUSION following default settings and adjust p values by BH method to get expected FDP. We have also used Bonferroni correction for p values.
+
 We ran coloc for all genes with TWAS p < 1e-4. We use PP4 (SNP associate with both traits).
@@
 are based on SMR p value adjusted by BH method to get expected FDP.
+
+We have tried to run MR-JTI.
+The results have higher false postive rate than TWAS.
+MR-JTI requires the SNPs be pruned before the analysis.
+It also requires that a gene has at least 20 eQTLs.
+This resulted in very few genes going into the analysis.
+Most genes left are in polymorphism dense regions, such as the MHC regions.
+I ran MR-JTI for top genes in TWAS, around 30-40% of them should be real.
+However, only a few genes pass MR-JTI’s 20 eQTL requirements and only 1 or 2 (5%) genes are real.
+We are showing MR-JTI results on this page.
 
@@
     f2 <- ncausalFUSIONp_plot(phenofs, fusioncolocfs)
-    f3 <- caliPP4_plot(phenofs, fusioncolocfs, twas.p = 0.05/J)
-    f4 <- ncausalPP4_plot(phenofs, fusioncolocfs, twas.p = 0.05/J)
+    f3 <- caliFUSIONbon_plot(phenofs, fusioncolocfs)
+    f4 <- ncausalFUSIONbon_plot(phenofs, fusioncolocfs)
+    f5 <- caliPP4_plot(phenofs, fusioncolocfs, twas.p = 0.05/J)
+    f6 <- ncausalPP4_plot(phenofs, fusioncolocfs, twas.p = 0.05/J)
     gridExtra::grid.arrange(f1, f2, ncol=2)
     gridExtra::grid.arrange(f3, f4, ncol=2)
+    gridExtra::grid.arrange(f5, f6, ncol=2)
 }
@@
 }
+
+plot_mrjti <- function(configtag, runtag,  simutags){
+    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+    mrfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.mrjti.result")
+
+    f1 <- caliMR_plot(phenofs, mrfs)
+    f2 <- ncausalMR_plot(phenofs, mrfs)
+    gridExtra::grid.arrange(f1, f2, ncol=2)
+}
 ```
@@
 plot_smr(configtag, runtag, simutags)
+plot_mrjti(configtag, runtag, simutags)
 

### 2021-07-22 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22_mixnormal_041721.Rmd
bonferroni Fusion, description

@@
   * **GWAS summary statistics** we simulated summary statistics data with different causal gene/SNP proportion and PVE. To simulate this data, we need the following:
-      * **genotype data** we used is from UKB biobank, randomly selecting `r as.character(n.ori)` samples.
-We then filtered samples based on relatedness, ethics and other qc metrics, that ended up with **n = `r as.character(n)` samples**.
-We use SNP genotype data from chr 1 to chr 22 combined from UKB.
-There are **total = `r as.character(p)` SNPs**.
+      * **genotype data** We used **n = `r as.character(n)` samples**. It is the same as used [here](simulation-ctwas-ukbWG-gtex.adipose_s40.22_041521.html)
 
@@
 
-We run FUSION following default settings and adjust p values by BH method to get expected FDP.
+We run FUSION following default settings and adjust p values by BH method to get expected FDP. We have also used Bonferroni correction for p values.
 we ran coloc for all genes with TWAS p < 1e-4. We use PP4 (SNP associate with both traits).
@@
     f2 <- ncausalFUSIONp_plot(phenofs, fusioncolocfs)
-    f3 <- caliPP4_plot(phenofs, fusioncolocfs, twas.p = 0.05/J)
-    f4 <- ncausalPP4_plot(phenofs, fusioncolocfs, twas.p = 0.05/J)
+    f3 <- caliFUSIONbon_plot(phenofs, fusioncolocfs)
+    f4 <- ncausalFUSIONbon_plot(phenofs, fusioncolocfs)
+    f5 <- caliPP4_plot(phenofs, fusioncolocfs, twas.p = 0.05/J)
+    f6 <- ncausalPP4_plot(phenofs, fusioncolocfs, twas.p = 0.05/J)
     gridExtra::grid.arrange(f1, f2, ncol=2)
     gridExtra::grid.arrange(f3, f4, ncol=2)
+    gridExtra::grid.arrange(f5, f6, ncol=2)
+}
 
-}
 plot_focus <- function(configtag, runtag,  simutags){

### 2021-07-22 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_041621.Rmd
LDR2.R
bonferroni Fusion, description

@@
   * **LD genotype reference**
-  We randomly selected 2000 samples to serve as the LD reference.
+  We used the genotype of 2k samples from UKbiobank (randomly selected from the samples used in simulations) to serve as the LD reference.
 
@@
 We run FUSION following default settings and adjust p values by BH method to get expected FDP.
+
+We run FUSION following default settings and adjust p values by BH method to get expected FDP. We have also used Bonferroni correction for p values.
+
 We ran coloc for all genes with TWAS p < 1e-4. We use PP4 (SNP associate with both traits).
@@
     f2 <- ncausalSMRp_plot(phenofs, smrfs)
+    gridExtra::grid.arrange(f1, f2, ncol=2)
+}
+
+plot_mrjti <- function(configtag, runtag,  simutags){
+    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+    mrfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.mrjti.result")
+
+    f1 <- caliMR_plot(phenofs, mrfs)
+    f2 <- ncausalMR_plot(phenofs, mrfs)
     gridExtra::grid.arrange(f1, f2, ncol=2)
@@
 plot_smr(configtag, runtag, simutags)
+```
 
+# ctwas results (LDR) 
+Using the R LD matrices as the LD reference input for ctwas, instead of the genotype type of randomly subsetted samples.
+The LD R matrices are generated using all 300k samples that passed the our filtering criteria (generated by Wes Crouse).
+A R matrice is provided for each LD block region.
+
+```{r param3, fig.height=3, fig.width=8, message=F}
+configtag <- "1_LDR"
+simutags <- paste(1, c(1,2,5), sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
+
+simutags <- paste(2, 1:5, sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
+
+simutags <- paste(3, 1:5, sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
+
+simutags <- paste(4, 1:5, sep = "-")
+plot_par(configtag, runtag, simutags)
+plot_PIP(configtag, runtag, simutags)
 ```

### 2021-07-22 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_mixnormal_041821.Rmd
bonferroni Fusion, description

@@
   * **GWAS summary statistics** we simulated summary statistics data with different causal gene/SNP proportion and PVE. To simulate this data, we need the following:
-      * **genotype data** we used is from UKB biobank, randomly selecting `r as.character(n.ori)` samples.
-We then filtered samples based on relatedness, ethics and other qc metrics, that ended up with **n = `r as.character(n)` samples**.
-We use SNP genotype data from chr 1 to chr 22 combined from UKB.
-There are **total = `r as.character(p)` SNPs**.
+      * **genotype data** We used **n = `r as.character(n)` samples**. It is the same as used [here](simulation-ctwas-ukbWG-gtex.adipose_s40.22_041621.html)
 

### 2021-07-22 analysis/simulation-ctwas-ukbWG-gtex.adipose_visual.Rmd
bonferroni Fusion, description

@@
 ```{r output dir}
-outputdir <- "~/causalTWAS/simulations/simulation_ctwas_rss_20210117/"
+outputdir <- "~/causalTWAS/simulations/simulation_ctwas_rss_20210415/"
 
@@
 ```{r sim 2, out.width = "98%", out.height="95%"}
-chrom <- 7
-runtag <- "ukb-s80.45-adi"
-simutag <- "2-1"
+chrom <- 10
+runtag <- "ukb-s40.22-adi"
+simutag <- "3-2"
 configtag <- 1

### 2021-07-22 analysis/temp.Rmd
bonferroni Fusion, description

--- 2021-05-09 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_041621.Rmd
@@
 ---
-title: "Simulations to test ctwas summary stats version, 45k samples"
+title: "Untitled"
 output: html_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE,  message= F, warning = F)
-```
-
-```{r libs}
-library(ctwas)
-library(data.table)
-suppressMessages({library(plotly)})
-library(tidyr)
-library(plyr)
-library(stringr)
-source("~/causalTWAS/causal-TWAS/analysis/summarize_ctwas_plots.R")
-source('~/causalTWAS/causal-TWAS/analysis/summarize_twas-coloc_plots.R')
-source('~/causalTWAS/causal-TWAS/analysis/summarize_focus_plots.R')
-source('~/causalTWAS/causal-TWAS/analysis/summarize_smr_plots.R')
-source('~/causalTWAS/causal-TWAS/code/qqplot.R')
-```
-
-```{r files}
-pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt"
-ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
-outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210418/" # /
-comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210418_compare/"
-runtag = "ukb-s80.45-adi"
-configtags = 1
-simutags = paste(rep(1:2, each = length(1:5)), 1:5, sep = "-")
-
-pgenfs <- read.table(pgenfn, header = F, stringsAsFactors = F)[,1]
-pvarfs <- sapply(pgenfs, prep_pvar, outputdir = outputdir)
-
-ld_pgenfs <- read.table(ld_pgenfn, header = F, stringsAsFactors = F)[,1]
-ld_pvarfs <- sapply(ld_pgenfs, prep_pvar, outputdir = outputdir)
-
-pgens <- lapply(1:length(pgenfs), function(x) prep_pgen(pgenf = pgenfs[x],pvarf = pvarfs[x]))
+knitr::opts_chunk$set(echo = TRUE)
 ```
@@
 ```{r parameters}
-n.ori <- 80000 # number of samples
-n <- pgenlibr::GetRawSampleCt(pgens[[1]])
-p <- sum(unlist(lapply(pgens, pgenlibr::GetVariantCt))) # number of SNPs
+n.ori <- 40000 # number of samples
+n <- 100
+p <-100 # number of SNPs
 J <- 8021 # number of genes
@@
 Lastly, run susie with given L for all regions and for all genes and SNPs using estimated prior and prior variance.
-
-# Power estimation
-
-```{r power}
-simutag <- "1-1"
-niter <- 1000
-snp.p <- 5e-8
-gene.p <- 1e-5
-source(paste0(outputdir, "simu", simutag, "_param.R"))
-load(paste0(outputdir, runtag, "_simu", simutag, "-pheno.Rd"))
-```
-
-We select run `r simutag`  as an example.
-
-* For SNPs. $\pi_1 =$ `r pi_theta` , effect size = `r phenores[["batch"]][[1]][["sigma_theta"]]`, PVE = `r phenores$param$pve.snp.truth`. Power at `r as.character(snp.p)` p value cutoff:
-
-```{r snp}
-load("data/power_s80.45.Rd")
-# p1 <- pow(niter, n, phenores[["batch"]][[1]][["sigma_theta"]], snp.p)
-print(p1)
-```
-
-* For genes. $\pi_1 = `r pi_beta`$ , effect size = `r phenores[["batch"]][[1]][["sigma_beta"]]`, PVE = `r phenores$param$pve.gene.truth`. Power at `r as.character(gene.p)` p value cutoff:
-
-```{r gene}
-# p2 <- pow(niter, n, phenores[["batch"]][[1]][["sigma_beta"]], gene.p)
-print(p2)
-# save(p1,p2, file = "data/power_s80.45.Rd")
-```
-
-# GWAS/TWAS p value distribution
-
-```{r pvalue}
-simutag <- "1-1"
-chrom <- 1
-source(paste0(outputdir, "simu", simutag, "_param.R"))
-load(paste0(outputdir, runtag, "_simu", simutag, "-pheno.Rd"))
-```
-We select run `r simutag`  as an example.
-
-* For genes. $\pi_1 = `r pi_beta`$ , effect size = `r phenores[["batch"]][[1]][["sigma_beta"]]`, PVE = `r phenores$param$pve.gene.truth`.
-TWAS p values and qqplot:
-
-```{r gene-p, message=F, fig.width= 7, fig.height=7}
-exprgwasf <- paste0(outputdir, runtag, "_simu", simutag, ".exprgwas.txt.gz")
-
-exprvarf <- paste0(outputdir, runtag, "_chr", chrom, ".exprvar")
-exprid <- read_exprvar(exprvarf)[, "id"]
-cau <- as.matrix(exprid[phenores[["batch"]][[chrom]][["idx.cgene"]]])
-pdist_plot(exprgwasf, chrom, cau)
-```
-```{r gene-qq, fig.width= 5, fig.height=5}
-exprgwas <- fread(exprgwasf, header =T)
-gg_qqplot(exprgwas$PVALUE)
-```
-
-* For SNPs. $\pi_1 =$ `r pi_theta` , effect size = `r phenores[["batch"]][[1]][["sigma_theta"]]`, PVE = `r phenores$param$pve.snp.truth`.
-GWAS p values and qqplot:
-
-```{r snp-p, fig.width= 7, fig.height=7}
-snpgwasf <-  paste0(outputdir, runtag, "_simu", simutag, ".snpgwas.txt.gz")
-
-pvarf <- pvarfs[chrom]
-snpid <- read_pvar(pvarf)[, "id"]
-cau <- as.matrix(snpid[phenores[["batch"]][[chrom]][["idx.cSNP"]]])
-pdist_plot(snpgwasf, chrom, cau, thin = 0.1)
-```
-
-```{r snp-qq, fig.width= 5, fig.height=5}
-snpgwas <- fread(snpgwasf, header =T)
-gg_qqplot(snpgwas$PVALUE, thin = 0.1)
-```
-
-# `ctwas` results
-
-Results: Each row shows parameter estimation results from 5 simulation runs with similar settings (i.e.
-pi1 and PVE for genes and SNPs).
-Results from each run were represented by one dot, dots with the same color come from the same run.
-`truth`: the true parameters,  `selected_truth`: the truth in selected regions that were used to estimate parameters, `ctwas`: ctwas estimated parameters (using summary statistics as input).
-We run FUSION following default settings and adjust p values by BH method to get expected FDP. 
-We ran coloc for all genes with TWAS p < 1e-4. We use PP4 (SNP associate with both traits).
-We ran SMR+HEIDI, using eQTL summary statistics GTEx v.7. We filter the results by requiring p_HEIDI > 0.05. The plots 
-are based on SMR p value adjusted by BH method to get expected FDP.
-
-```{r func}
-plot_par <- function(configtag, runtag, simutags){
-  source(paste0(outputdir, "config", configtag, ".R"))
-  phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
-  susieIfs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".s2.susieIrssres.Rd")
-  susieIfs2 <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".s2.susieIrss.txt")
-
-  mtx <- show_param(phenofs, susieIfs, susieIfs2, thin = thin)
-  par(mfrow=c(1,3))
-  cat("simulations ", paste(simutags, sep=",") , ": ")
-  cat("mean gene PVE:", mean(mtx[, "PVE.gene_truth"]), ",", "mean SNP PVE:", mean(mtx[, "PVE.SNP_truth"]), "\n")
-  plot_param(mtx)
-}
-
-plot_PIP <- function(configtag, runtag,  simutags){
-   phenofs <- paste0(outputdir, "ukb-s80.45-adi", "_simu", simutags, "-pheno.Rd")
-   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
-
-   f1 <- caliPIP_plot(phenofs, susieIfs)
-   f2 <- ncausal_plot(phenofs, susieIfs) 
-   gridExtra::grid.arrange(f1, f2, ncol =2)
-}
-
-plot_fusion_coloc <- function(configtag, runtag,  simutags){
-    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
-    fusioncolocfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.coloc.result")
-
-    f1 <- caliFUSIONp_plot(phenofs, fusioncolocfs)
-    f2 <- ncausalFUSIONp_plot(phenofs, fusioncolocfs)
-    f3 <- caliPP4_plot(phenofs, fusioncolocfs, twas.p = 0.05/J)
-    f4 <- ncausalPP4_plot(phenofs, fusioncolocfs, twas.p = 0.05/J)
-    gridExtra::grid.arrange(f1, f2, ncol=2)
-    gridExtra::grid.arrange(f3, f4, ncol=2)
-}
-
-plot_focus <- function(configtag, runtag,  simutags){
-    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
-    focusfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.focus.tsv")
-
-    f1 <- califocusPIP_plot(phenofs, focusfs)
-    f2 <- ncausalfocusPIP_plot(phenofs, focusfs)
-    gridExtra::grid.arrange(f1, f2, ncol=2)
-}
-
-plot_smr <- function(configtag, runtag,  simutags){
-    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
-    smrfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.smr")
-
-    f1 <- caliSMRp_plot(phenofs, smrfs)
-    f2 <- ncausalSMRp_plot(phenofs, smrfs)
-    gridExtra::grid.arrange(f1, f2, ncol=2)
-}
-```
-
-```{r param2, fig.height=3, fig.width=8, message=F}
-
-configtag <- 1
-runtag = "ukb-s80.45-adi"
-
-simutags <- paste(1, c(1:5), sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(2, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(3, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(4, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(5, c(1, 3:5), sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(6, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(7, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(8, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(9, 1:4, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(10, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-```

### 2021-07-22 code/run_MR-JTI.R

@@ -1,0 +1,99 @@
+library(data.table)
+args = commandArgs(trailingOnly=TRUE)
+if (length(args) != 5) {
+  stop(" 5 arguments:
+       * prune ld score file
+       * gwas file
+       * eqtl summary statistics
+       * TWAS results files
+       * out file name", call.=FALSE)
+}
+
+mrjti <- "/home/simingz/causalTWAS/MR-JTI/mr/MR-JTI.r"
+
+
+ldfile <- args[1]
+gwas <- args[2]
+eqtl <- args[3]
+twas <- args[4]
+outname <- args[5]
+
+harmonize_gwas_eqtl <- function(gwas, eqtl){
+  snpnames <- intersect(gwas$id, eqtl$id)
+  if (length(snpnames) != 0){
+    g.idx <- match(snpnames, gwas$id)
+    e.idx <- match(snpnames, eqtl$id)
+    qc <- ctwas:::allele.qc(gwas[g.idx, ]$alt, gwas[g.idx, ]$ref,
+                    eqtl[e.idx, ]$alt, eqtl[e.idx, ]$ref)
+
+    ifflip <- qc[["flip"]]
+    ifremove <- !qc[["keep"]]
+
+    flip.idx <- g.idx[ifflip]
+    gwas[flip.idx, c("alt", "ref")] <- gwas[flip.idx, c("ref", "alt")]
+    gwas[flip.idx, "Estimate"] <- - gwas[flip.idx, "Estimate"]
+
+    remove.idx <- g.idx[ifremove]
+    if (length(remove.idx) != 0) {
+      gwas <- gwas[-remove.idx, ,drop = F]
+    }
+  }
+  return(gwas)
+}
+
+# save.image("temp.Rd")
+# GWAS
+gwas <- fread(gwas, header = T)
+
+#  GTEx: the eQTL effect allele is the ALT allele
+eqtl <- fread(eqtl, header = T)
+eqtl <- eqtl[, c("rs_id_dbSNP147_GRCh37p13", "gene_name", "ref", "alt", "slope", "slope_se", "pval_nominal")]
+
+# harmonize gwas and eqtl
+gwas <- harmonize_gwas_eqtl(gwas, eqtl)
+
+# merge gwas, eqtl and ldinfo
+
+ldfiles <- fread(ldfile, header = F)
+
+ldinfolist <- list()
+for (i in 1:22){
+  prune <- fread(ldfiles[[i,1]], header = F)
[…]

### 2021-07-31 analysis/Paper_figures_simulation.Rmd
simulation paper figures

---
title: "Paper figures (Simulation)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup dir}
library(ctwas)
library(data.table)
source("~/causalTWAS/causal-TWAS/analysis/summarize_basic_plots.R")
source("~/causalTWAS/causal-TWAS/analysis/summarize_ctwas_plots.R")
source("~/causalTWAS/causal-TWAS/analysis/ld.R")

outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416/"
comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416_compare/"
runtag = "ukb-s80.45-adi"
configtag = 1
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt"
exprfn = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416//ukb-s80.45-adi.expr.txt"

pgenfs <- read.table(pgenfn, header = F, stringsAsFactors = F)[,1]
pvarfs <- sapply(pgenfs, prep_pvar, outputdir = outputdir)
pgens <- lapply(1:length(pgenfs), function(x) prep_pgen(pgenf = pgenfs[x],pvarf = pvarfs[x]))
exprfs <- read.table(exprfn, header = F, stringsAsFactors = F)[,1]
exprvarfs <- sapply(exprfs, prep_exprvar)

n <- pgenlibr::GetRawSampleCt(pgens[[1]])
p <- sum(unlist(lapply(pgens, pgenlibr::GetVariantCt))) # number of SNPs
J <- 8021 # number of genes

colorsall <- c("#7fc97f", "#beaed4", "#fdc086")
```

## Parameter estimation
### Parameter vs. true value

Each plot show one parameter: pi.gene, pi.gene/pi.SNP (enrichment), effectsize.gene, PVE.gene, horizontal bar shows true values, and multiple dots, each show one setting.

```{r paramplotfunc}

plot_single <- function(mtxlist, truecol, estcol, ...){
  truth <- do.call(rbind, lapply(1:length(mtxlist), function(x) cbind(x, mean(mtxlist[[x]][, truecol]))))
  est <- do.call(rbind, lapply(1:length(mtxlist), function(x) cbind(x, mtxlist[[x]][, estcol])))

  col = est[,1]
  est[,1] <- jitter(est[,1])

  plot(est, pch = 19, xaxt = "n", xlab="" ,frame.plot=FALSE, col = colorsall[col], ...)
  axis(side=1, at=1:2, labels = c("setting 1", "setting 2"), tick = F)
  #text(x=1:length(mtxlist), 0, labels = paste0("temp",1:length(mtxlist)), xpd = T, pos =1)
  for (t in 1:nrow(truth)){
    row <- truth[t,]
    segments(row[1]-0.2, row[2] , row[1] + 0.2, row[2],
       col = colorsall[t], lty = par("lty"), lwd = 2, xpd = FALSE)
  }
  grid()
}

plot_par <- function(configtag, runtag, simutaglist){
  mtxlist <- list()
  for (group in 1:length(simutaglist)){
    simutags <- simutaglist[[group]]
    source(paste0(outputdir, "config", configtag, ".R"))
    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
    susieIfs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".s2.susieIrssres.Rd")
    susieIfs2 <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".s2.susieIrss.txt")
    mtxlist[[group]] <- show_param(phenofs, susieIfs, susieIfs2, thin = thin)
  }

  # pi1 
  par(mfrow=c(1,4))
  plot_single(mtxlist, truecol = "pi1.gene_truth", estcol = "pi1.gene_est", ylab ="gene pi1", ylim = c(0,0.02), xlim = c(0.8,2.4))
  plot_single(mtxlist, truecol = "enrich_truth", estcol = "enrich_est",ylab ="gene enrichment", ylim = c(0,120), xlim = c(0.8,2.4) )
  plot_single(mtxlist, truecol = "sigma.gene_truth", estcol = "sigma.gene_est", ylab = "gene effect size", ylim = c(0.01, 0.03), xlim = c(0.8,2.4))
  plot_single(mtxlist, truecol = "PVE.gene_truth", estcol = "PVE.gene_est", ylab ="gene PVE", ylim = c(0, 0.1), xlim = c(0.8,2.4))

  return(mtxlist)
}

```

Figure (supplementary):
```{r paramplot, fig.width= 8, fig.height= 3}
simutaglist = list(paste(4, 1:5, sep = "-"), paste(10, 1:5, sep="-"))
mtxlist <- plot_par(configtag, runtag, simutaglist)
```

Back up figures

I have in total 10 x 2 = 20 settings to use generate the parameter figures.

### Comparison of PVE.gene vs. MESC estimates.
Leave it to Kevin.

## PIP calibration
### PIP plot under different settings.

```{r pipplotfunc}
plot_PIP <- function(configtag, runtag,  simutags, ...){
   phenofs <- paste0(outputdir, "ukb-s80.45-adi", "_simu", simutags, "-pheno.Rd")
   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")

   f1 <- caliPIP_plot(phenofs, susieIfs, ...) 
   return(f1)
}
```

```{r pipplot, fig.width= 8, fig.height= 4}
simutaglist = list(paste(4, 1:5, sep = "-"), paste(10, 1:5, sep="-"))
f1 <- plot_PIP(configtag, runtag, simutaglist[[1]], main = "setting 1")
f2 <-  plot_PIP(configtag, runtag, simutaglist[[2]], main = "setting 2")
gridExtra::grid.arrange(f1, f2, ncol =2)
```

### Comparison with other methods
Bar plot: each bar shows the number of genes, colored by causal status. Use a different color for each method. The method and cut off values:
* ctwas: PIP 0.8
* FUSION fdr: 0.05
* FUSION bonferroni: 0.05
* COLOC PP4: 0.8
* FOCUS PIP: 0.8
* SMR FDR: 0.05
* SMR HEIDI: HEIDI p > 0.05, SMR FDR < 0.05

Multiple bar plots, different settings: high gene power and low gene power.
```{r ncausalplotfunc}

get_ncausal_df <- function(pfiles, cau, cut = 0.8, 
                           method = c("ctwas", "fusionfdr", "fusionbon","coloc", "focus", "smr", "smrheidi")){
  df <- NULL
  for (i in 1:length(pfiles)) {
    res <- fread(pfiles[i], header = T)
    # res <- res[complete.cases(res),]
    if (method == "ctwas"){
        res <- data.frame(res[res$type  =="gene", ])
        res$ifcausal <- ifelse(res$id %in% cau[[i]], 1, 0)
        res <- res[res$susie_pip > cut,]
    } else if (method == "fusionfdr"){
        res$FDR <- p.adjust(res$TWAS.P, method = "fdr")
        res <- res[res$FDR < cut,]
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
    } else if  (method == "fusionbon"){
        res$FDR <- p.adjust(res$TWAS.P, method = "bonferroni")
        res <- res[res$FDR < cut,]
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
    } else if  (method == "coloc"){
        res <- res[res$COLOC.PP4 > cut,]
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
    } else if (method == "focus"){
        res <- res[res$mol_name != "NULL",]
        res <- res[res$pip > cut, ] 
        res$ifcausal <- ifelse(res$mol_name %in% cau[[i]], 1, 0)
    } else if (method == "smr"){
        res$FDR <- p.adjust(res$p_SMR, method = "fdr")
        res <- res[res$FDR < cut,]
        res$ifcausal <- ifelse(res$probeID %in% cau[[i]], 1, 0)
    } else if (method == "smrheidi"){
        res <- res[res$p_HEIDI > 0.05,]
        res$FDR <- p.adjust(res$p_SMR, method = "fdr")
        res <- res[res$FDR < cut,]
        res$ifcausal <- ifelse(res$probeID %in% cau[[i]], 1, 0)
    } else{
      stop("no such method")
    }

    df.rt <- rbind(c(nrow(res[res$ifcausal == 0, ]), 0, i),
                   c(nrow(res[res$ifcausal == 1, ]), 1, i))
    df <- rbind(df, df.rt)
  }
  colnames(df) <- c("count", "ifcausal", "runtag")
  df <- data.frame(df)
  df$method <- method
  return(df)
}

plot_ncausal <- function(configtag, runtag,  simutags, colors, ...){
  phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
  cau <- lapply(phenofs, function(x) {load(x);get_causal_id(phenores)})

  susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
  fusioncolocfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.coloc.result")
  focusfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.focus.tsv")
  smrfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.smr")

  ctwas_df <- get_ncausal_df(susieIfs, cau= cau, cut = 0.8, method ="ctwas")
  fusionfdr_df <- get_ncausal_df(fusioncolocfs, cau= cau, cut = 0.05, method = "fusionfdr")
  fusionbon_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = 0.05, method = "fusionbon")
  coloc_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = 0.8, method = "coloc")
  focus_df <- get_ncausal_df(focusfs , cau= cau, cut = 0.8, method = "focus")
  smr_df <- get_ncausal_df(smrfs, cau= cau, cut = 0.05, method = "smr")
  smrheidi_df <- get_ncausal_df(smrfs, cau= cau, cut = 0.05, method = "smrheidi")

  df <- rbind(ctwas_df, fusionfdr_df, fusionbon_df, coloc_df, focus_df, smr_df, smrheidi_df)
  df$ifcausal <- df$ifcausal + as.numeric(as.factor(df$method))*10
  df$ifcausal <- as.factor(df$ifcausal)
  fig <- ggbarplot(df, x = "method", y = "count", add = "mean_se", fill = "ifcausal", palette = colors, legend = "none", ...) + grids(linetype = "dashed")
  fig
}

```

```{r ncausalplot, fig.width= 10, fig.height= 4}
colset = c("#ebebeb", "#bebada" , "#ebebeb", "#fb8072", "#ebebeb","#ffffb3","#ebebeb", "#8dd3c7", "#ebebeb","#8dd3c7", "#ebebeb", "#87CEFA", "#ebebeb","#87CEFA") # coloc, ctwas, focus, fusionbon, fusionfdr, smrheidi, smr
f1 <- plot_ncausal(configtag, runtag,  simutaglist[[1]], colors = colset, ylim= c(0,150), main = "setting 1")
f2 <- plot_ncausal(configtag, runtag,  simutaglist[[2]], colors = colset, ylim= c(0,150), main = "setting 2")
gridExtra::grid.arrange(f1, f2, ncol=2)
```

## Examples
### FP example 1
cTWAS avoids the FP error. In this case, the false positive gene (from TWAS) is caused by LD of eQTLs with a causal gene's eQTLs.

```{r regionplotfunc}
plot_region <- function(runtag, simutag, configtag, chr, startpos = NULL, endpos = NULL){
 susief <- paste0(outputdir, runtag, "_simu", simutag, "_config", configtag, ".susieIrss.txt")
  pf <- paste0(outputdir, runtag, "_simu",simutag)
  phenof <- paste0(outputdir, runtag, "_simu", simutag, "-pheno.Rd")

  a <- fread(susief, header =T)
  b1 <- fread(paste0(pf, ".snpgwas.txt.gz"), header =T)
  setnames(b1, old = "pos", new = "p0")
  b2 <- fread(paste0(pf, ".exprgwas.txt.gz"), header =T)
  b <- rbind(b1, b2, fill = T)

  a <- a[a$chrom == chr,]
  b <- b[b$chrom == chr,]
  if (!is.null(startpos)){
    a <- a[a$pos > startpos & a$pos < endpos]
    b <- b[b$p0 > startpos & b$p0 < endpos]
  }

  load(phenof)
  cau <- get_causal_id(phenores)

  a <- merge(a, b, by = "id", all = T)
  a$ifcausal <- ifelse(a$id %in% cau, 1, 0)
  a[is.na(a$type),"type"] <- "SNP"

  a[, "PVALUE"] <- -log10(a[, "PVALUE"])
  a$r2max <- get_ld(ids =a$id, phenores = phenores, pgenfs = pgenfs, exprfs = exprfs, chrom = chr)

  r2cut <- 0.4

  layout(matrix(1:2, ncol = 1), widths = 1, heights = c(1.5,1.5), respect = FALSE)
  par(mar = c(0, 4.1, 4.1, 2.1))
  plot(a[a$type=="SNP"]$p0, a[a$type == "SNP"]$PVALUE, pch = 19, xlab="Genomic position" ,frame.plot=FALSE, col = "white", ylim= c(-0.1,1.1), ylab = "ctwas PIP", xaxt = 'n')
  grid()
  points(a[a$type=="SNP"]$p0, a[a$type == "SNP"]$susie_pip, pch = 21, xlab="Genomic position", bg = colorsall[1])
  points(a[a$type=="SNP" & a$r2max > r2cut, ]$p0, a[a$type == "SNP"  & a$r2max >r2cut]$susie_pip, pch = 21, bg = "purple")
  points(a[a$type=="SNP" & a$ifcausal == 1, ]$p0, a[a$type == "SNP" & a$ifcausal == 1]$susie_pip, pch = 21, bg = "salmon")
  points(a[a$type=="gene" ]$p0, a[a$type == "gene" ]$susie_pip, pch = 22, bg = colorsall[1], cex = 2)
  points(a[a$type=="gene" & a$r2max > r2cut, ]$p0, a[a$type == "gene"  & a$r2max > r2cut]$susie_pip, pch = 22, bg = "purple", cex = 2)
  points(a[a$type=="gene" & a$ifcausal == 1, ]$p0, a[a$type == "gene" & a$ifcausal == 1]$susie_pip, pch = 22, bg = "salmon", cex = 2)
  legend(min(a$p0), y= 1.3 ,c("Gene", "SNP"), pch = c(21,22), title="shape legend", bty ='n', cex =0.8, title.adj = 0)
  legend(min(a$p0), y= 0.8 ,c("Causal", "Noncausal, R2 > 0.4", "Noncausal, R2 <= 0.4"), pch = 19, col = c("salmon", "purple", colorsall[1]), title="color legend", bty ='n', cex =0.8, title.adj = 0)
  par(mar = c(4.1, 4.1, 0.5, 2.1))
  plot(a[a$type=="SNP"]$p0, a[a$type == "SNP"]$PVALUE, pch = 21, xlab="Genomic position" ,frame.plot=FALSE, bg = colorsall[1], ylab = "TWAS -log10(p value)", panel.first = grid(), ylim =c(0, max(a$PVALUE)*1.2))
  points(a[a$type=="SNP" & a$r2max > r2cut ]$p0, a[a$type == "SNP"  & a$r2max > r2cut]$PVALUE, pch = 21, bg = "purple")
  points(a[a$type=="SNP" & a$ifcausal == 1, ]$p0, a[a$type == "SNP" & a$ifcausal == 1]$PVALUE, pch = 21, bg = "salmon")
  points(a[a$type=="gene" ]$p0, a[a$type == "gene" ]$PVALUE, pch = 22, bg = colorsall[1], cex = 2)
  points(a[a$type=="gene" & a$r2max > r2cut, ]$p0, a[a$type == "gene"  & a$r2max > r2cut]$PVALUE, pch = 22, bg = "purple", cex = 2)
  points(a[a$type=="gene" & a$ifcausal == 1, ]$p0, a[a$type == "gene" & a$ifcausal == 1]$PVALUE, pch = 22, bg = "salmon", cex = 2)
  abline(h=-log10(0.05/J), col ="red", lty = 2)
  return(a)
}
```

```{r regionplot1}
simutag <- "4-4"
a <- plot_region(runtag, simutag, configtag, chr = 4, startpos = 43965045, endpos = 45189157)
```

### FP example 2
cTWAS avoids the FP error. In this case, the false positive gene (from TWAS) is caused by LD of eQTLs with a causal SNP nearby.

```{r regionplot2}
simutag <- "4-1"
a <- plot_region(runtag, simutag, configtag, chr = 5, startpos = 152867774, endpos = 153773088)
```

### TP example
cTWAS is able to find true positives. This gene has one eQTL, cTWAS choose this gene because it uses a  prior favoring genes, it didn't reach significance level by TWAS after bonferron correction.

```{r regionplot3}
simutag <- "4-4"
a <- plot_region(runtag, simutag, configtag, chr = 1, startpos = 37549183 , endpos =38731847)
```

### 2021-07-31 analysis/simulation-ctwas-ukbWG-gtex.adipose_visual.Rmd
simulation paper figures

--- 2021-02-10 analysis/simulation-ctwas-ukbWG-gtex.adipose_visual.Rmd
@@
 
-```{r output dir}
-outputdir <- "~/causalTWAS/simulations/simulation_ctwas_rss_20210117/"
-
+```{r functions file}
 source("~/causalTWAS/causal-TWAS/analysis/summarize_ctwas_plots.R")
 source("~/causalTWAS/causal-TWAS/analysis/visual_twas_plots.R")
-```
-
-```{r functions file}
 
@@
 
-  a <- merge(a, b, by = "name")
+  a <- merge(a, b, by = "name", all = T)
   a$ifcausal <- ifelse(a$name %in% cau, "Causal", "Non causal" )
+  a[is.na(a$type),"type"] <- "SNP"
 
   if (thin !=1){
-    b <- a[a$type == "gene" | (a$type == "SNP" & a$ifcausal == "Causal") | 
-             (a$type == "SNP" & a$ifcausal == "Non causal" & rbinom(nrow(a), 1, thin) ==1)]
+    a <- a[ !is.na(a$type) | a$ifcausal == "Causal" | 
+             (a$ifcausal == "Non causal" & rbinom(nrow(a), 1, thin) ==1)]
   }
@@
   colnames(a)[colnames(a) == "susie_pip"] <-  "susiePIP"
-  colnames(a)[colnames(a) == "pos"] <-  "p0"
 
@@
 
-## Simulation
-
-First three rows are for genes, last three rows are for SNPs.
+## Scatter plot
 
 ```{r sim 2, out.width = "98%", out.height="95%"}
-chrom <- 7
+outputdir <- "~/causalTWAS/simulations/simulation_ctwas_rss_20210416/"
 runtag <- "ukb-s80.45-adi"
-simutag <- "2-1"
+simutags <- paste(4, 1:5, sep = "-")
 configtag <- 1
 
-susief <- paste0(outputdir, runtag, "_simu", simutag, "_config", configtag, ".susieIrss.txt")
-pf <- paste0(outputdir, runtag, "_simu",simutag)
-phenof <- paste0(outputdir, runtag, "_simu", simutag, "-pheno.Rd")
-plot_all(susief, pf, phenof, chrom, thin = 0.01)
+susiefs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".susieIrss.txt")
+pfs <- paste0(outputdir, runtag, "_simu",simutags)
+gwasfs <- paste0(pfs, ".exprgwas.txt.gz")
+phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+scatter_plot_PIP_p(phenofs, susiefs, gwasfs, main ="PIP-p")
 ```
+
+## Zoom view
+FP example, LD with causal gene, GNPDA2
+
+```{r sim 2, out.width = "98%", out.height="95%"}
+chrom <- 4
+runtag <- 4
+plot_all(susiefs[runtag], pfs[runtag], phenofs[runtag], chrom, thin = 0.01)
+```
+
+FP example: LD with causal SNP, GALNT10
+```{r sim 2, out.width = "98%", out.height="95%"}
+runtag <- 1
+chrom <- 5
+plot_all(susiefs[runtag], pfs[runtag], phenofs[runtag], chrom, thin = 0.01)
+```
+
+True positive example: c1orf109
+```{r sim 2, out.width = "98%", out.height="95%"}
+runtag <- 4
+chrom <- 1
+plot_all(susiefs[runtag], pfs[runtag], phenofs[runtag], chrom, thin = 0.01)
+```

### 2021-08-04 analysis/Paper_figures_simulation.Rmd
update simulation figure

@@
 
-Each plot show one parameter: pi.gene, pi.gene/pi.SNP (enrichment), effectsize.gene, PVE.gene, horizontal bar shows true values, and multiple dots, each show one setting.
+In our simulations, the SNP PVE is always set to 0.5 in different settings.
+The gene PVE is shown as in figures.
+The number of causal SNPs is always set to 2.5e * $10^{-4}$.
+In the two settings shown below, number of samples is 45k.
+For other details about our simulation settings and procedures, please see [here](simulation-ctwas-ukbWG-gtex.adipose_s80.45_041621.html).
+Note, setting 1 is the high power setting and setting 2 is a low power setting, I will change the setting names manually later.
+
+Each plot show one parameter: pi.gene, pi.gene/pi.SNP (enrichment), effectsize.gene, PVE.gene.
+Horizontal bar shows mean true values across the 5 simulations with similar setting parameters.
+The results by ctwas for each simulation is shown by dots.
 
@@
 simutaglist = list(paste(4, 1:5, sep = "-"), paste(10, 1:5, sep="-"))
-f1 <- plot_PIP(configtag, runtag, simutaglist[[1]], main = "setting 1")
-f2 <-  plot_PIP(configtag, runtag, simutaglist[[2]], main = "setting 2")
+f1 <- plot_PIP(configtag, runtag, simutaglist[[1]], main = "high gene PVE")
+f2 <-  plot_PIP(configtag, runtag, simutaglist[[2]], main = "low gene PVE")
 gridExtra::grid.arrange(f1, f2, ncol =2)
@@
 colset = c("#ebebeb", "#bebada" , "#ebebeb", "#fb8072", "#ebebeb","#ffffb3","#ebebeb", "#8dd3c7", "#ebebeb","#8dd3c7", "#ebebeb", "#87CEFA", "#ebebeb","#87CEFA") # coloc, ctwas, focus, fusionbon, fusionfdr, smrheidi, smr
-f1 <- plot_ncausal(configtag, runtag,  simutaglist[[1]], colors = colset, ylim= c(0,150), main = "setting 1")
-f2 <- plot_ncausal(configtag, runtag,  simutaglist[[2]], colors = colset, ylim= c(0,150), main = "setting 2")
+f1 <- plot_ncausal(configtag, runtag,  simutaglist[[1]], colors = colset, ylim= c(0,150), main = "high gene PVE")
+f2 <- plot_ncausal(configtag, runtag,  simutaglist[[2]], colors = colset, ylim= c(0,150), main = "low gene PVE")
 gridExtra::grid.arrange(f1, f2, ncol=2)
@@
   points(a[a$type=="gene" & a$ifcausal == 1, ]$p0, a[a$type == "gene" & a$ifcausal == 1]$susie_pip, pch = 22, bg = "salmon", cex = 2)
-  legend(min(a$p0), y= 1.3 ,c("Gene", "SNP"), pch = c(21,22), title="shape legend", bty ='n', cex =0.8, title.adj = 0)
+  legend(min(a$p0), y= 1.3 ,c("Gene", "SNP"), pch = c(22,21), title="shape legend", bty ='n', cex =0.8, title.adj = 0)
   legend(min(a$p0), y= 0.8 ,c("Causal", "Noncausal, R2 > 0.4", "Noncausal, R2 <= 0.4"), pch = 19, col = c("salmon", "purple", colorsall[1]), title="color legend", bty ='n', cex =0.8, title.adj = 0)

### 2021-08-09 analysis/Paper_figures_simulation.Rmd
paper figure, examples rerun ctwas with full SNPs

@@
 pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt"
+ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
 exprfn = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416//ukb-s80.45-adi.expr.txt"
 
+ld_pgenfs <- read.table(ld_pgenfn, header = F, stringsAsFactors = F)[,1]
 pgenfs <- read.table(pgenfn, header = F, stringsAsFactors = F)[,1]
@@
 ```{r regionplotfunc}
-plot_region <- function(runtag, simutag, configtag, chr, startpos = NULL, endpos = NULL){
- susief <- paste0(outputdir, runtag, "_simu", simutag, "_config", configtag, ".susieIrss.txt")
+plot_region <- function(runtag, simutag, configtag, chr, startpos = NULL, endpos = NULL, rerun_ctwas = F, plot_ceqtl = F,
+                        plot_eqtl = F){
   pf <- paste0(outputdir, runtag, "_simu",simutag)
@@
 
-  a <- fread(susief, header =T)
   b1 <- fread(paste0(pf, ".snpgwas.txt.gz"), header =T)
@@
   b <- rbind(b1, b2, fill = T)
+  b <- b[b$chrom == chr,]
 
+  susief <- paste0(outputdir, runtag, "_simu", simutag, "_config", configtag, ".susieIrss.txt")
+  a <- fread(susief, header =T)
   a <- a[a$chrom == chr,]
-  b <- b[b$chrom == chr,]
+
   if (!is.null(startpos)){
@@
 
+  a <- merge(a, b, by = "id", all = T)
+  a[is.na(a$type),"type"] <- "SNP"
+
+  if (isTRUE(rerun_ctwas)){
+      ld_exprfs <- paste0(outputdir, runtag, "_simu", simutag, "_chr", 1:22, ".expr.gz")
+      load(paste0(outputdir, runtag, "_simu", simutag, "_config", configtag, ".s2.susieIrssres.Rd"))
+      source(paste0(outputdir, "config", configtag, ".R"))
+      group_prior <- group_prior_rec[, ncol(group_prior_rec)] 
+      group_prior[2] <- group_prior[2] * thin
+      group_prior_var <- group_prior_var_rec[, ncol(group_prior_var_rec)]
+      temp_reg <- data.frame("chr" = paste0("chr",chr), "start" = startpos, "stop" = endpos)
+      write.table(temp_reg, file= "temp_reg.txt" , row.names=F, col.names=T, sep="\t", quote = F)
+      z_gene <-  a[a$type == "gene", c("id", "t.value")]
+      colnames(z_gene) <- c("id", "z")
+      z_snp <-  a[a$type == "SNP", c("id", "alt", "ref", "t.value")]
+      colnames(z_snp) <- c("id", "A1", "A2", "z")
+      ctwas_rss(z_gene, z_snp, ld_exprfs= ld_exprfs, ld_pgenfs = ld_pgenfs, ld_R_dir = NULL, ld_regions_custom = "temp_reg.txt", thin = 1, outputdir = ".", outname = "temp", ncore = 1, ncore.rerun = 1, prob_single = 0,  group_prior = group_prior, group_prior_var = group_prior_var, estimate_group_prior = F, estimate_group_prior_var = F, recover_strand_ambig_z = F)
+      a2 <- fread("temp.susieIrss.txt", header = T)
+      a <- merge(a2, b, by = "id", all = T)
+      # file.remove("temp.susieIrss.txt")
+      # file.remove("temp.temp.susieIrssres.Rd")
+      # file.remove("temp.regions.txt")
+      # file.remove("temp_reg.txt")
+   }
+
   load(phenof)
@@
 
-  a <- merge(a, b, by = "id", all = T)
   a$ifcausal <- ifelse(a$id %in% cau, 1, 0)
@@
   a[, "PVALUE"] <- -log10(a[, "PVALUE"])
-  a$r2max <- get_ld(ids =a$id, phenores = phenores, pgenfs = pgenfs, exprfs = exprfs, chrom = chr)
+  a$r2max <- get_ld2(ids =a$id, phenores = phenores, pgenfs = pgenfs, exprfs = exprfs, chrom = chr)
 
@@
   points(a[a$type=="gene" & a$ifcausal == 1, ]$p0, a[a$type == "gene" & a$ifcausal == 1]$susie_pip, pch = 22, bg = "salmon", cex = 2)
+  if (isTRUE(plot_ceqtl)){
+      for (cgene in a[a$type=="gene" & a$ifcausal == 1, ]$id){
+     load(paste0(tools::file_path_sans_ext(exprfs[a[a$id == cgene, "chrom.x"][[1]]]), "qc.Rd"))
+     eqtls <- rownames(wgtlist[[cgene]])
+     points(a[a$id %in% eqtls,]$p0, rep( -0.15, nrow(a[a$id %in% eqtls,])), pch = "|", col = "salmon", cex = 1.5)
+    }
+  }
+
+  if (isTRUE(plot_eqtl)){
+    for (cgene in a[a$type=="gene" & a$PVALUE >5, ]$id){
+     load(paste0(tools::file_path_sans_ext(exprfs[a[a$id == cgene, "chrom.x"][[1]]]), "qc.Rd"))
+     eqtls <- rownames(wgtlist[[cgene]])
+     points(a[a$id %in% eqtls,]$p0, rep( -0.15, nrow(a[a$id %in% eqtls,])), pch = "|", col = "black", cex = 1.5)
+    }
+  }
+
   legend(min(a$p0), y= 1.3 ,c("Gene", "SNP"), pch = c(22,21), title="shape legend", bty ='n', cex =0.8, title.adj = 0)
@@
   abline(h=-log10(0.05/J), col ="red", lty = 2)
+
   return(a)
@@
 
-```{r regionplot1}
+```{r regionplot1, echo=FALSE,results='hide',fig.keep='all', message=F}
 simutag <- "4-4"
-a <- plot_region(runtag, simutag, configtag, chr = 4, startpos = 43965045, endpos = 45189157)
+a <- plot_region(runtag, simutag, configtag, chr = 4, startpos = 43965045, endpos = 45189157,  plot_ceqtl = T, rerun_ctwas = T)
 ```
@@
 
-```{r regionplot2}
-simutag <- "4-1"
-a <- plot_region(runtag, simutag, configtag, chr = 5, startpos = 152867774, endpos = 153773088)
+```{r regionplot2, echo=FALSE,results='hide',fig.keep='all', message=F}
+runtag = "ukb-s80.45-adi"
+simutag <- "3-2"         
+a <- plot_region(runtag, simutag, configtag, chr = 19, startpos = 16824416, endpos = 17559862, rerun_ctwas = T, plot_eqtl = T)
 ```
@@
 
-```{r regionplot3}
+```{r regionplot3, echo=FALSE,results='hide',fig.keep='all', message=F}
 simutag <- "4-4"
-a <- plot_region(runtag, simutag, configtag, chr = 1, startpos = 37549183 , endpos =38731847)
+a <- plot_region(runtag, simutag, configtag, chr = 1, startpos = 37549183 , endpos =38731847, rerun_ctwas = T)
 ```

### 2021-08-09 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_041621.Rmd
paper figure, examples rerun ctwas with full SNPs

@@
 2. Run ctwas_rss
-We used LDetect to define regions.
 `ctwas_rss` algorithm first runs on all regions to get rough estimate for gene and SNP prior.
 Then run on small regions (having small probablities of having > 1 causal signals based on rough estimates) to get more accurate estimate.
-To lower computational burden, we downsampled SNPs (0.1), estimate parameters and convert back to orginal scale.
-Lastly, run susie with given L for all regions and for all genes and SNPs using estimated prior and prior variance.
+To lower computational burden, we downsampled SNPs (0.1) to estimate parameters.
+With the estimated parameters, we then run susie for all regions using both genes and downsampled SNPs with specified $L$.
+After this, for regions with strong gene signals, we rerun susie with full SNPs using specified $L$.
+
+Configurations
+
+`ld_regions ='EUR'`, We used LDetect to define regions. To match UKbiobank data, we use the 'EUR' population
+
+`thin = 0.1`, downsampled SNPs to 1/10 for parameter estimation step
+
+`niter1 =3`, run `niter1 =3` iterations first to get some rough parameter estimates.
+
+`prob_single = 0.8`, the probability of a region having at most 1 singal has to be at least 0.8 to be selected for the parameter estimation step.
+This probability is obtained by using the PIPs from the first few iterations.
+
+`niter2 = 30`, run `niter2 = 30` for parameter estimation step
+
+`L = 3` , after parameter estimation, for running susie for all regions.
+
+`group_prior = NULL`, the intiating prior parameters we used for running susie for each region is uniform prior for genes and SNPs.
+
+`group_prior_var = NULL`, the intiating prior variance parameters we used for running susie for each region follows susie_rss's default (50).
+
+`max_SNP_region = 5000`, the maximum number of SNPs for re-running susie on strong gene signal regions is 5000.
 

### 2021-08-09 analysis/simulation-ctwas-ukbWG-gtex.adipose_visual.Rmd
paper figure, examples rerun ctwas with full SNPs

@@
 source("~/causalTWAS/causal-TWAS/analysis/visual_twas_plots.R")
+library(plyr)
 
@@
 runtag <- "ukb-s80.45-adi"
-simutags <- paste(4, 1:5, sep = "-")
+simutags <- paste(3, 1:5, sep = "-")
 configtag <- 1
@@
 FP example, LD with causal gene, GNPDA2
-
+under 4-1 to 4-5
 ```{r sim 2, out.width = "98%", out.height="95%"}
@@
 ```{r sim 2, out.width = "98%", out.height="95%"}
-runtag <- 1
-chrom <- 5
+runtag <- 5
+chrom <- 1
+plot_all(susiefs[runtag], pfs[runtag], phenofs[runtag], chrom, thin = 0.01)
+```
+
+under 10-1 to 10-5
+FP example: LD with causal SNP
+```{r sim 2, out.width = "98%", out.height="95%"}
+runtag <- 5
+chrom <- 1
 plot_all(susiefs[runtag], pfs[runtag], phenofs[runtag], chrom, thin = 0.01)

### 2021-08-09 analysis/summarize_ctwas_plots.R

@@ -50,7 +50,7 @@ scatter_plot_PIP_p <- function(phenofs, pipfs, gwasfs, main ="PIP-p"){
     df <- rbind(df, res)
   }
 
-  df <- rename(df, c( "PVALUE" = "TWAS.p"))
+  colnames(df)[colnames(df)== "PVALUE"] <- "TWAS.p"
   df[,"TWAS.p"] <- -log10(df[, "TWAS.p"])
 
   # df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1),to=c("darkgreen", "salmon"))

### 2023-03-22 analysis/Paper_Suppl_figures_simulation.Rmd
revision

---
title: "Paper figures (Simulation)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup dir}
library(ctwas)
library(data.table)
source("~/causalTWAS/causal-TWAS/analysis/summarize_basic_plots.R")
source("~/causalTWAS/causal-TWAS/analysis/summarize_ctwas_plots.R")
source("~/causalTWAS/causal-TWAS/analysis/ld.R")

outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210418/"
comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210418_compare/"
runtag = "ukb-s80.45-adi"
configtag = 1
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
exprfn = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416//ukb-s80.45-adi.expr.txt"

ld_pgenfs <- read.table(ld_pgenfn, header = F, stringsAsFactors = F)[,1]
pgenfs <- read.table(pgenfn, header = F, stringsAsFactors = F)[,1]
pvarfs <- sapply(pgenfs, prep_pvar, outputdir = outputdir)
pgens <- lapply(1:length(pgenfs), function(x) prep_pgen(pgenf = pgenfs[x],pvarf = pvarfs[x]))
exprfs <- read.table(exprfn, header = F, stringsAsFactors = F)[,1]
exprvarfs <- sapply(exprfs, prep_exprvar)

n <- pgenlibr::GetRawSampleCt(pgens[[1]])
p <- sum(unlist(lapply(pgens, pgenlibr::GetVariantCt))) # number of SNPs
J <- 8021 # number of genes

colorsall <- c("#7fc97f", "#beaed4", "#fdc086")
```

## Parameter estimation
### Parameter vs. true value

In our simulations, the SNP PVE is always set to 0.5 in different settings. The gene PVE is shown as in figures. The number of causal SNPs is always set to 2.5e * $10^{-4}$. In the two settings shown below, number of samples is 45k. For other details about our simulation settings and procedures, please see [here](simulation-ctwas-ukbWG-gtex.adipose_s80.45_mixnormal_041821.html).

Each plot show one parameter: pi.gene, pi.gene/pi.SNP (enrichment), PVE.gene, PVE.snp. Horizontal bar shows mean true values across the 5 simulations with similar setting parameters. The results by ctwas for each simulation is shown by dots.

```{r paramplotfunc}
require(latex2exp)
plot_single <- function(mtxlist, truecol, estcol, xlabels = NULL, ...){
  truth <- do.call(rbind, lapply(1:length(mtxlist), function(x) cbind(x, mean(mtxlist[[x]][, truecol]))))
  est <- do.call(rbind, lapply(1:length(mtxlist), function(x) cbind(x, mtxlist[[x]][, estcol])))

  col = est[,1]
  est[,1] <- jitter(est[,1])

  plot(est, pch = 21, xaxt = "n", xlab="" ,frame.plot=FALSE, col= "darkgreen", bg = colorsall[col], ...)
  axis(side=1, at=1, labels = xlabels, tick = F)
  #text(x=1:length(mtxlist), 0, labels = paste0("temp",1:length(mtxlist)), xpd = T, pos =1)
  for (t in 1:nrow(truth)){
    row <- truth[t,]
    segments(row[1]-0.2, row[2] , row[1] + 0.2, row[2],
       col = colorsall[t], lty = par("lty"), lwd = 2, xpd = FALSE)
  }
  grid()
}

get_params <- function(configtag, runtag, simutaglist){
    mtxlist <- list()
  for (group in 1:length(simutaglist)){
    simutags <- simutaglist[[group]]
    source(paste0(outputdir, "config", configtag, ".R"))
    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
    susieIfs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".s2.susieIrssres.Rd")
    susieIfs2 <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".s2.susieIrss.txt")
    mtxlist[[group]] <- show_param(phenofs, susieIfs, susieIfs2, thin = thin)
  }
    return(mtxlist)
}

plot_PIP <- function(configtag, runtag,  simutags, ...){
   phenofs <- paste0(outputdir, "ukb-s80.45-adi", "_simu", simutags, "-pheno.Rd")
   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")

   f1 <- caliPIP_plot(phenofs, susieIfs, ...) 
   return(f1)
}
```

## Parameters + PIP Figure
```{r paramplot, fig.width= 8, fig.height= 18}
simutaglist = lapply(1:10, function(x) paste(x, 1:5, sep ="-"))
mtxlist <- get_params(configtag, runtag, simutaglist)
pdf(file = "temp.pdf", height = 25, width =5)
par(mfrow=c(length(simutaglist),4))
for (i in 1:length(simutaglist)){
  plot_single(mtxlist[i], truecol = "enrich_truth", estcol = "enrich_est", main = TeX("Enrichment $\\pi_G/\\pi_V$"), ylim = c(0,200), xlim = c(0.8,1.2))
  plot_single(mtxlist[i], truecol = "pi1.gene_truth", estcol = "pi1.gene_est", main = TeX("Percent causal, $\\pi_G$"), ylim = c(0,0.1), xlim = c(0.8,1.2))
   plot_single(mtxlist[i], truecol = "PVE.gene_truth", estcol = "PVE.gene_est", main = "Gene PVE", ylim = c(0,1), xlim = c(0.8,1.2))
    plot_single(mtxlist[i], truecol = "PVE.SNP_truth", estcol = "PVE.SNP_est",main = "SNP PVE", ylim = c(0,1), xlim = c(0.8,1.2))
}
dev.off()
```

```{r PIPplot, fig.width= 8, fig.height= 18}
simutaglist = c(lapply(1:4, function(x) paste(x, 1:5, sep ="-")),
                list(paste(5, c(1,3,4,5), sep = "-")),
                lapply(6:10, function(x) paste(x, 1:5, sep ="-")))
pdf(file = "temp2.pdf", height = 25, width =3)
plotlist <- list()
for (i in 1:length(simutaglist)){
  # for (i in 1:2){
  plotlist[[i]] <- plot_PIP(configtag, runtag, simutaglist[[i]], main = "")
}
gridExtra::grid.arrange(plotlist[[1]], plotlist[[2]],
                        plotlist[[3]], plotlist[[4]],
                        plotlist[[5]], plotlist[[6]],
                        plotlist[[7]], plotlist[[8]],
                        plotlist[[9]], plotlist[[10]],
                        ncol =1)
dev.off()
```

### 2023-03-22 analysis/Paper_figures_simulation.Rmd
revision

@@
 ```{r paramplotfunc}
-
-plot_single <- function(mtxlist, truecol, estcol, ...){
+require(latex2exp)
+plot_single <- function(mtxlist, truecol, estcol, xlabels = c("setting 1", "setting 2"), ...){
   truth <- do.call(rbind, lapply(1:length(mtxlist), function(x) cbind(x, mean(mtxlist[[x]][, truecol]))))
@@
   plot(est, pch = 19, xaxt = "n", xlab="" ,frame.plot=FALSE, col = colorsall[col], ...)
-  axis(side=1, at=1:2, labels = c("setting 1", "setting 2"), tick = F)
+  axis(side=1, at=1:2, labels = xlabels, tick = F)
   #text(x=1:length(mtxlist), 0, labels = paste0("temp",1:length(mtxlist)), xpd = T, pos =1)
@@
 
-plot_par <- function(configtag, runtag, simutaglist){
+get_params <- function(configtag, runtag, simutaglist){
     mtxlist <- list()
@@
   }
+    return(mtxlist)
+}
 
-  # pi1 
+plot_par <- function(mtxlist){
   par(mfrow=c(1,4))
@@
   plot_single(mtxlist, truecol = "PVE.gene_truth", estcol = "PVE.gene_est", ylab ="gene PVE", ylim = c(0, 0.1), xlim = c(0.8,2.4))
-
-  return(mtxlist)
 }
@@
 
-Figure (supplementary):
+## Parameters Figure
 ```{r paramplot, fig.width= 8, fig.height= 3}
 simutaglist = list(paste(4, 1:5, sep = "-"), paste(10, 1:5, sep="-"))
-mtxlist <- plot_par(configtag, runtag, simutaglist)
+mtxlist <- get_params(configtag, runtag, simutaglist)
+plot_par(mtxlist)
 ```
 
-Back up figures
+### Supplementary-parameters
+```{r paramplot-suppl, fig.width= 8, fig.height= 3}
+plot_par_snps <- function(mtxlist){
+  par(mfrow=c(1,3))
+  plot_single(mtxlist, truecol = "pi1.SNP_truth", estcol = "pi1.SNP_est", xlabels = c("high gene PVE", "low gene PVE"), ylab =TeX('Percent causal, $\\pi_V$'), ylim = c(0,0.0005), xlim = c(0.8,2.4))
+  plot_single(mtxlist, truecol = "sigma.SNP_truth", estcol = "sigma.SNP_est",xlabels = c("high gene PVE", "low gene PVE"), ylab = "Variant effect size", ylim = c(0.01, 0.03), xlim = c(0.8,2.4))
+  plot_single(mtxlist, truecol = "PVE.SNP_truth", estcol = "PVE.SNP_est",xlabels = c("high gene PVE", "low gene PVE"), ylab ="Variant PVE", ylim = c(0, 0.8), xlim = c(0.8,2.4))
+}
 
-I have in total 10 x 2 = 20 settings to use generate the parameter figures.
-
-### Comparison of PVE.gene vs. MESC estimates.
-Leave it to Kevin.
+plot_par_snps(mtxlist)
+```
 
@@
 gridExtra::grid.arrange(f1, f2, ncol =2)
+```
+
+### Supplementary PIP plot
+
+```{r pipplot, fig.width= 8, fig.height= 16}
+simutaglist = lapply(c(1:3,5:9), function(x) paste(x, 1:5, sep ="-"))
+mtxlist <- get_params(configtag, runtag, simutaglist)
+simutaglist = c(list(paste(1, c(1,2,5), sep = "-")), 
+                lapply(c(2,3,5,6,7,8,9), function(x) paste(x, 1:5, sep ="-")))
+
+pdf(file = "temp3.pdf", height = 10, width =10)
+plotlist <- list()
+for (i in 1:length(simutaglist)){
+  plotlist[[i]] <- plot_PIP(configtag, runtag, simutaglist[[i]], main = 
+                              TeX(sprintf("\\overset{$\\pi_G$= %s, $PVE_G$ = %s;}{$\\pi_V$= %s, $PVE_G$ = %s}",
+                                  signif(mean(mtxlist[[i]][,"pi1.gene_truth"]), digits=2), 
+                                  signif(mean(mtxlist[[i]][,"PVE.gene_truth"]), digits=2),
+                                  signif(mean(mtxlist[[i]][,"pi1.SNP_truth"]), digits=2),
+                                  signif(mean(mtxlist[[i]][,"PVE.SNP_truth"]), digits=2))))
+}
+gridExtra::grid.arrange(plotlist[[1]], plotlist[[2]],
+                        plotlist[[3]], plotlist[[4]],
+                        plotlist[[5]], plotlist[[6]],
+                        plotlist[[7]], plotlist[[8]],ncol = 3)
+dev.off()
 ```
@@
 
+## Scatter plot
+```{r scatterplot}
+runtag <- "ukb-s80.45-adi"
+simutags <- paste(3, 2, sep = "-")
+phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+gwasfs <- paste0(outputdir, runtag, "_simu",simutags, ".exprgwas.txt.gz")
+pipfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
+df <- scatter_plot_PIP_p(phenofs, pipfs, gwasfs, main ="PIP-p")
+```
+
 ## Examples
@@
   if (isTRUE(plot_ceqtl)){
+    gi <- 0
       for (cgene in a[a$type=="gene" & a$ifcausal == 1, ]$id){
@@
      eqtls <- rownames(wgtlist[[cgene]])
-     points(a[a$id %in% eqtls,]$p0, rep( -0.15, nrow(a[a$id %in% eqtls,])), pch = "|", col = "salmon", cex = 1.5)
+     points(a[a$id %in% eqtls,]$p0, rep( -0.15 - gi, nrow(a[a$id %in% eqtls,])), pch = "|", col = "salmon", cex = 1.5)
+     gi <- gi + 0.2
     }
@@
   if (isTRUE(plot_eqtl)){
-    for (cgene in a[a$type=="gene" & a$PVALUE >5, ]$id){
+     gi <- 0
+    for (cgene in a[a$type=="gene" & a$PVALUE >4, ]$id){
      load(paste0(tools::file_path_sans_ext(exprfs[a[a$id == cgene, "chrom.x"][[1]]]), "qc.Rd"))
      eqtls <- rownames(wgtlist[[cgene]])
-     points(a[a$id %in% eqtls,]$p0, rep( -0.15, nrow(a[a$id %in% eqtls,])), pch = "|", col = "black", cex = 1.5)
+     points(a[a$id %in% eqtls,]$p0, rep( -0.15-gi, nrow(a[a$id %in% eqtls,])), pch = "|", col = "black", cex = 1.5)
+     gi <- gi + 0.2
     }
@@
 ```
+
+### coloc false positives
+```{r regionplot4, echo=FALSE,results='hide',fig.keep='all', message=F}
+simutag <- "3-4"
+a <- plot_region(runtag, simutag, configtag, chr = 1, startpos = 64000000 , endpos =66000000, rerun_ctwas = F, plot_eqtl = T )
+```
+```{r regionplot4, echo=FALSE,results='hide',fig.keep='all', message=F}
+simutag <- "3-4"
+a <- plot_region(runtag, simutag, configtag, chr = 1, startpos = 66000000 , endpos =70000000, rerun_ctwas = F, plot_eqtl = T )
+```
+```{r regionplot4, echo=FALSE,results='hide',fig.keep='all', message=F}
+simutag <- "3-4"
+a <- plot_region(runtag, simutag, configtag, chr = 1, startpos = 155000000 , endpos =170000000, rerun_ctwas = F, plot_eqtl = T )
+```
+```{r regionplot4, echo=FALSE,results='hide',fig.keep='all', message=F}
+simutag <- "3-4"
+a <- plot_region(runtag, simutag, configtag, chr = 5, startpos = 47000000 , endpos =50000000, rerun_ctwas = F, plot_eqtl = T )
+```

### 2023-03-22 analysis/simulation-ctwas-ukbWG-gtex.adipose_visual.Rmd
revision

@@
 
-plot_all <- function(susief, pf, phenof, chr, ldres = NULL, thin = 1){
+plot_all <- function(susief, pf, phenof, colocf = NULL, chr, ldres = NULL, thin = 1){
   a <- fread(susief, header =T)
@@
 
+  if (!is.null(colocf)) {
+    a <- fread(colocf, header = T)
+    setnames(a, old = "ID", new = "name")
+    dt.gene <- merge(dt.gene, a, by ="name")
+  }
+
   fig1 <- plot_susie_pip(dt.gene)  # %>%  add_trace(p, x = list(1), y = list(1), name = " ", showlegend = F)
@@
 
+  if (is.null(colocf)) {
      fig <- subplot(fig1, fig3, fig4, fig6, nrows=4, shareX = T, titleY = T)
+  } else {
+     fig7 <- plot_coloc_pp4(dt.gene)
+     fig <- subplot(fig1, fig3, fig4, fig6, fig7, nrows=5, shareX = T, titleY = T)
+
+  }
 
@@
 ```
+
+# region plot with coloc results
+```{r sim 2, out.width = "98%", out.height="95%"}
+runtag <- "ukb-s80.45-adi"
+i <- 4
+chrom <- 5
+comparedir <- "/project2/mstephens/causalTWAS/simulations/simulation_ctwas_rss_20210416_compare/"
+colocfs <- paste0(comparedir, runtag, "_simu", simutags,".Adipose_Subcutaneous.coloc.result")
+plot_all(susiefs[i], pfs[i], phenofs[i], chrom, colocf = colocfs[i], thin = 0.01)
+```

### 2023-03-22 code/gwas_wc.R

@@ -1,0 +1,95 @@
+library(foreach)
+library(doParallel)
+
+
+#' @description this is a function to run GWAS analysis for small genotype matrix (using glm),for faster results use plink (start from plink files). Credit to http://www.stat-gen.org/tut/tut_analysis.html, requires tabix and bgzip if compress is true.
+#'
+#' @param geno genotype matrix, columns: snps, rows: individuals.
+#' @param pheno phenotype, matrix, ncol = 1
+#' @param snpname a vector of N, correpond to geno type matrix, will be written to output.
+#' @param anno a data frame/matrix, anno column with N rows.
+#' @param compress T/F, if T will be compressed and indexed by tabix. (TODO)
+#'
+GWAA <- function(genof, mode = c("snp", "expr"), pheno, snpname, anno = NULL, outname, outputdir = getwd(), family = gaussian, ncore = 1, nSplits = 10, compress = F){
+
+  cl <- makeCluster(ncore)
+  show(cl)
+  registerDoParallel(cl)
+
+  colnames(pheno) <- "pheno"
+
+  mode <- match.arg(mode)
+
+  columns<- c(colnames(anno), "id", c("Estimate", "Std.Error", "t-value", "PVALUE"))
+  write.table(t(columns), paste0(outputdir, "/",outname), row.names=FALSE, col.names=FALSE, quote=FALSE, sep = "\t")
+
+  nSNPs <- length(snpname)
+  if (nSNPs > 0){
+
+    genosplit <- ceiling(nSNPs/nSplits) # number of SNPs in each subset
+
+    snp.start <- seq(1, nSNPs, genosplit) # index of first SNP in group
+    snp.stop <- pmin(snp.start+genosplit-1, nSNPs) # index of last SNP in group
+
+    res <- foreach(part = 1:length(snp.start),.combine='rbind', .packages = "ctwas")  %dopar% {
+
+      cat(sprintf("GWAS %s%% finished\n", part/length(snp.start)))
+
+      get_geno <- function(genof, mode = mode,
+                           variantidx = NULL, outputdir = NULL){
+        if (mode == "expr"){
+          geno <- read_expr(genof, variantidx = variantidx)
+        } else if (mode == "snp"){
+          pvarf <- prep_pvar(genof, outputdir = outputdir)
+          pgen <- prep_pgen(genof, pvarf)
+          geno <- read_pgen(pgen, variantidx = variantidx)
+        } else {
+          stop("undefined mode")
[…]

### 2023-03-22 code/run_PMR_batch_wc_pi.R

@@ -1,0 +1,279 @@
+library(PMR)
+library(data.table)
+
+args = commandArgs(trailingOnly=TRUE)
+if (length(args) != 8) {
+  stop(" 7 arguments:
+       * LD_R directory
+       * gwas file
+       * eqtl summary statistics
+       * weight dir
+       * gwas sample size
+       * eqtl sample size
+       * out file name", call.=FALSE)
+}
+
+ld_R_dir = as.character(args[1])
+gwas <- as.character(args[2])
+eqtl <- as.character(args[3])
+weightf <- as.character(args[4])
+gwasn = as.numeric(args[5])
+eqtln = as.numeric(args[6])
+outname <- as.character(args[7])
+batch <- as.numeric(args[8])
+
+batch_size <- 1000
+
+#window_size_kb <- NULL
+window_size_kb <- 100
+
+lambda <- 0
+#lambda <- 0.25
+
+#pi <- 1
+pi <- 0.80
+
+
+# simutag <- "4-1"
+#
+# ld_R_dir = "/project2/mstephens/causalTWAS/ukbiobank/ukb_LDR_s80.45/s80.45.2_LDR"
+# gwas <- paste0("/project2/mstephens/causalTWAS/simulations/simulation_ctwas_rss_20210416/ukb-s80.45-adi_simu", simutag, ".snpgwas.txt.gz")
+# eqtl <- "/project2/mstephens/causalTWAS/GTEx_v7_all/Adipose_Subcutaneous.allpairs_processed.txt.gz"
+# weightf <- "/project2/mstephens/causalTWAS/fusion_weights/Adipose_Subcutaneous.pos"
+# gwasn = 45000
+# eqtln = 380
+# outname <- paste0("/project2/mstephens/causalTWAS/simulations/simulation_ctwas_rss_20210416_compare/ukb-s80.45-adi_simu", simutag, ".Adipose_Subcutaneous.pmr.result.batch_2")
+# batch <- 2
+
+####################
+#define function for harmonization
+harmonize_sumstat_ld <- function(sumstats, ldref){
+  snpnames <- intersect(sumstats$id, ldref$id)
+
+  if (length(snpnames) != 0){
+    ss.idx <- which(sumstats$id %in% snpnames)
+    ld.idx <- match(sumstats$id[ss.idx], ldref$id)
+    qc <- ctwas:::allele.qc(sumstats$alt[ss.idx], sumstats$ref[ss.idx],
[…]

### 2023-03-22 code/run_gwas_snp_wc.R

@@ -1,0 +1,77 @@
+
+library(tools)
+library(ctwas)
+
+args = commandArgs(trailingOnly=TRUE)
+
+if (length(args) < 4 ) {
+  stop(" 6 arguments (last two optional):
+       * geno file (txt file)
+       * phenotype file (Rd file)
+       * out file name
+       * outputdir
+       * ncore
+       * nsplits", call.= FALSE)
+}
+
+codedir <- "/project2/mstephens/causalTWAS/causal-TWAS/code/"
+source(paste0(codedir, "gwas_wc.R"))
+
+outputdir <- args[4]
+
+pgenfs <- read.table(args[1], header = F, stringsAsFactors = F)[,1]
+pvarfs <- sapply(pgenfs, prep_pvar, outputdir = outputdir)
+
+phenofile <- args[2]
+load(phenofile)
+pheno <- phenores$Y
+
+outnames <- paste0(args[3], "-B", 1:length(pgenfs), ".snpgwas.txt")
+
+ncore <- as.numeric(args[5])
+if (is.na(args[5])) ncore <- 3
+
+nsplits <- as.numeric(args[6])
+if (is.na(args[6])) nsplits <- 100
+
+for (b in 1:length(pgenfs)){
+
+  outname <- outnames[b]
+
+  # load genotype data
+
+  snpinfo <- read_pvar(pvarfs[b])
+  anno <- snpinfo[, - "id"] # chrom pos alt ref
+
+  GWAA(pgenfs[b], mode = "snp", pheno, snpname = snpinfo$id, anno = anno,
+       outname, outputdir, family = gaussian,
+       ncore = ncore, nSplits = nsplits, compress = F)
+}
+
+# combine all results
+outdflist <- list()
+
+# for (b in 1:length(pgenfs)){
+#   outdflist[[b]] <- read.table(paste0(outnames[b], ".gz"),
+#                                header = T, stringsAsFactors = F,
+#                                comment.char = "")
+# }
+for (b in 1:length(pgenfs)){
+  outdflist[[b]] <- read.table(paste0(outputdir, outnames[b]),
+                               header = T, stringsAsFactors = F,
+                               comment.char = "")
+}
+outdf <- do.call(rbind, outdflist)
+
+outname <-  paste0(outputdir, "/", args[3], ".snpgwas.txt")
+write.table(outdf, file = outname , row.names=F, col.names=T, sep="\t", quote = F)
[…]

### 2023-03-22 code/run_impute_expr_multi_local.R

@@ -1,0 +1,36 @@
+library(ctwas)
+
+args = commandArgs(trailingOnly=TRUE)
+
+args <- c("/project2/mstephens/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt",
+          "/project2/compbio/predictdb/mashr_models/mashr_Liver.db",
+          "/project2/compbio/predictdb/mashr_models/mashr_Whole_Blood.db",
+          "ukb-s80.45-liv_wb",
+          "/project2/mstephens/causalTWAS/simulations/simulation_ctwas_rss_20220125_wc/")
+
+if (length(args) != 5) {
+  stop("5 arguments must be supplied:
+       * genotype file name (multiple file names in a .txt file)
+       * weight_1 file
+       * weight_2 file
+       * outfile name
+       * outputdir", call.=FALSE)
+}
+
+pgenfs <- read.table(args[1], header = F, stringsAsFactors = F)[,1]
+
+weight_1 <- args[2]
+weight_2 <- args[3]
+
+outname <- args[4]
+outputdir <- args[5]
+
+weight <- c(weight_1, weight_2)
+
+exprfs <- impute_expr(pgenfs = pgenfs, weight = weight,
+                           method = "lasso", outputdir = outputdir,
+                           outname = outname)
+
+out <- data.frame("file" = exprfs, stringsAsFactors = F)
+write.table(out, file = paste0(outputdir, "/", outname, ".expr.txt"),
+             row.names=F, col.names=F, sep="\t", quote = F)

### 2023-04-20 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_03212023.Rmd
more simulation settings for low SNP PVE

--- 2021-08-09 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_041621.Rmd
@@
 ---
-title: "Simulations to test ctwas summary stats version, 45k samples"
+title: "Simulations to test ctwas summary stats version, 45k samples, a few new setings required"
 output: html_document
@@
 ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
-outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416/" # /
-comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416_compare/"
+outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230321/" # /
+comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230321_compare/"
 runtag = "ukb-s80.45-adi"
 configtags = 1
-simutags = paste(rep(1:2, each = length(1:5)), 1:5, sep = "-")
+simutags = paste(rep(1:9, each = length(1:5)), 1:5, sep = "-")
 
@@
 ## Data
-
-  * **GWAS summary statistics** we simulated summary statistics data with different causal gene/SNP proportion and PVE. To simulate this data, we need the following:
-      * **genotype data** we used is from UKB biobank, randomly selecting `r as.character(n.ori)` samples. We then filtered samples.
-      The columns selected for filtering samples are as follows:
-        * sex (31)
-        * UK Biobank assessment centre (54)
-        * age (21022)
-        * genetic ethnic grouping (22006)
-        * genetic sex (22001)
-        * genotype measurement batch (22000)
-        * missingness (22005)
-        * genetic PCs (22009-0.1 - 22009-0.40)
-        * genetic relatedness pairing (22011)
-        * genetic kinship (22021)
-        * outliers (22027)
-
-        We filtered the samples based on the following criteria:  
-        1.
-Remove all rows in which one or more of the values are missing, aside from the in the "outlier" and "relatedness_genetic" columns.
-This step removes any samples that are not marked as being "White British".
-        2. Remove rows with mismatches between self-reported and genetic sex  
-        3. Remove "missingness" and "heterozygosity" outliers as defined by UK Biobank. 
-        4. Remove any individuals have at least one relative  
-        5. Remove any individuals that have close relatives identified from the "relatendess" calculations that weren't already identified using the kinship calculations.
-
-        This ended up with **n = `r as.character(n)` samples**.
-We use SNP genotype data from chr 1 to chr 22 combined from UKB.
-We select SNPs with minor allele frequency > 0.05 and SNPs with at least 95% genotyping rate (5% missing).
-There are **total = `r as.character(p)` SNPs** passed these filters and go into our analysis.
-    * **Expression models** The one we used in this analysis is GTEx Adipose tissue v7 dataset.
-This dataset contains ~ 380 samples, **`r as.character(J)` genes** with expression model.
-[FUSION/TWAS](http://gusevlab.org/projects/fusion/) were used to train expression model and we used their lasso results.
-SNPs included in eQTL anlaysis are restricted to cis-locus 500kb on either side of the gene boundary.
-eQTLs are defined as SNPs with abs(effectize) > 1e-8 in lasso results.
-
-      To simulate phenotype data, first we impute gene expression based on expression models, then we set gene/SNP pi1 and PVE, get rough effect size for causal SNPs and genes and simulate phenotype under the sparse model with spike and slab prior.
-      Then we performed GWAS for all SNPs and get z scores for each by univariate linear regression.
-
-  * **LD genotype reference**
-  We used the genotype of 2k samples from UKbiobank (randomly selected from the samples used in simulations) to serve as the LD reference.
-
-  * **Expression models**  
-  We used GTEx Adipose tissue v7 dataset, the same as used for simulating phenotypes.
+The same as 20210416. We tried a few additional settings with small SNP heritablity (0.2) or large gene heritablity (0.5) following the suggestions by reviewers.
 
@@
 
-`L = 3` , after parameter estimation, for running susie for all regions.
+`L = 5` , after parameter estimation, for running susie for all regions.
 
-`group_prior = NULL`, the intiating prior parameters we used for running susie for each region is uniform prior for genes and SNPs.
+`group_prior = NULL`, the initiating prior parameters we used for running susie for each region is uniform prior for genes and SNPs.
 
-`group_prior_var = NULL`, the intiating prior variance parameters we used for running susie for each region follows susie_rss's default (50).
+`group_prior_var = NULL`, the initiating prior variance parameters we used for running susie for each region follows susie_rss's default (50).
 
 `max_SNP_region = 5000`, the maximum number of SNPs for re-running susie on strong gene signal regions is 5000.
-
-# Power estimation
-
-```{r power}
-simutag <- "1-1"
-niter <- 1000
-snp.p <- 5e-8
-gene.p <- 1e-5
-source(paste0(outputdir, "simu", simutag, "_param.R"))
-load(paste0(outputdir, runtag, "_simu", simutag, "-pheno.Rd"))
-```
-
-We select run `r simutag`  as an example.
-
-* For SNPs. $\pi_1 =$ `r pi_theta` , effect size = `r phenores[["batch"]][[1]][["sigma_theta"]]`, PVE = `r phenores$param$pve.snp.truth`. Power at `r as.character(snp.p)` p value cutoff:
-
-```{r snp}
-load("data/power_s80.45.Rd")
-# p1 <- pow(niter, n, phenores[["batch"]][[1]][["sigma_theta"]], snp.p)
-print(p1)
-```
-
-* For genes. $\pi_1 = `r pi_beta`$ , effect size = `r phenores[["batch"]][[1]][["sigma_beta"]]`, PVE = `r phenores$param$pve.gene.truth`. Power at `r as.character(gene.p)` p value cutoff:
-
-```{r gene}
-# p2 <- pow(niter, n, phenores[["batch"]][[1]][["sigma_beta"]], gene.p)
-print(p2)
-# save(p1,p2, file = "data/power_s80.45.Rd")
-```
-
-# GWAS/TWAS p value distribution
-
-```{r pvalue}
-simutag <- "1-1"
-chrom <- 1
-source(paste0(outputdir, "simu", simutag, "_param.R"))
-load(paste0(outputdir, runtag, "_simu", simutag, "-pheno.Rd"))
-```
-We select run `r simutag`  as an example.
-
-* For genes. $\pi_1 = `r pi_beta`$ , effect size = `r phenores[["batch"]][[1]][["sigma_beta"]]`, PVE = `r phenores$param$pve.gene.truth`.
-TWAS p values and qqplot:
-
-```{r gene-p, message=F, fig.width= 7, fig.height=7}
-exprgwasf <- paste0(outputdir, runtag, "_simu", simutag, ".exprgwas.txt.gz")
-
-exprvarf <- paste0(outputdir, runtag, "_chr", chrom, ".exprvar")
-exprid <- read_exprvar(exprvarf)[, "id"]
-cau <- as.matrix(exprid[phenores[["batch"]][[chrom]][["idx.cgene"]]])
-pdist_plot(exprgwasf, chrom, cau)
-```
-```{r gene-qq, fig.width= 5, fig.height=5}
-exprgwas <- fread(exprgwasf, header =T)
-gg_qqplot(exprgwas$PVALUE)
-```
-
-* For SNPs. $\pi_1 =$ `r pi_theta` , effect size = `r phenores[["batch"]][[1]][["sigma_theta"]]`, PVE = `r phenores$param$pve.snp.truth`.
-GWAS p values and qqplot:
-
-```{r snp-p, fig.width= 7, fig.height=7}
-snpgwasf <-  paste0(outputdir, runtag, "_simu", simutag, ".snpgwas.txt.gz")
-
-pvarf <- pvarfs[chrom]
-snpid <- read_pvar(pvarf)[, "id"]
-cau <- as.matrix(snpid[phenores[["batch"]][[chrom]][["idx.cSNP"]]])
-pdist_plot(snpgwasf, chrom, cau, thin = 0.1)
-```
-
-```{r snp-qq, fig.width= 5, fig.height=5}
-snpgwas <- fread(snpgwasf, header =T)
-gg_qqplot(snpgwas$PVALUE, thin = 0.1)
-```
 
@@
 
-We have tried to run MR-JTI.
-The results have higher false postive rate than TWAS.
-MR-JTI requires the SNPs be pruned before the analysis.
-It also requires that a gene has at least 20 eQTLs.
-This resulted in very few genes going into the analysis.
-Most genes left are in polymorphism dense regions, such as the MHC regions.
-I ran MR-JTI for top genes in TWAS, around 30-40% of them should be real.
-However, only a few genes pass MR-JTI’s 20 eQTL requirements and only 1 or 2 (5%) genes are real.
-We are showing MR-JTI results on this page.
+FOCUS results contains nothing.
 
@@
 
-simutags <- paste(1, c(1,2,5), sep = "-")
+simutags <- paste(1, c(1:5), sep = "-")
 plot_par(configtag, runtag, simutags)
@@
 plot_fusion_coloc(configtag, runtag, simutags)
-plot_focus(configtag, runtag, simutags)
+# plot_focus(configtag, runtag, simutags)
+print("FOCUS result is null")
 plot_smr(configtag, runtag, simutags)
@@
 plot_fusion_coloc(configtag, runtag, simutags)
-plot_focus(configtag, runtag, simutags)
+#plot_focus(configtag, runtag, simutags)
 plot_smr(configtag, runtag, simutags)
@@
 plot_fusion_coloc(configtag, runtag, simutags)
-plot_focus(configtag, runtag, simutags)
+#plot_focus(configtag, runtag, simutags)
 plot_smr(configtag, runtag, simutags)
 
-simutags <- paste(5, c(1, 3:5), sep = "-")
+simutags <- paste(5, c(1:5), sep = "-")
 plot_par(configtag, runtag, simutags)
@@
 plot_fusion_coloc(configtag, runtag, simutags)
-plot_focus(configtag, runtag, simutags)
+#plot_focus(configtag, runtag, simutags)
 plot_smr(configtag, runtag, simutags)
 
-simutags <- paste(9, 1:4, sep = "-")
+simutags <- paste(9, 1:5, sep = "-")
 plot_par(configtag, runtag, simutags)
@@
 plot_fusion_coloc(configtag, runtag, simutags)
-plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(10, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-plot_focus(configtag, runtag, simutags)
+#plot_focus(configtag, runtag, simutags)
 plot_smr(configtag, runtag, simutags)
 ```
-
-# ctwas results (LDR) 
-Using the R LD matrices as the LD reference input for ctwas, instead of the genotype type of randomly subsetted samples.
-The LD R matrices are generated using all 300k samples that passed the our filtering criteria (generated by Wes Crouse).
-A R matrice is provided for each LD block region.
-
-```{r param3, fig.height=3, fig.width=8, message=F}
-configtag <- "1_LDR"
-simutags <- paste(1, c(1,2,5), sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-
-simutags <- paste(2, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-
-simutags <- paste(3, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-
-simutags <- paste(4, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-```

### 2023-04-20 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_03222023.Rmd
more simulation settings for low SNP PVE

--- 2023-04-20 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_03212023.Rmd
@@
 ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
-outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230321/" # /
-comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230321_compare/"
+outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230322/" # /
+comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230322_compare/"
 runtag = "ukb-s80.45-adi"
@@
 plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-# plot_focus(configtag, runtag, simutags)
-print("FOCUS result is null")
-plot_smr(configtag, runtag, simutags)
+# plot_fusion_coloc(configtag, runtag, simutags)
+# # plot_focus(configtag, runtag, simutags)
+# print("FOCUS result is null")
+# plot_smr(configtag, runtag, simutags)
 
@@
 plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
+# plot_fusion_coloc(configtag, runtag, simutags)
+# #plot_focus(configtag, runtag, simutags)
+# plot_smr(configtag, runtag, simutags)
 
@@
 plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
+# plot_fusion_coloc(configtag, runtag, simutags)
+# #plot_focus(configtag, runtag, simutags)
+# plot_smr(configtag, runtag, simutags)
 ```

### 2023-04-20 analysis/simulation-ctwas-ukbWG-gtex.adipose_visual.Rmd
more simulation settings for low SNP PVE

@@
 
-plot_all <- function(susief, pf, phenof, colocf = NULL, chr, ldres = NULL, thin = 1){
+plot_all <- function(susief, pf, phenof, chr, colocf = NULL,  ldres = NULL, thin = 1){
   a <- fread(susief, header =T)
@@
 under 4-1 to 4-5
-```{r sim 2, out.width = "98%", out.height="95%"}
+```{r sim GN, out.width = "98%", out.height="95%"}
 chrom <- 4
@@
 FP example: LD with causal SNP, GALNT10
-```{r sim 2, out.width = "98%", out.height="95%"}
+```{r sim GA, out.width = "98%", out.height="95%"}
 runtag <- 5
@@
 FP example: LD with causal SNP
-```{r sim 2, out.width = "98%", out.height="95%"}
+```{r sim 10-1, out.width = "98%", out.height="95%"}
 runtag <- 5
@@
 True positive example: c1orf109
-```{r sim 2, out.width = "98%", out.height="95%"}
+```{r sim c1, out.width = "98%", out.height="95%"}
 runtag <- 4
@@
 # region plot with coloc results
-```{r sim 2, out.width = "98%", out.height="95%"}
+```{r sim 3, out.width = "98%", out.height="95%"}
 runtag <- "ukb-s80.45-adi"
@@
 ```
+## Zoom view
+For new simulation settings in 20230322
+False positives
+```{r sim 4, out.width = "98%", out.height="95%"}
+outputdir <- "~/causalTWAS/simulations/simulation_ctwas_rss_20230322/"
+runtag <- "ukb-s80.45-adi"
+simutags <- paste(2, 1:5, sep = "-")
+configtag <- 1
+
+susiefs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".susieIrss.txt")
+pfs <- paste0(outputdir, runtag, "_simu",simutags)
+gwasfs <- paste0(pfs, ".exprgwas.txt.gz")
+phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+chrom <- 3
+runtag <- 1
+plot_all(susiefs[runtag], pfs[runtag], phenofs[runtag], chrom, thin = 0.01)
+```

### 2023-04-20 analysis/visual_twas_plots.R

@@ -1,8 +1,8 @@
-library(mr.ash.alpha)
+#library(mr.ash.alpha)
 library(data.table)
 suppressMessages({library(plotly)})
 
-source("~/causalTWAS/causal-TWAS/code/before_package/fit_mr.ash.R")
+#source("~/causalTWAS/causal-TWAS/code/before_package/fit_mr.ash.R")
 
 get_files <- function(tag, tag2){
   par <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt")

### 2023-05-08 analysis/Paper_figures_simulation.Rmd
110k simulation

@@
 exprfn = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416//ukb-s80.45-adi.expr.txt"
+weightf = "/project2/mstephens/causalTWAS/fusion_weights/Adipose_Subcutaneous.pos"
 
@@
 J <- 8021 # number of genes
+
+weights <- as.data.frame(fread(weightf, header = T))
+weights$ENSEMBL_ID <- sapply(weights$WGT, function(x){unlist(strsplit(unlist(strsplit(x,"/"))[2], "[.]"))[2]})
 
@@
 
-```{r pipplot, fig.width= 8, fig.height= 16}
+```{r pipplot-suppl, fig.width= 8, fig.height= 16, eval=F}
 simutaglist = lapply(c(1:3,5:9), function(x) paste(x, 1:5, sep ="-"))
@@
 * FUSION bonferroni: 0.05
+* FUSION permutation fdr: 0.05
 * COLOC PP4: 0.8
 * FOCUS PIP: 0.8
-* SMR FDR: 0.05
-* SMR HEIDI: HEIDI p > 0.05, SMR FDR < 0.05
+* SMR FDR: 0.05, updated using ensemble ID
+* SMR HEIDI: HEIDI p > 0.05, SMR FDR < 0.05, updated using ensemble ID
 
@@
 
-get_ncausal_df <- function(pfiles, cau, cut = 0.8, 
-                           method = c("ctwas", "fusionfdr", "fusionbon","coloc", "focus", "smr", "smrheidi")){
+get_ncausal_df <- function(pfiles, cau, cut = 0.8, useFDP= F,
+                           method = c("ctwas", "fusionfdr", "fusionbon","fusionperm", "coloc", "focus", "smr", "smrheidi")){
   df <- NULL
@@
         res$ifcausal <- ifelse(res$id %in% cau[[i]], 1, 0)
-        res <- res[res$susie_pip > cut,]
+        res <- res[order(res$susie_pip, decreasing = T ),] 
+        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
+        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
+          res <- res[res$susie_pip > cut,]}
     } else if (method == "fusionfdr"){
         res$FDR <- p.adjust(res$TWAS.P, method = "fdr")
-        res <- res[res$FDR < cut,]
+        res <- res[order(res$FDR, decreasing = F),] 
         res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
+        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
+        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
+          res <- res[res$FDR < cut,]}
+
     } else if  (method == "fusionbon"){
         res$FDR <- p.adjust(res$TWAS.P, method = "bonferroni")
+        res <- res[order(res$FDR, decreasing = F),] 
+        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
+        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
+        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
           res <- res[res$FDR < cut,]
+        }
+    } else if  (method == "fusionperm"){
+        res <- res[res$PERM.N!=0,]
+        res$FDR <- p.adjust(res$PERM.PV, method = "fdr")
+        res <- res[order(res$FDR, decreasing = F),] 
         res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
+        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
+        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
+          res <- res[res$FDR < cut,]
+        }
     } else if  (method == "coloc"){
+        res <- res[!is.na(res$COLOC.PP4),]
+        res <- res[order(res$COLOC.PP4, decreasing = T),] 
+        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
+        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
+        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[res$COLOC.PP4 > cut,]
-        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
+        }
     } else if (method == "focus"){
         res <- res[res$mol_name != "NULL",]
+        res$ifcausal <- ifelse(res$mol_name %in% cau[[i]], 1, 0)
+        res <- res[order(res$pip, decreasing = T),] 
+        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
+        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
           res <- res[res$pip > cut, ] 
-        res$ifcausal <- ifelse(res$mol_name %in% cau[[i]], 1, 0)
-    } else if (method == "smr"){
+        }
+    }  else if (method == "SMR"){
+        res <- as.data.frame(res)
+        res$probeID <- sapply(res$Gene, function(x){unlist(strsplit(x, "[.]"))[1]})
+        res <- res[res$probeID %in% weights$ENSEMBL_ID,]
+        res <- res[sapply(res$p_HEIDI > 0.05, isTRUE),]
         res$FDR <- p.adjust(res$p_SMR, method = "fdr")
+        res$ifcausal <- ifelse(res$probeID %in% cau[[i]], 1, 0)
+        res <- res[order(res$FDR, decreasing = F),] 
+        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
+        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
            res <- res[res$FDR < cut,]
-        res$ifcausal <- ifelse(res$probeID %in% cau[[i]], 1, 0)
-    } else if (method == "smrheidi"){
-        res <- res[res$p_HEIDI > 0.05,]
-        res$FDR <- p.adjust(res$p_SMR, method = "fdr")
-        res <- res[res$FDR < cut,]
-        res$ifcausal <- ifelse(res$probeID %in% cau[[i]], 1, 0)
+        }
+    } else if (method == "MR-JTI"){
+        res$ifcausal <- ifelse(res$variable %in% cau[[i]], 1, 0)
+        res <- res[order(abs(res$beta), decreasing = T),] 
+        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
+        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
+          res <- res[res$CI_significance=="sig",] 
+        }
+    } else if (method == "PMR-Egger"){
+        res <- as.data.frame(res)
+        res$ifcausal <- ifelse(res$gene_id %in% cau[[i]], 1, 0) 
+        res <- res[!is.na(res$causal_pvalue), ]
+        res <- res[order(res$causal_pvalue, decreasing = F),] 
+        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
+        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
+          res <- res[sapply(res$causal_pvalue < cut/sum(!is.na(res$causal_pvalue)), isTRUE),]
+        }
     } else{
@@
 
+  cau_ensembl <- cau
+
+  for (i in 1:length(cau_ensembl)){
+    cau_ensembl[[i]][cau_ensembl[[i]] %in% weights$ID] <- weights$ENSEMBL_ID[match(cau_ensembl[[i]][cau_ensembl[[i]] %in% weights$ID], weights$ID)]
+  }
+
   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
@@
   smrfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.smr")
+  mrjtifs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.mrjti.result")
+  pmrfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.pmr.result_pi_080")
 
   ctwas_df <- get_ncausal_df(susieIfs, cau= cau, cut = 0.8, method ="ctwas")
-  fusionfdr_df <- get_ncausal_df(fusioncolocfs, cau= cau, cut = 0.05, method = "fusionfdr")
+  #fusionfdr_df <- get_ncausal_df(fusioncolocfs, cau= cau, cut = 0.05, method = "fusionfdr")
   fusionbon_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = 0.05, method = "fusionbon")
+  fusionperm_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = 0.05, method = "fusionperm",)
   coloc_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = 0.8, method = "coloc")
   focus_df <- get_ncausal_df(focusfs , cau= cau, cut = 0.8, method = "focus")
-  smr_df <- get_ncausal_df(smrfs, cau= cau, cut = 0.05, method = "smr")
-  smrheidi_df <- get_ncausal_df(smrfs, cau= cau, cut = 0.05, method = "smrheidi")
+  #smr_df <- get_ncausal_df(smrfs, cau= cau, cut = 0.05, method = "smr")
+  smrheidi_df <- get_ncausal_df(smrfs, cau= cau_ensembl, cut = 0.05, method = "SMR")
+  mrjti_df <- get_ncausal_df(mrjtifs, cau= cau_ensembl, method = "MR-JTI")
+  pmr_df <- get_ncausal_df(pmrfs, cau= cau_ensembl, cut = 0.05, method = "PMR-Egger")
 
-  df <- rbind(ctwas_df, fusionfdr_df, fusionbon_df, coloc_df, focus_df, smr_df, smrheidi_df)
+  df <- rbind(ctwas_df, fusionbon_df,fusionperm_df, coloc_df, focus_df, smrheidi_df, mrjti_df, pmr_df)
+  df$ifcausal <- df$ifcausal + as.numeric(as.factor(df$method))*10
+  df$ifcausal <- as.factor(df$ifcausal)
+  fig <- ggbarplot(df, x = "method", y = "count", add = "mean_se", fill = "ifcausal", palette = colors, legend = "none", ...) + grids(linetype = "dashed")
+  fig
+}
+```
+
+```{r ncausalplot, fig.width= 10, fig.height= 4}
+colset = c("#ebebeb", "#ffffb3", # FOCUS
+           "#ebebeb", "#8dd3c7", # Fusion
+           "#ebebeb", "palegreen", # Fusion-permutation
+           "#ebebeb", "#CC79A7", # MR-JTI
+           "#ebebeb", "goldenrod", #PMR-Egger
+           "#ebebeb", "#87CEFA", # SMR
+           "#ebebeb", "#fb8072", # cTWAS
+           "#ebebeb", "#bebada") # coloc
+f1 <- plot_ncausal(configtag, runtag,  simutaglist[[1]], colors = colset, ylim= c(0,225), main = "high gene PVE")
+f2 <- plot_ncausal(configtag, runtag,  simutaglist[[2]], colors = colset, ylim= c(0,225), main = "low gene PVE")
+gridExtra::grid.arrange(f1, f2, ncol=2)
+```
+
+### Comparison with other methods at specified FDP
+We can compare the power of methods at a given false discovery proportion, say 10% or 25%. If a method cannot achieve that FDP no matter what the threshold is used, we will point out that fact.
+
+```{r ncausalplot-FDP, fig.width= 10, fig.height= 4}
+
+plot_ncausal_FDP <- function(configtag, runtag,  simutags, colors, cut =0.2, ...){
+  phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+  cau <- lapply(phenofs, function(x) {load(x);get_causal_id(phenores)})
+
+  cau_ensembl <- cau
+
+  for (i in 1:length(cau_ensembl)){
+    cau_ensembl[[i]][cau_ensembl[[i]] %in% weights$ID] <- weights$ENSEMBL_ID[match(cau_ensembl[[i]][cau_ensembl[[i]] %in% weights$ID], weights$ID)]
+  }
+
+  susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
+  fusioncolocfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.coloc.result")
+  focusfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.focus.tsv")
+  smrfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.smr")
+  mrjtifs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.mrjti.result")
+  pmrfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.pmr.result_pi_080")
+
+  ctwas_df <- get_ncausal_df(susieIfs, cau= cau, cut = cut, method ="ctwas", useFDP = T)
+  #fusionfdr_df <- get_ncausal_df(fusioncolocfs, cau= cau, cut = 0.05, method = "fusionfdr")
+  fusionbon_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = cut, method = "fusionbon", useFDP = T)
+  fusionperm_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = cut, method = "fusionperm", useFDP = T)
+  coloc_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = cut, method = "coloc", useFDP = T)
+  focus_df <- get_ncausal_df(focusfs , cau= cau, cut = cut, method = "focus", useFDP = T)
+  #smr_df <- get_ncausal_df(smrfs, cau= cau, cut = 0.05, method = "smr")
+  smrheidi_df <- get_ncausal_df(smrfs, cau= cau_ensembl, cut = cut, method = "SMR", useFDP = T)
+  mrjti_df <- get_ncausal_df(mrjtifs, cau= cau_ensembl, cut = cut, method = "MR-JTI", useFDP = T)
+  pmr_df <- get_ncausal_df(pmrfs, cau= cau_ensembl, cut = cut, method = "PMR-Egger", useFDP = T)
+
+  df <- rbind(ctwas_df, fusionbon_df,fusionperm_df, coloc_df, focus_df, smrheidi_df, mrjti_df, pmr_df)
   df$ifcausal <- df$ifcausal + as.numeric(as.factor(df$method))*10
@@
 
-```
+f1 <- plot_ncausal_FDP(configtag, runtag,  simutaglist[[1]], colors = colset, cut = 0.2, ylim= c(0,225), main = "high gene PVE")
+f2 <- plot_ncausal_FDP(configtag, runtag,  simutaglist[[2]], colors = colset, cut = 0.2, ylim= c(0,225), main = "low gene PVE")
+gridExtra::grid.arrange(f1, f2, ncol=2)
 
-```{r ncausalplot, fig.width= 10, fig.height= 4}
-colset = c("#ebebeb", "#bebada" , "#ebebeb", "#fb8072", "#ebebeb","#ffffb3","#ebebeb", "#8dd3c7", "#ebebeb","#8dd3c7", "#ebebeb", "#87CEFA", "#ebebeb","#87CEFA") # coloc, ctwas, focus, fusionbon, fusionfdr, smrheidi, smr
-f1 <- plot_ncausal(configtag, runtag,  simutaglist[[1]], colors = colset, ylim= c(0,150), main = "high gene PVE")
-f2 <- plot_ncausal(configtag, runtag,  simutaglist[[2]], colors = colset, ylim= c(0,150), main = "low gene PVE")
+f1 <- plot_ncausal_FDP(configtag, runtag,  simutaglist[[1]], colors = colset, cut = 0.4, ylim= c(0,225), main = "high gene PVE")
+f2 <- plot_ncausal_FDP(configtag, runtag,  simutaglist[[2]], colors = colset, cut = 0.4, ylim= c(0,225), main = "low gene PVE")
+gridExtra::grid.arrange(f1, f2, ncol=2)
+
+f1 <- plot_ncausal_FDP(configtag, runtag,  simutaglist[[1]], colors = colset, cut = 1, ylim= c(0,225), main = "high gene PVE")
+f2 <- plot_ncausal_FDP(configtag, runtag,  simutaglist[[2]], colors = colset, cut = 1, ylim= c(0,225), main = "low gene PVE")
 gridExtra::grid.arrange(f1, f2, ncol=2)
@@
       colnames(z_snp) <- c("id", "A1", "A2", "z")
-      ctwas_rss(z_gene, z_snp, ld_exprfs= ld_exprfs, ld_pgenfs = ld_pgenfs, ld_R_dir = NULL, ld_regions_custom = "temp_reg.txt", thin = 1, outputdir = ".", outname = "temp", ncore = 1, ncore.rerun = 1, prob_single = 0,  group_prior = group_prior, group_prior_var = group_prior_var, estimate_group_prior = F, estimate_group_prior_var = F, recover_strand_ambig_z = F)
+      ctwas_rss(z_gene, z_snp, ld_exprfs= ld_exprfs, ld_pgenfs = ld_pgenfs, ld_R_dir = NULL, ld_regions_custom = "temp_reg.txt", thin = 1, outputdir = ".", outname = "temp", ncore = 1, ncore.rerun = 1, prob_single = 0,  group_prior = group_prior, group_prior_var = group_prior_var, estimate_group_prior = F, estimate_group_prior_var = F)
       a2 <- fread("temp.susieIrss.txt", header = T)
@@
 ```
-```{r regionplot4, echo=FALSE,results='hide',fig.keep='all', message=F}
+```{r regionplot4-1, echo=FALSE,results='hide',fig.keep='all', message=F}
 simutag <- "3-4"
@@
 ```
-```{r regionplot4, echo=FALSE,results='hide',fig.keep='all', message=F}
+```{r regionplot4-2, echo=FALSE,results='hide',fig.keep='all', message=F}
 simutag <- "3-4"
@@
 ```
-```{r regionplot4, echo=FALSE,results='hide',fig.keep='all', message=F}
+```{r regionplot4-3, echo=FALSE,results='hide',fig.keep='all', message=F}
 simutag <- "3-4"

### 2023-05-08 analysis/index.Rmd
simulation paper figures
110k simulation

@@
 
+* [**Using UKbiobank data (45k samples), fusion weights Adipose, more settings with low SNP PVE **](simulation-ctwas-ukbWG-gtex.adipose_s80.45_03222023.html)
+
+* [**Using UKbiobank data (113k samples), fusion weights Adipose, more settings with low SNP PVE **](simulation-ctwas-ukbWG-gtex.adipose_s200.113_03212023.html)
+
 * [Visualization](simulation-ctwas-ukbWG-gtex.adipose_visual.html)
@@
 * [**Using UKbiobank data (45k samples), fusion weights Adipose**](simulation-ctwas-ukbWG-gtex.adipose_s80.45_mixnormal_041821.html)
+
+## Paper
+
+* [**Simulation figures**](Paper_figures_simulation.html)
 

### 2023-05-08 analysis/simulation-ctwas-ukbWG-gtex.adipose_s200.113_03212023.Rmd
110k simulation

--- 2023-04-20 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_03212023.Rmd
@@
 ---
-title: "Simulations to test ctwas summary stats version, 45k samples, a few new setings required"
+title: "Simulations to test ctwas summary stats version, 113k samples, a few new setings required"
 output: html_document
@@
 ```{r files}
-pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt"
-ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
+pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s200.113/ukb-s200.113_pgenfs.txt"
+ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s200.113/ukb-s200.113.2_pgenfs.txt"
 outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230321/" # /
 comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230321_compare/"
-runtag = "ukb-s80.45-adi"
+runtag = "ukb-s200.113-adi"
 configtags = 1
@@
 ```{r parameters}
-n.ori <- 80000 # number of samples
+n.ori <- 200000 # number of samples
 n <- pgenlibr::GetRawSampleCt(pgens[[1]])
@@
 plot_PIP <- function(configtag, runtag,  simutags){
-   phenofs <- paste0(outputdir, "ukb-s80.45-adi", "_simu", simutags, "-pheno.Rd")
-   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
+   phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".s1.susieIrss.txt")
 
@@
 configtag <- 1
-runtag = "ukb-s80.45-adi"
+runtag = "ukb-s200.113-adi"
 
-simutags <- paste(1, c(1:5), sep = "-")
+simutags <- paste(1, c(1:24,26:54, 56:70), sep = "-")
 plot_par(configtag, runtag, simutags)
 plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
+#plot_fusion_coloc(configtag, runtag, simutags)
 # plot_focus(configtag, runtag, simutags)
-print("FOCUS result is null")
-plot_smr(configtag, runtag, simutags)
+#print("FOCUS result is null")
+#plot_smr(configtag, runtag, simutags)
 
-simutags <- paste(2, 1:5, sep = "-")
+simutags <- paste(2, 1:30, sep = "-")
 plot_par(configtag, runtag, simutags)
 plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
+# plot_fusion_coloc(configtag, runtag, simutags)
+# #plot_focus(configtag, runtag, simutags)
+# plot_smr(configtag, runtag, simutags)
 
-simutags <- paste(3, 1:5, sep = "-")
+simutags <- paste(3, 1:10, sep = "-")
 plot_par(configtag, runtag, simutags)
 plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
+# plot_fusion_coloc(configtag, runtag, simutags)
+# #plot_focus(configtag, runtag, simutags)
+# plot_smr(configtag, runtag, simutags)
 
-simutags <- paste(4, 1:5, sep = "-")
+simutags <- paste(4, 1:10, sep = "-")
 plot_par(configtag, runtag, simutags)
 plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
+# plot_fusion_coloc(configtag, runtag, simutags)
+# #plot_focus(configtag, runtag, simutags)
+# plot_smr(configtag, runtag, simutags)
 
-simutags <- paste(5, c(1:5), sep = "-")
+simutags <- paste(5, c(1:10), sep = "-")
 plot_par(configtag, runtag, simutags)
 plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(6, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(7, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(8, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(9, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
+# plot_fusion_coloc(configtag, runtag, simutags)
+# #plot_focus(configtag, runtag, simutags)
+# plot_smr(configtag, runtag, simutags)
 ```

### 2023-05-08 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_03222023.Rmd
110k simulation

@@
 
-simutags <- paste(6, 1:5, sep = "-")
+simutags <- paste(6, 1:20, sep = "-")
 plot_par(configtag, runtag, simutags)
@@
 
-simutags <- paste(7, 1:5, sep = "-")
+simutags <- paste(7, 1:20, sep = "-")
 plot_par(configtag, runtag, simutags)
@@
 
-simutags <- paste(8, 1:5, sep = "-")
+simutags <- paste(8, 1:20, sep = "-")
 plot_par(configtag, runtag, simutags)
@@
 
-simutags <- paste(9, 1:5, sep = "-")
+simutags <- paste(9, 1:20, sep = "-")
 plot_par(configtag, runtag, simutags)

### 2023-05-08 analysis/summarize_basic_plots.R

@@ -135,9 +135,9 @@ cp_plot <- function(pips, ifcausal, runtag = NULL, mode = c("PIP", "FDR"), main 
       ifcausal.rt <- ifcausal[runtag == rt]
       dflist[[rt]] <- .exob(pips.rt, ifcausal.rt, mode = mode)
     }
-    mean_pip <- colMeans(do.call(rbind, lapply(dflist, '[[', "expected")))
-    observed_freq <- colMeans(do.call(rbind, lapply(dflist, '[[', "observed")))
-    se <- apply(do.call(rbind, lapply(dflist, '[[', "observed")), 2, plotrix::std.error)
+    mean_pip <- colMeans(do.call(rbind, lapply(dflist, '[[', "expected")),na.rm = T)
+    observed_freq <- colMeans(do.call(rbind, lapply(dflist, '[[', "observed")), na.rm =T)
+    se <- apply(do.call(rbind, lapply(dflist, '[[', "observed")), 2, plotrix::std.error, na.rm =T)
   }
 
   df <- data.frame("mean_pip" = mean_pip, "observed_freq"= observed_freq, "se" = se)

### 2023-05-10 analysis/simulation-ctwas-ukbWG-gtex.adipose_s200.113_03212023.Rmd
L=1 for ctwas rerun of 113k

@@
 runtag = "ukb-s200.113-adi"
-configtags = 1
+
 simutags = paste(rep(1:9, each = length(1:5)), 1:5, sep = "-")
@@
 
-`L = 5` , after parameter estimation, for running susie for all regions.
-
 `group_prior = NULL`, the initiating prior parameters we used for running susie for each region is uniform prior for genes and SNPs.
@@
 
+We have two configurations for `L`: config 1 the last step of ctwas was running susie with `L=5` in regions with big gene PIPs, config 2 the last step of ctwas was running susie with `L=1` in regions with big gene PIPs.
+
 # `ctwas` results
 
-Results: Each row shows parameter estimation results from 5 simulation runs with similar settings (i.e.
+Results: Each row shows parameter estimation results from >5 simulation runs with similar settings (i.e.
 pi1 and PVE for genes and SNPs).
@@
 `truth`: the true parameters,  `selected_truth`: the truth in selected regions that were used to estimate parameters, `ctwas`: ctwas estimated parameters (using summary statistics as input).
-We run FUSION following default settings and adjust p values by BH method to get expected FDP.
-
-We run FUSION following default settings and adjust p values by BH method to get expected FDP. We have also used Bonferroni correction for p values.
-
-We ran coloc for all genes with TWAS p < 1e-4. We use PP4 (SNP associate with both traits).
-We ran SMR+HEIDI, using eQTL summary statistics GTEx v.7. We filter the results by requiring p_HEIDI > 0.05. The plots 
-are based on SMR p value adjusted by BH method to get expected FDP.
-
-FOCUS results contains nothing.
 
@@
    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
-   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".s1.susieIrss.txt")
+   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
 
@@
 
-```{r param2, fig.height=3, fig.width=8, message=F}
+```{r config-1, fig.height=3, fig.width=8, message=F}
 
-configtag <- 1
 runtag = "ukb-s200.113-adi"
 
-simutags <- paste(1, c(1:24,26:54, 56:70), sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
+simutags <- paste(1, c(1:24, 26:54, 56:70), sep = "-")
+# configtag <- 1
+# plot_par(configtag, runtag, simutags)
+# plot_PIP(configtag, runtag, simutags)
+plot_par(1, runtag, simutags)
+print("When running with L= 5 in final step:")
+plot_PIP(1, runtag, simutags)
+print("When running with L= 1 in final step:")
+plot_PIP(2, runtag, simutags)
 #plot_fusion_coloc(configtag, runtag, simutags)
@@
 simutags <- paste(2, 1:30, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-# plot_fusion_coloc(configtag, runtag, simutags)
-# #plot_focus(configtag, runtag, simutags)
-# plot_smr(configtag, runtag, simutags)
+plot_par(1, runtag, simutags)
+print("When running with L= 5 in final step:")
+plot_PIP(1, runtag, simutags)
+print("When running with L= 1 in final step:")
+plot_PIP(2, runtag, simutags)
 
 simutags <- paste(3, 1:10, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-# plot_fusion_coloc(configtag, runtag, simutags)
-# #plot_focus(configtag, runtag, simutags)
-# plot_smr(configtag, runtag, simutags)
+plot_par(1, runtag, simutags)
+print("When running with L= 5 in final step:")
+plot_PIP(1, runtag, simutags)
+print("When running with L= 1 in final step:")
+plot_PIP(2, runtag, simutags)
 
 simutags <- paste(4, 1:10, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-# plot_fusion_coloc(configtag, runtag, simutags)
-# #plot_focus(configtag, runtag, simutags)
-# plot_smr(configtag, runtag, simutags)
+plot_par(1, runtag, simutags)
+print("When running with L= 5 in final step:")
+plot_PIP(1, runtag, simutags)
+print("When running with L= 1 in final step:")
+plot_PIP(2, runtag, simutags)
 
 simutags <- paste(5, c(1:10), sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-# plot_fusion_coloc(configtag, runtag, simutags)
-# #plot_focus(configtag, runtag, simutags)
-# plot_smr(configtag, runtag, simutags)
+plot_par(1, runtag, simutags)
+print("When running with L= 5 in final step:")
+plot_PIP(1, runtag, simutags)
+print("When running with L= 1 in final step:")
+plot_PIP(2, runtag, simutags)
 ```

### 2023-05-10 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_03222023.Rmd
L=1 for ctwas rerun of 113k

--- 2023-04-20 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_03212023.Rmd
@@
 ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
-outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230321/" # /
-comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230321_compare/"
+outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230322/" # /
+comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230322_compare/"
 runtag = "ukb-s80.45-adi"
-configtags = 1
+
 simutags = paste(rep(1:9, each = length(1:5)), 1:5, sep = "-")
@@
 `max_SNP_region = 5000`, the maximum number of SNPs for re-running susie on strong gene signal regions is 5000.
+
+We have two configurations for `L`: config 1 the last step of ctwas was running susie with `L=5` in regions with big gene PIPs, config 2 the last step of ctwas was running susie with `L=1` in regions with big gene PIPs.
 
@@
 
-configtag <- 1
+#configtag <- 1
 runtag = "ukb-s80.45-adi"
 
-simutags <- paste(1, c(1:5), sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-# plot_focus(configtag, runtag, simutags)
-print("FOCUS result is null")
-plot_smr(configtag, runtag, simutags)
+# simutags <- paste(1, c(1:5), sep = "-")
+# plot_par(configtag, runtag, simutags)
+# plot_PIP(configtag, runtag, simutags)
+# # plot_fusion_coloc(configtag, runtag, simutags)
+# # # plot_focus(configtag, runtag, simutags)
+# # print("FOCUS result is null")
+# # plot_smr(configtag, runtag, simutags)
+# 
+# simutags <- paste(2, 1:5, sep = "-")
+# plot_par(configtag, runtag, simutags)
+# plot_PIP(configtag, runtag, simutags)
+# # plot_fusion_coloc(configtag, runtag, simutags)
+# # #plot_focus(configtag, runtag, simutags)
+# # plot_smr(configtag, runtag, simutags)
+# 
+# simutags <- paste(3, 1:5, sep = "-")
+# plot_par(configtag, runtag, simutags)
+# plot_PIP(configtag, runtag, simutags)
+# # plot_fusion_coloc(configtag, runtag, simutags)
+# # #plot_focus(configtag, runtag, simutags)
+# # plot_smr(configtag, runtag, simutags)
+# 
+# simutags <- paste(4, 1:5, sep = "-")
+# plot_par(configtag, runtag, simutags)
+# plot_PIP(configtag, runtag, simutags)
+# # plot_fusion_coloc(configtag, runtag, simutags)
+# # #plot_focus(configtag, runtag, simutags)
+# # plot_smr(configtag, runtag, simutags)
+# 
+# simutags <- paste(5, c(1:5), sep = "-")
+# plot_par(configtag, runtag, simutags)
+# plot_PIP(configtag, runtag, simutags)
+# plot_fusion_coloc(configtag, runtag, simutags)
+# #plot_focus(configtag, runtag, simutags)
+# plot_smr(configtag, runtag, simutags)
 
-simutags <- paste(2, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
+simutags <- paste(6, 1:20, sep = "-")
+plot_par(1, runtag, simutags)
+print("When running with L= 5 in final step:")
+plot_PIP(1, runtag, simutags)
+print("When running with L= 1 in final step:")
+plot_PIP(2, runtag, simutags)
 
-simutags <- paste(3, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
+simutags <- paste(7, 1:20, sep = "-")
+plot_par(1, runtag, simutags)
+print("When running with L= 5 in final step:")
+plot_PIP(1, runtag, simutags)
+print("When running with L= 1 in final step:")
+plot_PIP(2, runtag, simutags)
 
-simutags <- paste(4, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
+simutags <- paste(8, 1:20, sep = "-")
+plot_par(1, runtag, simutags)
+print("When running with L= 5 in final step:")
+plot_PIP(1, runtag, simutags)
+print("When running with L= 1 in final step:")
+plot_PIP(2, runtag, simutags)
 
-simutags <- paste(5, c(1:5), sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(6, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(7, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(8, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
-
-simutags <- paste(9, 1:5, sep = "-")
-plot_par(configtag, runtag, simutags)
-plot_PIP(configtag, runtag, simutags)
-plot_fusion_coloc(configtag, runtag, simutags)
-#plot_focus(configtag, runtag, simutags)
-plot_smr(configtag, runtag, simutags)
+simutags <- paste(9, 1:20, sep = "-")
+plot_par(1, runtag, simutags)
+print("When running with L= 5 in final step:")
+plot_PIP(1, runtag, simutags)
+print("When running with L= 1 in final step:")
+plot_PIP(2, runtag, simutags)
 ```

### 2023-05-10 code/run_MRLocus_wc.R

@@ -3,15 +3,16 @@ library(mrlocus)
 
 args = commandArgs(trailingOnly=TRUE)
 
-if (length(args) != 7) {
-  stop(" 7 arguments:
+if (length(args) != 8) {
+  stop(" 8 arguments:
        * LD_R directory
        * gwas file
        * eqtl summary statistics
        * weight .pos file
        * output directory
        * out file name
-       * batch", call.=FALSE)
+       * batch
+       * prune clumps", call.=FALSE)
 }
 
 print(args)
@@ -23,8 +24,10 @@ weightf <- as.character(args[4])
 outputdir <- as.character(args[5])
 outname <- as.character(args[6])
 batch <- as.numeric(args[7])
+prune_clumps <- as.logical(args[8])
 
 batch_size <- 250
+ncores <- 2
 
 ####################
 
@@ -33,8 +36,9 @@ batch_size <- 250
 # eqtl <- "/project2/mstephens/causalTWAS/GTEx_v7_all/Adipose_Subcutaneous.allpairs_processed.txt.gz"
 # weightf <- "/project2/mstephens/causalTWAS/fusion_weights/Adipose_Subcutaneous.pos"
 # outputdir <- "/home/wcrouse/causalTWAS/simulations/simulation_ctwas_rss_20210416_compare/"
-# outname <- "/project2/mstephens/causalTWAS/simulations/simulation_ctwas_rss_20210416_compare/ukb-s80.45-adi_simu4-1.Adipose_Subcutaneous.mrlocus.result.batch_4"
-# batch <- 4
+# outname <- "/project2/mstephens/causalTWAS/simulations/simulation_ctwas_rss_20210416_compare/ukb-s80.45-adi_simu4-1.Adipose_Subcutaneous.mrlocus.result.batch_1"
+# batch <- 1
+# prune_clumps <- TRUE
 
 ####################
 
@@ -106,6 +110,7 @@ eqtl <- eqtl[eqtl$gene_id %in% weights$ENSEMBL_ID,]
 #drop entries not uniquely identified by gene_id and variant id (variant not biallelic)
 eqtl_unique_id_gene <- paste0(eqtl$id, eqtl$gene_id)
 eqtl <- eqtl[!(eqtl_unique_id_gene %in% eqtl_unique_id_gene[duplicated(eqtl_unique_id_gene)]),]
+rm(eqtl_unique_id_gene)
 
 ####################
 #LD files
[…]

### 2023-05-10 code/run_ctwas_rss2.R

@@ -29,6 +29,7 @@ ncore.rerun <- 1
 prob_single <- 0.8
 thin <- 1
 max_snp_region <- 1e4
+L <- 5
 
 source(args[4]) # config
 
@@ -36,9 +37,11 @@ source(args[4]) # config
 ld_exprfs <- paste0(outputdir, "/", outname.e, "_chr", 1:22, ".expr.gz")
 load(file = paste0(outputdir, "/", outname.e, "_z_gene.Rd"))
 
-load( paste0(outputdir, "/", outname, ".s2.susieIrssres.Rd"))
+load(paste0(outputdir, "/", gsub("config2", "config1", outname), ".s2.susieIrssres.Rd"))  # this is unusual. To check for each run.
+# load(paste0(outputdir, "/", outname, ".s2.susieIrssres.Rd"))
+
 group_prior <- group_prior_rec[, ncol(group_prior_rec)]
 group_prior_var <- group_prior_var_rec[, ncol(group_prior_var_rec)]
 
 # run ctwas_rss last step
-ctwas_rss(z_gene, z_snp, ld_exprfs, ld_pgenfs = ld_pgenfs, ld_R_dir = NULL, ld_regions = "EUR", ld_regions_custom = ld_regions_custom, thin = thin, max_snp_region = max_snp_region, outputdir = outputdir, outname = outname, ncore = ncore, ncore.rerun = ncore.rerun, prob_single = prob_single,  group_prior = group_prior, group_prior_var = group_prior_var, estimate_group_prior = F, estimate_group_prior_var = F)
+ctwas_rss(z_gene, z_snp, ld_exprfs, ld_pgenfs = ld_pgenfs, ld_R_dir = NULL, ld_regions = "EUR", ld_regions_custom = ld_regions_custom, thin = thin, L= L, max_snp_region = max_snp_region, outputdir = outputdir, outname = outname, ncore = ncore, ncore.rerun = ncore.rerun, prob_single = prob_single,  group_prior = group_prior, group_prior_var = group_prior_var, estimate_group_prior = F, estimate_group_prior_var = F)

### 2023-05-25 analysis/simulation-ctwas-ukbWG-gtex.adipose_visual.Rmd
PVEgene=0.02

@@
 ```
+
+## Results from `L=1` vs `L=5` for low SNP PVE settings.
+
+```{r sim 2, out.width = "98%", out.height="95%"}
+outputdir <- "~/causalTWAS/simulations/simulation_ctwas_rss_20230322/"
+runtag <- "ukb-s80.45-adi"
+simutags <- paste(8, 1, sep = "-")
+
+configtag <- 1
+susiefs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".susieIrss.txt")
+pfs <- paste0(outputdir, runtag, "_simu",simutags)
+gwasfs <- paste0(pfs, ".exprgwas.txt.gz")
+phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+#scatter_plot_PIP_p(phenofs, susiefs, gwasfs, main ="PIP-p")
+```
+Zoom in view
+```{r sim 3, out.width = "98%", out.height="95%"}
+chrom <- 1
+rtag <- 1
+plot_all(susiefs[rtag], pfs[rtag], phenofs[rtag], chrom, thin = 0.01)
+```
+```{r sim 2, out.width = "98%", out.height="95%"}
+configtag <- 2
+susiefs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".susieIrss.txt")
+pfs <- paste0(outputdir, runtag, "_simu",simutags)
+gwasfs <- paste0(pfs, ".exprgwas.txt.gz")
+phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+#scatter_plot_PIP_p(phenofs, susiefs, gwasfs, main ="PIP-p")
+
+chrom <- 1
+rtag <- 1
+plot_all(susiefs[rtag], pfs[rtag], phenofs[rtag], chrom, thin = 0.01)
+```
+
+```{r rerun}
+#a <- plot_region(runtag, simutags[1], configtag, chr = 1, startpos = 222230220 , endpos = 224938520 ,  plot_ceqtl = T, rerun_ctwas = T)
+
+load(phenofs[1])
+cau <- get_causal_id(phenores)
+load("~/temp2.Rd")
+susieres <- susie_rss(z, R, z_ld_weight = z_ld_weight, 
+                  L = L, prior_weights = prior, null_weight = nw, 
+                  prior_variance = V, estimate_prior_variance = F, 
+                  coverage = coverage)
+load("~/temp1.Rd")
+outdf.reg <- ctwas:::anno_susie(susieres, geneinfo, snpinfo, gidx, sidx , 1, 1)
+
+a <- plot_region(runtag, simutags[1], configtag, chr = 1, startpos =32438685 , endpos = 34799758,  plot_ceqtl = T, rerun_ctwas = T)
+
+```

### 2023-06-05 analysis/simulation-ctwas-ukbWG-gtex.adipose_s200.113_03212023.Rmd
PVEgene=0.02
low PVE simulation bug fix

@@
 
+Note, config 1 was not run correctly. Need to rerun with `harmonize_z = F` in `ctwas_rss`.
+
 ```{r func}
@@
 
-```{r config-1, fig.height=3, fig.width=8, message=F}
+```{r config-1, fig.height=3, fig.width=8, message=F, eval=F}
 
@@
 simutags <- paste(2, 1:30, sep = "-")
+plot_par(1, runtag, simutags)
+print("When running with L= 5 in final step:")
+plot_PIP(1, runtag, simutags)
+print("When running with L= 1 in final step:")
+plot_PIP(2, runtag, simutags)
+
+simutags <- paste(6, c(1:20), sep = "-")
 plot_par(1, runtag, simutags)

### 2023-06-05 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_03222023.Rmd
PVEgene=0.02
low PVE simulation bug fix

@@
 
-Results: Each row shows parameter estimation results from 5 simulation runs with similar settings (i.e.
+Results: Each row shows parameter estimation results from 20 simulation runs with similar settings (i.e.
 pi1 and PVE for genes and SNPs).
@@
 `truth`: the true parameters,  `selected_truth`: the truth in selected regions that were used to estimate parameters, `ctwas`: ctwas estimated parameters (using summary statistics as input).
+
+Note, we have three configurations, `config1`, `config2`, `config3`.
+Config 1 was run with wrong harmonziation configuration, they were rerun the last step as `config3`.
+This were run by v.0.1.29 (`harmonize_z = F` for config3, `harmonize_z = T` for config1.)
 
@@
 
-```{r param2, fig.height=3, fig.width=8, message=F}
-
+```{r config1, fig.height=3, fig.width=8, message=F}
 #configtag <- 1
@@
 print("When running with L= 5 in final step:")
-plot_PIP(1, runtag, simutags)
+plot_PIP(3, runtag, simutags)
 print("When running with L= 1 in final step:")
@@
 
-simutags <- paste(8, 1:20, sep = "-")
+simutags <- paste(11, 1:20, sep = "-")
 plot_par(1, runtag, simutags)
 print("When running with L= 5 in final step:")
-plot_PIP(1, runtag, simutags)
+plot_PIP(3, runtag, simutags)
+print("When running with L= 1 in final step:")
+plot_PIP(2, runtag, simutags)
+
+simutags <- paste(8, c(1:13,15:20), sep = "-")
+plot_par(1, runtag, simutags)
+print("When running with L= 5 in final step:")
+plot_PIP(3, runtag, simutags)
 print("When running with L= 1 in final step:")

### 2023-06-05 analysis/simulation-ctwas-ukbWG-gtex.adipose_visual.Rmd
low PVE simulation bug fix

--- 2023-04-20 analysis/simulation-ctwas-ukbWG-gtex.adipose_visual.Rmd
@@
 ```
+
+## Results from `L=1` vs `L=5` for low SNP PVE settings.
+
+```{r sim 2, out.width = "98%", out.height="95%"}
+simuoutputdir <- "~/causalTWAS/simulations/simulation_ctwas_rss_20230322/"
+runtag <- "ukb-s80.45-adi"
+simutags <- paste(8, 1, sep = "-")
+
+configtag <- 1
+susiefs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".susieIrss.txt")
+pfs <- paste0(outputdir, runtag, "_simu",simutags)
+gwasfs <- paste0(pfs, ".exprgwas.txt.gz")
+phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+scatter_plot_PIP_p(phenofs, susiefs, gwasfs, main ="PIP-p")
+
+configtag <- 2
+susiefs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".susieIrss.txt")
+pfs <- paste0(outputdir, runtag, "_simu",simutags)
+gwasfs <- paste0(pfs, ".exprgwas.txt.gz")
+phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+scatter_plot_PIP_p(phenofs, susiefs, gwasfs, main ="PIP-p")
+
+```
+Zoom in view
+```{r L=5, out.width = "98%", out.height="95%"}
+configtag <- 1
+susiefs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".susieIrss.txt")
+pfs <- paste0(outputdir, runtag, "_simu",simutags)
+gwasfs <- paste0(pfs, ".exprgwas.txt.gz")
+phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+chrom <- 17
+rtag <- 1
+plot_all(susiefs[rtag], pfs[rtag], phenofs[rtag], chrom, thin = 0.01)
+```
+```{r L=1, out.width = "98%", out.height="95%"}
+configtag <- 2
+susiefs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".susieIrss.txt")
+pfs <- paste0(outputdir, runtag, "_simu",simutags)
+gwasfs <- paste0(pfs, ".exprgwas.txt.gz")
+phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
+#scatter_plot_PIP_p(phenofs, susiefs, gwasfs, main ="PIP-p")
+
+chrom <- 17
+rtag <- 1
+plot_all(susiefs[rtag], pfs[rtag], phenofs[rtag], chrom, thin = 0.01)
+```
+
+```{r rerun}
+outputdir <- "~/causalTWAS/simulations/simulation_ctwas_rss_20230322/"
+runtag <- "ukb-s80.45-adi"
+source("~/causalTWAS/causal-TWAS/analysis/summarize_basic_plots.R")
+source("~/causalTWAS/causal-TWAS/analysis/summarize_ctwas_plots.R")
+source("~/causalTWAS/causal-TWAS/analysis/ld.R")
+pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt"
+ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
+exprfn = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230322//ukb-s80.45-adi.expr.txt"
+
+ld_pgenfs <- read.table(ld_pgenfn, header = F, stringsAsFactors = F)[,1]
+pgenfs <- read.table(pgenfn, header = F, stringsAsFactors = F)[,1]
+pvarfs <- sapply(pgenfs, prep_pvar, outputdir = outputdir)
+pgens <- lapply(1:length(pgenfs), function(x) prep_pgen(pgenf = pgenfs[x],pvarf = pvarfs[x]))
+exprfs <- read.table(exprfn, header = F, stringsAsFactors = F)[,1]
+exprvarfs <- sapply(exprfs, prep_exprvar)
+
+configtag <- 1
+# debug(susieI_rss) to generate data. save(list=ls(), file="~/temp.Rd")
+a <- plot_region(runtag, simutags[1], configtag, chr = 17, startpos = 9965921, endpos = 12534769,  plot_ceqtl = F, rerun_ctwas = T)
+
+load(phenofs[1])
+cau <- get_causal_id(phenores)
+
+susieres <- susie_rss(z, R, z_ld_weight = z_ld_weight, 
+                  L = L, prior_weights = prior, null_weight = nw, 
+                  prior_variance = V, estimate_prior_variance = F, 
+                  coverage = coverage)
+outdf.reg <- ctwas:::anno_susie(susieres, geneinfo, snpinfo, gidx, sidx , 1, 1) 
+```

### 2023-06-05 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_03222023.Rmd
remove L=1 results for low PVE

@@
 plot_PIP(3, runtag, simutags)
-print("When running with L= 1 in final step:")
-plot_PIP(2, runtag, simutags)
+# print("When running with L= 1 in final step:")
+# plot_PIP(2, runtag, simutags)
 
@@
 plot_PIP(3, runtag, simutags)
-print("When running with L= 1 in final step:")
-plot_PIP(2, runtag, simutags)
+# print("When running with L= 1 in final step:")
+# plot_PIP(2, runtag, simutags)
 ```

### 2023-06-10 analysis/index.Rmd
low PVE simulation bug fix
more settings for low SNP PVE

@@
 
-* [**Using UKbiobank data (113k samples), fusion weights Adipose, more settings with low SNP PVE **](simulation-ctwas-ukbWG-gtex.adipose_s200.113_03212023.html)
+* [**Using UKbiobank data (110k samples), fusion weights Adipose, more settings with low SNP PVE **](simulation-ctwas-ukbWG-gtex.adipose_s200.113_03221023.html)
 

### 2023-06-10 analysis/simulation-ctwas-ukbWG-gtex.adipose_s200.113_03212023.Rmd
more settings for low SNP PVE

@@
 
-```{r config-1, fig.height=3, fig.width=8, message=F, eval=F}
+```{r config-1-2, fig.height=3, fig.width=8, message=F}
 
@@
 
-simutags <- paste(1, c(1:24, 26:54, 56:70), sep = "-")
+# has not rerun last step for simu 1-x. 
+# simutags <- paste(1, c(1:24, 26:54, 56:70), sep = "-")
 # configtag <- 1 
@@
 # plot_PIP(configtag, runtag, simutags)
-plot_par(1, runtag, simutags)
-print("When running with L= 5 in final step:")
-plot_PIP(1, runtag, simutags)
-print("When running with L= 1 in final step:")
-plot_PIP(2, runtag, simutags)
+# plot_par(1, runtag, simutags)
+# print("When running with L= 5 in final step:")
+# plot_PIP(1, runtag, simutags)
+# print("When running with L= 1 in final step:")
+# plot_PIP(2, runtag, simutags)
 #plot_fusion_coloc(configtag, runtag, simutags)

### 2023-06-10 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_03222023.Rmd
more settings for low SNP PVE

--- 2023-06-05 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_03222023.Rmd
@@
 
-# simutags <- paste(1, c(1:5), sep = "-")
-# plot_par(configtag, runtag, simutags)
-# plot_PIP(configtag, runtag, simutags)
-# # plot_fusion_coloc(configtag, runtag, simutags)
-# # # plot_focus(configtag, runtag, simutags)
-# # print("FOCUS result is null")
-# # plot_smr(configtag, runtag, simutags)
-# 
-# simutags <- paste(2, 1:5, sep = "-")
-# plot_par(configtag, runtag, simutags)
-# plot_PIP(configtag, runtag, simutags)
-# # plot_fusion_coloc(configtag, runtag, simutags)
-# # #plot_focus(configtag, runtag, simutags)
-# # plot_smr(configtag, runtag, simutags)
-# 
-# simutags <- paste(3, 1:5, sep = "-")
-# plot_par(configtag, runtag, simutags)
-# plot_PIP(configtag, runtag, simutags)
-# # plot_fusion_coloc(configtag, runtag, simutags)
-# # #plot_focus(configtag, runtag, simutags)
-# # plot_smr(configtag, runtag, simutags)
-# 
-# simutags <- paste(4, 1:5, sep = "-")
-# plot_par(configtag, runtag, simutags)
-# plot_PIP(configtag, runtag, simutags)
-# # plot_fusion_coloc(configtag, runtag, simutags)
-# # #plot_focus(configtag, runtag, simutags)
-# # plot_smr(configtag, runtag, simutags)
-# 
-# simutags <- paste(5, c(1:5), sep = "-")
-# plot_par(configtag, runtag, simutags)
-# plot_PIP(configtag, runtag, simutags)
+simutags <- paste(1, c(1:5), sep = "-")
+plot_par(1, runtag, simutags)
+print("When running with L= 5 in final step:")
+plot_PIP(1, runtag, simutags)
+print("When running with L= 1 in final step:")
+plot_PIP(2, runtag, simutags)
+
+# plot_fusion_coloc(configtag, runtag, simutags)
+# # plot_focus(configtag, runtag, simutags)
+# print("FOCUS result is null")
+# plot_smr(configtag, runtag, simutags)
+
+simutags <- paste(2, 1:5, sep = "-")
+plot_par(1, runtag, simutags)
+print("When running with L= 5 in final step:")
+plot_PIP(1, runtag, simutags)
+print("When running with L= 1 in final step:")
+plot_PIP(2, runtag, simutags)
 # plot_fusion_coloc(configtag, runtag, simutags)
@@
 # plot_smr(configtag, runtag, simutags)
+
+simutags <- paste(3, 1:5, sep = "-")
+plot_par(1, runtag, simutags)
+print("When running with L= 5 in final step:")
+plot_PIP(1, runtag, simutags)
+print("When running with L= 1 in final step:")
+plot_PIP(2, runtag, simutags)
+# plot_fusion_coloc(configtag, runtag, simutags)
+# #plot_focus(configtag, runtag, simutags)
+# plot_smr(configtag, runtag, simutags)
+
+simutags <- paste(4, 1:5, sep = "-")
+plot_par(1, runtag, simutags)
+print("When running with L= 5 in final step:")
+plot_PIP(1, runtag, simutags)
+print("When running with L= 1 in final step:")
+plot_PIP(2, runtag, simutags)
+# plot_fusion_coloc(configtag, runtag, simutags)
+# #plot_focus(configtag, runtag, simutags)
+# plot_smr(configtag, runtag, simutags)
+
+simutags <- paste(5, c(1:5), sep = "-")
+plot_par(1, runtag, simutags)
+print("When running with L= 5 in final step:")
+plot_PIP(1, runtag, simutags)
+print("When running with L= 1 in final step:")
+plot_PIP(2, runtag, simutags)
 

### 2023-06-11 analysis/Paper_Suppl_figures_simulation.Rmd
fix link bug

@@
 ```
+
+## Additional plots for revision, just PIP plots
+```{r PIPplot2, fig.width= 8, fig.height= 18}
+outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20220322/"
+# simu 1-x and 2-x, SNP PVE = 0.1
+# simu 7,11,8,9, SNP PVE =0.2
+simutaglist = c(lapply(1:2, function(x) paste(x, 1:5, sep ="-")),
+                list(paste(c(7,11,8,9), 1:20, sep = "-")))
+pdf(file = "temp3.pdf", height = 25, width =3)
+plotlist <- list()
+for (i in 1:length(simutaglist)){
+  plotlist[[i]] <- plot_PIP(configtag, runtag, simutaglist[[i]], main = "")
+}
+
+outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20220321/"
+simutaglist = c(lapply(1:2, function(x) paste(x, 1:5, sep ="-")),
+                list(paste(c(7,11,8,9), 1:20, sep = "-")))
+pdf(file = "temp3.pdf", height = 25, width =3)
+plotlist <- list()
+for (i in 1:length(simutaglist)){
+  plotlist[[i]] <- plot_PIP(configtag, runtag, simutaglist[[i]], main = "")
+}
+
+gridExtra::grid.arrange(plotlist[[1]], plotlist[[2]],
+                        plotlist[[3]], plotlist[[4]],
+                        plotlist[[5]], plotlist[[6]],
+                        plotlist[[7]], plotlist[[8]],
+                        plotlist[[9]], plotlist[[10]],
+                        ncol = 2)
+dev.off()
+```

### 2023-06-11 analysis/index.Rmd
fix link bug

@@
 
-* [**Using UKbiobank data (110k samples), fusion weights Adipose, more settings with low SNP PVE **](simulation-ctwas-ukbWG-gtex.adipose_s200.113_03221023.html)
+* [**Using UKbiobank data (110k samples), fusion weights Adipose, more settings with low SNP PVE **](simulation-ctwas-ukbWG-gtex.adipose_s200.113_03212023.html)
 

### 2023-06-18 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45_03222023.Rmd
revision figures

@@
 print("When running with L= 5 in final step:")
-plot_PIP(3, runtag, simutags)
+plot_PIP(1, runtag, simutags)
 print("When running with L= 1 in final step:")
@@
 
-simutags <- paste(8, c(1:13,15:20), sep = "-")
+simutags <- paste(8, c(1:20), sep = "-")
 plot_par(1, runtag, simutags)
 print("When running with L= 5 in final step:")
-plot_PIP(3, runtag, simutags)
+plot_PIP(1, runtag, simutags)
 print("When running with L= 1 in final step:")

### 2023-06-21 analysis/Paper_figures_simulation.Rmd
low PVE simulation bug fix
fix fusion perm

@@
 * FUSION bonferroni: 0.05
-* FUSION permutation fdr: 0.05
+* FUSION permutation: bonferroni < 0.05 then use empirical permutation p < 0.01
 * COLOC PP4: 0.8
@@
     } else if  (method == "fusionperm"){
-        res <- res[res$PERM.N!=0,]
-        res$FDR <- p.adjust(res$PERM.PV, method = "fdr")
+        res$FDR <- p.adjust(res$TWAS.P, method = "bonferroni")
+        res <- res[res$PERM.N >0 & res$PERM.PV < 0.01,]
         res <- res[order(res$FDR, decreasing = F),] 
@@
            "#ebebeb", "#bebada") # coloc
+simutaglist = list(paste(4, 1:5, sep = "-"), paste(10, 1:5, sep="-"))
 f1 <- plot_ncausal(configtag, runtag,  simutaglist[[1]], colors = colset, ylim= c(0,225), main = "high gene PVE")
